### 2020-02-07 paper/paper.tex
Add skeleton for a paper

\documentclass{scrartcl}

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{todonotes}
\usepackage{enumitem}
\setlist{parsep=0pt,listparindent=\parindent}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[many]{tcolorbox}

% Links and their colors
\usepackage[
  colorlinks=true,
  linkcolor=darkgray,
  citecolor=darkgray,
  urlcolor=darkgray,
  ]{hyperref}

% Commands for the pi-calculus
\newcommand{\PO}{\mathbf{0}}
\newcommand{\comp}[2]{#1 \mid #2}
\newcommand{\new}[2]{(\boldsymbol{\nu} #1 #2)\,}
\newcommand{\cout}[2]{\overline{#1}\langle#2\rangle.}
\newcommand{\cin}[2]{#1(#2).}
\newcommand{\select}[2]{#1\triangleleft#2.}
\newcommand{\branch}[2]{#1\triangleright#2}

\newcommand{\subst}[3]{#1[#2/#3]}

\newcommand{\picalc}{$\pi$-calculus}
\newcommand{\Picalc}{$\pi$-Calculus}

\newcommand{\type}{\texttt}
\newcommand{\End}{\type{End}}
\newcommand{\Send}[1]{!#1.}
\newcommand{\Recv}[1]{?#1.}
\newcommand{\Select}{\oplus}
\newcommand{\Branch}{\&}
\newcommand{\dual}{\overline}

\newcommand{\reduce}{\rightarrow}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Machine verification of typed process calculi}
\subtitle{Research Proposal}
\author{Uma Zalakain}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\cite{empty}

\footnotesize
\bibliographystyle{alpha}
\bibliography{paper}
\end{document}

### 2020-02-07 paper/paper.tex
Some initial notes

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Machine verification of typed process calculi}
-\subtitle{Research Proposal}
+\title{Typechecking a linear \picalc}
 \author{Uma Zalakain}
@@
 
+\begin{abstract}
+  We present the syntax, operational semantics, and typing rules for a \picalc{} with linear and shared types.
+We use a typing with leftovers approach \cite{} to encode our typing rules.
+We build framing and weakening proofs that we then use to prove subject congruence and subject reduction.
+\end{abstract}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-\cite{empty}
+
+Goals: present nice model, prove subject reduction
+
+Intensional (correct by construction) vs extensional
+
+same syntax and semantics, different typing rules
+
+full formalisation available in Agda
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related work}
+
+\cite{previous-work} polymorphic tokens, HOAS
+
+\cite{typing-with-leftovers}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Syntax}
+
+variable references (strings, locally named, de Bruijn)
+
+strings to maybe de Bruijn
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Semantics}
+
+keeping track of the variable on which communication occurs
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Linear typing rules}
+
+using dependent types to encode typing rules
+
+removing from context vs keeping in context but marking it used
+
+two-layered approach: types on one hand, capabilities on the other
+
+n multiplicities as a generalisation
+
+properties of the underlaying monoid
+
+nat numbers + omega as a generalisation
+
+partial vs total
+
+\subsection{Typing with leftovers}
+
+Variable references as proofs of capability
+
+Context splits at each variable reference
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Subject reduction}
+
+\subsection{Framing}
+\subsection{Weakening}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Future work}
+
+Work that will be done time permiting:
+
+Decidable typechecking
+
+Proof of progress
+
+Product types
+
+Sum types
+
+Encoding of session types
 

### 2020-02-09 paper/paper.tex
Some more notes

@@
   We present the syntax, operational semantics, and typing rules for a \picalc{} with linear and shared types.
-We use a typing with leftovers approach \cite{} to encode our typing rules.
+We use leftover typing \cite{} to encode our typing rules.
 We build framing and weakening proofs that we then use to prove subject congruence and subject reduction.
@@
 
-Goals: present nice model, prove subject reduction
-
 Intensional (correct by construction) vs extensional
@@
 same syntax and semantics, different typing rules
+
+\subsection{Contribution}
+
+machine verified formalisation of the pi calculus
+
+typing with leftovers applied to the pi calculus
+
+unique quirks: multiple multiplicities per variable, shared types
 
@@
 
-using dependent types to encode typing rules
+\subsection{Variable references}
 
-removing from context vs keeping in context but marking it used
+\begin{description}
+  \item[Polarities]
+  \item[Multiple variables]
+  \item[Vectors]
+    Most general solution, $n$ possible multiplicities
+\end{description}
 
-two-layered approach: types on one hand, capabilities on the other
+\subsection{Multiplicities}
 
-n multiplicities as a generalisation
+$\inf$ and $\mathbb{N}$ as generalisations
 
@@
 
-nat numbers + omega as a generalisation
+\subsubsection{Total definition}
+
+\subsubsection{Partial definition}
 
@@
 
+\subsection{Variable types and multiplicities}
+
+two-layered approach: types on one hand, capabilities on the other
+removing from context vs keeping in context but marking it used
+
 \subsection{Typing with leftovers}
+
+\subsubsection{Typing relation}
 
@@
 \subsection{Framing}
+
+Definition
+Let $\Gamma \types P \boxtimes \Delta$ and $\Gamma - Delta \equiv \Xi - \phi$.
+Then $\Xi \types P \boxtimes \phi$.
+
+By defining of $\uplus$ as a total function, $\Gamma - \Delta$ is no longer
+functional: $\omega - \omega$ results in any multiplicity, including $0$. Therefore
+$\Xi \types P \boxtimes \phi$ would imply that such a variable cannot appear in $P$.
+
 \subsection{Weakening}

### 2020-02-10 paper/paper.tex
Change to LIPIcs

@@
-\documentclass{scrartcl}
+\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate,authorcolumns]{lipics-v2019}
 
@@
 \usepackage[many]{tcolorbox}
+\bibliographystyle{plainurl}
 
-% Links and their colors
-\usepackage[
-  colorlinks=true,
-  linkcolor=darkgray,
-  citecolor=darkgray,
-  urlcolor=darkgray,
-  ]{hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Typing a linear \picalc}
+\author{Uma Zalakain}{University of Glasgow, Scotland}
+       {uma.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
+\author{Ornela Dardha}{University of Glasgow, Scotland}
+       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
+\authorrunning{U. Zalakain and O. Dardha}
+\Copyright{Uma Zalakain and Ornela Dardha}
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003753.10003761.10003764</concept_id>
+<concept_desc>Theory of computation~Process calculi</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+\ccsdesc[300]{Theory of computation~Process calculi}
+\keywords{pi calculus, linear, types, concurrency}
+\supplement{\url{https://github/umazalakain/typing-linear-pi}}
+\acknowledgements{I want to thank \dots}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Typechecking a linear \picalc}
-\author{Uma Zalakain}
 \maketitle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Definition
-Let $\Gamma \types P \boxtimes \Delta$ and $\Gamma - Delta \equiv \Xi - \phi$.
-Then $\Xi \types P \boxtimes \phi$.
+Let $\Gamma \type P \boxtimes \Delta$ and $\Gamma - Delta \equiv \Xi - \phi$.
+Then $\Xi \type P \boxtimes \phi$.
 
@@
 functional: $\omega - \omega$ results in any multiplicity, including $0$. Therefore
-$\Xi \types P \boxtimes \phi$ would imply that such a variable cannot appear in $P$.
+$\Xi \type P \boxtimes \phi$ would imply that such a variable cannot appear in $P$.
 
@@
 
-\footnotesize
-\bibliographystyle{alpha}
 \bibliography{paper}

### 2020-02-24 paper/paper.tex
Proved subject congruence!!
Get Quantifiers module out, clean up imports
Add some notes on multiplicity algebras
Draft an abstract

@@
 \author{Uma Zalakain}{University of Glasgow, Scotland}
-       {uma.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
+       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
 \author{Ornela Dardha}{University of Glasgow, Scotland}
@@
 \begin{abstract}
-  We present the syntax, operational semantics, and typing rules for a \picalc{} with linear and shared types.
-We use leftover typing \cite{} to encode our typing rules.
-We build framing and weakening proofs that we then use to prove subject congruence and subject reduction.
+  We present the syntax, operational semantics, and typing rules of a \picalc{} with linear and shared types.
+  We abstract over types (using \emph{shapes} to untangle usage contexts from typing contexts) and generalize the algebras on multiplicities (using indexed sets of \emph{partial commutative monoids}).
+  We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
+  We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
+  We show that the type system is stable under substitution and prove subject reduction.
+
+  This formalization has been fully mechanized with Agda and is available at \url{https://github.com/umazalakain/typing-linear-pi}.
 \end{abstract}
@@
 
-Intensional (correct by construction) vs extensional
-
-same syntax and semantics, different typing rules
+extensional typing rules for a given syntax and operational semantics
 
@@
 
-unique quirks: multiple multiplicities per variable, shared types
+multiple multiplicities per variable
+
+multiple multiplicity types
 
@@
 
-strings to maybe de Bruijn
+allows to ignore alpha conversion, or proofs of inequality between strings
+
+strings to maybe de Bruijn, names can be kept in context as well, just not doing it
 
@@
 \section{Semantics}
+
+\subsection{Structural congruence}
+
+congruence relationship indexed by recursive tree
+
+\subsection{Reduction}
 
@@
 
-$\inf$ and $\mathbb{N}$ as generalisations
+Generalisation over indexed sets of \emph{partial commutative monoids}
 
-properties of the underlaying monoid
-
-\subsubsection{Total definition}
-
-\subsubsection{Partial definition}
-
-partial vs total
+given two multiplicities, the third is uniquely determined. this is better modelled through a function rather than an inductive type.
 
@@
 
+\subsection{Strengthening}
+
+\subsection{Swapping}
+
+\subsection{Substitution}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Decidable typechecking
+\begin{description}
 
-Proof of progress
+\item [Soundness and completeness with respect to an alternative formalization.]
 
-Product types
+\item [Proof of progress]
 
-Sum types
+\item [Product types]
 
-Encoding of session types
+\item [Sum types]
+
+\item [Decidable typechecking]
+
+\item [Encoding of session types]
+
+\end{description}
 

### 2020-03-09 paper/paper.tex
Use multiplicities with polarities, dummy
Write down some of the typing rules

@@
 \usepackage{amsmath}
+\usepackage{amsbsy}
 \usepackage[many]{tcolorbox}
 \bibliographystyle{plainurl}
+
+% Inference rules
+\usepackage{mathpartir}
+
+% TODO notes -- TODO: remove
+\usepackage{todonotes}
 
@@
 
-% Commands for the pi-calculus
+% Processes
 \newcommand{\PO}{\mathbf{0}}
-\newcommand{\comp}[2]{#1 \mid #2}
-\newcommand{\new}[2]{(\boldsymbol{\nu} #1 #2)\,}
-\newcommand{\cout}[2]{\overline{#1}\langle#2\rangle.}
-\newcommand{\cin}[2]{#1(#2).}
-\newcommand{\select}[2]{#1\triangleleft#2.}
-\newcommand{\branch}[2]{#1\triangleright#2}
+\newcommand{\comp}[2]{#1 \parallel #2}
+\newcommand{\new}{\boldsymbol{\nu} \,}
+\newcommand{\send}[2]{#1 \, \langle#2\rangle \,}
+\newcommand{\recv}[1]{#1 \, \mathbb{()} \,}
 
 \newcommand{\subst}[3]{#1[#2/#3]}
-
 \newcommand{\picalc}{$\pi$-calculus}
 \newcommand{\Picalc}{$\pi$-Calculus}
+\newcommand{\reduce}{\longrightarrow}
+\newcommand{\types}[4]{#1 \propto #2 \vdash #3 \boxtimes #4}
+\newcommand{\suc}{{\scriptscriptstyle 1+}}
 
-\newcommand{\type}{\texttt}
-\newcommand{\End}{\type{End}}
-\newcommand{\Send}[1]{!#1.}
-\newcommand{\Recv}[1]{?#1.}
-\newcommand{\Select}{\oplus}
-\newcommand{\Branch}{\&}
-\newcommand{\dual}{\overline}
-
-\newcommand{\reduce}{\rightarrow}
+% Types
+\newcommand{\Var}{\mathrm{VAR}}
+\newcommand{\Process}{\mathrm{PROCESS}}
+\newcommand{\Unused}{\mathrm{UNUSED}}
+\newcommand{\PreCtx}{\mathrm{PRECTX}}
+\newcommand{\Ctx}{\mathrm{CTX}}
+\newcommand{\Type}{\mathrm{TYPE}\;}
+\newcommand{\Idx}{\mathrm{IDX}\;}
+\newcommand{\Idxs}{\mathrm{IDXS}}
+\newcommand{\Carrier}{\mathrm{CARRIER}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\Channel}{\mathrm{CHANNEL}}
+\newcommand{\Rec}{\mathrm{REC}}
 
@@
 \maketitle
+\todo{remove todo notes}
 
@@
   We present the syntax, operational semantics, and typing rules of a \picalc{} with linear and shared types.
-  We abstract over types (using \emph{shapes} to untangle usage contexts from typing contexts) and generalize the algebras on multiplicities (using indexed sets of \emph{partial commutative monoids}).
   We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
+  We generalize the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, affine, gradual and shared typing.
   We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
@@
 
+The \picalc models communication.
+
+why resource-aware typing
+
 extensional typing rules for a given syntax and operational semantics
+
+leftover typing
 
@@
 
-machine verified formalisation of the pi calculus
+\begin{description}
+  \item [Machine verified formalisation of the linear pi calculus]
 
-typing with leftovers applied to the pi calculus
+  \item [Typing with leftovers applied to the pi calculus]
 
-multiple multiplicities per variable
+  \item [Abstraction over multiplicities]
 
-multiple multiplicity types
-
-full formalisation available in Agda
+  \item [Full formalisation available in Agda]
+\end{description}
 
@@
 
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\Var_n : Set}}
+
+  \inferrule
+  {n : \N}
+  {0 : \Var_n}
+
+  \inferrule
+  {x : \Var_n}
+  {\suc x : \Var_{\suc n}}
+\end{mathpar}
+\caption{Types of size $n$}
+\end{figure}
+
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\Process_n : Set}}
+\end{mathpar}
+
+\begin{equation*} \label{pi-calc-syntax}
+\begin{split}
+  \Process_n ::=& \; \PO_n                    \\
+              |& \; \new{} \Process_{\suc n}          \\
+              |& \; \comp{\Process_n}{\Process_n}          \\
+              |& \; \recv{\Var_n}{}\Process_{\suc n} \\
+              |& \; \send{\Var_n}{\Var_n}\Process_n
+\end{split}
+\end{equation*}
+\caption{Well-scoped grammar using de Bruijn indices}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {P \, Q : \Process_n}
+    {P \equiv_n Q : Set}}
+
+  \inferrule
+  {P \; Q \; R : \Process_n}
+  {comp-assoc : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}
+
+  \inferrule
+  {P \; Q : \Process_n}
+  {comp-sym : \comp{P}{Q} \equiv_n \comp{Q}{P}}
+
+  \inferrule
+  {P : \Process_n}
+  {comp-end : \comp{P}{\PO_n} \equiv_n P}
+
+  \inferrule
+  { }
+  {scope-end : \new \PO_{\suc n} \equiv_n \PO_n}
+
+  \inferrule
+  {P \, Q : \Process_{\suc n} \\ uQ : \Unused_0 \, Q}
+  {scope-ext : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
+
+  \inferrule
+  {P : \Process_{\suc \suc n}}
+  {scope-comm : \new \new P \equiv_n \new \new swap \; 0 \; P}
+\end{mathpar}
+\caption{Structural rewriting rules}
+\end{figure}
+
+\begin{figure}[h]
+\begin{mathpar}
+  \mprset{flushleft}
+
+  {\mprset{fraction={===}}
+    \inferrule
+    { }
+    {\Rec : Set}}
+
+  \inferrule
+  { }
+  {zero : \Rec}
+
+  \inferrule
+  {r : \Rec}
+  {one \; r : \Rec}
+
+  \inferrule
+  {r \; s : \Rec}
+  {two \; r \; s : \Rec}
+
+  {\mprset{fraction={===}}
+    \inferrule
+    {P \, Q : \Process_n \\ r : \Rec}
+    {P =_r Q : Set}}
+
+  \inferrule
+  {P \; Q : \Process_n \\ P \equiv_n Q}
+  {struct : P =_{zero} Q}
+
+  \inferrule
+  {P \; P' : \Process_{\suc n} \\ P =_r P'}
+  {cong-scope : \new P =_{one \; r} \new P'}
+
+  \inferrule
+  {P \; P' : \Process_n \\ P =_r P'}
+  {cong-comp : \comp{P}{Q} =_{one \; r} \comp{P'}{Q}}
+
+  \inferrule
+  {P \; P' : \Process_{\suc n} \\ P =_r P'}
+  {cong-recv : \recv{x}P =_{one \; r} \recv{x}P'}
+
+  \inferrule
+  {P \; P' : \Process_n \\ P =_r P'}
+  {cong-send : \send{x}{y}P =_{one \; r} \send{x}{y}P'}
+
+  \inferrule
+  {P : \Process_n}
+  {refl : P =_{zero} P}
+
+  \inferrule
+  {P \; Q : \Process_n \\ P =_r Q}
+  {sym : Q =_{one \; r} P}
+
+  \inferrule
+  {P \; Q \; R : \Process_n \\ P =_r Q \\ Q =_s R}
+  {trans : P =_{two \; r \; s} R}
+\end{mathpar}
+\caption{Structural rewriting rules lifted to a congruent equivalence relation}
+\end{figure}
+
 congruence relationship indexed by recursive tree
@@
 \subsection{Reduction}
+
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\Channel_n : Set}}
+
+  \inferrule
+  { }
+  {nothing : \Channel_n}
+
+  \inferrule
+  {i : \Var_n}
+  {just \; i : \Channel_n}
+
+  {\mprset{fraction={===}}
+    \inferrule
+    {i : \Channel_n \\ P \; Q : \Process_n}
+    {P \reduce_i Q : Set}}
+
+  \inferrule
+  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
+  {comm : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
+
+  \inferrule
+  {P \reduce_i P'}
+  {par : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
+
+  \inferrule
+  {P \reduce_i Q}
+  {res : \new P \reduce_{dec\; i} \new Q}
+
+  \inferrule
+  {P = P' \\ P' \reduce_i Q}
+  {struct : P \reduce_i Q}
+\end{mathpar}
+\caption{Operational semantics indexed by reducing channel}
+\end{figure}
 
@@
 
+\subsection{Multiplicities}
+
+A type system with both linear and shared resources has multiplicities $0$, $1$ and $\omega$.
+
 \subsection{Variable references}
 
-\begin{description}
-  \item[Polarities]
-  \item[Multiple variables]
-  \item[Vectors]
-    Most general solution, $n$ possible multiplicities
-\end{description}
+\subsection{Contexts}
 
-\subsection{Multiplicities}
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\PreCtx_n : Set}
+  }
 
-Generalisation over indexed sets of \emph{partial commutative monoids}
+  \inferrule
+  { }
+  {[] : \PreCtx_0}
 
-given two multiplicities, the third is uniquely determined. this is better modelled through a function rather than an inductive type.
+  \inferrule
+  {\gamma : \PreCtx_n \\ t : \Type}
+  {\gamma , t : \PreCtx_{\suc n}}
+\end{mathpar}
+\end{figure}
 
-\subsection{Variable types and multiplicities}
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\Idxs_n : Set}}
 
+  \inferrule
+  { }
+  {[] : \Idxs_0}
+
+  \inferrule
+  {\imath : \Idxs_n \\ x : \Idx}
+  {\imath , x : \Idxs_{\suc n}}
+
+  \\
+
+  {\mprset{fraction={===}}
+    \inferrule
+    {\imath : \Idxs n}
+    {\Ctx_\imath : Set}}
+
+  \inferrule
+  { }
+  {[] : Ctx_{[]}}
+
+  \inferrule
+  {\Gamma : \Ctx_\imath \\ c : \Carrier_x}
+  {\Gamma , c : \Ctx_{\imath , x}}
+\end{mathpar}
+\end{figure}
 two-layered approach: types on one hand, capabilities on the other
@@
 
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule{
+      \gamma : \PreCtx_n \\
+      \inferrule*{}{
+        \imath : \Idxs_n \\\\
+        \Gamma : \Ctx_\imath} \\
+      P : \Process_n \\
+      \Delta : \Ctx_\imath}
+    {\types{\gamma}{\Gamma}{P}{\Delta} : Set}}
+
+\end{mathpar}
+\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Definition
-Let $\Gamma \type P \boxtimes \Delta$ and $\Gamma - Delta \equiv \Xi - \phi$.
-Then $\Xi \type P \boxtimes \phi$.
+\subsection{Weakening}
 
-By defining of $\uplus$ as a total function, $\Gamma - \Delta$ is no longer
-functional: $\omega - \omega$ results in any multiplicity, including $0$. Therefore
-$\Xi \type P \boxtimes \phi$ would imply that such a variable cannot appear in $P$.
-
-\subsection{Weakening}
+Suffices to model as a series of insertions.
 
@@
 
-\item [Soundness and completeness with respect to an alternative formalization.]
+\item [Affine types]
 
@@
 
+\item [Soundness and completeness with respect to an alternative formalization.]
+
 \item [Encoding of session types]

### 2020-03-09 paper/paper.tex
Add process typing rules

@@
 \usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage[fleqn]{amsmath}
 \usepackage{amsbsy}
@@
 \newcommand{\types}[4]{#1 \propto #2 \vdash #3 \boxtimes #4}
+\newcommand{\contains}[5]{#1 \propto #2 \ni #3 \propto #4 \boxtimes #5}
 \newcommand{\suc}{{\scriptscriptstyle 1+}}
+\newcommand{\op}[3]{#1 := #2 \cdot #3}
 
@@
 \newcommand{\Rec}{\mathrm{REC}}
+\newcommand{\Quantifier}{\mathrm{QUANTIFIER}}
+\newcommand{\Quantifiers}{\mathrm{QUANTIFIERS}}
+\newcommand{\constr}[1]{\mathtt{#1}}
 
@@
 \end{description}
+
+\subsection{Notation}
+
+\begin{figure}[h]
+  \begin{mathpar}
+    {\mprset{fraction={===}}
+      \inferrule
+      { }
+      {\N : Set}}
+
+    \inferrule
+    { }
+    {0 : \N}
+
+    \inferrule
+    {n : \N}
+    {\suc n : \N}
+  \end{mathpar}
+  \caption{Notation used in this paper}
+\end{figure}
+
+Double rule for type-level definitions.
+We omit universe levels for brevity.
+Constructors are $\constr{teletyped}$ unless they are symbols.
+Some constructors are shared across types. They can however always be disambiguated through the type of the goal.
 
@@
     \inferrule
-    {P \, Q : \Process_n}
+    { }
     {P \equiv_n Q : Set}}
@@
   \inferrule
-  {P \; Q \; R : \Process_n}
-  {comp-assoc : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}
-
-  \inferrule
-  {P \; Q : \Process_n}
-  {comp-sym : \comp{P}{Q} \equiv_n \comp{Q}{P}}
-
-  \inferrule
-  {P : \Process_n}
-  {comp-end : \comp{P}{\PO_n} \equiv_n P}
+  { }
+  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}
 
@@
   { }
-  {scope-end : \new \PO_{\suc n} \equiv_n \PO_n}
+  {\constr{comp-sym} : \comp{P}{Q} \equiv_n \comp{Q}{P}}
+
+  \inferrule
+  { }
+  {\constr{comp-end} : \comp{P}{\PO_n} \equiv_n P}
+
+  \inferrule
+  { }
+  {\constr{scope-end} : \new \PO_{\suc n} \equiv_n \PO_n}
 
@@
   {P \, Q : \Process_{\suc n} \\ uQ : \Unused_0 \, Q}
-  {scope-ext : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
 
@@
   {P : \Process_{\suc \suc n}}
-  {scope-comm : \new \new P \equiv_n \new \new swap \; 0 \; P}
+  {\constr{scope-comm} : \new \new P \equiv_n \new \new swap \; 0 \; P}
 \end{mathpar}
-\caption{Structural rewriting rules}
+\caption{Structural rewriting rules indexed by recursion tree. $P \; Q \; R : \Process_n$ where $P$, $Q$, or $R$ have been omitted for brevity.}
 \end{figure}
@@
   { }
-  {zero : \Rec}
+  {\constr{zero} : \Rec}
 
@@
   {r : \Rec}
-  {one \; r : \Rec}
+  {\constr{one} \; r : \Rec}
 
@@
   {r \; s : \Rec}
-  {two \; r \; s : \Rec}
+  {\constr{two} \; r \; s : \Rec}
 
@@
   \inferrule
-  {P \; Q : \Process_n \\ P \equiv_n Q}
-  {struct : P =_{zero} Q}
+  {P \equiv_n Q}
+  {\constr{struct} : P =_{\constr{zero}} Q}
 
@@
   {P \; P' : \Process_{\suc n} \\ P =_r P'}
-  {cong-scope : \new P =_{one \; r} \new P'}
+  {\constr{cong-scope} : \new P =_{\constr{one} \; r} \new P'}
 
   \inferrule
-  {P \; P' : \Process_n \\ P =_r P'}
-  {cong-comp : \comp{P}{Q} =_{one \; r} \comp{P'}{Q}}
+  {P =_r P'}
+  {\constr{cong-comp} : \comp{P}{Q} =_{\constr{one} \; r} \comp{P'}{Q}}
 
@@
   {P \; P' : \Process_{\suc n} \\ P =_r P'}
-  {cong-recv : \recv{x}P =_{one \; r} \recv{x}P'}
+  {\constr{cong-recv} : \recv{x}P =_{\constr{one} \; r} \recv{x}P'}
 
   \inferrule
-  {P \; P' : \Process_n \\ P =_r P'}
-  {cong-send : \send{x}{y}P =_{one \; r} \send{x}{y}P'}
+  {P =_r P'}
+  {\constr{cong-send} : \send{x}{y}P =_{\constr{one} \; r} \send{x}{y}P'}
 
@@
   {P : \Process_n}
-  {refl : P =_{zero} P}
+  {\constr{refl} : P =_{\constr{zero}} P}
 
   \inferrule
-  {P \; Q : \Process_n \\ P =_r Q}
-  {sym : Q =_{one \; r} P}
+  {P =_r Q}
+  {\constr{sym} : Q =_{\constr{one} \; r} P}
 
   \inferrule
-  {P \; Q \; R : \Process_n \\ P =_r Q \\ Q =_s R}
-  {trans : P =_{two \; r \; s} R}
+  {P =_r Q \\ Q =_s R}
+  {\constr{trans} : P =_{\constr{two} \; r \; s} R}
 \end{mathpar}
-\caption{Structural rewriting rules lifted to a congruent equivalence relation}
+\caption{Structural rewriting rules lifted to a congruent equivalence relation. $P \; P' \; Q\; R : \Process_n$ where $P$, $P'$, $Q$, or $R$ have been omitted for brevity.}
 \end{figure}
@@
   { }
-  {nothing : \Channel_n}
+  {\constr{nothing} : \Channel_n}
 
@@
   {i : \Var_n}
-  {just \; i : \Channel_n}
+  {\constr{just} \; i : \Channel_n}
 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
-  {comm : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
 
@@
   {P \reduce_i P'}
-  {par : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
+  {\constr{par} : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
 
@@
   {P \reduce_i Q}
-  {res : \new P \reduce_{dec\; i} \new Q}
+  {\constr{res} : \new P \reduce_{dec\; i} \new Q}
 
@@
   {P = P' \\ P' \reduce_i Q}
-  {struct : P \reduce_i Q}
+  {\constr{struct} : P \reduce_i Q}
 \end{mathpar}
@@
 A type system with both linear and shared resources has multiplicities $0$, $1$ and $\omega$.
+
+\begin{figure}[h]
+\begin{equation}
+\begin{split}
+  \constr{0\cdot}          &: C \\
+  \constr{+\cdot}          &: C \\
+  \constr{-\cdot}          &: C \\
+  \constr{\op{\_}{\_}{\_}} &: C \to C \to C \to Set \\
+  \constr{1\cdot}          &: C \\
+  \constr{\cdot-join}      &: \forall \{x y z\} \to \op{x}{y}{+\cdot} \to \op{x}{z}{-\cdot} \to \exists w . (\op{x}{w}{1\cdot}) \\
+  \constr{\cdot-compute}   &: \forall y z \to Dec (\exists x . (\op{x}{y}{z})) \\
+  \constr{\cdot-compute^l} &: \forall x z \to Dec (\exists y . (\op{x}{y}{z})) \\
+  \constr{\cdot-unique}    &: \forall \{x x' y z\} \to \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+  \constr{\cdot-unique^l}  &: \forall \{x y y' z\} \to \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+  \constr{\cdot-id^l}      &: \forall x \to \op{x}{0\cdot}{x} \\
+  \constr{\cdot-comm}      &: \forall \{x y z\} \to \op{x}{y}{z} \to \op{x}{z}{y} \\
+  \constr{\cdot-assoc}     &: \forall \{x y z u v\} \to \op{x}{y}{z} \to \op{y}{u}{v} \to \exists w . (\op{x}{u}{w} \times \op{w}{v}{z})
+\end{split}
+\end{equation}
+\caption{Partial commutative monoid}
+\end{figure}
+\todo{alignment, get rid of $1\cdot$, $\cdot-join$, $\cdot-compute^l$}
+
+\begin{figure}[h]
+\begin{equation}
+\begin{split}
+  \constr{\Idx}          &: Set \\
+  \constr{\exists \Idx}  &: Idx \\
+  \constr{\Carrier}      &: \Idx \to Set \\
+  \constr{\Quantifiers}  &: \forall i : \Idx \to \Quantifier_{\Carrier_i}
+\end{split}
+\end{equation}
+\caption{Indexed set of partial commutative monoids}
+\end{figure}
+
+\subsubsection{Example type systems}
+
+Shared. Gradual. Affine. Linear.
 
@@
   \inferrule
-  {\imath : \Idxs_n \\ x : \Idx}
-  {\imath , x : \Idxs_{\suc n}}
+  {is : \Idxs_n \\ i : \Idx}
+  {is , i : \Idxs_{\suc n}}
 
@@
     \inferrule
-    {\imath : \Idxs n}
-    {\Ctx_\imath : Set}}
+    {is : \Idxs_n}
+    {\Ctx_{is} : Set}}
 
@@
   \inferrule
-  {\Gamma : \Ctx_\imath \\ c : \Carrier_x}
-  {\Gamma , c : \Ctx_{\imath , x}}
+  {\Gamma : \Ctx_{is} \\ x : \Carrier_i}
+  {\Gamma , x : \Ctx_{is , i}}
 \end{mathpar}
@@
       \inferrule*{}{
-        \imath : \Idxs_n \\\\
-        \Gamma : \Ctx_\imath} \\
+        is : \Idxs_n \\\\
+        \Gamma : \Ctx_{is}} \\
+      t : \Type \\
+      x : \Carrier_i \\
+      \Delta : \Ctx_{is}}
+    {\contains{\gamma}{\Gamma}{t}{x}{\Delta} : Set}}
+
+  \inferrule
+  {\Gamma : \Ctx_{is} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
+  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{t}{y}{\Gamma , z}}
+
+  \inferrule
+  {\Gamma : \Ctx_{is} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{is} \\\\
+    loc_x : \contains{\gamma}{\Gamma}{t}{x}{\Delta}}
+  {\constr{suc} : \contains{\gamma , t}{\Gamma , x'}{t}{x}{\Delta , x'}}
+\end{mathpar}
+\end{figure}
+
+\begin{figure}[h]
+\begin{mathpar}
+  \mprset{flushleft}
+
+  {\mprset{fraction={===}}
+    \inferrule{
+      \gamma : \PreCtx_n \\
+      \inferrule*{}{
+        is : \Idxs_n \\\\
+        \Gamma : \Ctx_{is}} \\
       P : \Process_n \\
-      \Delta : \Ctx_\imath}
+      \Delta : \Ctx_{is}}
     {\types{\gamma}{\Gamma}{P}{\Delta} : Set}}
 
+  \inferrule
+  { }
+  {\constr{end} : \types{\gamma}{\Gamma}{\PO}{\Gamma}}
+
+  \inferrule
+  {t : \Type \\ x : \Carrier_i \\ y : \Carrier_j \\\\
+   cont : \types{\gamma , C[ t \propto x ]}{\Gamma , y}{P}{\Delta , 0.}}
+  {\constr{chan} : \types{\gamma}{\Gamma}{\new P}{\Delta}}
+
+  \inferrule
+  {chan_x : \contains{\gamma}{\Gamma}{C[ t \propto x]}{+\cdot}{\Xi} \\\\
+   cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , 0\cdot}}
+  {\constr{recv} : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}
+
+  \inferrule
+  {chan_x : \contains{\gamma}{\Gamma}{C[t \propto x]}{-\cdot}{\Delta} \\\\
+   loc_y : \contains{\gamma}{\Delta}{t}{x}{\Xi} \\\\
+   cont : \types{\gamma}{\Xi}{P}{\Theta}}
+  {\constr{send} : \types{\gamma}{\Gamma}{\send{toFin \; chan_x}{toFin \; loc_y}P}{\Theta}}
+
+  \inferrule
+  {left : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
+    right : \types{\gamma}{\Delta}{Q}{\Xi}}
+  {\constr{comp} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
 \end{mathpar}
@@
 
-Suffices to model as a series of insertions.
+Order preserving embeddings model a series of insertions. We only ever need one insertion to prove subject congruence, but there is no loss of generality.
 

### 2020-03-09 paper/paper.tex
trying to see where one figure ends and the other begins...
Add draft on syntax

@@
 
-variable references (strings, locally named, de Bruijn)
+Abstraction is one of the key reasoning tools in a language: it allows for features to be defined over a range of inputs.
+Such is the case of the \picalc{} too, where both scope restriction and input introduce abstractions.
+The bodies of these constructs must have a way of referring to their argument.
+Using names as variable references is a popular option amongst humans.
+However, names are cumbersome to mechanize: inserting a new variable into an environment means proving that the name of such variable is different to all other variable names in the environment.
+Moreover, to a machine names are of no significance whatsoever.
 
-allows to ignore alpha conversion, or proofs of inequality between strings
+Machines prefer things they can algorithmically act upon. Like natural numbers!
+What does it mean to use a number as a variable reference?
+The idea de Bruijn had \cite{} was to use the index $n$ to refer to the variable introduced $n$ binders ago.
+The binders themselves introduce no names anymore.
+The expression $\lambda g . (\lambda f . f g) g$ in the $\lambda$-calculus would translate as $\lambda (\lambda 0 1) 0$.
+That is, terms at different \emph{depths} must use different indices to refer to the same binding.
+Humans find this often confusing.
+There is however no reason not to keep the original names bestowed by the humans together with the indices.
+Machines can then manipulate references mechanically and still use names to present them to humans.
 
-strings to maybe de Bruijn, names can be kept in context as well, just not doing it
+A variable occurring under $n$ abstractions has $n$ things to refer to.
+References outside of that range have no associated meaning.
+It is useful to rule out these nonsensical terms syntactically.
+In Figure \ref{var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 
@@
   {n : \N}
-  {0 : \Var_n}
+  {0 : \Var_{\suc n}}
 
@@
 \caption{Types of size $n$}
+\label{var}
 \end{figure}
+
+Every time we go under a binder, the number of binders a variable might refer to increments by one.
+To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+As shown in Figure \ref{process}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
 
@@
 \caption{Well-scoped grammar using de Bruijn indices}
+\label{process}
 \end{figure}
+
+Unlike with names, using type-level de Bruijn indices makes our syntax well-scoped by construction.
+As a consequence, the semantics of our language can be defined on the totality of the syntax.
+User-friendliness can still be recovered through a function that converts processes with names into (possibly) processes with indices.
+This function would keep track of what index is associated with what name, and would traverse the process recursively, taking note of new binders and substituting variable references.
+If the process is ill-scoped, the function would return nothing.
+To print things back to the user names can be substituted with an index together with the name.
 
@@
 \end{mathpar}
+\caption{This is...}
 \end{figure}

### 2020-03-09 paper/paper.tex
Add draft on structural congruence

@@
 \newcommand{\subst}[3]{#1[#2/#3]}
+\newcommand{\lamcalc}{$\lambda$-calculus}
 \newcommand{\picalc}{$\pi$-calculus}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related work}
-
-\cite{previous-work} polymorphic tokens, HOAS
-
-\cite{typing-with-leftovers}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Syntax}
@@
 The binders themselves introduce no names anymore.
-The expression $\lambda g . (\lambda f . f g) g$ in the $\lambda$-calculus would translate as $\lambda (\lambda 0 1) 0$.
+The expression $\lambda g . (\lambda f . f g) g$ in the \lamcalc{} would translate as $\lambda (\lambda 0 1) 0$.
 That is, terms at different \emph{depths} must use different indices to refer to the same binding.
@@
 
+In the \lamcalc{} $\beta$-reduction operates on syntactically adjacent terms.
+In the \picalc{} wowever, the syntax introduces unnecessary distinctions
+(e.g. semantically parallel composition is defined modulo associativity and commutativity).
+There are several ways around this, a structural congruence relation being one of the historical ones.
+(Others include labeled transition systems and higher inductive types.)
+
 \subsection{Structural congruence}
+
+Structural congruence is a congruent equivalence relation on processes.
+Any two structurally congruent processes are strongly bisimilar: they are can follow each other's reduction steps \cite{}.
+Figure \ref{struct-cong-base} lists the base cases of structural congruence.
 
@@
     { }
-    {P \equiv_n Q : Set}}
+    {P \equiv Q : Set}}
 
@@
   { }
-  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}
+  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \equiv \comp{\comp{P}{Q}}{R}}
 
@@
   { }
-  {\constr{comp-sym} : \comp{P}{Q} \equiv_n \comp{Q}{P}}
+  {\constr{comp-sym} : \comp{P}{Q} \equiv \comp{Q}{P}}
 
@@
   { }
-  {\constr{comp-end} : \comp{P}{\PO_n} \equiv_n P}
+  {\constr{comp-end} : \comp{P}{\PO_n} \equiv P}
 
@@
   { }
-  {\constr{scope-end} : \new \PO_{\suc n} \equiv_n \PO_n}
+  {\constr{scope-end} : \new \PO_{\suc n} \equiv \PO_n}
 
   \inferrule
-  {P \, Q : \Process_{\suc n} \\ uQ : \Unused_0 \, Q}
-  {\constr{scope-ext} : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
+  {uQ : \Unused_0 \, Q}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \equiv \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
 
   \inferrule
-  {P : \Process_{\suc \suc n}}
-  {\constr{scope-comm} : \new \new P \equiv_n \new \new swap \; 0 \; P}
+  { }
+  {\constr{scope-comm} : \new \new P \equiv \new \new swap \; 0 \; P}
 \end{mathpar}
-\caption{Structural rewriting rules indexed by recursion tree. $P \; Q \; R : \Process_n$ where $P$, $Q$, or $R$ have been omitted for brevity.}
+\caption{Structural rewriting rules. Premises $P$, $Q$ and $R$ are of type $\Process_n$ where $n$ can be inferred.}
+\label{struct-cong-base}
 \end{figure}
+
+Structural congruence is a congruent equivalence relation ship.
+As such, rewrites can happen at any point in a process' recursive definition, and they are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{struct-cong}.
+In \S \ref{subject-congruence} we will prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then $Q$ is well-typed.
+Specifically, in the case of transitivity we must prove that if $P$ is structurally congruent with $Q$ and $Q$ with $R$, and $P$ is well-typed, then so is $R$.
+To do so, we will have to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
+To show that the doubly recursive call terminates we index the equivalence relation $=$ by the type $\Rec$, which models the structure of the recursion.
 
@@
   \inferrule
-  {P \equiv_n Q}
-  {\constr{struct} : P =_{\constr{zero}} Q}
+  {eq : P \equiv Q}
+  {\constr{struct} \; eq : P =_{\constr{zero}} Q}
 
   \inferrule
-  {P \; P' : \Process_{\suc n} \\ P =_r P'}
-  {\constr{cong-scope} : \new P =_{\constr{one} \; r} \new P'}
+  {eq : P =_r P'}
+  {\constr{cong-scope} \; eq : \new P =_{\constr{one} \; r} \new P'}
 
   \inferrule
-  {P =_r P'}
-  {\constr{cong-comp} : \comp{P}{Q} =_{\constr{one} \; r} \comp{P'}{Q}}
+  {eq : P =_r P'}
+  {\constr{cong-comp} \; eq : \comp{P}{Q} =_{\constr{one} \; r} \comp{P'}{Q}}
 
   \inferrule
-  {P \; P' : \Process_{\suc n} \\ P =_r P'}
-  {\constr{cong-recv} : \recv{x}P =_{\constr{one} \; r} \recv{x}P'}
+  {eq : P =_r P'}
+  {\constr{cong-recv} \; eq : \recv{x}P =_{\constr{one} \; r} \recv{x}P'}
 
   \inferrule
-  {P =_r P'}
-  {\constr{cong-send} : \send{x}{y}P =_{\constr{one} \; r} \send{x}{y}P'}
+  {eq : P =_r P'}
+  {\constr{cong-send} \; eq : \send{x}{y}P =_{\constr{one} \; r} \send{x}{y}P'}
 
   \inferrule
-  {P : \Process_n}
+  { }
   {\constr{refl} : P =_{\constr{zero}} P}
@@
   \inferrule
-  {P =_r Q}
-  {\constr{sym} : Q =_{\constr{one} \; r} P}
+  {eq : P =_r Q}
+  {\constr{sym} \; eq : Q =_{\constr{one} \; r} P}
 
   \inferrule
-  {P =_r Q \\ Q =_s R}
-  {\constr{trans} : P =_{\constr{two} \; r \; s} R}
+  {eq_1 P =_r Q \\ \; eq_2 Q =_s R}
+  {\constr{trans} \; eq_1 \; eq_2 : P =_{\constr{two} \; r \; s} R}
 \end{mathpar}
-\caption{Structural rewriting rules lifted to a congruent equivalence relation. $P \; P' \; Q\; R : \Process_n$ where $P$, $P'$, $Q$, or $R$ have been omitted for brevity.}
+\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.
+  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred}
+\label{struct-cong}
 \end{figure}
-
-congruence relationship indexed by recursive tree
 
@@
   \inferrule
-  {P \reduce_i P'}
-  {\constr{par} : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
+  {red : P \reduce_i P'}
+  {\constr{par} \; red : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
 
   \inferrule
-  {P \reduce_i Q}
-  {\constr{res} : \new P \reduce_{dec\; i} \new Q}
+  {red : P \reduce_i Q}
+  {\constr{res} \; red : \new P \reduce_{dec\; i} \new Q}
 
   \inferrule
-  {P = P' \\ P' \reduce_i Q}
-  {\constr{struct} : P \reduce_i Q}
+  {eq : P = P' \\ red : P' \reduce_i Q}
+  {\constr{struct} \; eq \; red : P \reduce_i Q}
 \end{mathpar}
@@
     loc_x : \contains{\gamma}{\Gamma}{t}{x}{\Delta}}
-  {\constr{suc} : \contains{\gamma , t}{\Gamma , x'}{t}{x}{\Delta , x'}}
+  {\constr{suc} \; loc_x : \contains{\gamma , t}{\Gamma , x'}{t}{x}{\Delta , x'}}
 \end{mathpar}
@@
    cont : \types{\gamma , C[ t \propto x ]}{\Gamma , y}{P}{\Delta , 0.}}
-  {\constr{chan} : \types{\gamma}{\Gamma}{\new P}{\Delta}}
+  {\constr{chan} \; t \; x \; y : \types{\gamma}{\Gamma}{\new P}{\Delta}}
 
@@
    cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , 0\cdot}}
-  {\constr{recv} : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}
+  {\constr{recv} \; chan_x \; cont : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}
 
@@
    cont : \types{\gamma}{\Xi}{P}{\Theta}}
-  {\constr{send} : \types{\gamma}{\Gamma}{\send{toFin \; chan_x}{toFin \; loc_y}P}{\Theta}}
+  {\constr{send} \; chan_x \; loc_y \; cont : \types{\gamma}{\Gamma}{\send{toFin \; chan_x}{toFin \; loc_y}P}{\Theta}}
 
   \inferrule
-  {left : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
-    right : \types{\gamma}{\Delta}{Q}{\Xi}}
-  {\constr{comp} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
+  {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
+   r : \types{\gamma}{\Delta}{Q}{\Xi}}
+  {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
 \end{mathpar}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject reduction}
+\section{Subject reduction}\label{subject-reduction}
 
@@
 
+\subsection{Subject congruence}\label{subject-congruence}
+
 \subsection{Substitution}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\section{Related work}
+
+\cite{previous-work} polymorphic tokens, HOAS
+
+\cite{typing-with-leftovers}
+
+\cite{Higher-inductive-types-for-congruence}
+
+\cite{LTS-semantics}
+
+\cite{work-on-session-types}
 

### 2020-03-09 paper/paper.tex
Add bits on structural congruence and reduction

@@
 
-\subsection{Structural congruence}
+\subsection{Structural Congruence}
 
@@
 Figure \ref{struct-cong-base} lists the base cases of structural congruence.
+The type $\Unused_0 \; Q$ witnesses that index $0$ does not appear nor in the inputs nor in the outputs of process $Q$ --- it does so by traversing $Q$.
+The function $lower \; 0 \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $0$.
+Finally, $swap \; 0 \; P$ traverses $P$ (scoped under $\suc \suc n$) and swaps variable references $0$ and $\suc 0$.
 
@@
   \inferrule
-  {eq_1 P =_r Q \\ \; eq_2 Q =_s R}
+  {eq_1 : P =_r Q \\ \; eq_2 : Q =_s R}
   {\constr{trans} \; eq_1 \; eq_2 : P =_{\constr{two} \; r \; s} R}
@@
 \caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.
-  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred}
+  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred.}
 \label{struct-cong}
@@
 
-\subsection{Reduction}
+\subsection{Operational Semantics}
+
+Figure \ref{reduction} models the operational semantics of the \picalc{}.
+Processes put in parallel reduce when they communicate over a common variable.
+The receiving process substitutes references to its most immediate variable with references to the variable sent by the process doing the output, and is then \emph{lowered} --- all variable references are decreased by one.
+Reduction is closed under structural congruence and goes under parallel composition and scope restriction
+--- but notably not under input nor output, otherwise the sequencing of actions would not be preserved.
+In \S \ref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
+However, this reduction operation is effectful: it consumes the variable over which communication happens.
+If this variable is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
+To keep track of this information, we lift the variable index over which communication occurs to the type level.
+Every time we come out of a binder we decrement this variable.
+To do so, we make use of a $dec$ function that saturates at $\constr{internal}$ --- as opposed to an $inc$ function that would not.
 
@@
   { }
-  {\constr{nothing} : \Channel_n}
+  {\constr{internal} : \Channel_n}
 
@@
   {i : \Var_n}
-  {\constr{just} \; i : \Channel_n}
+  {\constr{external} \; i : \Channel_n}
 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{external \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
 
@@
 \end{mathpar}
-\caption{Operational semantics indexed by reducing channel}
+\caption{Operational semantics indexed by the channel over which reduction occurs.}
+\label{reduction}
 \end{figure}
 
-keeping track of the variable on which communication occurs
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Linear typing rules}
+\section{Resource-aware Typing System}
 
@@
 
-\subsubsection{Example type systems}
+\subsubsection{Example Type Systems}
 
 Shared. Gradual. Affine. Linear.
-
-\subsection{Variable references}
 
@@
 
-\subsection{Typing with leftovers}
-
-\subsubsection{Typing relation}
+\subsection{Typing with Leftovers}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject reduction}\label{subject-reduction}
+\section{Subject Reduction}\label{subject-reduction}
 
@@
 
-\subsection{Subject congruence}\label{subject-congruence}
+\subsection{Subject Congruence}\label{subject-congruence}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Related work}
+\section{Related Work}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Future work}
+\section{Future Work}
 

### 2020-03-15 paper/paper.tex
Some changes in nomenclature

@@
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
-\acknowledgements{I want to thank \dots}
+\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+We live in a concurrent world where any given state is often decided by the interaction between a myriad of parties --- people, machines, processors, services, networks.
+As humans, we aim to model, predict, and even construct such interactive systems.
+As mathematicians, abstraction is our tool of choice to do so.
+
 The \picalc models communication.
+
+The properties guaranteed by the picalc
 
@@
 In the \lamcalc{} $\beta$-reduction operates on syntactically adjacent terms.
-In the \picalc{} wowever, the syntax introduces unnecessary distinctions
+In the \picalc{} however, the syntax introduces unnecessary distinctions
 (e.g. semantically parallel composition is defined modulo associativity and commutativity).
@@
 
-Structural congruence is a congruent equivalence relation ship.
+Structural congruence is a congruent equivalence relation.
 As such, rewrites can happen at any point in a process' recursive definition, and they are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{struct-cong}.

### 2020-03-24 paper/paper.tex
Minor typesetting tweaks
Add some comments on capabilities

@@
 \usepackage{amsbsy}
+\usepackage{mathtools}
 \usepackage[many]{tcolorbox}
@@
 \newcommand{\suc}{{\scriptscriptstyle 1+}}
-\newcommand{\op}[3]{#1 := #2 \cdot #3}
+\newcommand{\op}[3]{#1 \coloneqq #2 \cdot #3}
 
@@
 
-\subsection{Multiplicities}
+A type system is in the business of further constraining a language, filtering out terms that (although well-formed) are meaningless.
+In the simply typed \lamcalc{}, a term has a meaning iff it has a type.
+This typing judgment is a proof of the form $\Gamma \vdash t : T$, where the term $t$ has type $T$ under the typing context $\Gamma$ -- the typing context associates variables (possibly occurring within $t$) to types.
+In the simply typed \picalc{} however, it is the channels that are typed, and a process has a meaning iff it \emph{uses} its channels according to their types.
+The typing judgment is therefore of the form $\Gamma \vdash p$, where the process $p$ (which has no type of its own) uses the channels in $\Gamma$ according to their types.
 
-A type system with both linear and shared resources has multiplicities $0$, $1$ and $\omega$.
+A resource-aware type system uses both types and usage counting to constraint a language.
+It uses typing contexts to keep track of both types and usage counts.
+In the case of a linear type system, resources must be used exactly once, and usages transition from available to consumed as soon as they are used.
+In the linear \picalc{} however, every channel must be used exactly once to receive and exactly once to send.
+In \S \ref{multiplicities}, we generalize this usage counting algebra so that the same type system can be used for a combination of linear, shared, gradual and affine types.
+
+Syntax directed
+Continuations.
+\ref{leftover-typing}
+
+\subsection{Multiplicities}\label{multiplicities}
+
+In its usual representation, the linear \picalc{} distinguishes between \emph{capabilities} -- $\ell_{\o}$ for cannot send or receive, $\ell_i$ for input, $\ell_o$ for output, and $\ell_\#$ for both input and output -- and multiplicities -- $0$ or $1$.
+This introduces unnecessary semantic duplication: the semantic meaning of a multiplicity of $0$ on a capability $c$ is independent of $c$; a multiplicity of 1 $\ell_\#$ is equivalent to a multiplicity of 1 $\ell_i$ and 1 $\ell_o$.
+In this paper, we merge multiplicities and capabilities into a single set and provide an algebra that generalizes the rules of the linear \picalc{} into rules that are suitable for shared, linear, affine and gradual typing.
+
+The algebra $\Quantifier_C$ on the set of capabilities $C$ presented in Figure \ref{capabilities} is based on a \emph{partial commutative monoid} with a few special elements.
+The operation of the partial monoid is defined as a relation $\op{x}{y}{z}$ on a carrier $C$.
+This relation must be commutative, associative, and absorb the neutral element $\ell_{\o}$.
+To recover the computability of the relation, $\constr{\cdot-compute}$ requires that given two elements $y$ and $z$, it is decidable whether a third element $x$ exists such that $\op{x}{y}{z}$.
+To recover the functionality, $\constr{\cdot-unique}$ requires that given any two elements $y$ and $z$, if a third one exists, it is uniquely defined.
+A consequence of using leftover typing (\S \ref{leftover-typing}) is that this algebra needs to be injective, and thus $\constr{\cdot-unique^l}$ requires for an operand to be uniquely defined as well.
+In addition to these rules, we define the special elements $\ell_i$ and $\ell_o$, to be subtracted every time an input or output occurs, respectively.
+Finally, we require for $\ell_i$ and $\ell_o$ to be compatible, and refer to the result as $\ell_\#$.
 
@@
 \begin{equation}
-\begin{split}
-  \constr{0\cdot}          &: C \\
-  \constr{+\cdot}          &: C \\
-  \constr{-\cdot}          &: C \\
-  \constr{\op{\_}{\_}{\_}} &: C \to C \to C \to Set \\
-  \constr{1\cdot}          &: C \\
-  \constr{\cdot-join}      &: \forall \{x y z\} \to \op{x}{y}{+\cdot} \to \op{x}{z}{-\cdot} \to \exists w . (\op{x}{w}{1\cdot}) \\
-  \constr{\cdot-compute}   &: \forall y z \to Dec (\exists x . (\op{x}{y}{z})) \\
-  \constr{\cdot-compute^l} &: \forall x z \to Dec (\exists y . (\op{x}{y}{z})) \\
-  \constr{\cdot-unique}    &: \forall \{x x' y z\} \to \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-  \constr{\cdot-unique^l}  &: \forall \{x y y' z\} \to \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  \constr{\cdot-id^l}      &: \forall x \to \op{x}{0\cdot}{x} \\
-  \constr{\cdot-comm}      &: \forall \{x y z\} \to \op{x}{y}{z} \to \op{x}{z}{y} \\
-  \constr{\cdot-assoc}     &: \forall \{x y z u v\} \to \op{x}{y}{z} \to \op{y}{u}{v} \to \exists w . (\op{x}{u}{w} \times \op{w}{v}{z})
-\end{split}
+\begin{aligned}
+  &\ell_{\o}                &:{} &                      &     & C \\
+  &\ell_i                   &:{} &                      &     & C \\
+  &\ell_o                   &:{} &                      &     & C \\
+  &\op{\_}{\_}{\_}          &:{} &                      &     & C \to C \to C \to Set \\
+  &\constr{\cdot-join}      &:{} &                      &     & \exists \ell_\# . \; \op{\ell_\#}{\ell_i}{\ell_o} \\
+  &\constr{\cdot-compute}   &:{} &\forall y z           & \to & Dec \; (\exists x . \; (\op{x}{y}{z})) \\
+  &\constr{\cdot-unique}    &:{} &\forall \{x x' y z\}  & \to & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+  &\constr{\cdot-unique^l}  &:{} &\forall \{x y y' z\}  & \to & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+  &\constr{\cdot-id^l}      &:{} &\forall x             & \to & \op{x}{\ell_{\o}}{x} \\
+  &\constr{\cdot-comm}      &:{} &\forall \{x y z\}     & \to & \op{x}{y}{z} \to \op{x}{z}{y} \\
+  &\constr{\cdot-assoc}     &:{} &\forall \{x y z u v\} & \to & \op{x}{y}{z} \to \op{y}{u}{v} \to \exists w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+\end{aligned}
 \end{equation}
-\caption{Partial commutative monoid}
+\caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid with special elements $\ell_i$, $\ell_o$ and $\ell_\#$.}
+\label{capabilities}
 \end{figure}
-\todo{alignment, get rid of $1\cdot$, $\cdot-join$, $\cdot-compute^l$}
+
+To allow for several capability algebras to coexist on the same type system, in Figure \ref{indexed-capabilities} we index the algebra carriers by an index $\Idx$.
+Such an index is a type with one element for each algebra of our type system.
+To be able to prove subject congruence, we require for at least one such algebra to exist.
+Finally, the field $\constr{\Quantifiers}$ requires for each index $i$ to have an algebra defined on the carrier set $\constr{\Carrier_i}$.
 
@@
 \caption{Indexed set of partial commutative monoids}
+\label{indexed-capabilities}
 \end{figure}
@@
 
-Shared. Gradual. Affine. Linear.
+\paragraph{Shared}
+
+The carrier $C$ is implemented as a type $\constr{One}$ with a single trivial constructor $\omega$.
+All of $\ell_{\o}$, $\ell_i$, $\ell_o$ and $\ell_\#$ are instantiated to $\omega$.
+The relation $\op{\omega}{\omega}{\omega}$ is inhabited.
+All other laws are trivially satisfied.
+
+\paragraph{Linear}
+
+The carrier $C$ is implemented as a type $\constr{Four}$ with the trivial elements $\ell_{\o}$, $\ell_i$, $\ell_o$ and $\ell_\#$.
+The partial monoid is defined with the element $\ell_{\o}$ being neutral on both sides, and the element $\ell_\#$ splitting into $\ell_i$ and $\ell_o$ or $\ell_o$ and $\ell_i$ -- and is uninhabited in every other case.
+All other rules follow trivially.
+
+\paragraph{Affine}
+
+The algebra of an affine system is exactly the one of a linear one, the only difference is the output context must satisfy a trivial predicate.
+
+\paragraph{Gradual}
+
+The carrier $C$ is implemented as a tuple $(\N , \N)$ of natural numbers.
+The element $\ell_{\o}$ corresponds to the tuple $(0 , 0)$, $\ell_i$ to $(1 , 0)$, $\ell_o$ to $(0 , 1)$, and $\ell_\#$ to $(1 , 1)$.
+The partial monoid is defined exactly when $\op{(x_l + y_l , x_r + y_r)}{(x_l , x_r)}{(y_l , y_r)}$.
+All other rules follow trivially from the algebraic rules for the addition of naturals.
 
@@
 
-\subsection{Typing with Leftovers}
+\subsection{Typing with Leftovers}\label{leftover typing}
 
@@
 \end{figure}
+
+\todo{comment on scope restriction rule and capabilities}
 

### 2020-03-24 paper/paper.tex
Add some multiplicity exposition, and some colors as well

--- 2020-03-15 paper/paper.tex
@@
 
-% Processes
-\newcommand{\PO}{\mathbf{0}}
-\newcommand{\comp}[2]{#1 \parallel #2}
-\newcommand{\new}{\boldsymbol{\nu} \,}
-\newcommand{\send}[2]{#1 \, \langle#2\rangle \,}
-\newcommand{\recv}[1]{#1 \, \mathbb{()} \,}
-
-\newcommand{\subst}[3]{#1[#2/#3]}
 \newcommand{\lamcalc}{$\lambda$-calculus}
@@
 \newcommand{\Picalc}{$\pi$-Calculus}
-\newcommand{\reduce}{\longrightarrow}
-\newcommand{\types}[4]{#1 \propto #2 \vdash #3 \boxtimes #4}
-\newcommand{\contains}[5]{#1 \propto #2 \ni #3 \propto #4 \boxtimes #5}
-\newcommand{\suc}{{\scriptscriptstyle 1+}}
-\newcommand{\op}[3]{#1 := #2 \cdot #3}
+
+\newcommand{\type}[1]{\textcolor{blue}{\mathtt{\operatorname{#1}}}}
+\newcommand{\constr}[1]{\textcolor{olive}{\mathtt{\operatorname{#1}}}}
+\newcommand{\field}[1]{\textcolor{red}{\mathtt{\operatorname{#1}}}}
+
+% Constructors
+\newcommand{\PO}{\constr{\mathbf{0}}}
+\newcommand{\comp}[2]{#1 \; \constr{\parallel} \; #2}
+\newcommand{\new}{\constr{\boldsymbol{\nu}} \;}
+\newcommand{\send}[2]{#1 \; \constr{\langle}#2\constr{\rangle} \;}
+\newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
+\newcommand{\suc}{\constr{\scriptstyle 1+}}
+
+% Functions
+\newcommand{\subst}[3]{#1[#2/#3]}
+
+% Fields
+\newcommand{\op}[3]{#1 \; \field{\coloneqq} \; #2 \; \field{\cdot} \; #3}
+\newcommand{\li}{\field{\ell_i}}
+\newcommand{\lo}{\field{\ell_o}}
+\newcommand{\lz}{\field{\ell_{\o}}}
+\newcommand{\lio}{\field{\ell_{\#}}}
 
 % Types
-\newcommand{\Var}{\mathrm{VAR}}
-\newcommand{\Process}{\mathrm{PROCESS}}
-\newcommand{\Unused}{\mathrm{UNUSED}}
-\newcommand{\PreCtx}{\mathrm{PRECTX}}
-\newcommand{\Ctx}{\mathrm{CTX}}
-\newcommand{\Type}{\mathrm{TYPE}\;}
-\newcommand{\Idx}{\mathrm{IDX}\;}
-\newcommand{\Idxs}{\mathrm{IDXS}}
-\newcommand{\Carrier}{\mathrm{CARRIER}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Channel}{\mathrm{CHANNEL}}
-\newcommand{\Rec}{\mathrm{REC}}
-\newcommand{\Quantifier}{\mathrm{QUANTIFIER}}
-\newcommand{\Quantifiers}{\mathrm{QUANTIFIERS}}
-\newcommand{\constr}[1]{\mathtt{#1}}
+\newcommand{\reduce}[1]{\; \type{\longrightarrow}_{#1} \;}
+\newcommand{\types}[4]{#1 \; \type{\propto} \; #2 \; \type{\vdash} \; #3 \; \type{\boxtimes} \; #4}
+\newcommand{\contains}[5]{#1 \; \type{\propto} \; #2 \; \type{\ni} \; #3 \; \type{\propto} \; #4 \; \type{\boxtimes} \; #5}
+\newcommand{\Var}{\type{VAR}}
+\newcommand{\Process}{\type{PROCESS}}
+\newcommand{\Unused}{\type{UNUSED}}
+\newcommand{\PreCtx}{\type{PRECTX}}
+\newcommand{\Ctx}{\type{CTX}}
+\newcommand{\Type}{\type{TYPE}\;}
+\newcommand{\Idx}{\type{IDX}\;}
+\newcommand{\Idxs}{\type{IDXS}}
+\newcommand{\Carrier}{\type{CARRIER}}
+\newcommand{\N}{\type{\mathbb{N}}}
+\newcommand{\Channel}{\type{CHANNEL}}
+\newcommand{\Rec}{\type{REC}}
+\newcommand{\Quantifier}{\type{QUANTIFIER}}
+\newcommand{\eq}[1]{\; \type{=}_{#1} \;}
+\newcommand{\eqeq}{\; \type{\equiv} \;}
 
@@
   We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
-  We generalize the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, affine, gradual and shared typing.
+  We generalize the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, gradual and shared typing.
   We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
@@
       { }
-      {\N : Set}}
+      {\type{\N} : \type{Set}}}
 
@@
     { }
-    {0 : \N}
+    {\constr{0} : \type{\N}}
 
     \inferrule
-    {n : \N}
-    {\suc n : \N}
+    {n : \type{\N}}
+    {\suc n : \type{\N}}
   \end{mathpar}
@@
 We omit universe levels for brevity.
-Constructors are $\constr{teletyped}$ unless they are symbols.
+$\type{TYPES}$ are in blue and uppercased (and their indices are subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red and variables and functions in black.
 Some constructors are shared across types. They can however always be disambiguated through the type of the goal.
@@
   {n : \N}
-  {0 : \Var_{\suc n}}
+  {\constr{0} : \Var_{\suc n}}
 
@@
     { }
-    {P \equiv Q : Set}}
+    {P \eqeq Q : Set}}
 
@@
   { }
-  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \equiv \comp{\comp{P}{Q}}{R}}
+  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \eqeq \comp{\comp{P}{Q}}{R}}
 
@@
   { }
-  {\constr{comp-sym} : \comp{P}{Q} \equiv \comp{Q}{P}}
+  {\constr{comp-sym} : \comp{P}{Q} \eqeq \comp{Q}{P}}
 
@@
   { }
-  {\constr{comp-end} : \comp{P}{\PO_n} \equiv P}
+  {\constr{comp-end} : \comp{P}{\PO_n} \eqeq P}
 
@@
   { }
-  {\constr{scope-end} : \new \PO_{\suc n} \equiv \PO_n}
+  {\constr{scope-end} : \new \PO_{\suc n} \eqeq \PO_n}
 
@@
   {uQ : \Unused_0 \, Q}
-  {\constr{scope-ext} : \new (\comp{P}{Q}) \equiv \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
 
@@
   { }
-  {\constr{scope-comm} : \new \new P \equiv \new \new swap \; 0 \; P}
+  {\constr{scope-comm} : \new \new P \eqeq \new \new swap \; 0 \; P}
 \end{mathpar}
@@
     {P \, Q : \Process_n \\ r : \Rec}
-    {P =_r Q : Set}}
+    {P \eq{r} Q : Set}}
 
   \inferrule
-  {eq : P \equiv Q}
-  {\constr{struct} \; eq : P =_{\constr{zero}} Q}
+  {eq : P \eqeq Q}
+  {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
 
   \inferrule
-  {eq : P =_r P'}
-  {\constr{cong-scope} \; eq : \new P =_{\constr{one} \; r} \new P'}
+  {eq : P \eq{r} P'}
+  {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}
 
   \inferrule
-  {eq : P =_r P'}
-  {\constr{cong-comp} \; eq : \comp{P}{Q} =_{\constr{one} \; r} \comp{P'}{Q}}
+  {eq : P \eq{r} P'}
+  {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}
 
   \inferrule
-  {eq : P =_r P'}
-  {\constr{cong-recv} \; eq : \recv{x}P =_{\constr{one} \; r} \recv{x}P'}
+  {eq : P \eq{r} P'}
+  {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}
 
   \inferrule
-  {eq : P =_r P'}
-  {\constr{cong-send} \; eq : \send{x}{y}P =_{\constr{one} \; r} \send{x}{y}P'}
+  {eq : P \eq{r} P'}
+  {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
 
@@
   { }
-  {\constr{refl} : P =_{\constr{zero}} P}
+  {\constr{refl} : P \eq{\constr{zero}} P}
 
   \inferrule
-  {eq : P =_r Q}
-  {\constr{sym} \; eq : Q =_{\constr{one} \; r} P}
+  {eq : P \eq{r} Q}
+  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
 
   \inferrule
-  {eq_1 : P =_r Q \\ \; eq_2 : Q =_s R}
-  {\constr{trans} \; eq_1 \; eq_2 : P =_{\constr{two} \; r \; s} R}
+  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
+  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
 \end{mathpar}
@@
     {i : \Channel_n \\ P \; Q : \Process_n}
-    {P \reduce_i Q : Set}}
+    {P \reduce{i} Q : Set}}
 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{external \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
 
   \inferrule
-  {red : P \reduce_i P'}
-  {\constr{par} \; red : \comp{P}{Q} \reduce_i \comp{P'}{Q}}
+  {red : P \reduce{i} P'}
+  {\constr{par} \; red : \comp{P}{Q} \reduce{i} \comp{P'}{Q}}
 
   \inferrule
-  {red : P \reduce_i Q}
-  {\constr{res} \; red : \new P \reduce_{dec\; i} \new Q}
+  {red : P \reduce{i} Q}
+  {\constr{res} \; red : \new P \reduce{dec\; i} \new Q}
 
   \inferrule
-  {eq : P = P' \\ red : P' \reduce_i Q}
-  {\constr{struct} \; eq \; red : P \reduce_i Q}
+  {eq : P = P' \\ red : P' \reduce{i} Q}
+  {\constr{struct} \; eq \; red : P \reduce{i} Q}
 \end{mathpar}
@@
 
-\subsection{Multiplicities}
+A type system is in the business of further constraining a language, filtering out terms that (although well-formed) are meaningless.
+In the simply typed \lamcalc{}, a term has a meaning iff it has a type.
+This typing judgment is a proof of the form $\Gamma \vdash t : T$, where the term $t$ has type $T$ under the typing context $\Gamma$ -- the typing context associates variables (possibly occurring within $t$) to types.
+In the simply typed \picalc{} however, it is the channels that are typed, and a process has a meaning iff it \emph{uses} its channels according to their types.
+The typing judgment is therefore of the form $\Gamma \vdash p$, where the process $p$ (which has no type of its own) uses the channels in $\Gamma$ according to their types.
 
-A type system with both linear and shared resources has multiplicities $0$, $1$ and $\omega$.
+A resource-aware type system uses both types and usage counting to constraint a language.
+It uses typing contexts to keep track of both types and usage counts.
+In the case of a linear type system, resources must be used exactly once, and usages transition from available to consumed as soon as they are used.
+In the linear \picalc{} however, every channel must be used exactly once to receive and exactly once to send.
+In \S \ref{multiplicities}, we generalize this usage counting algebra so that the same type system can be used for a combination of linear, shared, and gradual types.
+
+Syntax directed
+Continuations.
+\ref{leftover-typing}
+
+\subsection{Multiplicities and Capabilities}\label{multiplicities}
+
+In its usual representation, the linear \picalc{} distinguishes between \emph{capabilities} -- $\ell_{\o}$ for cannot send nor receive, $\ell_i$ for input, $\ell_o$ for output, and $\ell_\#$ for both input and output -- and multiplicities -- $0$ or $1$.
+This introduces some unnecessary semantic duplication: a channel with multiplicity 1 and capability $\cdot \ell_{\o}$ is indistinguishable, in usage, from a channel with multiplicity 0.
+In this paper, we merge multiplicities and capabilities into a single type and define the algebraic rules for a generalized resource-aware type system on the \picalc{}.
+As such, it can be used to model shared, linear, or gradual types.
+
+The algebra has a single binary operator $(\cdot)$ that crushes together capabilities, and must satisfy the following characteristics:
+\begin{description}
+
+\item [Partiality]
+  Not all capabilities are necessarily compatible: in the linear \picalc{} the capability $\ell_i$ cannot be split into $\ell_i$ and $\ell_i$.
+\item [Functionality]
+  Given two capabilities $y$ and $z$, it must be computable whether a third capability $x$ exists such that $\op{x}{y}{z}$.
+  If such a $x$ exists, it must be unique.
+
+\item [Injectivity]
+  To prove framing (\S \ref{framing}) we need to be able to compute the \emph{subtraction} of capabilities.
+  That is, $x - z \equiv y$ where $y$ is uniquely determined iff $\op{x}{y}{z}$ where $y$ is uniquely determined.
+
+\item [Associativity]
+  As a direct consequence of the structural congruence rule $\comp{P}{\comp{Q}{R}} \equiv \comp{\comp{P}{Q}}{R}$.
+
+\item [Commutativity]
+  As a direct consequence of the structural congruence rule $\comp{P}{Q} \equiv \comp{Q}{P}$.
+
+\item [Distinguished elements]
+  There must be a neutral element $\ell_{\o}$ absorbed on both sides.
+  There need to be two special elements $\ell_i$ and $\ell_o$ that will be subtracted every time input and output (respectively) happens.
+  These two elements need to be compatible, and we will refer to the result as $\ell_\#$.
+\end{description}
+
+In Figure \ref{capabilities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation, satisfying all the properties mentioned above.
+
+\begin{figure}[h]
+\begin{equation}
+\begin{aligned}
+  &\lz                    &:{} &                      &     & C \\
+  &\li                    &:{} &                      &     & C \\
+  &\lo                    &:{} &                      &     & C \\
+  &\op{\_}{\_}{\_}        &:{} &                      &     & C \to C \to C \to Set \\
+  &\field{\cdot-join}     &:{} &                      &     & \type{\exists} \lio . \; \op{\lio}{\li}{\lo} \\
+  &\field{\cdot-compute}  &:{} &\forall y z           & \to & \type{Dec} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+  &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+  &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+  &\field{\cdot-id^l}     &:{} &\forall x             & \to & \op{x}{\lz}{x} \\
+  &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to & \op{x}{y}{z} \to \op{x}{z}{y} \\
+  &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+\end{aligned}
+\end{equation}
+\caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid with special elements $\ell_i$, $\ell_o$ and $\ell_\#$.}
+\label{capabilities}
+\end{figure}
+
+To allow for several capability algebras to coexist on the same type system, in Figure \ref{indexed-capabilities} we index the algebra carriers by an index $\Idx$.
+Such an index is a type with one element for each algebra of our type system.
+To be able to prove subject congruence, we require for at least one such algebra to exist.
+Finally, the field $\field{QUANTIFIERS}$ requires for each index $i$ to have an algebra defined on the carrier set $\constr{\Carrier_i}$.
 
@@
 \begin{split}
-  \constr{0\cdot}          &: C \\
-  \constr{+\cdot}          &: C \\
-  \constr{-\cdot}          &: C \\
-  \constr{\op{\_}{\_}{\_}} &: C \to C \to C \to Set \\
-  \constr{1\cdot}          &: C \\
-  \constr{\cdot-join}      &: \forall \{x y z\} \to \op{x}{y}{+\cdot} \to \op{x}{z}{-\cdot} \to \exists w . (\op{x}{w}{1\cdot}) \\
-  \constr{\cdot-compute}   &: \forall y z \to Dec (\exists x . (\op{x}{y}{z})) \\
-  \constr{\cdot-compute^l} &: \forall x z \to Dec (\exists y . (\op{x}{y}{z})) \\
-  \constr{\cdot-unique}    &: \forall \{x x' y z\} \to \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-  \constr{\cdot-unique^l}  &: \forall \{x y y' z\} \to \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  \constr{\cdot-id^l}      &: \forall x \to \op{x}{0\cdot}{x} \\
-  \constr{\cdot-comm}      &: \forall \{x y z\} \to \op{x}{y}{z} \to \op{x}{z}{y} \\
-  \constr{\cdot-assoc}     &: \forall \{x y z u v\} \to \op{x}{y}{z} \to \op{y}{u}{v} \to \exists w . (\op{x}{u}{w} \times \op{w}{v}{z})
-\end{split}
-\end{equation}
-\caption{Partial commutative monoid}
-\end{figure}
-\todo{alignment, get rid of $1\cdot$, $\cdot-join$, $\cdot-compute^l$}
-
-\begin{figure}[h]
-\begin{equation}
-\begin{split}
-  \constr{\Idx}          &: Set \\
-  \constr{\exists \Idx}  &: Idx \\
-  \constr{\Carrier}      &: \Idx \to Set \\
-  \constr{\Quantifiers}  &: \forall i : \Idx \to \Quantifier_{\Carrier_i}
+  \field{IDX}          &: Set \\
+  \field{\exists IDX}  &: \field{IDX} \\
+  \field{CARRIER}      &: \field{IDX} \to Set \\
+  \field{QUANTIFIERS}  &: (i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
 \end{split}
@@
 
-Shared. Gradual. Affine. Linear.
+\begin{description}
+
+\item [Shared]
+
+The carrier $C$ is implemented as a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
+All of $\lz$, $\li$, $\lo$ and $\lio$ are instantiated to $\constr{\omega}$.
+The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
+All other laws are trivially satisfied.
+
+\item [Linear]
+
+The carrier $C$ is implemented as a type $\type{Four}$ with the trivial elements $\constr{\ell_{\o}}$, $constr{\ell_i}$, $\constr{\ell_o}$ and $\constr{\ell_\#}$.
+The monoidal operation has the element $\constr{\ell_{\o}}$ as neutral on both sides, and the element $\constr{\ell_\#}$ splitting into $\constr{\ell_i}$ and $\constr{\ell_o}$ (or $\constr{\ell_o}$ and $\constr{\ell_i}$ -- and is uninhabited in every other case.
+All other rules follow trivially.
+
+\item [Gradual]
+
+The carrier $C$ is implemented as a tuple $\N \type{\times} \N$ of natural numbers.
+The element $\lz$ corresponds to the tuple $(\constr{0} \constr{,} \constr{0})$, $\li$ to $(\constr{1} \constr{,} \constr{0})$, $\lo$ to $(\constr{0} \constr{,} \constr{1})$, and $\lio$ to $(\constr{1} \constr{,} \constr{1})$.
+The partial monoid is defined exactly when $\op{(x_l + y_l \constr{,} x_r + y_r)}{(x_l \constr{,} x_r)}{(y_l \constr{,} y_r)}$.
+All other rules follow trivially from the algebraic rules for the addition of naturals.
+
+\end{description}
 

### 2020-03-25 paper/paper.tex
Add index to capability proof

@@
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
+\newcommand{\base}[1]{\constr{B[} \; #1 \; \constr{]}}
+\newcommand{\channel}[2]{\constr{C[} \; #1 \; \constr{\propto} \; #2 \; \constr{]}}
 
@@
 % Types
+\newcommand{\Set}{\type{SET}}
 \newcommand{\reduce}[1]{\; \type{\longrightarrow}_{#1} \;}
@@
       { }
-      {\type{\N} : \type{Set}}}
+      {\type{\N} : \Set}}
 
@@
     {n : \N}
-    {\Var_n : Set}}
+    {\Var_n : \Set}}
 
@@
     {n : \N}
-    {\Process_n : Set}}
+    {\Process_n : \Set}}
 \end{mathpar}
@@
     { }
-    {P \eqeq Q : Set}}
+    {P \eqeq Q : \Set}}
 
@@
     { }
-    {\Rec : Set}}
+    {\Rec : \Set}}
 
@@
     {P \, Q : \Process_n \\ r : \Rec}
-    {P \eq{r} Q : Set}}
+    {P \eq{r} Q : \Set}}
 
@@
     {n : \N}
-    {\Channel_n : Set}}
+    {\Channel_n : \Set}}
 
@@
     {i : \Channel_n \\ P \; Q : \Process_n}
-    {P \reduce{i} Q : Set}}
+    {P \reduce{i} Q : \Set}}
 
@@
   &\lo                    &:{} &                      &        & C \\
-  &\op{\_}{\_}{\_}        &:{} &                      &     & C \to C \to C \to Set \\
+  &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
   &\field{\cdot-join}     &:{} &                      &        & \type{\exists} \lio . \; \op{\lio}{\li}{\lo} \\
-  &\field{\cdot-compute}  &:{} &\forall y z           & \to & \type{Dec} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
-  &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-  &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  &\field{\cdot-id^l}     &:{} &\forall x             & \to & \op{x}{\lz}{x} \\
-  &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to & \op{x}{y}{z} \to \op{x}{z}{y} \\
-  &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+  &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+  &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+  &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+  &\field{\cdot-id^l}     &:{} &\forall x             & \to \; & \op{x}{\lz}{x} \\
+  &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+  &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
 \end{aligned}
@@
 \begin{split}
-  \field{IDX}          &: Set \\
+  \field{IDX}          &: \Set \\
   \field{\exists IDX}  &: \field{IDX} \\
-  \field{CARRIER}      &: \field{IDX} \to Set \\
+  \field{CARRIER}      &: \field{IDX} \to \Set \\
   \field{QUANTIFIERS}  &: (i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
@@
 
+\todo{mention types are independent from capabilities}
+\todo{add index to in-type}
+\todo{update examples}
+
 \subsection{Contexts}
@@
     \inferrule
+    { }
+    {\Type : \Set}
+  }
+
+  \inferrule
   {n : \N}
-    {\PreCtx_n : Set}
+  {\base{n} : \Type}
+
+  \inferrule
+  {t : \Type \\ \inferrule*{} {i : \Idx \\\\ x : \Carrier_i}}
+  {\channel{t}{x} : \Type}
+\end{mathpar}
+\caption{This is...}
+\end{figure}
+
+\begin{figure}[h]
+\begin{mathpar}
+  {\mprset{fraction={===}}
+    \inferrule
+    {n : \N}
+    {\PreCtx_n : \Set}
   }
@@
     {n : \N}
-    {\Idxs_n : Set}}
+    {\Idxs_n : \Set}}
 
@@
     {is : \Idxs_n}
-    {\Ctx_{is} : Set}}
+    {\Ctx_{is} : \Set}}
 
@@
       \Delta : \Ctx_{is}}
-    {\contains{\gamma}{\Gamma}{t}{x}{\Delta} : Set}}
+    {\contains{\gamma}{\Gamma}{t}{x}{\Delta} : \Set}}
 
@@
       \Delta : \Ctx_{is}}
-    {\types{\gamma}{\Gamma}{P}{\Delta} : Set}}
+    {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}}
 
@@
   {t : \Type \\ x : \Carrier_i \\ y : \Carrier_j \\\\
-   cont : \types{\gamma , C[ t \propto x ]}{\Gamma , y}{P}{\Delta , 0.}}
-  {\constr{chan} \; t \; x \; y : \types{\gamma}{\Gamma}{\new P}{\Delta}}
+   cont : \types{\gamma , \channel{t}{x}}{\Gamma , y}{P}{\Delta , \lz}}
+  {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}
 
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{C[ t \propto x]}{+\cdot}{\Xi} \\\\
-   cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , 0\cdot}}
+  {chan_x : \contains{\gamma}{\Gamma}{\channel{t}{x}}{\li}{\Xi} \\\\
+   cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , \lz}}
   {\constr{recv} \; chan_x \; cont : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}
@@
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{C[t \propto x]}{-\cdot}{\Delta} \\\\
+  {chan_x : \contains{\gamma}{\Gamma}{\channel{t}{x}}{\lo}{\Delta} \\\\
    loc_y : \contains{\gamma}{\Delta}{t}{x}{\Xi} \\\\

### 2020-03-25 paper/paper.tex
Use commonly used symbols for capabilities

@@
 \newcommand{\new}{\constr{\boldsymbol{\nu}} \;}
-\newcommand{\send}[2]{#1 \; \constr{\langle}#2\constr{\rangle} \;}
+\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
@@
 \newcommand{\types}[4]{#1 \; \type{\propto} \; #2 \; \type{\vdash} \; #3 \; \type{\boxtimes} \; #4}
-\newcommand{\contains}[5]{#1 \; \type{\propto} \; #2 \; \type{\ni} \; #3 \; \type{\propto} \; #4 \; \type{\boxtimes} \; #5}
+\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{[} \; #3 \; \type{]\coloneqq} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
 \newcommand{\Var}{\type{VAR}}
@@
       \inferrule*{}{
-        is : \Idxs_n \\\\
-        \Gamma : \Ctx_{is}} \\
+        idxs : \Idxs_n \\\\
+        \Gamma : \Ctx_{idxs}} \\
+      i : Fin n
       t : \Type \\
@@
       \Delta : \Ctx_{is}}
-    {\contains{\gamma}{\Gamma}{t}{x}{\Delta} : \Set}}
+    {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}}
 
   \inferrule
-  {\Gamma : \Ctx_{is} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
-  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{t}{y}{\Gamma , z}}
+  {\Gamma : \Ctx_{idxs} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
+  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{zero}{t}{y}{\Gamma , z}}
 
   \inferrule
-  {\Gamma : \Ctx_{is} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{is} \\\\
-    loc_x : \contains{\gamma}{\Gamma}{t}{x}{\Delta}}
-  {\constr{suc} \; loc_x : \contains{\gamma , t}{\Gamma , x'}{t}{x}{\Delta , x'}}
+  {\Gamma : \Ctx_{idxs} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{idxs} \\\\
+    loc_x : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
+  {\constr{suc} \; loc_x : \contains{\gamma , t}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
 \end{mathpar}
@@
       \inferrule*{}{
-        is : \Idxs_n \\\\
-        \Gamma : \Ctx_{is}} \\
+        idxs : \Idxs_n \\\\
+        \Gamma : \Ctx_{idxs}} \\
       P : \Process_n \\
-      \Delta : \Ctx_{is}}
+      \Delta : \Ctx_{idxs}}
     {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}}
@@
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{\channel{t}{x}}{\li}{\Xi} \\\\
+  {chan_x : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
    cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , \lz}}
-  {\constr{recv} \; chan_x \; cont : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}
+  {\constr{recv} \; chan_x \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
 
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{\channel{t}{x}}{\lo}{\Delta} \\\\
-   loc_y : \contains{\gamma}{\Delta}{t}{x}{\Xi} \\\\
+  {chan_x : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+   loc_y : \contains{\gamma}{\Delta}{j}{t}{x}{\Xi} \\\\
    cont : \types{\gamma}{\Xi}{P}{\Theta}}
-  {\constr{send} \; chan_x \; loc_y \; cont : \types{\gamma}{\Gamma}{\send{toFin \; chan_x}{toFin \; loc_y}P}{\Theta}}
+  {\constr{send} \; chan_x \; loc_y \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
 

### 2020-03-26 paper/paper.tex
Update examples
note on capability and multiplicity

@@
 
+\todo{Ornela: in linear pi if you write $\li$ the $\ell$ already means that the multiplicity is 1, meaning linear. 
+We need to distinguish between \emph{capabilities} i, o or $\sharp$ and \emph{multiplicities} 0, 1 or $\omega$.
+So, the $\ell$ cannot be $\omega$, hence shared.
+You should use only i, o or $\sharp$ in Fig 7.
+and use $\ell$ as a syntactic sugar for i, o or $\sharp$ with multiplicity 1.}
+
 \begin{figure}[h]
@@
 \todo{mention types are independent from capabilities}
-\todo{add index to in-type}
-\todo{update examples}
 

### 2020-03-26 paper/paper.tex
Minor fixes

--- 2020-03-25 paper/paper.tex
@@
 \bibliographystyle{plainurl}
+
+% Less margins around figures
+\usepackage{changepage}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\stacked}[1]{\mprset{flushleft} \inferrule*{}{#1}}
+\newcommand{\datatype}[2]{{\mprset{fraction={===}} \inferrule{#1}{#2}}}
 
@@
   \begin{mathpar}
-    {\mprset{fraction={===}}
-      \inferrule
+    \datatype
     { }
-      {\type{\N} : \Set}}
+    {\type{\N} : \Set}
 
@@
 \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   {n : \N}
-    {\Var_n : \Set}}
+  {\Var_n : \Set}
 
@@
 \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   {n : \N}
-    {\Process_n : \Set}}
+  {\Process_n : \Set}
 \end{mathpar}
@@
 \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   { }
-    {P \eqeq Q : \Set}}
+  {P \eqeq Q : \Set}
 
@@
 \begin{mathpar}
-  \mprset{flushleft}
-
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   { }
-    {\Rec : \Set}}
+  {\Rec : \Set}
 
@@
 
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   {P \, Q : \Process_n \\ r : \Rec}
-    {P \eq{r} Q : \Set}}
+  {P \eq{r} Q : \Set}
 
@@
 \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   {n : \N}
-    {\Channel_n : \Set}}
+  {\Channel_n : \Set}
 
@@
 
-  {\mprset{fraction={===}}
-    \inferrule
+  \datatype
   {i : \Channel_n \\ P \; Q : \Process_n}
-    {P \reduce{i} Q : \Set}}
+  {P \reduce{i} Q : \Set}
 
@@
 \begin{split}
-  \field{IDX}          &: \Set \\
-  \field{\exists IDX}  &: \field{IDX} \\
-  \field{CARRIER}      &: \field{IDX} \to \Set \\
-  \field{QUANTIFIERS}  &: (i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
+  &\field{IDX}          &:{} &\Set \\
+  &\field{\exists IDX}  &:{} &\field{IDX} \\
+  &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
+  &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
 \end{split}
@@
 
-\todo{mention types are independent from capabilities}
-\todo{add index to in-type}
-\todo{update examples}
+\subsection{Contexts}
 
-\subsection{Contexts}
+In the linear \picalc{}, the capabilities of the variables in context need to be modified every time input or output occurs.
+However, the types of the variables in context need not.
+As neither capabilities nor their indices depend on types, we choose to keep these two kinds of information separate (Figure \ref{contexts}): a length-indexed context of types $\PreCtx_n$ on one side; a length-indexed context of indices $\Idxs_n$ on the other, with a context of capabilities $\Ctx_{idxs}$ on top.
 
@@
   \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
-    { }
-    {\Type : \Set}
-  }
-
-  \inferrule
-  {n : \N}
-  {\base{n} : \Type}
-
-  \inferrule
-  {t : \Type \\ \inferrule*{} {i : \Idx \\\\ x : \Carrier_i}}
-  {\channel{t}{x} : \Type}
-\end{mathpar}
-\caption{This is...}
-\end{figure}
-
-\begin{figure}[h]
-\begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+    \datatype
         {n : \N}
         {\PreCtx_n : \Set}
-  }
 
@@
   \end{mathpar}
-\caption{This is...}
-\end{figure}
 
-\begin{figure}[h]
   \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule
+    \datatype
         {n : \N}
-    {\Idxs_n : \Set}}
+        {\Idxs_n : \Set}
 
@@
         \inferrule
-  {is : \Idxs_n \\ i : \Idx}
-  {is , i : \Idxs_{\suc n}}
+            {idxs : \Idxs_n \\ idx : \Idx}
+            {idxs , idx : \Idxs_{\suc n}}
 
@@
 
-  {\mprset{fraction={===}}
-    \inferrule
-    {is : \Idxs_n}
-    {\Ctx_{is} : \Set}}
+            \datatype
+                {idxs : \Idxs_n}
+                {\Ctx_{idxs} : \Set}
 
@@
                 \inferrule
-  {\Gamma : \Ctx_{is} \\ x : \Carrier_i}
-  {\Gamma , x : \Ctx_{is , i}}
+                    {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx}}
+                    {\Gamma , x : \Ctx_{idxs , idx}}
   \end{mathpar}
-\caption{This is...}
+  \caption{Length-indexed typing contexts; length-indexed index context; index-indexed capability contexts.}
+  \label{contexts}
 \end{figure}
-two-layered approach: types on one hand, capabilities on the other
-removing from context vs keeping in context but marking it used
+
+We represent base types $\base{n}$ using a natural number $n$ that can be reified into a type by the host language.
+Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism, and it provides an additional proof of inhabitance, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
+
+Channel types $\channel{t}{x}$ use $t$ as the type of data sent over the channel and $x$ as the capability of the data sent over the channel.
+When data is received on a channel of type $\channel{t}{x}$, the type $t$ will be appended to the context of types $\PreCtx_n$ and the capability $x$ to the context of capabilities $\Ctx_{idxs}$.
+
+\begin{figure}[h]
+\begin{mathpar}
+  \datatype
+  { }
+  {\Type : \Set}
+
+  \inferrule
+  {n : \N}
+  {\base{n} : \Type}
+
+  \inferrule
+  {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
+  {\channel{t}{x} : \Type}
+\end{mathpar}
+\caption{Types for a resource-aware \picalc{}.}
+\label{types}
+\end{figure}
 
@@
 
+two-layered approach: types on one hand, capabilities on the other
+
 \begin{figure}[h]
+\begin{adjustwidth}{-1cm}{-1cm}
+\centering
 \begin{mathpar}
-  {\mprset{fraction={===}}
-    \inferrule{
+  \mprset{sep=0.5em}
+
+  \datatype{
     \gamma : \PreCtx_n \\
-      \inferrule*{}{
+    \stacked{
       idxs : \Idxs_n \\\\
       \Gamma : \Ctx_{idxs}} \\
-      i : Fin n
+    i : \Var_n \\
     t : \Type \\
-      x : \Carrier_i \\
-      \Delta : \Ctx_{is}}
-    {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}}
+    \stacked{
+      idx : \Idx \\\\
+      x : \Carrier_{idx}} \\
+    \Delta : \Ctx_{idxs}}
+  {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}
 
   \inferrule
-  {\Gamma : \Ctx_{idxs} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
+      {\gamma : \PreCtx_n \\
+        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
+        t : \Type \\
+        \stacked{idx : \Idx \\\\ x \; y \; z : \Carrier_{idx}}\\
+        \op{x}{y}{z}}
   {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{zero}{t}{y}{\Gamma , z}}
@@
   \inferrule
-  {\Gamma : \Ctx_{idxs} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{idxs} \\\\
+      {\gamma : \PreCtx_n \\
+        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
+        t \; t' \; : Type \\
+        \stacked{idx : \Idx \\\\ x : \Carrier_{idx}} \\
+        \stacked{idx' : \Idx \\\\ x' : \Carrier_{idx'}} \\
+        \Delta : \Ctx_{idxs} \\\\
     loc_x : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
-  {\constr{suc} \; loc_x : \contains{\gamma , t}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
+  {\constr{suc} \; loc_x : \contains{\gamma , t'}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
 \end{mathpar}
+\end{adjustwidth}
 \caption{This is...}
@@
 \begin{mathpar}
-  \mprset{flushleft}
-
-  {\mprset{fraction={===}}
-    \inferrule{
+  \datatype{
     \gamma : \PreCtx_n \\
-      \inferrule*{}{
+    \stacked{
       idxs : \Idxs_n \\\\
@@
     \Delta : \Ctx_{idxs}}
-    {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}}
+  {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}
 
@@
   \inferrule
-  {t : \Type \\ x : \Carrier_i \\ y : \Carrier_j \\\\
+  {t : \Type \\ x : \Carrier_{idx} \\ y : \Carrier_j \\\\
    cont : \types{\gamma , \channel{t}{x}}{\Gamma , y}{P}{\Delta , \lz}}
@@
 \end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-27 paper/paper.tex
Add a balanced predicate

@@
 \newcommand{\lio}{\field{\ell_{\#}}}
+\newcommand{\Balanced}{\field{Balanced}}
 
@@
   &\lo                    &:{} &                      &        & C \\
+  &\lio                   &:{} &                      &        & C \\
   &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
-  &\field{\cdot-join}     &:{} &                      &        & \type{\exists} \lio . \; \op{\lio}{\li}{\lo} \\
+  &\field{\cdot-join}     &:{} &                      &        & \op{\lio}{\li}{\lo} \\
   &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Progress}\label{progress}
+
+\begin{figure}[h]
+\begin{equation}
+\begin{aligned}
+  &\Balanced                &:{} &                      &        & C \to \Set \\
+  &\field{Balanced-}\lz     &:{} &                      &        & \Balanced \; \lz \\
+  &\field{Balanced-}\lio    &:{} &                      &        & \Balanced \; \lio \\
+  &\field{Balanced-\cdot^l} &:{} &\forall \{x y z\}     & \to \; & \Balanced \; x \to \Balanced \; z \to \op{x}{y}{z} \to \Balanced \; y \\
+  &\field{Balanced?}        &:{} &\forall x             &        & \type{DEC} \; (\field{Balanced} \; x)
+\end{aligned}
+\end{equation}
+\caption{}
+\label{progress-capabilities}
+\end{figure}
+
+\todo{scope extension rule}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \section{Related Work}

### 2020-03-27 paper/paper.tex
Minor stylistic changes

@@
 \newcommand{\channel}[2]{\constr{C[} \; #1 \; \constr{\propto} \; #2 \; \constr{]}}
+\newcommand{\comma}{\; \constr{,} \;}
 
@@
 \newcommand{\types}[4]{#1 \; \type{\propto} \; #2 \; \type{\vdash} \; #3 \; \type{\boxtimes} \; #4}
-\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{[} \; #3 \; \type{]\coloneqq} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
+\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{_[} _{#3} \type{_]=} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
 \newcommand{\Var}{\type{VAR}}
@@
             {\gamma : \PreCtx_n \\ t : \Type}
-            {\gamma , t : \PreCtx_{\suc n}}
+            {\gamma \comma t : \PreCtx_{\suc n}}
   \end{mathpar}
@@
             {idxs : \Idxs_n \\ idx : \Idx}
-            {idxs , idx : \Idxs_{\suc n}}
+            {idxs \comma idx : \Idxs_{\suc n}}
 
@@
                     {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx}}
-                    {\Gamma , x : \Ctx_{idxs , idx}}
+                    {\Gamma \comma x : \Ctx_{idxs \comma idx}}
   \end{mathpar}
@@
         \op{x}{y}{z}}
-  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{zero}{t}{y}{\Gamma , z}}
+  {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{zero}{t}{y}{\Gamma \comma z}}
 
@@
         \Delta : \Ctx_{idxs} \\\\
-    loc_x : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
-  {\constr{suc} \; loc_x : \contains{\gamma , t'}{\Gamma , x'}{\suc i}{t}{x}{\Delta , x'}}
+    loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
+  {\constr{suc} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
 \end{mathpar}
@@
   \inferrule
-  {t : \Type \\ x : \Carrier_{idx} \\ y : \Carrier_j \\\\
-   cont : \types{\gamma , \channel{t}{x}}{\Gamma , y}{P}{\Delta , \lz}}
+  {t : \Type \\ x : \Carrier_{idx} \\ y : \Carrier_{idx'} \\\\
+   cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma y}{P}{\Delta \comma \lz}}
   {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}
@@
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-   cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , \lz}}
-  {\constr{recv} \; chan_x \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
+  {chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
+   cont : \types{\gamma \comma t}{\Xi \comma x}{P}{\Theta \comma \lz}}
+  {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
 
   \inferrule
-  {chan_x : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-   loc_y : \contains{\gamma}{\Delta}{j}{t}{x}{\Xi} \\\\
+  {chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+   loc_j : \contains{\gamma}{\Delta}{j}{t}{x}{\Xi} \\\\
    cont : \types{\gamma}{\Xi}{P}{\Theta}}
-  {\constr{send} \; chan_x \; loc_y \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
+  {\constr{send} \; chan_i \; loc_j \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
 

### 2020-03-29 paper/paper.tex
Small improvements
Use tuples for io multiplicities

@@
 \newcommand{\op}[3]{#1 \; \field{\coloneqq} \; #2 \; \field{\cdot} \; #3}
+\newcommand{\zero}{\field{0\cdot}}
+\newcommand{\one}{\field{1\cdot}}
 \newcommand{\li}{\field{\ell_i}}
@@
 \newcommand{\lio}{\field{\ell_{\#}}}
-\newcommand{\Balanced}{\field{Balanced}}
 
@@
 
-In its usual representation, the linear \picalc{} distinguishes between \emph{capabilities} -- $\ell_{\o}$ for cannot send nor receive, $\ell_i$ for input, $\ell_o$ for output, and $\ell_\#$ for both input and output -- and multiplicities -- $0$ or $1$.
-This introduces some unnecessary semantic duplication: a channel with multiplicity 1 and capability $\cdot \ell_{\o}$ is indistinguishable, in usage, from a channel with multiplicity 0.
-In this paper, we merge multiplicities and capabilities into a single type and define the algebraic rules for a generalized resource-aware type system on the \picalc{}.
-As such, it can be used to model shared, linear, or gradual types.
-
-The algebra has a single binary operator $(\cdot)$ that crushes together capabilities, and must satisfy the following characteristics:
+In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 or 1.
+In this paper we generalize over linearity and provide an algebra on multiplicities that is capable of encoding shared and gradual types as well.
+The algebra has a single binary operator $(\cdot)$ that crushes multiplicities together.
+It must satisfy the following constraints:
 \begin{description}
@@
 \item [Partiality]
-  Not all capabilities are necessarily compatible: in the linear \picalc{} the capability $\ell_i$ cannot be split into $\ell_i$ and $\ell_i$.
+  Not all multiplicities are necessarily compatible: in the linear \picalc{} $1$ cannot be split into $1$ and $1$.
 \item [Functionality]
-  Given two capabilities $y$ and $z$, it must be computable whether a third capability $x$ exists such that $\op{x}{y}{z}$.
-  If such a $x$ exists, it must be unique.
+  Given two multiplicities $y$ and $z$, it must be computable whether a third multiplicity $x$ exists such that $\op{x}{y}{z}$.
+  If such an $x$ exists, it must be unique.
 
 \item [Injectivity]
-  To prove framing (\S \ref{framing}) we need to be able to compute the \emph{subtraction} of capabilities.
-  That is, $x - z \equiv y$ where $y$ is uniquely determined iff $\op{x}{y}{z}$ where $y$ is uniquely determined.
+  To prove framing (\S \ref{framing}) we need to be able to \emph{subtract} multiplicities.
+  For that end, we rely on $x - z \equiv y$ iff $\op{x}{y}{z}$.
+  As a consequence, we must constraint $y$ to be uniquely determined as well.
 
@@
 \item [Distinguished elements]
-  There must be a neutral element $\ell_{\o}$ absorbed on both sides.
-  There need to be two special elements $\ell_i$ and $\ell_o$ that will be subtracted every time input and output (respectively) happens.
-  These two elements need to be compatible, and we will refer to the result as $\ell_\#$.
+  There must be a neutral element $0$ absorbed on both sides.
+  There must be a special element $1$ to record usage.
 \end{description}
 
-In Figure \ref{capabilities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation, satisfying all the properties mentioned above.
-
-\todo{Ornela: in linear pi if you write $\li$ the $\ell$ already means that the multiplicity is 1, meaning linear. 
-We need to distinguish between \emph{capabilities} i, o or $\sharp$ and \emph{multiplicities} 0, 1 or $\omega$.
-So, the $\ell$ cannot be $\omega$, hence shared.
-You should use only i, o or $\sharp$ in Fig 7.
-and use $\ell$ as a syntactic sugar for i, o or $\sharp$ with multiplicity 1.}
+Together, these constraints give raise to a \emph{partial commutative monoid}.
+In Figure \ref{fig:multiplicities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation such that it forms a partially commutative monoid.
 
@@
 \begin{aligned}
-  &\lz                    &:{} &                      &        & C \\
-  &\li                    &:{} &                      &        & C \\
-  &\lo                    &:{} &                      &        & C \\
-  &\lio                   &:{} &                      &        & C \\
+  &\zero                  &:{} &                      &        & C \\
+  &\one                   &:{} &                      &        & C \\
   &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
-  &\field{\cdot-join}     &:{} &                      &        & \op{\lio}{\li}{\lo} \\
   &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
@@
   &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  &\field{\cdot-id^l}     &:{} &\forall x             & \to \; & \op{x}{\lz}{x} \\
+  &\field{\cdot-id^l}     &:{} &\forall x             & \to \; & \op{x}{\zero}{x} \\
   &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
@@
 \end{equation}
-\caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid with special elements $\ell_i$, $\ell_o$ and $\ell_\#$.}
-\label{capabilities}
+\caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid.}
+\label{fig:multiplicities}
 \end{figure}
 
-To allow for several capability algebras to coexist on the same type system, in Figure \ref{indexed-capabilities} we index the algebra carriers by an index $\Idx$.
-Such an index is a type with one element for each algebra of our type system.
-To be able to prove subject congruence, we require for at least one such algebra to exist.
-Finally, the field $\field{QUANTIFIERS}$ requires for each index $i$ to have an algebra defined on the carrier set $\constr{\Carrier_i}$.
+To allow for several multiplicity algebras to coexist on the same type system, we index the carriers of the algebras by an index $\Idx$ (see Figure \ref{fig:indexed-multiplicities}).
+Each index $i$ has a corresponding carrier type $\field{CARRIER}_i$ such that $\field{CARRIER}_i$ satisfies the algebra $\Quantifier_{\field{CARRIER}_i}$.
+We require the indexing type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
 
@@
 \end{equation}
-\caption{Indexed set of partial commutative monoids}
-\label{indexed-capabilities}
+\caption{Indexed set of partial commutative monoids.}
+\label{fig:indexed-multiplicities}
 \end{figure}
@@
 \subsubsection{Example Type Systems}
+
+\todo{Add typing derivations for example processes}
 
@@
 The carrier $C$ is implemented as a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
-All of $\lz$, $\li$, $\lo$ and $\lio$ are instantiated to $\constr{\omega}$.
+Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
 The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
@@
 
-The carrier $C$ is implemented as a type $\type{Four}$ with the trivial elements $\constr{\ell_{\o}}$, $constr{\ell_i}$, $\constr{\ell_o}$ and $\constr{\ell_\#}$.
-The monoidal operation has the element $\constr{\ell_{\o}}$ as neutral on both sides, and the element $\constr{\ell_\#}$ splitting into $\constr{\ell_i}$ and $\constr{\ell_o}$ (or $\constr{\ell_o}$ and $\constr{\ell_i}$ -- and is uninhabited in every other case.
+The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
+The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$ -- and is uninhabited in every other case.
 All other rules follow trivially.
@@
 
-The carrier $C$ is implemented as a tuple $\N \type{\times} \N$ of natural numbers.
-The element $\lz$ corresponds to the tuple $(\constr{0} \constr{,} \constr{0})$, $\li$ to $(\constr{1} \constr{,} \constr{0})$, $\lo$ to $(\constr{0} \constr{,} \constr{1})$, and $\lio$ to $(\constr{1} \constr{,} \constr{1})$.
-The partial monoid is defined exactly when $\op{(x_l + y_l \constr{,} x_r + y_r)}{(x_l \constr{,} x_r)}{(y_l \constr{,} y_r)}$.
+The carrier $C$ is implemented as the type of natural numbers $\N$.
+The element $\zero$ corresponds 0, and $\one$ to 1.
+The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
 All other rules follow trivially from the algebraic rules for the addition of naturals.
@@
 
-In the linear \picalc{}, the capabilities of the variables in context need to be modified every time input or output occurs.
-However, the types of the variables in context need not.
-As neither capabilities nor their indices depend on types, we choose to keep these two kinds of information separate (Figure \ref{contexts}): a length-indexed context of types $\PreCtx_n$ on one side; a length-indexed context of indices $\Idxs_n$ on the other, with a context of capabilities $\Ctx_{idxs}$ on top.
+In a resource-aware \picalc{} type multiplicities change every time input and output occurs; however, the types themselves remain constant.
+If multiplicities need not to depend on types, keeping types and multiplicities in independent contexts allows for a more modular approach to the algebra on contexts of multiplicities.
+In this paper we choose that approach, which entails accepting that base non-channel types have dedicated input and output multiplicities too -- we will take care of impeding that any actual input or output can occur through them in the typing rules.
+
+Figure \ref{fig:prectx} represents base types $\base{n}$ using a natural number $n$ that can be reified into a type by the host language.
+Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism, and it provides an additional proof of inhabitance, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
+
+Channel types $\channel{t}{x}$ use $t$ as the type of data sent over the channel and $x$ as the capability of the data sent over the channel.
+When data is received on a channel of type $\channel{t}{x}$, the type $t$ will be appended to the context of types $\PreCtx_n$ and the capability $x$ to the context of capabilities $\Ctx_{idxs}$.
+
+Types are then listed in a length-indexed \emph{pre-context} $\PreCtx_n$.
 
 \begin{figure}[h]
+\begin{mathpar}
+  \datatype
+  { }
+  {\Type : \Set}
+
+  \inferrule
+  {n : \N}
+  {\base{n} : \Type}
+
+  \inferrule
+  {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
+  {\channel{t}{x} : \Type}
+\end{mathpar}
+
 \begin{mathpar}
@@
 \end{mathpar}
+\caption{Types and length-indexed typing contexts.}
+\label{fig:prectx}
+\end{figure}
 
+To allow different multiplicity algebras to coexist we first define a length-indexed context of carrier indices $\Idxs_n$ (see \ref{fig:ctx}).
+On top of it, we define a context of multiplicities $\Ctx_{idxs}$.
+We lift the algebra on multiplicities to multiplicity contexts by constraining the algebra on multiplicity contexts to act on  contexts defined on the same underlying context of carrier indices.
+We use two multiplicities over a common index to represent multiplicities for both the input and the output capabilities ($\Carrier_{idx}^2$).
+
+\begin{figure}[h]
 \begin{mathpar}
@@
   \inferrule
-                    {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx}}
+      {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx} ^2}
       {\Gamma \comma x : \Ctx_{idxs \comma idx}}
 \end{mathpar}
-  \caption{Length-indexed typing contexts; length-indexed index context; index-indexed capability contexts.}
-  \label{contexts}
-\end{figure}
-
-We represent base types $\base{n}$ using a natural number $n$ that can be reified into a type by the host language.
-Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism, and it provides an additional proof of inhabitance, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
-
-Channel types $\channel{t}{x}$ use $t$ as the type of data sent over the channel and $x$ as the capability of the data sent over the channel.
-When data is received on a channel of type $\channel{t}{x}$, the type $t$ will be appended to the context of types $\PreCtx_n$ and the capability $x$ to the context of capabilities $\Ctx_{idxs}$.
-
-\begin{figure}[h]
-\begin{mathpar}
-  \datatype
-  { }
-  {\Type : \Set}
-
-  \inferrule
-  {n : \N}
-  {\base{n} : \Type}
-
-  \inferrule
-  {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
-  {\channel{t}{x} : \Type}
-\end{mathpar}
-\caption{Types for a resource-aware \picalc{}.}
-\label{types}
+\caption{Length-indexed context of carrier indices with a context of multiplicities on top.}
+\label{fig:ctx}
 \end{figure}
@@
 two-layered approach: types on one hand, capabilities on the other
+
+\todo{lift op to $^2$}
 
@@
       idx : \Idx \\\\
-      x : \Carrier_{idx}} \\
+      x : \Carrier_{idx}^2} \\
     \Delta : \Ctx_{idxs}}
@@
         t : \Type \\
-        \stacked{idx : \Idx \\\\ x \; y \; z : \Carrier_{idx}}\\
+        \stacked{idx : \Idx \\\\ x \; y \; z : \Carrier_{idx}^2}\\
         \op{x}{y}{z}}
@@
         t \; t' \; : Type \\
-        \stacked{idx : \Idx \\\\ x : \Carrier_{idx}} \\
-        \stacked{idx' : \Idx \\\\ x' : \Carrier_{idx'}} \\
+        \stacked{idx : \Idx \\\\ x : \Carrier_{idx}^2} \\
+        \stacked{idx' : \Idx \\\\ x' : \Carrier_{idx'}^2} \\
         \Delta : \Ctx_{idxs} \\\\
@@
 
-\todo{comment on scope restriction rule and capabilities}
+\todo{comment on balancing in chan rule}
 
@@
   \inferrule
-  {t : \Type \\ x : \Carrier_{idx} \\ y : \Carrier_{idx'} \\\\
-   cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma y}{P}{\Delta \comma \lz}}
+  {t : \Type \\ x : \Carrier_{idx}^2 \\ y : \Carrier_{idx'} \\\\
+   cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma (y \comma y) }{P}{\Delta \comma \lz}}
   {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Progress}\label{progress}
-
-\begin{figure}[h]
-\begin{equation}
-\begin{aligned}
-  &\Balanced                &:{} &                      &        & C \to \Set \\
-  &\field{Balanced-}\lz     &:{} &                      &        & \Balanced \; \lz \\
-  &\field{Balanced-}\lio    &:{} &                      &        & \Balanced \; \lio \\
-  &\field{Balanced-\cdot^l} &:{} &\forall \{x y z\}     & \to \; & \Balanced \; x \to \Balanced \; z \to \op{x}{y}{z} \to \Balanced \; y \\
-  &\field{Balanced?}        &:{} &\forall x             &        & \type{DEC} \; (\field{Balanced} \; x)
-\end{aligned}
-\end{equation}
-\caption{}
-\label{progress-capabilities}
-\end{figure}
-
-\todo{scope extension rule}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \section{Related Work}

### 2020-03-30 paper/paper.tex
Section on variable references

@@
 
-\newcommand{\lamcalc}{$\lambda$-calculus}
+\newcommand{\lamdacalc}{$\lambda$-calculus}
 \newcommand{\picalc}{$\pi$-calculus}
@@
 
+Type constructors as the names for the typing rules.
 Double rule for type-level definitions.
@@
 The binders themselves introduce no names anymore.
-The expression $\lambda g . (\lambda f . f g) g$ in the \lamcalc{} would translate as $\lambda (\lambda 0 1) 0$.
+The expression $\lambda g . (\lambda f . f g) g$ in the \lamdacalc{} would translate as $\lambda (\lambda 0 1) 0$.
 That is, terms at different \emph{depths} must use different indices to refer to the same binding.
@@
 
-In the \lamcalc{} $\beta$-reduction operates on syntactically adjacent terms.
+In the \lamdacalc{} $\beta$-reduction operates on syntactically adjacent terms.
 In the \picalc{} however, the syntax introduces unnecessary distinctions
@@
 A type system is in the business of further constraining a language, filtering out terms that (although well-formed) are meaningless.
-In the simply typed \lamcalc{}, a term has a meaning iff it has a type.
+In the simply typed \lamdacalc{}, a term has a meaning iff it has a type.
 This typing judgment is a proof of the form $\Gamma \vdash t : T$, where the term $t$ has type $T$ under the typing context $\Gamma$ -- the typing context associates variables (possibly occurring within $t$) to types.
@@
 Together, these constraints give raise to a \emph{partial commutative monoid}.
-In Figure \ref{fig:multiplicities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation such that it forms a partially commutative monoid.
+In Figure \ref{fig:multiplicities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation such that it forms a partial commutative monoid.
+
+\todo{SHOULD ALL THESE NOT BE DEFINITIONS?}
 
@@
 We require the indexing type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
+\todo{lift op to $^2$}
 
@@
 The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
-The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$ -- and is uninhabited in every other case.
+The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
 All other rules follow trivially.
@@
 
-Variable references as proofs of capability
+\todo{syntax directed typing rules}
 
-Context splits at each variable reference
+In a resource-aware calculi, typing rules modify the context of multiplicities under which their premises are typed -- e.g.
+in $\send{x}{y}P$ the multiplicity context under which $P$ is typed needs to be modified at at $x$ and $y$.
+This context transformations can be modeled as functions on contexts, however threading these transformations through the proofs on typing rules is cumbersome, and these transformations necessitate of a myriad of additional lemmas.
+Additionally, if the typing rules do not record the result of their actions on contexts, managing parallel composition entails a context split for which the user is required to provide an existential proof at the cutting point \todo{rephrase}.
 
-two-layered approach: types on one hand, capabilities on the other
+An alternative is to keep track in the typing judgment of the way in which the context has been transformed.
+One way of doing so is by mentioning both the \emph{input contexts} and the \emph{output contexts} over which the processes are typed.
+Parallel composition then uses the output context of the left hand process as the input context of the right hand.
+As an additional benefit, the associativity of parallel composition becomes trivial to prove.
+\S \ref{typing-judgments} introduces the typing rules for such a typing judgment.
+
+In the \picalc{} both input and output contain variable references.
+Both these actions need to modify their context, and thus using variable references that mention both their input and output contexts seems relevant.
+\S \ref{variable-references} introduces the typing rules for such variable references.
+
+Both the typing judgment and the variable references are of an inductive form.
+As a consequence, all proofs on them will need to perform a double induction: once on the typing derivation, once for each variable reference.
+
+Both the typing rules and the rules for variable references mention both contexts of types and contexts of multiplicities.
+The contexts of types are threaded through the premises unchanged; the contexts of multiplicities are threaded through the premises using one premise's output as the next premise's input.
+Both the context of types and the context of multiplicities are of a common length.
+
+\subsection{Variable references} \label{variable-references}
+
+Akin to the typing rules for processes, the rules for variable references presented in Figure \ref{fig:variable-references} are indexed over the typing context, the input multiplicity context, and the output multiplicity context.
+The base case $\constr{zero}$ splits the last multiplicity in the input context into $x$ and the last multiplicity in the output context, while also ensuring that $t$ is the last type in the typing context.
+The inductive case $\constr{suc}$ appends the same multiplicity to both the input context and the output context, while also appending an extra type to the typing context.
+
+\lemma{The typing context $\gamma$ at index $i$ is type $t$.}
+\lemma{The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.}
+\lemma{For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.}
+\corollary{Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.}
+\proof{By structural induction on the variable reference}
 
@@
       {\gamma : \PreCtx_n \\
-        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
+        \Gamma : \Ctx_{idxs}\\
         t : \Type \\
-        \stacked{idx : \Idx \\\\ x \; y \; z : \Carrier_{idx}^2}\\
+        x \; y \; z : \Carrier_{idx}^2\\
         \op{x}{y}{z}}
@@
       {\gamma : \PreCtx_n \\
-        \stacked{idxs : \Idxs_n \\\\ \Gamma : \Ctx_{idxs}}\\
-        t \; t' \; : Type \\
-        \stacked{idx : \Idx \\\\ x : \Carrier_{idx}^2} \\
-        \stacked{idx' : \Idx \\\\ x' : \Carrier_{idx'}^2} \\
+        \Gamma : \Ctx_{idxs}\\
+        t \; t' \; : \Type \\
+        x : \Carrier_{idx}^2 \\
+        x' : \Carrier_{idx'}^2 \\
         \Delta : \Ctx_{idxs} \\\\
@@
 \end{adjustwidth}
-\caption{This is...}
+\caption{Variable references as proofs on the typing context and input and output multiplicity contexts.}
+\label{fig:variable-references}
 \end{figure}
+
+\todo{example}
+
+\subsection{Typing judgments}\label{typing-judgments}
+
+The indices in the variable reference rules are then used as the variable references of the underlying process.
 

### 2020-03-31 paper/paper.tex
Add a unit type, use definition environment

@@
 \usepackage{todonotes}
-\usepackage{enumitem}
-\setlist{parsep=0pt,listparindent=\parindent}
 \usepackage{amssymb}
@@
 \bibliographystyle{plainurl}
+
+% Mathbb doesn't support digits
+\usepackage{bbm}
 
@@
 \usepackage{todonotes}
+
+% Do not use italics for definitions
+\theoremstyle{definition}
+\newtheorem{nidefinition}[theorem]{Definition}
+
+% No line numbers
+\nolinenumbers
 
@@
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
+\newcommand{\unit}{\constr{\mathbbm{1}}}
 \newcommand{\base}[1]{\constr{B[} \; #1 \; \constr{]}}
@@
     {\type{\N} : \Set}
+    \; \textsc{Nat}
 
@@
 Type constructors as the names for the typing rules.
+Datatype constructor rules might have additional index-free synonyms for ease of reference.
 Double rule for type-level definitions.
@@
   {\Var_n : \Set}
+  \; \textsc{Var}
 
@@
   {\Process_n : \Set}
+  \; \textsc{Process}
 \end{mathpar}
@@
   {P \eqeq Q : \Set}
+  \; \textsc{StructCong}
 
@@
   {\Rec : \Set}
+  \; \textsc{Rec}
 
@@
   {P \eq{r} Q : \Set}
+  \; \textsc{Equals}
 
@@
   {\Channel_n : \Set}
+  \; \textsc{Channel}
 
@@
   {P \reduce{i} Q : \Set}
+  \; \textsc{Reduces}
 
@@
 
-Together, these constraints give raise to a \emph{partial commutative monoid}.
-In Figure \ref{fig:multiplicities} we present an algebra $\Quantifier_C$ on the carrier set $C$ in the form of a terniary relation such that it forms a partial commutative monoid.
-
-\todo{SHOULD ALL THESE NOT BE DEFINITIONS?}
+\begin{nidefinition}
+  Together, the aforementioned constraints give raise to a \emph{partial commutative monoid}.
+  We deal with partiality by modeling the monoidal operation as a terniary relation.
+  An algebra on the carrier set $C$ forms a $\Quantifier_C$ iff it satisfies the fields in Figure \ref{fig:multiplicities}.
 
@@
   \end{figure}
+\end{nidefinition}
 
-To allow for several multiplicity algebras to coexist on the same type system, we index the carriers of the algebras by an index $\Idx$ (see Figure \ref{fig:indexed-multiplicities}).
-Each index $i$ has a corresponding carrier type $\field{CARRIER}_i$ such that $\field{CARRIER}_i$ satisfies the algebra $\Quantifier_{\field{CARRIER}_i}$.
-We require the indexing type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
+\begin{nidefinition}
+  An indexed set of partial commutative monoids is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it.
+  This allows for several algebras to coexist in the same type system.
+  Figure \ref{fig:indexed-multiplicities} captures these requirements: each index $i$ has a corresponding carrier type $\field{CARRIER}_i$ such that $\field{CARRIER}_i$ satisfies the algebra $\Quantifier_{\field{CARRIER}_i}$.
+  We require the index type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
   \todo{lift op to $^2$}
@@
   \end{figure}
+\end{nidefinition}
 
@@
 
-Figure \ref{fig:prectx} represents base types $\base{n}$ using a natural number $n$ that can be reified into a type by the host language.
-Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism, and it provides an additional proof of inhabitance, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
+\begin{nidefinition}
+  Figure \ref{fig:prectx} defines types as either unit types, base types or channel types.
+  Unit types $\unit$ provide a proof of inhabitance for types, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
+
+  Base types $\base{n}$ use a natural number $n$ that can be reified into a type by the host language.
+  Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism.
 
@@
 
-Types are then listed in a length-indexed \emph{pre-context} $\PreCtx_n$.
+  Types form length-indexed \emph{pre-contexts} $\PreCtx_n$.
 
@@
     {\Type : \Set}
+    \; \textsc{Type}
+
+    \inferrule
+    { }
+    {\unit : \Type}
 
@@
     {\PreCtx_n : \Set}
+    \; \textsc{PreCtx}
 
@@
   \end{figure}
+\end{nidefinition}
 
+\begin{nidefinition}
   To allow different multiplicity algebras to coexist we first define a length-indexed context of carrier indices $\Idxs_n$ (see \ref{fig:ctx}).
@@
     {\Idxs_n : \Set}
+    \; \textsc{Idxs}
 
@@
     {\Ctx_{idxs} : \Set}
+    \; \textsc{Ctx}
 
@@
 As an additional benefit, the associativity of parallel composition becomes trivial to prove.
-\S \ref{typing-judgments} introduces the typing rules for such a typing judgment.
 
@@
 Both these actions need to modify their context, and thus using variable references that mention both their input and output contexts seems relevant.
-\S \ref{variable-references} introduces the typing rules for such variable references.
 
@@
 
-\subsection{Variable references} \label{variable-references}
-
+\begin{nidefinition}
   Akin to the typing rules for processes, the rules for variable references presented in Figure \ref{fig:variable-references} are indexed over the typing context, the input multiplicity context, and the output multiplicity context.
@@
 
-\lemma{The typing context $\gamma$ at index $i$ is type $t$.}
-\lemma{The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.}
-\lemma{For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.}
-\corollary{Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.}
-\proof{By structural induction on the variable reference}
-
+  \todo{example}
   \todo{lift op to $^2$}
@@
   \begin{figure}[h]
-\begin{adjustwidth}{-1cm}{-1cm}
-\centering
   \begin{mathpar}
@@
     {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}
+    \; \textsc{VarRef}
 
     \inferrule
-      {\gamma : \PreCtx_n \\
-        \Gamma : \Ctx_{idxs}\\
-        t : \Type \\
+        {t : \Type \\
           x \; y \; z : \Carrier_{idx}^2\\
@@
     \inferrule
-      {\gamma : \PreCtx_n \\
-        \Gamma : \Ctx_{idxs}\\
-        t \; t' \; : \Type \\
+        {t \; t' \; : \Type \\
           x : \Carrier_{idx}^2 \\
-        x' : \Carrier_{idx'}^2 \\
-        \Delta : \Ctx_{idxs} \\\\
+          x' : \Carrier_{idx'}^2 \\\\
       loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
@@
   \end{mathpar}
-\end{adjustwidth}
   \caption{Variable references as proofs on the typing context and input and output multiplicity contexts.}
@@
   \end{figure}
+\end{nidefinition}
 
-\todo{example}
+\lemma{The typing context $\gamma$ at index $i$ is type $t$.}
+\lemma{The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.}
+\lemma{For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.}
+\corollary{Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.}
+\proof{By structural induction on the variable reference}
 
@@
 
+\begin{nidefinition}
   \begin{figure}[h]
@@
     {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}
+    \; \textsc{Types}
 

### 2020-03-31 paper/paper.tex
Draft on leftover typing, typesetting improvements

@@
 
-\newcommand{\type}[1]{\textcolor{blue}{\mathtt{\operatorname{#1}}}}
-\newcommand{\constr}[1]{\textcolor{olive}{\mathtt{\operatorname{#1}}}}
-\newcommand{\field}[1]{\textcolor{red}{\mathtt{\operatorname{#1}}}}
+\newcommand{\type}[1]{\textcolor{blue}{\operatorname{#1}}}
+\newcommand{\constr}[1]{\textcolor{olive}{\operatorname{#1}}}
+\newcommand{\field}[1]{\textcolor{red}{\operatorname{#1}}}
 
@@
 \newcommand{\subst}[3]{#1[#2/#3]}
+\newcommand{\opsquared}[3]{#1 \, \coloneqq \, #2 \, \cdot^2 \, #3}
+\newcommand{\opctx}[3]{#1 \, \coloneqq \, #2 \, \otimes \, #3}
 
@@
 \newcommand{\types}[4]{#1 \; \type{\propto} \; #2 \; \type{\vdash} \; #3 \; \type{\boxtimes} \; #4}
-\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{_[} _{#3} \type{_]=} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
+\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{\ni}_{#3} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
 \newcommand{\Var}{\type{VAR}}
@@
 
+\begin{note}
+  We will often work with input and output multiplicities pairs, and thus we use the notation $\Carrier^2$ to stand for the type $\Carrier \times \Carrier$, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all of its properties are trivially lifted.
+\end{note}
+
 \begin{nidefinition}
@@
   We require the index type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
-  \todo{lift op to $^2$}
 
@@
 
+\begin{note}
+  We lift the monoidal operation on pairs of multiplicities $\opsquared{x}{y}{z}$ to a monoidal operation on contexts with the same underlying vector of indices $\opctx{\Gamma}{\Delta}{\Xi}$, where all the properties of the partial monoid are trivially lifted to contexts.
+\end{note}
+
 \subsection{Typing with Leftovers}\label{leftover typing}
@@
   \todo{example}
-  \todo{lift op to $^2$}
 
@@
           x \; y \; z : \Carrier_{idx}^2\\
-          \op{x}{y}{z}}
+          \opsquared{x}{y}{z}}
     {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{zero}{t}{y}{\Gamma \comma z}}
@@
 
-\lemma{The typing context $\gamma$ at index $i$ is type $t$.}
-\lemma{The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.}
-\lemma{For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.}
-\corollary{Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.}
-\proof{By structural induction on the variable reference}
+\begin{lemma}
+  The typing context $\gamma$ at index $i$ is type $t$.
+\end{lemma}
+\begin{lemma}
+  The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.
+\end{lemma}
+\begin{lemma}
+  For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.
+\end{lemma}
+\begin{corollary}
+  Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.
+\end{corollary}
+\proof{All proofs are by structural induction on \textsc{VarRef}.}
 
-\subsection{Typing judgments}\label{typing-judgments}
+\begin{nidefinition}
+  The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed over a context of types $\gamma$, a context of input multiplicities $\Gamma$, the process $P$ that is being typed, and the context of output capabilities $\Delta$, and notated as $\types{\gamma}{\Gamma}{P}{\Delta}$.
+  \todo{empty output context}
+  The typing rules are in one-to-one correspondence with the underlying syntax for processes, and thus we say they are syntax-directed.
 
-The indices in the variable reference rules are then used as the variable references of the underlying process.
+  Process termination is typed under any typing context, and preserves the multiplicity context unchanged.
 
+  Scope restriction accepts three parameters: the type $t$ of data sent over the channel; the input and output multiplicities $x$ of the data sent over the channel; and the multiplicity $y$ of the channel itself, which is used both as input and as output.
+  The typing derivation of the continuation process $P$ has a channel of data type $t$ and data multiplicity $x$ added to the typing context, and the multiplicities $(y , y)$ added to the input multiplicity context.
+  The continuation must completely exhaust this channel, to that end we substitute $(y , y)$ by $(\zero , zero)$ in the output context.
   \todo{comment on balancing in chan rule}
 
-\begin{nidefinition}
+  Input requires a proof $chan_i$ locating a channel with data type $t$, data multiplicities $x$ and multiplicity $\ell_i$ \todo{} at index $i$ in the typing and multiplicity contexts $\gamma$ and $\Gamma$.
+  The result of exhausting $\ell_i$ is the output multiplicity $\Xi$.
+  The data type $t$ is then used to expand the typing context of the continuation $cont$, while the data multiplicities $x$ are appended to the output context $\Xi$ to form the input context over which $P$ is typed.
+  These multiplicities $x$ must be exhausted by typing derivation of the continuation process.
+  Note that the index $i$ is then used as part of the syntax.
+
+  \todo{note that every new multiplicity introduced in an input context must be exhausted in the output context}
+
+  Similar to input, output requires a proof $chan_i$ locating a channel with data type $t$ and data multiplicities $x$ in $\gamma$ at index $i$, however this time the channel must have multiplicity $\ell_o$.
+  The result of exhausting $\ell_o$ is the output multiplicity $\Delta$.
+  That output context $\Delta$ must have a variable of type $t$ and multiplicities $x$ at position $j$.
+  This second proof represents the data that is sent over the channel.
+  The result of exhausting the multiplicities $x$ of that variable is the context $\Xi$.
+  The continuation process must be typed under this context $\Xi$.
+
+  Finally, parallel composition requires its two subprocesses to be typed such that the output context of the first is the input context of the second.
+
   \begin{figure}[h]
@@
   \end{mathpar}
-  \caption{This is...}
+  \caption{Left-over typing for a resource-aware typing system.}
+  \label{fig:types}
   \end{figure}
@@
 \subsection{Framing}
+
+\todo{op but for contexts}
+\begin{theorem}
+  Let $\opctx{\Gamma_l}{\Delta}{\Xi_l}$ and $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  If $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$ then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+\end{theorem}
 
@@
 
-\item [Affine types]
-
 \item [Proof of progress]
@@
 \item [Decidable typechecking]
+
+  Bidirectional typing rules needed.
 

### 2020-04-06 paper/paper.tex
Think about substitution
Solve end case in subject reduction
Add a bit more to the introduction

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Typing a linear \picalc}
-\author{Uma Zalakain}{University of Glasgow, Scotland}
-       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
-\author{Ornela Dardha}{University of Glasgow, Scotland}
-       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
-\authorrunning{U. Zalakain and O. Dardha}
-\Copyright{Uma Zalakain and Ornela Dardha}
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10003753.10003761.10003764</concept_id>
-<concept_desc>Theory of computation~Process calculi</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-\ccsdesc[300]{Theory of computation~Process calculi}
-\keywords{pi calculus, linear, types, concurrency}
-\supplement{\url{https://github/umazalakain/typing-linear-pi}}
-\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Abbreviations
+\newcommand{\lambdacalc}{$\lambda$-calculus}
+\newcommand{\picalc}{$\pi$-calculus}
+\newcommand{\Picalc}{$\pi$-Calculus}
 
+% Typing rules
 \newcommand{\stacked}[1]{\mprset{flushleft} \inferrule*{}{#1}}
 \newcommand{\datatype}[2]{{\mprset{fraction={===}} \inferrule{#1}{#2}}}
-
-\newcommand{\lamdacalc}{$\lambda$-calculus}
-\newcommand{\picalc}{$\pi$-calculus}
-\newcommand{\Picalc}{$\pi$-Calculus}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Typing a linear \picalc}
+\author{Uma Zalakain}{University of Glasgow, Scotland}
+       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
+\author{Ornela Dardha}{University of Glasgow, Scotland}
+       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
+\authorrunning{U. Zalakain and O. Dardha}
+\Copyright{Uma Zalakain and Ornela Dardha}
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10003752.10003753.10003761.10003764</concept_id>
+<concept_desc>Theory of computation~Process calculi</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+\ccsdesc[300]{Theory of computation~Process calculi}
+\keywords{pi calculus, linear, types, concurrency}
+\supplement{\url{https://github/umazalakain/typing-linear-pi}}
+\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
@@
 
-We live in a concurrent world where any given state is often decided by the interaction between a myriad of parties --- people, machines, processors, services, networks.
-As humans, we aim to model, predict, and even construct such interactive systems.
-As mathematicians, abstraction is our tool of choice to do so.
+We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
+As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
+The \picalc{} is a formal language that abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
+Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted too.
+We define the syntax for the \picalc{} in \S \ref{syntax}.
+We use dependent types to make this syntax well-scoped by construction: every unbound variable is accounted for in the type of the process using it.
+\todo{de Bruijn}
 
-The \picalc models communication.
+In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however computations need not to happen on syntactically contiguous terms.
+One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence relation that acts as a quotient type.
+A term $A$ can then be freely rewritten into a term $B$ as long as $A$ is structurally congruent to $B$.
+We introduce this equivalence relation in \S \ref{structural-congruence}.
 
-The properties guaranteed by the picalc
+In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
+Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequentiality in behavior.
+We provide an operational semantics for the \picalc{} in \S \ref{operational-semantics}.
+Every computational step is the result of communication; we take care to keep information on what channel the communication is happening on so that we can later use it to state the theorem for subject reduction (aka type preservation).
+
+As with the \lambdacalc{}, the \picalc{} can be used to construct a wide variety of misbehaved processes.
+
+The properties guaranteed by the linear picalc
 
@@
 The binders themselves introduce no names anymore.
-The expression $\lambda g . (\lambda f . f g) g$ in the \lamdacalc{} would translate as $\lambda (\lambda 0 1) 0$.
+The expression $\lambda g . (\lambda f . f g) g$ in the \lambdacalc{} would translate as $\lambda (\lambda 0 1) 0$.
 That is, terms at different \emph{depths} must use different indices to refer to the same binding.
@@
 
-In the \lamdacalc{} $\beta$-reduction operates on syntactically adjacent terms.
+In the \lambdacalc{} $\beta$-reduction operates on syntactically adjacent terms.
 In the \picalc{} however, the syntax introduces unnecessary distinctions
@@
 A type system is in the business of further constraining a language, filtering out terms that (although well-formed) are meaningless.
-In the simply typed \lamdacalc{}, a term has a meaning iff it has a type.
+In the simply typed \lambdacalc{}, a term has a meaning iff it has a type.
 This typing judgment is a proof of the form $\Gamma \vdash t : T$, where the term $t$ has type $T$ under the typing context $\Gamma$ -- the typing context associates variables (possibly occurring within $t$) to types.
@@
 \end{description}
+
+\todo{add on 0 = y * z -> 0 = y or 0 = z}
 
@@
 
+\todo{stress what leftover typing is as a technique}
 \todo{syntax directed typing rules}

### 2020-04-07 paper/paper.tex
Go over the intro again. Still needs more work

@@
 As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-The \picalc{} is a formal language that abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
+The \picalc{} \cite{} is a formal language that abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
 Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted too.
-We define the syntax for the \picalc{} in \S \ref{syntax}.
-We use dependent types to make this syntax well-scoped by construction: every unbound variable is accounted for in the type of the process using it.
-\todo{de Bruijn}
+We type the syntax for the \picalc{} in \S \ref{syntax}.
+We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every unbound variable is accounted for in the type of the process that uses it.
 
-In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however computations need not to happen on syntactically contiguous terms.
-One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence relation that acts as a quotient type.
+In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
+One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence \cite{} relation that acts as a quotient type.
 A term $A$ can then be freely rewritten into a term $B$ as long as $A$ is structurally congruent to $B$.
-We introduce this equivalence relation in \S \ref{structural-congruence}.
+We type this equivalence relation in \S \ref{structural-congruence}.
 
 In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
-Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequentiality in behavior.
-We provide an operational semantics for the \picalc{} in \S \ref{operational-semantics}.
-Every computational step is the result of communication; we take care to keep information on what channel the communication is happening on so that we can later use it to state the theorem for subject reduction (aka type preservation).
+Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequential behavior.
+We type a reduction relation that acts as an operational semantics modulo subject congruence in \S \ref{operational-semantics}.
+The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it to state the theorem for subject reduction (aka type preservation).
 
-As with the \lambdacalc{}, the \picalc{} can be used to construct a wide variety of misbehaved processes.
+As with the untyped \lambdacalc{}, the untyped \picalc{} can be used to construct a wide variety of ill-behaved terms --- for some characterization of ill-behaved.
+One can use type systems to constructively obtain a subset of terms for which certain properties hold.
+As an example, the simply typed \lambdacalc{} ensures that the argument supplied to a function is of the expected type.
+Correspondingly, the simply typed \picalc{} \cite{} ensures that the data sent over a channel is of the expected type.
+This property is often referred to as \emph{communication safety}.
 
-The properties guaranteed by the linear picalc
+Another property of particular interest to concurrent systems is the avoidance of race conditions.
+A type system that excludes multiple usages of a resource excludes race conditions.
+In the linear \picalc{} \cite{} channels must be used exactly once.
+This not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
+It also ensures that every declared communication channel must be used.
+The linear \picalc{} is also interesting because together with product types it is isomorphic to session types \cite{} -- which are used to encode communication protocols.
+We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
 
-why resource-aware typing
+Linear type systems need of context splits.
+These are tricky to handle constructively without explicitly asking for them to the user.
+Leftover typing \cite{} is a type theoretical technique that addresses this.
+It does so by carrying both an input and an output context (the leftovers).
+The output context can then be used as the input context of the continuations.
+We apply this technique to the \picalc{} in \S \ref{leftover-typing}.
+We generalize our algebra on usage multiplicities in \S \ref{multiplicities}.
+As a result, we are able to model a mix of shared, linear, and gradual types.
+\todo{reformulate paragraph}
 
-extensional typing rules for a given syntax and operational semantics
-
-leftover typing
+\todo{subject reduction}
 
@@
 
-Type constructors as the names for the typing rules.
-Datatype constructor rules might have additional index-free synonyms for ease of reference.
-Double rule for type-level definitions.
+Data type definitions have double lines and index-free synonyms as rule names, for ease of reference.
+Otherwise, we use the constructor name as the name of a typing rule.
 We omit universe levels for brevity.
-$\type{TYPES}$ are in blue and uppercased (and their indices are subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red and variables and functions in black.
-Some constructors are shared across types. They can however always be disambiguated through the type of the goal.
+$\type{TYPES}$ are in blue and uppercased (with indices as subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red, and variables and functions in black.
+Some constructors can be shared across types and disambiguated by context.
 
@@
 \section{Syntax}
+\label{syntax}
 
@@
 \caption{Types of size $n$}
-\label{var}
+\label{fig:var}
 \end{figure}
@@
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-As shown in Figure \ref{process}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
+As shown in Figure \ref{fig:process}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
 
@@
 
-\begin{equation*} \label{pi-calc-syntax}
+\begin{equation*}
 \begin{split}
@@
 \caption{Well-scoped grammar using de Bruijn indices}
-\label{process}
+\label{fig:syntax}
 \end{figure}
@@
 \section{Semantics}
+\label{semantics}
 
@@
 \subsection{Structural Congruence}
+\label{structural-congruence}
 
@@
 Any two structurally congruent processes are strongly bisimilar: they are can follow each other's reduction steps \cite{}.
-Figure \ref{struct-cong-base} lists the base cases of structural congruence.
+Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.
 The type $\Unused_0 \; Q$ witnesses that index $0$ does not appear nor in the inputs nor in the outputs of process $Q$ --- it does so by traversing $Q$.
@@
 \caption{Structural rewriting rules. Premises $P$, $Q$ and $R$ are of type $\Process_n$ where $n$ can be inferred.}
-\label{struct-cong-base}
+\label{fig:struct-cong-base}
 \end{figure}
@@
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen at any point in a process' recursive definition, and they are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{struct-cong}.
+As such, rewrites can happen at any point in a process' recursive definition, and they are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{fig:struct-cong}.
 In \S \ref{subject-congruence} we will prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then $Q$ is well-typed.
@@
   Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred.}
-\label{struct-cong}
+\label{fig:struct-cong}
 \end{figure}
@@
 \subsection{Operational Semantics}
+\label{operational-semantics}
 
-Figure \ref{reduction} models the operational semantics of the \picalc{}.
+Figure \ref{fig:reduction} models the operational semantics of the \picalc{}.
 Processes put in parallel reduce when they communicate over a common variable.
@@
 \caption{Operational semantics indexed by the channel over which reduction occurs.}
-\label{reduction}
+\label{fig:reduction}
 \end{figure}
@@
 \section{Resource-aware Typing System}
+\label{type-system}
+
+\todo{why do we use a single relation for both type judgements and usage judgments}
 
@@
 
-\subsection{Multiplicities and Capabilities}\label{multiplicities}
+\subsection{Multiplicities and Capabilities}
+\label{multiplicities}
 
@@
 
-\subsection{Typing with Leftovers}\label{leftover typing}
+\subsection{Typing with Leftovers}
+\label{leftover typing}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject Reduction}\label{subject-reduction}
+\section{Subject Reduction}
+\label{subject-reduction}
 
@@
 
-\subsection{Subject Congruence}\label{subject-congruence}
+\subsection{Subject Congruence}
+\label{subject-congruence}
 

### 2020-04-08 paper/paper.tex
Add the algebraic rule for 0 being the minimal element

@@
   There must be a neutral element $0$ absorbed on both sides.
+  The neutral element $0$ must also be the minimal element.
   There must be a special element $1$ to record usage.
 \end{description}
-
-\todo{add on 0 = y * z -> 0 = y or 0 = z}
 
@@
     &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-    &\field{\cdot-id^l}     &:{} &\forall x             & \to \; & \op{x}{\zero}{x} \\
+    &\field{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+    &\field{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
     &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\

### 2020-04-10 paper/paper.tex
More ergonomic types, get closer to subst input

@@
 \subsection{Typing with Leftovers}
-\label{leftover typing}
+\label{leftover-typing}
 
@@
     \inferrule
-    {chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-     cont : \types{\gamma \comma t}{\Xi \comma x}{P}{\Theta \comma \lz}}
+        {\stacked{
+            chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
+            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P}{\Theta \comma \lz}}}
         {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
@@
     \inferrule
-    {chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-     loc_j : \contains{\gamma}{\Delta}{j}{t}{x}{\Xi} \\\\
-     cont : \types{\gamma}{\Xi}{P}{\Theta}}
+        {\stacked{
+            chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t}{x\hspace{4em}}{\Xi} \\\\
+            cont \hspace{0.6em} : \types{\gamma}{\Xi}{P\hspace{6em}}{\Theta}}}
         {\constr{send} \; chan_i \; loc_j \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}

### 2020-04-11 paper/paper.tex
Rewrite the intro to the typing system

@@
 
-\todo{why do we use a single relation for both type judgements and usage judgments}
+A \emph{resource-aware} typing system for the \emph{\picalc{}} is distinctive in several ways.
 
-A type system is in the business of further constraining a language, filtering out terms that (although well-formed) are meaningless.
-In the simply typed \lambdacalc{}, a term has a meaning iff it has a type.
-This typing judgment is a proof of the form $\Gamma \vdash t : T$, where the term $t$ has type $T$ under the typing context $\Gamma$ -- the typing context associates variables (possibly occurring within $t$) to types.
-In the simply typed \picalc{} however, it is the channels that are typed, and a process has a meaning iff it \emph{uses} its channels according to their types.
-The typing judgment is therefore of the form $\Gamma \vdash p$, where the process $p$ (which has no type of its own) uses the channels in $\Gamma$ according to their types.
+\begin{description}
+  \item [As a resource-aware typing system,] the language is constrained not just by types, but also by their usage annotations.
+    We must be able to add up and split usage contexts, to do so we introduce a generalized algebra in \S \ref{multiplicities}.
+    We provide an independent inductive definition of typing contexts and usage contexts in \S \ref{contexts}.
+    \todo{why independent}
 
-A resource-aware type system uses both types and usage counting to constraint a language.
-It uses typing contexts to keep track of both types and usage counts.
-In the case of a linear type system, resources must be used exactly once, and usages transition from available to consumed as soon as they are used.
-In the linear \picalc{} however, every channel must be used exactly once to receive and exactly once to send.
-In \S \ref{multiplicities}, we generalize this usage counting algebra so that the same type system can be used for a combination of linear, shared, and gradual types.
+  \item [As a typing system for the \picalc{},] the terms (processes) in the typing judgment do not come accompanied by a type -- processes have no type.
+    It is the communication channels that are annotated with types and usages.
+    Typing judgments assert that the process correctly uses the channels in the context under which it is typed.
+\end{description}
 
-Syntax directed
-Continuations.
-\ref{leftover-typing}
+In this paper, we combine typing and annotation judgments into a single judgment.
+Although the typing rules are for now completely syntax-directed, adding support for branching would introduce choice in the typing derivations.
+If we were to keep typing judgments and annotation judgments separate, we would need a coherence relation making sure that the same choice is taken by both.
+We avoid doing so by combining them into a single judgment.
+
+We make use \emph{leftover typing}, a technique consisting of adding an \emph{output context of usages} to the usual \emph{input context of usages}.
+This output context contains the leftovers: the multiplicities unused by the processes.
+Crucially, this output context can then be used as the input context of the continuation process.
+The typing rules are introduced in \S \ref{leftover-typing}.
 
@@
 \subsection{Contexts}
+\label{contexts}
 

### 2020-04-11 paper/paper.tex
Rewrite the intro to leftover typing

@@
 
-% Do not use italics for definitions
+% Do not use italics in definitions and theorems
 \theoremstyle{definition}
 \newtheorem{nidefinition}[theorem]{Definition}
+\newtheorem{nitheorem}[theorem]{Theorem}
 
@@
 
-\todo{stress what leftover typing is as a technique}
+We use leftover typing \cite{} to encode our typing rules.
+The insight of this approach is to use input and output contexts in the typing judgment.
+Our typing judgment on processes are of the form $\types{\gamma}{\Gamma}{P}{\Delta}$ (introduced in Figure \ref{fig:types}).
+This assertion is in fact both a typing judgment on $\gamma$ and a usage judgment on $\Gamma$ and $\Delta$.
+The typing context $\gamma$ needs to be explicit because the usage context $\Gamma$ and $\Delta$ do not in any way depend on it -- in fact, they are only related by their common length.
+Using both an input and an output usage context has several advantages.
+For one, the usage modifications effected by both input and output can be enclosed within a self-containing typing and usage judgment for variable references -- also with an output context.
+But more crucially, we avoid requesting an explicit constructive proof of context splitting for parallel composition:
+if $P$ is typed under $\Gamma_l$ and $Q$ under $\Gamma_r$, then to type $\comp{P}{Q}$ under $\Gamma_m$ we would need a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$.
+The search for such a proof can be automated, but its existence is still required.
+Moreover, providing a typing derivation for $P$ gets the user no usage annotations to type $Q$.
+By using output usage contexts, we get the leftovers from the derivation on $P$ and use them to type $Q$.
+
+Both the typing judgment and the variable references are of an inductive form.
+As a consequence, all proofs on them will need to perform a double induction: once on the typing derivation, once for each variable reference.
+
 \todo{syntax directed typing rules}
-
-In a resource-aware calculi, typing rules modify the context of multiplicities under which their premises are typed -- e.g.
-in $\send{x}{y}P$ the multiplicity context under which $P$ is typed needs to be modified at at $x$ and $y$.
-This context transformations can be modeled as functions on contexts, however threading these transformations through the proofs on typing rules is cumbersome, and these transformations necessitate of a myriad of additional lemmas.
-Additionally, if the typing rules do not record the result of their actions on contexts, managing parallel composition entails a context split for which the user is required to provide an existential proof at the cutting point \todo{rephrase}.
-
-An alternative is to keep track in the typing judgment of the way in which the context has been transformed.
-One way of doing so is by mentioning both the \emph{input contexts} and the \emph{output contexts} over which the processes are typed.
-Parallel composition then uses the output context of the left hand process as the input context of the right hand.
-As an additional benefit, the associativity of parallel composition becomes trivial to prove.
 
@@
 Both these actions need to modify their context, and thus using variable references that mention both their input and output contexts seems relevant.
-
-Both the typing judgment and the variable references are of an inductive form.
-As a consequence, all proofs on them will need to perform a double induction: once on the typing derivation, once for each variable reference.
 
@@
 \todo{op but for contexts}
-\begin{theorem}
+\begin{nitheorem}
   Let $\opctx{\Gamma_l}{\Delta}{\Xi_l}$ and $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
   If $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$ then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 
+\begin{nitheorem}
+  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+\end{nitheorem}
+
 \subsection{Strengthening}
 
+\begin{nitheorem}
+  If $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\Unused \; i \; P$, then $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+\end{nitheorem}
+
+\begin{remark}
+  The proof that index $i$ is unused in $P$ must be syntactical and cannot rely on the usage annotation of index $i$ being preserved between $\Gamma$ and $\Xi$: process $P$ can send variable reference $i$ with multiplicity $\zero$ over a channel.
+\end{remark}
+
 \subsection{Swapping}
+
+\begin{nitheorem}
+  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
+\end{nitheorem}
 
@@
 
+\begin{nitheorem}
+  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+\end{nitheorem}
+
 \subsection{Substitution}
+
+\subsection{Subject Reduction}
+
+\begin{nitheorem}
+  If $P \reduce{c} Q$ and $\types{\gamma}{\Gamma'}{P}{\Xi}$ where $\opctx{\Gamma'}{\ell_{\#} @ c}{\Gamma}$,
+  then $\types{\gamma}{\Gamma}{Q}{\Xi}$
+\end{nitheorem}
 

### 2020-04-11 paper/paper.tex
Minor improvements

@@
 Linear type systems need of context splits.
-These are tricky to handle constructively without explicitly asking for them to the user.
+These are tricky to handle constructively without explicitly asking the user to supply them.
 Leftover typing \cite{} is a type theoretical technique that addresses this.
@@
 This assertion is in fact both a typing judgment on $\gamma$ and a usage judgment on $\Gamma$ and $\Delta$.
+These two judgments could, in fact, happen separately: our typing rules are completely syntax directed.
+However, if we were to introduce branching, typing derivations would have to make choices, and the two separate judgments would need of a third relation to make sure that their choices agree.
+To avoid this, we merge both judgments into a single one.
 The typing context $\gamma$ needs to be explicit because the usage context $\Gamma$ and $\Delta$ do not in any way depend on it -- in fact, they are only related by their common length.
-Using both an input and an output usage context has several advantages.
-For one, the usage modifications effected by both input and output can be enclosed within a self-containing typing and usage judgment for variable references -- also with an output context.
-But more crucially, we avoid requesting an explicit constructive proof of context splitting for parallel composition:
-if $P$ is typed under $\Gamma_l$ and $Q$ under $\Gamma_r$, then to type $\comp{P}{Q}$ under $\Gamma_m$ we would need a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$.
+
+While the typing contexts are preserved through the typing derivation, the usage contexts change.
+We use input and output usage contexts to capture this transformation.
+This has two main advantages: for one, the usage modifications effected by both input and output can be enclosed in a self-containing typing and usage judgment for variable references (also with an output context); but more crucially, we avoid requesting an explicit constructive proof of context splitting for parallel composition -- if $P$ is typed under $\Gamma_l$ and $Q$ under $\Gamma_r$, then to type $\comp{P}{Q}$ under $\Gamma_m$ a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ would be needed.
 The search for such a proof can be automated, but its existence is still required.
@@
 By using output usage contexts, we get the leftovers from the derivation on $P$ and use them to type $Q$.
+This is the recurring theme in leftover typing: usage contexts are threaded through the premises.
 
-Both the typing judgment and the variable references are of an inductive form.
-As a consequence, all proofs on them will need to perform a double induction: once on the typing derivation, once for each variable reference.
-
-\todo{syntax directed typing rules}
-
-In the \picalc{} both input and output contain variable references.
-Both these actions need to modify their context, and thus using variable references that mention both their input and output contexts seems relevant.
-
-Both the typing rules and the rules for variable references mention both contexts of types and contexts of multiplicities.
-The contexts of types are threaded through the premises unchanged; the contexts of multiplicities are threaded through the premises using one premise's output as the next premise's input.
-Both the context of types and the context of multiplicities are of a common length.
+The typing and usage judgments on both processes and variable references are of an inductive form.
+As a result, the proofs on typing derivations will perform a double induction: once on the derivation of the process, once on the derivation of the variable reference.
 
 \begin{nidefinition}
-  Akin to the typing rules for processes, the rules for variable references presented in Figure \ref{fig:variable-references} are indexed over the typing context, the input multiplicity context, and the output multiplicity context.
-  The base case $\constr{zero}$ splits the last multiplicity in the input context into $x$ and the last multiplicity in the output context, while also ensuring that $t$ is the last type in the typing context.
-  The inductive case $\constr{suc}$ appends the same multiplicity to both the input context and the output context, while also appending an extra type to the typing context.
+  Akin to the typing and usage judgment on processes, in Figure \ref{fig:variable-references} we present a typing and usage judgment on variable references -- also indexed over a typing context and an input an output usage context.
+  However, in addition to this, variable references also record the index, the type, and the usage annotation of the variable reference.
+  The base case $\constr{zero}$ ensures that $t$ is the most immediate type in typing context, and splits the most immediate annotation in the input context $x$ into the usage annotation for the variable $y$ and the annotation in the output context $z$.
+  The inductive case $\constr{suc}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.
+
+  \begin{remark}
+    The constructor $\constr{zero}$ is the only place in the typing system that uses the algebra on multiplicities.
+  \end{remark}
 
@@
         idx : \Idx \\\\
-        x : \Carrier_{idx}^2} \\
+        y : \Carrier_{idx}^2} \\
       \Delta : \Ctx_{idxs}}
-    {\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta} : \Set}
+    {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : \Set}
     \; \textsc{VarRef}
@@
 \begin{lemma}
-  The input multiplicity context $\Gamma$ at index $i$ is the result of performing the monoidal operation on $x$ and the output multiplicity context $\Delta$ at index $i$.
+  The input annotation in context $\Gamma$ at index $i$ is the result of aggregating (as per the algebra) $y$ and the output annotation in context $\Delta$ at index $i$.
 \end{lemma}
 \begin{lemma}
-  For every index $j$ other than $i$, the input multiplicity context $\Gamma$ at index $i$ is the output multiplicity $\Delta$ at index $i$.
+  For every index $j$ other than $i$, the input annotation in context $\Gamma$ at index $i$ is the output annotation in $\Delta$ at index $i$.
 \end{lemma}
 \begin{corollary}
-  Given the witness $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$, $\Gamma \equiv \Delta$ iff $x \equiv \zero$.
+  Given the witness $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$, $\Gamma \equiv \Delta$ iff $y \equiv \zero$.
 \end{corollary}

### 2020-04-12 paper/paper.tex
Introduce substitution

@@
 \bibliographystyle{plainurl}
+
+% Diagrams
+\usepackage{tikz}
 
@@
 We omit universe levels for brevity.
+Implicit arguments are mentioned by type definitions but omitted in constructors.
 $\type{TYPES}$ are in blue and uppercased (with indices as subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red, and variables and functions in black.
@@
 
-  \begin{remark}
-    The constructor $\constr{zero}$ is the only place in the typing system that uses the algebra on multiplicities.
-  \end{remark}
-
-  \todo{example}
-
   \begin{figure}[h]
@@
     \inferrule
-        {t : \Type \\
-          x \; y \; z : \Carrier_{idx}^2\\
-          \opsquared{x}{y}{z}}
+    {\opsquared{x}{y}{z}}
     {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{zero}{t}{y}{\Gamma \comma z}}
@@
     \inferrule
-        {t \; t' \; : \Type \\
-          x : \Carrier_{idx}^2 \\
-          x' : \Carrier_{idx'}^2 \\\\
-      loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
+    {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
     {\constr{suc} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
@@
 \end{nidefinition}
+
+\begin{remark}
+  We use $\field{\cdot-unique}$ to relieve the user from the proof burden $\opsquared{x}{y}{z}$ in the base case $\constr{zero}$.
+  This is the only place in the typing system that uses the algebra on multiplicities.
+\end{remark}
+
+\begin{example}
+\todo{example}
+\end{example}
 
@@
   For every index $j$ other than $i$, the input annotation in context $\Gamma$ at index $i$ is the output annotation in $\Delta$ at index $i$.
+\end{lemma}
+\begin{lemma}
+  \todo{Add feedback and feedfront}
 \end{lemma}
@@
 
+Subject reduction (aka type preservation) relates the operational semantics and the typing system of a language.
+In its usual form, it states that if $P$ reduces to $Q$ and $P$ is well-typed under context $\Gamma$, then $Q$ is well-typed under $\Gamma$ too.
+In the linear \picalc{} however, reduction \emph{is} meant to transform the context: each time communication occurs over a channel, that channel is exhausted.
+In our case, and because we generalize linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity on both its input and output capabilities.
+In \S \ref{semantics}, we made sure to lift the index of this channel to the type level.
+If the channel is internal to the process, the context must not be modified; if it is external on index $i$, the input usage context at index $i$ must have $\ell_\#$ subtracted.
+
+\begin{nitheorem}
+  Let $P \reduce{c} Q$.
+  If $c$ is $\constr{internal}$, then let $\Delta \equiv \epsilon$;
+  if $c$ is $\constr{external}\; i$, then let $\Delta \equiv \ell_\# @ i$.
+  Let $\Gamma'$ and $\Gamma$ be usage contexts only differing in $\Delta$, so that $\opctx{\Gamma'}{\Delta}{\Gamma}$.
+  If $\types{\gamma}{\Gamma'}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$
+\end{nitheorem}
+
+\begin{proof}
+\end{proof}
+
 \subsection{Framing}
 
-\todo{op but for contexts}
 \begin{nitheorem}
@@
 
-\subsection{Subject Reduction}
+When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the innermost variable in the body of the abstraction gets replaced with the argument of the application.
+In the \picalc{} however substitution takes another form: every reference to the innermost variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
+In a resource-aware \picalc{}, that means that instead of exhausting all $m$ multiplicities of the innermost variable, we leave those unaltered and exhaust them from the element that is being sent.
 
 \begin{nitheorem}
-  If $P \reduce{c} Q$ and $\types{\gamma}{\Gamma'}{P}{\Xi}$ where $\opctx{\Gamma'}{\ell_{\#} @ c}{\Gamma}$,
-  then $\types{\gamma}{\Gamma}{Q}{\Xi}$
+  Let $P$ be a process where the innermost variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
+  Let $j$ be the index of the data that is being received.
+  This data has at least $m$ multiplicities, such that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute the innermost variable by $j$.
+  As a result, we will exhaust no multiplicities on the innermost variable and exhaust them in $j$ instead, such that $\types{\gamma , t}{\Gamma , m}{\subst{P}{j}{\constr{zero}}}{\Xi , m}$.
 \end{nitheorem}
+
+\begin{proof}
+  By induction on the typing derivation on $P$.
+  To be able to proceed under scope restriction, we need to generalize the theorem so that the variable that is being substituted need not be $\constr{zero}$, but any variable $i$.
+  We introduce such generalization in Theorem \ref{th:subst-generalization}.
+  \todo{explain arguments}
+\end{proof}
+
+\begin{nitheorem}
+  \label{th:subst-generalization}
+
+  We introduce an intermediate usage context $\Psi_m$.
+  This intermediate context is the result of exhausting some $\Delta$ from the input context $\Gamma$, such that $\opctx{\Gamma}{\Delta}{\Psi_m}$.
+  From $\Psi_m$, exhausting $m$ at index $i$ yields $\Psi_l$; exhausting it at index $j$ yields $\Psi_j$, such that $\contains{\gamma}{\Psi_m}{i}{t}{m}{\Psi_l}$ and $\contains{\gamma}{\Psi_m}{j}{t}{m}{\Psi_r}$.
+  The input derivation on $P$ takes context $\Gamma$ and yields context $\Psi_l$, such that $\types{\gamma}{\Gamma}{P}{\Psi_l}$.
+  The output derivation on $\subst{P}{j}{i}$ takes the same context $\Gamma$ but yields back $\Psi_r$, such that $\types{\gamma}{\Gamma}{\subst{P}{j}{i}}{\Psi_r}$.
+
+  \begin{remark}
+    \label{re:delta-zero}
+    It is of uttermost importance that the variable that is being substituted has \emph{all} of its multiplicities substituted.
+    To enforce this, we add an additional premise: that in going from $\Gamma$ to the intermediate context $\Psi_m$ we must exhaust no multiplicities at index $i$ -- so that $m$ indeed captures the totality of the multiplicities that are being exhausted between $\Gamma$ and $\Psi_l$.
+    We capture this requirement by enforcing $\Delta_i \equiv \ell_{\o}$.
+  \end{remark}
+\end{nitheorem}
+
+\begin{figure}[ht]
+  \begin{subfigure}{.5\textwidth}
+  \centering
+  \begin{tikzpicture}
+    \node (gamma) at (0,2) {$\Gamma$};
+    \node (psi-m) at (4,2) {$\Psi_m$};
+    \node (psi-l) at (4,4) {$\Psi_l$};
+    \node (psi-r) at (4,0) {$\Psi_r$};
+    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
+    \draw[->]                (gamma) -- node[above,sloped] {$\vdash P$}               (psi-l);
+    \draw[->,densely dotted] (gamma) -- node[below,sloped] {$\vdash \subst{P}{j}{i}$} (psi-r);
+    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
+    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
+  \end{tikzpicture}
+  \caption{Diagrammatic representation of the generalized substitution theorem.}
+  \label{fig:subst}
+  \end{subfigure}
+  % Do not leave empty line
+  \begin{subfigure}{.5\textwidth}
+  \centering
+  \begin{tikzpicture}
+    \node (gamma) at (0,2) {$\Gamma$};
+    \node (xi-l)  at (0,4) {$\Xi_l$};
+    \node (xi-r)  at (0,0) {$\Xi_r$};
+    \node (psi-m) at (4,2) {$\Psi_m$};
+    \node (psi-l) at (4,4) {$\Psi_l$};
+    \node (psi-r) at (4,0) {$\Psi_r$};
+    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
+    \draw[->]                (gamma) -- node[left]         {$\ell_i@i$}               (xi-l);
+    \draw[->,densely dotted] (gamma) -- node[left]         {$\ell_i@j$}               (xi-r);
+    \draw[->]                (xi-l)  -- node[above]        {$\vdash P$}               (psi-l);
+    \draw[->,densely dotted] (xi-r)  -- node[above]        {$\vdash \subst{P}{j}{i}$} (psi-r);
+    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
+    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
+  \end{tikzpicture}
+  \caption{
+    Diagrammatic representation of the generalized substitution for process input.
+    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, or $\opsquared{m}{\delta}{\ell_i}$.
+  }
+  \label{fig:subst-input}
+  \end{subfigure}
+\end{figure}
+
+\begin{proof}
+  \todo{BIG TODO}
+\end{proof}
 

### 2020-04-13 paper/paper.tex
Introduce subject congruence

@@
 
+\todo{lemma for lift-lower}
+
 Structural congruence is a congruent equivalence relation.
@@
 
+\begin{lemma}
+\todo{introduce entails-unused}
+\end{lemma}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+Structural congruence is an equivalence relation that aims to overcome the semantically meaningless difference introduced by the syntax of the \picalc{}.
+
 \begin{nitheorem}
+  Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
   If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
+
+\begin{proof}
+  By induction on \textsc{EQUALS}.
+  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
+  For those, we proceed by induction on \textsc{StructCong}.
+  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
+  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\S \ref{framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
+  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
+  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
+  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
+  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
+  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
+  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
+  We will then use strengthening (\S \ref{strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
+  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
+  We will then use Lemma \ref{lift-lower} to show that lifting a lowered $P$ leaves $P$ unchanged.
+  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \S \ref{swapping}.
+  In the reverse direction, we will use swapping and then show that swapping indices $i$ and $\suc i$ in process $P$ two times leaves $P$ unchanged.
+\end{proof}
+\todo{make sure constructor names in code coincide}
 
@@
     Diagrammatic representation of the generalized substitution for process input.
-    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, or $\opsquared{m}{\delta}{\ell_i}$.
+    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, i.e. $\opsquared{m}{\delta}{\ell_i}$ for some $\delta$.
   }

### 2020-04-13 paper/paper.tex
Slightly shorten the syntax section

@@
 \newtheorem{nitheorem}[theorem]{Theorem}
+\newtheorem{nilemma}[theorem]{Lemma}
 
@@
 \subsection{Contribution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Notation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Abstraction is one of the key reasoning tools in a language: it allows for features to be defined over a range of inputs.
-Such is the case of the \picalc{} too, where both scope restriction and input introduce abstractions.
-The bodies of these constructs must have a way of referring to their argument.
+In the \picalc{} both scope restriction and input introduce abstractions.
+The bodies of these constructs need a way of referring to their argument.
 Using names as variable references is a popular option amongst humans.
-However, names are cumbersome to mechanize: inserting a new variable into an environment means proving that the name of such variable is different to all other variable names in the environment.
-Moreover, to a machine names are of no significance whatsoever.
-
-Machines prefer things they can algorithmically act upon. Like natural numbers!
-What does it mean to use a number as a variable reference?
-The idea de Bruijn had \cite{} was to use the index $n$ to refer to the variable introduced $n$ binders ago.
-The binders themselves introduce no names anymore.
-The expression $\lambda g . (\lambda f . f g) g$ in the \lambdacalc{} would translate as $\lambda (\lambda 0 1) 0$.
-That is, terms at different \emph{depths} must use different indices to refer to the same binding.
-Humans find this often confusing.
-There is however no reason not to keep the original names bestowed by the humans together with the indices.
-Machines can then manipulate references mechanically and still use names to present them to humans.
+However, names are cumbersome to mechanize: inserting a new variable into a context means proving that the name of such variable is different from all other variable names in context.
+Moreover, names are of no significance to machines.
+
+Machines need constructs they can algorithmically act upon.
+The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago.
+Binders no longer introduce any names.
+\begin{example}
+The expression $\lambda g . (\lambda f . f g) g$ in the \lambdacalc{} would translate to $\lambda (\lambda 0 1) 0$.
+\end{example}
+That is, terms at different \emph{depths} use different indices to refer to the same binding.
+Humans often find this confusing.
+There is however no reason not to annotate these indices with the original names bestowed by humans.
+Machines can then manipulate references mechanically while still using names to present terms to humans.
 
@@
 It is useful to rule out these nonsensical terms syntactically.
-In Figure \ref{var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+In Figure \ref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 
@@
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-As shown in Figure \ref{fig:process}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
+As shown in Figure \ref{fig:syntax}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
 
@@
 
-Unlike with names, using type-level de Bruijn indices makes our syntax well-scoped by construction.
-As a consequence, the semantics of our language can be defined on the totality of the syntax.
+Using type-level de Bruijn indices makes our syntax well-scoped by construction, and the semantics of our language can now be defined on the totality of the syntax.
 User-friendliness can still be recovered through a function that converts processes with names into (possibly) processes with indices.
-This function would keep track of what index is associated with what name, and would traverse the process recursively, taking note of new binders and substituting variable references.
-If the process is ill-scoped, the function would return nothing.
-To print things back to the user names can be substituted with an index together with the name.
+This function keeps track of what index is associated with what name, and traverses the process recursively, taking note of new binders and substituting variable references.
+If the process is ill-scoped, the function returns nothing.
+To present things back to the user, we print indices as their named annotations.
 
@@
 \subsection{Structural Congruence}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{structural-congruence}
@@
 \subsection{Operational Semantics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{operational-semantics}
@@
 \subsection{Multiplicities and Capabilities}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{multiplicities}
@@
 \subsubsection{Example Type Systems}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Contexts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{contexts}
@@
 \subsection{Typing with Leftovers}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{leftover-typing}
@@
 
-\begin{lemma}
+\begin{nilemma}
   The typing context $\gamma$ at index $i$ is type $t$.
-\end{lemma}
-\begin{lemma}
+\end{nilemma}
+\begin{nilemma}
   The input annotation in context $\Gamma$ at index $i$ is the result of aggregating (as per the algebra) $y$ and the output annotation in context $\Delta$ at index $i$.
-\end{lemma}
-\begin{lemma}
+\end{nilemma}
+\begin{nilemma}
   For every index $j$ other than $i$, the input annotation in context $\Gamma$ at index $i$ is the output annotation in $\Delta$ at index $i$.
-\end{lemma}
-\begin{lemma}
+\end{nilemma}
+\begin{nilemma}
   \todo{Add feedback and feedfront}
-\end{lemma}
+\end{nilemma}
 \begin{corollary}
@@
 
-\begin{lemma}
+\begin{nilemma}
 \todo{introduce entails-unused}
-\end{lemma}
+\end{nilemma}
 
@@
 
+\begin{nilemma}
+  \label{lm:comm-capable}
+  \todo{reformulate}
+  A well-typed process on which reduction occurs externally, must have this external multiplicity be capable of $\ell_\#$.
+\end{nilemma}
+
 \begin{proof}
+  For the base case $\constr{comm}$ \todo{}
+
+  Reduction under composition with $\constr{par}$ \todo{}
+
+  Reduction under scope restriction case splits on \textsc{Channel} $c$ on which communication occurs.
+  If this channel is internal to the process, or if it is external but on a channel other then the one introduced by the recent scope restriction, subject reduction proceeds inductively.
+  However, if reduction happens on the most recently introduced channel, we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity, then apply the induction hypothesis.
+
+  Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and then the induction hypothesis.
+
 \end{proof}
@@
 \subsection{Framing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Weakening}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Strengthening}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Swapping}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \subsection{Subject Congruence}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{subject-congruence}
@@
 \begin{nitheorem}
+  \label{thm:subject-congruence}
   Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
@@
 \subsection{Substitution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2020-04-13 paper/paper.tex
Add weakening and strengthening theorems

@@
 In the \picalc{} however, the syntax introduces unnecessary distinctions
-(e.g. semantically parallel composition is defined modulo associativity and commutativity).
+(e.g. semantically, parallel composition is defined modulo associativity and commutativity).
 There are several ways around this, a structural congruence relation being one of the historical ones.
@@
 Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.
-The type $\Unused_0 \; Q$ witnesses that index $0$ does not appear nor in the inputs nor in the outputs of process $Q$ --- it does so by traversing $Q$.
-The function $lower \; 0 \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $0$.
-Finally, $swap \; 0 \; P$ traverses $P$ (scoped under $\suc \suc n$) and swaps variable references $0$ and $\suc 0$.
+The type $\Unused_i \; Q$ is an inductive proof on $Q$ that witnesses that the index $i$ does not appear nor in the inputs nor in the outputs of the process.
+The function $lower \; i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+Finally, $swap \; i \; P$ traverses $P$ (scoped under $\suc \suc n$) and swaps variable references $i$ and $\suc i$.
 
@@
 
-\todo{lemma for lift-lower}
+\begin{nilemma}
+  \label{lm:lower-lift}
+  The function $lower \; i \; P \; uP$ has an inverse $lift \; i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift \; i \; (lower \; i \; P \; uP) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:swap-swap}
+  The function $swap \; i \; P$ is its own inverse: $swap \; i \; (swap \; i \; P) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 In \S \ref{subject-congruence} we will prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then $Q$ is well-typed.
-Specifically, in the case of transitivity we must prove that if $P$ is structurally congruent with $Q$ and $Q$ with $R$, and $P$ is well-typed, then so is $R$.
+Specifically, in the case of transitivity, we must prove that if $P$ is structurally congruent with $Q$ and $Q$ with $R$, and $P$ is well-typed, then so is $R$.
 To do so, we will have to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
@@
 \begin{nilemma}
-\todo{introduce entails-unused}
+  \label{lm:types-unused}
+  If $P$ is well-typed under input context $\Gamma$ and output context $\Xi$, and variable $i$ is unused within $P$, then $\Gamma_i \equiv \Xi_i$.
 \end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:types-op}
+  For any well typed-process $P$ with input context $\Gamma$ and output context $\Xi$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 
+Framing captures the idea that the well-typedness of a process does not depend on its unused resources.
+These unused resources are not introduced as new variables -- the typing context remains unaltered -- but rather as extra multiplicities that are carried from the input context to the output context.
+
 \begin{nitheorem}
-  Let $\opctx{\Gamma_l}{\Delta}{\Xi_l}$ and $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-  If $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$ then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+  Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
+  Then, per Lemma \ref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
 \end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 
-Order preserving embeddings model a series of insertions. We only ever need one insertion to prove subject congruence, but there is no loss of generality.
+Weakening states that adding new variables to the typing and usage contexts preserves the well-typedness of a process.
+The multiplicities of these new variables will be carried from the input context to the output context.
+Usually, weakening is presented through order preserving embeddings.
+Order preserving embeddings model a series of insertions into a context.
+In this paper, we use weakening to prove subject congruence.
+To do so, we need of a single insertion into a context, and thus we will limit our theorem statement to that, but there is no loss of generality.
 
 \begin{nitheorem}
-  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let us insert an arbitrary type $t$ at position $i$ in the typing context and an arbitrary multiplicity $x$ at position $i$ into both the input and output contexts.
+  Let us reflect this on the process $P$ by incrementing every index greater than or equal to $i$.
+  The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
 \end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 
+Strengthening states that removing unused variables from the typing and usage contexts preserves the well-typedness of a process.
+The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations between the input and output contexts: a process can refer to a variable and still make no use of it -- e.g.
+by sending none of its multiplicities over a channel.
+
 \begin{nitheorem}
-  If $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\Unused \; i \; P$, then $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused \; i \; P$.
+  Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
 \end{nitheorem}
-
-\begin{remark}
-  The proof that index $i$ is unused in $P$ must be syntactical and cannot rely on the usage annotation of index $i$ being preserved between $\Gamma$ and $\Xi$: process $P$ can send variable reference $i$ with multiplicity $\zero$ over a channel.
-\end{remark}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
   For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
-  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
+  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
   We will then use strengthening (\S \ref{strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
   In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
-  We will then use Lemma \ref{lift-lower} to show that lifting a lowered $P$ leaves $P$ unchanged.
+  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
   Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \S \ref{swapping}.
-  In the reverse direction, we will use swapping and then show that swapping indices $i$ and $\suc i$ in process $P$ two times leaves $P$ unchanged.
+  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
 \end{proof}
 \todo{make sure constructor names in code coincide}
+
+\begin{remark}
+  We encode the recursive structure of \textsc{Equals} in its inductive definition -- by indexing its type with \textsc{Rec}.
+  This acts as a justification of the well-formedness of the recursion scheme, which Agda cannot immediately see as terminating due to the interaction between the transitive and symmetric cases.
+\end{remark}
 

### 2020-04-20 paper/paper.tex
some comments on the paper

@@
 \maketitle
-\todo{remove todo notes}
 
 \begin{abstract}
+  \todo{OD: the abstract requires an opening. Before diving into what we present, it is useful to give context of what the problem is and why it is important.}
   We present the syntax, operational semantics, and typing rules of a \picalc{} with linear and shared types.
@@
   This formalization has been fully mechanized with Agda and is available at \url{https://github.com/umazalakain/typing-linear-pi}.
+  \todo{OD: remove links and put them as reference under @misc}
 \end{abstract}
@@
 \section{Introduction}
-
+\todo{OD:
+It is important to emphasise that this is the first fully formalised pi-calculus from syntax to OS to typing rules, spanning from linear to shared types, "gradually".}
 We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
@@
 The \picalc{} \cite{} is a formal language that abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
+\todo{OD: The key point here is that the pi-calculus, especically the linear one, is a sufficcient core calculus where all research on concurrent and ditrubuted systems can be achieved.
+That is why we are inspecting it.
+The goal is to bring all distributed systems features (in the future) into the formalised world.}
 Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted too.
@@
 We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every unbound variable is accounted for in the type of the process that uses it.
-
+\todo{OD: don't put references to section just yet. This first part should describe the what and why of the problem.
+In the contributions section below, we can use references to sections to show where those contributions can be found.}
 In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
@@
 Some constructors can be shared across types and disambiguated by context.
-
+\todo{OD: the next part to be filled should be the notation, so that when I go through the paper, I can use this as a manual.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Binders no longer introduce any names.
+\todo{OD: IMPORTATNT!
+There is too much on lambda-calculus; I suggest you give examples and illustrate your points on the pi-calculus only from now on.
+It can get too broad and lack focus otherwise...
+It is fine in the Introduction as you are giving an overview.}
 \begin{example}
@@
 All other rules follow trivially from the algebraic rules for the addition of naturals.
-
+\todo{OD: in addition to the above, could we have examples of typing a process in each of the above type systems? It can be the same process everywhere and we can show how the typing changes...}
 \end{description}

### 2020-04-24 paper/paper.tex
Update substitution diagram

@@
 
-\begin{figure}[ht]
+\begin{figure}[h]
   \begin{subfigure}{.5\textwidth}
@@
     \begin{tikzpicture}
-    \node (gamma) at (0,2) {$\Gamma$};
-    \node (psi-m) at (4,2) {$\Psi_m$};
-    \node (psi-l) at (4,4) {$\Psi_l$};
-    \node (psi-r) at (4,0) {$\Psi_r$};
-    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
-    \draw[->]                (gamma) -- node[above,sloped] {$\vdash P$}               (psi-l);
-    \draw[->,densely dotted] (gamma) -- node[below,sloped] {$\vdash \subst{P}{j}{i}$} (psi-r);
-    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
-    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
-  \end{tikzpicture}
-  \caption{Diagrammatic representation of the generalized substitution theorem.}
-  \label{fig:subst}
-  \end{subfigure}
-  % Do not leave empty line
-  \begin{subfigure}{.5\textwidth}
-  \centering
-  \begin{tikzpicture}
-    \node (gamma) at (0,2) {$\Gamma$};
-    \node (xi-l)  at (0,4) {$\Xi_l$};
-    \node (xi-r)  at (0,0) {$\Xi_r$};
-    \node (psi-m) at (4,2) {$\Psi_m$};
-    \node (psi-l) at (4,4) {$\Psi_l$};
-    \node (psi-r) at (4,0) {$\Psi_r$};
-    \draw[->]                (gamma) -- node[above,sloped] {$\Delta$} node[below] {$\Delta_i = \ell_{\o}$} (psi-m);
-    \draw[->]                (gamma) -- node[left]         {$\ell_i@i$}               (xi-l);
-    \draw[->,densely dotted] (gamma) -- node[left]         {$\ell_i@j$}               (xi-r);
-    \draw[->]                (xi-l)  -- node[above]        {$\vdash P$}               (psi-l);
-    \draw[->,densely dotted] (xi-r)  -- node[above]        {$\vdash \subst{P}{j}{i}$} (psi-r);
-    \draw[->]                (psi-m) -- node[right]        {$m@i$}                    (psi-l);
-    \draw[->]                (psi-m) -- node[right]        {$m@j$}                    (psi-r);
+      \node (gamma-l) at (0,6)   {$\Gamma_l$};
+      \node (gamma-m)   at (0,3)   {$\Gamma$};
+      \node (gamma-r) at (0,0)   {$\Gamma_r$};
+      \node (xi-l)    at (3,5.5) {$\Xi_l$};
+      \node (theta)   at (3,3.3) {$\Theta$};
+      \node (delta-m) at (3,2.7) {};
+      \node (xi-r)    at (3,0.5) {$\Xi_r$};
+      \node (psi-l)   at (6,5)   {$\Psi_l$};
+      \node (psi-m)   at (6,3)   {$\Psi$};
+      \node (psi-r)   at (6,1)   {$\Psi_r$};
+
+      \draw[-]  (gamma-m) -- (delta-m.center);
+      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \ell_{\o}$}(psi-m);
+
+      \draw[->,densely dotted] (gamma-m) -- (theta);
+      \draw[->,densely dotted] (theta) -- (psi-m);
+
+      \draw[->] (gamma-l) -- node[left] {$\ni_i m$} (gamma-m);
+      \draw[->] (gamma-r) -- node[left] {$\ni_j m$} (gamma-m);
+      \draw[->] (psi-l) -- node[right] {$\ni_i n$} (psi-m);
+      \draw[->] (psi-r) -- node[right] {$\ni_j n$} (psi-m);
+
+      \draw[->] (gamma-l) -- node[above] {$\vdash P$} (xi-l);
+      \draw[->] (xi-l) -- node[above] {$\vdash Q$} (psi-l);
+      \draw[->,densely dotted] (gamma-r) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-r);
+      \draw[->,densely dotted] (xi-r) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-r);
+      \draw[->,densely dotted] (xi-l) -- node[left] {$\ni_i l$} (theta);
+      \draw[->,densely dotted] (xi-r) -- node[left] {$\ni_j l$} (theta);
     \end{tikzpicture}
@@
       Diagrammatic representation of the generalized substitution for process input.
-    Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $\ell_i$, i.e. $\opsquared{m}{\delta}{\ell_i}$ for some $\delta$.
+      Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $n$, i.e. $\opsquared{m}{\delta}{n}$ for some $\delta$.
     }
-  \label{fig:subst-input}
+    \label{fig:subst}
+  \end{subfigure}
+  \hspace{\fill} % DO NOT LEAVE EMPTY LINE
+  \begin{subfigure}{.4\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \node (gamma-l) at (0,4)   {};
+      \node (gamma-m) at (0,2)   {};
+      \node (gamma-r) at (0,0)   {};
+      \node (xi-l)    at (2,5)   {};
+      \node (xi-m)    at (2,3.5) {};
+      \node (xi-r)    at (2,2)   {};
+      \node (psi-m)   at (4,3.5) {};
+
+      \draw[->] (gamma-l) -- (xi-r);
+      \draw[->] (gamma-m) -- (xi-r);
+      \draw[->] (gamma-r) -- (xi-r);
+      \draw[->] (gamma-l) -- (gamma-m);
+      \draw[->] (gamma-r) -- (gamma-m);
+      \draw[->] (xi-l)    -- (xi-m);
+      \draw[->] (xi-r)    -- (xi-m);
+      \draw[->] (xi-l)    -- (psi-m);
+      \draw[->] (xi-m)    -- (psi-m);
+      \draw[->] (xi-r)    -- (psi-m);
+    \end{tikzpicture}
+    \caption{
+      Alternative approach without the $\ni_i n$ and $\ni_j n$ arrows.
+      These are not strictly necessary, but they make the inductive hypotheses compositional.
+      Without them, framing lemmas would have to be used extensively.
+    }
+    \label{fig:subst-compositionality}
   \end{subfigure}

### 2020-04-27 paper/paper.tex
Comment on subject reduction

@@
 % Functions
-\newcommand{\subst}[3]{#1[#2/#3]}
+\newcommand{\subst}[3]{#1 \; [ \; #2 \; / \;#3 \;]}
 \newcommand{\opsquared}[3]{#1 \, \coloneqq \, #2 \, \cdot^2 \, #3}
@@
 \newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{\ni}_{#3} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
+\newcommand{\containsusage}[4]{#1 \; \type{\ni}_{#2} \; #3 \; \type{\boxtimes} \; #4}
 \newcommand{\Var}{\type{VAR}}
@@
 Processes put in parallel reduce when they communicate over a common variable.
-The receiving process substitutes references to its most immediate variable with references to the variable sent by the process doing the output, and is then \emph{lowered} --- all variable references are decreased by one.
+The receiving process substitutes references to its most immediate variable with references to the variable sent by the process doing the output.
+The process is then \emph{lowered} --- all variable references are decreased by one.
 Reduction is closed under structural congruence and goes under parallel composition and scope restriction
@@
   \inferrule
-  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}
+  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower \; 0 \; (\subst{P}{j}{0})}{Q}}
 
@@
 In the linear \picalc{} however, reduction \emph{is} meant to transform the context: each time communication occurs over a channel, that channel is exhausted.
-In our case, and because we generalize linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity on both its input and output capabilities.
-In \S \ref{semantics}, we made sure to lift the index of this channel to the type level.
-If the channel is internal to the process, the context must not be modified; if it is external on index $i$, the input usage context at index $i$ must have $\ell_\#$ subtracted.
+In our case, and because we generalize linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity $\one$ on both its input and output capabilities.
+In \S \ref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
 
@@
   Let $P \reduce{c} Q$.
-  If $c$ is $\constr{internal}$, then let $\Delta \equiv \epsilon$;
-  if $c$ is $\constr{external}\; i$, then let $\Delta \equiv \ell_\# @ i$.
-  Let $\Gamma'$ and $\Gamma$ be usage contexts only differing in $\Delta$, so that $\opctx{\Gamma'}{\Delta}{\Gamma}$.
-  If $\types{\gamma}{\Gamma'}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$
+  If $c$ is $\constr{internal}$ to the process, the context must not be modified and thus $\Gamma' \equiv \Gamma$.
+  If $c$ is $\constr{external}\; i$, the input usage context at index $i$ must be decremented, and therefore $\containsusage{\Gamma'}{i}{\ell_\#}{\Gamma}$.
+  Then the typing derivation $\types{\gamma}{\Gamma'}{P}{\Xi}$ can be transformed into $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
+
+\begin{proof}
+  \hfill{}\\
+  \begin{itemize}
+    \item
+    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ from the following bits:
+    \begin{alignat*}{2}
+      \containsusage{&\Gamma'}{i}{\ell_\# &&}{\Gamma} \\
+      \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\ell_i &&}{\Psi} \\
+      \types{\gamma , t}{&\Psi , m}{P &&}{\Delta , \ell_{\o}} \\
+      \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\ell_o &&}{\Theta} \\
+      \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
+    \end{alignat*}
+
+    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel, so that we obtain the following:
+    \begin{alignat*}{2}
+      \types{\gamma , t}{& \Gamma , m}{P &&}{\Theta , \ell_{\o}} \\
+      \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
+    \end{alignat*}
+
+    We then use substitution (Theorem \ref{thm:substitution}) to obtain the following:
+    \begin{align*}
+      \types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}
+    \end{align*}
+
+    Finally, we use strengthening (Theorem \ref{thm:strengthening}) to lower the substituted variable index:
+    \begin{align*}
+      \types{\gamma , t}{\Gamma , m}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
+    \end{align*}
+
+    \item
+    Reduction under composition with $\constr{par}$ proceeds by induction on the process that is being reduced.
+
+    \item
+    Reduction under scope restriction case splits on \textsc{Channel} $c$ on which communication occurs.
+    If this channel is internal to the process, or if it is external but on a channel other then the one introduced by the recent scope restriction, subject reduction proceeds inductively.
+    However, if reduction happens on channel that is being introduced, we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity first, then apply the induction hypothesis.
+
+    \item
+    Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and then the induction hypothesis.
+  \end{itemize}
+\end{proof}
 
@@
   \label{lm:comm-capable}
-  \todo{reformulate}
-  A well-typed process on which reduction occurs externally, must have this external multiplicity be capable of $\ell_\#$.
+  Every input usage context $\Gamma$ on a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces such that $P \reduce{\constr{external} \; i} Q$ has a multiplicity of at least $\ell_\#$ at index $i$.
 \end{nilemma}
-
 \begin{proof}
-  For the base case $\constr{comm}$ \todo{}
-
-  Reduction under composition with $\constr{par}$ \todo{}
-
-  Reduction under scope restriction case splits on \textsc{Channel} $c$ on which communication occurs.
-  If this channel is internal to the process, or if it is external but on a channel other then the one introduced by the recent scope restriction, subject reduction proceeds inductively.
-  However, if reduction happens on the most recently introduced channel, we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity, then apply the induction hypothesis.
-
-  Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and then the induction hypothesis.
-
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
 \end{proof}
@@
 \begin{nitheorem}
+  \label{thm:framing}
   Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
@@
 \begin{nitheorem}
+  \label{thm:weakening}
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
@@
 \begin{nitheorem}
+  \label{thm:strengthening}
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
@@
 \begin{nitheorem}
+  \label{thm:swapping}
   If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
 \end{nitheorem}
+
+\todo{}
 
@@
 \begin{nitheorem}
+  \label{thm:substitution}
   Let $P$ be a process where the innermost variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
@@
   To be able to proceed under scope restriction, we need to generalize the theorem so that the variable that is being substituted need not be $\constr{zero}$, but any variable $i$.
-  We introduce such generalization in Theorem \ref{th:subst-generalization}.
+  We introduce such generalization in Theorem \ref{thm:subst-generalization}.
   \todo{explain arguments}
@@
 \begin{nitheorem}
-  \label{th:subst-generalization}
+  \label{thm:subst-generalization}
 

### 2020-04-27 paper/paper.tex
Clarify substitution diagrams

--- 2020-04-24 paper/paper.tex
@@
     \caption{
-      Diagrammatic representation of the generalized substitution for process input.
-      Thanks to the additional premise in Remark \ref{re:delta-zero}, we know that $m$ is at least capable of $n$, i.e. $\opsquared{m}{\delta}{n}$ for some $\delta$.
+      Diagrammatic representation of the inductive substitution lemma.
+      Continuous lines represent known facts, dotted lines proof obligations.
+      From $\Delta_i = \ell_{\o}$, we can deduce that there exists some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+      From that, we can find first $\ni_i l$ and then $\ni_j l$, splitting the diagram across its vertical axis, making it possible to inductively derive both $\vdash \subst{P}{j}{i}$ and $\vdash \subst{P}{j}{i}$.
     }
@@
     \caption{
-      Alternative approach without the $\ni_i n$ and $\ni_j n$ arrows.
-      These are not strictly necessary, but they make the inductive hypotheses compositional.
+      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
+      These are not strictly necessary, but they ease the composition of the inductive hypotheses.
       Without them, framing lemmas would have to be used extensively.

### 2020-04-28 paper/paper.tex
new title and abstract

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Typing a linear \picalc}
+\title{Mechanising a linear \picalc}
 \author{Uma Zalakain}{University of Glasgow, Scotland}
@@
 \ccsdesc[300]{Theory of computation~Process calculi}
-\keywords{pi calculus, linear, types, concurrency}
+\keywords{pi-calculus, linear types, concurrency, mecahnization, Agda}
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
@@
 \begin{abstract}
-  \todo{OD: the abstract requires an opening. Before diving into what we present, it is useful to give context of what the problem is and why it is important.}
-  We present the syntax, operational semantics, and typing rules of a \picalc{} with linear and shared types.
+  The \picalc{} is a computational model for communication and concurrency.
+  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channel resources must be used exactly once.
+  With the rise of session types as a formalism to model real-world distributed system, the linear \picalc{} re-emerged as the underlying theoretical and practical framework for session types and communication-based distributed programming \cite{DardhaGS12,Padovani17,ScalasY16,ScalasDHY17}.
+  While there have been several extensions to session typed \picalc{}, the linear \picalc{} remains essentially unchanged and a foundational calculus.
+
+  We present the \emph{first} fully mechanised \picalc{} featuring linear types, as well as gradual and shared types.
+  We present syntax, operational semantics, and typing rules and the corresponding type safety properties.
   We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
-  We generalize the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, gradual and shared typing.
+  We generalise the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, gradual and shared typing.
   We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
@@
 
-  This formalization has been fully mechanized with Agda and is available at \url{https://github.com/umazalakain/typing-linear-pi}.
-  \todo{OD: remove links and put them as reference under @misc}
+  The formalisation is fully mechanised in Agda and is available at \url{https://github.com/umazalakain/typing-linear-pi}.
 \end{abstract}

### 2020-04-28 paper/paper.tex
Minor work on subject reduction

--- 2020-04-27 paper/paper.tex
@@
     \item
-    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ from the following bits:
+    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in Figure \ref{fig:subject-reduction:a}.
+    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (Theorem \ref{thm:substitution}) and strengthening (Theorem \ref{thm:strengthening}).
+
+    \begin{figure}[h]
+      \setlength{\mathindent}{0pt}
+      \begin{subfigure}{.3\textwidth}
         \begin{alignat*}{2}
@@
         \end{alignat*}
-
-    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel, so that we obtain the following:
+        \caption{Initial assumptions.}
+        \label{fig:subject-reduction:a}
+      \end{subfigure}
+      % Do not leave empty space
+      \begin{subfigure}{.5\textwidth}
         \begin{alignat*}{2}
@@
         \end{alignat*}
-
-    We then use substitution (Theorem \ref{thm:substitution}) to obtain the following:
+        \caption{Rearrange $\ell_o$ witness, and merge with $\ell_i$ witness into $\ell_\#$.}
+      \end{subfigure}
+      % Do not leave empty space
+      \begin{subfigure}{.5\textwidth}
         \begin{align*}
@@
         \end{align*}
-
-    Finally, we use strengthening (Theorem \ref{thm:strengthening}) to lower the substituted variable index:
+        \caption{Substitution.}
+      \end{subfigure}
+      % Do not leave empty space
+      \begin{subfigure}{.5\textwidth}
         \begin{align*}
@@
         \end{align*}
+        \caption{Strengthening.}
+      \end{subfigure}
+      \caption{Proof steps for the $\constr{comm}$ case in subject reduction.}
+    \end{figure}
 
@@
 \begin{proof}
-  By induction on \textsc{EQUALS}.
+  By induction on \textsc{Equals}.
   The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
@@
   The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
-  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\S \ref{framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
+  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
   The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
@@
   In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
-  We will then use strengthening (\S \ref{strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
+  We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
   In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
   We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
-  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \S \ref{swapping}.
+  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
   In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.

### 2020-04-28 paper/paper.tex
some references and start of intro

@@
 \section{Introduction}
-\todo{OD:
-It is important to emphasise that this is the first fully formalised pi-calculus from syntax to OS to typing rules, spanning from linear to shared types, "gradually".}
 We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
 As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-The \picalc{} \cite{} is a formal language that abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
+The \picalc{} \cite{MilnerPW92,Milner99,Sangio01} is a computational model for communication and concurrency.
+It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
 \todo{OD: The key point here is that the pi-calculus, especically the linear one, is a sufficcient core calculus where all research on concurrent and ditrubuted systems can be achieved.
@@
 The goal is to bring all distributed systems features (in the future) into the formalised world.}
-Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted too.
+Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted, too.
+The \emph{linear} \picalc{} \cite{KPT96,Sangio01} is a typed version of the \picalc{} where the linear type system restricts the usage of channel resources to exactly ones.
+
 We type the syntax for the \picalc{} in \S \ref{syntax}.

### 2020-04-28 paper/paper.tex
Minor grammar changes to the abstract

--- 2020-04-28 paper/paper.tex
@@
   The \emph{linear} \picalc{} is a typed version of the \picalc{} where channel resources must be used exactly once.
-  With the rise of session types as a formalism to model real-world distributed system, the linear \picalc{} re-emerged as the underlying theoretical and practical framework for session types and communication-based distributed programming \cite{DardhaGS12,Padovani17,ScalasY16,ScalasDHY17}.
-  While there have been several extensions to session typed \picalc{}, the linear \picalc{} remains essentially unchanged and a foundational calculus.
+  With the rise of session types as a formalism to model real-world distributed systems, the linear \picalc{} has re-emerged as the underlying theoretical and practical framework for session types and communication-based distributed programming \cite{DardhaGS12,Padovani17,ScalasY16,ScalasDHY17}.
+  While there have been several extensions to the session typed \picalc{}, the linear \picalc{} remains essentially unchanged and stable as a foundational calculus.
 
   We present the \emph{first} fully mechanised \picalc{} featuring linear types, as well as gradual and shared types.
-  We present syntax, operational semantics, and typing rules and the corresponding type safety properties.
+  We present its syntax, operational semantics, and typing rules and the corresponding type safety properties.
   We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
@@
 
-  The formalisation is fully mechanised in Agda and is available at \url{https://github.com/umazalakain/typing-linear-pi}.
+  Our formalisation is fully mechanised in Agda. \cite{Zalakain2020Agda}
 \end{abstract}

### 2020-04-28 paper/paper.tex
Fix keywords
Finish introducing subject reduction

@@
 \ccsdesc[300]{Theory of computation~Process calculi}
-\keywords{pi-calculus, linear types, concurrency, mecahnization, Agda}
+\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanization, Agda}
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
@@
     \item
-    Reduction under scope restriction case splits on \textsc{Channel} $c$ on which communication occurs.
-    If this channel is internal to the process, or if it is external but on a channel other then the one introduced by the recent scope restriction, subject reduction proceeds inductively.
-    However, if reduction happens on channel that is being introduced, we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity first, then apply the induction hypothesis.
+    Reduction under scope restriction case splits on the channel $c$ on which communication occurs.
+    If the channel is $\constr{internal}$ to the process subject reduction proceeds inductively.
+    If the channel is $\constr{external}\; \constr{zero}$ -- i.e.
+the channel that is being introduced by the scope restriction -- we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity first, then apply the induction hypothesis.
+    If the channel is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
 

### 2020-04-28 paper/paper.tex
Draft swapping

@@
 
+Swapping two variables (in the typing context, input and output usage contexts and process) preserves the well-typedness of a process.
+
+\begin{remark}
+  We limit ourselves to the swapping of contiguous variables $i$ and $\suc i$.
+  Doing so is enough to prove subject reduction.
+  The swapping of any two variables can be modeled as a sequence of contiguous variable swappings.
+\end{remark}
+
 \begin{nitheorem}
   \label{thm:swapping}
-  If $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
+  Let $swap_i$ be a function that swaps variables $i$ and $\suc i$.
+  For simplicity, we overload $swap$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
 \end{nitheorem}
-
-\todo{}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 

### 2020-04-28 paper/paper.tex
Change the order of the theorems

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject Reduction}
-\label{subject-reduction}
+\section{Type Safety}
+\label{type-safety}
+
+\todo{intro}
+
+\subsection{Framing}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Framing captures the idea that the well-typedness of a process does not depend on its unused resources.
+These unused resources are not introduced as new variables -- the typing context remains unaltered -- but rather as extra multiplicities that are carried from the input context to the output context.
+
+\begin{nitheorem}
+  \label{thm:framing}
+  Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
+  Then, per Lemma \ref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+\end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\subsection{Weakening}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Weakening states that adding new variables to the typing and usage contexts preserves the well-typedness of a process.
+The multiplicities of these new variables will be carried from the input context to the output context.
+Usually, weakening is presented through order preserving embeddings.
+Order preserving embeddings model a series of insertions into a context.
+In this paper, we use weakening to prove subject congruence.
+To do so, we need of a single insertion into a context, and thus we will limit our theorem statement to that, but there is no loss of generality.
+
+\begin{nitheorem}
+  \label{thm:weakening}
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let us insert an arbitrary type $t$ at position $i$ in the typing context and an arbitrary multiplicity $x$ at position $i$ into both the input and output contexts.
+  Let us reflect this on the process $P$ by incrementing every index greater than or equal to $i$.
+  The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+\end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\subsection{Strengthening}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Strengthening states that removing unused variables from the typing and usage contexts preserves the well-typedness of a process.
+The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations between the input and output contexts: a process can refer to a variable and still make no use of it -- e.g.
+by sending none of its multiplicities over a channel.
+
+\begin{nitheorem}
+  \label{thm:strengthening}
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused \; i \; P$.
+  Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+\end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\subsection{Swapping}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Swapping two variables (in the typing context, input and output usage contexts and process) preserves the well-typedness of a process.
+
+\begin{remark}
+  We limit ourselves to the swapping of contiguous variables $i$ and $\suc i$.
+  Doing so is enough to prove subject reduction.
+  The swapping of any two variables can be modeled as a sequence of contiguous variable swappings.
+\end{remark}
+
+\begin{nitheorem}
+  \label{thm:swapping}
+  Let $swap_i$ be a function that swaps variables $i$ and $\suc i$.
+  For simplicity, we overload $swap$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
+\end{nitheorem}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\subsection{Subject Congruence}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\label{subject-congruence}
+
+Structural congruence is an equivalence relation that aims to overcome the semantically meaningless difference introduced by the syntax of the \picalc{}.
+
+\begin{nitheorem}
+  \label{thm:subject-congruence}
+  Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
+  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+\end{nitheorem}
+
+\begin{proof}
+  By induction on \textsc{Equals}.
+  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
+  For those, we proceed by induction on \textsc{StructCong}.
+  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
+  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
+  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
+  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
+  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
+  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
+  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
+  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
+  We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
+  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
+  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
+  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
+  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
+\end{proof}
+\todo{make sure constructor names in code coincide}
+
+\begin{remark}
+  We encode the recursive structure of \textsc{Equals} in its inductive definition -- by indexing its type with \textsc{Rec}.
+  This acts as a justification of the well-formedness of the recursion scheme, which Agda cannot immediately see as terminating due to the interaction between the transitive and symmetric cases.
+\end{remark}
+
+\subsection{Substitution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the innermost variable in the body of the abstraction gets replaced with the argument of the application.
+In the \picalc{} however substitution takes another form: every reference to the innermost variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
+In a resource-aware \picalc{}, that means that instead of exhausting all $m$ multiplicities of the most recent variable, we leave those unaltered and exhaust them from the variable that is being received.
+
+\begin{nitheorem}
+  \label{thm:substitution}
+  Let $P$ be a process where the innermost variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
+  Let $j$ be the index of the data that is being received.
+  This data has at least $m$ multiplicities, such that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute the most recent variable by $j$.
+  As a result, we will exhaust no multiplicities on the innermost variable and exhaust them in $j$ instead, therefore $\types{\gamma , t}{\Gamma , m}{\subst{P}{j}{\constr{zero}}}{\Xi , m}$.
+\end{nitheorem}
+
+\begin{proof}
+  By induction on the typing derivation on $P$.
+  To be able to proceed under scope restriction, we need to generalize the theorem so that the variable that is being substituted need not be $\constr{zero}$, but any variable $i$.
+  We introduce such generalization in Theorem \ref{thm:subst-generalization}.
+  \todo{explain arguments}
+\end{proof}
+
+\begin{nitheorem}
+  \label{thm:subst-generalization}
+
+  We introduce an intermediate usage context $\Psi_m$.
+  This intermediate context is the result of exhausting some $\Delta$ from the input context $\Gamma$, such that $\opctx{\Gamma}{\Delta}{\Psi_m}$.
+  From $\Psi_m$, exhausting $m$ at index $i$ yields $\Psi_l$; exhausting it at index $j$ yields $\Psi_j$, such that $\contains{\gamma}{\Psi_m}{i}{t}{m}{\Psi_l}$ and $\contains{\gamma}{\Psi_m}{j}{t}{m}{\Psi_r}$.
+  The input derivation on $P$ takes context $\Gamma$ and yields context $\Psi_l$, such that $\types{\gamma}{\Gamma}{P}{\Psi_l}$.
+  The output derivation on $\subst{P}{j}{i}$ takes the same context $\Gamma$ but yields back $\Psi_r$, such that $\types{\gamma}{\Gamma}{\subst{P}{j}{i}}{\Psi_r}$.
+
+  \begin{remark}
+    \label{re:delta-zero}
+    It is of uttermost importance that the variable that is being substituted has \emph{all} of its multiplicities substituted.
+    To enforce this, we add an additional premise: that in going from $\Gamma$ to the intermediate context $\Psi_m$ we must exhaust no multiplicities at index $i$ -- so that $m$ indeed captures the totality of the multiplicities that are being exhausted between $\Gamma$ and $\Psi_l$.
+    We capture this requirement by enforcing $\Delta_i \equiv \ell_{\o}$.
+  \end{remark}
+\end{nitheorem}
+
+\begin{figure}[h]
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \node (gamma-l) at (0,6)   {$\Gamma_l$};
+      \node (gamma-m)   at (0,3)   {$\Gamma$};
+      \node (gamma-r) at (0,0)   {$\Gamma_r$};
+      \node (xi-l)    at (3,5.5) {$\Xi_l$};
+      \node (theta)   at (3,3.3) {$\Theta$};
+      \node (delta-m) at (3,2.7) {};
+      \node (xi-r)    at (3,0.5) {$\Xi_r$};
+      \node (psi-l)   at (6,5)   {$\Psi_l$};
+      \node (psi-m)   at (6,3)   {$\Psi$};
+      \node (psi-r)   at (6,1)   {$\Psi_r$};
+
+      \draw[-]  (gamma-m) -- (delta-m.center);
+      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \ell_{\o}$}(psi-m);
+
+      \draw[->,densely dotted] (gamma-m) -- (theta);
+      \draw[->,densely dotted] (theta) -- (psi-m);
+
+      \draw[->] (gamma-l) -- node[left] {$\ni_i m$} (gamma-m);
+      \draw[->] (gamma-r) -- node[left] {$\ni_j m$} (gamma-m);
+      \draw[->] (psi-l) -- node[right] {$\ni_i n$} (psi-m);
+      \draw[->] (psi-r) -- node[right] {$\ni_j n$} (psi-m);
+
+      \draw[->] (gamma-l) -- node[above] {$\vdash P$} (xi-l);
+      \draw[->] (xi-l) -- node[above] {$\vdash Q$} (psi-l);
+      \draw[->,densely dotted] (gamma-r) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-r);
+      \draw[->,densely dotted] (xi-r) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-r);
+      \draw[->,densely dotted] (xi-l) -- node[left] {$\ni_i l$} (theta);
+      \draw[->,densely dotted] (xi-r) -- node[left] {$\ni_j l$} (theta);
+    \end{tikzpicture}
+    \caption{
+      Diagrammatic representation of the inductive substitution lemma.
+      Continuous lines represent known facts, dotted lines proof obligations.
+      From $\Delta_i = \ell_{\o}$, we can deduce that there exists some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+      From that, we can find first $\ni_i l$ and then $\ni_j l$, splitting the diagram across its vertical axis, making it possible to inductively derive both $\vdash \subst{P}{j}{i}$ and $\vdash \subst{P}{j}{i}$.
+    }
+    \label{fig:subst}
+  \end{subfigure}
+  \hspace{\fill} % DO NOT LEAVE EMPTY LINE
+  \begin{subfigure}{.4\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \node (gamma-l) at (0,4)   {};
+      \node (gamma-m) at (0,2)   {};
+      \node (gamma-r) at (0,0)   {};
+      \node (xi-l)    at (2,5)   {};
+      \node (xi-m)    at (2,3.5) {};
+      \node (xi-r)    at (2,2)   {};
+      \node (psi-m)   at (4,3.5) {};
+
+      \draw[->] (gamma-l) -- (xi-r);
+      \draw[->] (gamma-m) -- (xi-r);
+      \draw[->] (gamma-r) -- (xi-r);
+      \draw[->] (gamma-l) -- (gamma-m);
+      \draw[->] (gamma-r) -- (gamma-m);
+      \draw[->] (xi-l)    -- (xi-m);
+      \draw[->] (xi-r)    -- (xi-m);
+      \draw[->] (xi-l)    -- (psi-m);
+      \draw[->] (xi-m)    -- (psi-m);
+      \draw[->] (xi-r)    -- (psi-m);
+    \end{tikzpicture}
+    \caption{
+      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
+      These are not strictly necessary, but they ease the composition of the inductive hypotheses.
+      Without them, framing lemmas would have to be used extensively.
+    }
+    \label{fig:subst-compositionality}
+  \end{subfigure}
+\end{figure}
+
+\begin{proof}
+  \todo{BIG TODO}
+\end{proof}
+
+\subsection{Subject Reduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection{Framing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Framing captures the idea that the well-typedness of a process does not depend on its unused resources.
-These unused resources are not introduced as new variables -- the typing context remains unaltered -- but rather as extra multiplicities that are carried from the input context to the output context.
-
-\begin{nitheorem}
-  \label{thm:framing}
-  Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
-  Then, per Lemma \ref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
-  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{nitheorem}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\subsection{Weakening}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Weakening states that adding new variables to the typing and usage contexts preserves the well-typedness of a process.
-The multiplicities of these new variables will be carried from the input context to the output context.
-Usually, weakening is presented through order preserving embeddings.
-Order preserving embeddings model a series of insertions into a context.
-In this paper, we use weakening to prove subject congruence.
-To do so, we need of a single insertion into a context, and thus we will limit our theorem statement to that, but there is no loss of generality.
-
-\begin{nitheorem}
-  \label{thm:weakening}
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let us insert an arbitrary type $t$ at position $i$ in the typing context and an arbitrary multiplicity $x$ at position $i$ into both the input and output contexts.
-  Let us reflect this on the process $P$ by incrementing every index greater than or equal to $i$.
-  The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
-\end{nitheorem}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\subsection{Strengthening}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Strengthening states that removing unused variables from the typing and usage contexts preserves the well-typedness of a process.
-The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations between the input and output contexts: a process can refer to a variable and still make no use of it -- e.g.
-by sending none of its multiplicities over a channel.
-
-\begin{nitheorem}
-  \label{thm:strengthening}
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused \; i \; P$.
-  Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
-\end{nitheorem}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\subsection{Swapping}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Swapping two variables (in the typing context, input and output usage contexts and process) preserves the well-typedness of a process.
-
-\begin{remark}
-  We limit ourselves to the swapping of contiguous variables $i$ and $\suc i$.
-  Doing so is enough to prove subject reduction.
-  The swapping of any two variables can be modeled as a sequence of contiguous variable swappings.
-\end{remark}
-
-\begin{nitheorem}
-  \label{thm:swapping}
-  Let $swap_i$ be a function that swaps variables $i$ and $\suc i$.
-  For simplicity, we overload $swap$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
-\end{nitheorem}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\subsection{Subject Congruence}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{subject-congruence}
-
-Structural congruence is an equivalence relation that aims to overcome the semantically meaningless difference introduced by the syntax of the \picalc{}.
-
-\begin{nitheorem}
-  \label{thm:subject-congruence}
-  Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
-  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-\end{nitheorem}
-
-\begin{proof}
-  By induction on \textsc{Equals}.
-  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
-  For those, we proceed by induction on \textsc{StructCong}.
-  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
-  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
-  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
-  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
-  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
-  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
-  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
-  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
-  We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
-  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
-  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
-  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
-  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
-\end{proof}
-\todo{make sure constructor names in code coincide}
-
-\begin{remark}
-  We encode the recursive structure of \textsc{Equals} in its inductive definition -- by indexing its type with \textsc{Rec}.
-  This acts as a justification of the well-formedness of the recursion scheme, which Agda cannot immediately see as terminating due to the interaction between the transitive and symmetric cases.
-\end{remark}
-
-\subsection{Substitution}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the innermost variable in the body of the abstraction gets replaced with the argument of the application.
-In the \picalc{} however substitution takes another form: every reference to the innermost variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
-In a resource-aware \picalc{}, that means that instead of exhausting all $m$ multiplicities of the innermost variable, we leave those unaltered and exhaust them from the element that is being sent.
-
-\begin{nitheorem}
-  \label{thm:substitution}
-  Let $P$ be a process where the innermost variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
-  Let $j$ be the index of the data that is being received.
-  This data has at least $m$ multiplicities, such that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the innermost variable by $j$.
-  As a result, we will exhaust no multiplicities on the innermost variable and exhaust them in $j$ instead, such that $\types{\gamma , t}{\Gamma , m}{\subst{P}{j}{\constr{zero}}}{\Xi , m}$.
-\end{nitheorem}
-
-\begin{proof}
-  By induction on the typing derivation on $P$.
-  To be able to proceed under scope restriction, we need to generalize the theorem so that the variable that is being substituted need not be $\constr{zero}$, but any variable $i$.
-  We introduce such generalization in Theorem \ref{thm:subst-generalization}.
-  \todo{explain arguments}
-\end{proof}
-
-\begin{nitheorem}
-  \label{thm:subst-generalization}
-
-  We introduce an intermediate usage context $\Psi_m$.
-  This intermediate context is the result of exhausting some $\Delta$ from the input context $\Gamma$, such that $\opctx{\Gamma}{\Delta}{\Psi_m}$.
-  From $\Psi_m$, exhausting $m$ at index $i$ yields $\Psi_l$; exhausting it at index $j$ yields $\Psi_j$, such that $\contains{\gamma}{\Psi_m}{i}{t}{m}{\Psi_l}$ and $\contains{\gamma}{\Psi_m}{j}{t}{m}{\Psi_r}$.
-  The input derivation on $P$ takes context $\Gamma$ and yields context $\Psi_l$, such that $\types{\gamma}{\Gamma}{P}{\Psi_l}$.
-  The output derivation on $\subst{P}{j}{i}$ takes the same context $\Gamma$ but yields back $\Psi_r$, such that $\types{\gamma}{\Gamma}{\subst{P}{j}{i}}{\Psi_r}$.
-
-  \begin{remark}
-    \label{re:delta-zero}
-    It is of uttermost importance that the variable that is being substituted has \emph{all} of its multiplicities substituted.
-    To enforce this, we add an additional premise: that in going from $\Gamma$ to the intermediate context $\Psi_m$ we must exhaust no multiplicities at index $i$ -- so that $m$ indeed captures the totality of the multiplicities that are being exhausted between $\Gamma$ and $\Psi_l$.
-    We capture this requirement by enforcing $\Delta_i \equiv \ell_{\o}$.
-  \end{remark}
-\end{nitheorem}
-
-\begin{figure}[h]
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \node (gamma-l) at (0,6)   {$\Gamma_l$};
-      \node (gamma-m)   at (0,3)   {$\Gamma$};
-      \node (gamma-r) at (0,0)   {$\Gamma_r$};
-      \node (xi-l)    at (3,5.5) {$\Xi_l$};
-      \node (theta)   at (3,3.3) {$\Theta$};
-      \node (delta-m) at (3,2.7) {};
-      \node (xi-r)    at (3,0.5) {$\Xi_r$};
-      \node (psi-l)   at (6,5)   {$\Psi_l$};
-      \node (psi-m)   at (6,3)   {$\Psi$};
-      \node (psi-r)   at (6,1)   {$\Psi_r$};
-
-      \draw[-]  (gamma-m) -- (delta-m.center);
-      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \ell_{\o}$}(psi-m);
-
-      \draw[->,densely dotted] (gamma-m) -- (theta);
-      \draw[->,densely dotted] (theta) -- (psi-m);
-
-      \draw[->] (gamma-l) -- node[left] {$\ni_i m$} (gamma-m);
-      \draw[->] (gamma-r) -- node[left] {$\ni_j m$} (gamma-m);
-      \draw[->] (psi-l) -- node[right] {$\ni_i n$} (psi-m);
-      \draw[->] (psi-r) -- node[right] {$\ni_j n$} (psi-m);
-
-      \draw[->] (gamma-l) -- node[above] {$\vdash P$} (xi-l);
-      \draw[->] (xi-l) -- node[above] {$\vdash Q$} (psi-l);
-      \draw[->,densely dotted] (gamma-r) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-r);
-      \draw[->,densely dotted] (xi-r) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-r);
-      \draw[->,densely dotted] (xi-l) -- node[left] {$\ni_i l$} (theta);
-      \draw[->,densely dotted] (xi-r) -- node[left] {$\ni_j l$} (theta);
-    \end{tikzpicture}
-    \caption{
-      Diagrammatic representation of the inductive substitution lemma.
-      Continuous lines represent known facts, dotted lines proof obligations.
-      From $\Delta_i = \ell_{\o}$, we can deduce that there exists some $\delta$ such that $\opsquared{m}{\delta}{n}$.
-      From that, we can find first $\ni_i l$ and then $\ni_j l$, splitting the diagram across its vertical axis, making it possible to inductively derive both $\vdash \subst{P}{j}{i}$ and $\vdash \subst{P}{j}{i}$.
-    }
-    \label{fig:subst}
-  \end{subfigure}
-  \hspace{\fill} % DO NOT LEAVE EMPTY LINE
-  \begin{subfigure}{.4\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \node (gamma-l) at (0,4)   {};
-      \node (gamma-m) at (0,2)   {};
-      \node (gamma-r) at (0,0)   {};
-      \node (xi-l)    at (2,5)   {};
-      \node (xi-m)    at (2,3.5) {};
-      \node (xi-r)    at (2,2)   {};
-      \node (psi-m)   at (4,3.5) {};
-
-      \draw[->] (gamma-l) -- (xi-r);
-      \draw[->] (gamma-m) -- (xi-r);
-      \draw[->] (gamma-r) -- (xi-r);
-      \draw[->] (gamma-l) -- (gamma-m);
-      \draw[->] (gamma-r) -- (gamma-m);
-      \draw[->] (xi-l)    -- (xi-m);
-      \draw[->] (xi-r)    -- (xi-m);
-      \draw[->] (xi-l)    -- (psi-m);
-      \draw[->] (xi-m)    -- (psi-m);
-      \draw[->] (xi-r)    -- (psi-m);
-    \end{tikzpicture}
-    \caption{
-      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      These are not strictly necessary, but they ease the composition of the inductive hypotheses.
-      Without them, framing lemmas would have to be used extensively.
-    }
-    \label{fig:subst-compositionality}
-  \end{subfigure}
-\end{figure}
-
-\begin{proof}
-  \todo{BIG TODO}
-\end{proof}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

### 2020-04-28 paper/paper.tex
Change title
Another title!
Itemize subject congruence proof

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Mechanising a linear \picalc}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
 \author{Uma Zalakain}{University of Glasgow, Scotland}
@@
   \end{mathpar}
-  \caption{Left-over typing for a resource-aware typing system.}
+  \caption{Leftover typing for a resource-aware typing system.}
   \label{fig:types}
   \end{figure}
+  \todo{align typing rules to make them more readable}
 \end{nidefinition}
@@
   The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
-  For those, we proceed by induction on \textsc{StructCong}.
+  For those, we proceed by induction on \textsc{StructCong}:
+  \begin{itemize}
+    \item
       The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
+    \item
       The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
-  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
+      The usage context on which $Q$ and $P$ coincide in $\comp{Q}{P}$ will no longer be the same.
+    \item
       The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
+    \item
       In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
@@
       For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
+    \item
       In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
@@
       We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
+    \item
       Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
       In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
+  \end{itemize}
 \end{proof}

### 2020-04-28 paper/paper.tex
some more work on intro + bibs

--- 2020-04-28 paper/paper.tex
@@
 As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-The \picalc{} \cite{MilnerPW92,Milner99,Sangio01} is a computational model for communication and concurrency.
+The \picalc{} \cite{MilnerPW92,Milner99,Sangio01} is the most successful computational model for communication and concurrency.
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
@@
 The goal is to bring all distributed systems features (in the future) into the formalised world.}
-Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted, too.
+Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted.
+
+The discovery of linear logic by J.-Y.
+Girard \cite{Girard87} opened new avenues in computing science and introduced linear types and type systems for (functional) programming languages \cite{Wadler90,Bernardy2018} \todo{fill the gaps with more PLs}.
+
 The \emph{linear} \picalc{} \cite{KPT96,Sangio01} is a typed version of the \picalc{} where the linear type system restricts the usage of channel resources to exactly ones.
+Session types \cite{H93,THK94,HVK98}...
+Linearity is key in session types and the linear \picalc{} is the theoretica and practical framework and session types can be expressed in terms of linear \picalc{} \cite{DardhaGS12,DardhaGS17}.
 

### 2020-04-28 paper/paper.tex
Supply a running title without the newline
Rephrase the need for a cancelative monoid

@@
 \title{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
 \author{Uma Zalakain}{University of Glasgow, Scotland}
@@
 \item [Injectivity]
-  To prove framing (\S \ref{framing}) we need to be able to \emph{subtract} multiplicities.
-  For that end, we rely on $x - z \equiv y$ iff $\op{x}{y}{z}$.
-  As a consequence, we must constraint $y$ to be uniquely determined as well.
+  We need to be able to \emph{subtract} multiplicities.
+  To do so, we rely on $\op{x}{y}{z}$ uniquely determining $y$ for any $x$ and $z$.
 

### 2020-04-28 paper/paper.tex
Finish with substitution

@@
 \title{$\pi$ with leftovers: \\ a mechanisation in Agda}
-\titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
+\titlerunning{$\pi$ with leftovers: all mechanisation in Agda}
 \author{Uma Zalakain}{University of Glasgow, Scotland}
@@
 
-When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the innermost variable in the body of the abstraction gets replaced with the argument of the application.
-In the \picalc{} however substitution takes another form: every reference to the innermost variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
+When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the most recent variable in the body of the abstraction gets replaced with the argument of the application.
+In the \picalc{} however substitution takes another form: every reference to the most recent variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
 In a resource-aware \picalc{}, that means that instead of exhausting all $m$ multiplicities of the most recent variable, we leave those unaltered and exhaust them from the variable that is being received.
@@
   \label{thm:substitution}
-  Let $P$ be a process where the innermost variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
-  Let $j$ be the index of the data that is being received.
-  This data has at least $m$ multiplicities, such that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the most recent variable by $j$.
-  As a result, we will exhaust no multiplicities on the innermost variable and exhaust them in $j$ instead, therefore $\types{\gamma , t}{\Gamma , m}{\subst{P}{j}{\constr{zero}}}{\Xi , m}$.
+  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
+  Let $j$ be the index of the data with $m$ multiplicities that is being received after $P$ is done, so that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute variable $\constr{zero}$ in $P$ with $\suc j$.
+  As a result, the multiplicities at $\constr{zero}$ will be preserved and those at $j$ will be consumed instead, resulting in $\types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}$.
 \end{nitheorem}
-
 \begin{proof}
-  By induction on the typing derivation on $P$.
-  To be able to proceed under scope restriction, we need to generalize the theorem so that the variable that is being substituted need not be $\constr{zero}$, but any variable $i$.
-  We introduce such generalization in Theorem \ref{thm:subst-generalization}.
+  We need to proceed by induction on $P$, and to do so means going under scope restriction.
+  We use the generalised version in Theorem \ref{thm:subst-generalization} so that the variable that is being substituted is not $\constr{zero}$ but any variable $i$.
   \todo{explain arguments}
@@
 
-  We introduce an intermediate usage context $\Psi_m$.
-  This intermediate context is the result of exhausting some $\Delta$ from the input context $\Gamma$, such that $\opctx{\Gamma}{\Delta}{\Psi_m}$.
-  From $\Psi_m$, exhausting $m$ at index $i$ yields $\Psi_l$; exhausting it at index $j$ yields $\Psi_j$, such that $\contains{\gamma}{\Psi_m}{i}{t}{m}{\Psi_l}$ and $\contains{\gamma}{\Psi_m}{j}{t}{m}{\Psi_r}$.
-  The input derivation on $P$ takes context $\Gamma$ and yields context $\Psi_l$, such that $\types{\gamma}{\Gamma}{P}{\Psi_l}$.
-  The output derivation on $\subst{P}{j}{i}$ takes the same context $\Gamma$ but yields back $\Psi_r$, such that $\types{\gamma}{\Gamma}{\subst{P}{j}{i}}{\Psi_r}$.
-
-  \begin{remark}
-    \label{re:delta-zero}
-    It is of uttermost importance that the variable that is being substituted has \emph{all} of its multiplicities substituted.
-    To enforce this, we add an additional premise: that in going from $\Gamma$ to the intermediate context $\Psi_m$ we must exhaust no multiplicities at index $i$ -- so that $m$ indeed captures the totality of the multiplicities that are being exhausted between $\Gamma$ and $\Psi_l$.
-    We capture this requirement by enforcing $\Delta_i \equiv \ell_{\o}$.
-  \end{remark}
+  Let the variable that is being substituted be $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$ and the variable that it is being substituted with $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$.
+  As shown in Figure \ref{fig:subst}, the process that is being substituted will pivot from the input usage context $\Gamma_i$ to the input usage context $\Gamma_j$.
+  Let there be two \emph{adjustment} variables $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$ and $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$.
+  The process that is being substituted will pivot from the output usage context $\Psi_i$ to the output usage context $\Psi_j$.
+  As shown in Figure \ref{fig:subst-alternative}, without this \emph{adjustment} composing the substitution of subprocesses is not at all trivial.
+  Assume that $P$ uses all $m$ multiplicities at index $i$.
+  To model this, require $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$ where index $i$ is $\ell_{\o}$.
+  The process that is typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$ will be typed $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$ after substitution.
 \end{nitheorem}
@@
     \begin{tikzpicture}
-      \node (gamma-l) at (0,6)   {$\Gamma_l$};
+      \node (gamma-i) at (0,6)   {$\Gamma_i$};
       \node (gamma-m)   at (0,3)   {$\Gamma$};
-      \node (gamma-r) at (0,0)   {$\Gamma_r$};
-      \node (xi-l)    at (3,5.5) {$\Xi_l$};
+      \node (gamma-j) at (0,0)   {$\Gamma_j$};
+      \node (xi-i)    at (3,5.5) {$\Xi_i$};
       \node (theta)   at (3,3.3) {$\Theta$};
       \node (delta-m) at (3,2.7) {};
-      \node (xi-r)    at (3,0.5) {$\Xi_r$};
-      \node (psi-l)   at (6,5)   {$\Psi_l$};
+      \node (xi-j)    at (3,0.5) {$\Xi_j$};
+      \node (psi-i)   at (6,5)   {$\Psi_i$};
       \node (psi-m)   at (6,3)   {$\Psi$};
-      \node (psi-r)   at (6,1)   {$\Psi_r$};
+      \node (psi-j)   at (6,1)   {$\Psi_j$};
 
@@
 
-      \draw[->] (gamma-l) -- node[left] {$\ni_i m$} (gamma-m);
-      \draw[->] (gamma-r) -- node[left] {$\ni_j m$} (gamma-m);
-      \draw[->] (psi-l) -- node[right] {$\ni_i n$} (psi-m);
-      \draw[->] (psi-r) -- node[right] {$\ni_j n$} (psi-m);
-
-      \draw[->] (gamma-l) -- node[above] {$\vdash P$} (xi-l);
-      \draw[->] (xi-l) -- node[above] {$\vdash Q$} (psi-l);
-      \draw[->,densely dotted] (gamma-r) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-r);
-      \draw[->,densely dotted] (xi-r) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-r);
-      \draw[->,densely dotted] (xi-l) -- node[left] {$\ni_i l$} (theta);
-      \draw[->,densely dotted] (xi-r) -- node[left] {$\ni_j l$} (theta);
+      \draw[->] (gamma-i) -- node[left] {$\ni_i m$} (gamma-m);
+      \draw[->] (gamma-j) -- node[left] {$\ni_j m$} (gamma-m);
+      \draw[->] (psi-i) -- node[right] {$\ni_i n$} (psi-m);
+      \draw[->] (psi-j) -- node[right] {$\ni_j n$} (psi-m);
+
+      \draw[->] (gamma-i) -- node[above] {$\vdash P$} (xi-i);
+      \draw[->] (xi-i) -- node[above] {$\vdash Q$} (psi-i);
+      \draw[->,densely dotted] (gamma-j) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-j);
+      \draw[->,densely dotted] (xi-j) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-j);
+      \draw[->,densely dotted] (xi-i) -- node[left] {$\ni_i l$} (theta);
+      \draw[->,densely dotted] (xi-j) -- node[left] {$\ni_j l$} (theta);
     \end{tikzpicture}
@@
       Continuous lines represent known facts, dotted lines proof obligations.
-      From $\Delta_i = \ell_{\o}$, we can deduce that there exists some $\delta$ such that $\opsquared{m}{\delta}{n}$.
-      From that, we can find first $\ni_i l$ and then $\ni_j l$, splitting the diagram across its vertical axis, making it possible to inductively derive both $\vdash \subst{P}{j}{i}$ and $\vdash \subst{P}{j}{i}$.
     }
@@
     \begin{tikzpicture}
-      \node (gamma-l) at (0,4)   {};
+      \node (gamma-i) at (0,4)   {};
       \node (gamma-m) at (0,2)   {};
-      \node (gamma-r) at (0,0)   {};
-      \node (xi-l)    at (2,5)   {};
+      \node (gamma-j) at (0,0)   {};
+      \node (xi-i)    at (2,5)   {};
       \node (xi-m)    at (2,3.5) {};
-      \node (xi-r)    at (2,2)   {};
+      \node (xi-j)    at (2,2)   {};
       \node (psi-m)   at (4,3.5) {};
 
-      \draw[->] (gamma-l) -- (xi-r);
-      \draw[->] (gamma-m) -- (xi-r);
-      \draw[->] (gamma-r) -- (xi-r);
-      \draw[->] (gamma-l) -- (gamma-m);
-      \draw[->] (gamma-r) -- (gamma-m);
-      \draw[->] (xi-l)    -- (xi-m);
-      \draw[->] (xi-r)    -- (xi-m);
-      \draw[->] (xi-l)    -- (psi-m);
+      \draw[->] (gamma-i) -- (xi-j);
+      \draw[->] (gamma-m) -- (xi-j);
+      \draw[->,densely dotted] (gamma-j) -- (xi-j);
+      \draw[->] (gamma-i) -- (gamma-m);
+      \draw[->] (gamma-j) -- (gamma-m);
+      \draw[->] (xi-i)    -- (xi-m);
+      \draw[->] (xi-j)    -- (xi-m);
+      \draw[->] (xi-i)    -- (psi-m);
       \draw[->] (xi-m)    -- (psi-m);
-      \draw[->] (xi-r)    -- (psi-m);
+      \draw[->,densely dotted] (xi-j)    -- (psi-m);
     \end{tikzpicture}
@@
       Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      These are not strictly necessary, but they ease the composition of the inductive hypotheses.
-      Without them, framing lemmas would have to be used extensively.
+      Composition is non-trivial.
     }
-    \label{fig:subst-compositionality}
+    \label{fig:subst-alternative}
   \end{subfigure}
+  \caption{Substitution with and without the \emph{adjustment} arrows $\ni_i n$ and $\ni_j n$}
 \end{figure}
@@
 \begin{proof}
-  \todo{BIG TODO}
+  By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  \begin{itemize}
+    \item
+      From constructor $\constr{end}$ we derive $\Gamma_i \equiv \Psi_i$.
+      From $\Delta_i \equiv \ell_{\o}$ follows that $m \equiv n$.
+      Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.
+
+    \item
+      For constructor $\constr{chan}$ we proceed inductively, applying the constructor $\suc$ to all arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$.
+
+    \item
+      For constructor $\constr{recv}$ we split $\Delta$ so that we can proceed inductively on the continuation.
+      We first observe that given that there exists an arrow from $\Gamma_i$ to $\Psi_i$ and $\Delta$ at index $i$ is $\ell_{\o}$, for any $m$ and $n$, there is some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+ l     \begin{itemize}
+        \item
+          If the input occurs on the variable that is being substituted, we derive $\opsquared{m}{\ell_i}{l}$ for some $l$ and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ (used for the inductive call).
+          Similarly, we construct the arrows $\containsusage{\Gamma_j}{j}{\ell_i}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ (used for the inductive call) for some $\Xi_j$.
+        \item
+          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $containsusage{\Gamma}{x}{\ell_i}{\Theta}$ for some $\Theta$.
+          We can then find $\Xi_j$ and construct the two arrows $\containsusage{\Gamma_j}{x}{\ell_i}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (used for the inductive call).
+          From the fact that there exists a composition of arrows from $\Xi_i$ to $\Psi$, we can conclude that $\Theta$ is splitting $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+          As $\ell_{\o}$ is a minimal element, if $\Delta$ is $\ell_{\o}$ at index $i$, then it must be $\ell_{\o}$ at index $i$ for $\Delta_1$ and $\Delta_2$ as well.
+      \end{itemize}
+
+    \item
+      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the channel doing the output and the transmitted data.
+
+    \item
+      For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+      Given that $\Delta$ is $\ell_{\o}$ at index $i$, we know that $\Delta_1$ and $\Delta_2$ are so too.
+      We observe that $\opsquared{m}{\delta}{\omega}$ where $\omega$ is what subprocess $P$ consumes at index $i$.
+      We use this to construct an arrow $\containsusage{\Xi_j}{j}{\delta}{\Theta}$, for some $\Xi_j$.
+      We use these constructions to make two inductive calls (on the derivation of $P$ and $Q$) and compose their result.
+  \end{itemize}  
 \end{proof}

### 2020-04-28 paper/paper.tex
Minor terminology correction
Note to mention our Curry-style encoding (first semantics, then typing)

@@
 
+\todo{Mention that our semantics is prior to our typing (Curry-style)}
+
 Another property of particular interest to concurrent systems is the avoidance of race conditions.
@@
 
-We make use \emph{leftover typing}, a technique consisting of adding an \emph{output context of usages} to the usual \emph{input context of usages}.
+We make use \emph{leftover typing}, a technique consisting of adding an \emph{output usage context} to the usual \emph{input usage context}.
 This output context contains the leftovers: the multiplicities unused by the processes.

### 2020-04-29 paper/paper.tex
a first draft of the introduction

@@
 \newcommand{\eqeq}{\; \type{\equiv} \;}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Introduction}
+% The way I want to write the intro:
+% -- describe what is the picalc and why it is IMPORTATNT
+% -- describe what are linear types and why programming with them is IMPORTATNT
+% -- join together by saying that is why we focus on linear pi calculus but we do more than just that: we have linear, gradual and shared types
+% -- put emphasis on mechanisation in Agda
+
 We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
 As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-The \picalc{} \cite{MilnerPW92,Milner99,Sangio01} is the most successful computational model for communication and concurrency.
+
+The \picalc{} introduced by Milner et al. in the early '90s \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
-\todo{OD: The key point here is that the pi-calculus, especically the linear one, is a sufficcient core calculus where all research on concurrent and ditrubuted systems can be achieved.
-That is why we are inspecting it.
-The goal is to bring all distributed systems features (in the future) into the formalised world.}
 Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted.
-
-The discovery of linear logic by J.-Y.
-Girard \cite{Girard87} opened new avenues in computing science and introduced linear types and type systems for (functional) programming languages \cite{Wadler90,Bernardy2018} \todo{fill the gaps with more PLs}.
-
-The \emph{linear} \picalc{} \cite{KPT96,Sangio01} is a typed version of the \picalc{} where the linear type system restricts the usage of channel resources to exactly ones.
-Session types \cite{H93,THK94,HVK98}...
-Linearity is key in session types and the linear \picalc{} is the theoretica and practical framework and session types can be expressed in terms of linear \picalc{} \cite{DardhaGS12,DardhaGS17}.
-
+The \picalc{} is at the basis of programming language implementations for concurrency, such as Pict \cite{Pierce} and more recently Go \cite{Golang}.
+Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types, ranging from basic types, linear types, types for liveness properties, such as deadlock freedom or termination, and most notably, session types \cite{K07}, making the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
+
+In a parallel track in the early '90s, the discovery of linear logic by J.-Y.
+Girard \cite{Girard87} opened new avenues in computing science and introduced linear types and type systems for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}, allowing for resource-aware programming and efficient implementation.
+Linearity later led to uniqueness of typing in Clean \cite{BarendsenS96} or ownership typing in Rust \cite{MatsakisK14}.
+
+The \picalc{} was not an exception to the influence of linearity.
+Kobayashi et al. \cite{KPT96} defined the \emph{linear} \picalc{}, which is a typed version of the \picalc{} where the linear type system restricts the usage of channel resources to exactly ones.
+%
+With the rise of session types \cite{H93,THK94,HVK98}, linearity was reproposed in the \picalc{}, which is
+the most successful framework where session types have been used.
+Session types are a type formalism used in communication-structured modelling and programming and linearity is a key feature in order to guarantee type safety.
+Recent work brought back to light the linear \picalc{}, because session types can be expressed in terms of linear types \cite{DardhaGS12,DardhaGS17}.
+The encoding of session types into linear types not only has theoretical benefits in terms expressivity of session types and reusability of theoretical results from the linear \picalc{}, but also at the implementation level.
+The encoding has been used as a technique to implement session types languages such as OCaml \cite{Padovani17} or Scala \cite{ScalasY16,ScalasDHY17}.
+As such, linear \picalc{} is the underlying theoretical and practical framework of session types.
+
+Considering the relevance of the \picalc{} as a core calculus to model concurrent and distributed system and linearity for resource-aware programming, in this paper we take it a step further and investigate a \picalc{} featuring linear, gradual and shared types under the same unique framework and mechanise it all in Agda.
+We can use this framework as a core for type safe distributed programming with a range of types from linear to shared.
+
+\subsection{Contribution}
+Our contributions can be summarised as follows later explained in details, together with the corresponding sections where they can be found.
+\begin{itemize}
+  \item Machine verified formalisation of the syntax and semantics of linear \picalc{}
+
+  \item Typing with leftovers applied to the \picalc{}, where a range of type systems from linear to gradual and shared can be obtained
+
+  \item Abstraction over multiplicities
+
+  \item Full proofs of subject reduction and auxiliary lemmas
+
+  \item Full formalisation available in Agda
+\end{itemize}
+
+\paragraph*{Syntax}
 We type the syntax for the \picalc{} in \S \ref{syntax}.
 We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every unbound variable is accounted for in the type of the process that uses it.
-\todo{OD: don't put references to section just yet. This first part should describe the what and why of the problem.
-In the contributions section below, we can use references to sections to show where those contributions can be found.}
 In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
@@
 
+\paragraph*{Semantics}
 In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
@@
 
+\paragraph*{Typing with leftovers}
 Linear type systems need of context splits.
@@
 
-\todo{subject reduction}
-
-\subsection{Contribution}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{description}
-  \item [Machine verified formalisation of the linear pi calculus]
-
-  \item [Typing with leftovers applied to the pi calculus]
-
-  \item [Abstraction over multiplicities]
-
-  \item [Full formalisation available in Agda]
-\end{description}
+\paragraph*{Subject reduction}
+
+\paragraph*{On mechanisation in Agda}
 

### 2020-04-29 paper/paper.tex
Use the pi-calculus for the example of de Bruijn index usage

@@
 We type the syntax for the \picalc{} in \S \ref{syntax}.
-We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every unbound variable is accounted for in the type of the process that uses it.
+We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
 In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
@@
 
-Data type definitions have double lines and index-free synonyms as rule names, for ease of reference.
-Otherwise, we use the constructor name as the name of a typing rule.
-We omit universe levels for brevity.
+Data type definitions have double lines and index-free synonyms as rule names (for ease of reference).
+We otherwise use the constructor name as the name of a typing rule.
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
 Implicit arguments are mentioned by type definitions but omitted in constructors.
 $\type{TYPES}$ are in blue and uppercased (with indices as subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red, and variables and functions in black.
-Some constructors can be shared across types and disambiguated by context.
-\todo{OD: the next part to be filled should be the notation, so that when I go through the paper, I can use this as a manual.}
+Some constructor names are overloaded --- they and disambiguated by context.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Using names as variable references is a popular option amongst humans.
-However, names are cumbersome to mechanize: inserting a new variable into a context means proving that the name of such variable is different from all other variable names in context.
+However, names are cumbersome to mechanize: inserting a new variable into a context means proving that the name of such variable is different to all other variable names in context.
 Moreover, names are of no significance to machines.
 
-Machines need constructs they can algorithmically act upon.
-The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago.
-Binders no longer introduce any names.
-\todo{OD: IMPORTATNT!
-There is too much on lambda-calculus; I suggest you give examples and illustrate your points on the pi-calculus only from now on.
-It can get too broad and lack focus otherwise...
-It is fine in the Introduction as you are giving an overview.}
+The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and binders no longer introduce names.
 \begin{example}
-The expression $\lambda g . (\lambda f . f g) g$ in the \lambdacalc{} would translate to $\lambda (\lambda 0 1) 0$.
+The \picalc{} term $(\textbf{new} \; x) (x ( y ) .
+y \langle \texttt{"Hello world!"} \rangle .
+\mathbf{0} \parallel (\textbf{new} \; y) (x \langle y \rangle .
+y ( z ) .\mathbf{0}))$ would translate to $(\textbf{new}) (0 () .
+0 \langle \texttt{"Hello world"} \rangle .
+\mathbf{0} \parallel (\textbf{new}) (1 \langle 0 \rangle .
+0 () .
+\mathbf{0}))$.
 \end{example}

### 2020-04-29 paper/paper.tex
Pass on syntax

@@
 That is, terms at different \emph{depths} use different indices to refer to the same binding.
-Humans often find this confusing.
-There is however no reason not to annotate these indices with the original names bestowed by humans.
-Machines can then manipulate references mechanically while still using names to present terms to humans.
-
-A variable occurring under $n$ abstractions has $n$ things to refer to.
-References outside of that range have no associated meaning.
+Humans may find this confusing.
+There is however no reason not to annotate these indices with the original names bestowed upon them by humans.
+Machines can then manipulate references mechanically whilst using names to present terms to humans.
+
+A variable occurring under $n$ abstractions has $n$ possible variables to refer to.
+References outside of that range are meaningless.
 It is useful to rule out these nonsensical terms syntactically.
@@
 
-Using type-level de Bruijn indices makes our syntax well-scoped by construction, and the semantics of our language can now be defined on the totality of the syntax.
-User-friendliness can still be recovered through a function that converts processes with names into (possibly) processes with indices.
-This function keeps track of what index is associated with what name, and traverses the process recursively, taking note of new binders and substituting variable references.
+Using type-level de Bruijn indices makes our syntax well-scoped by construction --- the semantics of our language can now be defined on the totality of the syntax.
+User-friendliness can still be recovered through a function that converts processes with named variables into (possibly) processes with indexed variables.
+This function recursively traverses processes, keeping track of what name is associated with each index.
 If the process is ill-scoped, the function returns nothing.
-To present things back to the user, we print indices as their named annotations.
+To present things to the user, we substitute indices by their related names.
 

### 2020-04-29 paper/paper.tex
Give structural congruence another pass

@@
 
+\todo{rename Scoped to Process in code}
+\todo{make sure all citations are there}
+\todo{change colors, color functions}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 (e.g. semantically, parallel composition is defined modulo associativity and commutativity).
-There are several ways around this, a structural congruence relation being one of the historical ones.
-(Others include labeled transition systems and higher inductive types.)
+There are several ways around this, a structural congruence relation being one of the historical ones \cite{}.
+(Others include labeled transition systems \cite{} and higher inductive types \cite{}.)
 
@@
 Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.
-The type $\Unused_i \; Q$ is an inductive proof on $Q$ that witnesses that the index $i$ does not appear nor in the inputs nor in the outputs of the process.
-The function $lower \; i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
-Finally, $swap \; i \; P$ traverses $P$ (scoped under $\suc \suc n$) and swaps variable references $i$ and $\suc i$.
+The type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear nor in the inputs nor in the outputs of $Q$.
+The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+Finally, $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In all three, $i$ is incremented every time we go under a binder.
 
@@
 \end{mathpar}
-\caption{Structural rewriting rules. Premises $P$, $Q$ and $R$ are of type $\Process_n$ where $n$ can be inferred.}
+\caption{Base cases of structural congruence.}
 \label{fig:struct-cong-base}
@@
   \label{lm:lower-lift}
-  The function $lower \; i \; P \; uP$ has an inverse $lift \; i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift \; i \; (lower \; i \; P \; uP) \equiv P$.
+  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
 \end{nilemma}
@@
   \label{lm:swap-swap}
-  The function $swap \; i \; P$ is its own inverse: $swap \; i \; (swap \; i \; P) \equiv P$.
+  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
 \end{nilemma}
@@
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen at any point in a process' recursive definition, and they are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{fig:struct-cong}.
-In \S \ref{subject-congruence} we will prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then $Q$ is well-typed.
-Specifically, in the case of transitivity, we must prove that if $P$ is structurally congruent with $Q$ and $Q$ with $R$, and $P$ is well-typed, then so is $R$.
-To do so, we will have to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
-To show that the doubly recursive call terminates we index the equivalence relation $=$ by the type $\Rec$, which models the structure of the recursion.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{fig:struct-cong}.
+In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
+In the transitivity case, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
+To do so, we need to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
+To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
 
@@
 \end{mathpar}
-\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.
-  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred.}
+\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
 \label{fig:struct-cong}

### 2020-04-29 paper/paper.tex
Give the operational semantics another pass

@@
   {uQ : \Unused_0 \, Q}
-  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}
-
-  \inferrule
-  { }
-  {\constr{scope-comm} : \new \new P \eqeq \new \new swap \; 0 \; P}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower_0 \; \; Q \; uQ}}
+
+  \inferrule
+  { }
+  {\constr{scope-comm} : \new \new P \eqeq \new \new swap_0 \; P}
 \end{mathpar}
@@
 Figure \ref{fig:reduction} models the operational semantics of the \picalc{}.
-Processes put in parallel reduce when they communicate over a common variable.
-The receiving process substitutes references to its most immediate variable with references to the variable sent by the process doing the output.
+Processes put in parallel reduce when they communicate over a common channel.
+The receiving process substitutes references to its most immediate variable with references to the variable that is being sent by the process doing the output.
 The process is then \emph{lowered} --- all variable references are decreased by one.
 Reduction is closed under structural congruence and goes under parallel composition and scope restriction
---- but notably not under input nor output, otherwise the sequencing of actions would not be preserved.
+--- notably, not under input nor output, or otherwise the sequencing of actions would not be preserved.
 In \S \ref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
-However, this reduction operation is effectful: it consumes the variable over which communication happens.
-If this variable is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
-To keep track of this information, we lift the variable index over which communication occurs to the type level.
-Every time we come out of a binder we decrement this variable.
-To do so, we make use of a $dec$ function that saturates at $\constr{internal}$ --- as opposed to an $inc$ function that would not.
+However, this reduction operation is effectful: it consumes the channel over which communication occurs.
+If this channel is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
+To keep track of this information, we lift the index of the channel to the type level.
+We use a $dec$ function to decrement the index of the channel every time we come out of a binder.
+This function saturates at $\constr{internal}$ --- meaning communication happens internally to the process.
 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower \; 0 \; (\subst{P}{j}{0})}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower_0 \; (\subst{P}{j}{0})}{Q}}
 
@@
 \subsection{Framing}
+\label{framing}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Weakening}
+\label{weakening}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Strengthening}
+\label{strengthening}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused \; i \; P$.
+  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused_i \; P$.
   Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
@@
 \subsection{Swapping}
+\label{swapping}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Substitution}
+\label{substitution}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Subject Reduction}
+\label{subject-reduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \item
-    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in Figure \ref{fig:subject-reduction:a}.
+    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower_0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in Figure \ref{fig:subject-reduction:a}.
     We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (Theorem \ref{thm:substitution}) and strengthening (Theorem \ref{thm:strengthening}).
@@
         \begin{align*}
-          \types{\gamma , t}{\Gamma , m}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
+          \types{\gamma , t}{\Gamma , m}{lower_0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
         \end{align*}

### 2020-04-29 paper/paper.tex
Rearrange things, give multiplicities and typing contexts another pass

@@
 \todo{change colors, color functions}
+\todo{do not distinguish fields from functions, it is not worth it}
 
@@
 
-A \emph{resource-aware} typing system for the \emph{\picalc{}} is distinctive in several ways.
-
-\begin{description}
-  \item [As a resource-aware typing system,] the language is constrained not just by types, but also by their usage annotations.
-    We must be able to add up and split usage contexts, to do so we introduce a generalized algebra in \S \ref{multiplicities}.
-    We provide an independent inductive definition of typing contexts and usage contexts in \S \ref{contexts}.
-    \todo{why independent}
-
-  \item [As a typing system for the \picalc{},] the terms (processes) in the typing judgment do not come accompanied by a type -- processes have no type.
-    It is the communication channels that are annotated with types and usages.
-    Typing judgments assert that the process correctly uses the channels in the context under which it is typed.
-\end{description}
-
-In this paper, we combine typing and annotation judgments into a single judgment.
-Although the typing rules are for now completely syntax-directed, adding support for branching would introduce choice in the typing derivations.
-If we were to keep typing judgments and annotation judgments separate, we would need a coherence relation making sure that the same choice is taken by both.
-We avoid doing so by combining them into a single judgment.
-
-We make use \emph{leftover typing}, a technique consisting of adding an \emph{output usage context} to the usual \emph{input usage context}.
-This output context contains the leftovers: the multiplicities unused by the processes.
-Crucially, this output context can then be used as the input context of the continuation process.
-The typing rules are introduced in \S \ref{leftover-typing}.
+Typing systems for the \picalc{} ensure that processes make a correct use of their channels.
+Typing judgments are of the form $\Gamma \vdash P$ (for some typing context $\Gamma$ and some process $P$) --- types only appear in the contexts of typing judgments.
+A resource-aware typing system for the \picalc{} must take resource usages into account, and needs an algebra on these --- a way of \emph{splitting} resources on parallel composition, of \emph{ticking off} resources on input and output.
+We characterise such an algebra in \S \ref{multiplicities}.
+We define independent typing contexts and usage contexts in \S \ref{context}.
+We provide a typing judgment for a resource-aware \picalc{} in \S \ref{leftover-typing}.
 
@@
 
-In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 or 1.
-In this paper we generalize over linearity and provide an algebra on multiplicities that is capable of encoding shared and gradual types as well.
-The algebra has a single binary operator $(\cdot)$ that crushes multiplicities together.
-It must satisfy the following constraints:
-\begin{description}
-
-\item [Partiality]
-  Not all multiplicities are necessarily compatible: in the linear \picalc{} $1$ cannot be split into $1$ and $1$.
-\item [Functionality]
-  Given two multiplicities $y$ and $z$, it must be computable whether a third multiplicity $x$ exists such that $\op{x}{y}{z}$.
-  If such an $x$ exists, it must be unique.
-
-\item [Injectivity]
-  We need to be able to \emph{subtract} multiplicities.
-  To do so, we rely on $\op{x}{y}{z}$ uniquely determining $y$ for any $x$ and $z$.
-
-\item [Associativity]
-  As a direct consequence of the structural congruence rule $\comp{P}{\comp{Q}{R}} \equiv \comp{\comp{P}{Q}}{R}$.
-
-\item [Commutativity]
-  As a direct consequence of the structural congruence rule $\comp{P}{Q} \equiv \comp{Q}{P}$.
-
-\item [Distinguished elements]
-  There must be a neutral element $0$ absorbed on both sides.
-  The neutral element $0$ must also be the minimal element.
-  There must be a special element $1$ to record usage.
-\end{description}
+In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
+In this paper we generalize the type of multiplicities by defining an algebra any such type must comply with.
+Shared, gradual and linear usage types all satisfy these algebraic laws.
 
 \begin{nidefinition}
-  Together, the aforementioned constraints give raise to a \emph{partial commutative monoid}.
-  We deal with partiality by modeling the monoidal operation as a terniary relation.
-  An algebra on the carrier set $C$ forms a $\Quantifier_C$ iff it satisfies the fields in Figure \ref{fig:multiplicities}.
+  The algebra on multiplicities must be \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}, and needs a \emph{minimal neutral element}.
+  It also requires a special element that is to be used to count inputs and outputs.
+  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in Figure \ref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
 
@@
 \begin{note}
-  We will often work with input and output multiplicities pairs, and thus we use the notation $\Carrier^2$ to stand for the type $\Carrier \times \Carrier$, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all of its properties are trivially lifted.
+  We will often work with input and output multiplicities as pairs.
+  We use the notation $\Carrier^2$ to stand for the type $\Carrier \times \Carrier$, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
 \end{note}
+
+Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in Figure \ref{fig:multiplicities}.
+For examples of how these multiplicity types are used in typing derivations, refer to \S \ref{example-derivations}.
+
+\begin{description}
+  \item [Shared]
+    The carrier $C$ is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
+    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
+    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
+    All other laws are trivially satisfied.
+  \item [Gradual]
+    The carrier $C$ is implemented as the type of natural numbers $\N$.
+    The element $\zero$ corresponds to 0, and $\one$ to 1.
+    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
+    All other rules follow trivially from the algebraic rules for the addition of naturals.
+  \item [Linear]
+    The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
+    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
+    All other rules follow trivially.
+\end{description}
+
+\subsection{Contexts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\label{contexts}
+
+In a resource-aware \picalc{}, usage annotations are transformed as a result of input and output; types howeverr are preserved.
+In our case, usage annotations do not depend on types, and keeping types and usages in separate contexts allows for a more modular approach in our proofs.
+
+\subsubsection{Typing Contexts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{nidefinition}
+  Figure \ref{fig:prectx} defines types as either unit types, base types or channel types.
+  The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
+  The base type $\base{n}$ uses a natural number $n$ to represent types.
+  This natural number can then be interpreted as a type in the host language.
+  Using natural numbers instead of type universes is convenient: less universe polymorphism poluting specifications.
+
+  The channel type $\channel{t}{x}$ uses $t$ to stand for the type of data that is being transmitted over the channel, and $x$ to stand for what usages of that data are transmitted.
+
+  We keep track of types in a length-indexed \emph{pre-context} $\PreCtx_n$.
+
+  \begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    { }
+    {\Type : \Set}
+    \; \textsc{Type}
+
+    \inferrule
+    { }
+    {\unit : \Type}
+
+    \inferrule
+    {n : \N}
+    {\base{n} : \Type}
+
+    \inferrule
+    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
+    {\channel{t}{x} : \Type}
+  \end{mathpar}
+
+  \begin{mathpar}
+    \datatype
+    {n : \N}
+    {\PreCtx_n : \Set}
+    \; \textsc{PreCtx}
+
+    \inferrule
+        { }
+        {[] : \PreCtx_0}
+
+        \inferrule
+            {\gamma : \PreCtx_n \\ t : \Type}
+            {\gamma \comma t : \PreCtx_{\suc n}}
+  \end{mathpar}
+  \caption{Types and length-indexed typing contexts.}
+  \label{fig:prectx}
+  \end{figure}
+\end{nidefinition}
+
+\subsubsection{Usage Contexts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \end{nidefinition}
-
-\subsubsection{Example Type Systems}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\todo{Add typing derivations for example processes}
-
-\begin{description}
-
-\item [Shared]
-
-The carrier $C$ is implemented as a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
-Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
-The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
-All other laws are trivially satisfied.
-
-\item [Linear]
-
-The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
-The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
-All other rules follow trivially.
-
-\item [Gradual]
-
-The carrier $C$ is implemented as the type of natural numbers $\N$.
-The element $\zero$ corresponds 0, and $\one$ to 1.
-The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
-All other rules follow trivially from the algebraic rules for the addition of naturals.
-\todo{OD: in addition to the above, could we have examples of typing a process in each of the above type systems? It can be the same process everywhere and we can show how the typing changes...}
-\end{description}
-
-\subsection{Contexts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{contexts}
-
-In a resource-aware \picalc{} type multiplicities change every time input and output occurs; however, the types themselves remain constant.
-If multiplicities need not to depend on types, keeping types and multiplicities in independent contexts allows for a more modular approach to the algebra on contexts of multiplicities.
-In this paper we choose that approach, which entails accepting that base non-channel types have dedicated input and output multiplicities too -- we will take care of impeding that any actual input or output can occur through them in the typing rules.
-
-\begin{nidefinition}
-  Figure \ref{fig:prectx} defines types as either unit types, base types or channel types.
-  Unit types $\unit$ provide a proof of inhabitance for types, needed for the $\new \PO_{\suc n} \eqeq \PO_n$ rule in subject congruence.
-
-  Base types $\base{n}$ use a natural number $n$ that can be reified into a type by the host language.
-  Using natural numbers instead of type universes is convenient: there is no need to litter type types with universe polymorphism.
-
-  Channel types $\channel{t}{x}$ use $t$ as the type of data sent over the channel and $x$ as the capability of the data sent over the channel.
-  When data is received on a channel of type $\channel{t}{x}$, the type $t$ will be appended to the context of types $\PreCtx_n$ and the capability $x$ to the context of capabilities $\Ctx_{idxs}$.
-
-  Types form length-indexed \emph{pre-contexts} $\PreCtx_n$.
-
-  \begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    { }
-    {\Type : \Set}
-    \; \textsc{Type}
-
-    \inferrule
-    { }
-    {\unit : \Type}
-
-    \inferrule
-    {n : \N}
-    {\base{n} : \Type}
-
-    \inferrule
-    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
-    {\channel{t}{x} : \Type}
-  \end{mathpar}
-
-  \begin{mathpar}
-    \datatype
-    {n : \N}
-    {\PreCtx_n : \Set}
-    \; \textsc{PreCtx}
-
-    \inferrule
-        { }
-        {[] : \PreCtx_0}
-
-        \inferrule
-            {\gamma : \PreCtx_n \\ t : \Type}
-            {\gamma \comma t : \PreCtx_{\suc n}}
-  \end{mathpar}
-  \caption{Types and length-indexed typing contexts.}
-  \label{fig:prectx}
-  \end{figure}
-\end{nidefinition}
-
 \begin{nidefinition}
@@
 \label{leftover-typing}
+
+%%%% REARRANGE
+In this paper, we combine typing and annotation judgments into a single judgment.
+Although the typing rules are for now completely syntax-directed, adding support for branching would introduce choice in the typing derivations.
+If we were to keep typing judgments and annotation judgments separate, we would need a coherence relation making sure that the same choice is taken by both.
+We avoid doing so by combining them into a single judgment.
+%%%%
+
+We make use \emph{leftover typing}, a technique consisting of adding an \emph{output usage context} to the usual \emph{input usage context}.
+This output context contains the leftovers: the multiplicities unused by the processes.
+Crucially, this output context can then be used as the input context of the continuation process.
+The typing rules are introduced in \S \ref{leftover-typing}.
 
@@
     {\opsquared{x}{y}{z}}
-    {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{zero}{t}{y}{\Gamma \comma z}}
+    {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}
 
@@
 
+\subsubsection{Example derivations}
+
+\todo{OD: in addition to the above, could we have examples of typing a process in each of the above type systems? It can be the same process everywhere and we can show how the typing changes...}
+
 \begin{nilemma}
@@
   \label{thm:substitution}
-  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma , t}{\Gamma , m}{P}{\Psi , \ell_{\o}}$.
+  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \ell_{\o}}$.
   Let $j$ be the index of the data with $m$ multiplicities that is being received after $P$ is done, so that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
   Then, we can substitute variable $\constr{zero}$ in $P$ with $\suc j$.
-  As a result, the multiplicities at $\constr{zero}$ will be preserved and those at $j$ will be consumed instead, resulting in $\types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}$.
+  As a result, the multiplicities at $\constr{zero}$ will be preserved and those at $j$ will be consumed instead, resulting in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi \comma m}$.
 \end{nitheorem}
@@
           \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\ell_i &&}{\Psi} \\
-          \types{\gamma , t}{&\Psi , m}{P &&}{\Delta , \ell_{\o}} \\
+          \types{\gamma \comma t}{&\Psi \comma m}{P &&}{\Delta \comma \ell_{\o}} \\
           \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\ell_o &&}{\Theta} \\
@@
         \begin{alignat*}{2}
-          \types{\gamma , t}{& \Gamma , m}{P &&}{\Theta , \ell_{\o}} \\
+          \types{\gamma \comma t}{& \Gamma \comma m}{P &&}{\Theta \comma \ell_{\o}} \\
           \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
@@
         \begin{align*}
-          \types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}
+          \types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi \comma m}
         \end{align*}
@@
         \begin{align*}
-          \types{\gamma , t}{\Gamma , m}{lower_0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
+          \types{\gamma \comma t}{\Gamma \comma m}{lower_0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi \comma m}
         \end{align*}

### 2020-04-29 paper/paper.tex
Clean up usage contexts

@@
 \begin{example}
-The \picalc{} term $(\textbf{new} \; x) (x ( y ) .
+\hfill{}\\
+The \picalc{} term \hfill{} $(\textbf{new} \; x) (x ( y ) .
 y \langle \texttt{"Hello world!"} \rangle .
 \mathbf{0} \parallel (\textbf{new} \; y) (x \langle y \rangle .
-y ( z ) .\mathbf{0}))$ would translate to $(\textbf{new}) (0 () .
+y ( z ) .\mathbf{0}))$ \\ is represented as \hfill{} $(\textbf{new}) (0 () .
 0 \langle \texttt{"Hello world"} \rangle .
@@
 We characterise such an algebra in \S \ref{multiplicities}.
-We define independent typing contexts and usage contexts in \S \ref{context}.
+We define independent typing contexts and usage contexts in \S \ref{contexts}.
 We provide a typing judgment for a resource-aware \picalc{} in \S \ref{leftover-typing}.
@@
 
+\todo{we haven't introduced the forall and implicit syntax}
 \begin{nidefinition}
@@
 
+We choose to allow the use of multiple coexisting multiplicity algebras in our typing system.
+We do so by defining an indexed set of usage algebras.
+
 \begin{nidefinition}
-  An indexed set of partial commutative monoids is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it.
-  This allows for several algebras to coexist in the same type system.
-  Figure \ref{fig:indexed-multiplicities} captures these requirements: each index $i$ has a corresponding carrier type $\field{CARRIER}_i$ such that $\field{CARRIER}_i$ satisfies the algebra $\Quantifier_{\field{CARRIER}_i}$.
-  We require the index type to be inhabited -- necessary to prove type preservation on $\new \PO_{\suc n} \eqeq \PO_n$.
+  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by Figure \ref{fig:indexed-multiplicities}.
+  We require the index type to be inhabited -- proving subject congruence on $\new \PO_{\suc n} \eqeq \PO_n$ requires an inhabited carrier set.
 
@@
 \end{nidefinition}
+
 \begin{nidefinition}
-  To allow different multiplicity algebras to coexist we first define a length-indexed context of carrier indices $\Idxs_n$ (see \ref{fig:ctx}).
-  On top of it, we define a context of multiplicities $\Ctx_{idxs}$.
-  We lift the algebra on multiplicities to multiplicity contexts by constraining the algebra on multiplicity contexts to act on  contexts defined on the same underlying context of carrier indices.
-  We use two multiplicities over a common index to represent multiplicities for both the input and the output capabilities ($\Carrier_{idx}^2$).
-
+  A context of usages $\Ctx_{idxs}$ is indexed by a length-indexed context of indices $\Idxs_n$, as shown in Figure \ref{fig:ctx}.
   \begin{figure}[h]
@@
 \begin{note}
-  We lift the monoidal operation on pairs of multiplicities $\opsquared{x}{y}{z}$ to a monoidal operation on contexts with the same underlying vector of indices $\opctx{\Gamma}{\Delta}{\Xi}$, where all the properties of the partial monoid are trivially lifted to contexts.
+  We lift the monoidal operation on multiplicity pairs $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on contexts of usages with the same underlying context of indices.
+  All the properties of the partial monoid are trivially lifted.
 \end{note}

### 2020-04-29 paper/paper.tex
Give varref judgments a pass

@@
 \todo{do not distinguish fields from functions, it is not worth it}
+\todo{go over guillaume's paper again}
 
@@
 
-We choose to allow the use of multiple coexisting multiplicity algebras in our typing system.
+We choose to allow multiple multiplicity algebras to coexist in our typing system.
 We do so by defining an indexed set of usage algebras.
@@
 \subsection{Typing with Leftovers}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{leftover-typing}
-
-%%%% REARRANGE
-In this paper, we combine typing and annotation judgments into a single judgment.
-Although the typing rules are for now completely syntax-directed, adding support for branching would introduce choice in the typing derivations.
-If we were to keep typing judgments and annotation judgments separate, we would need a coherence relation making sure that the same choice is taken by both.
-We avoid doing so by combining them into a single judgment.
-%%%%
-
-We make use \emph{leftover typing}, a technique consisting of adding an \emph{output usage context} to the usual \emph{input usage context}.
-This output context contains the leftovers: the multiplicities unused by the processes.
-Crucially, this output context can then be used as the input context of the continuation process.
-The typing rules are introduced in \S \ref{leftover-typing}.
-
-We use leftover typing \cite{} to encode our typing rules.
-The insight of this approach is to use input and output contexts in the typing judgment.
-Our typing judgment on processes are of the form $\types{\gamma}{\Gamma}{P}{\Delta}$ (introduced in Figure \ref{fig:types}).
-This assertion is in fact both a typing judgment on $\gamma$ and a usage judgment on $\Gamma$ and $\Delta$.
-These two judgments could, in fact, happen separately: our typing rules are completely syntax directed.
-However, if we were to introduce branching, typing derivations would have to make choices, and the two separate judgments would need of a third relation to make sure that their choices agree.
-To avoid this, we merge both judgments into a single one.
-The typing context $\gamma$ needs to be explicit because the usage context $\Gamma$ and $\Delta$ do not in any way depend on it -- in fact, they are only related by their common length.
-
-While the typing contexts are preserved through the typing derivation, the usage contexts change.
-We use input and output usage contexts to capture this transformation.
-This has two main advantages: for one, the usage modifications effected by both input and output can be enclosed in a self-containing typing and usage judgment for variable references (also with an output context); but more crucially, we avoid requesting an explicit constructive proof of context splitting for parallel composition -- if $P$ is typed under $\Gamma_l$ and $Q$ under $\Gamma_r$, then to type $\comp{P}{Q}$ under $\Gamma_m$ a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ would be needed.
-The search for such a proof can be automated, but its existence is still required.
-Moreover, providing a typing derivation for $P$ gets the user no usage annotations to type $Q$.
-By using output usage contexts, we get the leftovers from the derivation on $P$ and use them to type $Q$.
-This is the recurring theme in leftover typing: usage contexts are threaded through the premises.
-
-The typing and usage judgments on both processes and variable references are of an inductive form.
-As a result, the proofs on typing derivations will perform a double induction: once on the derivation of the process, once on the derivation of the variable reference.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We make use \emph{leftover typing} \cite{}, an approach that, in addition to the usual typing context and input usage context, adds an extra output usage context to the typing rules.
+This output context contains the \emph{leftovers} of the process being typed --- its unused multiplicities.
+Crucially, the leftovers can then be used as an input to some othew typing derivation.
+Adopting this techinque has two main advantages:
+\begin{itemize}
+  \item
+    The input and output actions do no longer need to transform the usage context through functions, but can use a variable reference type with both an input and an output usage context to encapsulate these transformations.
+  \item
+    Parallel composition no longer needs an explicit context splitting proof --- a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    Such a proof could be obtained through automated search --- but it is better yet to avoid needing it.
+    With leftover typing, we use the leftovers of $P$ to type $Q$.
+\end{itemize}
+
+We introduce two judgments: one for variable references (Definition \ref{def:varref}), one for processes (Definition \ref{def:types}).
 
 \begin{nidefinition}
-  Akin to the typing and usage judgment on processes, in Figure \ref{fig:variable-references} we present a typing and usage judgment on variable references -- also indexed over a typing context and an input an output usage context.
-  However, in addition to this, variable references also record the index, the type, and the usage annotation of the variable reference.
-  The base case $\constr{zero}$ ensures that $t$ is the most immediate type in typing context, and splits the most immediate annotation in the input context $x$ into the usage annotation for the variable $y$ and the annotation in the output context $z$.
+  \label{def:varref}
+
+  Like processes, variable reference judgments (Figure \ref{fig:variable-references}) are indexed by a typing context and an input and output usage context.
+  In addition to these, variable reference judgments are also indexed by their position in the contexts, their type, and their usage annotation.
+
+  The base case $\constr{zero}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
+  This is the only place in the typing system where the algebra on multiplicities is used.
+
+  \begin{remark}
+    We use $\field{\cdot-compute}$ to aliviate the user from the proof burden $\opsquared{x}{y}{z}$.
+  \end{remark}
+
   The inductive case $\constr{suc}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.
@@
   \end{mathpar}
-  \caption{Variable references as proofs on the typing context and input and output multiplicity contexts.}
+  \caption{Variable reference judgments.}
   \label{fig:variable-references}
@@
 
-\begin{remark}
-  We use $\field{\cdot-unique}$ to relieve the user from the proof burden $\opsquared{x}{y}{z}$ in the base case $\constr{zero}$.
-  This is the only place in the typing system that uses the algebra on multiplicities.
-\end{remark}
-
 \begin{example}
@@
 
-\begin{nilemma}
-  The typing context $\gamma$ at index $i$ is type $t$.
-\end{nilemma}
-\begin{nilemma}
-  The input annotation in context $\Gamma$ at index $i$ is the result of aggregating (as per the algebra) $y$ and the output annotation in context $\Delta$ at index $i$.
-\end{nilemma}
-\begin{nilemma}
-  For every index $j$ other than $i$, the input annotation in context $\Gamma$ at index $i$ is the output annotation in $\Delta$ at index $i$.
-\end{nilemma}
-\begin{nilemma}
-  \todo{Add feedback and feedfront}
-\end{nilemma}
-\begin{corollary}
-  Given the witness $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$, $\Gamma \equiv \Delta$ iff $y \equiv \zero$.
-\end{corollary}
-\proof{All proofs are by structural induction on \textsc{VarRef}.}
-
 \begin{nidefinition}
+  \label{def:types}
+
+  Our judgment on processes is of the form $\types{\gamma}{\Gamma}{P}{\Delta}$ (Figure \ref{fig:types}).
+  This assertion combines a judgment on types ($\gamma$) and a judgment on usages ($\Gamma$ and $\Delta$).
+  They could be kept as two separate judgments, but doing so would risk incoherence: if there is any choice to be made, the judgment on types could choose to go left while the judgment on usages chooses to go right.
+  The constructions instroduced in this paper make the typing derivations completely syntax directed and thus this problem inexistent, but we nevertheless choose to keep both judgments as one for simplicity.
+
   The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed over a context of types $\gamma$, a context of input multiplicities $\Gamma$, the process $P$ that is being typed, and the context of output capabilities $\Delta$, and notated as $\types{\gamma}{\Gamma}{P}{\Delta}$.

### 2020-04-29 paper/paper.tex
Go through the typign rules again

@@
 However, names are cumbersome to mechanize: inserting a new variable into a context means proving that the name of such variable is different to all other variable names in context.
-Moreover, names are of no significance to machines.
-
-The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and binders no longer introduce names.
+Moreover, names are of no significance to machines whatsoever.
+
+The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and to have binders no longer introduce names.
 \begin{example}
@@
 y ( z ) .\mathbf{0}))$ \\ is represented as \hfill{} $(\textbf{new}) (0 () .
-0 \langle \texttt{"Hello world"} \rangle .
+0 \langle \texttt{"Hello world!"} \rangle .
 \mathbf{0} \parallel (\textbf{new}) (1 \langle 0 \rangle .
@@
   We will often work with input and output multiplicities as pairs.
-  We use the notation $\Carrier^2$ to stand for the type $\Carrier \times \Carrier$, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
+  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
 \end{note}
@@
 We introduce two judgments: one for variable references (Definition \ref{def:varref}), one for processes (Definition \ref{def:types}).
+Both of them combine a judgment on types and a judgment on usages.
+These could be kept separate, but doing so would risk incoherence: if typing derivations had any choice to make, the judgment on types could choose to do one thing while the judgment on usages chooses to do another.
+The constructions instroduced in this paper make the typing derivations completely syntax directed and render this problem inexistent, but we nevertheless choose to keep both judgments as one, for simplicity.
 
@@
 
-  Like processes, variable reference judgments (Figure \ref{fig:variable-references}) are indexed by a typing context and an input and output usage context.
-  In addition to these, variable reference judgments are also indexed by their position in the contexts, their type, and their usage annotation.
+  The \textsc{VarRef} relation presented in Figure \ref{fig:variable-references} is indexed by a typing context and an input and output usage context.
+  In addition to these, it also keeps track of its position in the contexts, its type, and its usage annotation.
 
@@
 
-  Our judgment on processes is of the form $\types{\gamma}{\Gamma}{P}{\Delta}$ (Figure \ref{fig:types}).
-  This assertion combines a judgment on types ($\gamma$) and a judgment on usages ($\Gamma$ and $\Delta$).
-  They could be kept as two separate judgments, but doing so would risk incoherence: if there is any choice to be made, the judgment on types could choose to go left while the judgment on usages chooses to go right.
-  The constructions instroduced in this paper make the typing derivations completely syntax directed and thus this problem inexistent, but we nevertheless choose to keep both judgments as one for simplicity.
-
-  The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed over a context of types $\gamma$, a context of input multiplicities $\Gamma$, the process $P$ that is being typed, and the context of output capabilities $\Delta$, and notated as $\types{\gamma}{\Gamma}{P}{\Delta}$.
-  \todo{empty output context}
-  The typing rules are in one-to-one correspondence with the underlying syntax for processes, and thus we say they are syntax-directed.
-
-  Process termination is typed under any typing context, and preserves the multiplicity context unchanged.
-
-  Scope restriction accepts three parameters: the type $t$ of data sent over the channel; the input and output multiplicities $x$ of the data sent over the channel; and the multiplicity $y$ of the channel itself, which is used both as input and as output.
-  The typing derivation of the continuation process $P$ has a channel of data type $t$ and data multiplicity $x$ added to the typing context, and the multiplicities $(y , y)$ added to the input multiplicity context.
-  The continuation must completely exhaust this channel, to that end we substitute $(y , y)$ by $(\zero , zero)$ in the output context.
-  \todo{comment on balancing in chan rule}
-
-  Input requires a proof $chan_i$ locating a channel with data type $t$, data multiplicities $x$ and multiplicity $\ell_i$ \todo{} at index $i$ in the typing and multiplicity contexts $\gamma$ and $\Gamma$.
-  The result of exhausting $\ell_i$ is the output multiplicity $\Xi$.
-  The data type $t$ is then used to expand the typing context of the continuation $cont$, while the data multiplicities $x$ are appended to the output context $\Xi$ to form the input context over which $P$ is typed.
-  These multiplicities $x$ must be exhausted by typing derivation of the continuation process.
-  Note that the index $i$ is then used as part of the syntax.
-
-  \todo{note that every new multiplicity introduced in an input context must be exhausted in the output context}
-
-  Similar to input, output requires a proof $chan_i$ locating a channel with data type $t$ and data multiplicities $x$ in $\gamma$ at index $i$, however this time the channel must have multiplicity $\ell_o$.
-  The result of exhausting $\ell_o$ is the output multiplicity $\Delta$.
-  That output context $\Delta$ must have a variable of type $t$ and multiplicities $x$ at position $j$.
-  This second proof represents the data that is sent over the channel.
-  The result of exhausting the multiplicities $x$ of that variable is the context $\Xi$.
-  The continuation process must be typed under this context $\Xi$.
-
-  Finally, parallel composition requires its two subprocesses to be typed such that the output context of the first is the input context of the second.
+  The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed by a typing context, an input and an output usage context, and the process $P$ being typed.
+
+  The process termination case is only concerned with usage contexts.
+  Its input usage context is returned as leftovers.
+
+  Scope restriction expects three arguments: the type $t$ of data being transmitted; the input and output multiplicities $x$ of the data being transmitted; and the multiplicity $y$ assigned to the channel itself.
+  The same multiplicity $y$ is used as the input multiplicity and as the output multiplicity of the channel, so these are balanced.
+  The typing derivation of the continuation process $P$ is provided with this channel with multiplicity $(y \comma y)$.
+  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\ell_{\o}$ in the output context.
+
+  \todo{comment on synonyms $\ell_i$ etc}
+
+  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\ell_i$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Note that this index $i$ is then used in the syntax of the typed process.
+  After consuming $\ell_i$ from this channel, we use the leftover context $\Xi$ to type the continuation.
+  The continuation is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
+  This multiplicity $x$ must be completely exhausted by the continuation process.
+
+  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\ell_o$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Note that this index $i$ is then used in the syntax of the typed process.
+  After consuming $\ell_o$ from this channel, we use the leftover context $\Delta$ to type the continuation.
+  The continuation must select the element at index $j$ that is being sent.
+  This element must have type $t$ and at least multiplicity $x$ --- as per the channel's type.
+  The result of consuming these multiplicity $x$ is the leftover context $\Xi$, which is then used to type the continuation process.
+  Note that both indices $i$ and $j$ are used in the syntax of the typed process.
+
+  Finally, parallel composition is only concerned with usage contexts.
+  The leftovers of the left-hand process are used to type the right-hand process.
 

### 2020-04-30 paper/paper.tex
Comment on pair synonyms
Align typing rules

@@
 \newcommand{\one}{\field{1\cdot}}
-\newcommand{\li}{\field{\ell_i}}
-\newcommand{\lo}{\field{\ell_o}}
-\newcommand{\lz}{\field{\ell_{\o}}}
-\newcommand{\lio}{\field{\ell_{\#}}}
+\newcommand{\li}{\ell_i}
+\newcommand{\lo}{\ell_o}
+\newcommand{\lz}{\ell_{\o}}
+\newcommand{\lio}{\ell_{\#}}
 
@@
   We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
+  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$.
 \end{note}
@@
   The typing derivation of the continuation process $P$ is provided with this channel with multiplicity $(y \comma y)$.
-  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\ell_{\o}$ in the output context.
-
-  \todo{comment on synonyms $\ell_i$ etc}
-
-  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\ell_i$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\lz$ in the output context.
+
+  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\li$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
   Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\ell_i$ from this channel, we use the leftover context $\Xi$ to type the continuation.
+  After consuming $\li$ from this channel, we use the leftover context $\Xi$ to type the continuation.
   The continuation is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
@@
 
-  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\ell_o$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\lo$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
   Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\ell_o$ from this channel, we use the leftover context $\Delta$ to type the continuation.
+  After consuming $\lo$ from this channel, we use the leftover context $\Delta$ to type the continuation.
   The continuation must select the element at index $j$ that is being sent.
@@
 
-  Finally, parallel composition is only concerned with usage contexts.
-  The leftovers of the left-hand process are used to type the right-hand process.
+  Parallel composition is only concerned with usage contexts.
+  It uses the leftovers of the left-hand process to type the right-hand process.
 
@@
         {\stacked{
-            chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P}{\Theta \comma \lz}}}
+            chan_i : \contains{\gamma \hspace{1.2em}}{\Gamma \hspace{1.3em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
+            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{5.6em}}{\Theta \comma \lz}}}
         {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
@@
         {\stacked{
-            chan_i : \contains{\gamma}{\Gamma}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t}{x\hspace{4em}}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma}{\Xi}{P\hspace{6em}}{\Theta}}}
+            chan_i : \contains{\gamma}{\Gamma \hspace{0.2em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t \hspace{3.7em}}{x \hspace{0.3em}}{\Xi} \\\\
+            cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{5.8em}}{\Theta}}}
         {\constr{send} \; chan_i \; loc_j \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
@@
     \inferrule
-    {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
+    {l \hspace{0.3em} : \types{\gamma}{\Gamma \hspace{0.2em}}{P}{\Delta} \\\\
      r : \types{\gamma}{\Delta}{Q}{\Xi}}
@@
   \end{figure}
-  \todo{align typing rules to make them more readable}
 \end{nidefinition}
@@
 \subsubsection{Example derivations}
+\label{example-derivations}
 
@@
     \item
-      In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
+      In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\lz$, making the proof trivial.
       In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
-      For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
+      For this, we will use type $\unit$ and multiplicities $\lz$.
     \item
@@
   \label{thm:substitution}
-  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \ell_{\o}}$.
+  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $j$ be the index of the data with $m$ multiplicities that is being received after $P$ is done, so that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
@@
   Assume that $P$ uses all $m$ multiplicities at index $i$.
-  To model this, require $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$ where index $i$ is $\ell_{\o}$.
+  To model this, require $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$ where index $i$ is $\lz$.
   The process that is typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$ will be typed $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$ after substitution.
@@
       \draw[-]  (gamma-m) -- (delta-m.center);
-      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \ell_{\o}$}(psi-m);
+      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \lz$}(psi-m);
 
@@
       From constructor $\constr{end}$ we derive $\Gamma_i \equiv \Psi_i$.
-      From $\Delta_i \equiv \ell_{\o}$ follows that $m \equiv n$.
+      From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
       Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.
@@
       For constructor $\constr{recv}$ we split $\Delta$ so that we can proceed inductively on the continuation.
-      We first observe that given that there exists an arrow from $\Gamma_i$ to $\Psi_i$ and $\Delta$ at index $i$ is $\ell_{\o}$, for any $m$ and $n$, there is some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+      We first observe that given that there exists an arrow from $\Gamma_i$ to $\Psi_i$ and $\Delta$ at index $i$ is $\lz$, for any $m$ and $n$, there is some $\delta$ such that $\opsquared{m}{\delta}{n}$.
  l     \begin{itemize}
         \item
-          If the input occurs on the variable that is being substituted, we derive $\opsquared{m}{\ell_i}{l}$ for some $l$ and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ (used for the inductive call).
-          Similarly, we construct the arrows $\containsusage{\Gamma_j}{j}{\ell_i}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ (used for the inductive call) for some $\Xi_j$.
-        \item
-          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $containsusage{\Gamma}{x}{\ell_i}{\Theta}$ for some $\Theta$.
-          We can then find $\Xi_j$ and construct the two arrows $\containsusage{\Gamma_j}{x}{\ell_i}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (used for the inductive call).
+          If the input occurs on the variable that is being substituted, we derive $\opsquared{m}{\li}{l}$ for some $l$ and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ (used for the inductive call).
+          Similarly, we construct the arrows $\containsusage{\Gamma_j}{j}{\li}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ (used for the inductive call) for some $\Xi_j$.
+        \item
+          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
+          We can then find $\Xi_j$ and construct the two arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (used for the inductive call).
           From the fact that there exists a composition of arrows from $\Xi_i$ to $\Psi$, we can conclude that $\Theta$ is splitting $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-          As $\ell_{\o}$ is a minimal element, if $\Delta$ is $\ell_{\o}$ at index $i$, then it must be $\ell_{\o}$ at index $i$ for $\Delta_1$ and $\Delta_2$ as well.
+          As $\lz$ is a minimal element, if $\Delta$ is $\lz$ at index $i$, then it must be $\lz$ at index $i$ for $\Delta_1$ and $\Delta_2$ as well.
       \end{itemize}
@@
       For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-      Given that $\Delta$ is $\ell_{\o}$ at index $i$, we know that $\Delta_1$ and $\Delta_2$ are so too.
+      Given that $\Delta$ is $\lz$ at index $i$, we know that $\Delta_1$ and $\Delta_2$ are so too.
       We observe that $\opsquared{m}{\delta}{\omega}$ where $\omega$ is what subprocess $P$ consumes at index $i$.
@@
   If $c$ is $\constr{internal}$ to the process, the context must not be modified and thus $\Gamma' \equiv \Gamma$.
-  If $c$ is $\constr{external}\; i$, the input usage context at index $i$ must be decremented, and therefore $\containsusage{\Gamma'}{i}{\ell_\#}{\Gamma}$.
+  If $c$ is $\constr{external}\; i$, the input usage context at index $i$ must be decremented, and therefore $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
   Then the typing derivation $\types{\gamma}{\Gamma'}{P}{\Xi}$ can be transformed into $\types{\gamma}{\Gamma}{Q}{\Xi}$.
@@
         \begin{alignat*}{2}
-          \containsusage{&\Gamma'}{i}{\ell_\# &&}{\Gamma} \\
-          \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\ell_i &&}{\Psi} \\
-          \types{\gamma \comma t}{&\Psi \comma m}{P &&}{\Delta \comma \ell_{\o}} \\
-          \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\ell_o &&}{\Theta} \\
+          \containsusage{&\Gamma'}{i}{\lio &&}{\Gamma} \\
+          \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\li &&}{\Psi} \\
+          \types{\gamma \comma t}{&\Psi \comma m}{P &&}{\Delta \comma \lz} \\
+          \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\lo &&}{\Theta} \\
           \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
@@
         \begin{alignat*}{2}
-          \types{\gamma \comma t}{& \Gamma \comma m}{P &&}{\Theta \comma \ell_{\o}} \\
+          \types{\gamma \comma t}{& \Gamma \comma m}{P &&}{\Theta \comma \lz} \\
           \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
         \end{alignat*}
-        \caption{Rearrange $\ell_o$ witness, and merge with $\ell_i$ witness into $\ell_\#$.}
+        \caption{Rearrange $\lo$ witness, and merge with $\li$ witness into $\lio$.}
       \end{subfigure}
@@
     If the channel is $\constr{external}\; \constr{zero}$ -- i.e.
-the channel that is being introduced by the scope restriction -- we use the Lemma \ref{lm:comm-capable} to subtract $\ell_\#$ from the channel's multiplicity first, then apply the induction hypothesis.
+the channel that is being introduced by the scope restriction -- we use the Lemma \ref{lm:comm-capable} to subtract $\lio$ from the channel's multiplicity first, then apply the induction hypothesis.
     If the channel is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
@@
   \label{lm:comm-capable}
-  Every input usage context $\Gamma$ on a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces such that $P \reduce{\constr{external} \; i} Q$ has a multiplicity of at least $\ell_\#$ at index $i$.
+  Every input usage context $\Gamma$ on a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces such that $P \reduce{\constr{external} \; i} Q$ has a multiplicity of at least $\lio$ at index $i$.
 \end{nilemma}

### 2020-04-30 paper/paper.tex
another pass to intro... getting there!

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda \thanks{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
+``Behavioural Application Program Interfaces'' (BehAPI).}
+}
 \titlerunning{$\pi$ with leftovers: all mechanisation in Agda}
@@
 We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
-As humans, we aim to model, predict, and even construct such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-
-The \picalc{} introduced by Milner et al. in the early '90s \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
+As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
+
+The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
 Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted.
-The \picalc{} is at the basis of programming language implementations for concurrency, such as Pict \cite{Pierce} and more recently Go \cite{Golang}.
-Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types, ranging from basic types, linear types, types for liveness properties, such as deadlock freedom or termination, and most notably, session types \cite{K07}, making the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
-
-In a parallel track in the early '90s, the discovery of linear logic by J.-Y.
-Girard \cite{Girard87} opened new avenues in computing science and introduced linear types and type systems for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}, allowing for resource-aware programming and efficient implementation.
-Linearity later led to uniqueness of typing in Clean \cite{BarendsenS96} or ownership typing in Rust \cite{MatsakisK14}.
-
-The \picalc{} was not an exception to the influence of linearity.
-Kobayashi et al. \cite{KPT96} defined the \emph{linear} \picalc{}, which is a typed version of the \picalc{} where the linear type system restricts the usage of channel resources to exactly ones.
-%
-With the rise of session types \cite{H93,THK94,HVK98}, linearity was reproposed in the \picalc{}, which is
-the most successful framework where session types have been used.
-Session types are a type formalism used in communication-structured modelling and programming and linearity is a key feature in order to guarantee type safety.
-Recent work brought back to light the linear \picalc{}, because session types can be expressed in terms of linear types \cite{DardhaGS12,DardhaGS17}.
-The encoding of session types into linear types not only has theoretical benefits in terms expressivity of session types and reusability of theoretical results from the linear \picalc{}, but also at the implementation level.
-The encoding has been used as a technique to implement session types languages such as OCaml \cite{Padovani17} or Scala \cite{ScalasY16,ScalasDHY17}.
-As such, linear \picalc{} is the underlying theoretical and practical framework of session types.
-
-Considering the relevance of the \picalc{} as a core calculus to model concurrent and distributed system and linearity for resource-aware programming, in this paper we take it a step further and investigate a \picalc{} featuring linear, gradual and shared types under the same unique framework and mechanise it all in Agda.
-We can use this framework as a core for type safe distributed programming with a range of types from linear to shared.
-
-\subsection{Contribution}
-Our contributions can be summarised as follows later explained in details, together with the corresponding sections where they can be found.
+The \picalc{} is at the basis of design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and more recently Go \cite{Golang}.
+Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types, ranging from basic types, linear types, types for liveness properties, such as deadlock freedom, livelock freedom or termination, and most notably, session types \cite{K07}, making the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
+
+In a parallel track, the discovery of linear logic by J.-Y.
+Girard \cite{Girard87} opened new avenues in computing science in the early '90s and introduced \emph{linear types} for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}.
+Linear type systems restrict the usage of resources to \emph{exactly ones}.
+Enforcing ``no duplicating'' and ``no discarding'' of resources via linearity, allows for resource-aware programming and efficient implementation \cite{Wadler90}.
+Later on, linearity inspired unique types, as in Clean \cite{BarendsenS96} or ownership types, as in Rust \cite{MatsakisK14}.
+
+The \picalc{} also benefited from this linearity wave in early '90s.
+Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, which is a typed version of the \picalc{} where the linear type system restricts the usage of channels to {exactly ones}.
+
+In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave linearity in the \picalc{} a different flavour.
+Session types are a type formalism used in communication-centric programming and the \picalc{} is the original and still
+the most successful framework where session types are defined.
+Linearity in session types means that a channel can be owned only by one communicating participant, but the channel itself can be used multiple times, as specified by the structure of a session type.
+Linearity in session types is a key ingredient in order to guarantee type safety.
+
+Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} turned again the spotlight on the linear \picalc{} \cite{KPT96}, via an encoding of session types into linear types.
+This encoding not only has theoretical benefits in terms expressivity of session types and reusability of theoretical results from the linear \picalc{}, but most importantly it is useful at a practical level.
+The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} or Scala \cite{ScalasY16,ScalasDHY17}.
+This means we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which we can build session types.
+
+Considering the importance of the \picalc{} and linearity in modelling concurrent and distributed systems, in this paper we focus our investigation on both and go beyond.
+We present for the first time a full mechanisation in Agda of a \picalc{} with linear, gradual and shared types all under the same unified framework.
+
+In order to obtain this unification of type systems, we use leftover typing for the first time in the \picalc{}, following Allais \cite{Allais2018a}.
+The user can choose among the possible type systems by instantiating our $\pi$ with leftovers, which gives as a result one of the three typed $\pi$s.
+Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modelling and programming with a range of type systems, on top of which we can build more advanced types, theories and languages.
+While linearity is needed for type safety, gradual types and shared types are needed to build real-world systems.
+
+\subsection{Contributions}
+Our contributions are given in the following together with the corresponding sections.
 \begin{itemize}
-  \item Machine verified formalisation of the syntax and semantics of linear \picalc{}
-
-  \item Typing with leftovers applied to the \picalc{}, where a range of type systems from linear to gradual and shared can be obtained
-
-  \item Abstraction over multiplicities
-
-  \item Full proofs of subject reduction and auxiliary lemmas
-
-  \item Full formalisation available in Agda
-\end{itemize}
-
-\paragraph*{Syntax}
-We type the syntax for the \picalc{} in \S \ref{syntax}.
+  \item \textbf{Machine verified formalisation of the syntax and semantics of the \picalc{}}: The syntax for the \picalc{} is given in \S \ref{syntax}.
   We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
-In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
-One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence \cite{} relation that acts as a quotient type.
-A term $A$ can then be freely rewritten into a term $B$ as long as $A$ is structurally congruent to $B$.
-We type this equivalence relation in \S \ref{structural-congruence}.
-
-\paragraph*{Semantics}
-In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
-Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequential behavior.
-We type a reduction relation that acts as an operational semantics modulo subject congruence in \S \ref{operational-semantics}.
-The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it to state the theorem for subject reduction (aka type preservation).
-
-As with the untyped \lambdacalc{}, the untyped \picalc{} can be used to construct a wide variety of ill-behaved terms --- for some characterization of ill-behaved.
-One can use type systems to constructively obtain a subset of terms for which certain properties hold.
-As an example, the simply typed \lambdacalc{} ensures that the argument supplied to a function is of the expected type.
-Correspondingly, the simply typed \picalc{} \cite{} ensures that the data sent over a channel is of the expected type.
-This property is often referred to as \emph{communication safety}.
-
-\todo{Mention that our semantics is prior to our typing (Curry-style)}
-
-Another property of particular interest to concurrent systems is the avoidance of race conditions.
-A type system that excludes multiple usages of a resource excludes race conditions.
-In the linear \picalc{} \cite{} channels must be used exactly once.
-This not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
-It also ensures that every declared communication channel must be used.
-The linear \picalc{} is also interesting because together with product types it is isomorphic to session types \cite{} -- which are used to encode communication protocols.
-We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
-
-\paragraph*{Typing with leftovers}
+  We define the operational semantics of the \picalc{} by a reduction relation in \S \ref{operational-semantics}.
+  The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it for subject reduction (aka type preservation) theorem.
+
+  \item \textbf{Machine verified typing with leftovers for the \picalc{} with linear, gradual and shared types}:
   Linear type systems need of context splits.
@@
   \todo{reformulate paragraph}
-
-\paragraph*{Subject reduction}
-
-\paragraph*{On mechanisation in Agda}
+  Another property of particular interest in concurrent and distributed systems is the avoidance of race conditions.
+  A type system that excludes multiple usages of a resource excludes race conditions.
+Linear \picalc{} not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
+  It also ensures that every declared communication channel must be used.
+  We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
+
+  \item \textbf{Machine verified proofs of subject reduction and auxiliary lemmas for all $\pi$ with leftovers}:
+
+  \item \textbf{Full formalisation in Agda available at \cite{Zalakain2020Agda}}:
+\end{itemize}
+
+%\paragraph*{Syntax}
+%In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
+%One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence \cite{} relation that acts as a quotient type.
+%A term $A$ can then be freely rewritten into a term $B$ as long as $A$ is structurally congruent to $B$.
+
+%\paragraph*{Semantics}
+% In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
+% Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequential behavior.
+% We type a reduction relation that acts as an operational semantics modulo subject congruence in \S \ref{operational-semantics}.
+% The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it to state the theorem for subject reduction (aka type preservation).
+
+% As with the untyped \lambdacalc{}, the untyped \picalc{} can be used to construct a wide variety of ill-behaved terms --- for some characterization of ill-behaved.
+% One can use type systems to constructively obtain a subset of terms for which certain properties hold.
+% As an example, the simply typed \lambdacalc{} ensures that the argument supplied to a function is of the expected type.
+% Correspondingly, the simply typed \picalc{} \cite{} ensures that the data sent over a channel is of the expected type.
+% This property is often referred to as \emph{communication safety}.
+
+% \todo{Mention that our semantics is prior to our typing (Curry-style)}
+
+% Another property of particular interest to concurrent systems is the avoidance of race conditions.
+% A type system that excludes multiple usages of a resource excludes race conditions.
+% In the linear \picalc{} \cite{} channels must be used exactly once.
+% This not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
+% It also ensures that every declared communication channel must be used.
+% The linear \picalc{} is also interesting because together with product types it is isomorphic to session types \cite{} -- which are used to encode communication protocols.
+% We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
+
+% \paragraph*{Typing with leftovers}
+% Linear type systems need of context splits.
+% These are tricky to handle constructively without explicitly asking the user to supply them.
+% Leftover typing \cite{} is a type theoretical technique that addresses this.
+% It does so by carrying both an input and an output context (the leftovers).
+% The output context can then be used as the input context of the continuations.
+% We apply this technique to the \picalc{} in \S \ref{leftover-typing}.
+% We generalize our algebra on usage multiplicities in \S \ref{multiplicities}.
+% As a result, we are able to model a mix of shared, linear, and gradual types.
+% \todo{reformulate paragraph}
 

### 2020-04-30 paper/paper.tex
Provide example derivation

@@
 \bibliographystyle{plainurl}
+
+% Example code
+\usepackage{listings}
 
@@
 Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in Figure \ref{fig:multiplicities}.
-For examples of how these multiplicity types are used in typing derivations, refer to \S \ref{example-derivations}.
+For examples of how these multiplicity types are used in typing derivations, refer to Example \ref{example-derivations}.
 
@@
 
-  The base case $\constr{zero}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
+  The base case $\constr{here}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
   This is the only place in the typing system where the algebra on multiplicities is used.
@@
 
-  The inductive case $\constr{suc}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.
+  The inductive case $\constr{there}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.
 
@@
     {\opsquared{x}{y}{z}}
-    {\constr{zero} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}
+    {\constr{here} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}
 
@@
     {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
-    {\constr{suc} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
+    {\constr{there} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
   \end{mathpar}
@@
 \begin{example}
-\todo{example}
+  The following is an example that asserts that the input usage context is capable of $\li$ on a $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{zero}$, and that consuming such $\li$ results in the stated leftovers.
+  \begin{lstlisting}[mathescape,caption=Here \_ introduces an anonymous declaration that must be immediately followed by its definition.]
+  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{zero}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
+  _ = $\constr{there} \; \constr{here}$
+  \end{lstlisting}
 \end{example}
@@
   The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed by a typing context, an input and an output usage context, and the process $P$ being typed.
-
-  The process termination case is only concerned with usage contexts.
-  Its input usage context is returned as leftovers.
-
-  Scope restriction expects three arguments: the type $t$ of data being transmitted; the input and output multiplicities $x$ of the data being transmitted; and the multiplicity $y$ assigned to the channel itself.
-  The same multiplicity $y$ is used as the input multiplicity and as the output multiplicity of the channel, so these are balanced.
-  The typing derivation of the continuation process $P$ is provided with this channel with multiplicity $(y \comma y)$.
-  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\lz$ in the output context.
-
-  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\li$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
-  Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\li$ from this channel, we use the leftover context $\Xi$ to type the continuation.
-  The continuation is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
-  This multiplicity $x$ must be completely exhausted by the continuation process.
-
-  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\lo$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
-  Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\lo$ from this channel, we use the leftover context $\Delta$ to type the continuation.
-  The continuation must select the element at index $j$ that is being sent.
-  This element must have type $t$ and at least multiplicity $x$ --- as per the channel's type.
-  The result of consuming these multiplicity $x$ is the leftover context $\Xi$, which is then used to type the continuation process.
-  Note that both indices $i$ and $j$ are used in the syntax of the typed process.
-
-  Parallel composition is only concerned with usage contexts.
-  It uses the leftovers of the left-hand process to type the right-hand process.
 
@@
   \end{figure}
+
+  The process termination case is only concerned with usage contexts.
+  Its input usage context is returned as leftovers.
+
+  Scope restriction expects three arguments: the type $t$ of data being transmitted; the input and output multiplicities $x$ of the data being transmitted; and the multiplicity $y$ assigned to the channel itself.
+  The same multiplicity $y$ is used as the input multiplicity and as the output multiplicity of the channel, so these are balanced.
+  The typing derivation of the continuation process $P$ is provided with this channel with multiplicity $(y \comma y)$.
+  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\lz$ in the output context.
+
+  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\li$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Note that this index $i$ is then used in the syntax of the typed process.
+  After consuming $\li$ from this channel, we use the leftover context $\Xi$ to type the continuation.
+  The continuation is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
+  This multiplicity $x$ must be completely exhausted by the continuation process.
+
+  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\lo$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
+  Note that this index $i$ is then used in the syntax of the typed process.
+  After consuming $\lo$ from this channel, we use the leftover context $\Delta$ to type the continuation.
+  The continuation must select the element at index $j$ that is being sent.
+  This element must have type $t$ and at least multiplicity $x$ --- as per the channel's type.
+  The result of consuming these multiplicity $x$ is the leftover context $\Xi$, which is then used to type the continuation process.
+  Note that both indices $i$ and $j$ are used in the syntax of the typed process.
+
+  Parallel composition is only concerned with usage contexts.
+  It uses the leftovers of the left-hand process to type the right-hand process.
 \end{nidefinition}
 
-\subsubsection{Example derivations}
+\begin{example}
   \label{example-derivations}
-
-\todo{OD: in addition to the above, could we have examples of typing a process in each of the above type systems? It can be the same process everywhere and we can show how the typing changes...}
+  The following example uses both linear and shared types to type the process $p$. 
+  We use $\omega$ as a synonym for a $(\one \comma \one)$ multiplicity in an algebra for shared types.
+  The usage context $\epsilon$ signifies the empty context where all multiplicities are $(\zero \comma \zero)$.
+  The typing derivation creates a new channel over which $\li$ multiplicities of a type $\channel{\unit}{\omega}$ are sent.
+  This channel is initialized with a linear multiplicity of $(\one \comma \one)$, and therefore must be used exactly once for sending and once for receiving.
+  The right-hand parallel subprocess then creates a new channel as per the specification $\channel{\unit}{\omega}$.
+
+  \begin{lstlisting}[mathescape]
+  $p$ : $\Process_1$
+  $p$ = $\new{} (\comp{\recv{\constr{zero}} (\recv{\constr{zero}} \PO)}
+                    {\new{} (\send{\constr{suc} \; \constr{zero}} {\constr{zero}} (\send{\constr{zero}}{\constr{suc} \; \constr{suc} \; \constr{zero}}) \; \PO)})$
+
+  _ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \omega}{p}{\epsilon}$
+  _ = $\constr{chan} \; \channel{\unit}{\omega} \; \li \; \one$
+      $(\constr{comp} \;
+       (\constr{recv} \; \constr{here} \; (\constr{recv} \; \constr{here} \; \constr{end}))$
+           $(\constr{chan} \; \unit \; \omega \; \one \; (\constr{send} \; (\constr{there} \; \constr{here}) \; \constr{here} \; (\constr{send} \; \constr{here} \; (\constr{there} \; \constr{here} \; \constr{there}) \; \constr{end})))
+       )$
+  \end{lstlisting}
+\end{example}
 
@@
   \label{lm:types-unused}
-  If $P$ is well-typed under input context $\Gamma$ and output context $\Xi$, and variable $i$ is unused within $P$, then $\Gamma_i \equiv \Xi_i$.
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
 \end{nilemma}
@@
   \label{lm:types-op}
-  For any well typed-process $P$ with input context $\Gamma$ and output context $\Xi$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
 \end{nilemma}

### 2020-04-30 paper/paper.tex
Use the funding macro provided by lipics

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{$\pi$ with leftovers: \\ a mechanisation in Agda \thanks{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
-``Behavioural Application Program Interfaces'' (BehAPI).}
-}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
 \titlerunning{$\pi$ with leftovers: all mechanisation in Agda}
@@
 \acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+\funding{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
+  ``Behavioural Application Program Interfaces'' (BehAPI).}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-30 paper/paper.tex
Give the intro another pass (mainly grammar)

@@
 Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted.
-The \picalc{} is at the basis of design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and more recently Go \cite{Golang}.
-Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types, ranging from basic types, linear types, types for liveness properties, such as deadlock freedom, livelock freedom or termination, and most notably, session types \cite{K07}, making the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
-
-In a parallel track, the discovery of linear logic by J.-Y.
-Girard \cite{Girard87} opened new avenues in computing science in the early '90s and introduced \emph{linear types} for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}.
-Linear type systems restrict the usage of resources to \emph{exactly ones}.
-Enforcing ``no duplicating'' and ``no discarding'' of resources via linearity, allows for resource-aware programming and efficient implementation \cite{Wadler90}.
-Later on, linearity inspired unique types, as in Clean \cite{BarendsenS96} or ownership types, as in Rust \cite{MatsakisK14}.
+The \picalc{} acts as a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
+Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types \cite{K07}.
+This makes the \picalc{} a fully-fledged tool for the modelling and verification of concurrent and distributed systems.
+
+In a parallel track, J.-Y.
+Girard's development of linear logic \cite{Girard87} in the early '90s opened new avenues in computing science by introducing \emph{linear types} for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}.
+Linear type systems guarantee that resources are used \emph{exactly once}.
+Enforcing ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
+Later on, linearity inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
 
 The \picalc{} also benefited from this linearity wave in early '90s.
-Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, which is a typed version of the \picalc{} where the linear type system restricts the usage of channels to {exactly ones}.
-
-In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave linearity in the \picalc{} a different flavour.
-Session types are a type formalism used in communication-centric programming and the \picalc{} is the original and still
-the most successful framework where session types are defined.
-Linearity in session types means that a channel can be owned only by one communicating participant, but the channel itself can be used multiple times, as specified by the structure of a session type.
-Linearity in session types is a key ingredient in order to guarantee type safety.
-
-Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} turned again the spotlight on the linear \picalc{} \cite{KPT96}, via an encoding of session types into linear types.
-This encoding not only has theoretical benefits in terms expressivity of session types and reusability of theoretical results from the linear \picalc{}, but most importantly it is useful at a practical level.
-The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} or Scala \cite{ScalasY16,ScalasDHY17}.
-This means we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which we can build session types.
-
-Considering the importance of the \picalc{} and linearity in modelling concurrent and distributed systems, in this paper we focus our investigation on both and go beyond.
-We present for the first time a full mechanisation in Agda of a \picalc{} with linear, gradual and shared types all under the same unified framework.
-
-In order to obtain this unification of type systems, we use leftover typing for the first time in the \picalc{}, following Allais \cite{Allais2018a}.
-The user can choose among the possible type systems by instantiating our $\pi$ with leftovers, which gives as a result one of the three typed $\pi$s.
-Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modelling and programming with a range of type systems, on top of which we can build more advanced types, theories and languages.
-While linearity is needed for type safety, gradual types and shared types are needed to build real-world systems.
+Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, a typed version of the \picalc{} where the linear type system restricts the usage of channels, which must be used exactly once.
+
+In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave a different flavour to the linearity in the \picalc{}.
+Session types are a type formalism used in communication-centric programming.
+In session types, linearity ensures that a channel must be owned by exactly one communicating participant, even if the channel itself can be used multiple times, as specified by the structure of its session type.
+Linearity in session types is the key ingredient that guarantees type safety.
+
+Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types.
+This encoding not only has theoretical benefits in terms of expressivity of session types and reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
+The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
+This means that we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
+
+Considering the importance of the \picalc{} and of linearity in modelling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
+We present for the first time a full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.
+We do so by defining a specification for an algebra on multiplicities and using it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{} for the first time.
+The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
+Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modelling and programming with a variety of type systems, on top of which more advanced types, theories and languages can be built.
+While linearity is needed for type safety, gradual types and shared types are needed for real-world systems.
 

### 2020-04-30 paper/paper.tex
Go through the contributions section

@@
 
-The \picalc{} also benefited from this linearity wave in early '90s.
+The \picalc{} also benefited from this linearity wave in the early '90s.
 Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, a typed version of the \picalc{} where the linear type system restricts the usage of channels, which must be used exactly once.
@@
 \subsection{Contributions}
-Our contributions are given in the following together with the corresponding sections.
+Our contributions are the following:
 \begin{itemize}
-  \item \textbf{Machine verified formalisation of the syntax and semantics of the \picalc{}}: The syntax for the \picalc{} is given in \S \ref{syntax}.
-  We use de Bruijn indices \cite{} as variable references, and dependent types to make the syntax well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
-  We define the operational semantics of the \picalc{} by a reduction relation in \S \ref{operational-semantics}.
-  The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it for subject reduction (aka type preservation) theorem.
-
-  \item \textbf{Machine verified typing with leftovers for the \picalc{} with linear, gradual and shared types}:
-  Linear type systems need of context splits.
-  These are tricky to handle constructively without explicitly asking the user to supply them.
-  Leftover typing \cite{} is a type theoretical technique that addresses this.
-  It does so by carrying both an input and an output context (the leftovers).
-  The output context can then be used as the input context of the continuations.
-  We apply this technique to the \picalc{} in \S \ref{leftover-typing}.
-  We generalize our algebra on usage multiplicities in \S \ref{multiplicities}.
-  As a result, we are able to model a mix of shared, linear, and gradual types.
-  \todo{reformulate paragraph}
-  Another property of particular interest in concurrent and distributed systems is the avoidance of race conditions.
-  A type system that excludes multiple usages of a resource excludes race conditions.
-Linear \picalc{} not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
-  It also ensures that every declared communication channel must be used.
-  We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
-
-  \item \textbf{Machine verified proofs of subject reduction and auxiliary lemmas for all $\pi$ with leftovers}:
-
-  \item \textbf{Full formalisation in Agda available at \cite{Zalakain2020Agda}}:
+\item \textbf{Formalisation of the syntax and semantics of the \picalc{}}:
+  Section \S \ref{syntax} uses type level de Bruijn indices \cite{} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
+  Section \S \ref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
+  This operational semantics is given as a reduction relation on processes.
+  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- ak-a type preservation.
+  It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+
+  \item \textbf{Leftover typing for the \picalc{}, with linear, gradual and shared types}:
+  Section \S \ref{type-system} provides the type system for a resource-aware \picalc{}.
+  The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
+  Linear types, gradual types and shared types are all instances of such a resource-aware algebra too.
+  Any \emph{partial commutative monoid} that is cancellative and has a minimal element is a valid such algebra.
+  Multiple of these algebras can be applied to a single type system --- contexts keep information about what algebra should be applied to which element.
+
+  The typing judgments use \emph{leftover typing} to model the resource-aware \picalc{}.
+  This approach adds a leftover usage context to the typing judgments.
+  Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
+  These technique renders explicit context splits unecessary.
+
+  \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
+  Section \S \ref{type-safety} builds on several theorems to proof subject reduction for our type system.
+  Here too, leftover typing renders the proofs for some of these lemmas straightforward, while for others it makes it possible to state them in the first place.
+  We prove \textbf{framing} (well-typedness is independent of the unused resources of a process), \textbf{weakening} (new variables that are added to both the input and output usage context preserve well-typedness), and \textbf{strengthening} (removing the variables that are unused by a process preserves well-typedness).
+  We use these to prove \textbf{subject congruence} (structural congruence rules can be freely applied while preserving well-typedness) and \textbf{substitution} (a variable can be substituted for another as long as this other variable is capable of doing everything the original one did), and use these to prove \textbf{subject reduction} (a well-typed process that reduces will still be well-typed).
+
+  \item \textbf{Fully mechanised formalisation available in Agda}:
+  This work has been fully formalised in Agda and is publicly avaible \cite{Zalakain2020Agda}.
+
 \end{itemize}
-
-%\paragraph*{Syntax}
-%In the \lambdacalc{} all computations are local and happen on syntactically contiguous terms; in the \picalc{} however communicating subprocesses can be non-contiguous and even have different parent processes.
-%One way to define a reduction relation modulo these syntactic minutiae is to define a structural congruence \cite{} relation that acts as a quotient type.
-%A term $A$ can then be freely rewritten into a term $B$ as long as $A$ is structurally congruent to $B$.
-
-%\paragraph*{Semantics}
-% In the \lambdacalc{} function application is the main computational engine; in the \picalc{} it is inter-process communication what drives computation.
-% Two subprocesses can communicate anywhere in a process but under input and output, which must be respected to ensure the sequential behavior.
-% We type a reduction relation that acts as an operational semantics modulo subject congruence in \S \ref{operational-semantics}.
-% The base case of every such reduction is communication over a channel; we take care to track this information at the type level so that we can then use it to state the theorem for subject reduction (aka type preservation).
-
-% As with the untyped \lambdacalc{}, the untyped \picalc{} can be used to construct a wide variety of ill-behaved terms --- for some characterization of ill-behaved.
-% One can use type systems to constructively obtain a subset of terms for which certain properties hold.
-% As an example, the simply typed \lambdacalc{} ensures that the argument supplied to a function is of the expected type.
-% Correspondingly, the simply typed \picalc{} \cite{} ensures that the data sent over a channel is of the expected type.
-% This property is often referred to as \emph{communication safety}.
-
-% \todo{Mention that our semantics is prior to our typing (Curry-style)}
-
-% Another property of particular interest to concurrent systems is the avoidance of race conditions.
-% A type system that excludes multiple usages of a resource excludes race conditions.
-% In the linear \picalc{} \cite{} channels must be used exactly once.
-% This not only avoids races, but ensures \emph{communication privacy}: if both endpoints of a channel are acquired no other process can eavesdrop.
-% It also ensures that every declared communication channel must be used.
-% The linear \picalc{} is also interesting because together with product types it is isomorphic to session types \cite{} -- which are used to encode communication protocols.
-% We provide a typing system for the linear \picalc{} in \S \ref{type-system} as extensional rules on sitting on top of the unruly well-scoped syntax.
-
-% \paragraph*{Typing with leftovers}
-% Linear type systems need of context splits.
-% These are tricky to handle constructively without explicitly asking the user to supply them.
-% Leftover typing \cite{} is a type theoretical technique that addresses this.
-% It does so by carrying both an input and an output context (the leftovers).
-% The output context can then be used as the input context of the continuations.
-% We apply this technique to the \picalc{} in \S \ref{leftover-typing}.
-% We generalize our algebra on usage multiplicities in \S \ref{multiplicities}.
-% As a result, we are able to model a mix of shared, linear, and gradual types.
-% \todo{reformulate paragraph}
 
@@
 
-Data type definitions have double lines and index-free synonyms as rule names (for ease of reference).
+Data type definitions use double lines and index-free synonyms as rule names (for ease of reference).
 We otherwise use the constructor name as the name of a typing rule.

### 2020-04-30 paper/paper.tex
Small fixes here and there

@@
 
-\subsubsection{Typing Contexts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{nidefinition}
-  Figure \ref{fig:prectx} defines types as either unit types, base types or channel types.
+  We define types (Figure \ref{fig:prectx}) as either unit types, base types or channel types.
   The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
@@
 
-\subsubsection{Usage Contexts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We choose to allow multiple multiplicity algebras to coexist in our typing system.
-We do so by defining an indexed set of usage algebras.
-
 \begin{nidefinition}
+  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
   An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by Figure \ref{fig:indexed-multiplicities}.
@@
 This output context contains the \emph{leftovers} of the process being typed --- its unused multiplicities.
-Crucially, the leftovers can then be used as an input to some othew typing derivation.
-Adopting this techinque has two main advantages:
+Crucially, the leftovers can then be used as an input to some other typing derivation.
+Adopting this techinque has three main advantages:
 \begin{itemize}
   \item
-    The input and output actions do no longer need to transform the usage context through functions, but can use a variable reference type with both an input and an output usage context to encapsulate these transformations.
+    The input and output actions no longer transform the usage context through functions.
+    Instead, they use a variable reference type with both an input and an output usage context that encapsulates these transformations.
   \item
@@
     With leftover typing, we use the leftovers of $P$ to type $Q$.
+  \item
+    Theorems like framing (\S \ref{framing}), weakening (\S \ref{weakening}) and strengthening (\S \ref{strengthening}) become suddenly stateable.
 \end{itemize}

### 2020-04-30 paper/paper.tex
Move a lemma so that it precedes the theorem that depends on it

@@
 In \S \ref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
+
+\begin{nilemma}
+  \label{lm:comm-capable}
+  As an auxiliary lemma, we must establish that every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
+\end{nilemma}
+
+\begin{proof}
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\end{proof}
 
@@
     \item
-    Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and then the induction hypothesis.
+    Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and the induction hypothesis.
   \end{itemize}
@@
 
-\begin{nilemma}
-  \label{lm:comm-capable}
-  Every input usage context $\Gamma$ on a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces such that $P \reduce{\constr{external} \; i} Q$ has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
-\begin{proof}
-  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
-\end{proof}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

### 2020-05-01 paper/paper.tex
Remove incorrect claim

@@
 \begin{itemize}
-\item \textbf{Formalisation of the syntax and semantics of the \picalc{}}:
+\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
   Section \S \ref{syntax} uses type level de Bruijn indices \cite{} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
@@
 Structural congruence is a congruent equivalence relation on processes.
-Any two structurally congruent processes are strongly bisimilar: they are can follow each other's reduction steps \cite{}.
 Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.

### 2020-05-01 paper/paper.tex
we can't assume people know pi

--- 2020-04-30 paper/paper.tex
@@
 \newcommand{\new}{\constr{\boldsymbol{\nu}} \;}
-\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
+\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}\newcommand{\sendp}[2]{#1 \constr{\langle} #2 \constr{\rangle}}
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
+\newcommand{\recvp}[2]{#1 \constr{(}#2\constr{)}}
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
 \label{syntax}
+The syntax of the \picalc{} \cite{Sangio01} is given by the following grammar.
+We let $x, y,\ldots$ denote channel \emph{names} and
+$P, Q,\ldots$ denote processes.
+\begin{equation*}
+  \begin{split}
+    P,Q ::=      & \; \PO       &\text{(terminated process)}  \\
+                |& \; (\new{}x)\  P   & \text{(restriction)} \\
+                |& \; \comp{P}{Q}  & \text{(parallel composition)} \\
+                |& \; \recvp{x}{y}.P  & \text{(receive)} \\
+                |& \; \sendp{x}{y}.P   &\text{(send)}
+  \end{split}
+\end{equation*}
 
@@
 
+\newpage
 \bibliography{paper}

### 2020-05-01 paper/paper.tex
updating syntax

--- 2020-05-01 paper/paper.tex
@@
 \label{syntax}
-
-In the \picalc{} both scope restriction and input introduce abstractions.
-The bodies of these constructs need a way of referring to their argument.
-Using names as variable references is a popular option amongst humans.
-However, names are cumbersome to mechanize: inserting a new variable into a context means proving that the name of such variable is different to all other variable names in context.
-Moreover, names are of no significance to machines whatsoever.
-
-The idea de Bruijn had \cite{} was to use a natural number $n$ (aka \emph{index}) to refer to the variable introduced $n$ binders ago --- and to have binders no longer introduce names.
+We let $x, y,\ldots$ range over \emph{channel names} or \emph{variables} and
+$P, Q,\ldots$ over processes.
+The syntax of the \picalc{} \cite{Sangio01} is given by the following grammar.
+\begin{equation*}
+  \begin{split}
+    P,Q ::=      & \; \PO       &\text{(terminated process)}  \\
+                |& \; (\new{}x)  P   & \text{(restriction)} \\
+                |& \; \comp{P}{Q}  & \text{(parallel composition)} \\
+                |& \; \recvp{x}{y}.P  & \text{(input)} \\
+                |& \; \sendp{x}{y}.P   &\text{(output)}
+  \end{split}
+\end{equation*}
+Process $\PO$ denotes the terminated process, where no operations can further occur.
+Process $(\new{}x)P$ is the restriction process where channel $x$ is created new and is bound with scope $P$.
+Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
+Processes $\recvp{x}{y}.P$ and $\sendp{x}{y}.P$ denote respectively the input and output processes of a variable/channel $y$ over a channel $x$, with continuation $P$.
+
+Scope restriction $(\new{}x)P$ and input $\recvp{x}{y}.P$ both introduce bound names in $P$.
+In order to mechanise the \picalc{} syntax in Agda, we need a way to deal with bound names in these continuation processes.
+Channel names are cumbersome to mechanise: inserting a new variable into a context means proving that the name of such variable is different to all other variable names in context.
+Instead we adopt de Bruijn indices \cite{} where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
+\todo{rewrite the sytanx below, following the syntax above}
 \begin{example}
@@
 That is, terms at different \emph{depths} use different indices to refer to the same binding.
-Humans may find this confusing.
-There is however no reason not to annotate these indices with the original names bestowed upon them by humans.
-Machines can then manipulate references mechanically whilst using names to present terms to humans.
 
@@
 It is useful to rule out these nonsensical terms syntactically.
-In Figure \ref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+\todo{replace ref with autoref everywhere...}
+In  \autoref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 
@@
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-As shown in Figure \ref{fig:syntax}, we increase the \emph{depth} counter every time we create a new channel or receive some input.
+Figure \ref{fig:syntax} shows the syntax of \picalc{} in Agda. We increase the \emph{depth} counter by 1 in channel restriction and input processes.
 
@@
 \end{equation*}
-\caption{Well-scoped grammar using de Bruijn indices}
+\caption{Well-scoped grammar for \picalc{} using de Bruijn indices}
 \label{fig:syntax}
@@
 Using type-level de Bruijn indices makes our syntax well-scoped by construction --- the semantics of our language can now be defined on the totality of the syntax.
-User-friendliness can still be recovered through a function that converts processes with named variables into (possibly) processes with indexed variables.
-This function recursively traverses processes, keeping track of what name is associated with each index.
-If the process is ill-scoped, the function returns nothing.
-To present things to the user, we substitute indices by their related names.
+In order to make the de Bruijn indices representation more human-readable, we can annotate these indices with the original channel names.
+Machines can then manipulate references mechanically whilst using channel names to present \picalc{} terms back to humans.
 

### 2020-05-01 paper/paper.tex
Use autoref

@@
 \item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
-  Section \S \ref{syntax} uses type level de Bruijn indices \cite{} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
-  Section \S \ref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
+  \autoref{syntax} uses type level de Bruijn indices \cite{} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
+  \autoref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
   This operational semantics is given as a reduction relation on processes.
@@
   \item \textbf{Leftover typing for the \picalc{}, with linear, gradual and shared types}:
-  Section \S \ref{type-system} provides the type system for a resource-aware \picalc{}.
+  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
   The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
@@
   \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
-  Section \S \ref{type-safety} builds on several theorems to proof subject reduction for our type system.
+  \autoref{type-safety} builds on several theorems to proof subject reduction for our type system.
   Here too, leftover typing renders the proofs for some of these lemmas straightforward, while for others it makes it possible to state them in the first place.
@@
     { }
-    {\type{\N} : \Set}
+    {\type{\N} : et}
     \; \textsc{Nat}
@@
 We otherwise use the constructor name as the name of a typing rule.
-Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $et$.
 Implicit arguments are mentioned by type definitions but omitted in constructors.
@@
   {n : \N}
-  {\Var_n : \Set}
+  {\Var_n : et}
   \; \textsc{Var}
@@
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-Figure \ref{fig:syntax} shows the syntax of \picalc{} in Agda. We increase the \emph{depth} counter by 1 in channel restriction and input processes.
+\autoref{fig:syntax} shows the syntax of \picalc{} in Agda. We increase the \emph{depth} counter by 1 in channel restriction and input processes.
 
@@
   {n : \N}
-  {\Process_n : \Set}
+  {\Process_n : et}
   \; \textsc{Process}
@@
 Structural congruence is a congruent equivalence relation on processes.
-Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.
+\autoref{fig:struct-cong-base} lists the base cases of structural congruence.
 The type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear nor in the inputs nor in the outputs of $Q$.
@@
   { }
-  {P \eqeq Q : \Set}
+  {P \eqeq Q : et}
   \; \textsc{StructCong}
@@
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{fig:struct-cong}.
-In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown in \autoref{fig:struct-cong}.
+In  \autoref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
 In the transitivity case, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
@@
   { }
-  {\Rec : \Set}
+  {\Rec : et}
   \; \textsc{Rec}
@@
   {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq{r} Q : \Set}
+  {P \eq{r} Q : et}
   \; \textsc{Equals}
@@
 
-Figure \ref{fig:reduction} models the operational semantics of the \picalc{}.
+\autoref{fig:reduction} models the operational semantics of the \picalc{}.
 Processes put in parallel reduce when they communicate over a common channel.
@@
 --- notably, not under input nor output, or otherwise the sequencing of actions would not be preserved.
-In \S \ref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
+In  \autoref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
 However, this reduction operation is effectful: it consumes the channel over which communication occurs.
@@
   {n : \N}
-  {\Channel_n : \Set}
+  {\Channel_n : et}
   \; \textsc{Channel}
@@
   {i : \Channel_n \\ P \; Q : \Process_n}
-  {P \reduce{i} Q : \Set}
+  {P \reduce{i} Q : et}
   \; \textsc{Reduces}
@@
 A resource-aware typing system for the \picalc{} must take resource usages into account, and needs an algebra on these --- a way of \emph{splitting} resources on parallel composition, of \emph{ticking off} resources on input and output.
-We characterise such an algebra in \S \ref{multiplicities}.
-We define independent typing contexts and usage contexts in \S \ref{contexts}.
-We provide a typing judgment for a resource-aware \picalc{} in \S \ref{leftover-typing}.
+We characterise such an algebra in  \autoref{multiplicities}.
+We define independent typing contexts and usage contexts in  \autoref{contexts}.
+We provide a typing judgment for a resource-aware \picalc{} in  \autoref{leftover-typing}.
 
@@
   It also requires a special element that is to be used to count inputs and outputs.
-  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in Figure \ref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
+  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
 
@@
     &\one                   &:{} &                      &        & C \\
-    &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
+    &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to et \\
     &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
@@
 
-Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in Figure \ref{fig:multiplicities}.
-For examples of how these multiplicity types are used in typing derivations, refer to Example \ref{example-derivations}.
+Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in \autoref{fig:multiplicities}.
+For examples of how these multiplicity types are used in typing derivations, refer to Example \autoref{example-derivations}.
 
@@
 \begin{nidefinition}
-  We define types (Figure \ref{fig:prectx}) as either unit types, base types or channel types.
+  We define types (\autoref{fig:prectx}) as either unit types, base types or channel types.
   The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
@@
     { }
-    {\Type : \Set}
+    {\Type : et}
     \; \textsc{Type}
@@
     {n : \N}
-    {\PreCtx_n : \Set}
+    {\PreCtx_n : et}
     \; \textsc{PreCtx}
@@
   We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by Figure \ref{fig:indexed-multiplicities}.
+  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by \autoref{fig:indexed-multiplicities}.
   We require the index type to be inhabited -- proving subject congruence on $\new \PO_{\suc n} \eqeq \PO_n$ requires an inhabited carrier set.
@@
   \begin{split}
-    &\field{IDX}          &:{} &\Set \\
+    &\field{IDX}          &:{} &et \\
     &\field{\exists IDX}  &:{} &\field{IDX} \\
-    &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
+    &\field{CARRIER}      &:{} &\field{IDX} \to et \\
     &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
@@
 \begin{nidefinition}
-  A context of usages $\Ctx_{idxs}$ is indexed by a length-indexed context of indices $\Idxs_n$, as shown in Figure \ref{fig:ctx}.
+  A context of usages $\Ctx_{idxs}$ is indexed by a length-indexed context of indices $\Idxs_n$, as shown in \autoref{fig:ctx}.
   \begin{figure}[h]
@@
     {n : \N}
-    {\Idxs_n : \Set}
+    {\Idxs_n : et}
     \; \textsc{Idxs}
@@
     {idxs : \Idxs_n}
-    {\Ctx_{idxs} : \Set}
+    {\Ctx_{idxs} : et}
     \; \textsc{Ctx}
@@
   \item
-    Theorems like framing (\S \ref{framing}), weakening (\S \ref{weakening}) and strengthening (\S \ref{strengthening}) become suddenly stateable.
+    Theorems like framing ( \autoref{framing}), weakening ( \autoref{weakening}) and strengthening ( \autoref{strengthening}) become suddenly stateable.
 \end{itemize}
 
-We introduce two judgments: one for variable references (Definition \ref{def:varref}), one for processes (Definition \ref{def:types}).
+We introduce two judgments: one for variable references (\autoref{def:varref}), one for processes (\autoref{def:types}).
 Both of them combine a judgment on types and a judgment on usages.
@@
 
-  The \textsc{VarRef} relation presented in Figure \ref{fig:variable-references} is indexed by a typing context and an input and output usage context.
+  The \textsc{VarRef} relation presented in \autoref{fig:variable-references} is indexed by a typing context and an input and output usage context.
   In addition to these, it also keeps track of its position in the contexts, its type, and its usage annotation.
@@
       \Delta : \Ctx_{idxs}}
-    {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : \Set}
+    {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : et}
     \; \textsc{VarRef}
@@
 
-  The typing relation \textsc{Types} presented in Figure \ref{fig:types} is indexed by a typing context, an input and an output usage context, and the process $P$ being typed.
+  The typing relation \textsc{Types} presented in \autoref{fig:types} is indexed by a typing context, an input and an output usage context, and the process $P$ being typed.
 
@@
       \Delta : \Ctx_{idxs}}
-    {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}
+    {\types{\gamma}{\Gamma}{P}{\Delta} : et}
     \; \textsc{Types}
@@
   Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
-  Then, per Lemma \ref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Then, per \autoref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
   Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
@@
     \item
-      The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
+      The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\autoref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
       The usage context on which $Q$ and $P$ coincide in $\comp{Q}{P}$ will no longer be the same.
@@
     \item
-      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
-      We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
+      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of $\autoref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
+      We then use strengthening (\autoref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
       In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
-      We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
-    \item
-      Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
-      In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
+      We will then use \autoref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
+    \item
+      Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \autoref{thm:swapping}.
+      In the reverse direction, we will use swapping and then use \autoref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
   \end{itemize}
@@
   We need to proceed by induction on $P$, and to do so means going under scope restriction.
-  We use the generalised version in Theorem \ref{thm:subst-generalization} so that the variable that is being substituted is not $\constr{zero}$ but any variable $i$.
+  We use the generalised version in \autoref{thm:subst-generalization} so that the variable that is being substituted is not $\constr{zero}$ but any variable $i$.
   \todo{explain arguments}
@@
   Let the variable that is being substituted be $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$ and the variable that it is being substituted with $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$.
-  As shown in Figure \ref{fig:subst}, the process that is being substituted will pivot from the input usage context $\Gamma_i$ to the input usage context $\Gamma_j$.
+  As shown in \autoref{fig:subst}, the process that is being substituted will pivot from the input usage context $\Gamma_i$ to the input usage context $\Gamma_j$.
   Let there be two \emph{adjustment} variables $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$ and $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$.
   The process that is being substituted will pivot from the output usage context $\Psi_i$ to the output usage context $\Psi_j$.
-  As shown in Figure \ref{fig:subst-alternative}, without this \emph{adjustment} composing the substitution of subprocesses is not at all trivial.
+  As shown in \autoref{fig:subst-alternative}, without this \emph{adjustment} composing the substitution of subprocesses is not at all trivial.
   Assume that $P$ uses all $m$ multiplicities at index $i$.
@@
 In our case, and because we generalize linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity $\one$ on both its input and output capabilities.
-In \S \ref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
+In  \autoref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
 
@@
     \item
-    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower_0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in Figure \ref{fig:subject-reduction:a}.
-    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (Theorem \ref{thm:substitution}) and strengthening (Theorem \ref{thm:strengthening}).
+    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower_0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in \autoref{fig:subject-reduction:a}.
+    We start by using framing (\autoref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
 
@@
     If the channel is $\constr{external}\; \constr{zero}$ -- i.e.
-the channel that is being introduced by the scope restriction -- we use the Lemma \ref{lm:comm-capable} to subtract $\lio$ from the channel's multiplicity first, then apply the induction hypothesis.
+the channel that is being introduced by the scope restriction -- we use the \autoref{lm:comm-capable} to subtract $\lio$ from the channel's multiplicity first, then apply the induction hypothesis.
     If the channel is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
@@
     \item
-    Finally, in the $\constr{struct}$ case, we apply subject congruence (Theorem \ref{thm:subject-congruence}) and the induction hypothesis.
+    Finally, in the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and the induction hypothesis.
   \end{itemize}

### 2020-05-01 paper/paper.tex
Adapt the syntax example

@@
 \newcommand{\new}{\constr{\boldsymbol{\nu}} \;}
-\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}\newcommand{\sendp}[2]{#1 \constr{\langle} #2 \constr{\rangle}}
+\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
+\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle} \;}
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
-\newcommand{\recvp}[2]{#1 \constr{(}#2\constr{)}}
+\newcommand{\recvp}[2]{#1 \; \constr{(}#2\constr{)} \;}
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
 Instead we adopt de Bruijn indices \cite{} where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
-\todo{rewrite the sytanx below, following the syntax above}
 \begin{example}
 \hfill{}\\
-The \picalc{} term \hfill{} $(\textbf{new} \; x) (x ( y ) .
-y \langle \texttt{"Hello world!"} \rangle .
-\mathbf{0} \parallel (\textbf{new} \; y) (x \langle y \rangle .
-y ( z ) .\mathbf{0}))$ \\ is represented as \hfill{} $(\textbf{new}) (0 () .
-0 \langle \texttt{"Hello world!"} \rangle .
-\mathbf{0} \parallel (\textbf{new}) (1 \langle 0 \rangle .
-0 () .
-\mathbf{0}))$.
+The \picalc{} term \hfill{} $(\new{}x ) (\comp {\recvp{x}{y} .
+\; \sendp{y}{\texttt{"Hello world!"}} .
+\; \PO} {(\new{} y) (\sendp{x}{y} .
+\; \recvp{y}{z} .
+\; \PO)})$ \\ is represented as \hfill{} $\new{} (\comp {\recv{0} .
+\; \send{0}{\texttt{"Hello world!"}} .
+\; \PO} {\new{} (\send{1}{0} .
+\; \recv{0} .
+\; \PO)})$.
 \end{example}
@@
 It is useful to rule out these nonsensical terms syntactically.
-\todo{replace ref with autoref everywhere...}
 In  \autoref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.

### 2020-05-01 paper/paper.tex
structural congruence

--- 2020-05-01 paper/paper.tex
@@
 The syntax of the \picalc{} \cite{Sangio01} is given by the following grammar.
+\todo{maybe put both syntaxes in two columns to save space}
 \begin{equation*}
@@
 In order to make the de Bruijn indices representation more human-readable, we can annotate these indices with the original channel names.
-Machines can then manipulate references mechanically whilst using channel names to present \picalc{} terms back to humans.
+Machines can then manipulate references mechanically whilst using channel names to present \picalc{} processes back to humans.
 
@@
 \label{semantics}
-
-In the \lambdacalc{} $\beta$-reduction operates on syntactically adjacent terms.
-In the \picalc{} however, the syntax introduces unnecessary distinctions
-(e.g. semantically, parallel composition is defined modulo associativity and commutativity).
-There are several ways around this, a structural congruence relation being one of the historical ones \cite{}.
-(Others include labeled transition systems \cite{} and higher inductive types \cite{}.)
+We introduce the operational semantics for the \picalc{}, starting with structural congruence.
 
@@
 
-Structural congruence is a congruent equivalence relation on processes.
-Figure \ref{fig:struct-cong-base} lists the base cases of structural congruence.
-The type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear nor in the inputs nor in the outputs of $Q$.
-The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
-Finally, $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
-In all three, $i$ is incremented every time we go under a binder.
-
+The complete structural congruence is given in \autoref{app:struct}.
+Figure \ref{fig:struct-cong-base} lists the base cases.
+
+\todo{remove space between inferrules}
 \begin{figure}[h]
@@
 
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:swap-swap}
-  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
+In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
+The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+In $\constr{scope-comm}$ function $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In all the above, $i$ is incremented every time we go under a binder.
 
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown in Figure \ref{fig:struct-cong}.
-In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
-In the transitivity case, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
-To do so, we need to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
-To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of Figure \ref{fig:struct-cong}.
+The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
+%In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
 
@@
 \begin{mathpar}
-  \datatype
-  { }
-  {\Rec : \Set}
-  \; \textsc{Rec}
-
-  \inferrule
-  { }
-  {\constr{zero} : \Rec}
-
-  \inferrule
-  {r : \Rec}
-  {\constr{one} \; r : \Rec}
-
-  \inferrule
-  {r \; s : \Rec}
-  {\constr{two} \; r \; s : \Rec}
-
-  \datatype
-  {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq{r} Q : \Set}
-  \; \textsc{Equals}
-
-  \inferrule
-  {eq : P \eqeq Q}
-  {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
+  \inferrule
+  { }
+  {\constr{refl} : P \eq{\constr{zero}} P}
+
+  \inferrule
+  {eq : P \eq{r} Q}
+  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
+
+  \inferrule
+  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
+  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
 
@@
   {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
-
-  \inferrule
-  { }
-  {\constr{refl} : P \eq{\constr{zero}} P}
-
-  \inferrule
-  {eq : P \eq{r} Q}
-  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
-
-  \inferrule
-  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
-  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
 \end{mathpar}
-\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
+\caption{Structural congruence under contexts.}
 \label{fig:struct-cong}
@@
 \bibliography{paper}
+
+%%
+\newpage
+\appendix
+\input{appendix.tex}
 \end{document}

### 2020-05-01 paper/appendix.tex
Create appendix.tex

%% Appendix
\section{Structural Congruence}
\label{app:struct}
\begin{nilemma}
  \label{lm:lower-lift}
  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{nilemma}
  \label{lm:swap-swap}
  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
\end{nilemma}
\begin{proof}
  By structural induction on \textsc{Process} and \textsc{Var}.
\end{proof}

\begin{figure}[h]
  \begin{mathpar}
    \datatype
    { }
    {\Rec : \Set}
    \; \textsc{Rec}

    \inferrule
    { }
    {\constr{zero} : \Rec}

    \inferrule
    {r : \Rec}
    {\constr{one} \; r : \Rec}

    \inferrule
    {r \; s : \Rec}
    {\constr{two} \; r \; s : \Rec}

    \datatype
    {P \, Q : \Process_n \\ r : \Rec}
    {P \eq{r} Q : \Set}
    \; \textsc{Equals}

    \inferrule
    {eq : P \eqeq Q}
    {\constr{struct} \; eq : P \eq{\constr{zero}} Q}

    \inferrule
    {eq : P \eq{r} P'}
    {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}

    \inferrule
    {eq : P \eq{r} P'}
    {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}

    \inferrule
    {eq : P \eq{r} P'}
    {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}

    \inferrule
    {eq : P \eq{r} P'}
    {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}

    \inferrule
    { }
    {\constr{refl} : P \eq{\constr{zero}} P}

    \inferrule
    {eq : P \eq{r} Q}
    {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}

    \inferrule
    {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
    {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
  \end{mathpar}
  \caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
  \label{fig:struct-cong1}
  \end{figure}

In the transitivity rule, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
To do so, we need to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.

### 2020-05-01 paper/paper.tex
Fix search and replace mistake

--- 2020-05-01 paper/paper.tex
@@
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
-\newcommand{\recvp}[2]{#1 \; \constr{(}#2\constr{)} \;}
+\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)} \;}
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
     { }
-    {\type{\N} : et}
+    {\type{\N} : \Set}
     \; \textsc{Nat}
@@
 We otherwise use the constructor name as the name of a typing rule.
-Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $et$.
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
 Implicit arguments are mentioned by type definitions but omitted in constructors.
@@
                 |& \; \comp{P}{Q}  & \text{(parallel composition)} \\
-                |& \; \recvp{x}{y}.P  & \text{(input)} \\
-                |& \; \sendp{x}{y}.P   &\text{(output)}
+                |& \; \recvp{x}{y}. \; P  & \text{(input)} \\
+                |& \; \sendp{x}{y}. \; P   &\text{(output)}
   \end{split}
@@
   {n : \N}
-  {\Var_n : et}
+  {\Var_n : \Set}
   \; \textsc{Var}
@@
   {n : \N}
-  {\Process_n : et}
+  {\Process_n : \Set}
   \; \textsc{Process}
@@
   { }
-  {P \eqeq Q : et}
+  {P \eqeq Q : \Set}
   \; \textsc{StructCong}
@@
   { }
-  {\Rec : et}
+  {\Rec : \Set}
   \; \textsc{Rec}
@@
   {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq{r} Q : et}
+  {P \eq{r} Q : \Set}
   \; \textsc{Equals}
@@
   {n : \N}
-  {\Channel_n : et}
+  {\Channel_n : \Set}
   \; \textsc{Channel}
@@
   {i : \Channel_n \\ P \; Q : \Process_n}
-  {P \reduce{i} Q : et}
+  {P \reduce{i} Q : \Set}
   \; \textsc{Reduces}
@@
     { }
-    {\Type : et}
+    {\Type : \Set}
     \; \textsc{Type}
@@
     {n : \N}
-    {\PreCtx_n : et}
+    {\PreCtx_n : \Set}
     \; \textsc{PreCtx}
@@
     {n : \N}
-    {\Idxs_n : et}
+    {\Idxs_n : \Set}
     \; \textsc{Idxs}
@@
     {idxs : \Idxs_n}
-    {\Ctx_{idxs} : et}
+    {\Ctx_{idxs} : \Set}
     \; \textsc{Ctx}
@@
       \Delta : \Ctx_{idxs}}
-    {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : et}
+    {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : \Set}
     \; \textsc{VarRef}
@@
       \Delta : \Ctx_{idxs}}
-    {\types{\gamma}{\Gamma}{P}{\Delta} : et}
+    {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}
     \; \textsc{Types}

### 2020-05-01 paper/paper.tex

--- 2020-05-01 paper/paper.tex
@@
 
-Structural congruence is a congruent equivalence relation on processes.
-\autoref{fig:struct-cong-base} lists the base cases of structural congruence.
-The type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear nor in the inputs nor in the outputs of $Q$.
-The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
-Finally, $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
-In all three, $i$ is incremented every time we go under a binder.
-
+The complete structural congruence is given in \autoref{app:struct}.
+Figure \ref{fig:struct-cong-base} lists the base cases.
+
+\todo{remove space between inferrules}
 \begin{figure}[h]
@@
 
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:swap-swap}
-  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
+In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
+The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+In $\constr{scope-comm}$ function $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In all the above, $i$ is incremented every time we go under a binder.
 
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown in \autoref{fig:struct-cong}.
-In  \autoref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
-In the transitivity case, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
-To do so, we need to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
-To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of Figure \ref{fig:struct-cong}.
+The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
+%In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
 
@@
 \begin{mathpar}
-  \datatype
-  { }
-  {\Rec : et}
-  \; \textsc{Rec}
-
-  \inferrule
-  { }
-  {\constr{zero} : \Rec}
-
-  \inferrule
-  {r : \Rec}
-  {\constr{one} \; r : \Rec}
-
-  \inferrule
-  {r \; s : \Rec}
-  {\constr{two} \; r \; s : \Rec}
-
-  \datatype
-  {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq{r} Q : et}
-  \; \textsc{Equals}
-
-  \inferrule
-  {eq : P \eqeq Q}
-  {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
+  \inferrule
+  { }
+  {\constr{refl} : P \eq{\constr{zero}} P}
+
+  \inferrule
+  {eq : P \eq{r} Q}
+  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
+
+  \inferrule
+  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
+  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
 

### 2020-05-01 paper/paper.tex
Update paper.tex

@@
 The complete structural congruence is given in \autoref{app:struct}.
-Figure \ref{fig:struct-cong-base} lists the base cases.
+\autoref{fig:struct-cong-base} lists the base cases.
 
@@
 Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of Figure \ref{fig:struct-cong}.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of \autoref{fig:struct-cong}.
 The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.

### 2020-05-01 paper/appendix.tex
more on polishing struct

@@
 \label{app:struct}
+
 \begin{nilemma}
@@
 \end{proof}
+
+Structural congruence is a congruent equivalence relation.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of \autoref{fig:struct-cong1}.
+The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
 
@@
 To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
+
+\section{Operational Semantics}
+\label{app:os}
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    {n : \N}
+    {\Channel_n : \Set}
+    \; \textsc{Channel}
+
+    \inferrule
+    { }
+    {\constr{internal} : \Channel_n}
+
+    \inferrule
+    {i : \Var_n}
+    {\constr{external} \; i : \Channel_n}
+
+    \datatype
+    {i : \Channel_n \\ P \; Q : \Process_n}
+    {P \reduce{i} Q : \Set}
+    \; \textsc{Reduces}
+
+    \inferrule
+    {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
+    {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower_0 \; (\subst{P}{j}{0})}{Q}}
+
+    \inferrule
+    {red : P \reduce{i} P'}
+    {\constr{par} \; red : \comp{P}{Q} \reduce{i} \comp{P'}{Q}}
+
+    \inferrule
+    {red : P \reduce{i} Q}
+    {\constr{res} \; red : \new P \reduce{dec\; i} \new Q}
+
+    \inferrule
+    {eq : P = P' \\ red : P' \reduce{i} Q}
+    {\constr{struct} \; eq \; red : P \reduce{i} Q}
+  \end{mathpar}
+  \caption{Operational semantics indexed by the channel over which reduction occurs.}
+  \label{fig:reduction1}
+  \end{figure}

### 2020-05-01 paper/paper.tex
more on polishing struct

@@
 In all the above, $i$ is incremented every time we go under a binder.
-
-Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of \autoref{fig:struct-cong}.
-The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
-%In \S \ref{subject-congruence} we prove that if two processes $P$ and $Q$ are structurally congruent and $P$ is well-typed, then so is $Q$.
-
-\begin{figure}[h]
-\begin{mathpar}
-  \inferrule
-  { }
-  {\constr{refl} : P \eq{\constr{zero}} P}
-
-  \inferrule
-  {eq : P \eq{r} Q}
-  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
-
-  \inferrule
-  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
-  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
-\end{mathpar}
-\caption{Structural congruence under contexts.}
-\label{fig:struct-cong}
-\end{figure}
 

### 2020-05-01 paper/paper.tex
pass on os

@@
 
-\autoref{fig:reduction} models the operational semantics of the \picalc{}.
-Processes put in parallel reduce when they communicate over a common channel.
-The receiving process substitutes references to its most immediate variable with references to the variable that is being sent by the process doing the output.
-The process is then \emph{lowered} --- all variable references are decreased by one.
-Reduction is closed under structural congruence and goes under parallel composition and scope restriction
---- notably, not under input nor output, or otherwise the sequencing of actions would not be preserved.
-In  \autoref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
-However, this reduction operation is effectful: it consumes the channel over which communication occurs.
-If this channel is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
+The operational semantics of the \picalc{} is defined via a reduction relation in \autoref{fig:reduction}.
+In rule $\constr{comm}$, processes in parallel reduce when they communicate over the same index ${i}$.
+The input process $\recv{i}P$ substitutes in $P$ references to the most immediate variable placeholder with references to the variable sent by the output process $\send{i}{j}{Q}$.
+The continuation $\subst{P}{j}{0}$ is then \emph{lowered} --- all variable references are decreased by one.
+Reduction is closed under parallel composition (rule $\constr{par}$), scope restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
+--- notably, not under input nor output, otherwise the sequencing of actions would not be preserved \cite{Sangio01}.
+%In  \autoref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
+
+\todo{this paragraph below is not clear}
+Notice that the reduction relation in \autoref{fig:reduction} is effectful: it consumes the channel over which communication occurs.
+If the channel is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
 To keep track of this information, we lift the index of the channel to the type level.

### 2020-05-01 paper/paper.tex
Update paper.tex

@@
 
-The complete structural congruence is given in \autoref{app:struct}.
+Structural congruence $\eqeq$ is fully defined in \autoref{app:struct}.
 \autoref{fig:struct-cong-base} lists the base cases.
@@
 
+The first three rules on $\constr{comp}$ state respectively, associativity, symmetry, and $\PO$ as the neutral element for parallel composition.
+The last three rules on $\constr{scope}$ state respectively, garbage collection, scope extrusion and commutativity of restrictions.
+
 In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
@@
 
-\subsection{Operational Semantics}
+\subsection{Reduction Relation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The operational semantics of the \picalc{} is defined via a reduction relation in \autoref{fig:reduction}.
+The operational semantics of the \picalc{} is defined via a reduction relation $\reduce{}$ in \autoref{fig:reduction}.
 In rule $\constr{comm}$, processes in parallel reduce when they communicate over the same index ${i}$.
@@
 The continuation $\subst{P}{j}{0}$ is then \emph{lowered} --- all variable references are decreased by one.
-Reduction is closed under parallel composition (rule $\constr{par}$), scope restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
---- notably, not under input nor output, otherwise the sequencing of actions would not be preserved \cite{Sangio01}.
+Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
+--- notably, not under input or output, otherwise the sequencing of actions would not be preserved \cite{Sangio01}.
 %In  \autoref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
@@
 \todo{this paragraph below is not clear}
-Notice that the reduction relation in \autoref{fig:reduction} is effectful: it consumes the channel over which communication occurs.
-If the channel is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
+Notice that reduction is effectful: it consumes the channel over which communication occurs.
+If the channel is $\constr{external}$ to $P$ (it resides in its typing context) then the context in which $Q$ is typed must change.
 To keep track of this information, we lift the index of the channel to the type level.

### 2020-05-01 paper/paper.tex
on types

@@
 
-Typing systems for the \picalc{} ensure that processes make a correct use of their channels.
-Typing judgments are of the form $\Gamma \vdash P$ (for some typing context $\Gamma$ and some process $P$) --- types only appear in the contexts of typing judgments.
+A typing context $\Gamma$ is a partial function from channel names to types.
+A typing judgment is of the form $\Gamma \vdash P$, reading ``process $P$ is well typed under the typing context $\Gamma$''.
 A resource-aware typing system for the \picalc{} must take resource usages into account, and needs an algebra on these --- a way of \emph{splitting} resources on parallel composition, of \emph{ticking off} resources on input and output.
@@
 We define independent typing contexts and usage contexts in  \autoref{contexts}.
-We provide a typing judgment for a resource-aware \picalc{} in  \autoref{leftover-typing}.
+We provide a typing system with leftovers for a resource-aware \picalc{} in  \autoref{leftover-typing}.
 
@@
 
-In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-In this paper we generalize the type of multiplicities by defining an algebra any such type must comply with.
-Shared, gradual and linear usage types all satisfy these algebraic laws.
+In the linear \picalc{} each channel has an input or an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
+We generalise multiplicities by defining an algebra such that linear, gradual and shared types all satisfy these algebraic laws.
 
@@
   We will often work with input and output multiplicities as pairs.
-  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple, and $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
-  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$.
+  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple, and $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
+  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
 \end{note}
 
-Shared, gradual and linear multiplicity types can all be defined as an instance of the properties in \autoref{fig:multiplicities}.
-For examples of how these multiplicity types are used in typing derivations, refer to Example \autoref{example-derivations}.
+\todo{note I've changed the order of presentation from linear to shared, check if it breaks anything!}
+Linear, gradual and shared types are all defined as an instance of the laws in \autoref{fig:multiplicities}.
+We illustrate these types in typing derivations in \autoref{example-derivations}.
 
 \begin{description}
+  \item [Linear]
+    The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
+    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
+    All other rules follow trivially.
+
+    \item [Gradual]
+    The carrier $C$ is implemented as the type of natural numbers $\N$.
+    The element $\zero$ corresponds to 0, and $\one$ to 1.
+    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
+    All other rules follow trivially from the algebraic rules for the addition of natural numbers.
+
     \item [Shared]
@@
     The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
-    All other laws are trivially satisfied.
-  \item [Gradual]
-    The carrier $C$ is implemented as the type of natural numbers $\N$.
-    The element $\zero$ corresponds to 0, and $\one$ to 1.
-    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
-    All other rules follow trivially from the algebraic rules for the addition of naturals.
-  \item [Linear]
-    The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
-    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
     All other rules follow trivially.
@@
 
-\subsection{Contexts}
+\subsection{Typing Contexts}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{contexts}
-
-In a resource-aware \picalc{}, usage annotations are transformed as a result of input and output; types howeverr are preserved.
+\todo{is usage annotation the same as multiplicities? if yes, just use the former as too many notions}
+In a resource-aware \picalc{}, usage annotations are transformed as a result of input and output; types however are preserved.
 In our case, usage annotations do not depend on types, and keeping types and usages in separate contexts allows for a more modular approach in our proofs.
@@
 In the linear \picalc{} however, reduction \emph{is} meant to transform the context: each time communication occurs over a channel, that channel is exhausted.
-In our case, and because we generalize linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity $\one$ on both its input and output capabilities.
+In our case, and because we generalise linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity $\one$ on both its input and output capabilities.
 In  \autoref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.

### 2020-05-01 paper/paper.tex
Add some references to related work

@@
 % -- put emphasis on mechanisation in Agda
+
+\todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
 
@@
         \item
-          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
+          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $\containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
           We can then find $\Xi_j$ and construct the two arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (used for the inductive call).
@@
 
+\cite{Allais2018a}
+typing with leftovers
+bidirectional typing rules
+decidable typechecking and type inference
+limited to a linear type system
+framing, weakening, substitution
+soundness and completeness with respect to a sequent calculus for intuitionistic linear logic
+
+\cite{Thiemann2019}
+\cite{Rouvoet2020}
+\cite{Veltri2020}
+
 \cite{previous-work} polymorphic tokens, HOAS
 
-\cite{typing-with-leftovers}
-
-\cite{Higher-inductive-types-for-congruence}
-
 \cite{LTS-semantics}
-
-\cite{work-on-session-types}
 

### 2020-05-01 paper/paper.tex
Go through the changes, until 4.1

@@
 \label{syntax}
-We let $x, y,\ldots$ range over \emph{channel names} or \emph{variables} and
-$P, Q,\ldots$ over processes.
+Let $x, y,\ldots$ range over \emph{variables} and $P, Q,\ldots$ over processes.
 The syntax of the \picalc{} \cite{Sangio01} is given by the following grammar.
-\todo{maybe put both syntaxes in two columns to save space}
 \begin{equation*}
@@
 Process $\PO$ denotes the terminated process, where no operations can further occur.
-Process $(\new{}x)P$ is the restriction process where channel $x$ is created new and is bound with scope $P$.
+Process $(\new{}x)P$ creates a new channel $x$ bound with scope $P$.
 Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recvp{x}{y}.P$ and $\sendp{x}{y}.P$ denote respectively the input and output processes of a variable/channel $y$ over a channel $x$, with continuation $P$.
+Processes $\recvp{x}{y}.P$ and $\sendp{x}{y}.P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
 
 Scope restriction $(\new{}x)P$ and input $\recvp{x}{y}.P$ both introduce bound names in $P$.
-In order to mechanise the \picalc{} syntax in Agda, we need a way to deal with bound names in these continuation processes.
-Channel names are cumbersome to mechanise: inserting a new variable into a context means proving that the name of such variable is different to all other variable names in context.
-Instead we adopt de Bruijn indices \cite{} where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
+In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
+Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
+Instead, we use de Bruijn indices \cite{}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
 \begin{example}
@@
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-\autoref{fig:syntax} shows the syntax of \picalc{} in Agda. We increase the \emph{depth} counter by 1 in channel restriction and input processes.
+We increase the \emph{depth} counter by 1 in channel restriction and input processes.
+\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction thanks to type-level de Bruijn indices.
 
@@
 
-Using type-level de Bruijn indices makes our syntax well-scoped by construction --- the semantics of our language can now be defined on the totality of the syntax.
-In order to make the de Bruijn indices representation more human-readable, we can annotate these indices with the original channel names.
-Machines can then manipulate references mechanically whilst using channel names to present \picalc{} processes back to humans.
+The semantics of our language can now be defined on the totality of the syntax.
 
@@
 
-Structural congruence $\eqeq$ is fully defined in \autoref{app:struct}.
-\autoref{fig:struct-cong-base} lists the base cases.
+We define the base cases of a structural congruence relation \textsc{StructCong} in \autoref{fig:struct-cong-base}.
+Its congruent equivalence closure \textsc{Equals} is provided in \autoref{app:struct}.
 
@@
   \inferrule
-  {uQ : \Unused_0 \, Q}
+  {uQ : \Unused_0 \; Q}
   {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower_0 \; \; Q \; uQ}}
@@
 
-The first three rules on $\constr{comp}$ state respectively, associativity, symmetry, and $\PO$ as the neutral element for parallel composition.
-The last three rules on $\constr{scope}$ state respectively, garbage collection, scope extrusion and commutativity of restrictions.
-
-In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof that witnesses that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
+The first three rules ($\constr{comp}-$*) state associativity, symmetry, and $\PO$ as the neutral element of parallel composition, respectively.
+The last three ($\constr{scope}-$*) state, garbage collection, scope extrusion and commutativity of restrictions, respectively.
+
+In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
 The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
-In $\constr{scope-comm}$ function $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In $\constr{scope-comm}$ the function $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
 In all the above, $i$ is incremented every time we go under a binder.
@@
 
-The operational semantics of the \picalc{} is defined via a reduction relation $\reduce{}$ in \autoref{fig:reduction}.
-In rule $\constr{comm}$, processes in parallel reduce when they communicate over the same index ${i}$.
-The input process $\recv{i}P$ substitutes in $P$ references to the most immediate variable placeholder with references to the variable sent by the output process $\send{i}{j}{Q}$.
-The continuation $\subst{P}{j}{0}$ is then \emph{lowered} --- all variable references are decreased by one.
+The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} indexed by the channel $i$ on which communication occurs (\autoref{fig:reduction}).
+In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel ${i}$.
+As a result of that communication, the continuation of the input process $P$ gets all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
+After substitution, $\subst{P}{j}{0}$ is \emph{lowered} --- all variable references are decreased by one.
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
 --- notably, not under input or output, otherwise the sequencing of actions would not be preserved \cite{Sangio01}.
-%In  \autoref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
-
-\todo{this paragraph below is not clear}
-Notice that reduction is effectful: it consumes the channel over which communication occurs.
-If the channel is $\constr{external}$ to $P$ (it resides in its typing context) then the context in which $Q$ is typed must change.
-To keep track of this information, we lift the index of the channel to the type level.
-We use a $dec$ function to decrement the index of the channel every time we come out of a binder.
+
+We keep track of the channel over which communication occurs so that we can state subject reduction.
+We use a $dec$ function to decrement the index of this channel every time we come out of a binder.
 This function saturates at $\constr{internal}$ --- meaning communication happens internally to the process.
@@
 
-A typing context $\Gamma$ is a partial function from channel names to types.
-A typing judgment is of the form $\Gamma \vdash P$, reading ``process $P$ is well typed under the typing context $\Gamma$''.
 A resource-aware typing system for the \picalc{} must take resource usages into account, and needs an algebra on these --- a way of \emph{splitting} resources on parallel composition, of \emph{ticking off} resources on input and output.
@@
 
-\todo{note I've changed the order of presentation from linear to shared, check if it breaks anything!}
 Linear, gradual and shared types are all defined as an instance of the laws in \autoref{fig:multiplicities}.
@@
 \label{contexts}
-\todo{is usage annotation the same as multiplicities? if yes, just use the former as too many notions}
 In a resource-aware \picalc{}, usage annotations are transformed as a result of input and output; types however are preserved.
@@
 \begin{note}
-  We lift the monoidal operation on multiplicity pairs $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on contexts of usages with the same underlying context of indices.
+  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with a common underlying context of indices.
   All the properties of the partial monoid are trivially lifted.

### 2020-05-01 paper/paper.tex
Show named and de Bruijn indexed processes side by side

@@
 Let $x, y,\ldots$ range over \emph{variables} and $P, Q,\ldots$ over processes.
-The syntax of the \picalc{} \cite{Sangio01} is given by the following grammar.
-\begin{equation*}
-  \begin{split}
-    P,Q ::=      & \; \PO       &\text{(terminated process)}  \\
-                |& \; (\new{}x)  P   & \text{(restriction)} \\
-                |& \; \comp{P}{Q}  & \text{(parallel composition)} \\
-                |& \; \recvp{x}{y}. \; P  & \text{(input)} \\
-                |& \; \sendp{x}{y}. \; P   &\text{(output)}
-  \end{split}
-\end{equation*}
-Process $\PO$ denotes the terminated process, where no operations can further occur.
-Process $(\new{}x)P$ creates a new channel $x$ bound with scope $P$.
-Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recvp{x}{y}.P$ and $\sendp{x}{y}.P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-
-Scope restriction $(\new{}x)P$ and input $\recvp{x}{y}.P$ both introduce bound names in $P$.
-In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
-Instead, we use de Bruijn indices \cite{}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
-\begin{example}
-\hfill{}\\
-The \picalc{} term \hfill{} $(\new{}x ) (\comp {\recvp{x}{y} .
-\; \sendp{y}{\texttt{"Hello world!"}} .
-\; \PO} {(\new{} y) (\sendp{x}{y} .
-\; \recvp{y}{z} .
-\; \PO)})$ \\ is represented as \hfill{} $\new{} (\comp {\recv{0} .
-\; \send{0}{\texttt{"Hello world!"}} .
-\; \PO} {\new{} (\send{1}{0} .
-\; \recv{0} .
-\; \PO)})$.
-\end{example}
-That is, terms at different \emph{depths} use different indices to refer to the same binding.
-
-A variable occurring under $n$ abstractions has $n$ possible variables to refer to.
-References outside of that range are meaningless.
-It is useful to rule out these nonsensical terms syntactically.
-In  \autoref{fig:var} we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+The syntax of the \picalc{} \cite{Sangio01} is given by grammar in \autoref{fig:syntax-names}.
 
 \begin{figure}[h]
+  \begin{subfigure}{1\textwidth}
     \begin{mathpar}
@@
       {\suc x : \Var_{\suc n}}
-\end{mathpar}
-\caption{Types of size $n$}
-\label{fig:var}
-\end{figure}
-
-Every time we go under a binder, the number of binders a variable might refer to increments by one.
-To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-We increase the \emph{depth} counter by 1 in channel restriction and input processes.
-\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction thanks to type-level de Bruijn indices.
-
-\begin{figure}[h]
-\begin{mathpar}
+
       \datatype
@@
     \end{mathpar}
-
+  \end{subfigure}
+  \begin{subfigure}[t]{.2\textwidth}
     \begin{equation*}
       \begin{split}
-  \Process_n ::=& \; \PO_n                    \\
-              |& \; \new{} \Process_{\suc n}          \\
-              |& \; \comp{\Process_n}{\Process_n}          \\
-              |& \; \recv{\Var_n}{}\Process_{\suc n} \\
-              |& \; \send{\Var_n}{\Var_n}\Process_n
+        P,Q ::=      & \; \PO                \\ 
+                    |& \; (\new{}x)  P       \\ 
+                    |& \; \comp{P}{Q}        \\ 
+                    |& \; \recvp{x}{y}. \; P \\ 
+                    |& \; \sendp{x}{y}. \; P 
       \end{split}
     \end{equation*}
-\caption{Well-scoped grammar for \picalc{} using de Bruijn indices}
+    \caption{Ill-scoped grammar with names}
+    \label{fig:syntax-names}
+  \end{subfigure}
+  \hspace{.1\textwidth}
+  \begin{subfigure}[t]{.7\textwidth}
+    \begin{equation*}
+      \begin{split}
+        \Process_n ::=& \; \PO_n               &&\text{(process termination)}   \\ 
+        |& \; \new{} \Process_{\suc n}         &&\text{(restriction)}          \\ 
+        |& \; \comp{\Process_n}{\Process_n}    &&\text{(parallel composition)} \\ 
+        |& \; \recv{\Var_n}{}\Process_{\suc n} &&\text{(input)}                \\ 
+        |& \; \send{\Var_n}{\Var_n}\Process_n  &&\text{(output)}                  
+      \end{split}
+    \end{equation*}
+    \caption{Well-scoped grammar with type-level de Bruijn indices}
     \label{fig:syntax}
+  \end{subfigure}
+  \caption{}
 \end{figure}
+
+Process $\PO$ denotes the terminated process, where no operations can further occur.
+Process $(\new{}x)P$ creates a new channel $x$ bound with scope $P$.
+Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
+Processes $\recvp{x}{y}. \; P$ and $\sendp{x}{y}. \; P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+
+Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y}. \; P$ both introduce bound names in $P$.
+In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
+Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
+Furthermore, names do not guarantee that the syntax well-scoped.
+Instead, we use de Bruijn indices \cite{}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
+\begin{example}
+  \begin{equation*}
+    \begin{split}
+      & (\new{}x ) && (\comp {\recvp{x}{y} && . \; \sendp{y}{\texttt{"Hello world!"}} . \; \PO} {(\new{} y) && (\sendp{x}{y} . \; \recvp{y}{z} && . \; \PO)})
+      & \text{names}
+      \\
+      & \new{} && (\comp {\recv{0} && . \; \send{0}{\texttt{"Hello world!"}} . \; \PO} {\new{} && (\send{1}{0} . \; \recv{0} && . \; \PO)})
+      & \text{de Bruijn}
+      \\
+    \end{split}
+  \end{equation*}
+\end{example}
+That is, terms at different \emph{depths} use different indices to refer to the same binding.
+
+A variable occurring under $n$ abstractions has $n$ possible variables to refer to.
+References outside of that range are meaningless.
+It is useful to rule out these nonsensical terms syntactically.
+In \autoref{fig:syntax}, we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+
+Every time we go under a binder, the number of binders a variable might refer to increments by one.
+To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+We increase the \emph{depth} counter by 1 in channel restriction and input processes.
+\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction thanks to type-level de Bruijn indices.
 

### 2020-05-01 paper/appendix.tex
Polish and shorten the section on multiplicities

--- 2020-05-01 paper/appendix.tex
@@
 \label{app:struct}
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
 
-\begin{nilemma}
-  \label{lm:swap-swap}
-  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
+Structural congruence is a congruent equivalence relation.
+As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of \autoref{fig:struct-cong1}.
+The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
 
@@
 To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
+
+\section{Usage Algebra}
+\label{app:usage-algebra}
+
+\todo{we haven't introduced the forall and implicit syntax}
+\begin{nidefinition}
+  The algebra on multiplicities must be \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}, and needs a \emph{minimal neutral element}.
+  It also requires a special element $\one$ that is to be used to count inputs and outputs.
+  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
+
+  \begin{figure}[h]
+  \begin{equation}
+  \begin{aligned}
+    &\zero                  &:{} &                      &        & C \\
+    &\one                   &:{} &                      &        & C \\
+    &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to et \\
+    &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+    &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+    &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+    &\field{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+    &\field{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
+    &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+    &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+  \end{aligned}
+  \end{equation}
+  \caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid.}
+  \label{fig:multiplicities}
+  \end{figure}
+\end{nidefinition}
+
+\section{Lemmas for Type Safety}
+\label{app:lemmas-type-safety}
+
+\begin{nilemma}
+  \label{lm:lower-lift}
+  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:swap-swap}
+  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}

### 2020-05-01 paper/paper.tex
Polish and shorten the section on multiplicities

@@
 We increase the \emph{depth} counter by 1 in channel restriction and input processes.
-\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction thanks to type-level de Bruijn indices.
+\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction using type-level de Bruijn indices.
 
@@
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
---- notably, not under input or output, otherwise the sequencing of actions would not be preserved \cite{Sangio01}.
+--- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
 
@@
 
-A resource-aware typing system for the \picalc{} must take resource usages into account, and needs an algebra on these --- a way of \emph{splitting} resources on parallel composition, of \emph{ticking off} resources on input and output.
-We characterise such an algebra in  \autoref{multiplicities}.
-We define independent typing contexts and usage contexts in  \autoref{contexts}.
-We provide a typing system with leftovers for a resource-aware \picalc{} in  \autoref{leftover-typing}.
+We characterise a usage algebra for our typing system in \autoref{multiplicities}.
+This algebra defines the way in which resources are \emph{split} on parallel composition; the way in which resources are \emph{consumed} on input and output.
+We then define typing and usage contexts in \autoref{contexts}.
+We provide a typing system for a resource-aware \picalc{} in \autoref{leftover-typing}.
 
@@
 
-In the linear \picalc{} each channel has an input or an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise multiplicities by defining an algebra such that linear, gradual and shared types all satisfy these algebraic laws.
-
-\todo{we haven't introduced the forall and implicit syntax}
+In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
+We generalise over this notion by defining an algebra that is satisfied by linear, gradual and shared types.
+
 \begin{nidefinition}
-  The algebra on multiplicities must be \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}, and needs a \emph{minimal neutral element}.
-  It also requires a special element that is to be used to count inputs and outputs.
-  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
-
-  \begin{figure}[h]
-  \begin{equation}
-  \begin{aligned}
-    &\zero                  &:{} &                      &        & C \\
-    &\one                   &:{} &                      &        & C \\
-    &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to et \\
-    &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
-    &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-    &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-    &\field{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
-    &\field{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
-    &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-    &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
-  \end{aligned}
-  \end{equation}
-  \caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid.}
-  \label{fig:multiplicities}
-  \end{figure}
+  Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}.
+  It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
+  It has an element $\one$ that is used to count inputs and outputs.
+  For a formal definition, refer to \autoref{fig:multiplicities} in \autoref{app:usage-algebra}.
 \end{nidefinition}
@@
 \begin{note}
-  We will often work with input and output multiplicities as pairs.
-  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple, and $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where all properties are trivially lifted.
+  We will often work with pairs of input and output multiplicities.
+  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple of input and output multiplicities, respectively.
+  We use $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are trivially lifted.
   Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
@@
 
-Linear, gradual and shared types are all defined as an instance of the laws in \autoref{fig:multiplicities}.
-We illustrate these types in typing derivations in \autoref{example-derivations}.
+Linear, gradual and shared types are all defined as an instance of our usage algebra.
+Their use in typing derivations is illustrated in \autoref{example-derivations}.
 
@@
   \item [Linear]
-    The carrier $C$ is implemented as a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
+    The carrier is a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
     The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
@@
     \item [Gradual]
-    The carrier $C$ is implemented as the type of natural numbers $\N$.
+    The carrier is the type of natural numbers $\N$.
     The element $\zero$ corresponds to 0, and $\one$ to 1.
@@
     \item [Shared]
-    The carrier $C$ is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
+    The carrier is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
     Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
@@
   \begin{split}
-    &\field{IDX}          &:{} &et \\
+    &\field{IDX}          &:{} &\Set \\
     &\field{\exists IDX}  &:{} &\field{IDX} \\
-    &\field{CARRIER}      &:{} &\field{IDX} \to et \\
+    &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
     &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}

### 2020-05-01 paper/appendix.tex
Keep the contexts subsection short

@@
 \todo{we haven't introduced the forall and implicit syntax}
-\begin{nidefinition}
-  The algebra on multiplicities must be \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}, and needs a \emph{minimal neutral element}.
-  It also requires a special element $\one$ that is to be used to count inputs and outputs.
-  We define the laws of such a \emph{partial commutative monoid} as a type $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities} --- we use a terniary relation to model the monoidal operation.
+\todo{introduce DEC and exists}
+
+Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}.
+It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
+It has an element $\one$ that is used to count inputs and outputs.
+These algebraic laws are defined as a record $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities}.
+We use a terniary relation to model the partial monoidal operation.
 
@@
 \end{figure}
-\end{nidefinition}
+
+\section{Indexed Set of Usage Algebras}
+\label{app:indexed-set-usage-algebra}
+
+\begin{figure}[h]
+  \begin{equation}
+    \begin{split}
+      &\field{IDX}          &:{} &\Set \\
+      &\field{\exists IDX}  &:{} &\field{IDX} \\
+      &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
+      &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
+    \end{split}
+  \end{equation}
+  \caption{Indexed set of usage algebras.}
+  \label{fig:indexed-multiplicities}
+\end{figure}
 

### 2020-05-01 paper/paper.tex
Keep the contexts subsection short

@@
 \item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
-  \autoref{syntax} uses type level de Bruijn indices \cite{} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
+  \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
   \autoref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
@@
 Furthermore, names do not guarantee that the syntax well-scoped.
-Instead, we use de Bruijn indices \cite{}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
+Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
 \begin{example}
@@
 \label{contexts}
-In a resource-aware \picalc{}, usage annotations are transformed as a result of input and output; types however are preserved.
-In our case, usage annotations do not depend on types, and keeping types and usages in separate contexts allows for a more modular approach in our proofs.
+
+We keep typing contexts and usage contexts separate.
+The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
 
 \begin{nidefinition}
-  We define types (\autoref{fig:prectx}) as either unit types, base types or channel types.
-  The unit type $\unit$ serves as a proof of inhabitance for types (we can trivially obtain it), and is used to such end in subject congruence --- for the $\new \PO_{\suc n} \eqeq \PO_n$ rule.
-  The base type $\base{n}$ uses a natural number $n$ to represent types.
-  This natural number can then be interpreted as a type in the host language.
-  Using natural numbers instead of type universes is convenient: less universe polymorphism poluting specifications.
-
-  The channel type $\channel{t}{x}$ uses $t$ to stand for the type of data that is being transmitted over the channel, and $x$ to stand for what usages of that data are transmitted.
-
+  \autoref{fig:prectx} defines types as either unit types, base types or channel types.
+  The unit type $\unit$ serves as a proof of inhabitance for types.
+  The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter.
+  This is for convenience, so that there is less universe polymorphism poluting specifications.
+  The type $\channel{t}{x}$ of a channel determines what type of data and what usage annotations are sent over that channel.
   We keep track of types in a length-indexed \emph{pre-context} $\PreCtx_n$.
@@
   We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a set of indices such that for every index there exists a carrier $C$ with a $\Quantifier_C$ algebra defined on it, as captured by \autoref{fig:indexed-multiplicities}.
-  We require the index type to be inhabited -- proving subject congruence on $\new \PO_{\suc n} \eqeq \PO_n$ requires an inhabited carrier set.
-
-  \begin{figure}[h]
-  \begin{equation}
-  \begin{split}
-    &\field{IDX}          &:{} &\Set \\
-    &\field{\exists IDX}  &:{} &\field{IDX} \\
-    &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
-    &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
-  \end{split}
-  \end{equation}
-  \caption{Indexed set of partial commutative monoids.}
-  \label{fig:indexed-multiplicities}
-  \end{figure}
+  An indexed set of usage algebras is a set of indices together with an interpretation of indices into types, and an interpretation of indices into usage algebras on that type.
+  This indexed set must be non-empty.
+  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:indexed-set-usage-algebra} for a formal definition.
 \end{nidefinition}
@@
 \begin{nidefinition}
-  A context of usages $\Ctx_{idxs}$ is indexed by a length-indexed context of indices $\Idxs_n$, as shown in \autoref{fig:ctx}.
+  We keep track of usages annotations in a context $\Ctx_{idxs}$ that is indexed by a length-indexed context of indices $\Idxs_n$, as shown in \autoref{fig:ctx}.
+
   \begin{figure}[h]
@@
 \begin{note}
-  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with a common underlying context of indices.
+  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
   All the properties of the partial monoid are trivially lifted.
@@
     Parallel composition no longer needs an explicit context splitting proof --- a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
-    Such a proof could be obtained through automated search --- but it is better yet to avoid needing it.
+    Such a proof could be obtained through automated search --- but it is better yet not to need it.
     With leftover typing, we use the leftovers of $P$ to type $Q$.

### 2020-05-01 paper/appendix.tex
Shorten intro to typing with leftovers

@@
 We use a terniary relation to model the partial monoidal operation.
+We then encapsulate a set carriers and their algebras in \autoref{fig:indexed-multiplicities}.
 
@@
 \end{figure}
-
-\section{Indexed Set of Usage Algebras}
-\label{app:indexed-set-usage-algebra}
 
@@
 \end{proof}
+
+\begin{nilemma}
+  \label{lm:types-unused}
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:types-op}
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}

### 2020-05-01 paper/paper.tex
Shorten intro to typing with leftovers

@@
   This indexed set must be non-empty.
-  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:indexed-set-usage-algebra} for a formal definition.
+  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
 \end{nidefinition}
@@
 
-We make use \emph{leftover typing} \cite{}, an approach that, in addition to the usual typing context and input usage context, adds an extra output usage context to the typing rules.
-This output context contains the \emph{leftovers} of the process being typed --- its unused multiplicities.
-Crucially, the leftovers can then be used as an input to some other typing derivation.
-Adopting this techinque has three main advantages:
+We use \emph{leftover typing} \cite{Allais2018a} for our typing rules, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
+This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
+These leftovers can then be used as an input to another typing derivation.
+This techinque presents three main advantages:
 \begin{itemize}
   \item
-    The input and output actions no longer transform the usage context through functions.
-    Instead, they use a variable reference type with both an input and an output usage context that encapsulates these transformations.
-  \item
-    Parallel composition no longer needs an explicit context splitting proof --- a proof $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$ where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
-    Such a proof could be obtained through automated search --- but it is better yet not to need it.
-    With leftover typing, we use the leftovers of $P$ to type $Q$.
-  \item
-    Theorems like framing ( \autoref{framing}), weakening ( \autoref{weakening}) and strengthening ( \autoref{strengthening}) become suddenly stateable.
+    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgment with both an input and an output usage context.
+  \item
+    The need for an extensional context splitting proof for parallel composition is eradicated.
+    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    This could be obtained through automated search --- but it is better yet not to need it.
+    Leftover typing uses the leftovers of $P$ to type $Q$.
+  \item
+    New theorems (like framing (\autoref{framing}), weakening (\autoref{weakening}) and strengthening (\autoref{strengthening}) can suddenly be state.
 \end{itemize}
 
-We introduce two judgments: one for variable references (\autoref{def:varref}), one for processes (\autoref{def:types}).
-Both of them combine a judgment on types and a judgment on usages.
-These could be kept separate, but doing so would risk incoherence: if typing derivations had any choice to make, the judgment on types could choose to do one thing while the judgment on usages chooses to do another.
-The constructions instroduced in this paper make the typing derivations completely syntax directed and render this problem inexistent, but we nevertheless choose to keep both judgments as one, for simplicity.
+We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
 
@@
 \end{example}
-
-\begin{nilemma}
-  \label{lm:types-unused}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:types-op}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
 

### 2020-05-02 paper/appendix.tex
Go over varref

@@
 \end{proof}
+
+\begin{nilemma}
+  \label{lm:comm-capable}
+  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
+\end{nilemma}
+
+\begin{proof}
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\end{proof}

### 2020-05-02 paper/paper.tex
Go over varref

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Structural Congruence}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{structural-congruence}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Reduction Relation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{operational-semantics}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \subsection{Typing Contexts}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{contexts}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \subsection{Typing with Leftovers}
 \label{leftover-typing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
   \item
-    New theorems (like framing (\autoref{framing}), weakening (\autoref{weakening}) and strengthening (\autoref{strengthening}) can suddenly be state.
+    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated.
 \end{itemize}
@@
 We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
+Both of these are indexed by a typing context, an input usage context, and an output usage context.
 
@@
 
-  The \textsc{VarRef} relation presented in \autoref{fig:variable-references} is indexed by a typing context and an input and output usage context.
-  In addition to these, it also keeps track of its position in the contexts, its type, and its usage annotation.
-
-  The base case $\constr{here}$ asserts that $t$ is the most immediate type in typing context, and splits the most immediate usage annotation $x$ in the input context into the usage annotation $y$ the usage annotation $z$ in the output context.
-  This is the only place in the typing system where the algebra on multiplicities is used.
+  The \textsc{VarRef} relation presented in \autoref{fig:variable-references} keeps track of the index, type, and usage annotation of a variable within a particular typing and usage context.
+  The leftover usage context is the result of subtracting the usage annotation from the input usage context.
+
+  The base case $\constr{here}$ splits the usage annotation $x$ into $y$ and $z$, the leftovers.
+  This splitting is as per the usage algebra defined on the index of the usage annotation.
 
   \begin{remark}
-    We use $\field{\cdot-compute}$ to aliviate the user from the proof burden $\opsquared{x}{y}{z}$.
+    We use the functionality of the monoidal relation to aliviate the user from the proof burden $\opsquared{x}{y}{z}$.
   \end{remark}
-
-  The inductive case $\constr{there}$ appends an extra type to the typing context, and an extra usage annotation to both the input and output usage context.
 
@@
 \begin{example}
-  The following is an example that asserts that the input usage context is capable of $\li$ on a $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{zero}$, and that consuming such $\li$ results in the stated leftovers.
-  \begin{lstlisting}[mathescape,caption=Here \_ introduces an anonymous declaration that must be immediately followed by its definition.]
+  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{zero}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
+  \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{zero}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
 In  \autoref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
-
-\begin{nilemma}
-  \label{lm:comm-capable}
-  As an auxiliary lemma, we must establish that every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
-
-\begin{proof}
-  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
-\end{proof}
 
@@
 
-\cite{Allais2018a}
-typing with leftovers
-bidirectional typing rules
-decidable typechecking and type inference
-limited to a linear type system
-framing, weakening, substitution
-soundness and completeness with respect to a sequent calculus for intuitionistic linear logic
+This work is heavily based on the work by Allais \cite{Allais2018a}.
+Allais uses leftover typing to build a bidirectional type system for the linear \lambdacalc{}.
+He proves type preservation and provides a procedure that decides type checking and type inference.
+He then goes on showing the soundness and completeness with respect to the sequent calculus for intuitionistic linear logic.
 

### 2020-05-02 paper/paper.tex
Finish revieweing the type system

@@
 
-  The typing relation \textsc{Types} presented in \autoref{fig:types} is indexed by a typing context, an input and an output usage context, and the process $P$ being typed.
+  The typing relation \textsc{Types} presented in \autoref{fig:types} asserts that process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$.
 
@@
 
-  The process termination case is only concerned with usage contexts.
-  Its input usage context is returned as leftovers.
-
-  Scope restriction expects three arguments: the type $t$ of data being transmitted; the input and output multiplicities $x$ of the data being transmitted; and the multiplicity $y$ assigned to the channel itself.
-  The same multiplicity $y$ is used as the input multiplicity and as the output multiplicity of the channel, so these are balanced.
-  The typing derivation of the continuation process $P$ is provided with this channel with multiplicity $(y \comma y)$.
-  Process $P$ must then completely exhaust this channel, and to that end we substitute $(y , y)$ in the input context with $\lz$ in the output context.
-
-  Receiving requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\li$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
-  Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\li$ from this channel, we use the leftover context $\Xi$ to type the continuation.
-  The continuation is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
-  This multiplicity $x$ must be completely exhausted by the continuation process.
-
-  Similar to input, output requires a proof $chan_i$ that selects a channel at index $i$ with multiplicity $\lo$ available, over which data with type $t$ and multiplicity $x$ is going to be sent.
-  Note that this index $i$ is then used in the syntax of the typed process.
-  After consuming $\lo$ from this channel, we use the leftover context $\Delta$ to type the continuation.
-  The continuation must select the element at index $j$ that is being sent.
-  This element must have type $t$ and at least multiplicity $x$ --- as per the channel's type.
-  The result of consuming these multiplicity $x$ is the leftover context $\Xi$, which is then used to type the continuation process.
+  Process termination threads usage annotations unchanged from input to output.
+
+  Scope restriction expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ that is being transmitted; and the multiplicity $y$ given to the channel itself.
+  This multiplicity $y$ is used for both input and output, so that these are balanced.
+  The continuation process $P$ is provided with the new channel with usage annotation $(y \comma y)$, which it must completely exhaust.
+
+  Receiving needs a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  Note that the index $i$ is used in the syntax of the typed process.
+  We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
+  The received element $x$ must be completely exhausted by the continuation process.
+
+  Similar to input, sending needs a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as specified by the channel $chan_i$.
+  The leftovers $\Xi$ are used to type the continuation process.
   Note that both indices $i$ and $j$ are used in the syntax of the typed process.
 
-  Parallel composition is only concerned with usage contexts.
-  It uses the leftovers of the left-hand process to type the right-hand process.
+  Parallel composition uses the leftovers of left-hand process to type the right-hand process.
 \end{nidefinition}
@@
   \label{example-derivations}
-  The following example uses both linear and shared types to type the process $p$. 
-  We use $\omega$ as a synonym for a $(\one \comma \one)$ multiplicity in an algebra for shared types.
-  The usage context $\epsilon$ signifies the empty context where all multiplicities are $(\zero \comma \zero)$.
-  The typing derivation creates a new channel over which $\li$ multiplicities of a type $\channel{\unit}{\omega}$ are sent.
-  This channel is initialized with a linear multiplicity of $(\one \comma \one)$, and therefore must be used exactly once for sending and once for receiving.
-  The right-hand parallel subprocess then creates a new channel as per the specification $\channel{\unit}{\omega}$.
+  The example below uses linear and shared types to type the process $p$. 
+  We define $\omega$ as a $(\one \comma \one)$ usage in an algebra for shared types.
+  We define $\epsilon$ as the empty usage context where all usages are $(\zero \comma \zero)$.
+  The typing derivation creates a new channel with usage $(\one \comma \one)$.
+  This channel transmits $\li$ multiplicities of a type $\channel{\unit}{\omega}$.
+  The right-hand subprocess creates a channel with usage $(\one \comma \one)$.
+  This channel transmits $\omega$ multiplicities of a type $\unit$.
+  Part of this channel ($(\one \comma \zero)$) is sent over the channel created by the parent process.
 

### 2020-05-02 paper/paper.tex
Fix silly mistakes -- thanks Erika!

@@
   This operational semantics is given as a reduction relation on processes.
-  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- ak-a type preservation.
+  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
   It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unecessary syntactic minutiae introduced by the syntax of the \picalc{}.
@@
   { }
-  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \eqeq \comp{\comp{P}{Q}}{R}}
+  {\constr{comp-assoc} : \comp{P}{(\comp{Q}{R})} \eqeq \comp{(\comp{P}{Q})}{R}}
 
@@
 The first three rules ($\constr{comp}-$*) state associativity, symmetry, and $\PO$ as the neutral element of parallel composition, respectively.
-The last three ($\constr{scope}-$*) state, garbage collection, scope extrusion and commutativity of restrictions, respectively.
+The last three ($\constr{scope}-$*) state garbage collection, scope extrusion and commutativity of restrictions, respectively.
 

### 2020-05-02 paper/paper.tex
Bring in Erika's suggestions

@@
 % Functions
-\newcommand{\subst}[3]{#1 \; [ \; #2 \; / \;#3 \;]}
+\newcommand{\subst}[3]{#1 \; [ \; #3 \; \mapsto \;#2 \;]}
 \newcommand{\opsquared}[3]{#1 \, \coloneqq \, #2 \, \cdot^2 \, #3}
@@
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
-\acknowledgements{We want to thank Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
 \funding{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
@@
 
-\todo{rename Scoped to Process in code}
 \todo{make sure all citations are there}
@@
 We use a $dec$ function to decrement the index of this channel every time we come out of a binder.
-This function saturates at $\constr{internal}$ --- meaning communication happens internally to the process.
+This function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
 
@@
   \datatype
-  {i : \Channel_n \\ P \; Q : \Process_n}
-  {P \reduce{i} Q : \Set}
+  {c : \Channel_n \\ P \; Q : \Process_n}
+  {P \reduce{c} Q : \Set}
   \; \textsc{Reduces}
@@
   \inferrule
-  {red : P \reduce{i} P'}
-  {\constr{par} \; red : \comp{P}{Q} \reduce{i} \comp{P'}{Q}}
-
-  \inferrule
-  {red : P \reduce{i} Q}
-  {\constr{res} \; red : \new P \reduce{dec\; i} \new Q}
-
-  \inferrule
-  {eq : P = P' \\ red : P' \reduce{i} Q}
-  {\constr{struct} \; eq \; red : P \reduce{i} Q}
+  {red : P \reduce{c} P'}
+  {\constr{par} \; red : \comp{P}{Q} \reduce{c} \comp{P'}{Q}}
+
+  \inferrule
+  {red : P \reduce{c} Q}
+  {\constr{res} \; red : \new P \reduce{dec\; c} \new Q}
+
+  \inferrule
+  {eq : P = P' \\ red : P' \reduce{c} Q}
+  {\constr{struct} \; eq \; red : P \reduce{c} Q}
 \end{mathpar}
@@
   We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple of input and output multiplicities, respectively.
-  We use $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are trivially lifted.
+  We use $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted elementwise.
   Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
@@
     The carrier is a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
-    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{\one}$), and is uninhabited in every other case.
-    All other rules follow trivially.
+    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited in every other case.
+    All other properties are lifted elementwise.
 
@@
   The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter.
-  This is for convenience, so that there is less universe polymorphism poluting specifications.
+  This is for convenience, so that there is less universe polymorphism polluting specifications.
   The type $\channel{t}{x}$ of a channel determines what type of data and what usage annotations are sent over that channel.
@@
     \inferrule
-    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}}}
+    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}^2}}
     {\channel{t}{x} : \Type}
@@
 
-Framing captures the idea that the well-typedness of a process does not depend on its unused resources.
+Framing captures the idea that the well-typedness of a process only depends on the resources consumed by it.
 These unused resources are not introduced as new variables -- the typing context remains unaltered -- but rather as extra multiplicities that are carried from the input context to the output context.

### 2020-05-02 paper/paper.tex
Change Carrier to Usage

@@
 \newcommand{\Idxs}{\type{IDXS}}
-\newcommand{\Carrier}{\type{CARRIER}}
+\newcommand{\Usage}{\type{USAGE}}
 \newcommand{\N}{\type{\mathbb{N}}}
@@
   We will often work with pairs of input and output multiplicities.
-  We use the notation $\Carrier^2$ to stand for a $\Carrier \times \Carrier$ tuple of input and output multiplicities, respectively.
+  We use the notation $\Usage^2$ to stand for a $\Usage \times \Usage$ tuple of input and output multiplicities, respectively.
   We use $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted elementwise.
@@
 
-We keep typing contexts and usage contexts separate.
+We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
 The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
@@
     \inferrule
-    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Carrier_{idx}^2}}
+    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Usage_{idx}^2}}
     {\channel{t}{x} : \Type}
@@
     \inferrule
-        {\Gamma : \Ctx_{idxs} \\ x : \Carrier_{idx} ^2}
+        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^2}
         {\Gamma \comma x : \Ctx_{idxs \comma idx}}
@@
         idx : \Idx \\\\
-        y : \Carrier_{idx}^2} \\
+        y : \Usage_{idx}^2} \\
       \Delta : \Ctx_{idxs}}
@@
     \inferrule
-    {t : \Type \\ x : \Carrier_{idx}^2 \\ y : \Carrier_{idx'} \\\\
+    {t : \Type \\ x : \Usage_{idx}^2 \\ y : \Usage_{idx'} \\\\
      cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma (y \comma y) }{P}{\Delta \comma \lz}}
@@
 
-Framing captures the idea that the well-typedness of a process only depends on the resources consumed by it.
-These unused resources are not introduced as new variables -- the typing context remains unaltered -- but rather as extra multiplicities that are carried from the input context to the output context.
+Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
 
@@
   \label{thm:framing}
-  Let $P$ be a well-typed process from context $\Gamma_l$ to $\Xi_l$.
+  Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
   Then, per \autoref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
@@
 
-Weakening states that adding new variables to the typing and usage contexts preserves the well-typedness of a process.
-The multiplicities of these new variables will be carried from the input context to the output context.
-Usually, weakening is presented through order preserving embeddings.
-Order preserving embeddings model a series of insertions into a context.
-In this paper, we use weakening to prove subject congruence.
-To do so, we need of a single insertion into a context, and thus we will limit our theorem statement to that, but there is no loss of generality.
+Weakening states that adding new variables to the typing and input and output usage contexts preserves the well-typedness of a process.
+The usage annotations of the new variables will be left untouched as leftovers.
 
@@
   \label{thm:weakening}
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let us insert an arbitrary type $t$ at position $i$ in the typing context and an arbitrary multiplicity $x$ at position $i$ into both the input and output contexts.
-  Let us reflect this on the process $P$ by incrementing every index greater than or equal to $i$.
+  Let $P$ be a well-typed process $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Insert an arbitrary type $t$ at position $i$ in the typing context, and an arbitrary usage annotation $x$ at position $i$ into both the input and output contexts.
+  Increment every referrence to a variable greater than or equal to $i$ in process $P$.
   The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.

### 2020-05-02 paper/appendix.tex
Fix several slips

@@
       &\field{\exists IDX}  &:{} &\field{IDX} \\
-      &\field{CARRIER}      &:{} &\field{IDX} \to \Set \\
-      &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{CARRIER}_i}
+      &\field{USAGE}        &:{} &\field{IDX} \to \Set \\
+      &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{USAGE}_i}
     \end{split}

### 2020-05-02 paper/paper.tex
Fix several slips

@@
 
-\begin{note}
-  We will often work with pairs of input and output multiplicities.
-  We use the notation $\Usage^2$ to stand for a $\Usage \times \Usage$ tuple of input and output multiplicities, respectively.
-  We use $\op{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted elementwise.
-  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
-\end{note}
-
 Linear, gradual and shared types are all defined as an instance of our usage algebra.
@@
 \label{contexts}
+
+\begin{nidefinition}
+  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
+  An indexed set of usage algebras is a set of indices $\Idx$ together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras on that type.
+  $\Idx$ must be non-empty.
+  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
+\end{nidefinition}
+
+\begin{note}
+  We will often work with pairs of input and output multiplicities.
+  We use the notation $\Usage_{idx}^2$ to stand for a $\Usage_{idx} \times \Usage_{idx}$ tuple of input and output multiplicities, respectively.
+  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted elementwise.
+  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
+\end{note}
 
@@
   \end{figure}
-\end{nidefinition}
-
-\begin{nidefinition}
-  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a set of indices together with an interpretation of indices into types, and an interpretation of indices into usage algebras on that type.
-  This indexed set must be non-empty.
-  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
 \end{nidefinition}

### 2020-05-03 paper/paper.tex
little black heart

@@
 \newtheorem{nilemma}[theorem]{Lemma}
+
+% An i with a heart on top
+\usepackage{pifont}
+\usepackage{graphics}
+\newcommand\iheart{%
+    \begingroup
+    \settowidth{\dimen0}{i}%
+    i%
+    \llap{\makebox[\dimen0][c]{%
+        \kern-0.000em%
+        \raisebox{1.1ex}{\scalebox{0.30}{\ding{170}}}%
+    }}%
+    \endgroup
+}
 
@@
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
-\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+\acknowledgements{We want to thank Er\iheart ka Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
 \funding{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233

### 2020-05-03 paper/paper.tex
All the way to subject congruence

@@
 
-Weakening states that adding new variables to the typing and input and output usage contexts preserves the well-typedness of a process.
-The usage annotations of the new variables will be left untouched as leftovers.
+Adding a new variable to the context preserves the well-typedness of a process as long as the usage annotation of that variable is preserved unchanged as a leftover.
 
@@
   Insert an arbitrary type $t$ at position $i$ in the typing context, and an arbitrary usage annotation $x$ at position $i$ into both the input and output contexts.
-  Increment every referrence to a variable greater than or equal to $i$ in process $P$.
-  The resulting process will be well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+  For simplicity, we overload $insert$ and use it to insert types in typing contexts and usage annotations in usage contexts.
+  Increment every variable referrence greater than or equal to $i$ in process $P$.
+  The resulting process is well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
 \end{nitheorem}
@@
 
-Strengthening states that removing unused variables from the typing and usage contexts preserves the well-typedness of a process.
-The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations between the input and output contexts: a process can refer to a variable and still make no use of it -- e.g.
-by sending none of its multiplicities over a channel.
+Removing an unused variable preserves the well-typedness of a process.
+The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations as leftovers: a process can refer to a variable and still make no use of it --- by sending $\lz$ usage annotations over a channel.
 
@@
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable index that is not mentioned by $P$, such that $\Unused_i \; P$.
-  Then removing index $i$ from the typing context and both of the usage contexts and decrementing every variable in $P$ greater than $i$ preserves its well typedness, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+  Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
+  Remove the elements at index $i$ from the typing and usage contexts and decrement every variable in $P$ greater than $i$.
+  For simplicity, we overload $delete$ and use it to delete types from typing contexts and usage annotations from usage contexts.
+  The result preserves the well-typedness of the process, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
 \end{nitheorem}
@@
 
-Swapping two variables (in the typing context, input and output usage contexts and process) preserves the well-typedness of a process.
+Swapping two variables preserves the well-typedness of a process.
 

### 2020-05-03 paper/paper.tex
Grammar mistake
Spell-check

@@
     \llap{\makebox[\dimen0][c]{%
-        \kern-0.000em%
-        \raisebox{1.1ex}{\scalebox{0.30}{\ding{170}}}%
+        \color{white}%
+        \raisebox{1.23ex}{\scalebox{0.40}{$\blacksquare$}}%
+      }}%
+    \llap{\makebox[\dimen0][c]{%
+        \raisebox{1.23ex}{\scalebox{0.20}{\ding{170}}}%
     }}%
@@
 Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types \cite{K07}.
-This makes the \picalc{} a fully-fledged tool for the modelling and verification of concurrent and distributed systems.
+This makes the \picalc{} a fully-fledged tool for the modeling and verification of concurrent and distributed systems.
 
@@
 
-Considering the importance of the \picalc{} and of linearity in modelling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
+Considering the importance of the \picalc{} and of linearity in modeling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
 We present for the first time a full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.
@@
 The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modelling and programming with a variety of type systems, on top of which more advanced types, theories and languages can be built.
+Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modeling and programming with a variety of type systems, on top of which more advanced types, theories and languages can be built.
 While linearity is needed for type safety, gradual types and shared types are needed for real-world systems.
@@
   The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
-  It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+  It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
 
@@
   Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
-  These technique renders explicit context splits unecessary.
+  These technique renders explicit context splits unnecessary.
 
@@
   \item \textbf{Fully mechanised formalisation available in Agda}:
-  This work has been fully formalised in Agda and is publicly avaible \cite{Zalakain2020Agda}.
+  This work has been fully formalised in Agda and is publicly available \cite{Zalakain2020Agda}.
 
@@
 \begin{nidefinition}
-  Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}.
+  Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
     The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited in every other case.
-    All other properties are lifted elementwise.
+    All other properties are lifted element-wise.
 
@@
   We use the notation $\Usage_{idx}^2$ to stand for a $\Usage_{idx} \times \Usage_{idx}$ tuple of input and output multiplicities, respectively.
-  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted elementwise.
-  Henceworth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
+  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
+  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
 \end{note}
@@
 These leftovers can then be used as an input to another typing derivation.
-This techinque presents three main advantages:
+This technique presents three main advantages:
 \begin{itemize}
@@
   \begin{remark}
-    We use the functionality of the monoidal relation to aliviate the user from the proof burden $\opsquared{x}{y}{z}$.
+    We use the functionality of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
   \end{remark}
@@
 
-Adding a new variable to the context preserves the well-typedness of a process as long as the usage annotation of that variable is preserved unchanged as a leftover.
+Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
 
@@
   For simplicity, we overload $insert$ and use it to insert types in typing contexts and usage annotations in usage contexts.
-  Increment every variable referrence greater than or equal to $i$ in process $P$.
+  Increment every variable reference greater than or equal to $i$ in process $P$.
   The resulting process is well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
@@
 \begin{remark}
-  We limit ourselves to the swapping of contiguous variables $i$ and $\suc i$.
+  We limit ourselves to the swapping of adjacent variables $i$ and $\suc i$.
   Doing so is enough to prove subject reduction.
@@
   \label{thm:substitution}
-  Let $P$ be a process where the most recent variable has all its $m$ multiplities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let $P$ be a process where the most recent variable has all its $m$ multiplicities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $j$ be the index of the data with $m$ multiplicities that is being received after $P$ is done, so that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
@@
 
-Work that will be done time permiting:
+Work that will be done time permitting:
 

### 2020-05-03 paper/paper.tex
Go over type safety

@@
   {uQ : \Unused_0 \; Q}
-  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower_0 \; \; Q \; uQ}}
-
-  \inferrule
-  { }
-  {\constr{scope-comm} : \new \new P \eqeq \new \new swap_0 \; P}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower_{\constr{0}} \; \; Q \; uQ}}
+
+  \inferrule
+  { }
+  {\constr{scope-comm} : \new \new P \eqeq \new \new swap_{\constr{0}} \; P}
 \end{mathpar}
@@
 As a result of that communication, the continuation of the input process $P$ gets all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After substitution, $\subst{P}{j}{0}$ is \emph{lowered} --- all variable references are decreased by one.
+After substitution, $\subst{P}{j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one.
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower_0 \; (\subst{P}{j}{0})}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower_{\constr{0}} \; (\subst{P}{j}{\constr{0}})}{Q}}
 
@@
 \begin{nidefinition}
-  Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associative}, and \emph{commutative}.
+  We define our usage algebra to be a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
         { }
-        {[] : \PreCtx_0}
+        {[] : \PreCtx_{\constr{0}}}
 
@@
         { }
-        {[] : \Idxs_0}
+        {[] : \Idxs_{\constr{0}}}
 
@@
     {\opsquared{x}{y}{z}}
-    {\constr{here} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{zero}}{t}{y}{\Gamma \comma z}}
+    {\constr{here} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{0}}{t}{y}{\Gamma \comma z}}
 
@@
 \begin{example}
-  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{zero}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
+  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
   \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]
-  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{zero}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
+  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
   _ = $\constr{there} \; \constr{here}$
@@
   Note that the index $i$ is used in the syntax of the typed process.
-  We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{zero}$.
+  We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{0}$.
   The received element $x$ must be completely exhausted by the continuation process.
@@
   $p$ : $\Process_1$
-  $p$ = $\new{} (\comp{\recv{\constr{zero}} (\recv{\constr{zero}} \PO)}
-                    {\new{} (\send{\constr{suc} \; \constr{zero}} {\constr{zero}} (\send{\constr{zero}}{\constr{suc} \; \constr{suc} \; \constr{zero}}) \; \PO)})$
+  $p$ = $\new{} (\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}
+                    {\new{} (\send{\constr{suc} \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\constr{suc} \; \constr{suc} \; \constr{0}}) \; \PO)})$
 
@@
   \label{thm:weakening}
-  Let $P$ be a well-typed process $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Insert an arbitrary type $t$ at position $i$ in the typing context, and an arbitrary usage annotation $x$ at position $i$ into both the input and output contexts.
-  For simplicity, we overload $insert$ and use it to insert types in typing contexts and usage annotations in usage contexts.
-  Increment every variable reference greater than or equal to $i$ in process $P$.
-  The resulting process is well-typed so that $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+  Let $insert_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
+  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then lifting every variable greater than or equal to $i$ in $P$ is well-typed under $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
 \end{nitheorem}
@@
 Removing an unused variable preserves the well-typedness of a process.
-The proof that a variable is unused within a process must be syntactical and cannot rely on the preservation of its usage annotations as leftovers: a process can refer to a variable and still make no use of it --- by sending $\lz$ usage annotations over a channel.
+We say a variable is \emph{unused} within a process if neither inputs nor outputs refer to it.
+
+\begin{remark}
+  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned within $P$ --- $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
+\end{remark}
 
@@
   \label{thm:strengthening}
+  Let $delete_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
   Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
-  Remove the elements at index $i$ from the typing and usage contexts and decrement every variable in $P$ greater than $i$.
-  For simplicity, we overload $delete$ and use it to delete types from typing contexts and usage annotations from usage contexts.
-  The result preserves the well-typedness of the process, so that $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed under $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
 \end{nitheorem}
@@
   Doing so is enough to prove subject reduction.
-  The swapping of any two variables can be modeled as a sequence of contiguous variable swappings.
+  More generally, any two variables can be swapped by performing a series of adjacent variable swappings.
 \end{remark}
@@
 
-Structural congruence is an equivalence relation that aims to overcome the semantically meaningless difference introduced by the syntax of the \picalc{}.
-
 \begin{nitheorem}
   \label{thm:subject-congruence}
-  Applying structural congruence rules to a well-typed process should preserve its well-typedness under the same typing and usage contexts.
-  If $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  Applying structural congruence rules to a well-typed process preserves its well-typedness.
+  That is, if $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
@@
   By induction on \textsc{Equals}.
-  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
+  The base cases under $\constr{struct}$ (and their symmetric variants under $\constr{sym}$) are the nontrivial ones.
   For those, we proceed by induction on \textsc{StructCong}:
@@
     \item
-      The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
-    \item
-      The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\autoref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
-      The usage context on which $Q$ and $P$ coincide in $\comp{Q}{P}$ will no longer be the same.
-    \item
-      The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
-    \item
-      In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\lz$, making the proof trivial.
-      In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
-      For this, we will use type $\unit$ and multiplicities $\lz$.
-    \item
-      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of $\autoref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
-      We then use strengthening (\autoref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
-      In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
-      We will then use \autoref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
-    \item
-      Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \autoref{thm:swapping}.
-      In the reverse direction, we will use swapping and then use \autoref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
+      The case $\constr{comp-assoc}$ is trivial: leftover typing is naturally associative.
+    \item
+      In the $\constr{comp-sym}$ case for a process $\comp{P}{Q}$, we need framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of of $Q$ to the one of $P$.
+      The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
+    \item
+      The case $\constr{comp-end}$ is trivial: the typing rule for a process $\PO$ has the same input and output contexts.
+    \item
+      In the $\constr{scope-end}$ case, we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
+      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
+    \item
+      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
+      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with typing derivation on $Q$ wrapped in $\constr{chan}$.
+      In the reverse direction, we weaken (\autoref{thm:weakening}) $P$ so that it accepts the additional variable.
+      We then use \autoref{lm:lower-lift} to show that lowering and then lifting $P$ results in $P$.
+    \item
+      In the $\constr{scope-comm}$ case, we use swapping (\autoref{thm:swapping}).
+      In the reverse direction, we use swapping and then \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
   \end{itemize}
 \end{proof}
-\todo{make sure constructor names in code coincide}
-
-\begin{remark}
-  We encode the recursive structure of \textsc{Equals} in its inductive definition -- by indexing its type with \textsc{Rec}.
-  This acts as a justification of the well-formedness of the recursion scheme, which Agda cannot immediately see as terminating due to the interaction between the transitive and symmetric cases.
-\end{remark}
 
@@
 
-When function abstraction and application come together in the \lambdacalc{} with de Bruijn indices, every reference to the most recent variable in the body of the abstraction gets replaced with the argument of the application.
-In the \picalc{} however substitution takes another form: every reference to the most recent variable in the receiving process gets replaced with the reference to the data that is being sent over the channel.
-In a resource-aware \picalc{}, that means that instead of exhausting all $m$ multiplicities of the most recent variable, we leave those unaltered and exhaust them from the variable that is being received.
-
-\begin{nitheorem}
-  \label{thm:substitution}
-  Let $P$ be a process where the most recent variable has all its $m$ multiplicities exhausted, such that $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
-  Let $j$ be the index of the data with $m$ multiplicities that is being received after $P$ is done, so that $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute variable $\constr{zero}$ in $P$ with $\suc j$.
-  As a result, the multiplicities at $\constr{zero}$ will be preserved and those at $j$ will be consumed instead, resulting in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi \comma m}$.
-\end{nitheorem}
-\begin{proof}
-  We need to proceed by induction on $P$, and to do so means going under scope restriction.
-  We use the generalised version in \autoref{thm:subst-generalization} so that the variable that is being substituted is not $\constr{zero}$ but any variable $i$.
-  \todo{explain arguments}
-\end{proof}
+\autoref{thm:subst-generalization} proves a generalized version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
+\autoref{thm:substitution} instantiates the generalized version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
 
@@
   \label{thm:subst-generalization}
-
-  Let the variable that is being substituted be $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$ and the variable that it is being substituted with $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$.
-  As shown in \autoref{fig:subst}, the process that is being substituted will pivot from the input usage context $\Gamma_i$ to the input usage context $\Gamma_j$.
-  Let there be two \emph{adjustment} variables $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$ and $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$.
-  The process that is being substituted will pivot from the output usage context $\Psi_i$ to the output usage context $\Psi_j$.
-  As shown in \autoref{fig:subst-alternative}, without this \emph{adjustment} composing the substitution of subprocesses is not at all trivial.
-  Assume that $P$ uses all $m$ multiplicities at index $i$.
-  To model this, require $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$ where index $i$ is $\lz$.
-  The process that is typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$ will be typed $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$ after substitution.
+  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$,
+  Let $\ni_i m$ be a \textsc{VarRef} $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$ for some $m$ and $\Gamma$.
+  Let $\ni_i n$ be a \textsc{VarRef} $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$ for some $m$ and $\Psi$.
+  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
+  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
+  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$ for some $\Gamma_j$.
+  Let $\ni_j n$ be a \textsc{VarRef} $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$ for some $\Psi_j$.
+  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Refer to \autoref{fig:subst} for a diagramatic representation.
 \end{nitheorem}
@@
       Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      Composition is non-trivial.
+      This approach makes composing typing derivations before and after substitution more difficult.
     }
@@
 
+\begin{nidefinition}
+  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+\end{nidefinition}
+
 \begin{proof}
@@
     \item
-      From constructor $\constr{end}$ we derive $\Gamma_i \equiv \Psi_i$.
+      For constructor $\constr{end}$ we get $\Gamma_i \equiv \Psi_i$.
       From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
@@
     \item
-      For constructor $\constr{chan}$ we proceed inductively, applying the constructor $\suc$ to all arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$.
-
-    \item
-      For constructor $\constr{recv}$ we split $\Delta$ so that we can proceed inductively on the continuation.
-      We first observe that given that there exists an arrow from $\Gamma_i$ to $\Psi_i$ and $\Delta$ at index $i$ is $\lz$, for any $m$ and $n$, there is some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\constr{there}$.
+
+    \item
+      For constructor $\constr{recv}$ we must split $\Delta$ to proceed inductively on the continuation.
+      Observe that given the arrow from $\Gamma_i$ to $\Psi_i$ and given that $\Delta$ is $\lz$ at index $i$, there must exist some $\delta$ such that $\opsquared{m}{\delta}{n}$.
  l     \begin{itemize}
         \item
-          If the input occurs on the variable that is being substituted, we derive $\opsquared{m}{\li}{l}$ for some $l$ and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ (used for the inductive call).
-          Similarly, we construct the arrows $\containsusage{\Gamma_j}{j}{\li}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ (used for the inductive call) for some $\Xi_j$.
-        \item
-          If the input occurs on some variable $x$ other than the one that is being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (used for the inductive call) and $\containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
-          We can then find $\Xi_j$ and construct the two arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (used as the new channel on which input occurs) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (used for the inductive call).
-          From the fact that there exists a composition of arrows from $\Xi_i$ to $\Psi$, we can conclude that $\Theta$ is splitting $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-          As $\lz$ is a minimal element, if $\Delta$ is $\lz$ at index $i$, then it must be $\lz$ at index $i$ for $\Delta_1$ and $\Delta_2$ as well.
+          If the input is on the variable being substituted, we split $m$ such that $\opsquared{m}{\li}{l}$ for some $l$, and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ for the inductive call.
+          Similarly, we construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{j}{\li}{\Xi_j}$ as the new input channel, and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ for the inductive call.
+        \item
+          If the input is on a variable $x$ other than the one being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (for the inductive call) and $\containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
+          We then construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (the new output channel) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (for the inductive call).
+          Given there exists a composition of arrows from $\Xi_i$ to $\Psi$, we conclude that $\Theta$ splits $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+          As $\lz$ is a minimal element, then $\Delta_1$ must be $\lz$ at index $i$, and so must $\Delta_2$.
       \end{itemize}
@@
     \item
-      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the channel doing the output and the transmitted data.
+      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the \textsc{VarRef} doing the output, and the \textsc{VarRef} for the sent data.
 
@@
       For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-      Given that $\Delta$ is $\lz$ at index $i$, we know that $\Delta_1$ and $\Delta_2$ are so too.
-      We observe that $\opsquared{m}{\delta}{\omega}$ where $\omega$ is what subprocess $P$ consumes at index $i$.
-      We use this to construct an arrow $\containsusage{\Xi_j}{j}{\delta}{\Theta}$, for some $\Xi_j$.
-      We use these constructions to make two inductive calls (on the derivation of $P$ and $Q$) and compose their result.
+      Given $\Delta$ is $\lz$ at index $i$, we conclude that $\Delta_1$ and $\Delta_2$ are too.
+      Observe that $\opsquared{m}{\delta}{\psi}$, where $\psi$ is the usage annotation at index $i$ consumed by the subprocess $P$.
+      We construct an arrow $\containsusage{\Xi_j}{j}{\delta}{\Theta}$, for some $\Xi_j$.
+      We can now make two inductive calls (on the derivation of $P$ and $Q$) and compose their results.
   \end{itemize}  
@@
 
+\begin{nitheorem}
+  \label{thm:substitution}
+  Let $P$ be a process typed under $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed under $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+\end{nitheorem}
+\begin{proof}
+  Use framing to derive $\contains{\gamma}{\Gamma}{j}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
+  Supply these to the generalised \autoref{thm:subst-generalization}.
+\end{proof}
+
 \subsection{Subject Reduction}
@@
 
-Subject reduction (aka type preservation) relates the operational semantics and the typing system of a language.
-In its usual form, it states that if $P$ reduces to $Q$ and $P$ is well-typed under context $\Gamma$, then $Q$ is well-typed under $\Gamma$ too.
-In the linear \picalc{} however, reduction \emph{is} meant to transform the context: each time communication occurs over a channel, that channel is exhausted.
-In our case, and because we generalise linearity to a partial monoid, the channel over which the communication occurs will loose a multiplicity $\one$ on both its input and output capabilities.
-In  \autoref{semantics}, we made sure to lift the index of the communicating channel to the type level, now we use it to state subject reduction.
+In the linear \picalc{} reduction affects the usage context: communicating over a channel takes away $\one$ input and $\one$ output multiplicity.
 
@@
   Let $P \reduce{c} Q$.
-  If $c$ is $\constr{internal}$ to the process, the context must not be modified and thus $\Gamma' \equiv \Gamma$.
-  If $c$ is $\constr{external}\; i$, the input usage context at index $i$ must be decremented, and therefore $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
-  Then the typing derivation $\types{\gamma}{\Gamma'}{P}{\Xi}$ can be transformed into $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  If $c$ is $\constr{internal}$, then $\Gamma' \equiv \Gamma$.
+  If however $c$ is $\constr{external}\; i$, then $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
+  Given $\types{\gamma}{\Gamma'}{P}{\Xi}$ then after reduction $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
@@
     \item
-    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower_0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in \autoref{fig:subject-reduction:a}.
-    We start by using framing (\autoref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
+    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions in \autoref{fig:subject-reduction:a} into the proofs in \autoref{fig:subject-reduction:b}.
+    We then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
 
     \begin{figure}[h]
-      \setlength{\mathindent}{0pt}
-      \begin{subfigure}{.3\textwidth}
+      \begin{subfigure}{.5\textwidth}
         \begin{alignat*}{2}
@@
         \end{alignat*}
-        \caption{Rearrange $\lo$ witness, and merge with $\li$ witness into $\lio$.}
+        \caption{Shift witness for $\lo$ and merge with the one for $\li$ into $\lio$.}
+        \label{fig:subject-reduction:b}
       \end{subfigure}
-      % Do not leave empty space
-      \begin{subfigure}{.5\textwidth}
-        \begin{align*}
-          \types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi \comma m}
-        \end{align*}
-        \caption{Substitution.}
-      \end{subfigure}
-      % Do not leave empty space
-      \begin{subfigure}{.5\textwidth}
-        \begin{align*}
-          \types{\gamma \comma t}{\Gamma \comma m}{lower_0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi \comma m}
-        \end{align*}
-        \caption{Strengthening.}
-      \end{subfigure}
-      \caption{Proof steps for the $\constr{comm}$ case in subject reduction.}
     \end{figure}
@@
     \item
-    Reduction under composition with $\constr{par}$ proceeds by induction on the process that is being reduced.
+    Reduction under $\constr{par}$ proceeds by induction on the process that is being reduced.
 
@@
     Reduction under scope restriction case splits on the channel $c$ on which communication occurs.
-    If the channel is $\constr{internal}$ to the process subject reduction proceeds inductively.
-    If the channel is $\constr{external}\; \constr{zero}$ -- i.e.
-the channel that is being introduced by the scope restriction -- we use the \autoref{lm:comm-capable} to subtract $\lio$ from the channel's multiplicity first, then apply the induction hypothesis.
-    If the channel is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
-
-    \item
-    Finally, in the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and the induction hypothesis.
+    If $c$ is $\constr{internal}$ subject reduction proceeds inductively.
+    If $c$ is $\constr{external}\; \constr{0}$ -- i.e.
+the channel that is being introduced by the scope restriction -- we use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and then apply the induction hypothesis.
+    If $c$ is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
+
+    \item
+    In the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and the induction hypothesis.
   \end{itemize}

### 2020-05-03 paper/appendix.tex
Color functions, change colors

@@
   &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to et \\
-  &\field{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
-  &\field{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-  &\field{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  &\field{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
-  &\field{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
-  &\field{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-  &\field{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+  &\func{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+  &\func{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+  &\func{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+  &\func{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+  &\func{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
+  &\func{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+  &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
 \end{aligned}
@@
     \begin{split}
-      &\field{IDX}          &:{} &\Set \\
-      &\field{\exists IDX}  &:{} &\field{IDX} \\
-      &\field{USAGE}        &:{} &\field{IDX} \to \Set \\
-      &\field{QUANTIFIERS}  &:{} &(i : \field{IDX}) \to \Quantifier_{\field{USAGE}_i}
+      &\func{IDX}          &:{} &\Set \\
+      &\func{\exists IDX}  &:{} &\func{IDX} \\
+      &\func{USAGE}        &:{} &\func{IDX} \to \Set \\
+      &\func{QUANTIFIERS}  &:{} &(i : \func{IDX}) \to \Quantifier_{\func{USAGE}_i}
     \end{split}
@@
   \label{lm:lower-lift}
-  The function $lower_i \; P \; uP$ has an inverse $lift_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $lift_i \; (lower_i \; P \; uP) \equiv P$.
+  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
 \end{nilemma}
@@
   \label{lm:swap-swap}
-  The function $swap_i \; P$ is its own inverse: $swap_i \; (swap_i \; P) \equiv P$.
+  The function $\func{swap}_i \; P$ is its own inverse: $\func{swap}_i \; (\func{swap}_i \; P) \equiv P$.
 \end{nilemma}

### 2020-05-03 paper/paper.tex
Color functions, change colors

@@
 
+\usepackage[dvipsnames]{xcolor}
 \usepackage[utf8]{inputenc}
@@
 
-\newcommand{\type}[1]{\textcolor{blue}{\operatorname{#1}}}
-\newcommand{\constr}[1]{\textcolor{olive}{\operatorname{#1}}}
-\newcommand{\field}[1]{\textcolor{red}{\operatorname{#1}}}
+\newcommand{\type}[1]{\textcolor{BlueViolet}{\operatorname{#1}}}
+\newcommand{\constr}[1]{\textcolor{BurntOrange}{\operatorname{#1}}}
+\newcommand{\func}[1]{\textcolor{OliveGreen}{\operatorname{#1}}}
 
@@
 % Functions
-\newcommand{\subst}[3]{#1 \; [ \; #3 \; \mapsto \;#2 \;]}
-\newcommand{\opsquared}[3]{#1 \, \coloneqq \, #2 \, \cdot^2 \, #3}
-\newcommand{\opctx}[3]{#1 \, \coloneqq \, #2 \, \otimes \, #3}
-
-% Fields
-\newcommand{\op}[3]{#1 \; \field{\coloneqq} \; #2 \; \field{\cdot} \; #3}
-\newcommand{\zero}{\field{0\cdot}}
-\newcommand{\one}{\field{1\cdot}}
-\newcommand{\li}{\ell_i}
-\newcommand{\lo}{\ell_o}
-\newcommand{\lz}{\ell_{\o}}
-\newcommand{\lio}{\ell_{\#}}
+\newcommand{\subst}[3]{#1 \; \func{[} \; #3 \; \func{\mapsto} \;#2 \; \func{]}}
+\newcommand{\op}[3]{#1 \; \func{\coloneqq} \; #2 \; \func{\cdot} \; #3}
+\newcommand{\opsquared}[3]{#1 \, \func{\coloneqq} \, #2 \, \func{\cdot^2} \, #3}
+\newcommand{\opctx}[3]{#1 \, \func{\coloneqq} \, #2 \, \func{\otimes} \, #3}
+\newcommand{\zero}{\func{0\cdot}}
+\newcommand{\one}{\func{1\cdot}}
+\newcommand{\li}{\func{\ell_i}}
+\newcommand{\lo}{\func{\ell_o}}
+\newcommand{\lz}{\func{\ell_{\o}}}
+\newcommand{\lio}{\func{\ell_{\#}}}
 
@@
 \todo{make sure all citations are there}
-\todo{change colors, color functions}
-\todo{do not distinguish fields from functions, it is not worth it}
-\todo{go over guillaume's paper again}
 
@@
 Implicit arguments are mentioned by type definitions but omitted in constructors.
-$\type{TYPES}$ are in blue and uppercased (with indices as subscripts), $\constr{constructors}$ are in olive, $\field{fields}$ in red, and variables and functions in black.
-Some constructor names are overloaded --- they and disambiguated by context.
+$\type{TYPES}$ are blue volet (uppercased, with indices as subscripts), $\constr{constructors}$ are burnt orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
 
@@
   {uQ : \Unused_0 \; Q}
-  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{lower_{\constr{0}} \; \; Q \; uQ}}
-
-  \inferrule
-  { }
-  {\constr{scope-comm} : \new \new P \eqeq \new \new swap_{\constr{0}} \; P}
+  {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{\func{lower}_{\constr{0}} \; \; Q \; uQ}}
+
+  \inferrule
+  { }
+  {\constr{scope-comm} : \new \new P \eqeq \new \new \func{swap}_{\constr{0}} \; P}
 \end{mathpar}
@@
 In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
-The function $lower_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
-In $\constr{scope-comm}$ the function $swap_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
 In all the above, $i$ is incremented every time we go under a binder.
@@
 We keep track of the channel over which communication occurs so that we can state subject reduction.
-We use a $dec$ function to decrement the index of this channel every time we come out of a binder.
+We use a $\func{dec}$ function to decrement the index of this channel every time we come out of a binder.
 This function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{lower_{\constr{0}} \; (\subst{P}{j}{\constr{0}})}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{j}{\constr{0}})}{Q}}
 
@@
   {red : P \reduce{c} Q}
-  {\constr{res} \; red : \new P \reduce{dec\; c} \new Q}
-
-  \inferrule
-  {eq : P = P' \\ red : P' \reduce{c} Q}
+  {\constr{res} \; red : \new P \reduce{\func{dec}\; c} \new Q}
+
+  \inferrule
+  {eq : P \eq{r} P' \\ red : P' \reduce{c} Q}
   {\constr{struct} \; eq \; red : P \reduce{c} Q}
@@
   We will often work with pairs of input and output multiplicities.
-  We use the notation $\Usage_{idx}^2$ to stand for a $\Usage_{idx} \times \Usage_{idx}$ tuple of input and output multiplicities, respectively.
+  We use the notation $\Usage_{idx}^{\func{2}}$ to stand for a $\Usage_{idx} \times \Usage_{idx}$ tuple of input and output multiplicities, respectively.
   We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
@@
     \inferrule
-    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Usage_{idx}^2}}
+    {t : \Type \\ \stacked{idx : \Idx \\\\ x : \Usage_{idx}^{\func{2}}}}
     {\channel{t}{x} : \Type}
@@
 
-        \\
-
     \datatype
@@
         { }
-        {[] : Ctx_{[]}}
-
-    \inferrule
-        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^2}
+        {[] : \Ctx_{[]}}
+
+    \inferrule
+        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^{\func{2}}}
         {\Gamma \comma x : \Ctx_{idxs \comma idx}}
@@
         idx : \Idx \\\\
-        y : \Usage_{idx}^2} \\
+        y : \Usage_{idx}^{\func{2}}} \\
       \Delta : \Ctx_{idxs}}
@@
     \inferrule
-    {t : \Type \\ x : \Usage_{idx}^2 \\ y : \Usage_{idx'} \\\\
+    {t : \Type \\ x : \Usage_{idx}^{\func{2}} \\ y : \Usage_{idx'} \\\\
      cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma (y \comma y) }{P}{\Delta \comma \lz}}
@@
   \label{thm:weakening}
-  Let $insert_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
+  Let $\func{insert}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then lifting every variable greater than or equal to $i$ in $P$ is well-typed under $\types{insert_i \; t \; \gamma}{insert_i \; x \; \Gamma}{lift_i \; P}{insert_i \; x \; \Xi}$.
+  Then lifting every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{insert}_i \; t \; \gamma}{\func{insert}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{insert}_i \; x \; \Xi}$.
 \end{nitheorem}
@@
   \label{thm:strengthening}
-  Let $delete_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
+  Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
   Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed under $\types{delete_i \; \gamma}{delete_i \; \Gamma}{lower_i \; P}{delete_i \; \Xi}$.
+  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
 \end{nitheorem}
@@
   \label{thm:swapping}
-  Let $swap_i$ be a function that swaps variables $i$ and $\suc i$.
-  For simplicity, we overload $swap$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
+  Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
+  For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
+  \todo{}
   Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then $\types{swap_i \; \gamma}{swap_i \; \Gamma}{swap_i \; P}{swap_i \; \Xi}$.
+  Then $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
 \end{nitheorem}

### 2020-05-03 paper/paper.tex
Go over contributions

@@
   \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
-  \autoref{semantics} provides an operational semantics for the untyped \picalc{}, prior to any typing.
+  \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
   This operational semantics is given as a reduction relation on processes.
   The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
-  It is also defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+  The relation is defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
 
@@
   The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
-  Linear types, gradual types and shared types are all instances of such a resource-aware algebra too.
-  Any \emph{partial commutative monoid} that is cancellative and has a minimal element is a valid such algebra.
-  Multiple of these algebras can be applied to a single type system --- contexts keep information about what algebra should be applied to which element.
+  Linear types, gradual types and shared types are all instances of such a resource-aware algebra.
+  Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
+  Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element.
 
@@
   \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
-  \autoref{type-safety} builds on several theorems to proof subject reduction for our type system.
-  Here too, leftover typing renders the proofs for some of these lemmas straightforward, while for others it makes it possible to state them in the first place.
-  We prove \textbf{framing} (well-typedness is independent of the unused resources of a process), \textbf{weakening} (new variables that are added to both the input and output usage context preserve well-typedness), and \textbf{strengthening} (removing the variables that are unused by a process preserves well-typedness).
-  We use these to prove \textbf{subject congruence} (structural congruence rules can be freely applied while preserving well-typedness) and \textbf{substitution} (a variable can be substituted for another as long as this other variable is capable of doing everything the original one did), and use these to prove \textbf{subject reduction} (a well-typed process that reduces will still be well-typed).
+  Leftover typing allows \emph{framing}, \emph{weakening} and \emph{strengthening} theorems to be stated.
+  We use these to prove \emph{subject congruence}, and together with \emph{substitution}, to prove \emph{subject reduction}.
 

### 2020-05-03 paper/paper.tex
Add subst-unused lemma, fix reduction

@@
 As a result of that communication, the continuation of the input process $P$ gets all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After substitution, $\subst{P}{j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one.
+After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one.
+To do so, we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$.
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{j}{\constr{0}})}{Q}}
+  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
 

### 2020-05-03 paper/paper.tex
Minor corrections

@@
 \begin{nidefinition}
-  We define our usage algebra to be a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associative}, and \emph{commutative}.
+  We define our usage algebra to be a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
   Similar to input, sending needs a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
-  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as specified by the channel $chan_i$.
+  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of channel $chan_i$.
   The leftovers $\Xi$ are used to type the continuation process.
@@
 
+Erika Kreuer: model usage algebra as a state machine?
+
 \begin{description}

### 2020-05-03 paper/paper.tex
Be careful not to confuse the reader with dots in the syntax

@@
         P,Q ::=      & \; \PO                \\ 
-                    |& \; (\new{}x)  P       \\ 
+                    |& \; (\new{}x) \; P       \\ 
                     |& \; \comp{P}{Q}        \\ 
-                    |& \; \recvp{x}{y}. \; P \\ 
-                    |& \; \sendp{x}{y}. \; P 
+                    |& \; \recvp{x}{y} P \\ 
+                    |& \; \sendp{x}{y} P 
       \end{split}
@@
 Process $\PO$ denotes the terminated process, where no operations can further occur.
-Process $(\new{}x)P$ creates a new channel $x$ bound with scope $P$.
+Process $(\new{}x)\; P$ creates a new channel $x$ bound with scope $P$.
 Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recvp{x}{y}. \; P$ and $\sendp{x}{y}. \; P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-
-Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y}. \; P$ both introduce bound names in $P$.
+Processes $\recvp{x}{y} P$ and $\sendp{x}{y} P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+
+Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names in $P$.
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
@@
     \begin{split}
-      & (\new{}x ) && (\comp {\recvp{x}{y} && . \; \sendp{y}{\texttt{"Hello world!"}} . \; \PO} {(\new{} y) && (\sendp{x}{y} . \; \recvp{y}{z} && . \; \PO)})
-      & \text{names}
+      & (\new{}x ) && (\comp {\recvp{x}{y} && \; \sendp{y}{a} \; \PO} {(\new{} y) && (\sendp{x}{y} \; \recvp{y}{z} && \; \PO)})
+      && \text{names}
       \\
-      & \new{} && (\comp {\recv{0} && . \; \send{0}{\texttt{"Hello world!"}} . \; \PO} {\new{} && (\send{1}{0} . \; \recv{0} && . \; \PO)})
-      & \text{de Bruijn}
+      & \new{} && (\comp {\recv{0} && \; \send{0}{a} \; \PO} {\new{} && (\send{1}{0} \; \recv{0} && \; \PO)})
+      && \text{de Bruijn}
       \\

### 2020-05-03 paper/paper.tex
Fix several slips

@@
 \usepackage{amsbsy}
+\usepackage{mathbbol}
 \usepackage{mathtools}
@@
 
-% An i with a heart on top
-\usepackage{pifont}
-\usepackage{graphics}
-\newcommand\iheart{%
-    \begingroup
-    \settowidth{\dimen0}{i}%
-    i%
-    \llap{\makebox[\dimen0][c]{%
-        \color{white}%
-        \raisebox{1.23ex}{\scalebox{0.40}{$\blacksquare$}}%
-      }}%
-    \llap{\makebox[\dimen0][c]{%
-        \raisebox{1.23ex}{\scalebox{0.20}{\ding{170}}}%
-    }}%
-    \endgroup
-}
-
 % No line numbers
@@
 % Constructors
-\newcommand{\PO}{\constr{\mathbf{0}}}
+\newcommand{\PO}{\constr{\mathbb{0}}}
 \newcommand{\comp}[2]{#1 \; \constr{\parallel} \; #2}
@@
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
-\acknowledgements{We want to thank Er\iheart ka Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
 \funding{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
@@
 
-\todo{remove space between inferrules}
 \begin{figure}[h]
@@
   The type $\channel{t}{x}$ of a channel determines what type of data and what usage annotations are sent over that channel.
-  We keep track of types in a length-indexed \emph{pre-context} $\PreCtx_n$.
+  We keep track of types in a length-indexed $\PreCtx_n$.
 

### 2020-05-03 paper/paper.tex
Move ^2 up to where multiplicities are introduced
Small clarification

@@
 
-Linear, gradual and shared types are all defined as an instance of our usage algebra.
+\begin{note}
+  We will often work with pairs of input and output multiplicities.
+  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \times \type{C}$ tuple of input and output multiplicities, respectively.
+  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
+  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
+\end{note}
+
+Linear, gradual and shared types (identified by a value in \Idx) are all defined as an instance of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.
@@
 \end{nidefinition}
-
-\begin{note}
-  We will often work with pairs of input and output multiplicities.
-  We use the notation $\Usage_{idx}^{\func{2}}$ to stand for a $\Usage_{idx} \times \Usage_{idx}$ tuple of input and output multiplicities, respectively.
-  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
-  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
-\end{note}
 

### 2020-05-03 paper/paper.tex
Pass to the intro

@@
 
-We live in a concurrent world where any given state is often computed interactively by a myriad of parties --- people, machines, processors, services, networks.
+We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
 As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
@@
 The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
-It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data through communication channels.
-Notably, it features \emph{channel mobility}: the channels themselves are first order constructs, and can therefore be transmitted.
+It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
+Notably, it features \emph{channel mobility}: channels themselves are first order constructs, and can therefore be transmitted.
 The \picalc{} acts as a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
-Moreover, at the state-of-the-art, the \picalc{} features a wide plethora of types: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types \cite{K07}.
+At the state-of-the-art, the \picalc{} features a wide plethora of types: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types \cite{K07}.
 This makes the \picalc{} a fully-fledged tool for the modeling and verification of concurrent and distributed systems.
@@
 
-In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave a different flavour to the linearity in the \picalc{}.
+In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave a different flavour to linearity in the \picalc{}.
 Session types are a type formalism used in communication-centric programming.
-In session types, linearity ensures that a channel must be owned by exactly one communicating participant, even if the channel itself can be used multiple times, as specified by the structure of its session type.
+In session types, linearity ensures that a channel is owned by exactly one communicating participant, even if the channel itself can be used multiple times, following the structure of its session type.
 Linearity in session types is the key ingredient that guarantees type safety.
 
-Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types.
-This encoding not only has theoretical benefits in terms of expressivity of session types and reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
+Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types.
+This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
 The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
-This means that we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
-
-Considering the importance of the \picalc{} and of linearity in modeling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
-We present for the first time a full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.
+This means that we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types are built.
+
+Considering the importance of the \picalc{} and of linearity in modeling concurrent and distributed systems, in this paper we focus our research on both and go beyond. \todo{}
+We present the first full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.
 We do so by defining a specification for an algebra on multiplicities and using it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{} for the first time.
 The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-Ultimately, we can use $\pi$ with leftovers as a unified framework for type safe distributed modeling and programming with a variety of type systems, on top of which more advanced types, theories and languages can be built.
-While linearity is needed for type safety, gradual types and shared types are needed for real-world systems.
+Ultimately, we can use this work as a unified framework for type safe distributed modeling and programming with a variety of type systems, and built more advanced types, theories and languages on top.
+While linearity is needed for type safety, gradual and shared types are needed for real-world systems.
 
@@
 
-Linear, gradual and shared types (identified by a value in \Idx) are all defined as an instance of our usage algebra.
+Linear, gradual and shared types (identified by a value in $\Idx$) are all defined as an instance of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.

### 2020-05-04 paper/paper.tex
a pass to introduction

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate,authorcolumns]{lipics-v2019}
+\documentclass[a4paper,UKenglish,cleveref,autoref,thm-restate,authorcolumns]{lipics-v2019}
 
@@
 \bibliographystyle{plainurl}
+
+\usepackage{hyperref}
+%%
+%% For section renaming in autoref
+\addto\extrasenglish{%
+\renewcommand{\sectionautorefname}{\S}%
+\renewcommand{\subsectionautorefname}{\S}%
+\def\exampleautorefname{Ex.}% 
+\renewcommand{\figureautorefname}{Fig.}%
+}
 
@@
 \section{Introduction}
-% The way I want to write the intro:
-% -- describe what is the picalc and why it is IMPORTATNT
-% -- describe what are linear types and why programming with them is IMPORTATNT
-% -- join together by saying that is why we focus on linear pi calculus but we do more than just that: we have linear, gradual and shared types
-% -- put emphasis on mechanisation in Agda
-
 \todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
-
+\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modeling is modelling}
+\todo{rename section and subsections in reference to \S (not sure why the commands above are not doing it so check later)}
 We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
@@
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
-Notably, it features \emph{channel mobility}: channels themselves are first order constructs, and can therefore be transmitted.
-The \picalc{} acts as a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
-At the state-of-the-art, the \picalc{} features a wide plethora of types: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types \cite{K07}.
-This makes the \picalc{} a fully-fledged tool for the modeling and verification of concurrent and distributed systems.
+Notably, it features channel mobility: channels themselves are first order values, and can therefore be transmitted.
+The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
+At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types.
+This makes the \picalc{} a fully-fledged tool for modeling and verification of concurrent and distributed systems.
 
@@
 
-The \picalc{} also benefited from this linearity wave in the early '90s.
-Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, a typed version of the \picalc{} where the linear type system restricts the usage of channels, which must be used exactly once.
-
-In the meantime, the rise of session types \cite{H93,THK94,HVK98} gave a different flavour to linearity in the \picalc{}.
+The \picalc{} also benefited from the linearity wave of the early '90s.
+Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
+%
+In the meantime, with the rise of \emph{session types} \cite{H93,THK94,HVK98} linearity got a different flavour in the \picalc{}.
 Session types are a type formalism used in communication-centric programming.
-In session types, linearity ensures that a channel is owned by exactly one communicating participant, even if the channel itself can be used multiple times, following the structure of its session type.
+In session types, linearity ensures that a channel is owned by exactly one communicating participant, whilst the channel itself is used multiple times by following the structure of its session type.
 Linearity in session types is the key ingredient that guarantees type safety.
 
-Recent work \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types.
-This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
+Recent work has pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17}.
+This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but it is useful at a practical level too.
 The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
-This means that we can use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types are built.
-
-Considering the importance of the \picalc{} and of linearity in modeling concurrent and distributed systems, in this paper we focus our research on both and go beyond. \todo{}
-We present the first full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.
-We do so by defining a specification for an algebra on multiplicities and using it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{} for the first time.
-The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-Ultimately, we can use this work as a unified framework for type safe distributed modeling and programming with a variety of type systems, and built more advanced types, theories and languages on top.
-While linearity is needed for type safety, gradual and shared types are needed for real-world systems.
+This allows us to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
+
+Considering the relevance of the \picalc{} and of linearity in modelling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
+
+\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.}
+
+We do so by defining a specification for an algebra on $0,1,n$ and $\omega$ \emph{multiplicities} (how many times a channel can be used) and use it to apply for the first time leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
+The developer is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
+Ultimately, we can exploit this formalisation as a unified framework for type safe distributed modelling and programming with a variety of type systems, and build on top of it more advanced types, theories and languages.
+In particular, while linearity is needed for type safety in communication-centric programming, gradual and shared types are needed to model real-world systems.
 
 \subsection{Contributions}
-Our contributions are the following:
+\begin{enumerate}
+\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
  \begin{itemize}
-\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
-  \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
-  \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
-  This operational semantics is given as a reduction relation on processes.
-  The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation.
-  The relation is defined modulo structural congruence --- a relation that acts as a quotient type that gets rid of the unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
+
+   \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
+   This operational semantics is given as a reduction relation on processes (in \autoref{operational-semantics}).
+   The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (in \autoref{subject-reduction}).
+   The reduction relation is defined modulo structural congruence (in \autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+ \end{itemize}
 
@@
   \autoref{type-system} provides the type system for a resource-aware \picalc{}.
-  The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
-  Linear types, gradual types and shared types are all instances of such a resource-aware algebra.
+  \begin{itemize}
+    \item \textbf{Algebras on multiplicities}: The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
+    Linear types, gradual types and shared types are all instances of such a resource-aware algebra (in \autoref{multiplicities}).
     Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
-  Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element.
-
-  The typing judgments use \emph{leftover typing} to model the resource-aware \picalc{}.
+    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (in \autoref{contexts}).
+
+    \item \textbf{Leftover typing}: The typing judgments use \emph{leftover typing} to model the resource-aware \picalc{} (in \autoref{leftover-typing}).
     This approach adds a leftover usage context to the typing judgments.
     Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
-  These technique renders explicit context splits unnecessary.
-
+    One of the benefits of this technique is that it removes the need for context splits, which is now not unnecessary.
+  \end{itemize}
   \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
-  Leftover typing allows \emph{framing}, \emph{weakening} and \emph{strengthening} theorems to be stated.
-  We use these to prove \emph{subject congruence}, and together with \emph{substitution}, to prove \emph{subject reduction}.
+  \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
+  Leftover typing allows \emph{framing} (\autoref{framing}), \emph{weakening} (\autoref{weakening}) and \emph{strengthening} (\autoref{strengthening}) theorems to be stated.
+  We use these to prove \emph{subject congruence} (\autoref{subject-congruence}), and together with \emph{substitution} (\autoref{substitution}), to prove \emph{subject reduction} (\autoref{subject-reduction}).
 
@@
 
-\end{itemize}
+\end{enumerate}
 

### 2020-05-04 paper/paper.tex
Notes on Allais and Thiemann

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+The notation used in this paper is given in \autoref{fig:notation}.
 \begin{figure}[h]
@@
   \caption{Notation used in this paper}
+  \label{fig:notation}
 \end{figure}
@@
 Data type definitions use double lines and index-free synonyms as rule names (for ease of reference).
-We otherwise use the constructor name as the name of a typing rule.
-Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
+For example ... \todo{complete the dots here, explain the rules above, in particular the 1+}
+We otherwise use the constructor name as the name of a typing rule. For example,...
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$. This means that ...
 Implicit arguments are mentioned by type definitions but omitted in constructors.
-$\type{TYPES}$ are blue volet (uppercased, with indices as subscripts), $\constr{constructors}$ are burnt orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
+
+We use colours to further distinguish the different entities in this paper.
+$\type{TYPES}$ are blue violet (uppercased, with indices as subscripts), $\constr{constructors}$ are burnt orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
 
@@
 
+Most of the work on mechanising resource-aware calculi is very recent.
+
 This work is heavily based on the work by Allais \cite{Allais2018a}.
 Allais uses leftover typing to build a bidirectional type system for the linear \lambdacalc{}.
-He proves type preservation and provides a procedure that decides type checking and type inference.
+He proves type preservation and provides a decision procedure for type checking and type inference.
 He then goes on showing the soundness and completeness with respect to the sequent calculus for intuitionistic linear logic.
+Also extrinsic.
+Also in Agda.
+Also de Bruijn.
+We apply this to the \picalc{}.
+We generalise the usage algebra.
+Our type system is not bidirectional --- scope restriction needs usage annotations.
+Adding bidirectionality would provide us with a decision procedure for type checking.
 
 \cite{Thiemann2019}
+intrinsicly typed
+linear session types with recursion and subtyping.
+also in Agda.
+Also de Bruijn.
+semantics as an intrinsically typed CEK machine
+explicit context splits
+MicroSession, minimal GV, lambda calculus with session types
+contexts remove elements with multiplicity 0
+keeps a global session context
+
+\cite{Goto2016a}
+first session-type paper with a mechanised proof of type safety
+Session types
+Coq
+Extrinisically typed
+Session type polymorphism
+
 \cite{Rouvoet2020}
 \cite{Veltri2020}
+\cite{Gay2001}
 

### 2020-05-04 paper/paper.tex
Write on Thiemann's paper

--- 2020-05-04 paper/paper.tex
@@
 
-This work is heavily based on the work by Allais \cite{Allais2018a}.
-Allais uses leftover typing to build a bidirectional type system for the linear \lambdacalc{}.
+Leftover typing as presented in this paper is based on work by Allais \cite{Allais2018a}.
+Allais uses leftover typing to mechanise in Agda a bidirectional type system for the linear \lambdacalc{}.
+This type system sits on top of an untyped but well-scoped syntax with de Bruijn indices.
 He proves type preservation and provides a decision procedure for type checking and type inference.
-He then goes on showing the soundness and completeness with respect to the sequent calculus for intuitionistic linear logic.
-Also extrinsic.
-Also in Agda.
-Also de Bruijn.
-We apply this to the \picalc{}.
-We generalise the usage algebra.
-Our type system is not bidirectional --- scope restriction needs usage annotations.
-Adding bidirectionality would provide us with a decision procedure for type checking.
-
-\cite{Thiemann2019}
-intrinsicly typed
-linear session types with recursion and subtyping.
-also in Agda.
-Also de Bruijn.
-semantics as an intrinsically typed CEK machine
-explicit context splits
-MicroSession, minimal GV, lambda calculus with session types
-contexts remove elements with multiplicity 0
-keeps a global session context
+In this paper, we apply his technique to the \picalc{}, but limit ourselves to type-checking (scope restriction needs usage annotations) terms, and focus on generalising the usage algebra.
+
+Taking an intrinsic approach, Thiemann formalises in Agda the MicroSession (minimal GV \cite{}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
+Context splits are given extrinsically, and exhausted elements are removed from typing contexts altogether.
+The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
 
@@
 
-\cite{LTS-semantics}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Erika Kreuer: model usage algebra as a state machine?
+
+Adding bidirectionality would provide us with a decision procedure for type checking.
 

### 2020-05-04 paper/paper.tex
a pass on the abstract

@@
   The \picalc{} is a computational model for communication and concurrency.
-  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channel resources must be used exactly once.
-  With the rise of session types as a formalism to model real-world distributed systems, the linear \picalc{} has re-emerged as the underlying theoretical and practical framework for session types and communication-based distributed programming \cite{DardhaGS12,Padovani17,ScalasY16,ScalasDHY17}.
-  While there have been several extensions to the session typed \picalc{}, the linear \picalc{} remains essentially unchanged and stable as a foundational calculus.
-
-  We present the \emph{first} fully mechanised \picalc{} featuring linear types, as well as gradual and shared types.
-  We present its syntax, operational semantics, and typing rules and the corresponding type safety properties.
-  We use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
+  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
+  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, including \emph{session types} \cite{H93,THK94,HVK98}.
+  Linearity is key for type safety in session communication.
+
+  We present the \emph{first} full mechanisation in Agda of a \picalc{} with \emph{linear}, \emph{gradual} and \emph{shared} types, all under the same unified framework.
+  While linearity is key for type safety in communication-centric programming, gradual and shared types are needed to model real-world software systems.
+
+  We present syntax, semantics, type system and the corresponding type safety properties.
+  For the first time in \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
   We generalise the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, gradual and shared typing.
@@
   We show that the type system is stable under substitution and prove subject reduction.
-
-  Our formalisation is fully mechanised in Agda. \cite{Zalakain2020Agda}
+%
+  Our formalisation is fully mechanised in Agda \cite{Zalakain2020Agda}.
 \end{abstract}
@@
 
-We do so by defining a specification for an algebra on $0,1,n$ and $\omega$ \emph{multiplicities} (how many times a channel can be used) and use it to apply for the first time leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
+We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply for the first time leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
 The developer is able to choose any mix of algebras for the type system --- as long as they comply with the specification.

### 2020-05-04 paper/paper.tex
Draft on Goto et al

@@
 
-Most of the work on mechanising resource-aware calculi is very recent.
+Considerable efforts in the mechanisation of resource-aware calculi have been made in recent years.
 
@@
 
-\cite{Goto2016a}
-first session-type paper with a mechanised proof of type safety
-Session types
-Coq
-Extrinisically typed
-Session type polymorphism
+The work by Goto et al. \cite{Goto2016a} is likely the first contribution to session-types that comes along with a mechanised proof of type safety (in Coq). 
+Their paper extends session types with polymorphism and pattern matching on session types.
+They use a locally-nameless encoding of variable references, a syntax prior to types, and a LTS semantics that encodes session-typed processes into the \picalc{}.
+Their type system uses reorering of contexts and extrinsic contex splits.
 
@@
 \cite{Gay2001}
-
-\cite{previous-work} polymorphic tokens, HOAS
 

### 2020-05-04 paper/paper.tex
syntax and semantics

@@
 \label{syntax}
-Let $x, y,\ldots$ range over \emph{variables} and $P, Q,\ldots$ over processes.
-The syntax of the \picalc{} \cite{Sangio01} is given by grammar in \autoref{fig:syntax-names}.
-
 \begin{figure}[h]
@@
   \end{subfigure}
-  \begin{subfigure}[t]{.2\textwidth}
+  \begin{subfigure}[t]{.3\textwidth}
     \begin{equation*}
@@
     \end{equation*}
-    \caption{Ill-scoped grammar with names}
+    \caption{Ill-scoped grammar with channel names}
     \label{fig:syntax-names}
   \end{subfigure}
-  \hspace{.1\textwidth}
-  \begin{subfigure}[t]{.7\textwidth}
+  \hspace{.05\textwidth}
+  \begin{subfigure}[t]{.65\textwidth}
     \begin{equation*}
       \begin{split}
-        \Process_n ::=& \; \PO_n               &&\text{(process termination)}   \\ 
+        \Process_n ::=& \; \PO_n               &&\text{(inaction)}   \\ 
         |& \; \new{} \Process_{\suc n}         &&\text{(restriction)}          \\ 
-        |& \; \comp{\Process_n}{\Process_n}    &&\text{(parallel composition)} \\ 
+        |& \; \comp{\Process_n}{\Process_n}    &&\text{(parallel)} \\ 
         |& \; \recv{\Var_n}{}\Process_{\suc n} &&\text{(input)}                \\ 
@@
 
+Let $x, y,\ldots$ range over \emph{channel names} or \emph{variables} and $P, Q,\ldots$ over processes.
+The syntax of the \picalc{} \cite{Sangio01} is given by the grammar in \autoref{fig:syntax-names}.
 Process $\PO$ denotes the terminated process, where no operations can further occur.
@@
 
-Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names in $P$.
+Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names, respectively, $x$ and $y$ in $P$.
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
 Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
-Furthermore, names do not guarantee that the syntax well-scoped.
+Furthermore, names do not guarantee that the syntax is well-scoped.
+\todo{It is not clear what ill vs well scoped means. Add a sentece saying: "By ill-scoped we mean... By well-scoped we mean... "}
 Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
@@
       & (\new{}x ) && (\comp {\recvp{x}{y} && \; \sendp{y}{a} \; \PO} {(\new{} y) && (\sendp{x}{y} \; \recvp{y}{z} && \; \PO)})
-      && \text{names}
+      && \text{channel names}
       \\
       & \new{} && (\comp {\recv{0} && \; \send{0}{a} \; \PO} {\new{} && (\send{1}{0} \; \recv{0} && \; \PO)})
-      && \text{de Bruijn}
+      && \text{de Bruijn indices}
       \\
@@
 \end{example}
-That is, terms at different \emph{depths} use different indices to refer to the same binding.
+That is, terms at different \emph{depths} use different indices to refer to the same binding introduced either by a restriction $\constr{\nu}$ or input $\constr{()}$.
 
@@
 It is useful to rule out these nonsensical terms syntactically.
-In \autoref{fig:syntax}, we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
-
-Every time we go under a binder, the number of binders a variable might refer to increments by one.
+In \autoref{fig:syntax} (including the top), we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+Every time we go under a binder, the number of binders a variable might refer to increments by one, in rules $\constr{0}$ and $\suc n$. \todo{check this is right what i just said}
 To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
@@
 
-The semantics of our language can now be defined on the totality of the syntax.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{semantics}
-We introduce the operational semantics for the \picalc{}, starting with structural congruence.
+Due to a well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax. We start by defining the structural congruence.
 
@@
 As a result of that communication, the continuation of the input process $P$ gets all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one.
-To do so, we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$.
+After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (to do so, we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 
-We keep track of the channel over which communication occurs so that we can state subject reduction.
+We keep track of the channel over which communication occurs so that we can state subject reduction (in \autoref{subject-reduction}).
 We use a $\func{dec}$ function to decrement the index of this channel every time we come out of a binder.

### 2020-05-04 paper/paper.tex
mutliplicities

@@
 
-We characterise a usage algebra for our typing system in \autoref{multiplicities}.
-This algebra defines the way in which resources are \emph{split} on parallel composition; the way in which resources are \emph{consumed} on input and output.
-We then define typing and usage contexts in \autoref{contexts}.
+In \autoref{multiplicities} we characterise a usage algebra for our typing system. It defines how resources are \emph{split} in parallel composition and \emph{consumed} in input and output.
+We define typing and usage contexts in \autoref{contexts}.
 We provide a typing system for a resource-aware \picalc{} in \autoref{leftover-typing}.
@@
 \begin{note}
+  \label{note:linearity}
   We will often work with pairs of input and output multiplicities.
@@
   We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
-  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$, $\li$ for the pair $(1 \comma 0)$, $\lo$ for $(0 \comma 1)$ and $\lio$ for $(1 \comma 1)$, borrowing from the linear \picalc{} notation \cite{KPT96,Sangio01}.
+
+  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$ for a channel which cannot be used furether, $\li$ for the pair $(1 \comma 0)$ for a channel to be used in input only once, $\lo$ for $(0 \comma 1)$ for a channel to be used in output only once, and $\lio$ for $(1 \comma 1)$, for a channel to be used once in input and once in output.
+This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
 \end{note}
 
+\todo{you cannot refer to IDX if it is in the appendix. Assume the reader will never look at the appendix. Maybe it is worth bringing back fig 12 or 13? Can only one be brought back?}
 Linear, gradual and shared types (identified by a value in $\Idx$) are all defined as an instance of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.
-
+\todo{Explain what are Two and N. The whole sentence the carrier is a type Two is not clear...}
 \begin{description}
@@
     All other properties are lifted element-wise.
+    For example, a linear channel must be used exactly once as specified by the pair and as detailed in \autoref{note:linearity}.
 
@@
     All other rules follow trivially from the algebraic rules for the addition of natural numbers.
+    For example, a gradual channel can be used $x$ times in total: $y$ times for an action (input or output) and $z$ times for the co-action (output or input).
 
@@
     All other rules follow trivially.
+    For example, a shared channel can be used unrestrictedly many times in either input or output.
 \end{description}
-
+\todo{Added a last for example sentence to all. check}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 \label{contexts}
-
+\todo{unify typing system and type system. only use one terminology in the paper.}
 \begin{nidefinition}

### 2020-05-04 paper/paper.tex
on contexts

@@
 \end{nidefinition}
+\todo{again, usage and idx occur here, but the formal defintion seems to be actually necessary. I'm thinking that both fig 12 and 13 are needed in the main text...}
 
@@
 The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
+\todo{where is the def of CTX?}
 
 \begin{nidefinition}
-  \autoref{fig:prectx} defines types as either unit types, base types or channel types.
+  \autoref{fig:prectx} defines types as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
+  %
   The unit type $\unit$ serves as a proof of inhabitance for types.
   The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter.
-  This is for convenience, so that there is less universe polymorphism polluting specifications.
-  The type $\channel{t}{x}$ of a channel determines what type of data and what usage annotations are sent over that channel.
-  We keep track of types in a length-indexed $\PreCtx_n$.
+\todo{I don't get this...
+give an example for int, and just talk about pi, rather than a host language.}
+  %This is for convenience, so that there is less universe polymorphism polluting specifications.
+  The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
+This notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
+  \todo{I don't really like the symbol between t and x. Why not a simple semicolon ; ?}
+  We keep track of types in a length-indexed $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
 
@@
 \end{nidefinition}
-
+\todo{Maybe fig 6 can go in appendix in order to make room for 12 and 13 to come back? This whole paragraph from def 6 below could be cut and brought to app? and only commented upon}
 \begin{note}

### 2020-05-04 paper/paper.tex
Comment on the work by Simon Gay

@@
 
-Considerable efforts in the mechanisation of resource-aware calculi have been made in recent years.
-
 Leftover typing as presented in this paper is based on work by Allais \cite{Allais2018a}.
@@
 In this paper, we apply his technique to the \picalc{}, but limit ourselves to type-checking (scope restriction needs usage annotations) terms, and focus on generalising the usage algebra.
+Drawing from quantitative type theory by McBride and Atkey \cite{McBride2016, Atkey2018}, it is crucial for us too to be able to talk about fully consumed resources.
+As a result, we are able to transmit fully exhausted channels over channels.
+
+Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but the formalisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001} is perhaps one of the earliest works.
+Gay encodes the \picalc{} with linear and shared types into a more general framework.
+As we do, he uses de Bruijn indices, a reduction relation and a type system posterior to the syntax.
+However, in his work context splits are explicit, and the types of used multiplicities erased.
 
@@
 
-The work by Goto et al. \cite{Goto2016a} is likely the first contribution to session-types that comes along with a mechanised proof of type safety (in Coq). 
+The work by Goto et al. \cite{Goto2016a} is likely the first formal contribution to session-types that comes along with a mechanised proof of type safety (in Coq). 
 Their paper extends session types with polymorphism and pattern matching on session types.
@@
 \cite{Rouvoet2020}
-\cite{Veltri2020}
-\cite{Gay2001}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Future Work}
-
-Work that will be done time permitting:
-
-Erika Kreuer: model usage algebra as a state machine?
-
-Adding bidirectionality would provide us with a decision procedure for type checking.
-
-\begin{description}
-
-\item [Proof of progress]
-
-\item [Product types]
-
-\item [Sum types]
-
-\item [Decidable typechecking]
-
-  Bidirectional typing rules needed.
-
-\item [Soundness and completeness with respect to an alternative formalization.]
-
-\item [Encoding of session types]
-
-\end{description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+
+Mechanisation of the linear, gradual and shared \picalc{} in less than 1500 lines of code.
+
+Adding product types and sum types.
+Proving that progress preserves the balancing of channels.
+Erika Kreuer has suggested modeling the usage algebra as a state machine.
+Adding bidirectionality would provide us with a decision procedure for type checking for a given context of algebras.
 

### 2020-05-04 paper/paper.tex
notes on leftover typing

--- 2020-05-04 paper/paper.tex
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our typing rules, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
+We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
 This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
 These leftovers can then be used as an input to another typing derivation.
-This technique presents three main advantages:
+Leftover typing presents three main advantages:
 \begin{itemize}
@@
   \item
-    The need for an extensional context splitting proof for parallel composition is eradicated.
-    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    No need for an extensional context splitting proof for parallel composition.
+    Alternatively, such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
     This could be obtained through automated search --- but it is better yet not to need it.
@@
   \item
-    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated.
+    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated. \todo{explain suddenly}
 \end{itemize}
@@
 We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
-Both of these are indexed by a typing context, an input usage context, and an output usage context.
+Both are indexed by a typing context, an input usage context, and an output usage context.
 

### 2020-05-04 paper/paper.tex
typing with leftovers

@@
 We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
-Both are indexed by a typing context, an input usage context, and an output usage context.
+Both are indexed by a typing context ($\gamma$), an input usage context ($\Gamma$), and an output usage context or leftovers ($\Delta$).
+A typing context $\gamma$ keeps track of...  An input or output usage context $\Gamma,\Delta$ keeps track of...
 
@@
 
-  The \textsc{VarRef} relation presented in \autoref{fig:variable-references} keeps track of the index, type, and usage annotation of a variable within a particular typing and usage context.
-  The leftover usage context is the result of subtracting the usage annotation from the input usage context.
-
-  The base case $\constr{here}$ splits the usage annotation $x$ into $y$ and $z$, the leftovers.
-  This splitting is as per the usage algebra defined on the index of the usage annotation.
-
+  The \textsc{VarRef} relation is presented in \autoref{fig:variable-references}.
+  The judgment $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ reads as "..." 
+  The premise of \textsc{VarRef} means that...
+  \todo{use parts from the commented sentences to complete the dots; refer to the greek letters.}
+  \todo{Again, not sure I like the notation of this judgement...Instead of the square I would put $\triangleright$ If the relation is symmetric, then use a symmetric symbol, otherwise a symbol that conveys direction.
+Also, it seems as if the direction changes with that epsilon inverted.}
+  % keeps track of the index, type, and usage annotation of a variable within a particular typing and usage context.
+  % The leftover usage context is the result of subtracting the usage annotation from the input usage context.
+
+  The base case $\constr{here}$ splits the usage annotation $x$ into $y$ and $z$, which are the leftovers.
+  This splitting is as per the usage algebra defined on the index of the usage annotation in \autoref{}.
+  $\constr{there}$...
+
+  \todo{integrate the remark somewhere else, rather than in a remark?}
   \begin{remark}
@@
 \begin{example}
-  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
-  \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]
+  Let us illustrate the use of channels with an example.
+  \begin{lstlisting}[label=lst:channels,mathescape,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
   \end{lstlisting}
+  \autoref{lst:channels} asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
+  The continuation means that...
+  Finally, $= \constr{there} \; \constr{here}$ means that...
 \end{example}
@@
 
-  The typing relation \textsc{Types} presented in \autoref{fig:types} asserts that process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$.
+  The typing relation \textsc{Types} is presented in \autoref{fig:types}.
+  The typing judgement $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that "process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$".
 
@@
 
-  Process termination threads usage annotations unchanged from input to output.
+  Some comments on the leftover typing system follow.
+  The inaction process does not change usage annotations.
 
@@
   Parallel composition uses the leftovers of left-hand process to type the right-hand process.
+  \todo{explain the rationale here as most people would expect a combination/splitting of typing context, which for us is not needed anymore. just add a sentence}
 \end{nidefinition}
@@
   \label{example-derivations}
-  The example below uses linear and shared types to type the process $p$. 
+  Let us illustrate a derivation of typing with leftovers by the following example, which uses linear and shared types to type the process $p$. 
   We define $\omega$ as a $(\one \comma \one)$ usage in an algebra for shared types.
@@
   \end{lstlisting}
+  \todo{not easy to follow.
+you need to expalin each bit of this derivation.
+when you comment you need to refer to what you are commenting on.
+either put things in bracket when commenting or refer to line number or whatnot}
 \end{example}

### 2020-05-04 paper/paper.tex
on properties

@@
 \todo{intro}
-
+\todo{I wonder if we can remove the proofs everywhere and comment on them in the intro and where they can be found.}
 \subsection{Framing}
@@
 Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
-
+\todo{Do not forward refer when it comes to results.
+Never refer to results in the appendix.
+Consider that, if the paper get publshed, it will get so without the appendix.
+It should be *completely* self contained}
 \begin{nitheorem}
@@
 
+\todo{Maybe all previous subsections could be put into one "Auxiliary Results"}
+\todo{Replace the proofs with an intuition on the theorem}
 \subsection{Subject Congruence}
@@
 
+\todo{considering that Equals is in the appendix, you cannot do induction on something that is not defined in the main text. Bring the proof in the appendix. Maybe briefly comment on the proof here}
 \begin{proof}
@@
 \autoref{thm:substitution} instantiates the generalized version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
-
+\todo{Structure the theorem (in general the long ones) a bit more. Use itemize or similar. It's a blob of text.}
 \begin{nitheorem}
@@
 
+\todo{I don't think a reviewer from CONCUR most likely will nto get the proof. Appendix! Replace it with a proof sketch where you comment on the diagram as well.}
 \begin{proof}
@@
 
+\todo{maybe only focus on the first case, internal and external and say the rest is by induction blah bla and details can be found in appenix/or github}
 \begin{proof}

### 2020-05-04 paper/paper.tex
Correct little slip

--- 2020-05-04 paper/paper.tex
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
+We use \emph{leftover typing} \cite{Allais2018a} for our typing rules, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
 This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
 These leftovers can then be used as an input to another typing derivation.
-Leftover typing presents three main advantages:
+This technique presents three main advantages:
 \begin{itemize}
@@
   \item
-    No need for an extensional context splitting proof for parallel composition.
-    Alternatively, such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    The need for an extensional context splitting proof for parallel composition is eradicated.
+    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
     This could be obtained through automated search --- but it is better yet not to need it.
@@
   \item
-    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated. \todo{explain suddenly}
+    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated.
 \end{itemize}
@@
 We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
-Both are indexed by a typing context, an input usage context, and an output usage context.
+Both of these are indexed by a typing context, an input usage context, and an output usage context.
 
@@
 \begin{example}
-  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
+  The following example asserts that the input $\channel{\unit}{\li}$ type at index $\suc \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
   \begin{lstlisting}[mathescape,caption=Here the underscore \_ introduces an anonymous declaration immediately followed by its definition.]

### 2020-05-04 paper/paper.tex
First pass to the comments

--- 2020-05-04 paper/paper.tex
@@
   The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
-  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, including \emph{session types} \cite{H93,THK94,HVK98}.
+  It is an underlying theoretical and practical framework on top of which more advanced types and theories are built, including \emph{session types} \cite{H93,THK94,HVK98}.
   Linearity is key for type safety in session communication.
@@
 
-  We present syntax, semantics, type system and the corresponding type safety properties.
-  For the first time in \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
-  We generalise the algebras on multiplicities using indexed sets of \emph{partial commutative monoids}, allowing the user to choose a mix of linear, gradual and shared typing.
+  We present the syntax, semantics, type system and corresponding type safety properties.
+  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
+  We generalise the algebras on multiplicities, allowing the developer to choose a mix of linear, gradual and shared typing.
   We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
@@
 %
-In the meantime, with the rise of \emph{session types} \cite{H93,THK94,HVK98} linearity got a different flavour in the \picalc{}.
+In the meantime, with the rise of \emph{session types} \cite{H93,THK94,HVK98} linearity acquired a different flavour in the \picalc{}.
 Session types are a type formalism used in communication-centric programming.
-In session types, linearity ensures that a channel is owned by exactly one communicating participant, whilst the channel itself is used multiple times by following the structure of its session type.
+In session types, linearity ensures that a channel is owned by exactly one communicating participant, whilst the channel itself is used multiple times as per its session type.
 Linearity in session types is the key ingredient that guarantees type safety.
@@
 
-We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply for the first time leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
+We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
 The developer is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
@@
   \begin{itemize}
-    \item \textbf{Algebras on multiplicities}: The type system is \emph{resource-aware} in that it allows the user to provide a resource-aware algebra that is then applied to the type-system.
-    Linear types, gradual types and shared types are all instances of such a resource-aware algebra (in \autoref{multiplicities}).
+    \item \textbf{Algebras on multiplicities}: The type system is \emph{resource-aware} in that it allows the user to provide resource-aware algebras that are then applied to the type-system (in \autoref{multiplicities}).
+    Linear types, gradual types and shared types are all instances of such a resource-aware algebra.
     Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
@@
     Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
-    One of the benefits of this technique is that it removes the need for context splits, which is now not unnecessary.
+    One of the benefits of this technique is that it removes the need for extrinsic context splits, which are rendered unnecessary.
   \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The notation used in this paper is given in \autoref{fig:notation}.
+\autoref{fig:notation} illustrates the notation used in this paper.
 \begin{figure}[h]
@@
 
-Data type definitions use double lines and index-free synonyms as rule names (for ease of reference).
-For example ... \todo{complete the dots here, explain the rules above, in particular the 1+}
-We otherwise use the constructor name as the name of a typing rule. For example,...
-Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$. This means that ...
+Data type definitions ($\type{\N}$) use double lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
+We otherwise use constructor names ($\constr{0}$ and $\suc$) to name our typing rules.
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
 Implicit arguments are mentioned by type definitions but omitted in constructors.
@@
     \end{equation*}
-    \caption{Ill-scoped grammar with channel names}
+    \caption{Ill-scoped grammar with names}
     \label{fig:syntax-names}
@@
 
-Let $x, y,\ldots$ range over \emph{channel names} or \emph{variables} and $P, Q,\ldots$ over processes.
+Let $x, y,\ldots$ range over \emph{variables} (including channel names) and $P, Q,\ldots$ over processes.
 The syntax of the \picalc{} \cite{Sangio01} is given by the grammar in \autoref{fig:syntax-names}.
@@
 
-Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names, respectively, $x$ and $y$ in $P$.
+Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names --- $x$ and $y$ in $P$, respectively.
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
@@
       & (\new{}x ) && (\comp {\recvp{x}{y} && \; \sendp{y}{a} \; \PO} {(\new{} y) && (\sendp{x}{y} \; \recvp{y}{z} && \; \PO)})
-      && \text{channel names}
+      && \text{names}
       \\
@@
 \end{example}
-That is, terms at different \emph{depths} use different indices to refer to the same binding introduced either by a restriction $\constr{\nu}$ or input $\constr{()}$.
+That is, terms at different \emph{depths} use different indices to refer to the same binding.
 
@@
 It is useful to rule out these nonsensical terms syntactically.
-In \autoref{fig:syntax} (including the top), we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
-Every time we go under a binder, the number of binders a variable might refer to increments by one, in rules $\constr{0}$ and $\suc n$. \todo{check this is right what i just said}
-To propagate this information, we index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-We increase the \emph{depth} counter by 1 in channel restriction and input processes.
+In \autoref{fig:syntax}, we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+Every time we go under a binder, we increase the index of the continuation process, allowing the variables within to refer to one more thing.
 \autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction using type-level de Bruijn indices.
@@
 \label{semantics}
-Due to a well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax. We start by defining the structural congruence.
+Thanks to our well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax.
+We start by defining the structural congruence.
 
@@
 
-In \autoref{multiplicities} we characterise a usage algebra for our typing system. It defines how resources are \emph{split} in parallel composition and \emph{consumed} in input and output.
+In \autoref{multiplicities} we characterise a usage algebra for our typing system.
+It defines how resources are \emph{split} in parallel composition and \emph{consumed} in input and output.
 We define typing and usage contexts in \autoref{contexts}.
@@
 
-  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$ for a channel which cannot be used furether, $\li$ for the pair $(1 \comma 0)$ for a channel to be used in input only once, $\lo$ for $(0 \comma 1)$ for a channel to be used in output only once, and $\lio$ for $(1 \comma 1)$, for a channel to be used once in input and once in output.
+  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$ for a channel which cannot be used any further, $\li$ for the pair $(1 \comma 0)$ for a channel to be used in input exactly once, $\lo$ for $(0 \comma 1)$ for a channel to be used in output exactly once, and $\lio$ for $(1 \comma 1)$, for a channel to be used exactly once in input and exactly once in output.
   This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
@@
 \todo{you cannot refer to IDX if it is in the appendix. Assume the reader will never look at the appendix. Maybe it is worth bringing back fig 12 or 13? Can only one be brought back?}
+
 Linear, gradual and shared types (identified by a value in $\Idx$) are all defined as an instance of our usage algebra.
@@
     The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited in every other case.
-    All other properties are lifted element-wise.
-    For example, a linear channel must be used exactly once as specified by the pair and as detailed in \autoref{note:linearity}.
+    All the properties of the algebra follow trivially.
+    As an example, the multiplicity pair $\ell_i$ is represented as $(\constr{one} \comma \constr{zero})$.
 
@@
     The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
-    All other rules follow trivially from the algebraic rules for the addition of natural numbers.
-    For example, a gradual channel can be used $x$ times in total: $y$ times for an action (input or output) and $z$ times for the co-action (output or input).
+    All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
+    As an example, we use $(5 \comma 3)$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.
 
@@
     The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
-    All other rules follow trivially.
-    For example, a shared channel can be used unrestrictedly many times in either input or output.
+    All the properties of the algebra follow trivially.
+    All shared channels have multiplicities $(\omega \comma \omega)$, which they will forever preserve.
 \end{description}
-\todo{Added a last for example sentence to all. check}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 \label{contexts}
+
 \todo{unify typing system and type system. only use one terminology in the paper.}
+
 \begin{nidefinition}
@@
 
-We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
+We keep typing contexts ($\PreCtx$, in \autoref{fig:prectx}) and usage contexts ($\Ctx$ in \autoref{fig:ctx}) separate.
 The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
-\todo{where is the def of CTX?}
 
@@
   \autoref{fig:prectx} defines types as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
-  %
   The unit type $\unit$ serves as a proof of inhabitance for types.
-  The base type $\base{n}$ uses natural numbers as placeholders for types --- the host language can interpret the former into the latter.
-\todo{I don't get this...
-give an example for int, and just talk about pi, rather than a host language.}
-  %This is for convenience, so that there is less universe polymorphism polluting specifications.
+  The base type $\base{n}$ uses natural numbers as placeholders for types, the host language can then interpret the former into the latter --- e.g. $0$ as booleans, $1$ as natural numbers.
+  This is for convenience, so that there is less universe polymorphism polluting specifications.
   The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
@@
 \end{nidefinition}
+
 \todo{Maybe fig 6 can go in appendix in order to make room for 12 and 13 to come back? This whole paragraph from def 6 below could be cut and brought to app? and only commented upon}
+
 \begin{note}
@@
     No need for an extensional context splitting proof for parallel composition.
-    Alternatively, such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
     This could be obtained through automated search --- but it is better yet not to need it.
@@
   \item
-    New theorems (\textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening})) can suddenly be stated. \todo{explain suddenly}
+    New theorems like \textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening}) can be stated thanks to the presence of a leftover context.
 \end{itemize}
@@
 We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
-Both are indexed by a typing context ($\gamma$), an input usage context ($\Gamma$), and an output usage context or leftovers ($\Delta$).
-A typing context $\gamma$ keeps track of...  An input or output usage context $\Gamma,\Delta$ keeps track of...
+Both are indexed by a $\PreCtx_n$, an input $\Ctx_{idxs}$, and an output $\Ctx_{idxs}$ (the leftovers).
 
@@
   The \textsc{VarRef} relation is presented in \autoref{fig:variable-references}.
-  The judgment $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ reads as "..." 
-  The premise of \textsc{VarRef} means that...
-  \todo{use parts from the commented sentences to complete the dots; refer to the greek letters.}
+  The judgment $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ reads as ``$\gamma$ at position $i$ is of type $t$, and subtracting $y$ at position $i$ from $\Gamma$ results in $\Delta$''.
+
   \todo{Again, not sure I like the notation of this judgement...Instead of the square I would put $\triangleright$ If the relation is symmetric, then use a symmetric symbol, otherwise a symbol that conveys direction.
 Also, it seems as if the direction changes with that epsilon inverted.}
-  % keeps track of the index, type, and usage annotation of a variable within a particular typing and usage context.
-  % The leftover usage context is the result of subtracting the usage annotation from the input usage context.
-
-  The base case $\constr{here}$ splits the usage annotation $x$ into $y$ and $z$, which are the leftovers.
-  This splitting is as per the usage algebra defined on the index of the usage annotation in \autoref{}.
-  $\constr{there}$...
-
-  \todo{integrate the remark somewhere else, rather than in a remark?}
-  \begin{remark}
-    We use the functionality of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
-  \end{remark}
+
+  The base case $\constr{here}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
+  This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
+  We use the functionality of the monoidal relation to alleviate the user from this proof burden.
+  The inductive case $\constr{there}$ preserves the usage annotation $x'$ from input to output.
 
@@
   \begin{lstlisting}[label=lst:channels,mathescape,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
-  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\constr{suc} \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
+  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
   _ = $\constr{there} \; \constr{here}$
   \end{lstlisting}
-  \autoref{lst:channels} asserts that the input $\channel{\unit}{\li}$ type at index $\constr{suc} \; \constr{0}$ is capable of $\li$, and that consuming such $\li$ results in $\lo$.
-  The continuation means that...
-  Finally, $= \constr{there} \; \constr{here}$ means that...
+  In \autoref{lst:channels}, the constructor $\constr{there}$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
+  The constructor $\constr{here}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving leftovers $\lo$.
 \end{example}
@@
   The typing relation \textsc{Types} is presented in \autoref{fig:types}.
-  The typing judgement $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that "process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$".
+  The typing judgement $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
 
@@
 
-  Parallel composition uses the leftovers of left-hand process to type the right-hand process.
-  \todo{explain the rationale here as most people would expect a combination/splitting of typing context, which for us is not needed anymore. just add a sentence}
+  Parallel composition uses the leftovers of the left-hand process to type the right-hand process.
+  By keeping track in the typing derivation of $P$ of the resources it uses, we save the user from having to provide an extrinsic proof of context split.
 \end{nidefinition}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Related Work}
+\section{Conclusion, Related and Future Work}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-
+To conclude, in this paper...
 Mechanisation of the linear, gradual and shared \picalc{} in less than 1500 lines of code.
 
+As future work we intend to...
 Adding product types and sum types.

### 2020-05-04 paper/paper.tex
Minor improvements

@@
 
-\todo{you cannot refer to IDX if it is in the appendix. Assume the reader will never look at the appendix. Maybe it is worth bringing back fig 12 or 13? Can only one be brought back?}
-
-Linear, gradual and shared types (identified by a value in $\Idx$) are all defined as an instance of our usage algebra.
+Linear, gradual and shared types are all defined as an instance of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.
-\todo{Explain what are Two and N. The whole sentence the carrier is a type Two is not clear...}
 \begin{description}
   \item [Linear]
-    The carrier is a type $\type{Two}$ with the trivial elements $\constr{zero}$ and $\constr{one}$.
+    The carrier is a type with two trivial elements $\constr{zero}$ and $\constr{one}$.
     The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited in every other case.
@@
     \item [Gradual]
-    The carrier is the type of natural numbers $\N$.
+    The carrier is the type of natural numbers.
     The element $\zero$ corresponds to 0, and $\one$ to 1.
@@
     \item [Shared]
-    The carrier is a type $\type{One}$ with a single trivial constructor $\constr{\omega}$.
+    The carrier is a type with a single trivial constructor $\constr{\omega}$.
     Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
@@
   We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a set of indices $\Idx$ together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras on that type.
+  An indexed set of usage algebras is a type $\Idx$ of indices together with an interpretation $\Usage_{idx} : \Idx \to \Set$ of indices into types, and an interpretation of indices into usage algebras.
   $\Idx$ must be non-empty.

### 2020-05-04 paper/paper.tex
Deal with the easy issues

@@
 \newcommand{\base}[1]{\constr{B[} \; #1 \; \constr{]}}
-\newcommand{\channel}[2]{\constr{C[} \; #1 \; \constr{\propto} \; #2 \; \constr{]}}
+\newcommand{\channel}[2]{\constr{C[} \; #1 \; \constr{;} \; #2 \; \constr{]}}
 \newcommand{\comma}{\; \constr{,} \;}
@@
 \newcommand{\reduce}[1]{\; \type{\longrightarrow}_{#1} \;}
-\newcommand{\types}[4]{#1 \; \type{\propto} \; #2 \; \type{\vdash} \; #3 \; \type{\boxtimes} \; #4}
-\newcommand{\contains}[6]{#1 \; \type{\propto} \; #2 \; \type{\ni}_{#3} \; #4 \; \type{\propto} \; #5 \; \type{\boxtimes} \; #6}
-\newcommand{\containsusage}[4]{#1 \; \type{\ni}_{#2} \; #3 \; \type{\boxtimes} \; #4}
+\newcommand{\types}[4]{#1 \; \type{;} \; #2 \; \type{\vdash} \; #3 \; \type{\triangleright} \; #4}
+\newcommand{\contains}[6]{#1 \; \type{;} \; #2 \; \type{\ni}_{#3} \; #4 \; \type{;} \; #5 \; \type{\triangleright} \; #6}
+\newcommand{\containsusage}[4]{#1 \; \type{\ni}_{#2} \; #3 \; \type{\triangleright} \; #4}
 \newcommand{\Var}{\type{VAR}}
@@
   \end{mathpar}
-  \caption{Notation used in this paper}
+  \caption{Notation used in this paper.}
   \label{fig:notation}
@@
     \end{equation*}
-    \caption{Ill-scoped grammar with names}
+    \caption{Ill-scoped grammar with names.}
     \label{fig:syntax-names}
@@
     \end{equation*}
-    \caption{Well-scoped grammar with type-level de Bruijn indices}
+    \caption{Well-scoped grammar with type-level de Bruijn indices.}
     \label{fig:syntax}
@@
 Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
-Furthermore, names do not guarantee that the syntax is well-scoped.
-\todo{It is not clear what ill vs well scoped means. Add a sentece saying: "By ill-scoped we mean... By well-scoped we mean... "}
 Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
@@
 References outside of that range are meaningless.
-It is useful to rule out these nonsensical terms syntactically.
+It is useful to rule out these ill-scoped nonsensical terms syntactically.
 In \autoref{fig:syntax}, we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-Every time we go under a binder, we increase the index of the continuation process, allowing the variables within to refer to one more thing.
-\autoref{fig:syntax} showcases a \picalc{} syntax that is well-scoped by construction using type-level de Bruijn indices.
+Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within it to refer to one more thing.
 
@@
 Thanks to our well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax.
-We start by defining the structural congruence.
+We start by defining structural congruence.
 
@@
 
-In \autoref{multiplicities} we characterise a usage algebra for our typing system.
+In \autoref{multiplicities} we characterise a usage algebra for our type system.
 It defines how resources are \emph{split} in parallel composition and \emph{consumed} in input and output.
 We define typing and usage contexts in \autoref{contexts}.
-We provide a typing system for a resource-aware \picalc{} in \autoref{leftover-typing}.
+We provide a type system for a resource-aware \picalc{} in \autoref{leftover-typing}.
 
@@
 
-\todo{unify typing system and type system. only use one terminology in the paper.}
-
 \begin{nidefinition}
@@
   This notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
-  \todo{I don't really like the symbol between t and x. Why not a simple semicolon ; ?}
   We keep track of types in a length-indexed $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
@@
 
-  \todo{Again, not sure I like the notation of this judgement...Instead of the square I would put $\triangleright$ If the relation is symmetric, then use a symmetric symbol, otherwise a symbol that conveys direction.
-Also, it seems as if the direction changes with that epsilon inverted.}
-
-  The base case $\constr{here}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
+  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
   We use the functionality of the monoidal relation to alleviate the user from this proof burden.
-  The inductive case $\constr{there}$ preserves the usage annotation $x'$ from input to output.
+  The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.
 
@@
     {\opsquared{x}{y}{z}}
-    {\constr{here} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{0}}{t}{y}{\Gamma \comma z}}
+    {\constr{0} : \contains{\gamma \comma t}{\Gamma \comma x}{\constr{0}}{t}{y}{\Gamma \comma z}}
 
@@
     {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
-    {\constr{there} \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
+    {\suc \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
   \end{mathpar}
@@
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
-  _ = $\constr{there} \; \constr{here}$
+  _ = $\suc \; \constr{0}$
   \end{lstlisting}
-  In \autoref{lst:channels}, the constructor $\constr{there}$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{here}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving leftovers $\lo$.
+  In \autoref{lst:channels}, the constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving leftovers $\lo$.
 \end{example}
@@
             chan_i : \contains{\gamma \hspace{1.2em}}{\Gamma \hspace{1.3em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{5.6em}}{\Theta \comma \lz}}}
+            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.6em}}{\Theta \comma \lz}}}
         {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
@@
             chan_i : \contains{\gamma}{\Gamma \hspace{0.2em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t \hspace{3.7em}}{x \hspace{0.3em}}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{5.8em}}{\Theta}}}
+            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t \hspace{3.2em}}{x \hspace{0.3em}}{\Xi} \\\\
+            cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{4.8em}}{\Theta}}}
         {\constr{send} \; chan_i \; loc_j \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
@@
 
-  Some comments on the leftover typing system follow.
+  Some comments on the leftover type system follow.
   The inaction process does not change usage annotations.
@@
   \label{example-derivations}
-  Let us illustrate a derivation of typing with leftovers by the following example, which uses linear and shared types to type the process $p$. 
-  We define $\omega$ as a $(\one \comma \one)$ usage in an algebra for shared types.
-  We define $\epsilon$ as the empty usage context where all usages are $(\zero \comma \zero)$.
-  The typing derivation creates a new channel with usage $(\one \comma \one)$.
-  This channel transmits $\li$ multiplicities of a type $\channel{\unit}{\omega}$.
-  The right-hand subprocess creates a channel with usage $(\one \comma \one)$.
-  This channel transmits $\omega$ multiplicities of a type $\unit$.
-  Part of this channel ($(\one \comma \zero)$) is sent over the channel created by the parent process.
+  Let us type a derivation of the process $\func{p}$ using linear and shared types. 
+  Let $\omega$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
+  Let $\epsilon$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
+  To type $\func{p}$, we need a channel over which we can transmit another channel.
+  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transiting data of type $\channel{\unit}{\omega}$, and declare (with $\li$) that we will only transmit its input capability.
+  The left-hand process is typed following its syntax.
+  On the right-hand, we need the channel that is going to be sent.
+  The type $\unit$ of this channel is entirely determined by the type of the channel over which it is going to be transmited.
+  Its usage annotation is $(\one \comma \one)$: we will transmit $\li$ and save $\lo$ for the continuation.
 
   \begin{lstlisting}[mathescape]
-  $p$ : $\Process_1$
-  $p$ = $\new{} (\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}
-                    {\new{} (\send{\constr{suc} \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\constr{suc} \; \constr{suc} \; \constr{0}}) \; \PO)})$
-
-  _ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \omega}{p}{\epsilon}$
-  _ = $\constr{chan} \; \channel{\unit}{\omega} \; \li \; \one$
-      $(\constr{comp} \;
-       (\constr{recv} \; \constr{here} \; (\constr{recv} \; \constr{here} \; \constr{end}))$
-           $(\constr{chan} \; \unit \; \omega \; \one \; (\constr{send} \; (\constr{there} \; \constr{here}) \; \constr{here} \; (\constr{send} \; \constr{here} \; (\constr{there} \; \constr{here} \; \constr{there}) \; \constr{end})))
+  $\func{p}$ : $\Process_1$
+  $\func{p}$ = $\new{} \; ($
+       $\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}$
+       ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}}) \; \PO)})$
+
+  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \omega}{\func{p}}{\epsilon}$
+  $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\omega} \; \li \; \one \; (\constr{comp}$
+       $(\constr{recv} \; \constr{0} \; (\constr{recv} \; \constr{0} \; \constr{end}))$
+       $(\constr{chan} \; \unit \; \omega \; \one \; (\constr{send} \; (\suc \; \constr{0}) \; \constr{0} \; (\constr{send} \; \constr{0} \; (\suc \; \suc \; \constr{0}) \; \constr{end})))
        )$
   \end{lstlisting}
-  \todo{not easy to follow.
-you need to expalin each bit of this derivation.
-when you comment you need to refer to what you are commenting on.
-either put things in bracket when commenting or refer to line number or whatnot}
 \end{example}
@@
     \item
-      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\constr{there}$.
+      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\suc$.
 

### 2020-05-04 paper/paper.tex
Get rid of rearrangement figure

@@
 \todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
-\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modeling is modelling}
+\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modelling is modelling}
 \todo{rename section and subsections in reference to \S (not sure why the commands above are not doing it so check later)}
@@
 At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types.
-This makes the \picalc{} a fully-fledged tool for modeling and verification of concurrent and distributed systems.
+This makes the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
 
@@
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
-  We use the functionality of the monoidal relation to alleviate the user from this proof burden.
+  We use the \emph{functionality} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.
@@
     \item
-    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions in \autoref{fig:subject-reduction:a} into the proofs in \autoref{fig:subject-reduction:b}.
-    We then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
-
-    \begin{figure}[h]
-      \begin{subfigure}{.5\textwidth}
-        \begin{alignat*}{2}
-          \containsusage{&\Gamma'}{i}{\lio &&}{\Gamma} \\
-          \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\li &&}{\Psi} \\
-          \types{\gamma \comma t}{&\Psi \comma m}{P &&}{\Delta \comma \lz} \\
-          \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\lo &&}{\Theta} \\
-          \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
-        \end{alignat*}
-        \caption{Initial assumptions.}
-        \label{fig:subject-reduction:a}
-      \end{subfigure}
-      % Do not leave empty space
-      \begin{subfigure}{.5\textwidth}
-        \begin{alignat*}{2}
-          \types{\gamma \comma t}{& \Gamma \comma m}{P &&}{\Theta \comma \lz} \\
-          \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
-        \end{alignat*}
-        \caption{Shift witness for $\lo$ and merge with the one for $\li$ into $\lio$.}
-        \label{fig:subject-reduction:b}
-      \end{subfigure}
-    \end{figure}
+    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
 
@@
 Proving that progress preserves the balancing of channels.
-Erika Kreuer has suggested modeling the usage algebra as a state machine.
+Erika Kreuer has suggested modelling the usage algebra as a state machine.
 Adding bidirectionality would provide us with a decision procedure for type checking for a given context of algebras.

### 2020-05-04 paper/paper.tex
Discard unnecessary lemma, group auxiliary results

@@
   \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
-  Leftover typing allows \emph{framing} (\autoref{framing}), \emph{weakening} (\autoref{weakening}) and \emph{strengthening} (\autoref{strengthening}) theorems to be stated.
+  Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
   We use these to prove \emph{subject congruence} (\autoref{subject-congruence}), and together with \emph{substitution} (\autoref{substitution}), to prove \emph{subject reduction} (\autoref{subject-reduction}).
@@
   \item
-    New theorems like \textbf{framing} (\autoref{framing}), \textbf{weakening} (\autoref{weakening}) and \textbf{strengthening} (\autoref{strengthening}) can be stated thanks to the presence of a leftover context.
+    New theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can be stated thanks to the presence of a leftover context.
 \end{itemize}
@@
 \todo{I wonder if we can remove the proofs everywhere and comment on them in the intro and where they can be found.}
-\subsection{Framing}
-\label{framing}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
-\todo{Do not forward refer when it comes to results.
-Never refer to results in the appendix.
-Consider that, if the paper get publshed, it will get so without the appendix.
-It should be *completely* self contained}
+
+\subsection{Auxiliaty Results}
+\label{auxiliary-results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{nitheorem}
   \label{thm:framing}
+  Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
   Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
-  Then, per \autoref{lm:types-op}, there exists some $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Let there be a $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
   Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
@@
 
-\subsection{Weakening}
-\label{weakening}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
-
 \begin{nitheorem}
   \label{thm:weakening}
+  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+
   Let $\func{insert}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
@@
 
-\subsection{Strengthening}
-\label{strengthening}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Removing an unused variable preserves the well-typedness of a process.
-We say a variable is \emph{unused} within a process if neither inputs nor outputs refer to it.
-
-\begin{remark}
-  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned within $P$ --- $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
-\end{remark}
-
 \begin{nitheorem}
   \label{thm:strengthening}
+  Removing an unused variable preserves the well-typedness of a process.
+
   Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
@@
 
-\subsection{Swapping}
-\label{swapping}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Swapping two variables preserves the well-typedness of a process.
-
 \begin{remark}
-  We limit ourselves to the swapping of adjacent variables $i$ and $\suc i$.
-  Doing so is enough to prove subject reduction.
-  More generally, any two variables can be swapped by performing a series of adjacent variable swappings.
+  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned within $P$ --- $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
 \end{remark}
@@
   \label{thm:swapping}
+  Swapping two variables preserves the well-typedness of a process.
   Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
@@
 
-\todo{Maybe all previous subsections could be put into one "Auxiliary Results"}
+\begin{remark}
+  We limit ourselves to the swapping of adjacent variables $i$ and $\suc i$.
+  Doing so is enough to prove subject reduction.
+  More generally, any two variables can be swapped by performing a series of adjacent variable swappings.
+\end{remark}
+
 \todo{Replace the proofs with an intuition on the theorem}
@@
 
-\todo{maybe only focus on the first case, internal and external and say the rest is by induction blah bla and details can be found in appenix/or github}
 \begin{proof}

### 2020-05-04 paper/appendix.tex
Add subst-unused lemma, fix reduction
Minor improvements
Discard unnecessary lemma, group auxiliary results
Move formal def of algebra and indexed algs back into the paper

@@
 
-\section{Usage Algebra}
-\label{app:usage-algebra}
+\section{Lemmas}
+\label{app:lemmas}
 
-\todo{we haven't introduced the forall and implicit syntax}
-\todo{introduce DEC and exists}
-
-Our usage algebra is a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancelative}, \emph{associtive}, and \emph{commutative}.
-It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
-It has an element $\one$ that is used to count inputs and outputs.
-These algebraic laws are defined as a record $\Quantifier_C$ on a carrier set $C$ in \autoref{fig:multiplicities}.
-We use a terniary relation to model the partial monoidal operation.
-We then encapsulate a set carriers and their algebras in \autoref{fig:indexed-multiplicities}.
-
-\begin{figure}[h]
-\begin{equation}
-\begin{aligned}
-  &\zero                  &:{} &                      &        & C \\
-  &\one                   &:{} &                      &        & C \\
-  &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to et \\
-  &\func{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
-  &\func{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-  &\func{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-  &\func{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
-  &\func{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
-  &\func{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-  &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
-\end{aligned}
-\end{equation}
-\caption{Quantifier algebra $\Quantifier_C$ algebra on a partial commutative monoid.}
-\label{fig:multiplicities}
-\end{figure}
-
-\begin{figure}[h]
-  \begin{equation}
-    \begin{split}
-      &\func{IDX}          &:{} &\Set \\
-      &\func{\exists IDX}  &:{} &\func{IDX} \\
-      &\func{USAGE}        &:{} &\func{IDX} \to \Set \\
-      &\func{QUANTIFIERS}  &:{} &(i : \func{IDX}) \to \Quantifier_{\func{USAGE}_i}
-    \end{split}
-  \end{equation}
-  \caption{Indexed set of usage algebras.}
-  \label{fig:indexed-multiplicities}
-\end{figure}
-
-\section{Lemmas for Type Safety}
-\label{app:lemmas-type-safety}
+\begin{nilemma}
+  \label{lm:subst-unused}
+  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 \begin{nilemma}
-  \label{lm:types-op}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, there exists some $\Delta$ such that $\opctx{\Gamma}{\Delta}{\Xi}$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
   \label{lm:comm-capable}

### 2020-05-04 paper/paper.tex
Move formal def of algebra and indexed algs back into the paper

@@
 \newcommand{\Rec}{\type{REC}}
-\newcommand{\Quantifier}{\type{QUANTIFIER}}
+\newcommand{\Algebra}{\type{ALGEBRA}}
 \newcommand{\eq}[1]{\; \type{=}_{#1} \;}
@@
 In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise over this notion by defining an algebra that is satisfied by linear, gradual and shared types.
+We generalise over this notion by defining an algebra that is satisfied by linear, gradual and shared types alike.
 
@@
   It has an element $\one$ that is used to count inputs and outputs.
-  For a formal definition, refer to \autoref{fig:multiplicities} in \autoref{app:usage-algebra}.
+  \autoref{fig:multiplicities} uses a terniary relation to model such an algebra as a record $\Algebra_C$ on a carrier $C$.
+
+  \todo{we haven't introduced the forall and implicit syntax}
+  \todo{introduce DEC and exists}
+
+  \begin{figure}[h]
+  \begin{equation*}
+  \begin{aligned}
+    &\zero                  &:{} &                      &        & C \\
+    &\one                   &:{} &                      &        & C \\
+    &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
+    &\func{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+    &\func{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+    &\func{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+    &\func{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+    &\func{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
+    &\func{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+    &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+  \end{aligned}
+  \end{equation*}
+  \caption{Algebra $\Algebra_C$ algebra on a partial commutative monoid.}
+  \label{fig:multiplicities}
+  \end{figure}
 \end{nidefinition}
@@
   We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a type $\Idx$ of indices together with an interpretation $\Usage_{idx} : \Idx \to \Set$ of indices into types, and an interpretation of indices into usage algebras.
+  An indexed set of usage algebras is a nonempty type $\Idx$ of indices together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras.
   $\Idx$ must be non-empty.
-  Refer to \autoref{fig:indexed-multiplicities} in \autoref{app:usage-algebra} for a formal definition.
+  Refer to \autoref{fig:indexed-multiplicities} for a formal definition.
+  \begin{figure}[h]
+    \begin{equation}
+      \begin{split}
+        &\Idx               &:{} &\Set \\
+        &\type{\exists IDX} &:{} &\Idx \\
+        &\Usage             &:{} &\Idx \to \Set \\
+        &\type{ALGEBRAS}    &:{} &(idx : \Idx) \to \Algebra_{\Usage_{idx}}
+      \end{split}
+    \end{equation}
+    \caption{Indexed set of usage algebras.}
+    \label{fig:indexed-multiplicities}
+  \end{figure}
+
 \end{nidefinition}
-\todo{again, usage and idx occur here, but the formal defintion seems to be actually necessary. I'm thinking that both fig 12 and 13 are needed in the main text...}
 
@@
   Some comments on the leftover type system follow.
+
   The inaction process does not change usage annotations.
@@
   $\func{p}$ : $\Process_1$
-  $\func{p}$ = $\new{} \; ($
-       $\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}$
+  $\func{p}$ = $\new{}$
+       $(\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}$
        ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}}) \; \PO)})$

### 2020-05-04 paper/paper.tex
Minor stuff

@@
 
-  \todo{we haven't introduced the forall and implicit syntax}
-  \todo{introduce DEC and exists}
-
   \begin{figure}[h]
@@
   \end{figure}
+  \todo{we haven't introduced the forall and implicit syntax}
+  \todo{introduce DEC and exists}
 \end{nidefinition}
@@
   We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a nonempty type $\Idx$ of indices together with an interpretation $\Usage_{idx}$ of indices into types, and an interpretation of indices into usage algebras.
-  $\Idx$ must be non-empty.
-  Refer to \autoref{fig:indexed-multiplicities} for a formal definition.
+  An indexed set of usage algebras is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
   \begin{figure}[h]
@@
   The unit type $\unit$ serves as a proof of inhabitance for types.
-  The base type $\base{n}$ uses natural numbers as placeholders for types, the host language can then interpret the former into the latter --- e.g. $0$ as booleans, $1$ as natural numbers.
-  This is for convenience, so that there is less universe polymorphism polluting specifications.
+  The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g.
+$0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
   The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
@@
 \end{nidefinition}
-
-\todo{Maybe fig 6 can go in appendix in order to make room for 12 and 13 to come back? This whole paragraph from def 6 below could be cut and brought to app? and only commented upon}
 
@@
   Let us type a derivation of the process $\func{p}$ using linear and shared types. 
-  Let $\omega$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
+  Let $\func{\omega}$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
   Let $\epsilon$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
   To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transiting data of type $\channel{\unit}{\omega}$, and declare (with $\li$) that we will only transmit its input capability.
+  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transiting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability.
   The left-hand process is typed following its syntax.
@@
 
-  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \omega}{\func{p}}{\epsilon}$
-  $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\omega} \; \li \; \one \; (\constr{comp}$
+  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\epsilon}$
+  $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\func{\omega}} \; \li \; \one \; (\constr{comp}$
        $(\constr{recv} \; \constr{0} \; (\constr{recv} \; \constr{0} \; \constr{end}))$
-       $(\constr{chan} \; \unit \; \omega \; \one \; (\constr{send} \; (\suc \; \constr{0}) \; \constr{0} \; (\constr{send} \; \constr{0} \; (\suc \; \suc \; \constr{0}) \; \constr{end})))
+       $(\constr{chan} \; \unit \; \func{\omega} \; \one \; (\constr{send} \; (\suc \; \constr{0}) \; \constr{0} \; (\constr{send} \; \constr{0} \; (\suc \; \suc \; \constr{0}) \; \constr{end})))
        )$
@@
   \label{thm:weakening}
-  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+  Weakening captures the idea that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
 
@@
   \label{thm:strengthening}
-  Removing an unused variable preserves the well-typedness of a process.
+  Strengthening captures the idea that removing an unused variable preserves the well-typedness of a process.
 

### 2020-05-05 paper/paper.tex
Solve some more todos

@@
 %% For section renaming in autoref
-\addto\extrasenglish{%
+\addto\extrasUKenglish{%
 \renewcommand{\sectionautorefname}{\S}%
@@
 \ccsdesc[300]{Theory of computation~Process calculi}
-\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanization, Agda}
+\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
 \supplement{\url{https://github/umazalakain/typing-linear-pi}}
@@
 
+\todo{spellcheck}
 \todo{make sure all citations are there}
@@
 \todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
-\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modelling is modelling}
-\todo{rename section and subsections in reference to \S (not sure why the commands above are not doing it so check later)}
 We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
@@
     &\op{\_}{\_}{\_}        &:{} &                      &        & C \to C \to C \to \Set \\
-    &\func{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x . \; (\op{x}{y}{z})) \\
+    &\func{\cdot-compute}  &:{} &\forall y z           & \to \; & \type{DEC} \; (\type{\exists} x  \; (\op{x}{y}{z})) \\
     &\func{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
@@
     &\func{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-    &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w . \; (\op{x}{u}{w} \times \op{w}{v}{z})
+    &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
   \end{aligned}
   \end{equation*}
-  \caption{Algebra $\Algebra_C$ algebra on a partial commutative monoid.}
+  \caption{
+    Algebra $\Algebra_C$ algebra on a partial commutative monoid.
+    In Agda $\forall$ introduces universal quantification, while $\{ \text{brackets} \}$ make the variables within them implicit.
+    The dependent product $\type{\exists}$ uses the value of its first argument in the type of its second.
+    The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.
+  }
   \label{fig:multiplicities}
   \end{figure}
-  \todo{we haven't introduced the forall and implicit syntax}
-  \todo{introduce DEC and exists}
 \end{nidefinition}
@@
 Drawing from quantitative type theory by McBride and Atkey \cite{McBride2016, Atkey2018}, it is crucial for us too to be able to talk about fully consumed resources.
-As a result, we are able to transmit fully exhausted channels over channels.
+As a by-product, we are able to transmit over a channel all none multiplicities of a fully exhausted channel over.
 

### 2020-05-05 paper/paper.tex
a pass on contributions

@@
 As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-
+\todo{rename gradual to something else...}
 The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
@@
  \begin{itemize}
-   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is well-scoped by construction: every free variable is accounted for in the type of the process that uses it.
+   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is \emph{well-scoped by construction}: every free variable is accounted for in the type of the process that uses it.
 
    \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
-   This operational semantics is given as a reduction relation on processes (in \autoref{operational-semantics}).
-   The reduction relation tracks at the type level what channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (in \autoref{subject-reduction}).
-   The reduction relation is defined modulo structural congruence (in \autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+   This operational semantics is given as a reduction relation on processes (\autoref{operational-semantics}).
+   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (\autoref{subject-reduction}).
+   The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
  \end{itemize}
 
-  \item \textbf{Leftover typing for the \picalc{}, with linear, gradual and shared types}:
+  \item \textbf{Leftover typing for the \picalc{} with linear, gradual and shared types}:
   \autoref{type-system} provides the type system for a resource-aware \picalc{}.
   \begin{itemize}
-    \item \textbf{Algebras on multiplicities}: The type system is \emph{resource-aware} in that it allows the user to provide resource-aware algebras that are then applied to the type-system (in \autoref{multiplicities}).
-    Linear types, gradual types and shared types are all instances of such a resource-aware algebra.
+    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
+    Linear types, gradual types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
     Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
-    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (in \autoref{contexts}).
-
-    \item \textbf{Leftover typing}: The typing judgments use \emph{leftover typing} to model the resource-aware \picalc{} (in \autoref{leftover-typing}).
+
+    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
+    This allows for type systems combining linear, gradual and shared types.
+
+    \item \textbf{\picalc{} with leftovers}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
     This approach adds a leftover usage context to the typing judgments.
@@
   \item \textbf{Fully mechanised formalisation available in Agda}:
-  This work has been fully formalised in Agda and is publicly available \cite{Zalakain2020Agda}.
+  This work has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
 

### 2020-05-05 paper/paper.tex
Shorten substitution proof, move the long version to the appendix

@@
 }
+
+% Two column itemize
+\usepackage{multicol}
 
@@
 \autoref{thm:substitution} instantiates the generalized version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
-\todo{Structure the theorem (in general the long ones) a bit more. Use itemize or similar. It's a blob of text.}
-\begin{nitheorem}
-  \label{thm:subst-generalization}
-  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$,
-  Let $\ni_i m$ be a \textsc{VarRef} $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$ for some $m$ and $\Gamma$.
-  Let $\ni_i n$ be a \textsc{VarRef} $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$ for some $m$ and $\Psi$.
-  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
-  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
-  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$ for some $\Gamma_j$.
-  Let $\ni_j n$ be a \textsc{VarRef} $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$ for some $\Psi_j$.
-  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-  Refer to \autoref{fig:subst} for a diagramatic representation.
-\end{nitheorem}
 
@@
 
-\begin{nidefinition}
-  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-\end{nidefinition}
-
-\todo{I don't think a reviewer from CONCUR most likely will nto get the proof. Appendix! Replace it with a proof sketch where you comment on the diagram as well.}
+\begin{nitheorem}
+  \label{thm:subst-generalization}
+  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
+  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
+  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
+  \begin{multicols}{2}
+  \begin{itemize}
+    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
+    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
+  \end{itemize}
+  \end{multicols}
+  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
+  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
+  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Refer to \autoref{fig:subst} for a diagramatic representation.
+\end{nitheorem}
+
 \begin{proof}
-  By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
   \begin{itemize}
     \item
-      For constructor $\constr{end}$ we get $\Gamma_i \equiv \Psi_i$.
-      From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
-      Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.
-
-    \item
-      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\suc$.
-
-    \item
-      For constructor $\constr{recv}$ we must split $\Delta$ to proceed inductively on the continuation.
-      Observe that given the arrow from $\Gamma_i$ to $\Psi_i$ and given that $\Delta$ is $\lz$ at index $i$, there must exist some $\delta$ such that $\opsquared{m}{\delta}{n}$.
- l     \begin{itemize}
-        \item
-          If the input is on the variable being substituted, we split $m$ such that $\opsquared{m}{\li}{l}$ for some $l$, and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ for the inductive call.
-          Similarly, we construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{j}{\li}{\Xi_j}$ as the new input channel, and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ for the inductive call.
-        \item
-          If the input is on a variable $x$ other than the one being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (for the inductive call) and $\containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
-          We then construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (the new output channel) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (for the inductive call).
-          Given there exists a composition of arrows from $\Xi_i$ to $\Psi$, we conclude that $\Theta$ splits $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-          As $\lz$ is a minimal element, then $\Delta_1$ must be $\lz$ at index $i$, and so must $\Delta_2$.
-      \end{itemize}
-
-    \item
-      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the \textsc{VarRef} doing the output, and the \textsc{VarRef} for the sent data.
-
-    \item
-      For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
-      Given $\Delta$ is $\lz$ at index $i$, we conclude that $\Delta_1$ and $\Delta_2$ are too.
-      Observe that $\opsquared{m}{\delta}{\psi}$, where $\psi$ is the usage annotation at index $i$ consumed by the subprocess $P$.
-      We construct an arrow $\containsusage{\Xi_j}{j}{\delta}{\Theta}$, for some $\Xi_j$.
-      We can now make two inductive calls (on the derivation of $P$ and $Q$) and compose their results.
+      For constructor $\constr{end}$ observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
+      The details are found in \autoref{app:subst-generalization}.
+
+    \item
+      For constructor $\constr{chan}$ we proceed inductively.
+
+    \item
+      For constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$ we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
+      The details are found in \autoref{app:subst-generalization}.
   \end{itemize}  

### 2020-05-05 paper/paper.tex
Use different counters for definitions, lemmas, and theorems
Name our theorems

@@
 \theoremstyle{definition}
-\newtheorem{nidefinition}[theorem]{Definition}
-\newtheorem{nitheorem}[theorem]{Theorem}
-\newtheorem{nilemma}[theorem]{Lemma}
+\newtheorem{nidefinition}{Definition}
+\newtheorem{nitheorem}{Theorem}
+\newtheorem{nilemma}{Lemma}
 
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Framing]
   \label{thm:framing}
-  Framing captures the idea that the well-typedness of a process is independent of its leftover resources.
+  The well-typedness of a process is independent of its leftover resources.
   Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Weakening]
   \label{thm:weakening}
-  Weakening captures the idea that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
 
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Strengthening]
   \label{thm:strengthening}
-  Strengthening captures the idea that removing an unused variable preserves the well-typedness of a process.
+  Removing an unused variable preserves the well-typedness of a process.
 
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Swapping]
   \label{thm:swapping}
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Subject congruence]
   \label{thm:subject-congruence}
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Generalised substitution]
   \label{thm:subst-generalization}
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Substitution]
   \label{thm:substitution}
@@
 
-\begin{nitheorem}
+\begin{nitheorem}[Subject reduction]
   Let $P \reduce{c} Q$.

### 2020-05-05 paper/paper.tex
Provide single proof for auxiliary theorems
pass on syntax

--- 2020-05-05 paper/paper.tex
@@
 \newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
-\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle} \;}
+\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle}. \;}
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
-\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)} \;}
+\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)}.\; }
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
 
-Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ both introduce bound names --- $x$ and $y$ in $P$, respectively.
+Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ are \emph{binders}, they are the only constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
@@
 
-A variable occurring under $n$ abstractions has $n$ possible variables to refer to.
+A variable occurring under $n$ binders can refer to other $n$ possible variables.
 References outside of that range are meaningless.

### 2020-05-05 paper/appendix.tex

@@
 \end{proof}
+
+\section{Substitution}
+\label{app:subst-generalization}
+
+\begin{nitheorem}
+  \label{thm:subst-generalization}
+  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
+  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
+  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
+  \begin{multicols}{2}
+  \begin{itemize}
+    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
+    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
+  \end{itemize}
+  \end{multicols}
+  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
+  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
+  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Refer to \autoref{fig:subst} for a diagramatic representation.
+\end{nitheorem}
+
+\begin{nidefinition}
+  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+\end{nidefinition}
+
+\begin{proof}
+  By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  \begin{itemize}
+    \item
+      For constructor $\constr{end}$ we get $\Gamma_i \equiv \Psi_i$.
+      From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
+      Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.
+
+    \item
+      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\suc$.
+
+    \item
+      For constructor $\constr{recv}$ we must split $\Delta$ to proceed inductively on the continuation.
+      Observe that given the arrow from $\Gamma_i$ to $\Psi_i$ and given that $\Delta$ is $\lz$ at index $i$, there must exist some $\delta$ such that $\opsquared{m}{\delta}{n}$.
+ l     \begin{itemize}
+        \item
+          If the input is on the variable being substituted, we split $m$ such that $\opsquared{m}{\li}{l}$ for some $l$, and construct an arrow $\containsusage{\Xi_i}{i}{l}{\Gamma}$ for the inductive call.
+          Similarly, we construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{j}{\li}{\Xi_j}$ as the new input channel, and $\containsusage{\Xi_j}{j}{l}{\Gamma}$ for the inductive call.
+        \item
+          If the input is on a variable $x$ other than the one being substituted, we construct the arrows $\containsusage{\Xi_i}{i}{m}{\Theta}$ (for the inductive call) and $\containsusage{\Gamma}{x}{\li}{\Theta}$ for some $\Theta$.
+          We then construct for some $\Xi_j$ the arrows $\containsusage{\Gamma_j}{x}{\li}{\Xi_j}$ (the new output channel) and $\containsusage{Xi_j}{j}{m}{\Theta}$ (for the inductive call).
+          Given there exists a composition of arrows from $\Xi_i$ to $\Psi$, we conclude that $\Theta$ splits $\Delta$ such that $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+          As $\lz$ is a minimal element, then $\Delta_1$ must be $\lz$ at index $i$, and so must $\Delta_2$.
+      \end{itemize}
+
+    \item
+      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the \textsc{VarRef} doing the output, and the \textsc{VarRef} for the sent data.
+
+    \item
+      For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+      Given $\Delta$ is $\lz$ at index $i$, we conclude that $\Delta_1$ and $\Delta_2$ are too.
+      Observe that $\opsquared{m}{\delta}{\psi}$, where $\psi$ is the usage annotation at index $i$ consumed by the subprocess $P$.
+      We construct an arrow $\containsusage{\Xi_j}{j}{\delta}{\Theta}$, for some $\Xi_j$.
+      We can now make two inductive calls (on the derivation of $P$ and $Q$) and compose their results.
+  \end{itemize}  
+\end{proof}

### 2020-05-05 paper/paper.tex

--- 2020-05-05 paper/paper.tex
@@
 \end{nitheorem}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
 
@@
 
-\todo{Replace the proofs with an intuition on the theorem}
+\begin{proof}
+  All the above theorems are by induction on \textsc{Process} and \textsc{Var}.
+  For details, refer to the mechanisation in Agda \cite{Zalakain2020Agda}.
+\end{proof}
+
 \subsection{Subject Congruence}

### 2020-05-05 paper/paper.tex
Rephrase subject reduction

@@
 \todo{intro}
-\todo{I wonder if we can remove the proofs everywhere and comment on them in the intro and where they can be found.}
 
@@
 
-In the linear \picalc{} reduction affects the usage context: communicating over a channel takes away $\one$ input and $\one$ output multiplicity.
+\begin{nidefinition}
+  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+\end{nidefinition}
 
 \begin{nitheorem}[Subject reduction]
-  Let $P \reduce{c} Q$.
-  If $c$ is $\constr{internal}$, then $\Gamma' \equiv \Gamma$.
-  If however $c$ is $\constr{external}\; i$, then $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$.
-  Given $\types{\gamma}{\Gamma'}{P}{\Xi}$ then after reduction $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  If $P$ reduces to $Q$ and $P$ is well-typed, then so is $Q$.
+  In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ by communication on a channel created within $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ by communication on a channel created outside of $P$.
+  \begin{itemize}
+    \item If the reduction is internal, we prove that if $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If the reduction is external and on channel $i$, we prove that if $\types{\gamma}{\Gamma'}{P}{\Xi}$ and $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$ , then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  \end{itemize}
 \end{nitheorem}
@@
 \begin{proof}
+  By induction on the typing derivation on $P$.
   \hfill{}\\

### 2020-05-05 paper/paper.tex
on semantics

--- 2020-05-05 paper/paper.tex
@@
 Thanks to our well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax.
-We start by defining structural congruence.
+We start by defining structural congruence in \autoref{structural-congruence} and then the reduction relation in \autoref{operational-semantics}.
 
@@
 
-We define the base cases of a structural congruence relation \textsc{StructCong} in \autoref{fig:struct-cong-base}.
-Its congruent equivalence closure \textsc{Equals} is provided in \autoref{app:struct}.
+We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ in \autoref{fig:struct-cong-base}.
+Its congruent equivalence closure \textsc{Equals} $=$ is defined1 in \autoref{app:struct}.
 
@@
 
-The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} indexed by the channel $i$ on which communication occurs (\autoref{fig:reduction}).
-In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel ${i}$.
-As a result of that communication, the continuation of the input process $P$ gets all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
+The operational semantics of the \picalc{} is defined by a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs (\autoref{fig:reduction}).
+We keep track of channel $c$ so we can state subject reduction (\autoref{subject-reduction}).
+We identify two kinds of channels, $\constr{internal}$ and $\constr{external}\ i$, respectively standing for a channel which ... \todo{complete}
+
+In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel with index ${i}$.
+As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
 After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (to do so, we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
@@
 
-We keep track of the channel over which communication occurs so that we can state subject reduction (in \autoref{subject-reduction}).
-We use a $\func{dec}$ function to decrement the index of this channel every time we come out of a binder.
-This function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
+We use a $\func{dec}$ function to decrement the index of this channel every time we come out of a binder in rule $\constr{res}$. \todo{this seems more getting 'into' a binder than coming out...}
+This $\func{dec}$ function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
 

### 2020-05-05 paper/paper.tex
pass on usages and contexts

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Resource-aware Typing System}
+\section{Resource-aware Type System}
 \label{type-system}
@@
 
-\begin{nidefinition}
-  We define our usage algebra to be a ternary relation $\op{x}{y}{z}$ that is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
+\begin{nidefinition}[Usage algebra]
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, which is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
     The carrier is a type with two trivial elements $\constr{zero}$ and $\constr{one}$.
-    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited in every other case.
+    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited otherwise.
     All the properties of the algebra follow trivially.
@@
     The element $\zero$ corresponds to 0, and $\one$ to 1.
-    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$.
+    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$. \todo{why isnt $\equiv$ here just =?}
     All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
@@
 
-\begin{nidefinition}
-  We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system.
-  An indexed set of usage algebras is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
+We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
+\begin{nidefinition}[IDX set of usage algebra]
+  An \emph{indexed set of usage algebras} is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
   \begin{figure}[h]
@@
   \end{figure}
-
 \end{nidefinition}
@@
 
-\begin{nidefinition}
-  \autoref{fig:prectx} defines types as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
-  The unit type $\unit$ serves as a proof of inhabitance for types.
-  The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g.
-$0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
-  The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
-  This notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
-  We keep track of types in a length-indexed $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
-
+\begin{nidefinition}[Types and length-indexed typing contexts]
+  \emph{Types} are defined in \autoref{fig:prectx} as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
+
+  We keep track of types in a \emph{length-indexed typing context} $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
+  A typing context is either empty $[]$ or building upon an existing typing context $\gamma$, it associates a type $t$ to an the next index in the context.
+  \todo{check the last sentence}
   \begin{figure}[h]
@@
 
-\begin{nidefinition}
-  We keep track of usages annotations in a context $\Ctx_{idxs}$ that is indexed by a length-indexed context of indices $\Idxs_n$, as shown in \autoref{fig:ctx}.
+The unit type $\unit$ serves as a proof of inhabitance for types.
+The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g.
+$0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
+The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
+
+Notice that our notation of channel types is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
+
+\begin{nidefinition}[Usage context]
+  A \emph{usage context} in a context $\Ctx_{idxs}$ that is indexed by a length-indexed context of indices $\Idxs_n$, where we keep track of usage annotations (\autoref{fig:ctx}).
+
+  A usage context is either empty $[]$ or building upon an existing context $\Gamma$, it associates a usage to the current index.
+  \todo{check the last sentence}
 
@@
   \end{mathpar}
-  \caption{Leftover typing for a resource-aware typing system.}
+  \caption{Leftover typing for a resource-aware type system.}
   \label{fig:types}

### 2020-05-05 paper/paper.tex
Comment on Arjen's work

@@
 
-Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but the formalisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001} is perhaps one of the earliest works.
+Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but perhaps one of the earliest works is the formalisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
 Gay encodes the \picalc{} with linear and shared types into a more general framework.
@@
 However, in his work context splits are explicit, and the types of used multiplicities erased.
-
-Taking an intrinsic approach, Thiemann formalises in Agda the MicroSession (minimal GV \cite{}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
-Context splits are given extrinsically, and exhausted elements are removed from typing contexts altogether.
-The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
 
@@
 
-\cite{Rouvoet2020}
+Taking an intrinsic approach, Thiemann formalises in Agda the MicroSession (minimal GV \cite{}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
+Context splits are given extrinsically, and exhausted elements are removed from typing contexts altogether.
+The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
+
+Most recently, Rouvoet et al. provide an intrinsic encoding (where type safety follows by construction) of a \lambdacalc{} with session types \cite{Rouvoet2020}.
+They use proof relevant separation logic and a notion of a supply and demand \emph{market} to make context splits transparent to the user.
+Their separation logic is based on a partial commutative monoid that need not be functional nor cancellative.
+Their typing rules preserve the balance between supply and demand, and are very elegant.
+They distill their typing rules even further by modelling the supply and demand market as a state monad.
 

### 2020-05-05 paper/paper.tex
Pass to intro, contrib, syntax, semantics, and contexts

@@
 
+\todo{dots}
 Let $x, y,\ldots$ range over \emph{variables} (including channel names) and $P, Q,\ldots$ over processes.
@@
 We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ in \autoref{fig:struct-cong-base}.
-Its congruent equivalence closure \textsc{Equals} $=$ is defined1 in \autoref{app:struct}.
+Its congruent equivalence closure \textsc{Equals} $\eq{}$ is defined in \autoref{app:struct}.
 
@@
 We keep track of channel $c$ so we can state subject reduction (\autoref{subject-reduction}).
-We identify two kinds of channels, $\constr{internal}$ and $\constr{external}\ i$, respectively standing for a channel which ... \todo{complete}
+We distinguish between channels created inside a process ($\constr{internal}$) and channels created outside ($\constr{external}\ i$, where $i$ is the de Bruijn index of the variable).
 
@@
 
-We use a $\func{dec}$ function to decrement the index of this channel every time we come out of a binder in rule $\constr{res}$. \todo{this seems more getting 'into' a binder than coming out...}
+In $\constr{res}$, we use a $\func{dec}$ function to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
 This $\func{dec}$ function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
@@
 \begin{nidefinition}[Usage algebra]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, which is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
     The element $\zero$ corresponds to 0, and $\one$ to 1.
-    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x \equiv y + z$. \todo{why isnt $\equiv$ here just =?}
+    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x = y + z$.
     All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
@@
 We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
-\begin{nidefinition}[IDX set of usage algebra]
+\begin{nidefinition}[Indexed set of usage algebras]
   An \emph{indexed set of usage algebras} is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
@@
   We keep track of types in a \emph{length-indexed typing context} $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
-  A typing context is either empty $[]$ or building upon an existing typing context $\gamma$, it associates a type $t$ to an the next index in the context.
-  \todo{check the last sentence}
+  A typing context is either empty $[]$ or the result of appending a type $t$ to an existing context$\gamma$.
   \begin{figure}[h]
@@
 
-\begin{nidefinition}[Usage context]
-  A \emph{usage context} in a context $\Ctx_{idxs}$ that is indexed by a length-indexed context of indices $\Idxs_n$, where we keep track of usage annotations (\autoref{fig:ctx}).
-
-  A usage context is either empty $[]$ or building upon an existing context $\Gamma$, it associates a usage to the current index.
-  \todo{check the last sentence}
+\begin{nidefinition}[Usage contexts]
+  A \emph{usage context} is a context $\Ctx_{idxs}$ of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
+
+  A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to the existing context $\Gamma$.
 

### 2020-05-05 paper/paper.tex
small updates to contexts

@@
 \begin{nidefinition}[Types and length-indexed typing contexts]
-  \emph{Types} are defined in \autoref{fig:prectx} as either unit types (rule $\unit$), base types (rule $\base{n}$) or channel types (rule $\channel{t}{x}$).
-
-  We keep track of types in a \emph{length-indexed typing context} $\PreCtx_n$, defined by the last row of rules in \autoref{fig:prectx}.
+  A \emph{Type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) defined in \autoref{fig:prectx}.
+
+  A \emph{length-indexed typing context} $\PreCtx_n$, is defined by the last row of rules in \autoref{fig:prectx}.
   A typing context is either empty $[]$ or the result of appending a type $t$ to an existing context$\gamma$.
@@
 
-Notice that our notation of channel types is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output \cite{K07}.
+Notice that our notation of channel types is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
 

### 2020-05-05 paper/paper.tex
Alignment in the example syntax

@@
 \newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
-\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle}. \;}
+\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle} \; . \;}
 \newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
-\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)}.\; }
+\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)} \; . \; }
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
 \todo{rename gradual to something else...}
+\todo{replace judgment with judgement everywhere!}
+
 The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
@@
 
-\todo{dots}
 Let $x, y,\ldots$ range over \emph{variables} (including channel names) and $P, Q,\ldots$ over processes.
@@
     \begin{split}
-      & (\new{}x ) && (\comp {\recvp{x}{y} && \; \sendp{y}{a} \; \PO} {(\new{} y) && (\sendp{x}{y} \; \recvp{y}{z} && \; \PO)})
+      & (\new{}x ) && (\comp {\recvp{x}{y} && \sendp{y}{a} && \PO} {(\new{} y) && (\sendp{x}{y} && \recvp{y}{z} && \PO)})
       && \text{names}
       \\
-      & \new{} && (\comp {\recv{0} && \; \send{0}{a} \; \PO} {\new{} && (\send{1}{0} \; \recv{0} && \; \PO)})
+      & \new{} && (\comp {\recv{0} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0} && \PO)})
       && \text{de Bruijn indices}
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context and usage context, adds an extra output usage context to the typing rules.
+We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra {output usage context} to the typing rules.
+
 This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
@@
   \item
-    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgment with both an input and an output usage context.
+    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgement with both an input and an output usage context.
   \item
@@
 \end{itemize}
-
-We introduce two judgments: one on variable references (\autoref{def:varref}), one on processes (\autoref{def:types}).
-Both are indexed by a $\PreCtx_n$, an input $\Ctx_{idxs}$, and an output $\Ctx_{idxs}$ (the leftovers).
-
-\begin{nidefinition}
+  \todo{I still don't get the last point above, as weakening and strengthening were already possible to state in \picalc{} before. What do you mean with can be stated here?}
+
+Our type system with leftovers is composed of two typing relations: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
+Both relations are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (containing the leftovers).
+
+The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in $\gamma$, and subtracting $y$ at position $i$ from $\Gamma$ results in $\Delta$''.
+The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
+
+\begin{nidefinition}[\textsc{VarRef}: typing variable references]
   \label{def:varref}
 
-  The \textsc{VarRef} relation is presented in \autoref{fig:variable-references}.
-  The judgment $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ reads as ``$\gamma$ at position $i$ is of type $t$, and subtracting $y$ at position $i$ from $\Gamma$ results in $\Delta$''.
-
-  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
-  This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
-  We use the \emph{functionality} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
-  The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.
-
+  The \textsc{VarRef} typing relation for variable references is presented in \autoref{fig:variable-references}.
   \begin{figure}[h]
@@
   \end{mathpar}
-  \caption{Variable reference judgments.}
+  \caption{Typing rules for variable references.}
   \label{fig:variable-references}
   \end{figure}
+
+  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
+  This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
+  We use the \emph{functionality} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
+  The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts. 
 \end{nidefinition}
 
-\begin{example}
+\begin{example}[Channels]
   Let us illustrate the use of channels with an example.
@@
 
-\begin{nidefinition}
+\begin{nidefinition}[\textsc{Types}: typing processes]
   \label{def:types}
 
-  The typing relation \textsc{Types} is presented in \autoref{fig:types}.
-  The typing judgement $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
+  The \textsc{Types} typing relation for \picalc{} processes is presented in \autoref{fig:types}.
 
@@
 
-  Some comments on the leftover type system follow.
-
   The inaction process does not change usage annotations.
-
   Scope restriction expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ that is being transmitted; and the multiplicity $y$ given to the channel itself.
@@
 
-\begin{example}
+\begin{example}[Typing derivation]
   \label{example-derivations}

### 2020-05-05 paper/paper.tex
a pass on leftovers
First conclusions and future work

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-To conclude, in this paper...
-Mechanisation of the linear, gradual and shared \picalc{} in less than 1500 lines of code.
-
-As future work we intend to...
-Adding product types and sum types.
-Proving that progress preserves the balancing of channels.
-Erika Kreuer has suggested modelling the usage algebra as a state machine.
-Adding bidirectionality would provide us with a decision procedure for type checking for a given context of algebras.
+In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, define on top of the syntax a type system capable of handling linear, gradual and shared types, and show that reduction preserves the well-typedness of a process.
+We avoid the need for extrinsic context splits by defining a type system with leftovers \cite{Allais2018a}.
+All this is mechanised in less than 1500 lines of code.
+
+We consider proving that progress preserves the balancing of channels long-hanging fruit for the future.
+More involved, we intend to add support for products and sum types to both the syntax and the type system.
+Orthogonally, making our typing rules bidirectional would allows us to provide a decision procedure for type checking processes for a given set of algebras.
+Following Erika Kreuer's suggestion, it might also be worth identifying an equivalence between our usage algebra and particular state machines.
 

### 2020-05-05 paper/paper.tex
Work on the usage algebra figure

@@
     &\func{\cdot-compute}  &:{} &\forall y z       & \to \; & \type{DEC} \; (\type{\exists} x  \; (\op{x}{y}{z})) \\
-    &\func{\cdot-unique}   &:{} &\forall \{x x' y z\}  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
-    &\func{\cdot-unique^l} &:{} &\forall \{x y y' z\}  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-    &\func{0\cdot-min^l}   &:{} &\forall \{y z\}       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
-    &\func{\cdot-id^l}     &:{} &\forall \{x\}         & \to \; & \op{x}{\zero}{x} \\
-    &\func{\cdot-comm}     &:{} &\forall \{x y z\}     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-    &\func{\cdot-assoc}    &:{} &\forall \{x y z u v\} & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
+    &\func{\cdot-unique}   &:{} &\forall x x' y z  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
+    &\func{\cdot-unique^l} &:{} &\forall x y y' z  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+    &\func{0\cdot-min^l}   &:{} &\forall y z       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+    &\func{\cdot-id^l}     &:{} &\forall x         & \to \; & \op{x}{\zero}{x} \\
+    &\func{\cdot-comm}     &:{} &\forall x y z     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+    &\func{\cdot-assoc}    &:{} &\forall x y z u v & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
   \end{aligned}
@@
   \caption{
-    Algebra $\Algebra_C$ algebra on a partial commutative monoid.
-    In Agda $\forall$ introduces universal quantification, while $\{ \text{brackets} \}$ make the variables within them implicit.
+    Usage algebra $\Algebra_C$ on a carrier $C$.
+    We use $\forall$ for universal quantification.
     The dependent product $\type{\exists}$ uses the value of its first argument in the type of its second.

### 2020-05-05 paper/paper.tex
Small fixes

@@
   We will often work with pairs of input and output multiplicities.
-  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \times \type{C}$ tuple of input and output multiplicities, respectively.
+  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ tuple of input and output multiplicities, respectively.
   We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
@@
     All the properties of the algebra follow trivially.
-    All shared channels have multiplicities $(\omega \comma \omega)$, which they will forever preserve.
+    Shared channels have multiplicities $(\constr{\omega} \comma \constr{\omega})$, which they will forever preserve.
 \end{description}

### 2020-05-05 paper/paper.tex
Minor changes
more on leftovers

@@
 \todo{replace judgment with judgement everywhere!}
+\todo{Theorem to Thm. Definiton to Def. when referencing.}
 
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra {output usage context} to the typing rules.
+We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{output usage context} to the typing rules.
 
@@
 \begin{itemize}
-  \item
+  \item \textbf{Encapsulation}:
     The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgement with both an input and an output usage context.
-  \item
+  \item \textbf{No context split}:
     No need for an extensional context splitting proof for parallel composition.
@@
     Leftover typing uses the leftovers of $P$ to type $Q$.
-  \item
-    New theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can be stated thanks to the presence of a leftover context.
+  \item \textbf{New theorems}:
+    For the first time, theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can now be stated within the linear \picalc{} --- as well as gradual and shared, thanks to the presence of a leftover context.
 \end{itemize}
-  \todo{I still don't get the last point above, as weakening and strengthening were already possible to state in \picalc{} before. What do you mean with can be stated here?}
 
 Our type system with leftovers is composed of two typing relations: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
-Both relations are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (containing the leftovers).
-
-The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in $\gamma$, and subtracting $y$ at position $i$ from $\Gamma$ results in $\Delta$''.
+Both relations are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).
+
+The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in typing context $\gamma$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftover $\Delta$''.
 The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
@@
 
-  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
+  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$ (the leftovers).
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
@@
 
-  The inaction process does not change usage annotations.
-  Scope restriction expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ that is being transmitted; and the multiplicity $y$ given to the channel itself.
+  The inaction process in rule $\constr{end}$ does not change usage annotations.
+  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ that is being transmitted; and the multiplicity $y$ given to the channel itself.
   This multiplicity $y$ is used for both input and output, so that these are balanced.
@@
 
-  Receiving needs a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  The input process in rule $\constr{recv}$ requires a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
   Note that the index $i$ is used in the syntax of the typed process.
@@
 
-  Similar to input, sending needs a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  Similarly to input, the output processes in rule $\constr{send}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
   We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of channel $chan_i$.
@@
 
-  Parallel composition uses the leftovers of the left-hand process to type the right-hand process.
-  By keeping track in the typing derivation of $P$ of the resources it uses, we save the user from having to provide an extrinsic proof of context split.
+  Parallel composition in rule $\constr{comp}$ uses the leftovers of the left-hand process to type the right-hand process.
+  By keeping track in the typing derivation of $P$ the resources it uses, we save the user from having to provide an extrinsic proof of context split.
 \end{nidefinition}
@@
   \label{example-derivations}
-  Let us type a derivation of the process $\func{p}$ using linear and shared types. 
+  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types, given in line ?? of Listing \autoref{}.
+
   Let $\func{\omega}$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
@@
   To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transiting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability.
-  The left-hand process is typed following its syntax.
-  On the right-hand, we need the channel that is going to be sent.
-  The type $\unit$ of this channel is entirely determined by the type of the channel over which it is going to be transmited.
-  Its usage annotation is $(\one \comma \one)$: we will transmit $\li$ and save $\lo$ for the continuation.
+  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability.
+  The left-hand process (line ??) is typed following its syntax.
+  On the right-hand (line ??), we need the channel that is going to be sent.
+  The type $\unit$ of this channel is entirely determined by the type of the channel over which it is going to be transmitted.
+  Its usage annotation is $(\one \comma \one)$: we will transmit $\li$ and save $\lo$ for the continuation process.
 

### 2020-05-05 paper/paper.tex
on auxiliary results

@@
 
-\todo{intro}
+In \autoref{auxiliary-results} we present some auxiliary results: framing (\autoref{thm:framing}), weakening (\autoref{thm:weakening}), strengthening (\autoref{thm:strengthening}) and swapping of variables (\autoref{thm:swapping}).
+We present subject congruence in \autoref{subject-congruence} and substitution in \autoref{substitution}.
+Finally, we present subject reduction in \autoref{subject-reduction}.
 
@@
 \label{auxiliary-results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Framing asserts that the well-typedness of a process is independent of its leftover resources.
 \begin{nitheorem}[Framing]
   \label{thm:framing}
-  The well-typedness of a process is independent of its leftover resources.
-  Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
-  Let there be a $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
-  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
+  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+
   Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
@@
 
+Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
 \begin{nitheorem}[Weakening]
   \label{thm:weakening}
-  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
-
-  Let $\func{insert}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then lifting every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{insert}_i \; t \; \gamma}{\func{insert}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{insert}_i \; x \; \Xi}$.
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, lifting every variable greater than or equal to $i$ in $P$ is well-typed in
+  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
 \end{nitheorem}
 
+Removing an unused variable preserves the well-typedness of a process.
+Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
 \begin{nitheorem}[Strengthening]
   \label{thm:strengthening} 
-  Removing an unused variable preserves the well-typedness of a process.
-
-  Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
+  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed in $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
 \end{nitheorem}
@@
 
-\begin{nitheorem}[Swapping]
-  \label{thm:swapping}
 Swapping two variables preserves the well-typedness of a process.
@@
 For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
-  \todo{}
-  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
+\begin{nitheorem}[Swapping]
+  \label{thm:swapping}
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
 \end{nitheorem}
@@
 \begin{proof}
-  All the above theorems are by induction on \textsc{Process} and \textsc{Var}.
+  All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
   For details, refer to the mechanisation in Agda \cite{Zalakain2020Agda}.

### 2020-05-05 paper/paper.tex
on congruence

@@
 
+Subject congruence theorem states that applying structural congruence rules to a well-typed process preserves its well-typedness.
 \begin{nitheorem}[Subject congruence]
   \label{thm:subject-congruence}
-  Applying structural congruence rules to a well-typed process preserves its well-typedness.
-  That is, if $P \eq{r} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
 
-\todo{considering that Equals is in the appendix, you cannot do induction on something that is not defined in the main text. Bring the proof in the appendix. Maybe briefly comment on the proof here}
-\begin{proof}
-  By induction on \textsc{Equals}.
-  The base cases under $\constr{struct}$ (and their symmetric variants under $\constr{sym}$) are the nontrivial ones.
-  For those, we proceed by induction on \textsc{StructCong}:
+\begin{proof}[Proof (Sketch)]
+  The proof is induction on \textsc{Equals} $\eq{}$.
+  Here we will only consider the base cases for $\constr{struct}$ (and their symmetric variants under $\constr{sym}$), which are nontrivial. Full proof in \cite{Zalakain2020Agda}.
+  We proceed by induction on \textsc{StructCong} $\equiv$:
   \begin{itemize}
     \item
-      The case $\constr{comp-assoc}$ is trivial: leftover typing is naturally associative.
-    \item
-      In the $\constr{comp-sym}$ case for a process $\comp{P}{Q}$, we need framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of of $Q$ to the one of $P$.
+      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
+    \item
+      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
       The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
     \item
-      The case $\constr{comp-end}$ is trivial: the typing rule for a process $\PO$ has the same input and output contexts.
-    \item
-      In the $\constr{scope-end}$ case, we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
+      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
+    \item
+      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
       In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
     \item
-      In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
+      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
       We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with typing derivation on $Q$ wrapped in $\constr{chan}$.
@@
     \item
-      In the $\constr{scope-comm}$ case, we use swapping (\autoref{thm:swapping}).
+      Case $\constr{scope-comm}$: in one direction we use swapping (\autoref{thm:swapping}).
       In the reverse direction, we use swapping and then \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.

### 2020-05-05 paper/paper.tex
Go over the last changes

@@
 \todo{make sure all citations are there}
+\todo{rename gradual to something else...}
+\todo{Theorem to Thm. Definiton to Def. when referencing. UMA: it's currently set by the lipics class, should we change it?}
 
@@
 \section{Introduction}
-\todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
+
 We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
 As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-\todo{rename gradual to something else...}
-\todo{replace judgment with judgement everywhere!}
-\todo{Theorem to Thm. Definiton to Def. when referencing.}
 
@@
     \item \textbf{\picalc{} with leftovers}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    This approach adds a leftover usage context to the typing judgments.
+    This approach adds a leftover usage context to the typing judgements.
     Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
@@
   \item \textbf{New theorems}:
-    For the first time, theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can now be stated within the linear \picalc{} --- as well as gradual and shared, thanks to the presence of a leftover context.
+    For the first time, theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can be stated within the linear \picalc{} thanks to the presence of a leftover context.
 \end{itemize}
@@
   Let us illustrate the use of channels with an example.
-  \begin{lstlisting}[label=lst:channels,mathescape,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
+  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
 
-  Similarly to input, the output processes in rule $\constr{send}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  Similarly to input, the output process in rule $\constr{send}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
   We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of channel $chan_i$.
@@
   \label{example-derivations}
-  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types, given in line ?? of Listing \autoref{}.
+  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types (Listing \autoref{lst:derivation}).
 
@@
   To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability.
-  The left-hand process (line ??) is typed following its syntax.
-  On the right-hand (line ??), we need the channel that is going to be sent.
+  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability (line 7).
+  The left-hand process is typed following its syntax (line 8).
+  On the right-hand we need the channel that is going to be sent (line 9).
   The type $\unit$ of this channel is entirely determined by the type of the channel over which it is going to be transmitted.
@@
 
-  \begin{lstlisting}[mathescape]
+  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left]
   $\func{p}$ : $\Process_1$
@@
 Removing an unused variable preserves the well-typedness of a process.
-Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
+Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
 \begin{nitheorem}[Strengthening]
@@
   Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed in $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
+  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
 \end{nitheorem}
@@
 Swapping two variables preserves the well-typedness of a process.
-Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
-For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
+Let $\func{sw}_i$ be a function that swaps variables $i$ and $\suc i$.
+For simplicity, we overload $\func{sw}$ and use it to swap types in a typing context --- we swap variable references in a process with $swap_i \; P$, as first used in structural congruence (\autoref{structural-congruence}).
 \begin{nitheorem}[Swapping]
@@
   Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
+  Then, $\types{\func{sw}_i \; \gamma}{\func{sw}_i \; \Gamma}{\func{swap}_i \; P}{\func{sw}_i \; \Xi}$.
 \end{nitheorem}

### 2020-05-05 paper/paper.tex
Shorten autoref names

@@
 \usepackage{hyperref}
-%%
-%% For section renaming in autoref
-\addto\extrasUKenglish{%
-\renewcommand{\sectionautorefname}{\S}%
-\renewcommand{\subsectionautorefname}{\S}%
-\def\exampleautorefname{Ex.}% 
-\renewcommand{\figureautorefname}{Fig.}%
-}
 
@@
 \newtheorem{nilemma}{Lemma}
+\renewcommand{\nitheoremautorefname}{Thm.}%
+\renewcommand{\nidefinitionautorefname}{Def.}%
+\renewcommand{\nilemmaautorefname}{Lem.}%
+\renewcommand{\exampleautorefname}{Ex.}%
+\addto\extrasUKenglish{%
+  \renewcommand{\sectionautorefname}{\S}%
+  \renewcommand{\subsectionautorefname}{\S}%
+  \renewcommand{\figureautorefname}{Fig.}%
+}
 
@@
 \todo{rename gradual to something else...}
-\todo{Theorem to Thm. Definiton to Def. when referencing. UMA: it's currently set by the lipics class, should we change it?}
 
@@
 \begin{nidefinition}[Types and length-indexed typing contexts]
-  A \emph{Type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) defined in \autoref{fig:prectx}.
+  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) defined in \autoref{fig:prectx}.
 

### 2020-05-05 paper/appendix.tex
Full pass

@@
 \begin{nitheorem}
-  \label{thm:subst-generalization}
   Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
 \end{nitheorem}
-
-\begin{nidefinition}
-  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-\end{nidefinition}
 

### 2020-05-05 paper/paper.tex
Full pass

@@
 \todo{rename gradual to something else...}
+\todo{lemmas in appendix are referred to from here}
 
@@
 
-Recent work has pushed the linear \picalc{} into the spotlight again \cite{KPT96}, thanks to an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17}.
-This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but it is useful at a practical level too.
+Recent work has pushed the linear \picalc{} into the spotlight again \cite{KPT96} thanks to an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17}.
+This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
 The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
@@
  \begin{itemize}
-   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972} to introduce a syntax that is \emph{well-scoped by construction}: every free variable is accounted for in the type of the process that uses it.
+   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well-scoped by construction}: every free variable is accounted for in the type of the process that uses it.
 
@@
     Linear types, gradual types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
-    Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a minimal element is a valid such algebra.
+    Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
 
@@
 Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
-Implicit arguments are mentioned by type definitions but omitted in constructors.
+Implicit arguments are mentioned by type definitions but omitted by constructors.
 
@@
 
-A variable occurring under $n$ binders can refer to other $n$ possible variables.
+A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
 References outside of that range are meaningless.
 It is useful to rule out these ill-scoped nonsensical terms syntactically.
-In \autoref{fig:syntax}, we do so by introducing the indexed family of types $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+In \autoref{fig:syntax}, we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within it to refer to one more thing.
+Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
 
@@
 
-The first three rules ($\constr{comp}-$*) state associativity, symmetry, and $\PO$ as the neutral element of parallel composition, respectively.
+The first three rules ($\constr{comp}-$*) state associativity, symmetry, and $\PO$ as being the neutral element of parallel composition, respectively.
 The last three ($\constr{scope}-$*) state garbage collection, scope extrusion and commutativity of restrictions, respectively.
@@
 In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
-The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $i$.
+The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
 In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
@@
 
-The operational semantics of the \picalc{} is defined by a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs (\autoref{fig:reduction}).
+The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs (\autoref{fig:reduction}).
 We keep track of channel $c$ so we can state subject reduction (\autoref{subject-reduction}).
-We distinguish between channels created inside a process ($\constr{internal}$) and channels created outside ($\constr{external}\ i$, where $i$ is the de Bruijn index of the variable).
+We distinguish between channels that are created inside of the process ($\constr{internal}$) and channels that are created outside ($\constr{external}\ i$, where $i$ is the index of the channel variable).
 
@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (to do so, we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
+After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 
-In $\constr{res}$, we use a $\func{dec}$ function to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
+In $\constr{res}$, we use $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
 This $\func{dec}$ function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
@@
   \item [Linear]
-    The carrier is a type with two trivial elements $\constr{zero}$ and $\constr{one}$.
+    The carrier is a type with two trivial constructors $\constr{zero}$ and $\constr{one}$.
     The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited otherwise.
@@
 We keep typing contexts ($\PreCtx$, in \autoref{fig:prectx}) and usage contexts ($\Ctx$ in \autoref{fig:ctx}) separate.
-The former are preserved throughout a typing derivation; the latter are transformed as a result of input, output, and context splits.
-
-\begin{nidefinition}[Types and length-indexed typing contexts]
-  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) defined in \autoref{fig:prectx}.
-
-  A \emph{length-indexed typing context} $\PreCtx_n$, is defined by the last row of rules in \autoref{fig:prectx}.
-  A typing context is either empty $[]$ or the result of appending a type $t$ to an existing context$\gamma$.
+The former are preserved throughout typing derivations; the latter are transformed as a result of input, output, and context splits.
+
+\begin{nidefinition}[Types and typing contexts]
+  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (\autoref{fig:prectx}).
+
   \begin{figure}[h]
@@
 The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
-
-Notice that our notation of channel types is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
+Notice that this notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
+
+A \emph{typing context} (last row in \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
 
@@
   \end{mathpar}
-  \caption{Length-indexed context of carrier indices with a context of multiplicities on top.}
+  \caption{Length-indexed context of carrier indices with a context of usage annotations on top.}
   \label{fig:ctx}
@@
   We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
-  All the properties of the partial monoid are trivially lifted.
+  All the properties of the partial monoid are lifted element-wise.
 \end{note}
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{output usage context} to the typing rules.
+We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{output usage context} $\Ctx_{idxs}$ to the typing rules.
 
@@
     The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgement with both an input and an output usage context.
-  \item \textbf{No context split}:
+  \item \textbf{No extensional context split}:
     No need for an extensional context splitting proof for parallel composition.
@@
 
-The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in typing context $\gamma$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftover $\Delta$''.
+The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in typing context $\gamma$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
 The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
@@
 
-\begin{example}[Channels]
-  Let us illustrate the use of channels with an example.
+\begin{example}[Variable reference]
+  Let us illustrate the use of variable reference relations with an example.
   \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
@@
   In \autoref{lst:channels}, the constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving leftovers $\lo$.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover.
 \end{example}
@@
   The inaction process in rule $\constr{end}$ does not change usage annotations.
-  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ that is being transmitted; and the multiplicity $y$ given to the channel itself.
-  This multiplicity $y$ is used for both input and output, so that these are balanced.
+  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ of what is being transmitted; and the multiplicity $y$ given to the channel itself.
+  This multiplicity $y$ is used for both input and output, so that they are balanced.
   The continuation process $P$ is provided with the new channel with usage annotation $(y \comma y)$, which it must completely exhaust.
@@
   Similarly to input, the output process in rule $\constr{send}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
-  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of channel $chan_i$.
+  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan_i$.
   The leftovers $\Xi$ are used to type the continuation process.
@@
   Parallel composition in rule $\constr{comp}$ uses the leftovers of the left-hand process to type the right-hand process.
-  By keeping track in the typing derivation of $P$ the resources it uses, we save the user from having to provide an extrinsic proof of context split.
+  By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide an extrinsic proof of context split.
 \end{nidefinition}
@@
   \label{example-derivations}
-  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types (Listing \autoref{lst:derivation}).
+  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types (\autoref{lst:derivation}).
 
@@
   To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we will only transmit its input capability (line 7).
+  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
   The left-hand process is typed following its syntax (line 8).
   On the right-hand we need the channel that is going to be sent (line 9).
-  The type $\unit$ of this channel is entirely determined by the type of the channel over which it is going to be transmitted.
-  Its usage annotation is $(\one \comma \one)$: we will transmit $\li$ and save $\lo$ for the continuation process.
+  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
+  Its usage annotation is $(\one \comma \one)$: we will transmit $(\one \comma \zero)$ and save $(\zero \comma \one)$ for the continuation process.
 
@@
   Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
+  Then lowering every variable greater than $i$ in $P$ is well-typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
 \end{nitheorem}
@@
 \begin{remark}
-  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned within $P$ --- $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
+  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned by $P$: $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
 \end{remark}
@@
 Let $\func{sw}_i$ be a function that swaps variables $i$ and $\suc i$.
-For simplicity, we overload $\func{sw}$ and use it to swap types in a typing context --- we swap variable references in a process with $swap_i \; P$, as first used in structural congruence (\autoref{structural-congruence}).
+For simplicity, we overload $\func{sw}$ and use it to swap types in typing contexts and usage annotations in usage contexts.
+We use $\func{swap}_i \; P$ to swap variables $i$ and $\suc i$ within a process $P$, as introduced in structural congruence (\autoref{structural-congruence}).
 \begin{nitheorem}[Swapping]
@@
   Doing so is enough to prove subject reduction.
-  More generally, any two variables can be swapped by performing a series of adjacent variable swappings.
+  More generally, any two variables can be swapped by a series of adjacent variable swappings.
 \end{remark}
 
-\begin{proof}
+\begin{proof}[Proof (Sketch)]
   All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
@@
 
-Subject congruence theorem states that applying structural congruence rules to a well-typed process preserves its well-typedness.
+Subject congruence states that applying structural congruence rules to a well-typed process preserves its well-typedness.
 \begin{nitheorem}[Subject congruence]
@@
 \begin{proof}[Proof (Sketch)]
-  The proof is induction on \textsc{Equals} $\eq{}$.
+  The proof is by induction on \textsc{Equals} $\eq{}$.
   Here we will only consider the base cases for $\constr{struct}$ (and their symmetric variants under $\constr{sym}$), which are nontrivial. Full proof in \cite{Zalakain2020Agda}.
-  We proceed by induction on \textsc{StructCong} $\equiv$:
+  We proceed by induction on \textsc{StructCong} $\eqeq$:
   \begin{itemize}
@@
       Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
-      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with typing derivation on $Q$ wrapped in $\constr{chan}$.
+      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with the typing derivation on $Q$ wrapped in $\constr{chan}$.
       In the reverse direction, we weaken (\autoref{thm:weakening}) $P$ so that it accepts the additional variable.
@@
     \caption{
-      Diagrammatic representation of the inductive substitution lemma.
+      Diagrammatic representation of the $\constr{comp}$ case for substitution.
       Continuous lines represent known facts, dotted lines proof obligations.
@@
   \label{thm:subst-generalization}
-  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
+  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
   Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
@@
   Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
-  Let $\Delta$ have a usage annotation $\lz$ at position $i$.
-  Then substituting $i$ with $j$ in $P$ will be well-typed such that $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
+  Then substituting $i$ with $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
   Refer to \autoref{fig:subst} for a diagramatic representation.
@@
 
-\begin{proof}
+\begin{proof}[Proof (Sketch)]
   By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
   \label{thm:substitution}
-  Let $P$ be a process typed under $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
-  Let $\ni_j m$ be a \textsc{VarRef} $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed under $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+  Let $P$ be a well-typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
 \end{nitheorem}
-\begin{proof}
-  Use framing to derive $\contains{\gamma}{\Gamma}{j}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
-  Supply these to the generalised \autoref{thm:subst-generalization}.
+\begin{proof}[Proof (Sketch)]
+  Use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
+  Use these to apply \autoref{thm:subst-generalization}.
 \end{proof}
@@
   If $P$ reduces to $Q$ and $P$ is well-typed, then so is $Q$.
-  In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ by communication on a channel created within $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ by communication on a channel created outside of $P$.
+  In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$.
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
   \begin{itemize}
-    \item If the reduction is internal, we prove that if $\types{\gamma}{\Gamma}{P}{\Xi}$ then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-    \item If the reduction is external and on channel $i$, we prove that if $\types{\gamma}{\Gamma'}{P}{\Xi}$ and $\containsusage{\Gamma'}{i}{\lio}{\Gamma}$ , then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If the reduction is internal, we prove $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If the reduction is external and on channel $i$, we prove that if $\containsusage{\Gamma}{i}{\lio}{\Delta}$ , then $\types{\gamma}{\Delta}{Q}{\Xi}$.
   \end{itemize}
@@
 \begin{proof}
-  By induction on the typing derivation on $P$.
+  By induction on $\types{\gamma}{\Gamma}{P}{\Xi}$.
   \hfill{}\\
@@
     Reduction under scope restriction case splits on the channel $c$ on which communication occurs.
-    If $c$ is $\constr{internal}$ subject reduction proceeds inductively.
+    If $c$ is $\constr{internal}$ we proceed inductively.
     If $c$ is $\constr{external}\; \constr{0}$ -- i.e.
-the channel that is being introduced by the scope restriction -- we use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and then apply the induction hypothesis.
-    If $c$ is $\constr{external}\; (\suc i)$ subject reduction proceeds inductively.
-
-    \item
-    In the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and the induction hypothesis.
+the channel that is being introduced by the scope restriction -- we use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and proceed inductively.
+    If $c$ is $\constr{external}\; (\suc i)$ we proceed inductively.
+
+    \item
+    In the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively.
   \end{itemize}
@@
 Allais uses leftover typing to mechanise in Agda a bidirectional type system for the linear \lambdacalc{}.
-This type system sits on top of an untyped but well-scoped syntax with de Bruijn indices.
+This type system sits on top of an untyped but well-scoped syntax.
 He proves type preservation and provides a decision procedure for type checking and type inference.
-In this paper, we apply his technique to the \picalc{}, but limit ourselves to type-checking (scope restriction needs usage annotations) terms, and focus on generalising the usage algebra.
-Drawing from quantitative type theory by McBride and Atkey \cite{McBride2016, Atkey2018}, it is crucial for us too to be able to talk about fully consumed resources.
-As a by-product, we are able to transmit over a channel all none multiplicities of a fully exhausted channel over.
-
-Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but perhaps one of the earliest works is the formalisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
+In this paper, we apply his technique to the \picalc{}, but limit ourselves to type-checking terms (scope restriction needs usage annotations), and focus on generalising the usage algebra.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), it is crucial for us too to be able to talk about fully consumed resources.
+As a by-product, we are able to transmit over a channel all none multiplicities of a fully exhausted channel.
+
+Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but perhaps one of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
 Gay encodes the \picalc{} with linear and shared types into a more general framework.
@@
 Their paper extends session types with polymorphism and pattern matching on session types.
-They use a locally-nameless encoding of variable references, a syntax prior to types, and a LTS semantics that encodes session-typed processes into the \picalc{}.
-Their type system uses reorering of contexts and extrinsic contex splits.
-
-Taking an intrinsic approach, Thiemann formalises in Agda the MicroSession (minimal GV \cite{}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
-Context splits are given extrinsically, and exhausted elements are removed from typing contexts altogether.
+They use a locally-nameless encoding for variable references, a syntax prior to types, and a LTS semantics that encodes session-typed processes into the \picalc{}.
+Their type system uses reordering of contexts and extrinsic contex splits.
+
+Taking an intrinsic approach to typing, Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
+Context splits are still given extrinsically, and exhausted elements are removed from typing contexts altogether.
 The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
 
-Most recently, Rouvoet et al. provide an intrinsic encoding (where type safety follows by construction) of a \lambdacalc{} with session types \cite{Rouvoet2020}.
+Most recently, Rouvoet et al. provide an intrinsic type system for a \lambdacalc{} with session types \cite{Rouvoet2020}.
 They use proof relevant separation logic and a notion of a supply and demand \emph{market} to make context splits transparent to the user.
 Their separation logic is based on a partial commutative monoid that need not be functional nor cancellative.
-Their typing rules preserve the balance between supply and demand, and are very elegant.
+Their typing rules preserve the balance between supply and demand, and are extremely elegant.
 They distill their typing rules even further by modelling the supply and demand market as a state monad.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, define on top of the syntax a type system capable of handling linear, gradual and shared types, and show that reduction preserves the well-typedness of a process.
-We avoid the need for extrinsic context splits by defining a type system with leftovers \cite{Allais2018a}.
-All this is mechanised in less than 1500 lines of code.
-
-We consider proving that progress preserves the balancing of channels long-hanging fruit for the future.
-More involved, we intend to add support for products and sum types to both the syntax and the type system.
-Orthogonally, making our typing rules bidirectional would allows us to provide a decision procedure for type checking processes for a given set of algebras.
-Following Erika Kreuer's suggestion, it might also be worth identifying an equivalence between our usage algebra and particular state machines.
+In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, gradual and shared types, and show that reduction preserves the well-typedness of processes.
+We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.
+Our work is mechanised in roughly 1500 lines of code.
+
+For the future, we consider proving that reduction preserves the balancing of channels long-hanging fruit.
+More involved, we intend to add support for products and sum types to both our syntax and our type system.
+Orthogonally, making our typing rules bidirectional would allows us to provide a decision procedure for type checking processes in a given set of algebras.
+Following Erika Kreuer's suggestion, it might also be worth identifying correspondences between our usage algebra and particular state machines.
 

### 2020-05-05 paper/paper.tex
Replace gradual with graded

@@
 
-  We present the \emph{first} full mechanisation in Agda of a \picalc{} with \emph{linear}, \emph{gradual} and \emph{shared} types, all under the same unified framework.
-  While linearity is key for type safety in communication-centric programming, gradual and shared types are needed to model real-world software systems.
+  We present the \emph{first} full mechanisation in Agda of a \picalc{} with \emph{linear}, \emph{graded} and \emph{shared} types, all under the same unified framework.
+  While linearity is key for type safety in communication-centric programming, graded and shared types are needed to model real-world software systems.
 
@@
   For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
-  We generalise the algebras on multiplicities, allowing the developer to choose a mix of linear, gradual and shared typing.
+  We generalise the algebras on multiplicities, allowing the developer to choose a mix of linear, graded and shared typing.
   We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
@@
 \todo{make sure all citations are there}
-\todo{rename gradual to something else...}
 \todo{lemmas in appendix are referred to from here}
@@
 
-\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, gradual and shared types, all under the same unified framework.}
+\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under the same unified framework.}
 
@@
 Ultimately, we can exploit this formalisation as a unified framework for type safe distributed modelling and programming with a variety of type systems, and build on top of it more advanced types, theories and languages.
-In particular, while linearity is needed for type safety in communication-centric programming, gradual and shared types are needed to model real-world systems.
+In particular, while linearity is needed for type safety in communication-centric programming, graded and shared types are needed to model real-world systems.
 
@@
 
-  \item \textbf{Leftover typing for the \picalc{} with linear, gradual and shared types}:
+  \item \textbf{Leftover typing for the \picalc{} with linear, graded and shared types}:
   \autoref{type-system} provides the type system for a resource-aware \picalc{}.
@@
     \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-    Linear types, gradual types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
+    Linear types, graded types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
     Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
@@
     Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, gradual and shared types.
+    This allows for type systems combining linear, graded and shared types.
 
@@
 In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise over this notion by defining an algebra that is satisfied by linear, gradual and shared types alike.
+We generalise over this notion by defining an algebra that is satisfied by linear, graded and shared types alike.
 
@@
 
-Linear, gradual and shared types are all defined as an instance of our usage algebra.
+Linear, graded and shared types are all defined as an instance of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.
@@
 
-    \item [Gradual]
+    \item [Graded]
     The carrier is the type of natural numbers.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, gradual and shared types, and show that reduction preserves the well-typedness of processes.
+In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types, and show that reduction preserves the well-typedness of processes.
 We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.

### 2020-05-05 paper/appendix.tex
Update subst theorem in the appendix

@@
 
-\begin{nitheorem}
-  Let process $P$ be well-typed such that $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  The substituted variable $i$ has a usage $m$ in $\Gamma_i$, and a usage $n$ in $\Psi_i$.
+\begin{nitheorem}[Generalised substitution]
+  \label{thm:subst-generalization}
+  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
   Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.

### 2020-05-06 paper/paper.tex
Update subst theorem in the appendix
on benefits of leftovers

@@
 \todo{lemmas in appendix are referred to from here}
+\todo{shorten: we are slightly over the limit}
 
@@
     Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
-    One of the benefits of this technique is that it removes the need for extrinsic context splits, which are rendered unnecessary.
+
+    Benefits of typing with leftovers include: removing the need for extrinsic context splits, which are rendered unnecessary; and for the first time we can state theorems like \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) in the linear \picalc{}.
   \end{itemize}

### 2020-05-06 paper/paper.tex
on examples

@@
 \begin{example}[Variable reference]
-  Let us illustrate the use of variable reference relations with an example.
-  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=The underscore \_ introduces an anonymous declaration immediately followed by its definition.]
+  Let us illustrate the use of variable reference relations with an example in \autoref{lst:channels}.
+  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Variable referece]
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
   \end{lstlisting}
-  In \autoref{lst:channels}, the constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover.
+  The underscore \_ introduces an anonymous declaration immediately followed by its definition.
+  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
+  The constructor $\constr{0}$ asserts that the next variable is of a channel type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover, to be used later on.
+  \todo{use the caption to give a name to the example, not to comment much. Complete/change the incomplete name that I gave}
 \end{example}
@@
 
-  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left]
+  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption= A typing derivation]
   $\func{p}$ : $\Process_1$
@@
   \end{lstlisting}
+  \todo{a better caption}
 \end{example}

### 2020-05-06 paper/paper.tex
Complete example todos

@@
   Let us illustrate the use of variable reference relations with an example in \autoref{lst:channels}.
-  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Variable referece]
+  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
   _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
 
-  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption= A typing derivation]
+  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing a process that sends part of one channel over another.]
   $\func{p}$ : $\Process_1$

### 2020-05-06 paper/paper.tex
more on aux results

@@
   The constructor $\constr{0}$ asserts that the next variable is of a channel type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover, to be used later on.
-  \todo{use the caption to give a name to the example, not to comment much. Complete/change the incomplete name that I gave}
 \end{example}
@@
   \end{lstlisting}
-  \todo{a better caption}
 \end{example}
@@
 
+\todo{sw and swap? Just use one}
+
 Swapping two variables preserves the well-typedness of a process.
@@
 
-\begin{proof}[Proof (Sketch)]
 All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
-  For details, refer to the mechanisation in Agda \cite{Zalakain2020Agda}.
-\end{proof}
+For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
 
@@
       Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
-      The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
+      %The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
     \item
@@
     \item
-      Case $\constr{scope-comm}$: in one direction we use swapping (\autoref{thm:swapping}).
-      In the reverse direction, we use swapping and then \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
+      Case $\constr{scope-comm}$:
+      we use swapping (\autoref{thm:swapping}) (and \autoref{lm:swap-swap} for the reverse direction) to show that swapping two elements in $P$ twice leaves $P$ unchanged.
   \end{itemize}
 \end{proof}
-
+\todo{remove the space between the itemize and the qed at the end}
 \subsection{Substitution}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\todo{generaliz* to generalis* everywhere}
 \autoref{thm:subst-generalization} proves a generalized version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
 \autoref{thm:substitution} instantiates the generalized version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
+\todo{explain what i is 0 mean}
 

### 2020-05-06 paper/paper.tex
on substitution

@@
   \renewcommand{\figureautorefname}{Fig.}%
+  \renewcommand{\appendixautorefname}{App.}%
 }
@@
   \end{multicols}
-  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
-  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
+  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
+  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
   Then substituting $i$ with $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-  Refer to \autoref{fig:subst} for a diagramatic representation.
+  \autoref{fig:subst} gives a diagramatic representation.
 \end{nitheorem}
@@
     \item
-      For constructor $\constr{end}$ observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
-      The details are found in \autoref{app:subst-generalization}.
-
-    \item
-      For constructor $\constr{chan}$ we proceed inductively.
-
-    \item
-      For constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$ we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
-      The details are found in \autoref{app:subst-generalization}.
+      Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
+
+    \item
+      Constructor $\constr{chan}$: we proceed inductively.
+
+    \item
+      Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
+      Full details are found in \autoref{app:subst-generalization}.
   \end{itemize}
@@
 \begin{proof}[Proof (Sketch)]
-  Use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
-  Use these to apply \autoref{thm:subst-generalization}.
+  W use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
+  Then, we use these to apply \autoref{thm:subst-generalization}.
 \end{proof}

### 2020-05-06 paper/paper.tex
Shorten swapping and subject congruence

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\autoref{fig:notation} illustrates the notation used in this paper.
 \begin{figure}[h]
@@
 
+\autoref{fig:notation} illustrates the notation used in this paper.
 Data type definitions ($\type{\N}$) use double lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
@@
 \begin{remark}
-  It is possible for the usage annotations of a variable $i$ within a process $P$ to be preserved as leftovers, yet for the variable $i$ to be mentioned by $P$: $P$ can choose to send $\lz$ usage annotations of $i$ over a channel.
+  Usage annotations of a variable $i$ within a process $P$ can be preserved as leftovers, yet the variable $i$ be mentioned by $P$: $P$ may send $\lz$ usage annotations of $i$ over a channel.
 \end{remark}
 
-\todo{sw and swap? Just use one}
-
 Swapping two variables preserves the well-typedness of a process.
-Let $\func{sw}_i$ be a function that swaps variables $i$ and $\suc i$.
-For simplicity, we overload $\func{sw}$ and use it to swap types in typing contexts and usage annotations in usage contexts.
-We use $\func{swap}_i \; P$ to swap variables $i$ and $\suc i$ within a process $P$, as introduced in structural congruence (\autoref{structural-congruence}).
+For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
 \begin{nitheorem}[Swapping]
@@
   Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, $\types{\func{sw}_i \; \gamma}{\func{sw}_i \; \Gamma}{\func{swap}_i \; P}{\func{sw}_i \; \Xi}$.
+  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
 \end{nitheorem}
-
-\begin{remark}
-  We limit ourselves to the swapping of adjacent variables $i$ and $\suc i$.
-  Doing so is enough to prove subject reduction.
-  More generally, any two variables can be swapped by a series of adjacent variable swappings.
-\end{remark}
 
@@
   The proof is by induction on \textsc{Equals} $\eq{}$.
-  Here we will only consider the base cases for $\constr{struct}$ (and their symmetric variants under $\constr{sym}$), which are nontrivial. Full proof in \cite{Zalakain2020Agda}.
+  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
+  Full proof in \cite{Zalakain2020Agda}.
   We proceed by induction on \textsc{StructCong} $\eqeq$:
@@
       Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
-      %The point at which $Q$ and $P$ join in $\comp{Q}{P}$ will no longer be the same.
     \item
@@
     \item
-      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we use \autoref{lm:types-unused} to show that $P$ preserves the usage annotations of the unused variable.
-      We then use strengthening (\autoref{thm:strengthening}) to delete the unused variable from $P$, and compose the result with the typing derivation on $Q$ wrapped in $\constr{chan}$.
-      In the reverse direction, we weaken (\autoref{thm:weakening}) $P$ so that it accepts the additional variable.
-      We then use \autoref{lm:lower-lift} to show that lowering and then lifting $P$ results in $P$.
+      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
+      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
     \item
       Case $\constr{scope-comm}$:
-      we use swapping (\autoref{thm:swapping}) (and \autoref{lm:swap-swap} for the reverse direction) to show that swapping two elements in $P$ twice leaves $P$ unchanged.
+      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
   \end{itemize}
@@
 \todo{remove the space between the itemize and the qed at the end}
+
 \subsection{Substitution}

### 2020-05-06 paper/paper.tex
Use \qedhere on proofs ending in itemize
Make the diagrams slightly smaller
on SR

--- 2020-05-06 paper/paper.tex
@@
 
-\begin{nidefinition}
-  For convenience, we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-\end{nidefinition}
-
+\begin{note}
+  We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+\end{note}
+\todo{use of well-typed:
+is well-typed replace with is well typed.
+A well-typed P is ok.}
+Subject reduction states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
+In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
 \begin{nitheorem}[Subject reduction]
-  If $P$ reduces to $Q$ and $P$ is well-typed, then so is $Q$.
-  In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$.
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$
   \begin{itemize}
-    \item If the reduction is internal, we prove $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-    \item If the reduction is external and on channel $i$, we prove that if $\containsusage{\Gamma}{i}{\lio}{\Delta}$ , then $\types{\gamma}{\Delta}{Q}{\Xi}$.
+    \item If $P \reduce{\constr{internal}} Q$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If $P \reduce{\constr{external} \; i} Q$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
   \end{itemize}
@@
     \item
-    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
-
-    \item
-    Reduction under $\constr{par}$ proceeds by induction on the process that is being reduced.
-
-    \item
-    Reduction under scope restriction case splits on the channel $c$ on which communication occurs.
+    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) to rearrange the assumptions and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
+
+    \item
+    Case $\constr{par}$: by induction on the process that is being reduced.
+
+    \item
+    Case $\constr{res}$: reduction under scope restriction case splits on the channel $c$ on which communication occurs.
     If $c$ is $\constr{internal}$ we proceed inductively.
@@
     \item
-    In the $\constr{struct}$ case, we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively.
+    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively.
   \end{itemize}

### 2020-05-06 paper/paper.tex

--- 2020-05-06 paper/paper.tex
@@
       Case $\constr{scope-comm}$:
-      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged.
+      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged. \qedhere
   \end{itemize}
 \end{proof}
-\todo{remove the space between the itemize and the qed at the end}
-
 \subsection{Substitution}
@@
     \begin{tikzpicture}
-      \node (gamma-i) at (0,6)   {$\Gamma_i$};
-      \node (gamma-m)   at (0,3)   {$\Gamma$};
+      \node (gamma-i) at (0,4)   {$\Gamma_i$};
+      \node (gamma-m) at (0,2)   {$\Gamma$};
       \node (gamma-j) at (0,0)   {$\Gamma_j$};
-      \node (xi-i)    at (3,5.5) {$\Xi_i$};
-      \node (theta)   at (3,3.3) {$\Theta$};
-      \node (delta-m) at (3,2.7) {};
+      \node (xi-i)    at (3,3.5) {$\Xi_i$};
+      \node (theta)   at (3,2.3) {$\Theta$};
+      \node (delta-m) at (3,1.7) {};
       \node (xi-j)    at (3,0.5) {$\Xi_j$};
-      \node (psi-i)   at (6,5)   {$\Psi_i$};
-      \node (psi-m)   at (6,3)   {$\Psi$};
+      \node (psi-i)   at (6,3)   {$\Psi_i$};
+      \node (psi-m)   at (6,2)   {$\Psi$};
       \node (psi-j)   at (6,1)   {$\Psi_j$};
@@
     \begin{tikzpicture}
-      \node (gamma-i) at (0,4)   {};
-      \node (gamma-m) at (0,2)   {};
+      \node (gamma-i) at (0,2)   {};
+      \node (gamma-m) at (0,1)   {};
       \node (gamma-j) at (0,0)   {};
-      \node (xi-i)    at (2,5)   {};
-      \node (xi-m)    at (2,3.5) {};
-      \node (xi-j)    at (2,2)   {};
-      \node (psi-m)   at (4,3.5) {};
+      \node (xi-i)    at (2,3)   {};
+      \node (xi-m)    at (2,2) {};
+      \node (xi-j)    at (2,1)   {};
+      \node (psi-m)   at (4,2) {};
 
@@
       Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
-      Full details are found in \autoref{app:subst-generalization}.
+      Full details are found in \autoref{app:subst-generalization}. \qedhere
   \end{itemize}

### 2020-05-06 paper/paper.tex
Fix small slip

@@
 \begin{proof}[Proof (Sketch)]
-  W use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
+  We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
   Then, we use these to apply \autoref{thm:subst-generalization}.

### 2020-05-06 paper/paper.tex
Change well-typed to well typed where relevant

@@
 The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in typing context $\gamma$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
-The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well-typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
+The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
 
@@
   \label{thm:framing}
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
+  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
   Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
@@
   \label{thm:weakening}
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, lifting every variable greater than or equal to $i$ in $P$ is well-typed in
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
   $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
@@
   \label{thm:strengthening} 
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
   Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than $i$ in $P$ is well-typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
+  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
 \end{nitheorem}
@@
   \label{thm:swapping}
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
   Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
@@
 
-Subject congruence states that applying structural congruence rules to a well-typed process preserves its well-typedness.
+Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.
 \begin{nitheorem}[Subject congruence]
@@
   \label{thm:subst-generalization}
-  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
   The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
@@
   Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
-  Then substituting $i$ with $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Then substituting $i$ with $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
   \autoref{fig:subst} gives a diagramatic representation.
@@
   \label{thm:substitution}
-  Let $P$ be a well-typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well-typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
 \end{nitheorem}
@@
 \end{note}
-\todo{use of well-typed:
-is well-typed replace with is well typed.
-A well-typed P is ok.}
 Subject reduction states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
@@
 \begin{nitheorem}[Subject reduction]
-  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$
   \begin{itemize}

### 2020-05-06 paper/paper.tex
generalize -> generalise

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\todo{generaliz* to generalis* everywhere}
-\autoref{thm:subst-generalization} proves a generalized version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
-\autoref{thm:substitution} instantiates the generalized version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
+\autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
+\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
 \todo{explain what i is 0 mean}

### 2020-05-06 paper/paper.tex
Clarify what 0 is

@@
 \autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
-\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is $\constr{0}$, as required by subject reduction.
-\todo{explain what i is 0 mean}
+\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
 

### 2020-05-06 paper/paper.tex
Do not number equations
Clarifications

@@
   \begin{figure}[h]
-    \begin{equation}
+    \begin{equation*}
       \begin{split}
@@
       \end{split}
-    \end{equation}
+    \end{equation*}
     \caption{Indexed set of usage algebras.}
@@
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
+  Let $i$ be a variable not in $P$, such that $\Unused_i \; P$.
   Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
@@
 \begin{remark}
-  Usage annotations of a variable $i$ within a process $P$ can be preserved as leftovers, yet the variable $i$ be mentioned by $P$: $P$ may send $\lz$ usage annotations of $i$ over a channel.
+  The \emph{unusedness} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
 \end{remark}
@@
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
+  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
 \end{nitheorem}

### 2020-05-06 paper/paper.tex
Be more careful with the example derivation

@@
   Let $\func{\omega}$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
-  Let $\epsilon$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
+  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
   To type $\func{p}$, we need a channel over which we can transmit another channel.
@@
   \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing a process that sends part of one channel over another.]
-  $\func{p}$ : $\Process_1$
+  $\func{p}$ : $\Process_{\suc \constr{zero}}$
   $\func{p}$ = $\new{}$
@@
 
-  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\epsilon}$
+  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{epsilon}}$
   $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\func{\omega}} \; \li \; \one \; (\constr{comp}$

### 2020-05-06 paper/paper.tex
No need for parentheses around tuples

@@
 
-  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$ for a channel which cannot be used any further, $\li$ for the pair $(1 \comma 0)$ for a channel to be used in input exactly once, $\lo$ for $(0 \comma 1)$ for a channel to be used in output exactly once, and $\lio$ for $(1 \comma 1)$, for a channel to be used exactly once in input and exactly once in output.
+  Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
   This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
@@
     All the properties of the algebra follow trivially.
-    As an example, the multiplicity pair $\ell_i$ is represented as $(\constr{one} \comma \constr{zero})$.
+    As an example, the multiplicity pair $\ell_i$ is represented as $\constr{one} \comma \constr{zero}$.
 
@@
     All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
-    As an example, we use $(5 \comma 3)$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.
+    As an example, we use $5 \comma 3$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.
 
@@
     All the properties of the algebra follow trivially.
-    Shared channels have multiplicities $(\constr{\omega} \comma \constr{\omega})$, which they will forever preserve.
+    Shared channels have multiplicities $\constr{\omega} \comma \constr{\omega}$, which they will forever preserve.
 \end{description}
@@
   This multiplicity $y$ is used for both input and output, so that they are balanced.
-  The continuation process $P$ is provided with the new channel with usage annotation $(y \comma y)$, which it must completely exhaust.
+  The continuation process $P$ is provided with the new channel with usage annotation $y \comma y$, which it must completely exhaust.
 
@@
 
-  Let $\func{\omega}$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
-  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
+  Let $\func{\omega}$ be a shortcut for $\one \comma \one$ in an algebra of shared types.
+  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $\zero \comma \zero$.
   To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
+  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
   The left-hand process is typed following its syntax (line 8).
@@
   The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
-  Its usage annotation is $(\one \comma \one)$: we will transmit $(\one \comma \zero)$ and save $(\zero \comma \one)$ for the continuation process.
+  Its usage annotation is $\one \comma \one$: we will transmit $\one \comma \zero$ and save $\zero \comma \one$ for the continuation process.
 

### 2020-05-06 paper/paper.tex
Mention work by Ferreira

@@
     \item
-    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively.
+    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qedhere
   \end{itemize}
@@
 
+Castro et al.
+provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- where de Bruijn indices are not as popular as dependent patter matching is not as straightforward to handle.
+As an example use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
+
 Taking an intrinsic approach to typing, Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.

### 2020-05-06 paper/paper.tex
Fix slip
Fix mistake in subject reduction

@@
 \begin{nitheorem}[Subject reduction]
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
   \begin{itemize}
-    \item If $P \reduce{\constr{internal}} Q$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-    \item If $P \reduce{\constr{external} \; i} Q$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
+    \item If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
   \end{itemize}
@@
 \begin{proof}
-  By induction on $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  By induction on $P \reduce{c} Q$.
   \hfill{}\\
@@
 As we do, he uses de Bruijn indices, a reduction relation and a type system posterior to the syntax.
-However, in his work context splits are explicit, and the types of used multiplicities erased.
+However, in his work context splits are extrinsic, and the types of used multiplicities erased.
 

### 2020-05-06 paper/paper.tex
Shorten subject reduction a little

@@
     \item
-    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) to rearrange the assumptions and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
+    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
 
@@
     \item
-    Case $\constr{res}$: reduction under scope restriction case splits on the channel $c$ on which communication occurs.
-    If $c$ is $\constr{internal}$ we proceed inductively.
-    If $c$ is $\constr{external}\; \constr{0}$ -- i.e.
-the channel that is being introduced by the scope restriction -- we use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and proceed inductively.
-    If $c$ is $\constr{external}\; (\suc i)$ we proceed inductively.
+    Case $\constr{res}$: case split on channel $c$:
+    if $\constr{internal}$ proceed inductively;
+    if $\constr{external}\; \constr{0}$ (i.e. the channel introduced by scope restriction) use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and proceed inductively;
+    if $\constr{external}\; (\suc i)$ proceed inductively.
 

### 2020-05-06 paper/paper.tex
Better phrasing
Keep conventions up

@@
     &\func{\cdot-unique^l} &:{} &\forall x y y' z  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
-    &\func{0\cdot-min^l}   &:{} &\forall y z       & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
+    &\func{\cdot-min^l}   &:{} &\forall y z        & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
     &\func{\cdot-id^l}     &:{} &\forall x         & \to \; & \op{x}{\zero}{x} \\
@@
 The work by Goto et al. \cite{Goto2016a} is likely the first formal contribution to session-types that comes along with a mechanised proof of type safety (in Coq). 
-Their paper extends session types with polymorphism and pattern matching on session types.
+In their paper, they extend session types with polymorphism and pattern matching on session types.
 They use a locally-nameless encoding for variable references, a syntax prior to types, and a LTS semantics that encodes session-typed processes into the \picalc{}.

### 2020-05-06 paper/paper.tex
Improve the wain in which \constains is spelled out

@@
 
-The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ is of type $t$ in typing context $\gamma$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
+The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
 The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.

### 2020-05-06 paper/paper.tex
For uniformity, always mention textsc-d type name in definition titles

@@
 
-\begin{nidefinition}[Types and typing contexts]
+\begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
   A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (\autoref{fig:prectx}).
@@
 
-\begin{nidefinition}[Usage contexts]
+\begin{nidefinition}[\textsc{Ctx}: usage contexts]
   A \emph{usage context} is a context $\Ctx_{idxs}$ of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).

### 2020-05-06 paper/appendix.tex
a first pass to results

@@
 %% Appendix
+\section{Lemmas}
+\label{app:lemmas}
+
+\begin{nilemma}
+  \label{lm:subst-unused}
+  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:lower-lift}
+  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:swap-swap}
+  The function $\func{swap}_i \; P$ is its own inverse: $\func{swap}_i \; (\func{swap}_i \; P) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:types-unused}
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:comm-capable}
+  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
+\end{nilemma}
+
+\begin{proof}
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\end{proof}
+
 \section{Structural Congruence}
@@
 
-\section{Lemmas}
-\label{app:lemmas}
+Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.
+\begin{nitheorem}[Subject congruence]
+  \label{thm:subject-congruence1}
+  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+\end{nitheorem}
 
-\begin{nilemma}
-  \label{lm:subst-unused}
-  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:swap-swap}
-  The function $\func{swap}_i \; P$ is its own inverse: $\func{swap}_i \; (\func{swap}_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:types-unused}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:comm-capable}
-  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
-
-\begin{proof}
-  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\begin{proof}[Proof (Sketch)]
+  The proof is by induction on \textsc{Equals} $\eq{}$.
+  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
+  Full proof in \cite{Zalakain2020Agda}.
+  We proceed by induction on \textsc{StructCong} $\eqeq$:
+  \begin{itemize}
+    \item
+      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
+    \item
+      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
+    \item
+      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
+    \item
+      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
+      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
+    \item
+      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
+      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
+    \item
+      Case $\constr{scope-comm}$:
+      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged. \qedhere
+  \end{itemize}
 \end{proof}
@@
 \begin{nitheorem}[Generalised substitution]
-  \label{thm:subst-generalization}
+  \label{thm:subst-generalization1}
   Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.

### 2020-05-06 paper/paper.tex
a first pass to results

@@
 
-In \autoref{auxiliary-results} we present some auxiliary results: framing (\autoref{thm:framing}), weakening (\autoref{thm:weakening}), strengthening (\autoref{thm:strengthening}) and swapping of variables (\autoref{thm:swapping}).
-We present subject congruence in \autoref{subject-congruence} and substitution in \autoref{substitution}.
-Finally, we present subject reduction in \autoref{subject-reduction}.
+We now discuss the type safety of our type system, starting with auxiliary results.
 
@@
 
-\subsection{Subject Congruence}
+\subsection{Subject Reduction}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{subject-congruence}
+\label{substitution}
+\label{subject-reduction}
+
+In this section we move towards subject reduction.
+We start with two important results: subject congruence (\autoref{thm:subject-congruence}) and substitution (\autoref{thm:subst-generalization} and \autoref{thm:substitution}).
 
@@
   Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
-  Full proof in \cite{Zalakain2020Agda}.
-  We proceed by induction on \textsc{StructCong} $\eqeq$:
-  \begin{itemize}
-    \item
-      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
-    \item
-      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
-    \item
-      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
-    \item
-      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
-      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
-    \item
-      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
-      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
-    \item
-      Case $\constr{scope-comm}$:
-      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged. \qedhere
-  \end{itemize}
+  Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
 \end{proof}
-\subsection{Substitution}
-\label{substitution}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
-  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
-  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
+  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
   \begin{multicols}{2}
@@
   \end{multicols}
+\end{nitheorem}
+
+\begin{proof}[Proof (Sketch)]
+  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$. Full details are found in \autoref{app:subst-generalization}.
+\end{proof}
+
+The above theorem means that the substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
+  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
+
   Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
@@
   \autoref{fig:subst} gives a diagramatic representation.
-\end{nitheorem}
-
-\begin{proof}[Proof (Sketch)]
-  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  \begin{itemize}
-    \item
-      Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
-
-    \item
-      Constructor $\constr{chan}$: we proceed inductively.
-
-    \item
-      Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
-      Full details are found in \autoref{app:subst-generalization}. \qedhere
-  \end{itemize}
-\end{proof}
 
@@
 
-\subsection{Subject Reduction}
-\label{subject-reduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 \begin{note}

### 2020-05-06 paper/paper.tex
First pass to type safety

@@
 
-We now discuss the type safety of our type system, starting with auxiliary results.
-
-\subsection{Auxiliaty Results}
-\label{auxiliary-results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subparagraph*{Framing}
 Framing asserts that the well-typedness of a process is independent of its leftover resources.
+
 \begin{nitheorem}[Framing]
@@
 
+\subparagraph*{Weakening}
 Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
@@
 
+\subparagraph*{Strengthening}
 Removing an unused variable preserves the well-typedness of a process.
@@
 
+\subparagraph*{Swapping}
 Swapping two variables preserves the well-typedness of a process.
@@
 
+\begin{proof}[Proof (Sketch)]
   All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
   For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
-
-\subsection{Subject Reduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{subject-congruence}
-\label{substitution}
-\label{subject-reduction}
-
-In this section we move towards subject reduction.
-We start with two important results: subject congruence (\autoref{thm:subject-congruence}) and substitution (\autoref{thm:subst-generalization} and \autoref{thm:substitution}).
-
+\end{proof}
+
+\subparagraph*{Subject Congruence}
 Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.
+
 \begin{nitheorem}[Subject congruence]
@@
   The proof is by induction on \textsc{Equals} $\eq{}$.
-  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
+  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:swapping}.
   Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
 \end{proof}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subparagraph*{Substitution}
 \autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
 \autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
+
+\begin{nitheorem}[Generalised substitution]
+  \label{thm:subst-generalization}
+  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
+  \begin{multicols}{2}
+  \begin{itemize}
+    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
+    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
+    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
+  \end{itemize}
+  \end{multicols}
+  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
+  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
+  Then substituting $i$ with $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+\end{nitheorem}
+
+\begin{proof}[Proof (Sketch)]
+  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
+  Full details are found in \autoref{app:subst-generalization}.
+  \begin{itemize}
+   \item
+     Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
+   \item
+     Constructor $\constr{chan}$: we proceed inductively.
+   \item
+     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
+ \end{itemize}
+\end{proof}
 
@@
 
-\begin{nitheorem}[Generalised substitution]
-  \label{thm:subst-generalization}
-  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
-  \begin{multicols}{2}
-  \begin{itemize}
-    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
-    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
-    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
-    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
-  \end{itemize}
-  \end{multicols}
-\end{nitheorem}
-
-\begin{proof}[Proof (Sketch)]
-  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$. Full details are found in \autoref{app:subst-generalization}.
-\end{proof}
-
-The above theorem means that the substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
-  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
-
-  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
-  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
-  Then substituting $i$ with $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-  \autoref{fig:subst} gives a diagramatic representation.
-
 \begin{nitheorem}[Substitution]
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{note}
@@
 \end{note}
+
+\subparagraph*{Subject Reduction}
 Subject reduction states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
 In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
+
 \begin{nitheorem}[Subject reduction]

### 2020-05-06 paper/paper.tex
Add "(Sketch)" to SR proof
Use subparagraph title as the start of intro sentences

@@
 \subparagraph*{Framing}
-Framing asserts that the well-typedness of a process is independent of its leftover resources.
+asserts that the well-typedness of a process is independent of its leftover resources.
 
@@
 \subparagraph*{Weakening}
-Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
 Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
@@
 \subparagraph*{Strengthening}
-Removing an unused variable preserves the well-typedness of a process.
+states that removing an unused variable preserves the well-typedness of a process.
 Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
@@
 \subparagraph*{Swapping}
-Swapping two variables preserves the well-typedness of a process.
+two variables preserves the well-typedness of a process.
 For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
@@
 \subparagraph*{Subject Congruence}
-Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.
+states that applying structural congruence rules to a well typed process preserves its well-typedness.
 
@@
 \subparagraph*{Substitution}
+\todo{intro}
 \autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
@@
    \item
-     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction.
+     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
  \end{itemize}
@@
 \subparagraph*{Subject Reduction}
-Subject reduction states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
+states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
 In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
@@
 
-\begin{proof}
+\begin{proof}[Proof (Sketch)]
   By induction on $P \reduce{c} Q$.

### 2020-05-06 paper/paper.tex
Replace references to sections in type safety with references to theorems

@@
    This operational semantics is given as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (\autoref{subject-reduction}).
+   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
    The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
@@
   Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
-  We use these to prove \emph{subject congruence} (\autoref{subject-congruence}), and together with \emph{substitution} (\autoref{substitution}), to prove \emph{subject reduction} (\autoref{subject-reduction}).
+  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
 
@@
 The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs (\autoref{fig:reduction}).
-We keep track of channel $c$ so we can state subject reduction (\autoref{subject-reduction}).
+We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
 We distinguish between channels that are created inside of the process ($\constr{internal}$) and channels that are created outside ($\constr{external}\ i$, where $i$ is the index of the channel variable).
@@
 \begin{nitheorem}[Subject reduction]
+  \label{thm:subject-reduction}
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.

### 2020-05-06 paper/paper.tex
pass on conclusions

--- 2020-05-06 paper/paper.tex
@@
 
-Leftover typing as presented in this paper is based on work by Allais \cite{Allais2018a}.
-Allais uses leftover typing to mechanise in Agda a bidirectional type system for the linear \lambdacalc{}.
-This type system sits on top of an untyped but well-scoped syntax.
+Allais \cite{Allais2018a} uses leftover typing to mechanise in Agda a bidirectional type system for the linear \lambdacalc{}.
 He proves type preservation and provides a decision procedure for type checking and type inference.
-In this paper, we apply his technique to the \picalc{}, but limit ourselves to type-checking terms (scope restriction needs usage annotations), and focus on generalising the usage algebra.
-Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), it is crucial for us too to be able to talk about fully consumed resources.
-As a by-product, we are able to transmit over a channel all none multiplicities of a fully exhausted channel.
-
-Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi, but perhaps one of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
-Gay encodes the \picalc{} with linear and shared types into a more general framework.
-As we do, he uses de Bruijn indices, a reduction relation and a type system posterior to the syntax.
-However, in his work context splits are extrinsic, and the types of used multiplicities erased.
-
-The work by Goto et al. \cite{Goto2016a} is likely the first formal contribution to session-types that comes along with a mechanised proof of type safety (in Coq). 
-In their paper, they extend session types with polymorphism and pattern matching on session types.
-They use a locally-nameless encoding for variable references, a syntax prior to types, and a LTS semantics that encodes session-typed processes into the \picalc{}.
-Their type system uses reordering of contexts and extrinsic contex splits.
-
+In our paper, we apply leftover typing from Allais \cite{Allais2018a} to the \picalc{} and focus on type checking.
+We give a more general usage algebra, leading to linear, graded and shared type systems.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work we are also able to talk about fully consumed resources.
+As such, we can transmit over a channel all none multiplicities of a fully exhausted channel, namely a channel of multiplicity $\lz$.
+
+Recent years have seen an increase in effort to mechanise resource-aware concurrency calculi.
+One of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
+Gay encodes the \picalc{} with linear and shared types using de Bruijn indices, a reduction relation and a type system posterior to the syntax.
+However, in his work context splits are extrinsic, and the types of used multiplicities erased, while we adopt a more general usage algebra.
+Our framework integrates three different type systems and their combination in one and our linear \picalc{} uses leftover typing for the first time.
+%
+The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first mechanisation of session types and their type safety proof (in Coq). 
+The authors extend session types with polymorphism and pattern matching.
+They use a locally-nameless encoding for variable references, a syntax prior to types, and an LTS semantics that encodes session-typed processes into the \picalc{}.
+Their type system uses reordering of contexts and extrinsic context splits. 
+%
 Castro et al.
-provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- where de Bruijn indices are not as popular as dependent patter matching is not as straightforward to handle.
-As an example use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
-
+provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- while de Bruijn indices are not as popular as dependent pattern matching, the latter is not as straightforward to handle.
+As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
+%
 Taking an intrinsic approach to typing, Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
@@
 The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
-
+%
 Most recently, Rouvoet et al. provide an intrinsic type system for a \lambdacalc{} with session types \cite{Rouvoet2020}.
@@
 They distill their typing rules even further by modelling the supply and demand market as a state monad.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types, and show that reduction preserves the well-typedness of processes.
-We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.
-Our work is mechanised in roughly 1500 lines of code.
-
-For the future, we consider proving that reduction preserves the balancing of channels long-hanging fruit.
-More involved, we intend to add support for products and sum types to both our syntax and our type system.
+\todo{related work above is so thoroughly presented. well done!
+more to be added. Check what Cinzia di Giusto, Kirsten Peters, Antonio Ravara, Jorge Perez if they have any work published on mecahnisation (I know they are working on it).
+Check Jesper, and Ivan Scagneto who has worked on mechanisiation of pi.
+Check Brigite Pienta and her work on Beluga.
+Next, check theories on linear pi or other languages and what is the world saying on linearity.
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+To conclude, in this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types, under the same unified framework and show that reduction preserves the well typedness of processes.
+We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.
+As a result, theorems like weakening and strengthening can now we stated for the linear \picalc{}.
+Our work is fully mechanised in around 1500 lines of code in Agda \cite{Zalakain2020Agda}.
+
+As future work, we intend to prove further properties of our type system, such as reduction preserves the balancing of channels.
+We intend to add support for products, sum types and recursion to both our syntax and our type system.
 Orthogonally, making our typing rules bidirectional would allows us to provide a decision procedure for type checking processes in a given set of algebras.
-Following Erika Kreuer's suggestion, it might also be worth identifying correspondences between our usage algebra and particular state machines.
+Furthermore, it might also be worth identifying correspondences between our usage algebra and particular state machines.
+Finally, we will use this our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
 

### 2020-05-06 paper/paper.tex
new bib

@@
 Notably, it features channel mobility: channels themselves are first order values, and can therefore be transmitted.
-The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
+The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and (more recently) Go \footnote{https://golang.org}.
 At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types.
@@
 In a parallel track, J.-Y.
-Girard's development of linear logic \cite{Girard87} in the early '90s opened new avenues in computing science by introducing \emph{linear types} for (functional) programming languages \cite{Curry-Howard,Wadler90,Bernardy2018}.
+Girard's development of linear logic \cite{Girard87} in the early '90s opened new avenues in computing science by introducing \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018}.
 Linear type systems guarantee that resources are used \emph{exactly once}.

### 2020-05-06 paper/paper.tex
Last touches

@@
 \end{abstract}
-
-\todo{spellcheck}
-\todo{make sure all citations are there}
-\todo{lemmas in appendix are referred to from here}
-\todo{shorten: we are slightly over the limit}
 
@@
 
-Recent years have seen an increase in effort to mechanise resource-aware concurrency calculi.
+Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi.
 One of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
 Gay encodes the \picalc{} with linear and shared types using de Bruijn indices, a reduction relation and a type system posterior to the syntax.
-However, in his work context splits are extrinsic, and the types of used multiplicities erased, while we adopt a more general usage algebra.
-Our framework integrates three different type systems and their combination in one and our linear \picalc{} uses leftover typing for the first time.
+However, in his work context splits are extrinsic, and the types of used multiplicities erased.
+We make context splits intrinsic, preserve the ability to talk about used resources, and adopt a more general usage algebra: our framework integrates three different type systems into one.
+
 %
-The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first mechanisation of session types and their type safety proof (in Coq). 
+The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types that includes a type safety proof (in Coq). 
 The authors extend session types with polymorphism and pattern matching.
@@
 Castro et al.
-provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- while de Bruijn indices are not as popular as dependent pattern matching, the latter is not as straightforward to handle.
+provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
 As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
@@
 Taking an intrinsic approach to typing, Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
-Context splits are still given extrinsically, and exhausted elements are removed from typing contexts altogether.
+Context splits are still given extrinsically, and exhausted resources are removed from typing contexts altogether.
 The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
@@
 They distill their typing rules even further by modelling the supply and demand market as a state monad.
-\todo{related work above is so thoroughly presented. well done!
-more to be added. Check what Cinzia di Giusto, Kirsten Peters, Antonio Ravara, Jorge Perez if they have any work published on mecahnisation (I know they are working on it).
-Check Jesper, and Ivan Scagneto who has worked on mechanisiation of pi.
-Check Brigite Pienta and her work on Beluga.
-Next, check theories on linear pi or other languages and what is the world saying on linearity.
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-To conclude, in this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types, under the same unified framework and show that reduction preserves the well typedness of processes.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+To conclude, in this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
 We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.
-As a result, theorems like weakening and strengthening can now we stated for the linear \picalc{}.
+As a result, theorems like framing, weakening and strengthening can now we stated for the linear \picalc{}.
 Our work is fully mechanised in around 1500 lines of code in Agda \cite{Zalakain2020Agda}.
 
-As future work, we intend to prove further properties of our type system, such as reduction preserves the balancing of channels.
+As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
 We intend to add support for products, sum types and recursion to both our syntax and our type system.
-Orthogonally, making our typing rules bidirectional would allows us to provide a decision procedure for type checking processes in a given set of algebras.
+Orthogonally, making our typing rules bidirectional would allow us to provide a decision procedure for type checking processes in a given set of algebras.
 Furthermore, it might also be worth identifying correspondences between our usage algebra and particular state machines.
-Finally, we will use this our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
+Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
 

### 2020-05-06 paper/paper.tex
Forgotten todo
Fix github link
Introduce some suggested changes

@@
 \usepackage{url}
-\usepackage{todonotes}
 \usepackage{amssymb}
@@
 \usepackage{mathpartir}
-
-% TODO notes -- TODO: remove
-\usepackage{todonotes}
 
@@
 \keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
-\supplement{\url{https://github/umazalakain/typing-linear-pi}}
+\supplement{\url{https://github.com/umazalakain/typing-linear-pi}}
 \acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
@@
   It has an element $\one$ that is used to count inputs and outputs.
-  \autoref{fig:multiplicities} uses a terniary relation to model such an algebra as a record $\Algebra_C$ on a carrier $C$.
+  \autoref{fig:multiplicities} uses a ternary relation to model such an algebra as a record $\Algebra_C$ on a carrier $C$.
 
@@
 
-  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{epsilon}}$
+  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{\epsilon}}$
   $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\func{\omega}} \; \li \; \one \; (\constr{comp}$
@@
 \subparagraph*{Substitution}
-\todo{intro}
 \autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
@@
 Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work we are also able to talk about fully consumed resources.
-As such, we can transmit over a channel all none multiplicities of a fully exhausted channel, namely a channel of multiplicity $\lz$.
+As such, we can transmit over a channel $\lz$ multiplicities of a fully exhausted channel.
 

### 2020-05-07 paper/paper.tex
abstract and intro

@@
   The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
-  It is an underlying theoretical and practical framework on top of which more advanced types and theories are built, including \emph{session types} \cite{H93,THK94,HVK98}.
-  Linearity is key for type safety in session communication.
+  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, including \emph{session types} \cite{H93,THK94,HVK98}.
+  %Linearity is key for type safety in session communication.
 
@@
   For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
-  We generalise the algebras on multiplicities, allowing the developer to choose a mix of linear, graded and shared typing.
-  We provide framing, weakening and strengthening proofs that we then use to prove subject congruence.
+  We generalise the algebras on multiplicities, allowing the user to choose a mix of linear, graded and shared typing.
+  We provide framing, weakening and strengthening theorems, which can now be stated for the linear \picalc{} and use them to prove subject congruence.
   We show that the type system is stable under substitution and prove subject reduction.
@@
 In a parallel track, J.-Y.
-Girard's development of linear logic \cite{Girard87} in the early '90s opened new avenues in computing science by introducing \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018}.
+Girard's development of linear logic \cite{Girard87} opened new avenues in computing science in the early '90s by introducing \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018}.
 Linear type systems guarantee that resources are used \emph{exactly once}.
@@
 Recent work has pushed the linear \picalc{} into the spotlight again \cite{KPT96} thanks to an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17}.
-This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful at a practical level too.
+This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice, too.
 The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
@@
 
-We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{}.
-The developer is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
+We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{} for the first time.
+The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
+This allows for linear, graded and shared types to be used either separately or mixed in a type system.
 Ultimately, we can exploit this formalisation as a unified framework for type safe distributed modelling and programming with a variety of type systems, and build on top of it more advanced types, theories and languages.
-In particular, while linearity is needed for type safety in communication-centric programming, graded and shared types are needed to model real-world systems.
+In particular, while linearity is needed for type safety in communication-centric programming, graded and shared types are needed to model real-world software systems.
 
@@
  \begin{itemize}
-   \item \textbf{Syntax}: \autoref{syntax} uses type level de Bruijn indices \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well-scoped by construction}: every free variable is accounted for in the type of the process that uses it.
+   \item \textbf{Syntax}: \autoref{syntax} uses type level \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it.
 
    \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
-   This operational semantics is given as a reduction relation on processes (\autoref{operational-semantics}).
+   This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
    The reduction relation tracks at the type level the channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
@@
 
-    Benefits of typing with leftovers include: removing the need for extrinsic context splits, which are rendered unnecessary; and for the first time we can state theorems like \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) in the linear \picalc{}.
+    Benefits of typing with leftovers include: removing the need for extrinsic context splits, which are rendered unnecessary; and for the first time we can state theorems like \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the linear \picalc{}.
   \end{itemize}
-  \item \textbf{Machine verified proofs of subject reduction and auxiliary theorems}:
+  \item \textbf{Fully mechanised formalisation and proofs of the \picalc{} with leftovers}:
+  The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
+
   \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
@@
   We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
-
-  \item \textbf{Fully mechanised formalisation available in Agda}:
-  This work has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
 

### 2020-05-07 paper/paper.tex
a pass on type system

@@
   We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
-
+%
   Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
@@
 The unit type $\unit$ serves as a proof of inhabitance for types.
-The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g.
-$0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
-The type $\channel{t}{x}$ of a channel determines what type of data ${t}$ and what usage annotations ${x}$ are sent over that channel.
-Notice that this notation is similar to Kobayashi's $[t]\mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
+The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
+The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.
+This notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
 
@@
 
-  A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to the existing context $\Gamma$.
+  A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to an existing context $\Gamma$.
 
@@
     No need for an extensional context splitting proof for parallel composition.
-    Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
-    This could be obtained through automated search --- but it is better yet not to need it.
+    % Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
+    % This could be obtained through automated search --- but it is better yet not to need it.
     Leftover typing uses the leftovers of $P$ to type $Q$.
   \item \textbf{New theorems}:
-    For the first time, theorems like \textbf{framing} (\autoref{thm:framing}), \textbf{weakening} (\autoref{thm:weakening}) and \textbf{strengthening} (\autoref{thm:strengthening}) can be stated within the linear \picalc{} thanks to the presence of a leftover context.
+    For the first time, theorems like \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) can be stated within the linear \picalc{} thanks to the presence of a leftover context.
 \end{itemize}
@@
   \label{example-derivations}
-  We will now illustrate a typing derivation of the process $\func{p}$ using linear and shared types (\autoref{lst:derivation}).
+  In \autoref{lst:derivation} we illustrate a typing derivation of the process $\func{p}$ (lines 1-4) using linear and shared types.
 
@@
   We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
-  The left-hand process is typed following its syntax (line 8).
-  On the right-hand we need the channel that is going to be sent (line 9).
+  The left-hand process (line 3) is typed following its syntax (line 8).
+  For the right-hand (line 4) we need the channel that is going to be sent (line 9).
   The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.

### 2020-05-08 paper/paper.tex
more structure needed
type safety was too long and needed to be more structured...
functional -> deterministic and decidable

@@
     Linear types, graded types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
-    Any \emph{partial commutative monoid} that is \emph{functional}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
+    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
 
@@
 \begin{nidefinition}[Usage algebra]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{functional}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{decidable}, \emph{deterministic}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
-  We use the \emph{functionality} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
+  We use the \emph{decidability} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts. 
@@
 \label{type-safety}
+\subsection{Auxiliary Results}
 
@@
 
+\subsection{Towards Subject Reduction}
 \subparagraph*{Subject Congruence}
@@
 They use proof relevant separation logic and a notion of a supply and demand \emph{market} to make context splits transparent to the user.
-Their separation logic is based on a partial commutative monoid that need not be functional nor cancellative.
+Their separation logic is based on a partial commutative monoid that need not be deterministic nor cancellative.
 Their typing rules preserve the balance between supply and demand, and are extremely elegant.
@@
 Our work is fully mechanised in around 1500 lines of code in Agda \cite{Zalakain2020Agda}.
-
+%
 As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.

### 2020-05-08 paper/paper.tex
Comment on Pientka's work and further work on nonlinear pi calculus

@@
 
+Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
+The result is a meta-theory in which HOAS encodings with both linear and dependent types can be described.
+The developer obtains for free an equational theory of substitution and decidable typechecking without having to encode context splits within the object language.
+
+% Orchard
+\cite{Orchard}
+
+Further work on the mechanisation of the \picalc{} include \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, where the authors mechanise non-linear variations of the \picalc{}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-05-08 paper/paper.tex
Draft on Orchard's work

@@
 
+Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged language with graded modal types, linear types, indexed types and polymorphism.
+Modalities include exact usages, security levels and intervals.
+Grade terms allow for grade polymorphism and their resource algebra is based on a pre-ordered semiring with monotonic multiplication and partial addition.
+They provide bidirectional typing rules, and show the type safety of their semantics.
+
 Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi.
@@
 
-Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
+More generally, based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
 The result is a meta-theory in which HOAS encodings with both linear and dependent types can be described.
 The developer obtains for free an equational theory of substitution and decidable typechecking without having to encode context splits within the object language.
-
-% Orchard
-\cite{Orchard}
 

### 2020-05-08 paper/paper.tex
Rephrase slightly
a pass to related work

--- 2020-05-08 paper/paper.tex
@@
 \subparagraph*{Swapping}
-two variables preserves the well-typedness of a process.
-For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
+of two variables preserves the well-typedness of a process.
+We extend $\func{swap}_i$  introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
 \begin{nitheorem}[Swapping]
@@
 He proves type preservation and provides a decision procedure for type checking and type inference.
-In our paper, we apply leftover typing from Allais \cite{Allais2018a} to the \picalc{} and focus on type checking.
+In our paper, we apply leftover typing to the \picalc{} following Allais \cite{Allais2018a}, and focus on type checking.
 We give a more general usage algebra, leading to linear, graded and shared type systems.
-Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work we are also able to talk about fully consumed resources.
-As such, we can transmit over a channel $\lz$ multiplicities of a fully exhausted channel.
-
-Recent years have seen an increase in the efforts to mechanise resource-aware concurrency calculi.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work, we too are able to talk about fully consumed resources.
+As such, we can transmit over a channel the $\lz$ multiplicities of a fully exhausted channel.
+%
+Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged functional language with graded modal types, linear types, indexed types and polymorphism.
+Modalities include exact usages, security levels and intervals.
+%Grade terms allow for grade polymorphism and their resource algebra is based on a pre-ordered semiring where addition is partial.
+%They provide bidirectional typing rules, and show the type safety of their semantics.
+Their types resemble our graded types, but we focus on concurrency and one of our goals is mechanisation in Agda.
+
+Recent years have seen an increase in the efforts to mechanise resource-aware process algebras.
 One of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
@@
 We make context splits intrinsic, preserve the ability to talk about used resources, and adopt a more general usage algebra: our framework integrates three different type systems into one.
-
 %
@@
 Castro et al.
-provide tooling for locally-nameless representations of process calculi in Coq \cite{Castro2020} --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
+\cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq  --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
 As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
@@
 They distill their typing rules even further by modelling the supply and demand market as a state monad.
-
+%
 Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
@@
 
-% Orchard
-\cite{Orchard}
-
-Further work on the mechanisation of the \picalc{} include \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, where the authors mechanise non-linear variations of the \picalc{}.
+Further work on mechanising the \picalc{} \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, focuses on  non-linear variations, whether we present a range of linear, graded and shared types.
 

### 2020-05-08 paper/paper.tex
On Orchard's work

@@
 Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged functional language with graded modal types, linear types, indexed types and polymorphism.
-Modalities include exact usages, security levels and intervals.
-%Grade terms allow for grade polymorphism and their resource algebra is based on a pre-ordered semiring where addition is partial.
-%They provide bidirectional typing rules, and show the type safety of their semantics.
-Their types resemble our graded types, but we focus on concurrency and one of our goals is mechanisation in Agda.
+Modalities include exact usages, security levels and intervals; resource algebras are pre-ordered semirings with partial addition.
+They provide bidirectional typing rules, and show the type safety of their semantics.
 

### 2020-05-25 paper/paper.tex
back to long names
Small corrections
Fix silly mistake in running title
Make z be decidable in x := y · z

@@
 
-% Inference rules
+% inference rules
 \usepackage{mathpartir}
@@
 \newtheorem{nilemma}{Lemma}
-\renewcommand{\nitheoremautorefname}{Thm.}%
-\renewcommand{\nidefinitionautorefname}{Def.}%
-\renewcommand{\nilemmaautorefname}{Lem.}%
-\renewcommand{\exampleautorefname}{Ex.}%
+% \renewcommand{\nitheoremautorefname}{Thm.}%
+% \renewcommand{\nidefinitionautorefname}{Def.}%
+% \renewcommand{\nilemmaautorefname}{Lem.}%
+% \renewcommand{\exampleautorefname}{Ex.}%
 \addto\extrasUKenglish{%
@@
   \renewcommand{\subsectionautorefname}{\S}%
-  \renewcommand{\figureautorefname}{Fig.}%
-  \renewcommand{\appendixautorefname}{App.}%
+  % \renewcommand{\figureautorefname}{Fig.}%
+  % \renewcommand{\appendixautorefname}{App.}%
 }
@@
 \newcommand{\Algebra}{\type{ALGEBRA}}
-\newcommand{\eq}[1]{\; \type{=}_{#1} \;}
-\newcommand{\eqeq}{\; \type{\equiv} \;}
+\newcommand{\eq}[1]{\; \type{\simeq}_{#1} \;}
+\newcommand{\eqeq}{\; \type{\cong} \;}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title{$\pi$ with leftovers: \\ a mechanisation in Agda}
-\titlerunning{$\pi$ with leftovers: all mechanisation in Agda}
+\titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
 \author{Uma Zalakain}{University of Glasgow, Scotland}
@@
   \inferrule
-  {uQ : \Unused_0 \; Q}
+  {uQ : \Unused_{\constr{0} \; Q}}
   {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{\func{lower}_{\constr{0}} \; \; Q \; uQ}}
@@
   \inferrule
-  {eq : P \eq{r} P' \\ red : P' \reduce{c} Q}
+  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
   {\constr{struct} \; eq \; red : P \reduce{c} Q}
@@
     &\op{\_}{\_}{\_}         &:{} &                 &        & C \to C \to C \to \Set \\
-    &\func{\cdot-compute}  &:{} &\forall y z       & \to \; & \type{DEC} \; (\type{\exists} x  \; (\op{x}{y}{z})) \\
+    &\func{\cdot-compute^r}  &:{} &\forall x y       & \to \; & \type{DEC} \; (\type{\exists} z  \; (\op{x}{y}{z})) \\
     &\func{\cdot-unique}     &:{} &\forall x x' y z  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\

### 2020-06-22 paper/paper.tex
Add work by Ciccone and Padovani

@@
 %
+In their recent paper (unpublished at the time of writing), Ciccone and Padovani mechanise a dependently-typed linear \picalc{} in Agda \cite{Ciccone}.
+They do so by defining an intrinsicly-typed calculus, which allows them to leverage Agda to provide a dependently-typed interpretation of messages.
+Contexts contain not only types, but also the messages that inhabit their corresponding interpretation in Agda.
+To avoid linearity violations, channel types as interpreted as unit types --- and therefore only present in the context in the object language.
+This approach allows them to encode message predicates, branching, and variable-length conversations, as message input is modelled as dependent function abstraction in Agda.
+They use the concrete set of multiplicities $0$, $1$, $\omega$, define a partial algebra on contexts and require extrinsic proofs of context split.
+%
 Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged functional language with graded modal types, linear types, indexed types and polymorphism.
@@
 Gay encodes the \picalc{} with linear and shared types using de Bruijn indices, a reduction relation and a type system posterior to the syntax.
-However, in his work context splits are extrinsic, and the types of used multiplicities erased.
+However, in his work context splits are extrinsic, and used multiplicities erased from context.
 We make context splits intrinsic, preserve the ability to talk about used resources, and adopt a more general usage algebra: our framework integrates three different type systems into one.

### 2020-06-30 paper/paper.tex
Apply ACM SIGPLAN style to the paper

@@
-\documentclass[a4paper,UKenglish,cleveref,autoref,thm-restate,authorcolumns]{lipics-v2019}
-
-\usepackage[dvipsnames]{xcolor}
+\documentclass[sigplan,10pt,anonymous,review]{acmart}
+\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
 \usepackage[utf8]{inputenc}
 \usepackage{url}
-\usepackage{amssymb}
-\usepackage[fleqn]{amsmath}
-\usepackage{amsbsy}
+\usepackage{amsmath}
 \usepackage{mathbbol}
 \usepackage{mathtools}
-\usepackage[many]{tcolorbox}
-\bibliographystyle{plainurl}
-
 \usepackage{hyperref}
@@
 \usepackage{bbm}
-
-% Less margins around figures
-\usepackage{changepage}
 
@@
 \newtheorem{nilemma}{Lemma}
-% \renewcommand{\nitheoremautorefname}{Thm.}%
-% \renewcommand{\nidefinitionautorefname}{Def.}%
-% \renewcommand{\nilemmaautorefname}{Lem.}%
-% \renewcommand{\exampleautorefname}{Ex.}%
-\addto\extrasUKenglish{%
 \renewcommand{\sectionautorefname}{\S}%
 \renewcommand{\subsectionautorefname}{\S}%
-  % \renewcommand{\figureautorefname}{Fig.}%
-  % \renewcommand{\appendixautorefname}{App.}%
-}
-
-% No line numbers
-\nolinenumbers
 
@@
 
-\newcommand{\type}[1]{\textcolor{BlueViolet}{\operatorname{#1}}}
-\newcommand{\constr}[1]{\textcolor{BurntOrange}{\operatorname{#1}}}
-\newcommand{\func}[1]{\textcolor{OliveGreen}{\operatorname{#1}}}
+\newcommand{\type}[1]{\textcolor{blue}{\operatorname{#1}}}
+\newcommand{\constr}[1]{\textcolor{orange}{\operatorname{#1}}}
+\newcommand{\func}[1]{\textcolor{olive}{\operatorname{#1}}}
 
@@
 \newcommand{\PO}{\constr{\mathbb{0}}}
-\newcommand{\comp}[2]{#1 \; \constr{\parallel} \; #2}
-\newcommand{\new}{\constr{\boldsymbol{\nu}} \;}
-\newcommand{\send}[2]{#1 \; \constr{\langle} \; #2 \;\constr{\rangle} \;}
-\newcommand{\sendp}[2]{#1 \; \constr{\langle} \; #2 \; \constr{\rangle} \; . \;}
-\newcommand{\recv}[1]{#1 \; \constr{\mathbb{()}} \;}
-\newcommand{\recvp}[2]{#1 \; \constr{(} \; #2 \; \constr{)} \; . \; }
+\newcommand{\comp}[2]{#1 \, \constr{\parallel} \, #2}
+\newcommand{\new}{\constr{\boldsymbol{\nu}} \,}
+\newcommand{\send}[2]{#1 \, \constr{\langle} \, #2 \,\constr{\rangle} \,}
+\newcommand{\sendp}[2]{#1 \, \constr{\langle} \, #2 \, \constr{\rangle} \, . \,}
+\newcommand{\recv}[1]{#1 \, \constr{\mathbb{()}} \,}
+\newcommand{\recvp}[2]{#1 \, \constr{(} \, #2 \, \constr{)} \, . \, }
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
 \newcommand{\unit}{\constr{\mathbbm{1}}}
-\newcommand{\base}[1]{\constr{B[} \; #1 \; \constr{]}}
-\newcommand{\channel}[2]{\constr{C[} \; #1 \; \constr{;} \; #2 \; \constr{]}}
-\newcommand{\comma}{\; \constr{,} \;}
+\newcommand{\base}[1]{\constr{B[} \, #1 \, \constr{]}}
+\newcommand{\channel}[2]{\constr{C[} \, #1 \, \constr{,} \, #2 \, \constr{]}}
+\newcommand{\comma}{\, \constr{,} \,}
 
 % Functions
-\newcommand{\subst}[3]{#1 \; \func{[} \; #3 \; \func{\mapsto} \;#2 \; \func{]}}
-\newcommand{\op}[3]{#1 \; \func{\coloneqq} \; #2 \; \func{\cdot} \; #3}
+\newcommand{\subst}[3]{#1 \, \func{[} \, #3 \, \func{\mapsto} \,#2 \, \func{]}}
+\newcommand{\op}[3]{#1 \, \func{\coloneqq} \, #2 \, \func{\cdot} \, #3}
 \newcommand{\opsquared}[3]{#1 \, \func{\coloneqq} \, #2 \, \func{\cdot^2} \, #3}
@@
 \newcommand{\Set}{\type{SET}}
-\newcommand{\reduce}[1]{\; \type{\longrightarrow}_{#1} \;}
-\newcommand{\types}[4]{#1 \; \type{;} \; #2 \; \type{\vdash} \; #3 \; \type{\triangleright} \; #4}
-\newcommand{\contains}[6]{#1 \; \type{;} \; #2 \; \type{\ni}_{#3} \; #4 \; \type{;} \; #5 \; \type{\triangleright} \; #6}
-\newcommand{\containsusage}[4]{#1 \; \type{\ni}_{#2} \; #3 \; \type{\triangleright} \; #4}
+\newcommand{\reduce}[1]{\, \type{\longrightarrow}_{#1} \,}
+\newcommand{\types}[4]{#1 \, \type{;} \, #2 \, \type{\vdash} \, #3 \, \type{\triangleright} \, #4}
+\newcommand{\contains}[6]{#1 \, \type{;} \, #2 \, \type{\ni}_{#3} \, #4 \, \type{;} \, #5 \, \type{\triangleright} \, #6}
+\newcommand{\containsusage}[4]{#1 \, \type{\ni}_{#2} \, #3 \, \type{\triangleright} \, #4}
 \newcommand{\Var}{\type{VAR}}
@@
 \newcommand{\Ctx}{\type{CTX}}
-\newcommand{\Type}{\type{TYPE}\;}
-\newcommand{\Idx}{\type{IDX}\;}
+\newcommand{\Type}{\type{TYPE}\,}
+\newcommand{\Idx}{\type{IDX}\,}
 \newcommand{\Idxs}{\type{IDXS}}
@@
 \newcommand{\Algebra}{\type{ALGEBRA}}
-\newcommand{\eq}[1]{\; \type{\simeq}_{#1} \;}
-\newcommand{\eqeq}{\; \type{\cong} \;}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
-\titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
-\author{Uma Zalakain}{University of Glasgow, Scotland}
-       {u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
-\author{Ornela Dardha}{University of Glasgow, Scotland}
-       {ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
-\authorrunning{U. Zalakain and O. Dardha}
-\Copyright{Uma Zalakain and Ornela Dardha}
+\newcommand{\eq}[1]{\, \type{\simeq}_{#1} \,}
+\newcommand{\eqeq}{\, \type{\cong} \,}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title[$\pi$ with leftovers: a mechanisation in Agda]{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\author{Uma Zalakain}
+\affiliation{University of Glasgow, Scotland}
+\email{u.zalakain.1@research.gla.ac.uk}
+\orcid{https://orcid.org/0000-0002-3268-9338}{}
+\author{Ornela Dardha}
+\affiliation{University of Glasgow, Scotland}
+\email{ornela.dardha@glasgow.ac.uk}
+\orcid{https://orcid.org/0000-0001-9927-7875}{}
+\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
+\setcopyright{acmlicensed}
+\ccsdesc[300]{Theory of computation~Process calculi}
+% \supplement{\url{https://github.com/umazalakain/typing-linear-pi}}
 \begin{CCSXML}
@@
 \end{CCSXML}
-\ccsdesc[300]{Theory of computation~Process calculi}
-\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
-\supplement{\url{https://github.com/umazalakain/typing-linear-pi}}
-\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
-\funding{Supported by the UK EPSRC grant EP/K034413/1, ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233
-  ``Behavioural Application Program Interfaces'' (BehAPI).}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\maketitle
+\begin{acks}
+We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.
+
+The authors would also like to thank the anonymous referees fortheir valuable comments and helpful suggestions.
+This work is supported by the \grantsponsor{UKEPSRC}{UK EPSRC}{} grant \grantnum{UKEPSRC}{EP/K034413/1} ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the \grantsponsor{EUHORIZON}{EU HORIZON 2020 MSCA RISE}{} project \grantnum{EUHORIZON}{778233} ``Behavioural Application Program Interfaces'' (BehAPI).
+\end{acks}
 
@@
 
+\maketitle
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{syntax}
-\begin{figure}[h]
-  \begin{subfigure}{1\textwidth}
+
+\begin{figure}[h]
     \begin{mathpar}
@@
     \end{mathpar}
-  \end{subfigure}
-  \begin{subfigure}[t]{.3\textwidth}
+\end{figure}
+
+\begin{figure}[h]
     \begin{equation*}
@@
     \label{fig:syntax-names}
-  \end{subfigure}
-  \hspace{.05\textwidth}
-  \begin{subfigure}[t]{.65\textwidth}
+\end{figure}
+
+\begin{figure}[h]
     \begin{equation*}
-      \begin{split}
+      \begin{aligned}
         \Process_n ::=& \; \PO_n               &&\text{(inaction)}   \\ 
@@
         |& \; \send{\Var_n}{\Var_n}\Process_n  &&\text{(output)}                  
-      \end{split}
+      \end{aligned}
     \end{equation*}
@@
     \label{fig:syntax}
-  \end{subfigure}
-  \caption{}
 \end{figure}
@@
 \begin{example}
+  Example process using names (top) and de Bruijn indices (bottom).
+  \small
   \begin{equation*}
-    \begin{split}
-      & (\new{}x ) && (\comp {\recvp{x}{y} && \sendp{y}{a} && \PO} {(\new{} y) && (\sendp{x}{y} && \recvp{y}{z} && \PO)})
-      && \text{names}
-      \\
-      & \new{} && (\comp {\recv{0} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0} && \PO)})
-      && \text{de Bruijn indices}
-      \\
-    \end{split}
+    \begin{aligned}
+      & (\new{}x ) && (\comp {\recvp{x}{y} && \sendp{y}{a} && \PO} {(\new{} y) && (\sendp{x}{y} && \recvp{y}{z} && \PO)}) \\
+      & \new{} && (\comp {\recv{0} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0} && \PO)}) \\
+    \end{aligned}
   \end{equation*}
@@
 The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
-In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\allowbreak \Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
 In all the above, $i$ is incremented every time we go under a binder.
@@
     &\func{\cdot-comm}       &:{} &\forall x y z     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-    &\func{\cdot-assoc}      &:{} &\forall x y z u v & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
+    &\func{\cdot-assoc}      &:{} &\forall x y z u v & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \\
+    &                        &    &                  &        & \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
   \end{aligned}
@@
 
-\begin{note}
+%\begin{note}
   \label{note:linearity}
@@
   This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
-\end{note}
+%\end{note}
 
@@
     \begin{equation*}
-      \begin{split}
+      \begin{aligned}
         &\Idx               &:{} &\Set \\
@@
         &\type{ALGEBRAS}    &:{} &(idx : \Idx) \to \Algebra_{\Usage_{idx}}
-      \end{split}
+      \end{aligned}
     \end{equation*}
@@
 
-\begin{note}
+% \begin{note}
   We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
   All the properties of the partial monoid are lifted element-wise.
-\end{note}
+% \end{note}
 
@@
 
-\begin{remark}
+% \begin{remark}
   The \emph{unusedness} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
-\end{remark}
+% \end{remark}
 
@@
 \begin{figure}[h]
-  \begin{subfigure}{.5\textwidth}
     \centering
@@
     \label{fig:subst}
-  \end{subfigure}
-  \hspace{\fill} % DO NOT LEAVE EMPTY LINE
-  \begin{subfigure}{.4\textwidth}
+\end{figure}
+
+\begin{figure}[h]
     \centering
@@
     \label{fig:subst-alternative}
-  \end{subfigure}
-  \caption{Substitution with and without the \emph{adjustment} arrows $\ni_i n$ and $\ni_j n$}
 \end{figure}
@@
 
-\begin{note}
+% \begin{note}
   We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-\end{note}
+% \end{note}
 
@@
 \newpage
+\bibliographystyle{plainurl}
 \bibliography{paper}

### 2020-07-10 paper/paper.tex
Introduce the de Bruijn indices to names and back subsection

@@
 \newcommand{\send}[2]{#1 \, \constr{\langle} \, #2 \,\constr{\rangle} \,}
-\newcommand{\sendp}[2]{#1 \, \constr{\langle} \, #2 \, \constr{\rangle} \, . \,}
-\newcommand{\recv}[1]{#1 \, \constr{\mathbb{()}} \,}
-\newcommand{\recvp}[2]{#1 \, \constr{(} \, #2 \, \constr{)} \, . \, }
+\newcommand{\recv}[2]{#1 \, \constr{\mathbb{(}} \, #2 \, \constr{\mathbb{)}} \,}
 \newcommand{\suc}{\constr{\scriptstyle 1+}}
@@
 \newcommand{\Process}{\type{PROCESS}}
+\newcommand{\Raw}{\type{RAW}}
+\newcommand{\Name}{\type{NAME}}
+\newcommand{\Names}{\type{NAMES}}
+\newcommand{\Fresh}{\type{FRESH}}
+\newcommand{\WellScoped}{\type{WELLSCOPED}}
 \newcommand{\Unused}{\type{UNUSED}}
@@
 Data type definitions ($\type{\N}$) use double lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
-We otherwise use constructor names ($\constr{0}$ and $\suc$) to name our typing rules.
+Constructors ($\constr{0}$ and $\suc$) are given as inference rules, although we occasionally use a BNF style definition of types for brevity.
+We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda --- where premises become argument types and conclusions return types.
 Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
@@
 \label{syntax}
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    { }
+    {\Raw : \Set}
+    \; \textsc{Raw}
+  \end{mathpar}
+
+  \begin{equation*}
+    \begin{aligned}
+      \Raw ::= &\; \PO              &&\text{(inaction)}    \\ 
+      |& \; (\new{\Name}) \; \Raw         &&\text{(restriction)} \\ 
+      |& \; \comp{\Raw}{\Raw}       &&\text{(parallel)}    \\ 
+      |& \; \recv{\Name}{\Name}\Raw &&\text{(input)}       \\ 
+      |& \; \send{\Name}{\Name}\Raw &&\text{(output)}      \\
+    \end{aligned}
+  \end{equation*}
+  \caption{Grammar with names.}
+  \label{fig:syntax-names}
+\end{figure}
 
@@
   \end{mathpar}
-\end{figure}
-
-\begin{figure}[h]
-    \begin{equation*}
-      \begin{split}
-        P,Q ::=      & \; \PO                \\ 
-                    |& \; (\new{}x) \; P       \\ 
-                    |& \; \comp{P}{Q}        \\ 
-                    |& \; \recvp{x}{y} P \\ 
-                    |& \; \sendp{x}{y} P 
-      \end{split}
-    \end{equation*}
-    \caption{Ill-scoped grammar with names.}
-    \label{fig:syntax-names}
-\end{figure}
-
-\begin{figure}[h]
+
   \begin{equation*}
@@
 
-Let $x, y,\ldots$ range over \emph{variables} (including channel names) and $P, Q,\ldots$ over processes.
 The syntax of the \picalc{} \cite{Sangio01} is given by the grammar in \autoref{fig:syntax-names}.
@@
 Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recvp{x}{y} P$ and $\sendp{x}{y} P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-
-Scope restriction $(\new{}x) \; P$ and input $\recvp{x}{y} \; P$ are \emph{binders}, they are the only constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
+Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+
+Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: inserting a new variable into a context means proving that its name differs from all other variable names in context.
-Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago --- hence binders no longer introduce names.
+Names are cumbersome to mechanise: they offer no guarantee of being well-scoped, one has to deal with alpha-conversion, and inserting new variables into context entails proving that their names differ from all other names in context.
+Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
+That is, terms at different \emph{depths} use different indices to refer to the same binding, and binders no longer introduce names.
 \begin{example}
@@
     \begin{aligned}
-      & (\new{}x ) && (\comp {\recvp{x}{y} && \sendp{y}{a} && \PO} {(\new{} y) && (\sendp{x}{y} && \recvp{y}{z} && \PO)}) \\
-      & \new{} && (\comp {\recv{0} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0} && \PO)}) \\
+      & (\new{}x ) && (\comp {\recv{x}{y} && \send{y}{a} && \PO} {(\new{} y) && (\send{x}{y} && \recv{y}{z} && \PO)}) \\
+      & \new{} && (\comp {\recv{0}{} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
     \end{aligned}
@@
 \end{example}
-That is, terms at different \emph{depths} use different indices to refer to the same binding.
 
@@
 References outside of that range are meaningless.
-It is useful to rule out these ill-scoped nonsensical terms syntactically.
+It is useful to rule out these ill-scoped terms syntactically.
 In \autoref{fig:syntax}, we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
@@
 Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
+
+\subsection{From names to de Bruijn indices and back}
+
+De Bruijn indices are easier to manipulate, but as humans we rather deal with names.
+To that end, we provide the user with named processes, while using de Bruijn indexed processes internally.
+We do so by defining translation functions in both directions and proving that they are inverses of each other up to $\alpha$-conversion.
+
+\paragraph{From names to de Bruijn indices}
+We keep the original binder names around as name hints when we translate into processes with de Bruijn names --- these names will serve as hints when we translate back.
+The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps between the first $n$ indices and their names, and substituting named references within the process with their corresponding de Bruijn index.
+We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate it.
+\begin{equation*}
+  \begin{aligned}
+    &\func{fromRaw} && : (ctx : \Names_n) \; (P : \Raw) \\
+    &               && \to \WellScoped \; ctx \; P \to \Process_n \\
+  \end{aligned}
+\end{equation*}
+
+\paragraph{From de Bruijn indices to names}
+The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps between $n$ de Bruijn indices and names, and substituting indices within the process with their corresponding names.
+As many widely-used languages do, we produce unique variable names but use the original variable names kept around throughout the de Bruijn representation as hints, to make the correspondence visually more straightforward.
+For this, we use the naming scheme $<name-hint>\textasciicircum<n>$, where $<n>$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
+\begin{equation*}
+  \begin{aligned}
+    &\func{toRaw} && : (ctx : \Names_n) \to \Process_n \to \Raw
+  \end{aligned}
+\end{equation*}
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names results in a well-scoped process.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names results in a process with no shadowed variables.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names and back results in the same process modulo variable name hints.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from names to de Bruijn indices and back results in the same process modulo $\alpha$-conversion.
+\end{nilemma}
+
+\begin{proof}
+  TODO
+\end{proof}
 
@@
 We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
-We distinguish between channels that are created inside of the process ($\constr{internal}$) and channels that are created outside ($\constr{external}\ i$, where $i$ is the index of the channel variable).
+We distinguish between channels that are created inside of the process ($\constr{internal}$) and channels that are created outside ($\constr{external}\ i$), where $i$ is the index of the channel variable).
 
@@
   {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
+  {\constr{comm} : \comp{\recv{i}{}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
 
@@
             cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.6em}}{\Theta \comma \lz}}}
-        {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{P}}{\Theta}}
+        {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
 
@@
   $\func{p}$ = $\new{}$
-       $(\comp{\recv{\constr{0}} (\recv{\constr{0}} \PO)}$
+       $(\comp{\recv{\constr{0}}{} (\recv{\constr{0}}{} \PO)}$
        ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}}) \; \PO)})$

### 2020-07-13 paper/paper.tex
First batch of mostly small changes addressing reviews

@@
 \usepackage{mathpartir}
+
+% Newline inside table
+\usepackage{makecell}
+
+% Fix autoref's example treatment 
+\newtheorem{example}{Example}
+\newcommand{\exampleautorefname}{Example}
 
@@
 \newcommand{\lo}{\func{\ell_o}}
-\newcommand{\lz}{\func{\ell_{\o}}}
+\newcommand{\lz}{\func{\ell_{\varnothing}}}
 \newcommand{\lio}{\func{\ell_{\#}}}
@@
 We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
-As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
+As humans, we aim to model, build, and predict such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
 
@@
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
-Notably, it features channel mobility: channels themselves are first order values, and can therefore be transmitted.
+Notably, it features channel mobility: channels themselves are first class values, and can therefore be transmitted.
 The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and (more recently) Go \footnote{https://golang.org}.
@@
 Ultimately, we can exploit this formalisation as a unified framework for type safe distributed modelling and programming with a variety of type systems, and build on top of it more advanced types, theories and languages.
-In particular, while linearity is needed for type safety in communication-centric programming, graded and shared types are needed to model real-world software systems.
 
@@
    \item \textbf{Syntax}: \autoref{syntax} uses type level \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it.
+     For convenience, the user can convert from and to a syntax with names.
+     These two conversion functions are shown to be inverses of one another modulo $\alpha$-conversion.
 
@@
 Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
-Implicit arguments are mentioned by type definitions but omitted by constructors.
+Implicit arguments are mentioned in type definitions but omitted by constructors.
 
 We use colours to further distinguish the different entities in this paper.
-$\type{TYPES}$ are blue violet (uppercased, with indices as subscripts), $\constr{constructors}$ are burnt orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
+$\type{TYPES}$ are blue (uppercased, with indices as subscripts), $\constr{constructors}$ are orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
 
@@
 
-\begin{proof}
-  TODO
+\begin{proof}[Sketch]
+  All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
+  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
 \end{proof}
@@
   It has an element $\one$ that is used to count inputs and outputs.
-  \autoref{fig:multiplicities} uses a ternary relation to model such an algebra as a record $\Algebra_C$ on a carrier $C$.
+  \autoref{fig:multiplicities} models such an algebra as a record $\Algebra_C$ on a carrier $C$.
 
@@
 
-Linear, graded and shared types are all defined as an instance of our usage algebra.
+\autoref{fig:algebra-instances} sketches the implementation of linear, graded and shared types as instances of our usage algebra.
 Their use in typing derivations is illustrated in \autoref{example-derivations}.
-\begin{description}
-  \item [Linear]
-    The carrier is a type with two trivial constructors $\constr{zero}$ and $\constr{one}$.
-    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited otherwise.
-    All the properties of the algebra follow trivially.
-    As an example, the multiplicity pair $\ell_i$ is represented as $\constr{one} \comma \constr{zero}$.
-
-    \item [Graded]
-    The carrier is the type of natural numbers.
-    The element $\zero$ corresponds to 0, and $\one$ to 1.
-    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x = y + z$.
-    All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
-    As an example, we use $5 \comma 3$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.
-
-    \item [Shared]
-    The carrier is a type with a single trivial constructor $\constr{\omega}$.
-    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
-    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
-    All the properties of the algebra follow trivially.
-    Shared channels have multiplicities $\constr{\omega} \comma \constr{\omega}$, which they will forever preserve.
-\end{description}
+
+\begin{figure}[h]
+\begin{tabular}{l | l | l}
+  & carrier & operation \\
+  \hline
+  \textbf{linear} & \makecell[cl]{$\constr{0} \, : \, \type{Lin}$ \\ $\constr{1} \, : \, \type{Lin}$} & \makecell[cl]{$\op{\constr{0}}{\constr{0}}{\constr{0}}$ \\ $\op{\constr{1}}{\constr{1}}{\constr{0}}$ \\ $\op{\constr{1}}{\constr{0}}{\constr{1}}$} \\
+  \hline
+  \textbf{graded} & \makecell[cl]{$\constr{0} \, : \type{Gra}$ \\ $\suc \, : \type{Gra} \to \type{Gra}$} & \makecell[cl]{$\forall \, x \, y \, z$ \\ $\to x \, \type{\equiv} \, y \, \func{+} \, z$ \\ $\to \op{x}{y}{z}$} \\
+  \hline
+  \textbf{shared} & $\constr{\omega} \, : \, \type{Sha}$ & $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ \\
+\end{tabular}
+\caption{Example instances of $\Usage$ and a ternary relation that satisfies our $\Algebra$ interface.}
+\label{fig:algebra-instances}
+\end{figure}
 
@@
 \begin{nidefinition}[Indexed set of usage algebras]
-  An \emph{indexed set of usage algebras} is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
+  An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type (\autoref{fig:indexed-multiplicities}).
   \begin{figure}[h]
@@
 \begin{proof}[Proof (Sketch)]
-  All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
+  All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
   For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
@@
 He proves type preservation and provides a decision procedure for type checking and type inference.
-In our paper, we apply leftover typing to the \picalc{} following Allais \cite{Allais2018a}, and focus on type checking.
+In this paper, we follow Allais \cite{Allais2018a} and apply leftover typing to the \picalc{}.
 We give a more general usage algebra, leading to linear, graded and shared type systems.
-Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work, we too are able to talk about fully consumed resources.
-As such, we can transmit over a channel the $\lz$ multiplicities of a fully exhausted channel.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work, we too are able to talk about fully consumed resources --- e.g., we can transmit over a channel $\lz$ multiplicities of a fully exhausted channel.
 %
@@
 
-Further work on mechanising the \picalc{} \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, focuses on  non-linear variations, whether we present a range of linear, graded and shared types.
+Further work on mechanising the \picalc{} \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, focuses on non-linear variations, whereas we present a range of linear, graded and shared types.
 

### 2020-07-13 paper/appendix.tex
Last touches
Reuse constructor names from the syntax in the typing rules

@@
 
-\begin{proof}
+\begin{proof}[Proof (Sketch)]
   By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
     \item
-      For constructor $\constr{end}$ we get $\Gamma_i \equiv \Psi_i$.
+      For constructor $\PO$ we get $\Gamma_i \equiv \Psi_i$.
       From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
@@
     \item
-      For constructor $\constr{chan}$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\suc$.
+      For constructor $\new$ we proceed inductively, wrapping arrows $\ni_i m$, $\ni_j m$, $\ni_i n$ and $\ni_j n$ with $\suc$.
 
     \item
-      For constructor $\constr{recv}$ we must split $\Delta$ to proceed inductively on the continuation.
+      For constructor $\recv{}{}$ we must split $\Delta$ to proceed inductively on the continuation.
       Observe that given the arrow from $\Gamma_i$ to $\Psi_i$ and given that $\Delta$ is $\lz$ at index $i$, there must exist some $\delta$ such that $\opsquared{m}{\delta}{n}$.
@@
     \item
-      $\constr{send}$ applies the ideas outlined for the $\constr{recv}$ constructor to both the \textsc{VarRef} doing the output, and the \textsc{VarRef} for the sent data.
+      $\send{}{}$ applies the ideas outlined for the $\recv{}{}$ constructor to both the \textsc{VarRef} doing the output, and the \textsc{VarRef} for the sent data.
 
     \item
-      For $\constr{comp}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
+      For $\comp{}{}$ we first find a $\delta$, $\Theta$, $\Delta_1$ and $\Delta_2$ such that $\containsusage{\Xi_i}{i}{\delta}{\Theta}$ and $\opctx{\Gamma}{\Delta_1}{\Theta}$ and $\opctx{\Theta}{\Delta_2}{\Psi}$.
       Given $\Delta$ is $\lz$ at index $i$, we conclude that $\Delta_1$ and $\Delta_2$ are too.

### 2020-07-13 paper/paper.tex
Reuse constructor names from the syntax in the typing rules

@@
   The \textsc{Types} typing relation for \picalc{} processes is presented in \autoref{fig:types}.
+  For convenience, we reuse the constructor names introduced for the syntax in \autoref{fig:syntax}.
 
@@
     { }
-    {\constr{end} : \types{\gamma}{\Gamma}{\PO}{\Gamma}}
+    {\PO : \types{\gamma}{\Gamma}{\PO}{\Gamma}}
 
@@
      cont : \types{\gamma \comma \channel{t}{x}}{\Gamma \comma (y \comma y) }{P}{\Delta \comma \lz}}
-    {\constr{chan} \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}
+    {\new \; t \; x \; y \; cont : \types{\gamma}{\Gamma}{\new P}{\Delta}}
 
@@
             cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.6em}}{\Theta \comma \lz}}}
-        {\constr{recv} \; chan_i \; cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
+        {\recv{chan_i}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
 
@@
             cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{4.8em}}{\Theta}}}
-        {\constr{send} \; chan_i \; loc_j \; cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
+        {\send{chan_i}{loc_j} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
 
@@
      r : \types{\gamma}{\Delta}{Q}{\Xi}}
-    {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
+    {\comp{l}{r} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
   \end{mathpar}
@@
 
-  The inaction process in rule $\constr{end}$ does not change usage annotations.
-  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ of what is being transmitted; and the multiplicity $y$ given to the channel itself.
+  The inaction process in rule $\PO$ does not change usage annotations.
+  The scope restriction in rule $\new$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ of what is being transmitted; and the multiplicity $y$ given to the channel itself.
   This multiplicity $y$ is used for both input and output, so that they are balanced.
@@
 
-  The input process in rule $\constr{recv}$ requires a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  The input process in rule $\recv{}{}$ requires a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
   Note that the index $i$ is used in the syntax of the typed process.
@@
 
-  Similarly to input, the output process in rule $\constr{send}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  Similarly to input, the output process in rule $\send{}{}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
   We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan_i$.
@@
 
-  Parallel composition in rule $\constr{comp}$ uses the leftovers of the left-hand process to type the right-hand process.
+  Parallel composition in rule $\comp{}{}$ uses the leftovers of the left-hand process to type the right-hand process.
   By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide an extrinsic proof of context split.
@@
        $(\comp{\recv{\constr{0}}{} (\recv{\constr{0}}{} \PO)}$
-       ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}}) \; \PO)})$
+       ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO)})$
 
   $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{\epsilon}}$
-  $\func{\_}$ = $\constr{chan} \; \channel{\unit}{\func{\omega}} \; \li \; \one \; (\constr{comp}$
-       $(\constr{recv} \; \constr{0} \; (\constr{recv} \; \constr{0} \; \constr{end}))$
-       $(\constr{chan} \; \unit \; \func{\omega} \; \one \; (\constr{send} \; (\suc \; \constr{0}) \; \constr{0} \; (\constr{send} \; \constr{0} \; (\suc \; \suc \; \constr{0}) \; \constr{end})))
-       )$
+  $\func{\_}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; \one$
+       $(\comp{\recv{\constr{0}}{} \; (\recv{\constr{0}}{} \; \PO)}$
+       ${\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO))}$
   \end{lstlisting}
@@
    \item
-     Constructor $\constr{end}$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
+     Constructor $\PO$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
    \item
-     Constructor $\constr{chan}$: we proceed inductively.
+     Constructor $\new$: we proceed inductively.
    \item
-     Constructors $\constr{recv}$, $\constr{send}$ and $\constr{comp}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
+     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
  \end{itemize}
@@
     \caption{
-      Diagrammatic representation of the $\constr{comp}$ case for substitution.
+      Diagrammatic representation of the $\comp{}{}$ case for substitution.
       Continuous lines represent known facts, dotted lines proof obligations.

### 2020-07-13 paper/paper.tex
Clarify decidability/computability

@@
 \begin{nidefinition}[Usage algebra]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{decidable}, \emph{deterministic}, \emph{cancellative}, \emph{associative}, and \emph{commutative}.
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative}, \emph{commutative}, and in which the leftovers can be \emph{computed}.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
@@
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
-  We use the \emph{decidability} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
+  We use the \emph{computability} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts. 

### 2020-07-17 paper/paper.tex
Typeset typing rules to avoid consuming too much space
Revise syntax

@@
 Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-
 Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
+
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: they offer no guarantee of being well-scoped, one has to deal with alpha-conversion, and inserting new variables into context entails proving that their names differ from all other names in context.
+Names are cumbersome to mechanise: they are not inherently well-scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
 Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
-That is, terms at different \emph{depths} use different indices to refer to the same binding, and binders no longer introduce names.
+That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
 \begin{example}
@@
 
-De Bruijn indices are easier to manipulate, but as humans we rather deal with names.
-To that end, we provide the user with named processes, while using de Bruijn indexed processes internally.
-We do so by defining translation functions in both directions and proving that they are inverses of each other up to $\alpha$-conversion.
+In order to demonstrate the correspondence between an encoding that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
 
 \paragraph{From names to de Bruijn indices}
-We keep the original binder names around as name hints when we translate into processes with de Bruijn names --- these names will serve as hints when we translate back.
-The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps between the first $n$ indices and their names, and substituting named references within the process with their corresponding de Bruijn index.
+We keep the original binder names around as name hints when we translate to processes with de Bruijn names --- these names will serve as hints for when we translate back.
+The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
+Named references within the process are substituted with their corresponding de Bruijn index.
 We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate it.
@@
 \paragraph{From de Bruijn indices to names}
-The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps between $n$ de Bruijn indices and names, and substituting indices within the process with their corresponding names.
-As many widely-used languages do, we produce unique variable names but use the original variable names kept around throughout the de Bruijn representation as hints, to make the correspondence visually more straightforward.
-For this, we use the naming scheme $<name-hint>\textasciicircum<n>$, where $<n>$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
+The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
+As many widely-used languages do, this translation function produces unique variable names.
+These unique variable names use the naming scheme $<name-hint>\textasciicircum<n>$, where $<n>$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
 \begin{equation*}
   \begin{aligned}
-    &\func{toRaw} && : (ctx : \Names_n) \to \Process_n \to \Raw
+    \func{toRaw} : (ctx : \Names_n) \to \Process_n \to \Raw
   \end{aligned}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names and back results in the same process modulo variable name hints.
+  Translating from de Bruijn indices to names and back results in the same process modulo internal variable name hints.
 \end{nilemma}
@@
 
-\begin{proof}[Sketch]
+\begin{proof}[Prove (Sketch)]
   All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
@@
       \gamma : \PreCtx_n \\
+      i : \Var_n \\
+      \stacked{
+        t : \Type \\\\
+        idx : \Idx \\\\
+        y : \Usage_{idx}^{\func{2}}} \\
       \stacked{
         idxs : \Idxs_n \\\\
-        \Gamma : \Ctx_{idxs}} \\
-      i : \Var_n \\
-      t : \Type \\
-      \stacked{
-        idx : \Idx \\\\
-        y : \Usage_{idx}^{\func{2}}} \\
-      \Delta : \Ctx_{idxs}}
+        \Gamma \; \Delta : \Ctx_{idxs}}}
     {\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta} : \Set}
@@
       \gamma : \PreCtx_n \\
+      P : \Process_n \\
       \stacked{
         idxs : \Idxs_n \\\\
-        \Gamma : \Ctx_{idxs}} \\
-      P : \Process_n \\
-      \Delta : \Ctx_{idxs}}
+        \Gamma \; \Delta : \Ctx_{idxs}}}
     {\types{\gamma}{\Gamma}{P}{\Delta} : \Set}

### 2020-07-17 paper/appendix.tex
Rename substitution to renaming and swapping to exchange

@@
 \begin{nilemma}
-  \label{lm:subst-unused}
+  \label{lm:renaming-unused}
   For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
@@
 \begin{nilemma}
-  \label{lm:swap-swap}
-  The function $\func{swap}_i \; P$ is its own inverse: $\func{swap}_i \; (\func{swap}_i \; P) \equiv P$.
+  \label{lm:exchange-exchange}
+  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
 \end{nilemma}
@@
       Case $\constr{scope-comm}$:
-      we use swapping (\autoref{thm:swapping}), and for the reverse direction swapping and \autoref{lm:swap-swap} to show that swapping two elements in $P$ twice leaves $P$ unchanged. \qedhere
+      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged.
+\qedhere
   \end{itemize}
@@
 
-\section{Substitution}
-\label{app:subst-generalization}
+\section{Renaming}
+\label{app:renaming-generalization}
 
-\begin{nitheorem}[Generalised substitution]
+\begin{nitheorem}[Generalised renaming]
   \label{thm:subst-generalization1}
@@
   The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
-  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
+  Renaming will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
   In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
@@
   Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
-  Then substituting $i$ with $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Then renaming $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
   Refer to \autoref{fig:subst} for a diagramatic representation.

### 2020-07-17 paper/paper.tex
Rename substitution to renaming and swapping to exchange

@@
   We provide framing, weakening and strengthening theorems, which can now be stated for the linear \picalc{} and use them to prove subject congruence.
-  We show that the type system is stable under substitution and prove subject reduction.
+  We show that the type system is stable under renaming and prove subject reduction.
 %
@@
   Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
-  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
+  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{renaming} (\autoref{thm:renaming}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
 
@@
   { }
-  {\constr{scope-comm} : \new \new P \eqeq \new \new \func{swap}_{\constr{0}} \; P}
+  {\constr{scope-comm} : \new \new P \eqeq \new \new \func{exchange}_{\constr{0}} \; P}
 \end{mathpar}
@@
 The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
-In $\constr{scope-comm}$ the function $\func{swap}_i \; P$ traverses $P$ (of type $\allowbreak \Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
+In $\constr{scope-comm}$ the function $\func{exchange}_i \; P$ traverses $P$ (of type $\allowbreak \Process_{\suc \suc n}$) and swaps variable references $i$ and $\suc i$.
 In all the above, $i$ is incremented every time we go under a binder.
@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:subst-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
+After this renaming, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:renaming-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 
-\subparagraph*{Swapping}
+\subparagraph*{Exchange}
 of two variables preserves the well-typedness of a process.
-We extend $\func{swap}_i$  introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
-\begin{nitheorem}[Swapping]
-  \label{thm:swapping}
+We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
+\begin{nitheorem}[Exchange]
+  \label{thm:exchange}
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
+  Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
 \end{nitheorem}
@@
   The proof is by induction on \textsc{Equals} $\eq{}$.
-  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:swapping}.
+  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:exchange}.
   Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
@@
 
-\subparagraph*{Substitution}
-\autoref{thm:subst-generalization} proves a generalised version of substitution, where the substitution $\subst{P}{j}{i}$ is on any variable $i$.
-\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
-
-\begin{nitheorem}[Generalised substitution]
-  \label{thm:subst-generalization}
+\subparagraph*{Renaming}
+\autoref{thm:renaming-generalization} proves a generalised version of renaming, where the renaming $\subst{P}{j}{i}$ is on any variable $i$.
+\autoref{thm:renaming} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
+
+\begin{nitheorem}[Generalised renaming]
+  \label{thm:renaming-generalization}
   Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
   Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
-  Then substituting $i$ with $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Then renaming $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
 \end{nitheorem}
@@
   By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Full details are found in \autoref{app:subst-generalization}.
+  Full details are found in \autoref{app:renaming-generalization}.
   \begin{itemize}
@@
     \caption{
-      Diagrammatic representation of the $\comp{}{}$ case for substitution.
+      Diagrammatic representation of the $\comp{}{}$ case for renaming.
       Continuous lines represent known facts, dotted lines proof obligations.
@@
       Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      This approach makes composing typing derivations before and after substitution more difficult.
+      This approach makes composing typing derivations before and after renaming more difficult.
     }
@@
 
-\begin{nitheorem}[Substitution]
-  \label{thm:substitution}
+\begin{nitheorem}[Renaming]
+  \label{thm:renaming}
   Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+  Then, we can rename the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
 \end{nitheorem}
@@
   We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
-  Then, we use these to apply \autoref{thm:subst-generalization}.
+  Then, we use these to apply \autoref{thm:renaming-generalization}.
 \end{proof}
@@
     \item
-    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
+    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), renaming (\autoref{thm:renaming}) and strengthening (\autoref{thm:strengthening}).
 

### 2020-07-17 paper/paper.tex
Restructure parts of the type system section

@@
 
-%\begin{note}
-  \label{note:linearity}
-  We will often work with pairs of input and output multiplicities.
-  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ tuple of input and output multiplicities, respectively.
-  We use $\opsquared{x}{y}{z}$ to stand for a monoidal operation on pairs of multiplicities -- where the algebraic laws are lifted element-wise.
-%
-  Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
-  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
-%\end{note}
-
 \autoref{fig:algebra-instances} sketches the implementation of linear, graded and shared types as instances of our usage algebra.
@@
 \begin{nidefinition}[\textsc{Ctx}: usage contexts]
-  A \emph{usage context} is a context $\Ctx_{idxs}$ of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
+  A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
 
   A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to an existing context $\Gamma$.
+
+  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ tuple of input and output multiplicities, respectively.
+  Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
+  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
 
@@
 \end{nidefinition}
-
-% \begin{note}
-  We lift the monoidal operation on usage annotations $\opsquared{x}{y}{z}$ to a monoidal operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have a common underlying context of indices.
-  All the properties of the partial monoid are lifted element-wise.
-% \end{note}
 
@@
   The \textsc{VarRef} typing relation for variable references is presented in \autoref{fig:variable-references}.
+
   \begin{figure}[h]
@@
   \end{figure}
+
+  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{x_l \comma x_r}{y_l \comma y_r}{z_l \comma z_r}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
 

### 2020-07-17 paper/paper.tex
Clarify that we use pairs of multiplicities for channels

@@
 In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise over this notion by defining an algebra that is satisfied by linear, graded and shared types alike.
+We generalise over this notion of multiplicities by defining an algebra for them that is satisfied by linear, graded and shared types alike.
+We then use pairs of multiplicities as usage annotations for channels.
 

### 2020-07-17 paper/paper.tex
Comment on the algebraic laws for the usage algebra

@@
 \begin{nidefinition}[Usage algebra]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative}, \emph{commutative}, and in which the leftovers can be \emph{computed}.
-  It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal}.
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational resoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
+  It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal} (so that new resources cannot arbitrarily spring into life).
   It has an element $\one$ that is used to count inputs and outputs.

### 2020-07-17 paper/paper.tex
Take care to anonsuppress references to our mechanisation

@@
 %
-  Our formalisation is fully mechanised in Agda \cite{Zalakain2020Agda}.
+  Our formalisation is fully mechanised in Agda.
 \end{abstract}
@@
   \item \textbf{Fully mechanised formalisation and proofs of the \picalc{} with leftovers}:
-  The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
+    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda and is publicly available in our GitHub repository%
+\begin{anonsuppress}
+\cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
 
@@
   All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
-  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
+  For details, refer to our mechanisation in Agda%
+\begin{anonsuppress}
+\cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
 \end{proof}
@@
   All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
-  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
+  For details, refer to our mechanisation in Agda%
+\begin{anonsuppress}
+\cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
 \end{proof}
@@
   For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:exchange}.
-  Further details can be found in \autoref{app:struct} and the full proof in \cite{Zalakain2020Agda}.
+  Further details can be found in \autoref{app:struct} and the full proof in%
+\begin{anonsuppress}
+\cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
 \end{proof}
@@
 As a result, theorems like framing, weakening and strengthening can now we stated for the linear \picalc{}.
-Our work is fully mechanised in around 1500 lines of code in Agda \cite{Zalakain2020Agda}.
+Our work is fully mechanised in around 1500 lines of code in Agda%
+\begin{anonsuppress}
+\cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
 %

### 2020-07-17 paper/paper.tex
Better explain the benefits of leftover typing

@@
 \newcommand{\exampleautorefname}{Example}
+\newcommand{\nitheoremautorefname}{Theorem}
+\newcommand{\nidefinitionautorefname}{Definition}
 
@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{output usage context} $\Ctx_{idxs}$ to the typing rules.
-
+We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
 This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
 These leftovers can then be used as an input to another typing derivation.
-Leftover typing presents three main advantages:
-\begin{itemize}
-  \item \textbf{Encapsulation}:
-    The transformations to the usage context (consequence of receiving and sending over a channel) are encapsulated into a variable reference judgement with both an input and an output usage context.
-  \item \textbf{No extensional context split}:
-    No need for an extensional context splitting proof for parallel composition.
-    % Such a proof would need to construct a witness $\opctx{\Gamma_m}{\Gamma_l}{\Gamma_r}$, where $\Gamma_l \vdash P$, $\Gamma_r \vdash Q$ and $\Gamma_m \vdash \comp{P}{Q}$.
-    % This could be obtained through automated search --- but it is better yet not to need it.
-    Leftover typing uses the leftovers of $P$ to type $Q$.
-  \item \textbf{New theorems}:
-    For the first time, theorems like \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) can be stated within the linear \picalc{} thanks to the presence of a leftover context.
-\end{itemize}
+
+Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent explicit proofs of context splitting --- e.g., the typing rule for parallel composition would look as follows:
+\begin{mathpar}
+  \inferrule
+  {\opctx{\Gamma}{\Delta}{\Xi} \\ \Delta \; \type{\vdash} \; P \\ \Xi \; \type{\vdash} \; Q}
+  {\Gamma \; \type{\vdash} \; \comp{P}{Q}}
+\end{mathpar}
+Input requires a similar context split; output requires two such context splits.
+This is rather cumbersome: for input and output, the proof of context split carries \emph{exactly} the same information as the typing rule for variables; for parallel composition it means that the typing derivations of both subprocesses will have no context information whatsoever until the context split is determined.
+
+Leftover typing solves this problem by inverting the flow of information so that \textbf{the typing derivations of the subprocesses determine the context split}.
+As a result, context information is threaded left to right through the typing derivation, and explicit context split proofs are no longer needed.
+In addition to this, it allows a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and for the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to acquire a more general form.
 
@@
 
-The typing judgement for variables $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
-The typing judgement for processes $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
+The \textbf{typing judgement for variables} $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
+The \textbf{typing judgement for processes} $\types{\gamma}{\Gamma}{P}{\Delta}$ asserts that ``process $P$ is well typed under typing context $\gamma$, usage input context $\Gamma$ and leftovers $\Delta$''.
 
@@
 
-\newpage
 \bibliographystyle{plainurl}
@@
 %%
-\newpage
 \appendix

### 2020-07-17 paper/paper.tex
Small improvements

@@
   It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, including \emph{session types} \cite{H93,THK94,HVK98}.
-  %Linearity is key for type safety in session communication.
 
@@
   We present the syntax, semantics, type system and corresponding type safety properties.
-  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates linearity constraints into process continuations.
+  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates usage constraints into process continuations and requires of no top-down context splits.
   We generalise the algebras on multiplicities, allowing the user to choose a mix of linear, graded and shared typing.
-  We provide framing, weakening and strengthening theorems, which can now be stated for the linear \picalc{} and use them to prove subject congruence.
+  Thanks to leftover typing, we provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.
   We show that the type system is stable under renaming and prove subject reduction.
@@
      For convenience, the user can convert from and to a syntax with names.
-     These two conversion functions are shown to be inverses of one another modulo $\alpha$-conversion.
+     These conversion functions are shown to be inverses of one another modulo $\alpha$-conversion.
 
@@
 
-    Benefits of typing with leftovers include: removing the need for extrinsic context splits, which are rendered unnecessary; and for the first time we can state theorems like \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the linear \picalc{}.
+    Leftover typing renders top-down context splits unnecessary, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
   \end{itemize}
+
   \item \textbf{Fully mechanised formalisation and proofs of the \picalc{} with leftovers}:
-    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda and is publicly available in our GitHub repository%
+    \autoref{type-safety} presents the type safety properties of our \picalc{} with leftovers.
+    Leftover typing for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and generalises the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{renaming} (\autoref{thm:renaming}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
+
+    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda%
 \begin{anonsuppress}
-\cite{Zalakain2020Agda}
+and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
 \end{anonsuppress}
 .
-
-  \autoref{type-safety} details the type safety properties of our \picalc{} with leftovers.
-  Leftover typing allows \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to be stated.
-  We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{renaming} (\autoref{thm:renaming}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
-
 \end{enumerate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Semantics}
+\section{Operational Semantics}
 \label{semantics}
 Thanks to our well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax.
-We start by defining structural congruence in \autoref{structural-congruence} and then the reduction relation in \autoref{operational-semantics}.
+We define structural congruence in \autoref{structural-congruence} and provide a reduction relation in \autoref{operational-semantics}.
 
@@
 
-Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent explicit proofs of context splitting --- e.g., the typing rule for parallel composition would look as follows:
+Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look as follows:
 \begin{mathpar}
@@
 Leftover typing solves this problem by inverting the flow of information so that \textbf{the typing derivations of the subprocesses determine the context split}.
-As a result, context information is threaded left to right through the typing derivation, and explicit context split proofs are no longer needed.
+As a result, usage information is threaded bottom-up through the typing derivation, and top-down context split proofs are no longer needed.
 In addition to this, it allows a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and for the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to acquire a more general form.

### 2020-07-17 paper/appendix.tex
Take care to anonsuppress references to our mechanisation
Add the definition of unused to the appendix

@@
 %% Appendix
+\section{Definitions}
+\label{app:definitions}
+
+\begin{nidefinition}
+  A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
+
+  \begin{mathpar}
+    \datatype
+    {i : \Var_n \\ P : \Process_n}
+    {\Unused_i \; P}
+    \; \textsc{Unused}
+
+    \inferrule
+    { }
+    {\Unused_i \; \PO}
+
+    \inferrule
+    {\Unused_{\suc i} \; P}
+    {\Unused_i \; \new{}P}
+
+    \inferrule
+    {\Unused_i \; P \\ \Unused_i \; Q}
+    {\Unused_i \; \comp{P}{Q}}
+
+    \inferrule
+    {i \type{\not\equiv} x \\ \Unused_{\suc i} \; P}
+    {\Unused_i \; \recv{x}{} P}
+
+    \inferrule
+    {i \type{\not\equiv} x \\ i \type{\not\equiv} y \\ \Unused_i \; P}
+    {\Unused_i \; \send{x}{y} P}
+  \end{mathpar}
+
+  The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
+\end{nidefinition}
+
 \section{Lemmas}
@@
   Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
+\begin{anonsuppress}
 Full proof in \cite{Zalakain2020Agda}.
+\end{anonsuppress}
   We proceed by induction on \textsc{StructCong} $\eqeq$:

### 2020-07-17 paper/paper.tex
Add the definition of unused to the appendix

@@
 
-    Leftover typing renders top-down context splits unnecessary, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
   \end{itemize}
 
-  \item \textbf{Fully mechanised formalisation and proofs of the \picalc{} with leftovers}:
+  \item \textbf{Fully mechanised formalisation and type safety proofs}:
     \autoref{type-safety} presents the type safety properties of our \picalc{} with leftovers.
@@
   \inferrule
-  {uQ : \Unused_{\constr{0} \; Q}}
+  {uQ : \Unused_{\constr{0}} \; Q}
   {\constr{scope-ext} : \new (\comp{P}{Q}) \eqeq \comp{(\new P)}{\func{lower}_{\constr{0}} \; \; Q \; uQ}}
@@
 
-In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$.
+In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$ (see \autoref{app:definitions}).
 The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.

### 2020-07-23 paper/paper.tex
Clarify the variable typing example

@@
 
-  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{x_l \comma x_r}{y_l \comma y_r}{z_l \comma z_r}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
+  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{(x_l \comma x_r)}{(y_l \comma y_r)}{(z_l \comma z_r)}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
 
@@
 \begin{example}[Variable reference]
-  Let us illustrate the use of variable reference relations with an example in \autoref{lst:channels}.
+  \autoref{lst:channels} types a variable reference under a linear algebra.
   \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
-  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \; \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
-  _ = $\suc \; \constr{0}$
+  _ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
+  _ = $\suc \constr{0}$
   \end{lstlisting}
   The underscore \_ introduces an anonymous declaration immediately followed by its definition.
+  We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
+  Which \textsc{VarRef} constructors to use is completely determined by the $\suc \; \constr{0}$ in the type.
   The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of a channel type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover, to be used later on.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover.
 \end{example}

### 2020-07-23 paper/paper.tex
More involved example

@@
 \newcommand{\base}[1]{\constr{B[} \, #1 \, \constr{]}}
-\newcommand{\channel}[2]{\constr{C[} \, #1 \, \constr{,} \, #2 \, \constr{]}}
+\newcommand{\channel}[2]{\constr{C[} \, #1 \, \constr{;} \, #2 \, \constr{]}}
 \newcommand{\comma}{\, \constr{,} \,}
@@
   \label{example-derivations}
-  In \autoref{lst:derivation} we illustrate a typing derivation of the process $\func{p}$ (lines 1-4) using linear and shared types.
-
-  Let $\func{\omega}$ be a shortcut for $\one \comma \one$ in an algebra of shared types.
-  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $\zero \comma \zero$.
-  To type $\func{p}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
-  The left-hand process (line 3) is typed following its syntax (line 8).
-  For the right-hand (line 4) we need the channel that is going to be sent (line 9).
+  In \autoref{lst:derivation} we illustrate a typing derivation using shared, graded and linear types.
+  We use $\func{\omega}$ as a shortcut for $\one \comma \one$ in an algebra of shared types.
+
+  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
+$\func{P}$ : $\Process_{\suc \suc \constr{0}}$
+$\func{P}$ = $\comp{(\new{} \; (\send{\suc \; \constr{0}} {\constr{0}} \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))}{\recv{\constr{0}}{\recv{\constr{0}}{}} \PO}$
+
+$\func{\gamma}$ : $\PreCtx_{\suc \suc \constr{0}}$
+$\func{\gamma}$ = $\constr{[]} \comma \unit \comma \channel{\channel{\unit}{\func{\omega}}}{\li}$
+
+$\func{\vdash P}$ : $\forall \{n\} \to \types{\func{\gamma}}{\constr{[]} \comma \func{\omega} \comma (\suc n \comma \suc n)}{\func{P}}{\constr{[]} \comma \func{\omega} \comma (n \comma n)}$
+$\func{\vdash P}$ = $\comp{(\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))} {\recv{\constr{0}}{} \; \recv{\constr{0}}{} \PO}$
+
+$\func{\vdash PP}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\new{} \comp{\func{P}}{\func{P}}}{\constr{[]} \comma \func{\omega}}$
+$\func{\vdash PP}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; (\suc \suc \constr{0}) \; (\comp{\vdash P}{\vdash P})$
+  \end{lstlisting}
+
+  Notice that the \emph{only} difference between the definition of the untyped process $\func{P}$ and its typing derivation are the extra type annotations that accompany scope restriction.
+  This extra information requires a (limited) choice.
+  To type $\func{P}$, we need a channel over which we can transmit another channel.
+  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 9).
+  The left-hand process (line 3) is typed following its syntax (line 10).
+  For the right-hand (lines 4-5) we need the channel that is going to be sent (lines 11-12).
   The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
@@
 
-  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing a process that sends part of one channel over another.]
-  $\func{p}$ : $\Process_{\suc \constr{zero}}$
-  $\func{p}$ = $\new{}$
-       $(\comp{\recv{\constr{0}}{} (\recv{\constr{0}}{} \PO)}$
-       ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO)})$
-
-  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{\epsilon}}$
-  $\func{\_}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; \one$
-       $(\comp{\recv{\constr{0}}{} \; (\recv{\constr{0}}{} \; \PO)}$
-       ${\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO))}$
-  \end{lstlisting}
 \end{example}

### 2020-07-27 paper/paper.tex
Add example of conversion from names to de Bruijn and back

@@
 That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
-\begin{example}
-  Example process using names (top) and de Bruijn indices (bottom).
-  \small
-  \begin{equation*}
-    \begin{aligned}
-      & (\new{}x ) && (\comp {\recv{x}{y} && \send{y}{a} && \PO} {(\new{} y) && (\send{x}{y} && \recv{y}{z} && \PO)}) \\
-      & \new{} && (\comp {\recv{0}{} && \send{0}{a} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
-    \end{aligned}
-  \end{equation*}
-\end{example}
 
@@
 
+\begin{example}
+  Example process using names (top) translated to de Bruijn indices (middle) and back to names (bottom).
+The result is a process that follows the Barendregt convention and is $\alpha$-equivalent to the original.
+  \small
+  \begin{equation*}
+    \begin{aligned}
+      & (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
+      & \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
+      & (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)}) \\
+    \end{aligned}
+  \end{equation*}
+\end{example}
+
 \paragraph{From names to de Bruijn indices}
-We keep the original binder names around as name hints when we translate to processes with de Bruijn names --- these names will serve as hints for when we translate back.
+When we translate into de Bruijn indices we keep the original binder names around --- they will serve as name hints for when we translate back.
 The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
 Named references within the process are substituted with their corresponding de Bruijn index.
-We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate it.
+We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such proof term.
 \begin{equation*}
@@
 The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
-As many widely-used languages do, this translation function produces unique variable names.
-These unique variable names use the naming scheme $<name-hint>\textasciicircum<n>$, where $<n>$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
+As some widely-used languages do, this translation function produces unique variable names.
+These unique variable names use the naming scheme $<name-hint>^{<n>}$, where $^{<n>}$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
 \begin{equation*}

### 2020-09-02 paper/appendix.tex
Go over the intro, rename renaming back to substitution

@@
 \begin{nilemma}
-  \label{lm:renaming-unused}
+  \label{lm:substitution-unused}
   For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
@@
 
-\section{Renaming}
-\label{app:renaming-generalization}
+\section{Substitution}
+\label{app:substitution-generalization}
 
-\begin{nitheorem}[Generalised renaming]
+\begin{nitheorem}[Generalised substitution]
   \label{thm:subst-generalization1}
@@
   The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
-  Renaming will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
+  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
   In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
@@
   Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
-  Then renaming $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Then substituting $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
   Refer to \autoref{fig:subst} for a diagramatic representation.

### 2020-09-02 paper/paper.tex
Go over the intro, rename renaming back to substitution

@@
 
-  We present the \emph{first} full mechanisation in Agda of a \picalc{} with \emph{linear}, \emph{graded} and \emph{shared} types, all under the same unified framework.
-  While linearity is key for type safety in communication-centric programming, graded and shared types are needed to model real-world software systems.
-
-  We present the syntax, semantics, type system and corresponding type safety properties.
-  For the first time in the \picalc{}, we use leftover typing \cite{Allais2018a} to encode our typing rules in a way that propagates usage constraints into process continuations and requires of no top-down context splits.
-  We generalise the algebras on multiplicities, allowing the user to choose a mix of linear, graded and shared typing.
-  Thanks to leftover typing, we provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.
-  We show that the type system is stable under renaming and prove subject reduction.
+  We present the syntax, semantics, type system and corresponding type safety properties for the \picalc{}.
+  We generalise the algebras on multiplicities, and allow the user to use a mix of \emph{linear}, \emph{graded} and \emph{shared} types.
+  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs so common in type systems for linear calculi.
+  We provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.
+  We show that the type system is stable under substitution and prove subject reduction.
 %
@@
 
-In a parallel track, J.-Y.
+J.-Y.
 Girard's development of linear logic \cite{Girard87} opened new avenues in computing science in the early '90s by introducing \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018}.
@@
 Enforcing ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Later on, linearity inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
-
-The \picalc{} also benefited from the linearity wave of the early '90s.
-Kobayashi et al. \cite{KPT96} defined the {linear} \picalc{}, a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
+Linearity later on inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
+
+Kobayashi et al. \cite{KPT96} brought linearity to the \picalc{} and provided a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
 %
-In the meantime, with the rise of \emph{session types} \cite{H93,THK94,HVK98} linearity acquired a different flavour in the \picalc{}.
-Session types are a type formalism used in communication-centric programming.
-In session types, linearity ensures that a channel is owned by exactly one communicating participant, whilst the channel itself is used multiple times as per its session type.
+In the meantime, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98} --- a formalism used in communication-centric programming.
+With session types, linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself is used multiple times, as per its session type.
 Linearity in session types is the key ingredient that guarantees type safety.
 
-Recent work has pushed the linear \picalc{} into the spotlight again \cite{KPT96} thanks to an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17}.
-This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice, too.
-The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}.
-This allows us to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
-
-Considering the relevance of the \picalc{} and of linearity in modelling concurrent and distributed systems, in this paper we focus our research on both and go beyond.
-
-\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under the same unified framework.}
+Recently, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} into the spotlight again \cite{KPT96}.
+This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice too.
+The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, and it also allows us to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
+
+\emph{In this paper we present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
 
@@
     Leftover typing for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and generalises the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
-    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{renaming} (\autoref{thm:renaming}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
+    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
 
@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After this renaming, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:renaming-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
+After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:substitution-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 
-\subparagraph*{Renaming}
-\autoref{thm:renaming-generalization} proves a generalised version of renaming, where the renaming $\subst{P}{j}{i}$ is on any variable $i$.
-\autoref{thm:renaming} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
-
-\begin{nitheorem}[Generalised renaming]
-  \label{thm:renaming-generalization}
+\subparagraph*{Substitution}
+\autoref{thm:substitution-generalization} proves a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
+\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
+
+\begin{nitheorem}[Generalised substitution]
+  \label{thm:substitution-generalization}
   Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
   Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
-  Then renaming $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
+  Then substituting $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
 \end{nitheorem}
@@
   By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Full details are found in \autoref{app:renaming-generalization}.
+  Full details are found in \autoref{app:substitution-generalization}.
   \begin{itemize}
@@
     \caption{
-      Diagrammatic representation of the $\comp{}{}$ case for renaming.
+      Diagrammatic representation of the $\comp{}{}$ case for substitution.
       Continuous lines represent known facts, dotted lines proof obligations.
@@
       Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      This approach makes composing typing derivations before and after renaming more difficult.
+      This approach makes composing typing derivations before and after subsitution more difficult.
     }
@@
 
-\begin{nitheorem}[Renaming]
-  \label{thm:renaming}
+\begin{nitheorem}[Substitution]
+  \label{thm:substitution}
   Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can rename the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
 \end{nitheorem}
@@
   We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
-  Then, we use these to apply \autoref{thm:renaming-generalization}.
+  Then, we use these to apply \autoref{thm:substitution-generalization}.
 \end{proof}
@@
     \item
-    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), renaming (\autoref{thm:renaming}) and strengthening (\autoref{thm:strengthening}).
+    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
 

### 2020-09-02 paper/paper.tex
Better organize related work

@@
 _ = $\suc \constr{0}$
+
+_ : $\contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
+_ = $\constr{0}$
   \end{lstlisting}
@@
   We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
-  Which \textsc{VarRef} constructors to use is completely determined by the $\suc \; \constr{0}$ in the type.
-  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that we can split its input annotation $\lio$, taking away $\li$ and leaving $\lo$ as the leftover.
+  The \textsc{VarRef} constructors we must use are completely determined by $\suc \; \constr{0}$ in the type.
+  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output, the result is the obligation on line 4.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to fulfill this proof obligation automatically.
 \end{example}
@@
 
-  Notice that the \emph{only} difference between the definition of the untyped process $\func{P}$ and its typing derivation are the extra type annotations that accompany scope restriction.
-  This extra information requires a (limited) choice.
+  Notice that the \emph{only} differences between the definition of the untyped process $\func{P}$ and its typing derivation are the extra type annotations that accompany scope restriction.
+  This extra information demands a (limited) choice.
   To type $\func{P}$, we need a channel over which we can transmit another channel.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Conclusion, Related and Future Work}
+\section{Related Work}
+
+\paragraph{Extrinsic Encodings}
+
+Extrinsic encodings define a syntax (often well-scoped) and a runtime semantics prior to any type system.
+This allows one to talk about ill-typed terms, and defers the proof of subject reduction to a later stage.
 
@@
 In this paper, we follow Allais \cite{Allais2018a} and apply leftover typing to the \picalc{}.
-We give a more general usage algebra, leading to linear, graded and shared type systems.
-Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work, we too are able to talk about fully consumed resources --- e.g., we can transmit over a channel $\lz$ multiplicities of a fully exhausted channel.
-%
-In their recent paper (unpublished at the time of writing), Ciccone and Padovani mechanise a dependently-typed linear \picalc{} in Agda \cite{Ciccone}.
-They do so by defining an intrinsicly-typed calculus, which allows them to leverage Agda to provide a dependently-typed interpretation of messages.
-Contexts contain not only types, but also the messages that inhabit their corresponding interpretation in Agda.
-To avoid linearity violations, channel types as interpreted as unit types --- and therefore only present in the context in the object language.
-This approach allows them to encode message predicates, branching, and variable-length conversations, as message input is modelled as dependent function abstraction in Agda.
-They use the concrete set of multiplicities $0$, $1$, $\omega$, define a partial algebra on contexts and require extrinsic proofs of context split.
-%
+We generalise the usage algebra, leading to linear, graded and shared type systems.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work we too are able to talk about fully consumed resources --- e.g., we can transmit $\lz$ multiplicities of a fully exhausted channel.
+
+Recent years have seen an increase in the efforts to mechanise resource-aware process algebras, but one of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
+Gay encodes the \picalc{} with linear and shared types using de Bruijn indices, a reduction relation and a type system posterior to the syntax.
+However, in his work typing rules demand user-provided context splits, and variables with consumed usage annotations are erased from context.
+We remove the demand for context splits, preserve the ability to talk about consumed resources, and adopt a more general usage algebra.
+
 Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged functional language with graded modal types, linear types, indexed types and polymorphism.
@@
 
-Recent years have seen an increase in the efforts to mechanise resource-aware process algebras.
-One of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
-Gay encodes the \picalc{} with linear and shared types using de Bruijn indices, a reduction relation and a type system posterior to the syntax.
-However, in his work context splits are extrinsic, and used multiplicities erased from context.
-We make context splits intrinsic, preserve the ability to talk about used resources, and adopt a more general usage algebra: our framework integrates three different type systems into one.
-%
-The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types that includes a type safety proof (in Coq). 
+The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types to come along with a mechanised proof of type safety (in Coq).
 The authors extend session types with polymorphism and pattern matching.
@@
 Their type system uses reordering of contexts and extrinsic context splits.
-%
-Castro et al.
-\cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq  --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
-As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
-%
-Taking an intrinsic approach to typing, Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
-Context splits are still given extrinsically, and exhausted resources are removed from typing contexts altogether.
-The semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
-%
-Most recently, Rouvoet et al. provide an intrinsic type system for a \lambdacalc{} with session types \cite{Rouvoet2020}.
+
+\paragraph{Intrinsic Encodings}
+
+Intrinsic encodings merge syntax and type system.
+As a result, one can only ever talk about well-typed terms, and the reduction relation by construction carries a proof of subject reduction.
+Significantly, by merging the syntax and static semantics of the object language one can fully use the expressive power of the host language.
+
+Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
+As Gay does \cite{Gay2001}, context splits are given extrinsically, and exhausted resources are removed from typing contexts altogether.
+The runtime semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
+
+In their recent paper, Ciccone and Padovani mechanise a dependently-typed linear \picalc{} in Agda \cite{Ciccone}.
+They do so by leveraging Agda's dependent types to provide a dependently-typed interpretation of messages.
+Contexts contain not only types, but also the messages that inhabit their corresponding interpretation in Agda.
+To avoid linearity violations, the interpretation of channel types is erased --- they are only present in the context of the typing judgment.
+They model message input as a dependent function abstraction in Agda, and as a result are able to encode message predicates, branching, and variable-length conversations.
+They use the concrete set of multiplicities $0$, $1$, $\omega$, define a partial algebra on contexts and require user-provided proofs of context split.
+
+In another recent work, Rouvoet et al. provide an intrinsic type system for a \lambdacalc{} with session types \cite{Rouvoet2020}.
 They use proof relevant separation logic and a notion of a supply and demand \emph{market} to make context splits transparent to the user.
@@
 They distill their typing rules even further by modelling the supply and demand market as a state monad.
-%
+
+\paragraph{Other Work}
+
+Castro et al.
+\cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq  --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
+As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
+
 Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-To conclude, in this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
+\section{Conclusion and Future Work}
+
+In this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
 We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.
@@
 .
-%
+
 As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.

### 2020-09-03 paper/paper.tex
Provide a more realistic example

@@
 
-\emph{In this paper we present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
+\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
 
@@
   \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
-_ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
-_ = $\suc \constr{0}$
-
-_ : $\contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
-_ = $\constr{0}$
+$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
+$\_ \; = \; \suc \constr{0}$
+
+$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
+$\_ \; = \; \constr{0}$
   \end{lstlisting}
@@
         {\stacked{
-            chan_i : \contains{\gamma \hspace{1.2em}}{\Gamma \hspace{1.3em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.6em}}{\Theta \comma \lz}}}
+            chan_i : \contains{\gamma \hspace{1.0em}}{\Gamma \hspace{1.0em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
+            cont \hspace{0.5em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.2em}}{\Theta \comma \lz}}}
         {\recv{chan_i}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
@@
         {\stacked{
-            chan_i : \contains{\gamma}{\Gamma \hspace{0.2em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t \hspace{3.2em}}{x \hspace{0.3em}}{\Xi} \\\\
-            cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{4.8em}}{\Theta}}}
+            chan_i : \contains{\gamma}{\Gamma \hspace{0.1em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+            loc_j \hspace{0.7em} : \contains{\gamma}{\Delta}{j}{t \hspace{2.8em}}{x \hspace{0.2em}}{\Xi} \\\\
+            cont \hspace{0.5em} : \types{\gamma}{\Xi}{\hspace{0.4em}P\hspace{4.0em}}{\Theta}}}
         {\send{chan_i}{loc_j} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
@@
     \inferrule
-    {l \hspace{0.3em} : \types{\gamma}{\Gamma \hspace{0.2em}}{P}{\Delta} \\\\
+    {l : \types{\gamma}{\Gamma}{P\hspace{0.3em}}{\Delta} \\\\
      r : \types{\gamma}{\Delta}{Q}{\Xi}}
@@
   \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
-$\func{P}$ : $\Process_{\suc \suc \constr{0}}$
-$\func{P}$ = $\comp{(\new{} \; (\send{\suc \; \constr{0}} {\constr{0}} \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))}{\recv{\constr{0}}{\recv{\constr{0}}{}} \PO}$
-
-$\func{\gamma}$ : $\PreCtx_{\suc \suc \constr{0}}$
-$\func{\gamma}$ = $\constr{[]} \comma \unit \comma \channel{\channel{\unit}{\func{\omega}}}{\li}$
-
-$\func{\vdash P}$ : $\forall \{n\} \to \types{\func{\gamma}}{\constr{[]} \comma \func{\omega} \comma (\suc n \comma \suc n)}{\func{P}}{\constr{[]} \comma \func{\omega} \comma (n \comma n)}$
-$\func{\vdash P}$ = $\comp{(\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))} {\recv{\constr{0}}{} \; \recv{\constr{0}}{} \PO}$
-
-$\func{\vdash PP}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\new{} \comp{\func{P}}{\func{P}}}{\constr{[]} \comma \func{\omega}}$
-$\func{\vdash PP}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; (\suc \suc \constr{0}) \; (\comp{\vdash P}{\vdash P})$
+  $\func{sync} \; : \; \forall \{n\} \to \Var_n \to \Var_n \to \Var_n \to \Process_n$
+  $\func{sync} \; i0 \; i1 \; o \; =$
+    $\recv{i0}{} \recv{(\suc \; i1)}{} \send{(\suc \suc \; o)}{\suc \constr{0}} \send{(\suc \suc \; o)}{\constr{0}} \PO$
+
+  $\func{send} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
+  $\func{send} \; c \; = \new{} (\send{\suc c}{\constr{0}} \PO)$
+
+  $\func{recv} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
+  $\func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO)$
+
+  $\func{example} \; : \; \Process_{\constr{0}}$
+  $\func{example} \; = \new{} ( \func{send} \; \constr{0}$
+           $\comp{}{ \new{} ( \func{send} \; \constr{0} }$
+             $\comp{}{ \new{} ( \func{recv} \; \constr{0} }$
+               $\comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) }$
+
+  $\func{\_} \; : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}}$
+  $\func{\_} = \new{} \; \channel{\unit}{\omega} \; \lz \; 1$
+    $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
+    $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 1 }$
+      $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
+      $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 2 }$
+        $( (\recv{\constr{0}}{} (\recv{\suc \constr{0}}{} \PO))$
+        $\comp{}{( ( \recv{\suc (\suc \constr{0}))}{}}$
+          $( \recv{\suc (\suc \constr{0}))}{}$
+          $( \send{\suc (\suc \constr{0}))}{\suc \constr{0}}$
+          $( \send{\suc (\suc \constr{0}))}{\constr{0}} \PO))))$
   \end{lstlisting}

### 2020-09-04 paper/paper.tex
First go at an improved example

@@
 \label{example-derivations}
-  In \autoref{lst:derivation} we illustrate a typing derivation using shared, graded and linear types.
+
+We present an example with four parallel processes: two processes that send data over a channel; one that receives data over a channel twice; and a fourth process that sticks the previous processes together and synchronizes them.
+
+We start by defining the process doing the synchronization.
+This process sequentially receives data on both input channels and then sends on the output channel both pieces of received data.
+For simplicity, when we type this process we instantiate it to variable references and usage annotations.
+We expect both input channels to be graded and capable of performing input once.
+The output channel must be capable of performing output twice.
+Notice that the well-typedness proof has no choice but to blindly follow the syntax.
+
+\begin{flalign*}
+& \func{sync} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
+& \func{sync} \; in0 \; in1 \; out \; &&= \\
+& && \recv{in0}{} \\
+& && \recv{(\suc \; in1)}{} \\
+& && \send{(\suc \suc \; out)}{\suc \constr{0}} \\
+& && \send{(\suc \suc \; out)}{\constr{0}} \PO \\
+\end{flalign*}
+\begin{flalign*}
+& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (1 \comma 0) \comma (1 \comma 0) \comma (0 \comma 2)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (0 \comma 0) \comma (0 \comma 0) \comma (0 \comma 0)} & \\
+& \func{\vdash sync} && = \\
+& && \recv{\suc \suc \constr{0}}{} \\
+& && \recv{\suc \suc \constr{0}}{} \\
+& && \send{\suc \suc \constr{0}}{\suc \constr{0}} \\
+& && \send{\suc \suc \constr{0}}{\constr{0}} \PO \\
+\end{flalign*}
+
+Next, we define the processes doing the input and output.
+This time too, we instantiate them to concrete variable references for typing.
+
+\begin{flalign*}
+& \func{recv} \; : \; \Var_n \to \Process_n & \\
+& \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO) \\
+& \\
+& \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (2 , 2)}{\func{recv} \; \constr{0}}{\Gamma \comma (0 \comma 2)} & \\
+& \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
+& \\
+& \func{send} \; : \; \Var_n \to \Var_n \to \Process_n & \\
+& \func{send} \; c \; v \; = \send{c}{v} \PO \\
+& \\
+& \func{\vdash send} \; \types{\gamma \comma \channel{t}{\lz} \comma t}{\Gamma \comma (1 , 1) \comma \lz}{\func{send} \; (\suc \constr{0}) \; \constr{0}}{\Gamma \comma (1 \comma 0)\comma \lz} & \\
+& \func{\vdash send} \; = \; \send{(\suc \constr{0})}{\constr{0}} \PO
+\end{flalign*}
+
+Finally, we put the four processes together in parallel.
+It is here that the well-typedness proof must make a choice as to what types and usage annotations to instantiate new channels with.
+In this case, these types and usage annotations were determined by the types of the previous subproofs, and we can thus omit most of them with $\_$ and have Agda find them.
 We use $\func{\omega}$ as a shortcut for $\one \comma \one$ in an algebra of shared types.
 
-  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
-  $\func{sync} \; : \; \forall \{n\} \to \Var_n \to \Var_n \to \Var_n \to \Process_n$
-  $\func{sync} \; i0 \; i1 \; o \; =$
-    $\recv{i0}{} \recv{(\suc \; i1)}{} \send{(\suc \suc \; o)}{\suc \constr{0}} \send{(\suc \suc \; o)}{\constr{0}} \PO$
-
-  $\func{send} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
-  $\func{send} \; c \; = \new{} (\send{\suc c}{\constr{0}} \PO)$
-
-  $\func{recv} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
-  $\func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO)$
-
-  $\func{example} \; : \; \Process_{\constr{0}}$
-  $\func{example} \; = \new{} ( \func{send} \; \constr{0}$
-           $\comp{}{ \new{} ( \func{send} \; \constr{0} }$
-             $\comp{}{ \new{} ( \func{recv} \; \constr{0} }$
-               $\comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) }$
-
-  $\func{\_} \; : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}}$
-  $\func{\_} = \new{} \; \channel{\unit}{\omega} \; \lz \; 1$
-    $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
-    $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 1 }$
-      $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
-      $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 2 }$
-        $( (\recv{\constr{0}}{} (\recv{\suc \constr{0}}{} \PO))$
-        $\comp{}{( ( \recv{\suc (\suc \constr{0}))}{}}$
-          $( \recv{\suc (\suc \constr{0}))}{}$
-          $( \send{\suc (\suc \constr{0}))}{\suc \constr{0}}$
-          $( \send{\suc (\suc \constr{0}))}{\constr{0}} \PO))))$
-  \end{lstlisting}
-
-  Notice that the \emph{only} differences between the definition of the untyped process $\func{P}$ and its typing derivation are the extra type annotations that accompany scope restriction.
-  This extra information demands a (limited) choice.
-  To type $\func{P}$, we need a channel over which we can transmit another channel.
-  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 9).
-  The left-hand process (line 3) is typed following its syntax (line 10).
-  For the right-hand (lines 4-5) we need the channel that is going to be sent (lines 11-12).
-  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
-  Its usage annotation is $\one \comma \one$: we will transmit $\one \comma \zero$ and save $\zero \comma \one$ for the continuation process.
+\begin{flalign*}
+& \func{example} && : \; \Process_{\constr{0}} & \\
+& \func{example} && = \new{} ( ( \new{} (\func{send} \; \suc \constr{0} \; \constr{0})) \\
+& && \comp{}{ \new{} ( (\new{} (\func{send} \; \suc \constr{0} \; \constr{0} })) \\
+& && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
+& && \comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
+\end{flalign*}
+\begin{flalign*}
+& \func{\vdash example} && : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}} & \\
+& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
+& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
+& && \comp{}{ \new{} \; \channel{\unit}{\func{\omega}} \; \_ \; \_ } \; ( \func{\vdash recv} \\
+& && \comp{}{ \func{\vdash sync} }))) \\
+\end{flalign*}
 

### 2020-09-07 paper/paper.tex
Move the example process to the introduction

@@
 
+\begin{example}[Example process]
+\label{example-process}
+
+We present an example that highlights the ways in which graded types can be used to prevent mistakes in the definition of \picalc{} processes.
+The syntax uses Bruijn indices as variable references and is introduced in detail in \autoref{syntax}.
+In short: $\new{} P$ creates a new channel and makes it available at index $0$ in $P$; $\comp{P}{Q}$ composes $P$ and $Q$ in parallel; $\recv{x}{}P$ receives along channel $x$ and makes the data available at index $0$ in $P$; and $\send{x}{y}{P}$ sends $y$ over channel $x$ and continues with $P$.
+Both variable references and processes are indexed by the number of free variables they can refer to (for any natural number $n$, $\Var_n$ and $\Process_n$, respectively).
+For the sake of clarity, we define the example subprocesses separately and parametrise them by the channels on which they operate.
+
+\begin{figure}[h]
+  \tiny
+  \centering
+  \begin{tikzpicture}[draw,circle]
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=orange] (send-x) at (0,2)   {$\func{send} \, x$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=orange] (send-y) at (2,2)   {$\func{send} \, y$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=green] (recv-z) at (4,2)   {$\func{recv} \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown,align=center] (sync) at (2,0)   {$\func{sync}$\\$ \, x \, y \, z$};
+
+    \draw[->] (send-x) -- node[left] {$x$} (sync);
+    \draw[->] (send-y) -- node[right] {$y$} (sync);
+    \draw[->] (sync.35) -- node[right] {$z$} (recv-z.240);
+    \draw[->] (sync) -- (recv-z);
+  \end{tikzpicture}
+  \caption{Four communicating processes composed in parallel.}
+  \label{fig:example-process}
+\end{figure}
+
+The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous processes (\autoref{fig:example-process}).
+The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
+Thanks for linear and graded types, the processes below have their interactions safely restricted: the usage annotation $(0 \comma 1)$ on channel $c$ (no input, output once) restricts what both $\func{send}$ subprocesses can do; and $(2 \comma 0)$ (input twice, no output) restricts what $\func{send}$ can do.
+Dually, the remaining annotations $(1 \comma 0)$, $(1 \comma 0)$ and $(0 \comma 2)$ are given to the subprocess $\func{sync}$.
+We use these types to give a suitable typing derivation in \autoref{example-derivation}.
+
+\begin{flalign*}
+& \func{recv} \; : \; \Var_n \to \Process_n & \\
+& \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO) \\
+& \\
+& \func{send} \; : \; \Var_n \to \Process_n & \\
+& \func{send} \; c = \new{} \send{\suc c}{\constr{0}} \PO \\
+\end{flalign*}
+\begin{flalign*}
+& \func{sync} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
+& \func{sync} \; in0 \; in1 \; out \; &&= \\
+& && \recv{in0}{} \\
+& && \recv{(\suc \; in1)}{} \\
+& && \send{(\suc \suc \; out)}{\suc \constr{0}} \\
+& && \send{(\suc \suc \; out)}{\constr{0}} \PO \\
+\end{flalign*}
+\begin{flalign*}
+& \func{example} && : \; \Process_{\constr{0}} & \\
+& \func{example} && = \new{} ( \func{send} \; \constr{0} \\
+& && \comp{}{ \new{} ( \func{send} \; \constr{0} } \\
+& && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
+& && \comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
+\end{flalign*}
+
+\end{example}
+
 \subsection{Notation}
@@
 \autoref{fig:algebra-instances} sketches the implementation of linear, graded and shared types as instances of our usage algebra.
-Their use in typing derivations is illustrated in \autoref{example-derivations}.
+Their use in typing derivations is illustrated in \autoref{example-derivation}.
 
@@
 \begin{example}[Typing derivation]
-\label{example-derivations}
+\label{example-derivation}
 
@@
 
-\begin{flalign*}
-& \func{sync} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
-& \func{sync} \; in0 \; in1 \; out \; &&= \\
-& && \recv{in0}{} \\
-& && \recv{(\suc \; in1)}{} \\
-& && \send{(\suc \suc \; out)}{\suc \constr{0}} \\
-& && \send{(\suc \suc \; out)}{\constr{0}} \PO \\
-\end{flalign*}
 \begin{flalign*}
@@
 \begin{flalign*}
-& \func{recv} \; : \; \Var_n \to \Process_n & \\
-& \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO) \\
-& \\
 & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (2 , 2)}{\func{recv} \; \constr{0}}{\Gamma \comma (0 \comma 2)} & \\
 & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
-& \\
-& \func{send} \; : \; \Var_n \to \Var_n \to \Process_n & \\
-& \func{send} \; c \; v \; = \send{c}{v} \PO \\
 & \\
@@
 
-\begin{flalign*}
-& \func{example} && : \; \Process_{\constr{0}} & \\
-& \func{example} && = \new{} ( ( \new{} (\func{send} \; \suc \constr{0} \; \constr{0})) \\
-& && \comp{}{ \new{} ( (\new{} (\func{send} \; \suc \constr{0} \; \constr{0} })) \\
-& && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
-& && \comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
-\end{flalign*}
 \begin{flalign*}

### 2020-09-07 paper/paper.tex
More general type for the example derivation

@@
 The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
-Thanks for linear and graded types, the processes below have their interactions safely restricted: the usage annotation $(0 \comma 1)$ on channel $c$ (no input, output once) restricts what both $\func{send}$ subprocesses can do; and $(2 \comma 0)$ (input twice, no output) restricts what $\func{send}$ can do.
-Dually, the remaining annotations $(1 \comma 0)$, $(1 \comma 0)$ and $(0 \comma 2)$ are given to the subprocess $\func{sync}$.
+Thanks for linear and graded types, the processes below have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
+Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
 We use these types to give a suitable typing derivation in \autoref{example-derivation}.
@@
 
-We present an example with four parallel processes: two processes that send data over a channel; one that receives data over a channel twice; and a fourth process that sticks the previous processes together and synchronizes them.
-
-We start by defining the process doing the synchronization.
-This process sequentially receives data on both input channels and then sends on the output channel both pieces of received data.
-For simplicity, when we type this process we instantiate it to variable references and usage annotations.
+We provide an example typing derivation for the processes defined in \autoref{example-process}.
+For the sake of simplicity, we instantiate these processes with concrete variable references before typing them.
+
+\begin{flalign*}
+  & \func{\vdash send} \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}}{\Gamma \comma (l \comma \suc r)}{\func{send} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO \\
+  & \\
+  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
+\end{flalign*}
+
 We expect both input channels to be graded and capable of performing input once.
@@
 \begin{flalign*}
-& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (1 \comma 0) \comma (1 \comma 0) \comma (0 \comma 2)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (0 \comma 0) \comma (0 \comma 0) \comma (0 \comma 0)} & \\
+& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
 & \func{\vdash sync} && = \\
@@
 
-Next, we define the processes doing the input and output.
-This time too, we instantiate them to concrete variable references for typing.
-
-\begin{flalign*}
-& \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (2 , 2)}{\func{recv} \; \constr{0}}{\Gamma \comma (0 \comma 2)} & \\
-& \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
-& \\
-& \func{\vdash send} \; \types{\gamma \comma \channel{t}{\lz} \comma t}{\Gamma \comma (1 , 1) \comma \lz}{\func{send} \; (\suc \constr{0}) \; \constr{0}}{\Gamma \comma (1 \comma 0)\comma \lz} & \\
-& \func{\vdash send} \; = \; \send{(\suc \constr{0})}{\constr{0}} \PO
-\end{flalign*}
-
 Finally, we put the four processes together in parallel.
@@
 & \func{\vdash example} && : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}} & \\
-& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
-& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; (( \new{} \; \_ \; \_ \; \_ \; \func{\vdash send}) \\
-& && \comp{}{ \new{} \; \channel{\unit}{\func{\omega}} \; \_ \; \_ } \; ( \func{\vdash recv} \\
+& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; ( \func{\vdash send} \\
+& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash send} \\
+& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash recv} \\
 & && \comp{}{ \func{\vdash sync} }))) \\

### 2020-09-08 paper/paper.tex
Update the explanations on the typing derivation example

@@
 We present an example that highlights the ways in which graded types can be used to prevent mistakes in the definition of \picalc{} processes.
-The syntax uses Bruijn indices as variable references and is introduced in detail in \autoref{syntax}.
-In short: $\new{} P$ creates a new channel and makes it available at index $0$ in $P$; $\comp{P}{Q}$ composes $P$ and $Q$ in parallel; $\recv{x}{}P$ receives along channel $x$ and makes the data available at index $0$ in $P$; and $\send{x}{y}{P}$ sends $y$ over channel $x$ and continues with $P$.
+The syntax uses de Bruijn indices for variables and is introduced in detail in \autoref{syntax}.
+In short: $\new{} P$ creates a new channel and makes it available at index $0$ in the continuation $P$; $\comp{P}{Q}$ composes $P$ and $Q$ in parallel; $\recv{x}{}P$ receives along channel $x$ and makes the data available at index $0$ in the continuation $P$; and $\send{x}{y}{P}$ sends $y$ over channel $x$ and continues with $P$.
 Both variable references and processes are indexed by the number of free variables they can refer to (for any natural number $n$, $\Var_n$ and $\Process_n$, respectively).
@@
 
-The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous processes (\autoref{fig:example-process}).
+The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous three processes (\autoref{fig:example-process}).
 The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
-Thanks for linear and graded types, the processes below have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
+Thanks to linear and graded types, the processes below can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
 Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
@@
 
+The $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$.
+After receiving twice, the process ends, so we should not be left with unused multiplicities: $u$ needs to be $\lz$.
+We will use graded types to keep track of the exact number of times communication happens.
+Whatever the input multiplicity of the channel, we will consume $2$ of it and leave the remaining as leftovers.
+Once the types are given, the typing derivation is completely syntax directed.
+
+The $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$.
+We instantiate the type of transmitted data $t$ to $\channel{\unit}{\func{\omega}}$, as that is something we can construct.
+As by the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
+We will transmit once, thus using a single output multiplicity, and leave the rest as leftovers.
+
 \begin{flalign*}
+  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
+  & \\
   & \func{\vdash send} \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}}{\Gamma \comma (l \comma \suc r)}{\func{send} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
   & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO \\
-  & \\
-  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
 \end{flalign*}
 
-We expect both input channels to be graded and capable of performing input once.
-The output channel must be capable of performing output twice.
-Notice that the well-typedness proof has no choice but to blindly follow the syntax.
+Dually, the process doing the synchronisation expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
+The typing derivation here is uniquely determined by the type and syntax of the process.
 
@@
 
-Finally, we put the four processes together in parallel.
-It is here that the well-typedness proof must make a choice as to what types and usage annotations to instantiate new channels with.
-In this case, these types and usage annotations were determined by the types of the previous subproofs, and we can thus omit most of them with $\_$ and have Agda find them.
-We use $\func{\omega}$ as a shortcut for $\one \comma \one$ in an algebra of shared types.
+We can now put these processes together and omit specifying the types of the channels that they share: Agda can infer them from the types of the processes.
 

### 2020-09-08 paper/appendix.tex
Move the equivalent congruence relation to the body

@@
 \label{app:struct}
-
-Structural congruence is a congruent equivalence relation.
-As such, rewrites can happen anywhere inside a process, and are closed under reflexivity, symmetry and transitivity as shown by the first row of \autoref{fig:struct-cong1}.
-The rest of the rules defines structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
-
-\begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    { }
-    {\Rec : \Set}
-    \; \textsc{Rec}
-
-    \inferrule
-    { }
-    {\constr{zero} : \Rec}
-
-    \inferrule
-    {r : \Rec}
-    {\constr{one} \; r : \Rec}
-
-    \inferrule
-    {r \; s : \Rec}
-    {\constr{two} \; r \; s : \Rec}
-
-    \datatype
-    {P \, Q : \Process_n \\ r : \Rec}
-    {P \eq{r} Q : \Set}
-    \; \textsc{Equals}
-
-    \inferrule
-    {eq : P \eqeq Q}
-    {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
-
-    \inferrule
-    {eq : P \eq{r} P'}
-    {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}
-
-    \inferrule
-    {eq : P \eq{r} P'}
-    {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}
-
-    \inferrule
-    {eq : P \eq{r} P'}
-    {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}
-
-    \inferrule
-    {eq : P \eq{r} P'}
-    {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
-
-    \inferrule
-    { }
-    {\constr{refl} : P \eq{\constr{zero}} P}
-
-    \inferrule
-    {eq : P \eq{r} Q}
-    {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
-
-    \inferrule
-    {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
-    {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
-  \end{mathpar}
-  \caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
-  \label{fig:struct-cong1}
-  \end{figure}
-
-In the transitivity rule, we must show that if $P$ is structurally congruent to $Q$ and $Q$ is to $R$, and $P$ is well-typed, then so is $R$.
-To do so, we need to proceed by induction and first get a proof of the well-typedness of $Q$, then use that to reach $R$.
-To show the typechecker that the doubly recursive call terminates we index the equivalence relation $=$ by a type $\Rec$ that models the structure of the recursion.
 

### 2020-09-08 paper/paper.tex
Move the equivalent congruence relation to the body

@@
 We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ in \autoref{fig:struct-cong-base}.
-Its congruent equivalence closure \textsc{Equals} $\eq{}$ is defined in \autoref{app:struct}.
 
@@
 In all the above, $i$ is incremented every time we go under a binder.
+
+We lift the relation \textsc{StrucCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$ as shown in \autoref{fig:struct-cong1}.
+The $\func{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
+The remaining three rules close structural congruence under reflexivity, symmetry and transitivity. 
+The transitivity rule makes it tricky to prove predicates by induction on \textsc{Equals}: Agda cannot immediately see that such doubly recursive calls terminate.
+In order to remedy this, we model the inductive structure syntactically and index \textsc{Equals} by a type \textsc{Rec}.
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    { }
+    {\Rec : \Set}
+    \; \textsc{Rec}
+
+    \inferrule
+    { }
+    {\constr{zero} : \Rec}
+
+    \inferrule
+    {r : \Rec}
+    {\constr{one} \; r : \Rec}
+
+    \inferrule
+    {r \; s : \Rec}
+    {\constr{two} \; r \; s : \Rec}
+
+    \datatype
+    {P \, Q : \Process_n \\ r : \Rec}
+    {P \eq{r} Q : \Set}
+    \; \textsc{Equals}
+
+    \inferrule
+    {eq : P \eqeq Q}
+    {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
+
+    \inferrule
+    {eq : P \eq{r} P'}
+    {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}
+
+    \inferrule
+    {eq : P \eq{r} P'}
+    {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}
+
+    \inferrule
+    {eq : P \eq{r} P'}
+    {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}
+
+    \inferrule
+    {eq : P \eq{r} P'}
+    {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
+
+    \inferrule
+    { }
+    {\constr{refl} : P \eq{\constr{zero}} P}
+
+    \inferrule
+    {eq : P \eq{r} Q}
+    {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
+
+    \inferrule
+    {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
+    {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
+\end{mathpar}
+\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
+\label{fig:struct-cong1}
+\end{figure}
 

### 2020-09-08 paper/appendix.tex
Move subject congruence to the body

@@
 
-\section{Structural Congruence}
-\label{app:struct}
-
-Subject congruence states that applying structural congruence rules to a well typed process preserves its well-typedness.
-\begin{nitheorem}[Subject congruence]
-  \label{thm:subject-congruence1}
-  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-\end{nitheorem}
-
-\begin{proof}[Proof (Sketch)]
-  The proof is by induction on \textsc{Equals} $\eq{}$.
-  Here we only consider the nontrivial base cases for $\constr{struct}$ and their symmetric variants.
-\begin{anonsuppress}
-Full proof in \cite{Zalakain2020Agda}.
-\end{anonsuppress}
-  We proceed by induction on \textsc{StructCong} $\eqeq$:
-  \begin{itemize}
-    \item
-      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
-    \item
-      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
-    \item
-      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
-    \item
-      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
-      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$..
-    \item
-      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
-      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
-    \item
-      Case $\constr{scope-comm}$:
-      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged.
-\qedhere
-  \end{itemize}
-\end{proof}
-
 \section{Substitution}

### 2020-09-08 paper/paper.tex
Move subject congruence to the body

@@
 
-\subparagraph*{Framing}
+\paragraph*{Framing}
 asserts that the well-typedness of a process is independent of its leftover resources.
@@
 
-\subparagraph*{Weakening}
+\paragraph*{Weakening}
 states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
@@
 
-\subparagraph*{Strengthening}
+\paragraph*{Strengthening}
 states that removing an unused variable preserves the well-typedness of a process.
@@
 
-\subparagraph*{Exchange}
+\paragraph*{Exchange}
 of two variables preserves the well-typedness of a process.
@@
 \subsection{Towards Subject Reduction}
-\subparagraph*{Subject Congruence}
+\paragraph*{Subject Congruence}
 states that applying structural congruence rules to a well typed process preserves its well-typedness.
@@
   The proof is by induction on \textsc{Equals} $\eq{}$.
-  For the nontrivial base cases and their symmetric variants, we apply a combination of \autoref{thm:framing}, \autoref{thm:strengthening}, \autoref{thm:weakening} and \autoref{thm:exchange}.
-  Further details can be found in \autoref{app:struct} and the full proof in%
+  Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
 \begin{anonsuppress}
-\cite{Zalakain2020Agda}
+Full proof in \cite{Zalakain2020Agda}.
 \end{anonsuppress}
-.
+  We proceed by induction on \textsc{StructCong} $\eqeq$:
+  \begin{itemize}
+    \item
+      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
+    \item
+      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
+    \item
+      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
+    \item
+      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
+      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$.
+    \item
+      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
+      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
+    \item
+      Case $\constr{scope-comm}$:
+      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged.
+\qedhere
+  \end{itemize}
 \end{proof}
 
-\subparagraph*{Substitution}
+\paragraph*{Substitution}
 \autoref{thm:substitution-generalization} proves a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
@@
 
-\subparagraph*{Subject Reduction}
+\paragraph*{Subject Reduction}
 states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.

### 2020-09-08 paper/appendix.tex
Bring lemmas forward to the body

@@
 \end{nidefinition}
-
-\section{Lemmas}
-\label{app:lemmas}
-
-\begin{nilemma}
-  \label{lm:substitution-unused}
-  For every variables $i$ and $j$, if $i \not\equiv j$ then $\Unused_i (\subst{P}{j}{i})$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:exchange-exchange}
-  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:types-unused}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:comm-capable}
-  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external to it (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
-
-\begin{proof}
-  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
-\end{proof}
 

### 2020-09-08 paper/paper.tex
Bring lemmas forward to the body

@@
 \newtheorem{nilemma}{Lemma}
+\renewcommand{\nidefinitionautorefname}{Definition}%
+\renewcommand{\nitheoremautorefname}{Theorem}%
+\newcommand{\nilemmaautorefname}{Lemma}%
 \renewcommand{\sectionautorefname}{\S}%
@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:substitution-unused} in \autoref{app:lemmas} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
+After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:substitution-unused} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
 --- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
+
+\begin{nilemma}
+  \label{lm:substitution-unused}
+  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 states that applying structural congruence rules to a well typed process preserves its well-typedness.
+We must first introduce lemmas that establish that some syntactic manipulations can be inverted and how unused variables relate to the preservation of leftovers.
+
+\begin{nilemma}
+  \label{lm:lower-lift}
+  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:exchange-exchange}
+  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:types-unused}
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+We are now in a position to prove subject congruence.
 
@@
 In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
+We first introduce a helper lemma:
+
+\begin{nilemma}
+  \label{lm:comm-capable}
+  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
+\end{nilemma}
+
+\begin{proof}
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\end{proof}
 

### 2020-09-10 paper/paper.tex
Minor fix
Another pass to the introduction

@@
 
-  We present the syntax, semantics, type system and corresponding type safety properties for the \picalc{}.
-  We generalise the algebras on multiplicities, and allow the user to use a mix of \emph{linear}, \emph{graded} and \emph{shared} types.
+  We present an untyped syntax and semantics for the \picalc{}.
+  We provide a resource-aware type system parametrised by a set of multiplicity algebras.
+  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types.
   We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs so common in type systems for linear calculi.
@@
 Notably, it features channel mobility: channels themselves are first class values, and can therefore be transmitted.
-The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and (more recently) Go \footnote{https://golang.org}.
+The \picalc{} is a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and (more recently) Go \footnote{https://golang.org}.
 At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types.
-This makes the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.
-
-J.-Y.
-Girard's development of linear logic \cite{Girard87} opened new avenues in computing science in the early '90s by introducing \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018}.
+These make the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
+
+In the early '90s, \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018} were introduced, following J.-Y. Girard's development of linear logic \cite{Girard87}.
 Linear type systems guarantee that resources are used \emph{exactly once}.
-Enforcing ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Linearity later on inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
-
-Kobayashi et al. \cite{KPT96} brought linearity to the \picalc{} and provided a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
+Enforcing the ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
+Later on, linearity inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
+
+Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
 %
-In the meantime, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98} --- a formalism used in communication-centric programming.
-With session types, linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself is used multiple times, as per its session type.
+Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98} (a formalism used in communication-centric programming), where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself is used multiple times, as per its session type.
 Linearity in session types is the key ingredient that guarantees type safety.
@@
 This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice too.
-The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, and it also allows us to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
+The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, and it allows to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
 
@@
 The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-This allows for linear, graded and shared types to be used either separately or mixed in a type system.
-Ultimately, we can exploit this formalisation as a unified framework for type safe distributed modelling and programming with a variety of type systems, and build on top of it more advanced types, theories and languages.
+This allows for linear, graded and shared types to be mixed in the same type system.
+Ultimately, we can exploit this formalisation as a unified framework for type-safe distributed modelling and programming using a variety of type systems, and build on top of it more advanced types, theories and languages.
 
@@
      For convenience, the user can convert from and to a syntax with names.
-     These conversion functions are shown to be inverses of one another modulo $\alpha$-conversion.
+     These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
 
@@
    This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can then be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
+   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can later be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
    The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
@@
 
-  \item \textbf{Leftover typing for the \picalc{} with linear, graded and shared types}:
+  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
   \autoref{type-system} provides the type system for a resource-aware \picalc{}.
@@
     \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-    Linear types, graded types and shared types are all instances of the same resource-aware algebra, giving rise to three different type systems under the same framework.
-    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
+    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
 
@@
 
-    \item \textbf{\picalc{} with leftovers}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
     This approach adds a leftover usage context to the typing judgements.
-    Typing derivations take the resources of their input usage context, consume some of them, and leave the rest as leftovers in the output usage context.
+    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
 
@@
   \end{tikzpicture}
-  \caption{Four communicating processes composed in parallel.}
+  \caption{Four communicating example processes composed in parallel.}
   \label{fig:example-process}
@@
 Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
-We use these types to give a suitable typing derivation in \autoref{example-derivation}.
+We use these types to give a suitable typing derivation in \autoref{example-derivation} (\autoref{leftover-typing}).
 
@@
 \autoref{fig:notation} illustrates the notation used in this paper.
-Data type definitions ($\type{\N}$) use double lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
-Constructors ($\constr{0}$ and $\suc$) are given as inference rules, although we occasionally use a BNF style definition of types for brevity.
-We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda --- where premises become argument types and conclusions return types.
+Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
+Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
+We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: premises become argument types and conclusions return types.
 Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
@@
 We use colours to further distinguish the different entities in this paper.
-$\type{TYPES}$ are blue (uppercased, with indices as subscripts), $\constr{constructors}$ are orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
+$\type{TYPES}$ are blue and uppercased, with indices as subscripts, $\constr{constructors}$ are orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
 
@@
     {eq : P \eq{r} P'}
-    {\constr{cong-recv} \; eq : \recv{x}P \eq{\constr{one} \; r} \recv{x}P'}
+    {\constr{cong-recv} \; eq : \recv{x}{}P \eq{\constr{one} \; r} \recv{x}{}P'}
 

### 2020-09-10 paper/paper.tex
Go over the example derivation

@@
 
-The $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$.
-After receiving twice, the process ends, so we should not be left with unused multiplicities: $u$ needs to be $\lz$.
+The $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+After receiving twice, the process ends, and we should not be left with any unused multiplicities: $u$ needs to be $\lz$.
 We will use graded types to keep track of the exact number of times communication happens.
@@
 
-The $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$.
-We instantiate the type of transmitted data $t$ to $\channel{\unit}{\func{\omega}}$, as that is something we can construct.
-As by the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
+The $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+We instantiate $t$ (the type of data that we transmit) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
+As per the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
 We will transmit once, thus using a single output multiplicity, and leave the rest as leftovers.
@@
 
-We can now put these processes together and omit specifying the types of the channels that they share: Agda can infer them from the types of the processes.
+We can now put these processes together.
+Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
 

### 2020-09-11 paper/paper.tex
a pass on abstract

@@
   The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
-  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, including \emph{session types} \cite{H93,THK94,HVK98}.
-
-  We present an untyped syntax and semantics for the \picalc{}.
-  We provide a resource-aware type system parametrised by a set of multiplicity algebras.
-  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types.
-  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs so common in type systems for linear calculi.
+  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, such as \emph{session types} \cite{H93,THK94,HVK98}.
+
+  We present an untyped syntax and semantics for the \picalc{}, on top of which we provide a resource-aware type system parametrised by a set of multiplicity algebras.
+  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types, under the same unified framework.
+  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs, which are mandatory when dealing with linear type systems.
   We provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.

### 2020-09-14 paper/paper.tex
Move example before contribs

@@
   This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types, under the same unified framework.
-  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs, which are mandatory when dealing with linear type systems.
+  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs common when dealing with linear type systems.
   We provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.
@@
 Ultimately, we can exploit this formalisation as a unified framework for type-safe distributed modelling and programming using a variety of type systems, and build on top of it more advanced types, theories and languages.
-
-\subsection{Contributions}
-\begin{enumerate}
-\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
- \begin{itemize}
-   \item \textbf{Syntax}: \autoref{syntax} uses type level \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it.
-     For convenience, the user can convert from and to a syntax with names.
-     These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
-
-   \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
-   This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can later be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
-   The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
- \end{itemize}
-
-  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
-  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
-  \begin{itemize}
-    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
-
-    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, graded and shared types.
-
-    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    This approach adds a leftover usage context to the typing judgements.
-    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
-
-    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
-  \end{itemize}
-
-  \item \textbf{Fully mechanised formalisation and type safety proofs}:
-    \autoref{type-safety} presents the type safety properties of our \picalc{} with leftovers.
-    Leftover typing for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and generalises the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
-    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
-
-    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda%
-\begin{anonsuppress}
-and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-\end{anonsuppress}
-.
-\end{enumerate}
 
@@
 \end{figure}
-
-The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous three processes (\autoref{fig:example-process}).
-The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
-Thanks to linear and graded types, the processes below can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
-Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
-We use these types to give a suitable typing derivation in \autoref{example-derivation} (\autoref{leftover-typing}).
 
@@
 
+The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous three processes (\autoref{fig:example-process}).
+The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
+Thanks to linear and graded types, the processes below can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
+Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
+We use these types to give a suitable typing derivation in \autoref{example-derivation} (\autoref{leftover-typing}).
+
 \end{example}
+
+\subsection{Contributions}
+\begin{enumerate}
+\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
+ \begin{itemize}
+   \item \textbf{Syntax}: \autoref{syntax} uses type level \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it.
+     For convenience, the user can convert from and to a syntax with names.
+     These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
+
+   \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
+   This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
+   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can later be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
+   The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+ \end{itemize}
+
+  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
+  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
+  \begin{itemize}
+    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
+    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
+
+    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
+    This allows for type systems combining linear, graded and shared types.
+
+    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+    This approach adds a leftover usage context to the typing judgements.
+    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
+
+    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+  \end{itemize}
+
+  \item \textbf{Fully mechanised formalisation and type safety proofs}:
+    \autoref{type-safety} presents the type safety properties of our \picalc{} with leftovers.
+    Leftover typing for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and generalises the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
+
+    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda%
+\begin{anonsuppress}
+and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
+\end{anonsuppress}
+.
+\end{enumerate}
 

### 2020-09-15 paper/paper.tex
Add work by Ian Mackie
pass on the intro

@@
 
-We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
+We live in a concurrent and distributed world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
 As humans, we aim to model, build, and predict such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
@@
 It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
-Notably, it features channel mobility: channels themselves are first class values, and can therefore be transmitted.
-The \picalc{} is a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and (more recently) Go \footnote{https://golang.org}.
-At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types.
+Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
+The \picalc{} is at the basis of the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and more recently Go \footnote{https://golang.org}.
+At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties, such as deadlock freedom, livelock freedom or termination, and most notably, session types.
 These make the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
@@
 Linear type systems guarantee that resources are used \emph{exactly once}.
-Enforcing the ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Later on, linearity inspired unique types (as in Clean \cite{BarendsenS96}) and ownership types (as in Rust \cite{MatsakisK14}).
-
+Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
+Later on, linearity inspired unique types, as in Clean \cite{BarendsenS96} and ownership types, as in Rust \cite{MatsakisK14}.
+
+Linearity also influenced the $\pi$-calculus.
 Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
 %
-Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98} (a formalism used in communication-centric programming), where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself is used multiple times, as per its session type.
-Linearity in session types is the key ingredient that guarantees type safety.
-
-Recently, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} into the spotlight again \cite{KPT96}.
+Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming, where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, in particular by following its session type.
+Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
+
+In recent years, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} \cite{KPT96} into the spotlight again.
 This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice too.
-The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, and it allows to use the linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be built.
-
-\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
-
-We do so by defining a specification for an algebra on \emph{multiplicities} (how many times a channel can be used) and use it to apply leftover typing (following Allais \cite{Allais2018a}) to the \picalc{} for the first time.
+The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, thus allowing to use the target linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be defined and implemented.
+
+In this paper we present \emph{the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
+
+While linear types are useful for resource-aware programming, graded and shared types are useful for a more flexible and 
+mainstream programming.
+
+We define a specification for an algebra on \emph{multiplicities} --- how many times a channel can be used --- and use it to apply \emph{leftover typing} for the first time to the \picalc{}, along the lines of Allais \cite{Allais2018a}.
 The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
@@
 
-\begin{example}[Example process]
+\begin{example}[The courier system]
 \label{example-process}
 
-We present an example that highlights the ways in which graded types can be used to prevent mistakes in the definition of \picalc{} processes.
-The syntax uses de Bruijn indices for variables and is introduced in detail in \autoref{syntax}.
-In short: $\new{} P$ creates a new channel and makes it available at index $0$ in the continuation $P$; $\comp{P}{Q}$ composes $P$ and $Q$ in parallel; $\recv{x}{}P$ receives along channel $x$ and makes the data available at index $0$ in the continuation $P$; and $\send{x}{y}{P}$ sends $y$ over channel $x$ and continues with $P$.
-Both variable references and processes are indexed by the number of free variables they can refer to (for any natural number $n$, $\Var_n$ and $\Process_n$, respectively).
-For the sake of clarity, we define the example subprocesses separately and parametrise them by the channels on which they operate.
-
-\begin{figure}[h]
-  \tiny
+We present an example that highlights the ways in which linear, graded and shared types can be used to prevent errors in the definition of \picalc{} processes.
+
+A courier system consists of three roles: a sender, who wants to send a package; a receiver, who receives the package sent by the sender and a courier, who carries the package from the sender to the receiver.
+Our courier system consists of four \picalc{} processes composed in parallel:
+two sender processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
+one receiver process, $R$, which receives the data from each of the senders;
+and finally a courier process $C$, which synchronises communication among the previous three process and
+first receives data from $S_1$ and $S_2$ on its input channels, respectively $x$ and $y$ and then sends the received data to $R$ along its output channel $z$.
+A graphical representation of the courier system is given in \autoref{fig:example-process}.
+
+Let us now consider the $\pi$-calculus process behind the courier system.
+The syntax of \picalc{} processes adopts de Bruijn indices \cite{deBruijn1972} for channel names and variables and is introduced in detail in \autoref{syntax}.
+In short, for the purposes of this example:
+process $\new{} P$ creates a new channel and makes it available at index $0$ in the continuation process $P$;
+$\comp{P}{Q}$ composes processes $P$ and $Q$ in parallel; process $\recv{x}{}P$ receives data along channel $x$ and makes that data available at index $0$ in the continuation process $P$;
+process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
+Both variable references and processes are indexed by the number of free variables they can refer to, respectively, $\Var_n$ and $\Process_n$, for any natural number $n$.
+
+For the sake of clarity, we will define the four subprocesses in our system separately and parametrise them by the channels on which they operate.
+
+\begin{figure}[t]
+  \small
   \centering
@@
     \node[draw,circle,fill opacity=0.3,text opacity=1,fill=green] (recv-z) at (4,2)   {$\func{recv} \, z$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown,align=center] (sync) at (2,0)   {$\func{sync}$\\$ \, x \, y \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=blue,align=center] (sync) at (2,0)   {$\func{cour}$\\$ \, x \, y \, z$};
 
@@
   \end{tikzpicture}
-  \caption{Four communicating example processes composed in parallel.}
+  \caption{The courier system: two senders $S_1$ and $S_2$, one receiver $R$ and one courier $C$.}
   \label{fig:example-process}
@@
 
+The receiver process $R$ is defined as follows:
 \begin{flalign*}
@@
   & \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO) \\
-  & \\
+\end{flalign*}
+
+The sender processes sender processes $S_1$ and $S_2$ are defined as follows:
+\begin{flalign*}
   & \func{send} \; : \; \Var_n \to \Process_n & \\
@@
 \end{flalign*}
+
+The courier process $C$ is defined as follows:
 \begin{flalign*}
-  & \func{sync} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
-  & \func{sync} \; in0 \; in1 \; out \; &&= \\
+  & \func{cour} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
+  & \func{cour} \; in0 \; in1 \; out \; &&= \\
   & && \recv{in0}{} \\
@@
 \begin{flalign*}
-  & \func{example} && : \; \Process_{\constr{0}} & \\
-  & \func{example} && = \new{} ( \func{send} \; \constr{0} \\
+  & \func{courier} && : \; \Process_{\constr{0}} & \\
+  & \func{courier} && = \new{} ( \func{send} \; \constr{0} \\
   & && \comp{}{ \new{} ( \func{send} \; \constr{0} } \\
   & && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
-  & && \comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
+  & && \comp{}{ \func{cour} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
 \end{flalign*}
-
-The example contains four processes that are composed in parallel: two of them send data over a channel; one receives data twice; the fourth sticks together and synchronizes the previous three processes (\autoref{fig:example-process}).
-The process doing the synchronisation first receives on both input channels and then sends the received data along the output channel.
-Thanks to linear and graded types, the processes below can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
-Dually, the process $\func{sync}$ consumes one input multiplicity on the channels shared with $\func{send}$, and two output multiplicities on the channel shared with $\func{recv}$.
+[ALIGN FIG AND PROCESSES IN PARTICULAR send c and send x,y in THE FIGURE]
+[COMMENT ON THE PROCESSES AND WHAT THEY DO]
+[EXPLAIN HERE WHAT IS TYPED WITH LIN, GRADED AND SHARED]
+Thanks to linear and graded types, the above processes can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
+Dually, the process $\func{cour}$ consumes one input multiplicity on the shared channels with $\func{send}$, respectively $x$ and $y$ in \autoref{fig:example-process} and two output multiplicities on the channel shared with $\func{recv}$, channel $z$ in \autoref{fig:example-process}.
 We use these types to give a suitable typing derivation in \autoref{example-derivation} (\autoref{leftover-typing}).
-
 \end{example}
@@
 
+To the best of our knowledge, leftover typing makes its appearance in 1994, when Ian Mackie first uses it to formulate intuitionistic linear logic \cite{Mackie}.
 Allais \cite{Allais2018a} uses leftover typing to mechanise in Agda a bidirectional type system for the linear \lambdacalc{}.

### 2020-09-16 paper/paper.tex
Preliminary pass to the example

@@
 
-% Fix autoref's example treatment 
+% Continue examples later on
+\usepackage{thmtools}
 \newtheorem{example}{Example}
-\newcommand{\exampleautorefname}{Example}
-\newcommand{\nitheoremautorefname}{Theorem}
-\newcommand{\nidefinitionautorefname}{Definition}
+\renewcommand\thmcontinues[1]{Continued}
 
@@
 \renewcommand{\nitheoremautorefname}{Theorem}%
-\newcommand{\nilemmaautorefname}{Lemma}%
+\renewcommand{\nilemmaautorefname}{Lemma}%
 \renewcommand{\sectionautorefname}{\S}%
 \renewcommand{\subsectionautorefname}{\S}%
+\renewcommand{\exampleautorefname}{Example}
 
@@
 %
-Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming, where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, in particular by following its session type.
+Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming, where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, as specified by its session type.
 Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
@@
 
-\begin{example}[The courier system]
+\begin{example}[The courier system, label=exa:cont]
 \label{example-process}
@@
 
-A courier system consists of three roles: a sender, who wants to send a package; a receiver, who receives the package sent by the sender and a courier, who carries the package from the sender to the receiver.
+A courier system consists of three roles: a sender, who wants to send a package; a receiver, who receives the package sent by the sender; and a courier, who carries the package from the sender to the receiver.
 Our courier system consists of four \picalc{} processes composed in parallel:
@@
 one receiver process, $R$, which receives the data from each of the senders;
-and finally a courier process $C$, which synchronises communication among the previous three process and
-first receives data from $S_1$ and $S_2$ on its input channels, respectively $x$ and $y$ and then sends the received data to $R$ along its output channel $z$.
+and a courier process $C$, which synchronises communication among the previous three process.
+The courier process $C$ first receives data from $S_1$ and $S_2$ on its input channels $x$ and $y$, respectively, and then sends the received data to $R$ along its output channel $z$.
 A graphical representation of the courier system is given in \autoref{fig:example-process}.
@@
     \node[draw,circle,fill opacity=0.3,text opacity=1,fill=green] (recv-z) at (4,2)   {$\func{recv} \, z$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=blue,align=center] (sync) at (2,0)   {$\func{cour}$\\$ \, x \, y \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=blue,align=center] (sync) at (2,0)   {$\func{carry}$\\$ \, x \, y \, z$};
 
@@
 
-The receiver process $R$ is defined as follows:
+The receiver process $R$, defined below, receives data twice on a channel $c$ and then terminates.
 \begin{flalign*}
@@
 
-The sender processes sender processes $S_1$ and $S_2$ are defined as follows:
+The sender processes $S_1$ and $S_2$, defined below, create a new channel that acts as the data that is being sent, send it over the channel $c$, and then terminate.
 \begin{flalign*}
@@
 
-The courier process $C$ is defined as follows:
+The courier process $C$, defined below, sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
 \begin{flalign*}
-  & \func{cour} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
-  & \func{cour} \; in0 \; in1 \; out \; &&= \\
+  & \func{carry} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
+  & \func{carry} \; in0 \; in1 \; out \; &&= \\
   & && \recv{in0}{} \\
@@
 \end{flalign*}
+
+Finally, we fit these processes together and create three communication channels: the first shared between the sender $S_1$ and the courier $C$, the second between the sender $S_2$ and the courier $C$, and the third between the receiver $R$ and the courier $C$.
+The result is the process defined below.
 \begin{flalign*}
-  & \func{courier} && : \; \Process_{\constr{0}} & \\
-  & \func{courier} && = \new{} ( \func{send} \; \constr{0} \\
+  & \func{system} && : \; \Process_{\constr{0}} & \\
+  & \func{system} && = \new{} ( \func{send} \; \constr{0} \\
   & && \comp{}{ \new{} ( \func{send} \; \constr{0} } \\
   & && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
-  & && \comp{}{ \func{cour} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
+  & && \comp{}{ \func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
 \end{flalign*}
+
 [ALIGN FIG AND PROCESSES IN PARTICULAR send c and send x,y in THE FIGURE]
-[COMMENT ON THE PROCESSES AND WHAT THEY DO]
-[EXPLAIN HERE WHAT IS TYPED WITH LIN, GRADED AND SHARED]
-Thanks to linear and graded types, the above processes can have their interactions safely restricted: both $\func{send}$ processes consume $0$ input and $1$ output multiplicities; $\func{recv}$ consumes $2$ input and $0$ output multiplicities.
-Dually, the process $\func{cour}$ consumes one input multiplicity on the shared channels with $\func{send}$, respectively $x$ and $y$ in \autoref{fig:example-process} and two output multiplicities on the channel shared with $\func{recv}$, channel $z$ in \autoref{fig:example-process}.
-We use these types to give a suitable typing derivation in \autoref{example-derivation} (\autoref{leftover-typing}).
+
+We continue this example by providing typing derivations for these processes in \autoref{leftover-typing}.
+We use graded types for the usage annotations of the communication channels $x$, $y$, and $z$: both processes $\func{send} \; x$ and $\func{send} \; y$ consume $1$ output multiplicity; $\func{recv} \; z$ consumes $2$ input multiplicities.
+Dually, the process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and $y$, and two output multiplicities on $z$.
+
+[TODO: type of data being transmitted]
 \end{example}
@@
 
-\begin{example}[Typing derivation]
+\begin{example}[Typing derivation, continues=exa:cont]
 \label{example-derivation}
@@
 \begin{flalign*}
-& \func{\vdash sync} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
-& \func{\vdash sync} && = \\
+& \func{\vdash carry} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
+& \func{\vdash carry} && = \\
 & && \recv{\suc \suc \constr{0}}{} \\
@@
 \begin{flalign*}
-& \func{\vdash example} && : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}} & \\
-& \func{\vdash example} && = \new{} \; \_ \; \_ \; \_ \; ( \func{\vdash send} \\
+& \func{\vdash system} && : \; \types{\constr{[]}}{\constr{[]}}{\func{system}}{\constr{[]}} & \\
+& \func{\vdash system} && = \new{} \; \_ \; \_ \; \_ \; ( \func{\vdash send} \\
 & && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash send} \\
 & && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash recv} \\
-& && \comp{}{ \func{\vdash sync} }))) \\
+& && \comp{}{ \func{\vdash carry} }))) \\
 \end{flalign*}

### 2020-09-16 paper/paper.tex
Second pass to example

@@
   \end{tikzpicture}
-  \caption{The courier system: two senders $S_1$ and $S_2$, one receiver $R$ and one courier $C$.}
+  \caption{The courier system: two \textcolor{orange}{senders} $S_1$ and $S_2$, one \textcolor{green}{receiver} $R$ and one \textcolor{blue}{courier} $C$.}
   \label{fig:example-process}
@@
 
-We continue this example by providing typing derivations for these processes in \autoref{leftover-typing}.
-We use graded types for the usage annotations of the communication channels $x$, $y$, and $z$: both processes $\func{send} \; x$ and $\func{send} \; y$ consume $1$ output multiplicity; $\func{recv} \; z$ consumes $2$ input multiplicities.
+We continue this example and provide typing derivations for these processes in \autoref{leftover-typing}.
+We use \textbf{shared types} for the type of data that is being sent over these channels, and \textbf{linear types} for its usage annotations.
+We use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$ themselves: both processes $\func{send} \; x$ and $\func{send} \; y$ consume $1$ output multiplicity; $\func{recv} \; z$ consumes $2$ input multiplicities.
 Dually, the process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and $y$, and two output multiplicities on $z$.
-
-[TODO: type of data being transmitted]
 \end{example}

### 2020-09-16 paper/paper.tex
Add Orchard's encoding of effects into session types
Swap green for darkgray

@@
     \node[draw,circle,fill opacity=0.3,text opacity=1,fill=orange] (send-y) at (2,2)   {$\func{send} \, y$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=green] (recv-z) at (4,2)   {$\func{recv} \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=darkgray] (recv-z) at (4,2)   {$\func{recv} \, z$};
     \node[draw,circle,fill opacity=0.3,text opacity=1,fill=blue,align=center] (sync) at (2,0)   {$\func{carry}$\\$ \, x \, y \, z$};
@@
   \end{tikzpicture}
-  \caption{The courier system: two \textcolor{orange}{senders} $S_1$ and $S_2$, one \textcolor{green}{receiver} $R$ and one \textcolor{blue}{courier} $C$.}
+  \caption{The courier system: two \textcolor{orange}{senders} $S_1$ and $S_2$, one \textcolor{darkgray}{receiver} $R$ and one \textcolor{blue}{courier} $C$.}
   \label{fig:example-process}
@@
 
+Orchard and Yoshida \cite{OrchardY16} embed a small effectul imperative language into the session-typed \picalc{}, showing that session types are expressive enough to encode effect systems.
+
 Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.

### 2020-09-16 paper/paper.tex
Clarify instantiations in the example

@@
 
-The receiver process $R$, defined below, receives data twice on a channel $c$ and then terminates.
+The receiver process $R$, defined below and instantiated as $\func{recv} \; z$, receives data twice on a channel $c$ and then terminates.
 \begin{flalign*}
@@
 
-The sender processes $S_1$ and $S_2$, defined below, create a new channel that acts as the data that is being sent, send it over the channel $c$, and then terminate.
+The sender processes $S_1$ and $S_2$, defined below and instantiated as $\func{send} \; x$ and $\func{send} \; y$, create a new channel that acts as the data that is being sent, send it over the channel $c$, and then terminate.
 \begin{flalign*}
@@
 
-The courier process $C$, defined below, sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
+The courier process $C$, defined below and instantiated as $\func{carry} \; x \; y \; z$, sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
 \begin{flalign*}
@@
 \end{flalign*}
-
-[ALIGN FIG AND PROCESSES IN PARTICULAR send c and send x,y in THE FIGURE]
 

### 2020-09-17 paper/paper.tex
Be careful to preserve the example title
another pass to the example

--- 2020-09-16 paper/paper.tex
@@
 
-\begin{example}[The courier system, label=exa:cont]
+\begin{example}[The courier system]
 \label{example-process}
-
-We present an example that highlights the ways in which linear, graded and shared types can be used to prevent errors in the definition of \picalc{} processes.
-
-A courier system consists of three roles: a sender, who wants to send a package; a receiver, who receives the package sent by the sender; and a courier, who carries the package from the sender to the receiver.
-Our courier system consists of four \picalc{} processes composed in parallel:
-two sender processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
-one receiver process, $R$, which receives the data from each of the senders;
-and a courier process $C$, which synchronises communication among the previous three process.
-The courier process $C$ first receives data from $S_1$ and $S_2$ on its input channels $x$ and $y$, respectively, and then sends the received data to $R$ along its output channel $z$.
+\newcommand{\sender}{\textcolor{brown}{sender}}
+\newcommand{\receiver}{\textcolor{purple}{receiver}}
+\newcommand{\courier}{\textcolor{cyan}{courier}}
+
+We present an example that highlights the ways in which linear, graded and shared types can be used to prevent errors in a concurrent and distributed system defined by \picalc{} processes in parallel.
+
+A courier system consists of three roles:
+a \sender{}, who wants to send a package;
+a \receiver{}, who receives the package sent by the sender;
+and a \courier{}, who carries the package from the sender to the receiver.
+Our courier system is defined by four \picalc{} processes in parallel inhabiting the above three roles:
+we have two \sender{} processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
+one \receiver{} process, $R$, which receives the data sent from each of the senders--hence receives twice;
+and a \courier{} process $C$, which synchronises communication among $S_1$, $S_2$ and $R$.
+The \courier{} process $C$ first receives data from $S_1$ and $S_2$ on its input channels $x$ and $y$, respectively, and then sends the received data to $R$ along its output channel $z$.
 A graphical representation of the courier system is given in \autoref{fig:example-process}.
 
-Let us now consider the $\pi$-calculus process behind the courier system.
-The syntax of \picalc{} processes adopts de Bruijn indices \cite{deBruijn1972} for channel names and variables and is introduced in detail in \autoref{syntax}.
+We will now define the courier system as \picalc{} processes.
+In our framework, the syntax of \picalc{} processes adopts de Bruijn indices \cite{deBruijn1972} for channel names and variables and is introduced in detail in \autoref{syntax}.
 In short, for the purposes of this example:
@@
 process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
-Both variable references and processes are indexed by the number of free variables they can refer to, respectively, $\Var_n$ and $\Process_n$, for any natural number $n$.
-
-For the sake of clarity, we will define the four subprocesses in our system separately and parametrise them by the channels on which they operate.
+Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to, for any natural number $n$.
+
+For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
 
@@
   \begin{tikzpicture}[draw,circle]
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=orange] (send-x) at (0,2)   {$\func{send} \, x$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=orange] (send-y) at (2,2)   {$\func{send} \, y$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=darkgray] (recv-z) at (4,2)   {$\func{recv} \, z$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=blue,align=center] (sync) at (2,0)   {$\func{carry}$\\$ \, x \, y \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown] (send-x) at (0,2)   {$\func{send} \, x$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown] (send-y) at (2,2)   {$\func{send} \, y$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=purple] (recv-z) at (4,2)   {$\func{recv} \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=cyan,align=center] (sync) at (2,0)   {$\func{carry}$\\$ \, x \, y \, z$};
 
@@
   \end{tikzpicture}
-  \caption{The courier system: two \textcolor{orange}{senders} $S_1$ and $S_2$, one \textcolor{darkgray}{receiver} $R$ and one \textcolor{blue}{courier} $C$.}
+  \caption{The courier system: two \sender{} $S_1$ and $S_2$, one \receiver{} $R$ and one \courier{} $C$.}
   \label{fig:example-process}
@@
 
-The receiver process $R$, defined below, receives data twice on a channel $c$ and then terminates.
+Each of the roles \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$ with signature $\Var_n \to \Process_n$.
+
+The \sender{} role is defined below.
+The \sender{} creates a new channel $\new{} \suc c$ denoting the data to be sent, and sends it over channel $c$, and then terminates, denoted by $\P0$.
+\begin{flalign*}
+  & \func{send} \; : \; \Var_n \to \Process_n & \\
+  & \func{send} \; c = \new{} \send{\suc c}{\constr{0}} \PO
+\end{flalign*}
+Each \sender{} process $S_1$ and $S_2$ will  be an instantiation of $\func{send} \; c$.
+In particular
+$S_1$ will be defined as $\func{send} \; x$ and $S_2$ as $\func{send} \; y$, where data is sent over channels $x$ and $y$, respectively (see \autoref{fig:example-process}).
+
+The \receiver{} role is defined below, where $\func{recv} \; c \;$ receives data \emph{twice} on a channel $c$ and then terminates with $\PO$.
 \begin{flalign*}
   & \func{recv} \; : \; \Var_n \to \Process_n & \\
-  & \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO) \\
+  & \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO)
 \end{flalign*}
-
-The sender processes $S_1$ and $S_2$, defined below, create a new channel that acts as the data that is being sent, send it over the channel $c$, and then terminate.
-\begin{flalign*}
-  & \func{send} \; : \; \Var_n \to \Process_n & \\
-  & \func{send} \; c = \new{} \send{\suc c}{\constr{0}} \PO \\
-\end{flalign*}
-
-The courier process $C$, defined below, sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
+The \receiver{} process $R$ will be instantiated as $\func{recv} \; z$, where $z$ is the channel used to receive data from the \courier{} (see \autoref{fig:example-process}).
+
+The \courier{} role is defined below.
 \begin{flalign*}
@@
   & && \send{(\suc \suc \; out)}{\suc \constr{0}} \\
-  & && \send{(\suc \suc \; out)}{\constr{0}} \PO \\
+  & && \send{(\suc \suc \; out)}{\constr{0}} \PO
 \end{flalign*}
-
-Finally, we fit these processes together and create three communication channels: the first shared between the sender $S_1$ and the courier $C$, the second between the sender $S_2$ and the courier $C$, and the third between the receiver $R$ and the courier $C$.
-The result is the process defined below.
+The \courier{} process $C$ will be instantiated as $\func{carry} \; x \; y \; z$, which sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
+Here channels $x,y,z$ instantiate respectively $in0,in1,out$.
+
+Finally, we compose these processes together and create three communication channels: the first channel is shared between the \sender{} $S_1$ and the \courier{} $C$, the second between the \sender{} $S_2$ and the \courier{} $C$, and the third between the \receiver{} $R$ and the \courier{} $C$.
+The result is the courier systems defined below.
 \begin{flalign*}
@@
 
-[ALIGN FIG AND PROCESSES IN PARTICULAR send c and send x,y in THE FIGURE]
-
-We continue this example and provide typing derivations for these processes in \autoref{leftover-typing}.
-We use \textbf{shared types} for the type of data that is being sent over these channels, and \textbf{linear types} for its usage annotations.
-We use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$ themselves: both processes $\func{send} \; x$ and $\func{send} \; y$ consume $1$ output multiplicity; $\func{recv} \; z$ consumes $2$ input multiplicities.
-Dually, the process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and $y$, and two output multiplicities on $z$.
+In order to type the courier system we will use all three typing disciplines: linear, graded and shared typing.
+We will continue this example and provide the typing derivations in \autoref{leftover-typing}.
+We use \textbf{shared types} to type the data that is being sent from the \sender{} processes to the \receiver{};
+we use \textbf{linear types} for the usage annotations of [FILL IN, NOT CLEAR BEFORE];
+and finally, we use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$: the \sender{} processes $\func{send} \; x$ and $\func{send} \; y$ each consume $1$ output multiplicity and the \receiver{} process $\func{recv} \; z$ consumes $2$ input multiplicities, as it receives twice on channel $z$.
+Dually, the \courier{} process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and one input multiplicity on $y$, and two output multiplicities on $z$.
 \end{example}
@@
 
-\begin{example}[Typing derivation, continues=exa:cont]
+\begin{example}[Typing derivation, continues=example-process]
 \label{example-derivation}

### 2020-09-17 paper/paper.tex
because colours are important!

@@
 \newcommand{\constr}[1]{\textcolor{orange}{\operatorname{#1}}}
-\newcommand{\func}[1]{\textcolor{olive}{\operatorname{#1}}}
+\newcommand{\func}[1]{\textcolor{teal}{\operatorname{#1}}}
 
@@
 We use colours to further distinguish the different entities in this paper.
-$\type{TYPES}$ are blue and uppercased, with indices as subscripts, $\constr{constructors}$ are orange, $\func{functions}$ are olive green, variables are black, and some constructor names are overloaded --- and disambiguated by context.
+$\type{TYPES}$ are blue and uppercased, with indices as subscripts, $\constr{constructors}$ are orange, $\func{functions}$ are teal, variables are black, and some constructor names are overloaded --- and disambiguated by context.
 

### 2020-09-17 paper/paper.tex
a pass to contributions

@@
 We will now define the courier system as \picalc{} processes.
-In our framework, the syntax of \picalc{} processes adopts de Bruijn indices \cite{deBruijn1972} for channel names and variables and is introduced in detail in \autoref{syntax}.
+In our framework, the syntax of \picalc{} processes adopts \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} for channel names and variables and is introduced in detail in \autoref{syntax}.
 In short, for the purposes of this example:
@@
  \begin{itemize}
-   \item \textbf{Syntax}: \autoref{syntax} uses type level \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it.
-     For convenience, the user can convert from and to a syntax with names.
+   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+   For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
    These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
 
-   \item \textbf{Semantics}: \autoref{semantics} provides an operational semantics for the \picalc{}, prior to any typing.
+   \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
    This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel the communication occurs on, so that this information can later be used to state subject reduction --- aka type preservation (\autoref{thm:subject-reduction}).
-   The reduction relation is defined modulo structural congruence (\autoref{structural-congruence}) --- a relation that acts as a quotient type that removes unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+   The reduction relation tracks at the type level the channel on which communication occurs.
+This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
+   The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
  \end{itemize}
@@
   \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
-  \autoref{type-system} provides the type system for a resource-aware \picalc{}.
+  we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
   \begin{itemize}
@@
     Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, graded and shared types.
+    This allows for type systems combining linear, graded and shared types under the same framework.
 
@@
     Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
-
-    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and makes it possible to generalise the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
   \end{itemize}
 
-  \item \textbf{Fully mechanised formalisation and type safety proofs}:
-    \autoref{type-safety} presents the type safety properties of our \picalc{} with leftovers.
-    Leftover typing for a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and generalises the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
-    We use these to prove \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove \emph{subject reduction} (\autoref{thm:subject-reduction}).
-
-    The formalisation of the \picalc{} with leftovers has been fully mechanised in Agda%
+  \item \textbf{Fully mechanised formalisation and type safety results}:
+    The formalisation of the \picalc{} with leftovers typing, from syntax to semantics and the type system, has been fully mechanised in Agda.
+
+    We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
+    We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
+    We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
+\end{enumerate}
 \begin{anonsuppress}
@@
 \end{anonsuppress}
-.
-\end{enumerate}
 
@@
 \subsection{From names to de Bruijn indices and back}
-
+\label{from_to_deBruijn}
 In order to demonstrate the correspondence between an encoding that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.

### 2020-09-17 paper/paper.tex
pass on syntax

@@
 
-The syntax of the \picalc{} \cite{Sangio01} is given by the grammar in \autoref{fig:syntax-names}.
-Process $\PO$ denotes the terminated process, where no operations can further occur.
+The syntax of the \picalc{} \cite{Sangio01} adopting channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
+Channel names and variables range over $x,y,z$ in $\Name$ and processes over $P,Q,R$ in $\Raw$.
+
+Process $\PO$ denotes the terminated process, where no further communications can occur.
 Process $(\new{}x)\; P$ creates a new channel $x$ bound with scope $P$.
 Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
+Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only process constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
 
@@
 Names are cumbersome to mechanise: they are not inherently well-scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
-Instead, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
+To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
 That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
@@
 It is useful to rule out these ill-scoped terms syntactically.
-In \autoref{fig:syntax}, we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+In \autoref{fig:syntax} --- which presents the syntax of the \picalc{} using de Bruijn indices --- we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
@@
 \label{from_to_deBruijn}
-In order to demonstrate the correspondence between an encoding that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
-
-\begin{example}
-  Example process using names (top) translated to de Bruijn indices (middle) and back to names (bottom).
-The result is a process that follows the Barendregt convention and is $\alpha$-equivalent to the original.
-  \small
-  \begin{equation*}
-    \begin{aligned}
-      & (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
-      & \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
-      & (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)}) \\
-    \end{aligned}
-  \end{equation*}
-\end{example}
+In order to demonstrate the correspondence between an \picalc{} that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
 
@@
 
+\begin{example}[$\func{fromRaw}$ and $\func{toRaw}$]
+  \begin{figure}[t]
+    \small
+    \begin{equation*}
+      \begin{aligned}
+        &P = (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
+        &Q = \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
+        &R = (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)}) \\
+      \end{aligned}
+    \end{equation*}
+    \caption{From names to de Bruijn indices}
+    \label{fig:conversion}
+    \end{figure}
+  We illustrate the conversion functions from names to de Bruijn indices $\func{fromRaw}$ and back $\func{toRaw}$ with three processes $P,Q,R$ given in \autoref{fig:conversion}.
+
+  Process $P$ uses names $x,y,z$ and it is translated via the conversion function $\func{fromRaw}$ into process $Q$, which uses de Bruijn indices.
+  Process $Q$ is then translated via $\func{toRaw}$ into process $R$, which follows the \emph{Barendregt convention}\footnote{The Barendregt variable convention states that all bound variables/names in a process are distinct among each other and from the free variables/names.} and is $\alpha$-equivalent to the original process $P$.
+  [ADJUST SPACING OF THE EXAMPLE IN FIG 5]
+\end{example}
+
+In the following we present the main results that our conversion functions satisfy.
+
 \begin{nilemma}
-  Translating from de Bruijn indices to names results in a well-scoped process.
+  Translating from de Bruijn indices to names via $\func{fromRaw}$ results in a well-scoped process.
 \end{nilemma}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names results in a process with no shadowed variables.
+  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process with no shadowed variables.
 \end{nilemma}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names and back results in the same process modulo internal variable name hints.
+  Translating from de Bruijn indices to names and back via via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
 \end{nilemma}
@@
 \begin{nilemma}
-  Translating from names to de Bruijn indices and back results in the same process modulo $\alpha$-conversion.
+  Translating from names to de Bruijn indices and back via $\func{toRaw} \circ \func{fromRaw}$ results in the same process modulo $\alpha$-conversion.
 \end{nilemma}
 
-\begin{proof}[Prove (Sketch)]
-  All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
-  For details, refer to our mechanisation in Agda%
+\begin{proof}[Proof (Sketch)]
+  All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
+  Complete details can be found in our mechanisation in Agda%
 \begin{anonsuppress}
-\cite{Zalakain2020Agda}
+in \cite{Zalakain2020Agda}
 \end{anonsuppress}

### 2020-09-18 paper/paper.tex
Remove italics from examples, better align syntax example

@@
 \usepackage{thmtools}
-\newtheorem{example}{Example}
 \renewcommand\thmcontinues[1]{Continued}
@@
 \newtheorem{nilemma}{Lemma}
+\newtheorem{example}{Example}
 \renewcommand{\nidefinitionautorefname}{Definition}%
@@
 
-The syntax of the \picalc{} \cite{Sangio01} adopting channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
+The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
 Channel names and variables range over $x,y,z$ in $\Name$ and processes over $P,Q,R$ in $\Raw$.
@@
 We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such proof term.
-\begin{equation*}
-  \begin{aligned}
+\begin{alignat*}{2}
     &\func{fromRaw} && : (ctx : \Names_n) \; (P : \Raw) \\
-    &               && \to \WellScoped \; ctx \; P \to \Process_n \\
-  \end{aligned}
-\end{equation*}
+    &               && \to \WellScoped \; ctx \; P \to \Process_n
+\end{alignat*}
 
@@
 \begin{equation*}
-  \begin{aligned}
     \func{toRaw} : (ctx : \Names_n) \to \Process_n \to \Raw
-  \end{aligned}
 \end{equation*}
@@
   \begin{figure}[t]
-    \small
-    \begin{equation*}
-      \begin{aligned}
+    \begin{alignat*}{7}
       &P = (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
       &Q = \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
-        &R = (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)}) \\
-      \end{aligned}
-    \end{equation*}
+      &R = (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)})
+    \end{alignat*}
     \caption{From names to de Bruijn indices}
@@
   Process $Q$ is then translated via $\func{toRaw}$ into process $R$, which follows the \emph{Barendregt convention}\footnote{The Barendregt variable convention states that all bound variables/names in a process are distinct among each other and from the free variables/names.} and is $\alpha$-equivalent to the original process $P$.
-  [ADJUST SPACING OF THE EXAMPLE IN FIG 5]
 \end{example}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names and back via via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
+  Translating from de Bruijn indices to names and back via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
 \end{nilemma}

### 2020-09-21 paper/paper.tex
Minor changes

@@
   & \func{send} \; : \; \Var_n \to \Process_n & \\
-  & \func{send} \; c = \new{} \send{\suc c}{\constr{0}} \PO
+  & \func{send} \; c = \new{} \send{(\suc c)}{\constr{0}} \PO
 \end{flalign*}
@@
   & \func{recv} \; : \; \Var_n \to \Process_n & \\
-  & \func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO)
+  & \func{recv} \; c \; = \recv{c}{} \recv{(\suc c)} \PO
 \end{flalign*}
@@
 & \func{\vdash carry} && = \\
-& && \recv{\suc \suc \constr{0}}{} \\
-& && \recv{\suc \suc \constr{0}}{} \\
-& && \send{\suc \suc \constr{0}}{\suc \constr{0}} \\
-& && \send{\suc \suc \constr{0}}{\constr{0}} \PO \\
+& && \recv{(\suc \suc \constr{0})}{} \\
+& && \recv{(\suc \suc \constr{0})}{} \\
+& && \send{(\suc \suc \constr{0})}{\suc \constr{0}} \\
+& && \send{(\suc \suc \constr{0})}{\constr{0}} \PO \\
 \end{flalign*}

### 2020-09-21 paper/paper.tex
Update paper.tex

@@
 
-A courier system consists of three roles:
+A courier system consists of three \emph{roles}:
 a \sender{}, who wants to send a package;
@@
 and a \courier{}, who carries the package from the sender to the receiver.
-Our courier system is defined by four \picalc{} processes in parallel inhabiting the above three roles:
+
+Our courier system (given in \autoref{fig:example-process}) is defined by four \picalc{} processes in parallel instantiating the above three roles:
 we have two \sender{} processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
-one \receiver{} process, $R$, which receives the data sent from each of the senders--hence receives twice;
+one \receiver{} process, $R$, which receives over channel $z$ the data sent from each of the senders--hence receives twice;
 and a \courier{} process $C$, which synchronises communication among $S_1$, $S_2$ and $R$.
 The \courier{} process $C$ first receives data from $S_1$ and $S_2$ on its input channels $x$ and $y$, respectively, and then sends the received data to $R$ along its output channel $z$.
-A graphical representation of the courier system is given in \autoref{fig:example-process}.
 
 We will now define the courier system as \picalc{} processes.
-In our framework, the syntax of \picalc{} processes adopts \emph{de Bruijn indices} \cite{deBruijn1972, Dybjer1994} for channel names and variables and is introduced in detail in \autoref{syntax}.
+In our work, the syntax of \picalc{} processes adopts \emph{de Bruijn indices} \cite{deBruijn1972} for channel names and variables (details in \autoref{syntax}).
 In short, for the purposes of this example:
@@
 Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to, for any natural number $n$.
-
-For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
 
@@
 \end{figure}
-
-Each of the roles \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$ with signature $\Var_n \to \Process_n$.
-
-The \sender{} role is defined below.
-The \sender{} creates a new channel $\new{} \suc c$ denoting the data to be sent, and sends it over channel $c$, and then terminates, denoted by $\P0$.
+For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
+Each roles \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$ with signature $\Var_n \to \Process_n$.
+
+The \sender{} \emph{role} is defined below.
+The \sender{} creates a new channel $\new{} \suc c$ denoting the data to be sent, and sends it over channel $c$, and then terminates, denoted by process $\PO$.
 \begin{flalign*}
@@
 \end{flalign*}
-Each \sender{} process $S_1$ and $S_2$ will  be an instantiation of $\func{send} \; c$.
-In particular
-$S_1$ will be defined as $\func{send} \; x$ and $S_2$ as $\func{send} \; y$, where data is sent over channels $x$ and $y$, respectively (see \autoref{fig:example-process}).
+Each \sender{} \emph{process} $S_1$ and $S_2$ is an instantiation of $\func{send} \; c$.
+In particular,
+$S_1$ is defined as $\func{send} \; x$ and $S_2$ as $\func{send} \; y$, where data is sent over channels $x$ and $y$, respectively (see \autoref{fig:example-process}).
 
@@
 \end{flalign*}
-The \receiver{} process $R$ will be instantiated as $\func{recv} \; z$, where $z$ is the channel used to receive data from the \courier{} (see \autoref{fig:example-process}).
+The \receiver{} process $R$ is defined as  $\func{recv} \; z$, with channel $z$ instantiating $c$, which is the channel used to receive data from \courier{} (see \autoref{fig:example-process}).
 
 The \courier{} role is defined below.
+The \courier{} process $C$ is defined as $\func{carry} \; x \; y \; z$.
+It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$ (see \autoref{fig:example-process}).
 \begin{flalign*}
@@
 \end{flalign*}
-The \courier{} process $C$ will be instantiated as $\func{carry} \; x \; y \; z$, which sequentially receives on the two input channels $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $out$.
-Here channels $x,y,z$ instantiate respectively $in0,in1,out$.
 
@@
 We will continue this example and provide the typing derivations in \autoref{leftover-typing}.
-We use \textbf{shared types} to type the data that is being sent from the \sender{} processes to the \receiver{};
-we use \textbf{linear types} for the usage annotations of [FILL IN, NOT CLEAR BEFORE];
-and finally, we use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$: the \sender{} processes $\func{send} \; x$ and $\func{send} \; y$ each consume $1$ output multiplicity and the \receiver{} process $\func{recv} \; z$ consumes $2$ input multiplicities, as it receives twice on channel $z$.
+We use \textbf{linear types} for the \emph{usage annotations} of data and \textbf{shared types} for the \emph{data} itself that is being sent from the \sender{} processes to the \receiver{}.
+We use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$: the \sender{} processes $\func{send} \; x$ and $\func{send} \; y$ each consume $1$ output multiplicity and the \receiver{} process $\func{recv} \; z$ consumes $2$ input multiplicities, as it receives twice on channel $z$.
 Dually, the \courier{} process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and one input multiplicity on $y$, and two output multiplicities on $z$.
@@
    \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
-   This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
+   Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
    The reduction relation tracks at the type level the channel on which communication occurs.
@@
 \autoref{fig:notation} illustrates the notation used in this paper.
-Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names (for ease of reference).
+Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.
 Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
@@
     \end{alignat*}
-    \caption{From names to de Bruijn indices}
+    \caption{From names to de Bruijn indices and back}
     \label{fig:conversion}
     \end{figure}
-  We illustrate the conversion functions from names to de Bruijn indices $\func{fromRaw}$ and back $\func{toRaw}$ with three processes $P,Q,R$ given in \autoref{fig:conversion}.
-
-  Process $P$ uses names $x,y,z$ and it is translated via the conversion function $\func{fromRaw}$ into process $Q$, which uses de Bruijn indices.
+  We illustrate the conversion functions from names to de Bruijn indices $(\func{fromRaw}$) and back ($\func{toRaw}$) with three processes $P,Q,R$  in \autoref{fig:conversion}.
+
+  Process $P$ uses names $x,y,z$ and is translated via the conversion function $\func{fromRaw}$ into process $Q$, which uses de Bruijn indices.
   Process $Q$ is then translated via $\func{toRaw}$ into process $R$, which follows the \emph{Barendregt convention}\footnote{The Barendregt variable convention states that all bound variables/names in a process are distinct among each other and from the free variables/names.} and is $\alpha$-equivalent to the original process $P$.
@@
 The last three ($\constr{scope}-$*) state garbage collection, scope extrusion and commutativity of restrictions, respectively.
-
 In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$ (see \autoref{app:definitions}).
@@
 We lift the relation \textsc{StrucCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$ as shown in \autoref{fig:struct-cong1}.
-The $\func{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively restriction, composition, input and output.
-The remaining three rules close structural congruence under reflexivity, symmetry and transitivity. 
+The $\constr{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively scope restriction, parallel composition, input and output.
+The remaining three rules close structural congruence under reflexivity, symmetry and transitivity, respectively. 
 The transitivity rule makes it tricky to prove predicates by induction on \textsc{Equals}: Agda cannot immediately see that such doubly recursive calls terminate.
@@
 We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
-We distinguish between channels that are created inside of the process ($\constr{internal}$) and channels that are created outside ($\constr{external}\ i$), where $i$ is the index of the channel variable).
-
+
+We distinguish between channels that are created inside of the process ($\constr{internal}$), and channels that are created outside ($\constr{external}\ i$), where $i$ is the index of the channel variable.
 In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel with index ${i}$.
@@
 
-In $\constr{res}$, we use $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
-This $\func{dec}$ function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
+Rule $\constr{res}$ uses $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
+This function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
 
@@
 \autoref{fig:algebra-instances} sketches the implementation of linear, graded and shared types as instances of our usage algebra.
-Their use in typing derivations is illustrated in \autoref{example-derivation}.
+Their use in typing derivations is illustrated in \autoref{example-derivation} (Continued).
 

### 2020-09-21 paper/paper.tex
on typing and the example

@@
 
+% Example
+\newcommand{\sender}{\textcolor{brown}{sender}}
+\newcommand{\receiver}{\textcolor{purple}{receiver}}
+\newcommand{\courier}{\textcolor{cyan}{courier}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{example-process}
-\newcommand{\sender}{\textcolor{brown}{sender}}
-\newcommand{\receiver}{\textcolor{purple}{receiver}}
-\newcommand{\courier}{\textcolor{cyan}{courier}}
 
@@
 In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise over this notion of multiplicities by defining an algebra for them that is satisfied by linear, graded and shared types alike.
-We then use pairs of multiplicities as usage annotations for channels.
+We generalise over this notion by defining an algebra for multiplicities that is satisfied by linear, graded and shared types alike.
+We then use pairs of multiplicities as usage annotations for a channel's input and output capabilities.
 
 \begin{nidefinition}[Usage algebra]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$, that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational resoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal} (so that new resources cannot arbitrarily spring into life).
@@
 The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.
-This notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
+This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
 
@@
 
-Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look as follows:
+Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look like:
 \begin{mathpar}
@@
 As a result, usage information is threaded bottom-up through the typing derivation, and top-down context split proofs are no longer needed.
-In addition to this, it allows a \emph{framing} (\autoref{thm:framing}) theorem to be stated, and for the \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems to acquire a more general form.
+In addition, it allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.
 
@@
   The continuation process $P$ is provided with the new channel with usage annotation $y \comma y$, which it must completely exhaust.
-
+%
   The input process in rule $\recv{}{}$ requires a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
@@
   The received element $x$ must be completely exhausted by the continuation process.
-
+%
   Similarly to input, the output process in rule $\send{}{}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
@@
   Note that both indices $i$ and $j$ are used in the syntax of the typed process.
-
+%
   Parallel composition in rule $\comp{}{}$ uses the leftovers of the left-hand process to type the right-hand process.
@@
 
-We provide an example typing derivation for the processes defined in \autoref{example-process}.
+We provide the typing derivation for the courier system defined in \autoref{example-process}.
 For the sake of simplicity, we instantiate these processes with concrete variable references before typing them.
 
-The $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+The \receiver{} defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
 After receiving twice, the process ends, and we should not be left with any unused multiplicities: $u$ needs to be $\lz$.
@@
 
-The $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+The \sender{} defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
 We instantiate $t$ (the type of data that we transmit) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
@@
 
-Dually, the process doing the synchronisation expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
+Dually, the \courier{} defined by the $\func{carry}$ process expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
 The typing derivation here is uniquely determined by the type and syntax of the process.
@@
 
-We can now put these processes together.
+We can now put these processes together to type the courier system.
 Specifying the types of the channels is not required: Agda can infer them from the types of the processes.

### 2020-09-22 paper/paper.tex
a pass on technical parts

@@
 
-\begin{proof}[Proof (Sketch)]
+\begin{proof}[Proof]
   All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
@@
   It has an element $\one$ that is used to count inputs and outputs.
-  \autoref{fig:multiplicities} models such an algebra as a record $\Algebra_C$ on a carrier $C$.
+  \autoref{fig:multiplicities} defines such an algebra as a record $\Algebra_C$ on a carrier $C$.
 
@@
 \begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
-  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (\autoref{fig:prectx}).
+  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two rows of \autoref{fig:prectx}).
 
@@
   \end{figure}
-\end{nidefinition}
-
 The unit type $\unit$ serves as a proof of inhabitance for types.
@@
 The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.
-This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input and output, respectively \cite{K07}.
-
-A \emph{typing context} (last row in \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
+This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input $i$ and output $o$ capabilities, respectively \cite{K07}.
+
+A \emph{typing context} (last two rows of \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
+\end{nidefinition}
 
@@
 
-  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ tuple of input and output multiplicities, respectively.
+  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
   Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
@@
 
-The \receiver{} defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+The \receiver{} defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
 After receiving twice, the process ends, and we should not be left with any unused multiplicities: $u$ needs to be $\lz$.
@@
 Once the types are given, the typing derivation is completely syntax directed.
-
-The \sender{} defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
-We instantiate $t$ (the type of data that we transmit) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
+\begin{flalign*}
+  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO
+\end{flalign*}
+
+The \sender{} defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
+We instantiate $t$ (the type of data that the \sender{} sends) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
 As per the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
 We will transmit once, thus using a single output multiplicity, and leave the rest as leftovers.
-
 \begin{flalign*}
-  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
-  & \\
   & \func{\vdash send} \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}}{\Gamma \comma (l \comma \suc r)}{\func{send} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO \\
+  & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
 \end{flalign*}
@@
 The typing derivation here is uniquely determined by the type and syntax of the process.
-
 \begin{flalign*}
@@
 & && \send{(\suc \suc \constr{0})}{\suc \constr{0}} \\
-& && \send{(\suc \suc \constr{0})}{\constr{0}} \PO \\
+& && \send{(\suc \suc \constr{0})}{\constr{0}} \PO
 \end{flalign*}
 
-We can now put these processes together to type the courier system.
+We can now compose these processes in parallel and type the courier system.
 Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
-
 \begin{flalign*}
@@
 & && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash recv} \\
-& && \comp{}{ \func{\vdash carry} }))) \\
+& && \comp{}{ \func{\vdash carry} })))
 \end{flalign*}
-
 \end{example}
@@
 \subsection{Auxiliary Results}
+We will start this section with auxiliary results which are needed to prove two main theorems: subject congruence (\autoref{thm:subject-congruence}) and subject reduction (\autoref{thm:subject-reduction}) for our \picalc{} with leftovers.
 
 \paragraph*{Framing}
-asserts that the well-typedness of a process is independent of its leftover resources.
+This property asserts that the well-typedness of a process is independent of its leftover resources.
 
@@
   Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-
   Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
@@
 \paragraph*{Weakening}
-states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+This property states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
 Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
@@
 \paragraph*{Strengthening}
-states that removing an unused variable preserves the well-typedness of a process.
+This property states that removing an unused variable preserves the well-typedness of a process.
 Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
@@
 % \begin{remark}
-  The \emph{unusedness} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
+  Notice that the \emph{unused} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
 % \end{remark}
@@
 \paragraph*{Exchange}
-of two variables preserves the well-typedness of a process.
+This property states that the exchange of two variables preserves the well-typedness of a process.
 We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
@@
 
-\begin{proof}[Proof (Sketch)]
+\begin{proof}[Proof]
   All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
@@
 \paragraph*{Subject Congruence}
-states that applying structural congruence rules to a well typed process preserves its well-typedness.
-We must first introduce lemmas that establish that some syntactic manipulations can be inverted and how unused variables relate to the preservation of leftovers.
+This property states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
+To prove this result, we must first introduce lemmas that establish that some syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
 
@@
 
-\begin{proof}[Proof (Sketch)]
+\begin{proof}[Proof]
   The proof is by induction on \textsc{Equals} $\eq{}$.
@@
 \paragraph*{Substitution}
-\autoref{thm:substitution-generalization} proves a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
-\autoref{thm:substitution} instantiates the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
+This result is key to proving subject reduction.
+In \autoref{thm:substitution-generalization} we prove a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
+Then, in \autoref{thm:substitution} we instantiate the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
 
@@
 
-\begin{proof}[Proof (Sketch)]
+\begin{proof}[Proof]
   By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Full details are found in \autoref{app:substitution-generalization}.
+  Full proof can be found in \autoref{app:substitution-generalization}.
   \begin{itemize}
@@
   \label{thm:substitution}
-  Let $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let process $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
   Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
@@
 \end{nitheorem}
-\begin{proof}[Proof (Sketch)]
-  We use framing to derive $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$.
+\begin{proof}[Proof]
+  For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
   Then, we use these to apply \autoref{thm:substitution-generalization}.
@@
 \paragraph*{Subject Reduction}
-states that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
+Finally we are ready to present our main result, stating  that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
+The relation between the typing contexts used to type $P$ and $Q$ will be explained in \autoref{thm:subject-reduction}.
 In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
-We first introduce a helper lemma:
+We first introduce an auxiliary lemma:
 
@@
 
-\begin{proof}[Proof (Sketch)]
-  By induction on $P \reduce{c} Q$.
-  \hfill{}\\
+\begin{proof}[Proof]
+  By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  
   \begin{itemize}

### 2020-09-22 paper/paper.tex
pass on related and conclusion

@@
 In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: they are not inherently well-scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
+Names are cumbersome to mechanise: they are not inherently well scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
 To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
@@
 Named references within the process are substituted with their corresponding de Bruijn index.
-We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such proof term.
+We demand that the original process is well scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such proof term.
 \begin{alignat*}{2}
@@
 He proves type preservation and provides a decision procedure for type checking and type inference.
-In this paper, we follow Allais \cite{Allais2018a} and apply leftover typing to the \picalc{}.
+In this paper, we follow Allais \cite{Allais2018a} and apply leftover typing to the \picalc{} for the first time.
 We generalise the usage algebra, leading to linear, graded and shared type systems.
@@
 Modalities include exact usages, security levels and intervals; resource algebras are pre-ordered semirings with partial addition.
-They provide bidirectional typing rules, and show the type safety of their semantics.
-
-The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types to come along with a mechanised proof of type safety (in Coq).
+The authors provide bidirectional typing rules, and show the type safety of their semantics.
+
+The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types which comes along with a mechanised proof of type safety in Coq.
 The authors extend session types with polymorphism and pattern matching.
 They use a locally-nameless encoding for variable references, a syntax prior to types, and an LTS semantics that encodes session-typed processes into the \picalc{}.
-Their type system uses reordering of contexts and extrinsic context splits.
+Their type system uses reordering of contexts and extrinsic context splits, whether they are not needed in our work.
 
@@
 
-In this paper we provide a well scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
+In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
 We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.

### 2020-09-22 paper/paper.tex
Fix minor errors

@@
   We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs common when dealing with linear type systems.
-  We provide a framing theorem, and generalise the weakening and strengthening theorems, and use them to prove subject congruence.
+  We provide a framing theorem, generalise the weakening and strengthening theorems, and use them to prove subject congruence.
   We show that the type system is stable under substitution and prove subject reduction.
@@
 Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Later on, linearity inspired unique types, as in Clean \cite{BarendsenS96} and ownership types, as in Rust \cite{MatsakisK14}.
-
-Linearity also influenced the $\pi$-calculus.
+Later on, linearity also inspired unique types, as in Clean \cite{BarendsenS96}, and ownership types, as in Rust \cite{MatsakisK14}.
+
+Linearity influenced the $\pi$-calculus as well.
 Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
 %
-Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming, where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, as specified by its session type.
+Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, as specified by its session type.
 Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
@@
 we have two \sender{} processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
-one \receiver{} process, $R$, which receives over channel $z$ the data sent from each of the senders--hence receives twice;
+one \receiver{} process, $R$, which receives over channel $z$ the data sent from each of the senders -- hence receives twice;
 and a \courier{} process $C$, which synchronises communication among $S_1$, $S_2$ and $R$.
@@
 process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
-Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to, for any natural number $n$.
+Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to.
 
@@
     \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown] (send-x) at (0,2)   {$\func{send} \, x$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown] (send-y) at (2,2)   {$\func{send} \, y$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=purple] (recv-z) at (4,2)   {$\func{recv} \, z$};
-    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=cyan,align=center] (sync) at (2,0)   {$\func{carry}$\\$ \, x \, y \, z$};
-
-    \draw[->] (send-x) -- node[left] {$x$} (sync);
-    \draw[->] (send-y) -- node[right] {$y$} (sync);
-    \draw[->] (sync.35) -- node[right] {$z$} (recv-z.240);
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=brown] (send-y) at (0,0)   {$\func{send} \, y$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=purple] (recv-z) at (4,1)   {$\func{recv} \, z$};
+    \node[draw,circle,fill opacity=0.3,text opacity=1,fill=cyan,align=center] (sync) at (2,1)   {$\func{carry}$\\$ \, x \, y \, z$};
+
+    \draw[->] (send-x) -- node[above] {$x$} (sync);
+    \draw[->] (send-y) -- node[below] {$y$} (sync);
+    \draw[->] (sync.10) -- node[above] {$z$} (recv-z.170);
     \draw[->] (sync) -- (recv-z);
@@
 For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
-Each roles \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$ with signature $\Var_n \to \Process_n$.
+Each role \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$.
 
 The \sender{} \emph{role} is defined below.
-The \sender{} creates a new channel $\new{} \suc c$ denoting the data to be sent, and sends it over channel $c$, and then terminates, denoted by process $\PO$.
+The \sender{} creates a new channel to be sent as data, and sends it over channel $c$, and then terminates, denoted by process $\PO$.
 \begin{flalign*}
@@
 We use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$: the \sender{} processes $\func{send} \; x$ and $\func{send} \; y$ each consume $1$ output multiplicity and the \receiver{} process $\func{recv} \; z$ consumes $2$ input multiplicities, as it receives twice on channel $z$.
-Dually, the \courier{} process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$ and one input multiplicity on $y$, and two output multiplicities on $z$.
+Dually, the \courier{} process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$, one input multiplicity on $y$, and two output multiplicities on $z$.
 \end{example}
@@
   \item \textbf{Fully mechanised formalisation and type safety results}:
-    The formalisation of the \picalc{} with leftovers typing, from syntax to semantics and the type system, has been fully mechanised in Agda.
+    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda.
 
@@
 Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
-We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: premises become argument types and conclusions return types.
+We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: inference rules become type constructors, premises become argument types and conclusions return types.
 Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
@@
 The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
+
 Channel names and variables range over $x,y,z$ in $\Name$ and processes over $P,Q,R$ in $\Raw$.
-
 Process $\PO$ denotes the terminated process, where no further communications can occur.
@@
 Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only process constructs which introduce bound names --- $x$ and $y$ in $P$, respectively.
+Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs that introduce bound names --- $x$ and $y$ in $P$, respectively.
 
@@
 Named references within the process are substituted with their corresponding de Bruijn index.
-We demand that the original process is well scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such proof term.
+We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such a proof term.
 \begin{alignat*}{2}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names via $\func{fromRaw}$ results in a well-scoped process.
+  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a well-scoped process.
 \end{nilemma}
@@
 \begin{nilemma}
-  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process with no shadowed variables.
+  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process that follows the \emph{Barendregt convention}.
 \end{nilemma}
@@
 The transitivity rule makes it tricky to prove predicates by induction on \textsc{Equals}: Agda cannot immediately see that such doubly recursive calls terminate.
-In order to remedy this, we model the inductive structure syntactically and index \textsc{Equals} by a type \textsc{Rec}.
+In order to remedy this, we model the recursive call structure syntactically and index \textsc{Equals} by a type \textsc{Rec}.
 

### 2020-09-22 paper/paper.tex
Pass to most of the body

@@
   We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
-  Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
+  Henceforth, we use $\lz$ to denote the multiplicity pair $0 \comma 0$, $\li$ for the pair $1 \comma 0$, $\lo$ for $0 \comma 1$, and $\lio$ for $1 \comma 1$.
   This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
@@
   The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$ (the leftovers).
+  Note that the remaining context $\Gamma$ is preserved unused as a leftover.
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
@@
   \end{lstlisting}
-  The underscore \_ introduces an anonymous declaration immediately followed by its definition.
+  In Agda, the underscore \_ introduces an anonymous declaration immediately followed by its definition.
   We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
@@
   Note that the index $i$ is used in the syntax of the typed process.
-  We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ at index $\constr{0}$.
+  We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ --- at index $\constr{0}$.
   The received element $x$ must be completely exhausted by the continuation process.
@@
   Similarly to input, the output process in rule $\send{}{}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
-  We use the leftover context $\Delta$ to type the continuation process, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan_i$.
+  We use the leftover context $\Delta$ to type the transmitted data, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan_i$.
   The leftovers $\Xi$ are used to type the continuation process.
@@
   Parallel composition in rule $\comp{}{}$ uses the leftovers of the left-hand process to type the right-hand process.
-  By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide an extrinsic proof of context split.
+  By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide a top-down proof of context split.
 \end{nidefinition}
@@
 As per the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
-We will transmit once, thus using a single output multiplicity, and leave the rest as leftovers.
+We will transmit once, thus use a single output multiplicity, and leave the rest as leftovers.
+Agda can uniquely determine the arguments required by the $\new{}$ constructor, and thus we can omit them using underscores.
 \begin{flalign*}
@@
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable not in $P$, such that $\Unused_i \; P$.
-  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
+  Let $i$ be a variable not in $P$, such that $uP \; : \; \Unused_i \; P$.
+  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
 \end{nitheorem}
-
-% \begin{remark}
-  Notice that the \emph{unused} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
-% \end{remark}
 
@@
 This property states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
-To prove this result, we must first introduce lemmas that establish that some syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
+To prove this result, we must first introduce lemmas that establish that certain syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
 

### 2020-09-22 paper/paper.tex
Update LOCs
Keep line numbers within column margins

@@
   \autoref{lst:channels} types a variable reference under a linear algebra.
-  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
+  \begin{lstlisting}[label=lst:channels,xleftmargin=15pt,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
 $\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
@@
 As a result, theorems like framing, weakening and strengthening can now we stated for the linear \picalc{}.
-Our work is fully mechanised in around 1500 lines of code in Agda%
+Our work is fully mechanised in around 1850 lines of code in Agda%
 \begin{anonsuppress}

### 2020-09-22 paper/paper.tex
Clarify Ciccone and Padovani's work

@@
 In their recent paper, Ciccone and Padovani mechanise a dependently-typed linear \picalc{} in Agda \cite{Ciccone}.
-They do so by leveraging Agda's dependent types to provide a dependently-typed interpretation of messages.
-Contexts contain not only types, but also the messages that inhabit their corresponding interpretation in Agda.
-To avoid linearity violations, the interpretation of channel types is erased --- they are only present in the context of the typing judgment.
-They model message input as a dependent function abstraction in Agda, and as a result are able to encode message predicates, branching, and variable-length conversations.
-They use the concrete set of multiplicities $0$, $1$, $\omega$, define a partial algebra on contexts and require user-provided proofs of context split.
+Their intrinsic encoding allows them to leverage Agda's dependent types to provide a dependently-typed interpretation of messages --- to avoid linearity violations the interpretation of channel types is erased.
+Message input is modeled as a dependent function in Agda, and as a result message predicates, branching, and variable-length conversations can be encoded.
+In contrast to our work, their algebra is on the multiplicities $0$, $1$, $\omega$, and top-down context splitting proofs must be provided.
 

### 2020-09-22 paper/paper.tex
Add own work

@@
 
+In her master's thesis, Zalakain uses a HOAS to embed a session-typed \picalc{} in Coq \cite{Zalakain2019}.
+Channel type polymorphism (preventing channels from being constructed out of thin air) and a process-traversing predicate ensure that channels are used linearly --- subject reduction is then shown to preserve this predicate.
+
 \paragraph{Other Work}

### 2020-09-22 paper/paper.tex
Anon supress previous work

@@
 
-The authors would also like to thank the anonymous referees fortheir valuable comments and helpful suggestions.
+The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
 This work is supported by the \grantsponsor{UKEPSRC}{UK EPSRC}{} grant \grantnum{UKEPSRC}{EP/K034413/1} ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the \grantsponsor{EUHORIZON}{EU HORIZON 2020 MSCA RISE}{} project \grantnum{EUHORIZON}{778233} ``Behavioural Application Program Interfaces'' (BehAPI).
@@
 
+\begin{anonsuppress}
 In her master's thesis, Zalakain uses a HOAS to embed a session-typed \picalc{} in Coq \cite{Zalakain2019}.
 Channel type polymorphism (preventing channels from being constructed out of thin air) and a process-traversing predicate ensure that channels are used linearly --- subject reduction is then shown to preserve this predicate.
+\end{anonsuppress}
 

### 2020-11-30 paper/paper.tex
Apply LIPIcs

--- 2020-09-22 paper/paper.tex
@@
-\documentclass[sigplan,10pt,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[]{lipics-v2019}
 
@@
 \newtheorem{nilemma}{Lemma}
-\newtheorem{example}{Example}
+% \newtheorem{example}{Example}
 \renewcommand{\nidefinitionautorefname}{Definition}%
@@
 \renewcommand{\subsectionautorefname}{\S}%
-\renewcommand{\exampleautorefname}{Example}
+% \renewcommand{\exampleautorefname}{Example}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title[$\pi$ with leftovers: a mechanisation in Agda]{$\pi$ with leftovers: \\ a mechanisation in Agda}
-\author{Uma Zalakain}
-\affiliation{University of Glasgow, Scotland}
-\email{u.zalakain.1@research.gla.ac.uk}
-\orcid{https://orcid.org/0000-0002-3268-9338}{}
-\author{Ornela Dardha}
-\affiliation{University of Glasgow, Scotland}
-\email{ornela.dardha@glasgow.ac.uk}
-\orcid{https://orcid.org/0000-0001-9927-7875}{}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
+\author{Uma Zalakain}{University of Glasgow, Scotland}{u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
+\author{Ornela Dardha}{University of Glasgow, Scotland}{ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
 \keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
-\setcopyright{acmlicensed}
+\Copyright{U. Zalakain, and O. Dardha}
 \ccsdesc[300]{Theory of computation~Process calculi}
@@
 
-\begin{acks}
-We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.
-
-The authors would also like to thank the anonymous referees fortheir valuable comments and helpful suggestions.
-This work is supported by the \grantsponsor{UKEPSRC}{UK EPSRC}{} grant \grantnum{UKEPSRC}{EP/K034413/1} ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the \grantsponsor{EUHORIZON}{EU HORIZON 2020 MSCA RISE}{} project \grantnum{EUHORIZON}{778233} ``Behavioural Application Program Interfaces'' (BehAPI).
-\end{acks}
+\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
+
+\funding{The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
+This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
+
+\maketitle
 
@@
 \end{abstract}
-
-\maketitle
 
@@
   \item \textbf{Fully mechanised formalisation and type safety results}:
-    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda.
+    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
 
@@
 \end{enumerate}
-\begin{anonsuppress}
-and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-\end{anonsuppress}
 
@@
 
-\paragraph{From names to de Bruijn indices}
+\paragraph*{From names to de Bruijn indices}
 When we translate into de Bruijn indices we keep the original binder names around --- they will serve as name hints for when we translate back.
@@
 
-\paragraph{From de Bruijn indices to names}
+\paragraph*{From de Bruijn indices to names}
 The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
@@
   All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
-  Complete details can be found in our mechanisation in Agda%
-\begin{anonsuppress}
-in \cite{Zalakain2020Agda}
-\end{anonsuppress}
-.
+  Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
 \end{proof}
@@
   All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
-  For details, refer to our mechanisation in Agda%
-\begin{anonsuppress}
-\cite{Zalakain2020Agda}
-\end{anonsuppress}
-.
+  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
 \end{proof}
@@
   Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
-\begin{anonsuppress}
   Full proof in \cite{Zalakain2020Agda}.
-\end{anonsuppress}
   We proceed by induction on \textsc{StructCong} $\eqeq$:
@@
 
-\paragraph{Extrinsic Encodings}
+\paragraph*{Extrinsic Encodings}
 
@@
 
-\paragraph{Intrinsic Encodings}
+\paragraph*{Intrinsic Encodings}
 
@@
 
-\paragraph{Other Work}
+\paragraph*{Other Work}
 
@@
 As a result, theorems like framing, weakening and strengthening can now we stated for the linear \picalc{}.
-Our work is fully mechanised in around 1850 lines of code in Agda%
-\begin{anonsuppress}
-\cite{Zalakain2020Agda}
-\end{anonsuppress}
-.
+Our work is fully mechanised in around 1850 lines of code in Agda \cite{Zalakain2020Agda}.
 

### 2020-12-03 paper/appendix.tex
a pass on technical parts
Start working on the CPP reviews

@@
 %% Appendix
+
+\section{From names to de Bruijn indices and back}
+\label{from_to_deBruijn}
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    { }
+    {\Raw : \Set}
+    \; \textsc{Raw}
+  \end{mathpar}
+
+  \begin{equation*}
+    \begin{aligned}
+      \Raw ::= &\; \PO              &&\text{(inaction)}    \\ 
+      |& \; (\new{\Name}) \; \Raw         &&\text{(restriction)} \\ 
+      |& \; \comp{\Raw}{\Raw}       &&\text{(parallel)}    \\ 
+      |& \; \recv{\Name}{\Name}\Raw &&\text{(input)}       \\ 
+      |& \; \send{\Name}{\Name}\Raw &&\text{(output)}      \\
+    \end{aligned}
+  \end{equation*}
+  \caption{Grammar with names.}
+  \label{fig:syntax-names}
+\end{figure}
+
+The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
+
+Channel names and variables range over $x,y,z$ in $\Name$ and processes over $P,Q,R$ in $\Raw$.
+Process $\PO$ denotes the terminated process, where no further communications can occur.
+Process $(\new{}x)\; P$ creates a new channel $x$ bound with scope $P$.
+Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
+Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs that introduce bound names --- $x$ and $y$ in $P$, respectively.
+
+In order to demonstrate the correspondence between an \picalc{} that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
+
+\paragraph*{From names to de Bruijn indices}
+When we translate into de Bruijn indices we keep the original binder names around --- they will serve as name hints for when we translate back.
+The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
+Named references within the process are substituted with their corresponding de Bruijn index.
+We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such a proof term.
+\begin{alignat*}{2}
+    &\func{fromRaw} && : (ctx : \Names_n) \; (P : \Raw) \\
+    &               && \to \WellScoped \; ctx \; P \to \Process_n
+\end{alignat*}
+
+\paragraph*{From de Bruijn indices to names}
+The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
+As some widely-used languages do, this translation function produces unique variable names.
+These unique variable names use the naming scheme $<name-hint>^{<n>}$, where $^{<n>}$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
+\begin{equation*}
+    \func{toRaw} : (ctx : \Names_n) \to \Process_n \to \Raw
+\end{equation*}
+
+\begin{example}[$\func{fromRaw}$ and $\func{toRaw}$]
+  \begin{figure}[t]
+    \begin{alignat*}{7}
+      &P = (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
+      &Q = \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
+      &R = (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)})
+    \end{alignat*}
+    \caption{From names to de Bruijn indices and back}
+    \label{fig:conversion}
+    \end{figure}
+  We illustrate the conversion functions from names to de Bruijn indices $(\func{fromRaw}$) and back ($\func{toRaw}$) with three processes $P,Q,R$  in \autoref{fig:conversion}.
+
+  Process $P$ uses names $x,y,z$ and is translated via the conversion function $\func{fromRaw}$ into process $Q$, which uses de Bruijn indices.
+  Process $Q$ is then translated via $\func{toRaw}$ into process $R$, which follows the \emph{Barendregt convention}\footnote{The Barendregt variable convention states that all bound variables/names in a process are distinct among each other and from the free variables/names.} and is $\alpha$-equivalent to the original process $P$.
+\end{example}
+
+In the following we present the main results that our conversion functions satisfy.
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a well-scoped process.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process that follows the \emph{Barendregt convention}.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from de Bruijn indices to names and back via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
+\end{nilemma}
+
+\begin{nilemma}
+  Translating from names to de Bruijn indices and back via $\func{toRaw} \circ \func{fromRaw}$ results in the same process modulo $\alpha$-conversion.
+\end{nilemma}
+
+\begin{proof}[Proof]
+  All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
+  Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
+\end{proof}
+
 \section{Definitions}
@@
 
-\begin{proof}[Proof (Sketch)]
+\begin{proof}[Proof]
   By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.

### 2020-12-03 paper/paper.tex
Start working on the CPP reviews

@@
 \usepackage{hyperref}
+\usepackage{todonotes}
 
@@
 
-The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
-It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
+The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
 Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
@@
 
+\todo{introduce shared and graded types sooner}
+\todo{too much on session types}
 Linearity influenced the $\pi$-calculus as well.
@@
 
-In this paper we present \emph{the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}
-
-While linear types are useful for resource-aware programming, graded and shared types are useful for a more flexible and 
-mainstream programming.
-
-We define a specification for an algebra on \emph{multiplicities} --- how many times a channel can be used --- and use it to apply \emph{leftover typing} for the first time to the \picalc{}, along the lines of Allais \cite{Allais2018a}.
+In this paper we make the following contributions:
+
+\todo{complete}
+\begin{enumerate}
+  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+
+  \item We \textbf{abstract our type system over a set of resource-aware usage algebras}:
     The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
     This allows for linear, graded and shared types to be mixed in the same type system.
-Ultimately, we can exploit this formalisation as a unified framework for type-safe distributed modelling and programming using a variety of type systems, and build on top of it more advanced types, theories and languages.
+    While linear types are useful for resource-aware programming, graded and shared types are useful for a more flexible and mainstream programming.
+
+  \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system} along the lines of Allais \cite{Allais2018a}.
+\end{enumerate}
+
+\subsection{Summary}
+\todo{shorten}
+\begin{enumerate}
+\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
+ \begin{itemize}
+   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+   For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
+   These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
+
+   \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
+   Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
+   The reduction relation tracks at the type level the channel on which communication occurs.
+This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
+   The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+ \end{itemize}
+
+  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
+  we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
+  \begin{itemize}
+    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
+    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
+
+    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
+    This allows for type systems combining linear, graded and shared types under the same framework.
+
+    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+    This approach adds a leftover usage context to the typing judgements.
+    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
+    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
+  \end{itemize}
+
+  \item \textbf{Fully mechanised formalisation and type safety results}:
+    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
+
+    We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
+    We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
+    We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
+\end{enumerate}
+
+\subsection{Notation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    { }
+    {\type{\N} : \Set}
+    \; \textsc{Nat}
+
+    \inferrule
+    { }
+    {\constr{0} : \type{\N}}
+
+    \inferrule
+    {n : \type{\N}}
+    {\suc n : \type{\N}}
+  \end{mathpar}
+  \caption{Notation used in this paper.}
+  \label{fig:notation}
+\end{figure}
+
+\autoref{fig:notation} illustrates the notation used in this paper.
+Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.
+Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
+We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: inference rules become type constructors, premises become argument types and conclusions return types.
+Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
+Implicit arguments are mentioned in type definitions but omitted by constructors.
+
+We use colours to further distinguish the different entities in this paper.
+$\type{TYPES}$ are blue and uppercased, with indices as subscripts, $\constr{constructors}$ are orange, $\func{functions}$ are teal, variables are black, and some constructor names are overloaded --- and disambiguated by context.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Syntax}
+\label{syntax}
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \datatype
+    {n : \N}
+    {\Var_n : \Set}
+    \; \textsc{Var}
+
+    \inferrule
+    {n : \N}
+    {\constr{0} : \Var_{\suc n}}
+
+    \inferrule
+    {x : \Var_n}
+    {\suc x : \Var_{\suc n}}
+
+    \datatype
+    {n : \N}
+    {\Process_n : \Set}
+    \; \textsc{Process}
+  \end{mathpar}
+
+  \begin{equation*}
+    \begin{aligned}
+      \Process_n ::=& \; \PO_n               &&\text{(inaction)}   \\ 
+      |& \; \new{} \Process_{\suc n}         &&\text{(restriction)}          \\ 
+      |& \; \comp{\Process_n}{\Process_n}    &&\text{(parallel)} \\ 
+      |& \; \recv{\Var_n}{}\Process_{\suc n} &&\text{(input)}                \\ 
+      |& \; \send{\Var_n}{\Var_n}\Process_n  &&\text{(output)}                  
+    \end{aligned}
+  \end{equation*}
+  \caption{Well-scoped grammar with type-level de Bruijn indices.}
+  \label{fig:syntax}
+\end{figure}
+
+In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
+Names are cumbersome to mechanise: they are not inherently well scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
+To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
+That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
+
+A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
+References outside of that range are meaningless.
+It is useful to rule out these ill-scoped terms syntactically.
+In \autoref{fig:syntax} --- which presents the syntax of the \picalc{} using de Bruijn indices --- we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
+
+\todo{adapt to de bruijn}
+Process $\PO$ denotes the terminated process, where no further communications can occur.
+Process $\new{}\; P$ creates a new channel bound with scope $P$.
+Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
+Processes $\recv{x}{} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
+Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs that introduce bound names --- $x$ and $y$ in $P$, respectively.
 
@@
 \end{example}
-
-\subsection{Contributions}
-\begin{enumerate}
-\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
- \begin{itemize}
-   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-   For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
-   These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
-
-   \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
-   Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel on which communication occurs.
-This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
-   The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
- \end{itemize}
-
-  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
-  we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
-  \begin{itemize}
-    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
-
-    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, graded and shared types under the same framework.
-
-    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    This approach adds a leftover usage context to the typing judgements.
-    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
-    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
-  \end{itemize}
-
-  \item \textbf{Fully mechanised formalisation and type safety results}:
-    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-
-    We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
-    We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
-    We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
-\end{enumerate}
-
-\subsection{Notation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    { }
-    {\type{\N} : \Set}
-    \; \textsc{Nat}
-
-    \inferrule
-    { }
-    {\constr{0} : \type{\N}}
-
-    \inferrule
-    {n : \type{\N}}
-    {\suc n : \type{\N}}
-  \end{mathpar}
-  \caption{Notation used in this paper.}
-  \label{fig:notation}
-\end{figure}
-
-\autoref{fig:notation} illustrates the notation used in this paper.
-Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.
-Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
-We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: inference rules become type constructors, premises become argument types and conclusions return types.
-Universe levels and universe polymorphism are omitted for brevity --- all our types are of type $\Set$.
-Implicit arguments are mentioned in type definitions but omitted by constructors.
-
-We use colours to further distinguish the different entities in this paper.
-$\type{TYPES}$ are blue and uppercased, with indices as subscripts, $\constr{constructors}$ are orange, $\func{functions}$ are teal, variables are black, and some constructor names are overloaded --- and disambiguated by context.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Syntax}
-\label{syntax}
-
-\begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    { }
-    {\Raw : \Set}
-    \; \textsc{Raw}
-  \end{mathpar}
-
-  \begin{equation*}
-    \begin{aligned}
-      \Raw ::= &\; \PO              &&\text{(inaction)}    \\ 
-      |& \; (\new{\Name}) \; \Raw         &&\text{(restriction)} \\ 
-      |& \; \comp{\Raw}{\Raw}       &&\text{(parallel)}    \\ 
-      |& \; \recv{\Name}{\Name}\Raw &&\text{(input)}       \\ 
-      |& \; \send{\Name}{\Name}\Raw &&\text{(output)}      \\
-    \end{aligned}
-  \end{equation*}
-  \caption{Grammar with names.}
-  \label{fig:syntax-names}
-\end{figure}
-
-\begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    {n : \N}
-    {\Var_n : \Set}
-    \; \textsc{Var}
-
-    \inferrule
-    {n : \N}
-    {\constr{0} : \Var_{\suc n}}
-
-    \inferrule
-    {x : \Var_n}
-    {\suc x : \Var_{\suc n}}
-
-    \datatype
-    {n : \N}
-    {\Process_n : \Set}
-    \; \textsc{Process}
-  \end{mathpar}
-
-  \begin{equation*}
-    \begin{aligned}
-      \Process_n ::=& \; \PO_n               &&\text{(inaction)}   \\ 
-      |& \; \new{} \Process_{\suc n}         &&\text{(restriction)}          \\ 
-      |& \; \comp{\Process_n}{\Process_n}    &&\text{(parallel)} \\ 
-      |& \; \recv{\Var_n}{}\Process_{\suc n} &&\text{(input)}                \\ 
-      |& \; \send{\Var_n}{\Var_n}\Process_n  &&\text{(output)}                  
-    \end{aligned}
-  \end{equation*}
-  \caption{Well-scoped grammar with type-level de Bruijn indices.}
-  \label{fig:syntax}
-\end{figure}
-
-The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
-
-Channel names and variables range over $x,y,z$ in $\Name$ and processes over $P,Q,R$ in $\Raw$.
-Process $\PO$ denotes the terminated process, where no further communications can occur.
-Process $(\new{}x)\; P$ creates a new channel $x$ bound with scope $P$.
-Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recv{x}{y} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs that introduce bound names --- $x$ and $y$ in $P$, respectively.
-
-In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: they are not inherently well scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
-To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
-That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
-
-A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
-References outside of that range are meaningless.
-It is useful to rule out these ill-scoped terms syntactically.
-In \autoref{fig:syntax} --- which presents the syntax of the \picalc{} using de Bruijn indices --- we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
-We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
-
-\subsection{From names to de Bruijn indices and back}
-\label{from_to_deBruijn}
-In order to demonstrate the correspondence between an \picalc{} that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
-
-\paragraph*{From names to de Bruijn indices}
-When we translate into de Bruijn indices we keep the original binder names around --- they will serve as name hints for when we translate back.
-The translation function $\func{fromRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
-Named references within the process are substituted with their corresponding de Bruijn index.
-We demand that the original process is well-scoped: that all its free variable names appear in $ctx$ --- this is decidable and we therefore automate the construction of such a proof term.
-\begin{alignat*}{2}
-    &\func{fromRaw} && : (ctx : \Names_n) \; (P : \Raw) \\
-    &               && \to \WellScoped \; ctx \; P \to \Process_n
-\end{alignat*}
-
-\paragraph*{From de Bruijn indices to names}
-The translation function $\func{toRaw}$ works recursively, keeping a context $ctx : \Names_n$ that maps the first $n$ indices to their names.
-As some widely-used languages do, this translation function produces unique variable names.
-These unique variable names use the naming scheme $<name-hint>^{<n>}$, where $^{<n>}$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
-\begin{equation*}
-    \func{toRaw} : (ctx : \Names_n) \to \Process_n \to \Raw
-\end{equation*}
-
-\begin{example}[$\func{fromRaw}$ and $\func{toRaw}$]
-  \begin{figure}[t]
-    \begin{alignat*}{7}
-      &P = (\new{x} ) && (\comp {\recv{x}{x} && \send{x}{z} && \PO} {(\new{y}) && (\send{x}{y} && \recv{y}{y} && \PO)}) \\
-      &Q = \new{} && (\comp {\recv{0}{} && \send{0}{2} && \PO} {\new{} && (\send{1}{0} && \recv{0}{} && \PO)}) \\
-      &R = (\new{x^0} ) && (\comp {\recv{x^0}{x^1} && \send{x^1}{z^0} && \PO} {(\new{y^0}) && (\send{x^0}{y^0} && \recv{y^0}{y^1} && \PO)})
-    \end{alignat*}
-    \caption{From names to de Bruijn indices and back}
-    \label{fig:conversion}
-    \end{figure}
-  We illustrate the conversion functions from names to de Bruijn indices $(\func{fromRaw}$) and back ($\func{toRaw}$) with three processes $P,Q,R$  in \autoref{fig:conversion}.
-
-  Process $P$ uses names $x,y,z$ and is translated via the conversion function $\func{fromRaw}$ into process $Q$, which uses de Bruijn indices.
-  Process $Q$ is then translated via $\func{toRaw}$ into process $R$, which follows the \emph{Barendregt convention}\footnote{The Barendregt variable convention states that all bound variables/names in a process are distinct among each other and from the free variables/names.} and is $\alpha$-equivalent to the original process $P$.
-\end{example}
-
-In the following we present the main results that our conversion functions satisfy.
-
-\begin{nilemma}
-  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a well-scoped process.
-\end{nilemma}
-
-\begin{nilemma}
-  Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process that follows the \emph{Barendregt convention}.
-\end{nilemma}
-
-\begin{nilemma}
-  Translating from de Bruijn indices to names and back via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
-\end{nilemma}
-
-\begin{nilemma}
-  Translating from names to de Bruijn indices and back via $\func{toRaw} \circ \func{fromRaw}$ results in the same process modulo $\alpha$-conversion.
-\end{nilemma}
-
-\begin{proof}[Proof]
-  All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
-  Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
-\end{proof}
 
@@
 They use a locally-nameless encoding for variable references, a syntax prior to types, and an LTS semantics that encodes session-typed processes into the \picalc{}.
-Their type system uses reordering of contexts and extrinsic context splits, whether they are not needed in our work.
+Their type system uses reordering of contexts and extrinsic context splits, which are not needed in our work.
 
@@
 
-In her master's thesis, Zalakain uses a HOAS to embed a session-typed \picalc{} in Coq \cite{Zalakain2019}.
-Channel type polymorphism (preventing channels from being constructed out of thin air) and a process-traversing predicate ensure that channels are used linearly --- subject reduction is then shown to preserve this predicate.
-
 \paragraph*{Other Work}
 
-Castro et al.
-\cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq  --- in Coq de Bruijn indices are not as popular as in Agda or Idris because dependent pattern matching is far from straightforward.
+Castro et al. \cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq, where de Bruijn indices are not as popular as in Agda or Idris.
 As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
@@
 We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.
-As a result, theorems like framing, weakening and strengthening can now we stated for the linear \picalc{}.
+As a result, theorems like framing, weakening and strengthening can now be stated for the linear \picalc{}.
 Our work is fully mechanised in around 1850 lines of code in Agda \cite{Zalakain2020Agda}.

### 2020-12-08 paper/paper.tex
Take focus away of session types

@@
 
-We live in a concurrent and distributed world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
-As humans, we aim to model, build, and predict such interactive systems; as mathematicians, abstraction is our tool of choice to do so.
-
 The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
@@
 The \picalc{} is at the basis of the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and more recently Go \footnote{https://golang.org}.
-At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties, such as deadlock freedom, livelock freedom or termination, and most notably, session types.
-These make the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
+At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, session types, and types for liveness properties, such as deadlock freedom, livelock freedom or termination.
+These make of the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
 
@@
 \todo{introduce shared and graded types sooner}
-\todo{too much on session types}
+
 Linearity influenced the $\pi$-calculus as well.
@@
 %
-Meanwhile, linearity acquired a different flavour with the rise of \emph{session types} \cite{H93,THK94,HVK98}, a formalism used in communication-centric programming where linearity ensures that a channel is owned by exactly one communicating participant --- the channel itself may be used multiple times, as specified by its session type.
+With the rise of session types \cite{H93,THK94,HVK98} linearity acquired a different flavour: linearity ensures that each channel endpoint is owned by exactly one communicating participant, even though the channel itself may be used multiple times, as specified by its session type.
 Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
@@
 In recent years, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} \cite{KPT96} into the spotlight again.
-This encoding not only has theoretical benefits in terms of the expressivity of session types and the reusability of theoretical results from the linear \picalc{}, but is useful in practice too.
 The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, thus allowing to use the target linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be defined and implemented.
+\todo{go back to linear \picalc{}}
 

### 2020-12-08 paper/paper.tex
Expand a bit on contributions

@@
 In this paper we make the following contributions:
-
-\todo{complete}
 \begin{enumerate}
-  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-
-  \item We \textbf{abstract our type system over a set of resource-aware usage algebras}:
+  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
+    Every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+    \todo{Do Ciccone \& Padovani also use type level de Bruijn indices?}
+
+  \item We \textbf{parametrise our type system over a set of resource-aware usage algebras}:
+    A \emph{usage algebra} is defined as a ternary relation that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative} and \emph{commutative}.
     The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
     This allows for linear, graded and shared types to be mixed in the same type system.
-    While linear types are useful for resource-aware programming, graded and shared types are useful for a more flexible and mainstream programming.
-
-  \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system} along the lines of Allais \cite{Allais2018a}.
+
+  \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system}:
+    \todo{along the lines of Allais \cite{Allais2018a}.
 \end{enumerate}

### 2020-12-09 paper/paper.tex
Include definition of dec

@@
   \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system}:
-    \todo{along the lines of Allais \cite{Allais2018a}.
+    \todo{along the lines of Allais \cite{Allais2018a}}.
 \end{enumerate}
@@
 
-We lift the relation \textsc{StrucCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$ as shown in \autoref{fig:struct-cong1}.
+We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$ as shown in \autoref{fig:struct-cong1}.
 The $\constr{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively scope restriction, parallel composition, input and output.
@@
 --- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
+
+\begin{figure}[h]
+\begin{mathpar}
+  \datatype
+  {n : \N}
+  {\Channel_n : \Set}
+  \; \textsc{Channel}
+
+  \inferrule
+  { }
+  {\constr{internal} : \Channel_n}
+
+  \inferrule
+  {i : \Var_n}
+  {\constr{external} \; i : \Channel_n}
+
+  \datatype
+  {c : \Channel_n \\ P \; Q : \Process_n}
+  {P \reduce{c} Q : \Set}
+  \; \textsc{Reduces}
+
+  \inferrule
+  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
+  {\constr{comm} : \comp{\recv{i}{}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
+
+  \inferrule
+  {red : P \reduce{c} P'}
+  {\constr{par} \; red : \comp{P}{Q} \reduce{c} \comp{P'}{Q}}
+
+  \inferrule
+  {red : P \reduce{c} Q}
+  {\constr{res} \; red : \new P \reduce{\func{dec}\; c} \new Q}
+
+  \inferrule
+  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
+  {\constr{struct} \; eq \; red : P \reduce{c} Q}
+\end{mathpar}
+\caption{Operational semantics indexed by the channel over which reduction occurs.}
+\label{fig:reduction}
+\end{figure}
+
+Rule $\constr{res}$ uses $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
+It is defined as follows:
+\begin{equation*}
+  \begin{aligned}
+    & \func{dec} : \Channel_(\suc n) && \to \Channel_n \\
+    & \func{dec} \; \constr{internal}                 && \hookrightarrow \constr{internal} \\
+    & \func{dec} \; (\constr{external} \; \constr{0}) && \hookrightarrow \constr{internal} \\
+    & \func{dec} \; (\constr{external} \; (\suc n))   && \hookrightarrow \constr{external} \; n \\
+  \end{aligned}
+\end{equation*}
+\todo{introduce $\hookrightarrow$ notation?}
 
@@
 \end{proof}
-
-Rule $\constr{res}$ uses $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
-This function saturates at $\constr{internal}$: wrapping an internally reducing process with a binder will result in an internally reducing process.
-
-\begin{figure}[h]
-\begin{mathpar}
-  \datatype
-  {n : \N}
-  {\Channel_n : \Set}
-  \; \textsc{Channel}
-
-  \inferrule
-  { }
-  {\constr{internal} : \Channel_n}
-
-  \inferrule
-  {i : \Var_n}
-  {\constr{external} \; i : \Channel_n}
-
-  \datatype
-  {c : \Channel_n \\ P \; Q : \Process_n}
-  {P \reduce{c} Q : \Set}
-  \; \textsc{Reduces}
-
-  \inferrule
-  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}{}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
-
-  \inferrule
-  {red : P \reduce{c} P'}
-  {\constr{par} \; red : \comp{P}{Q} \reduce{c} \comp{P'}{Q}}
-
-  \inferrule
-  {red : P \reduce{c} Q}
-  {\constr{res} \; red : \new P \reduce{\func{dec}\; c} \new Q}
-
-  \inferrule
-  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
-  {\constr{struct} \; eq \; red : P \reduce{c} Q}
-\end{mathpar}
-\caption{Operational semantics indexed by the channel over which reduction occurs.}
-\label{fig:reduction}
-\end{figure}
 

### 2020-12-09 paper/paper.tex
Bring the def of Unused into the paper

@@
 The last three ($\constr{scope}-$*) state garbage collection, scope extrusion and commutativity of restrictions, respectively.
-In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$ (see \autoref{app:definitions}).
+In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$ (see \autoref{def:unused}).
 The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
@@
 In all the above, $i$ is incremented every time we go under a binder.
+
+\begin{nidefinition}
+  \label{def:unused}
+  A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
+  \todo{should we have this as a figure instead?}
+
+  \begin{mathpar}
+    \datatype
+    {i : \Var_n \\ P : \Process_n}
+    {\Unused_i \; P}
+    \; \textsc{Unused}
+
+    \inferrule
+    { }
+    {\Unused_i \; \PO}
+
+    \inferrule
+    {\Unused_{\suc i} \; P}
+    {\Unused_i \; \new{}P}
+
+    \inferrule
+    {\Unused_i \; P \\ \Unused_i \; Q}
+    {\Unused_i \; \comp{P}{Q}}
+
+    \inferrule
+    {i \type{\not\equiv} x \\ \Unused_{\suc i} \; P}
+    {\Unused_i \; \recv{x}{} P}
+
+    \inferrule
+    {i \type{\not\equiv} x \\ i \type{\not\equiv} y \\ \Unused_i \; P}
+    {\Unused_i \; \send{x}{y} P}
+  \end{mathpar}
+
+  The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
+\end{nidefinition}
 

### 2020-12-14 paper/appendix.tex
Adapt to LLNCS style

@@
 
-\begin{nilemma}
+\begin{lemma}
   Translating from de Bruijn indices to names via $\func{toRaw}$ results in a well-scoped process.
-\end{nilemma}
+\end{lemma}
 
-\begin{nilemma}
+\begin{lemma}
   Translating from de Bruijn indices to names via $\func{toRaw}$ results in a process that follows the \emph{Barendregt convention}.
-\end{nilemma}
+\end{lemma}
 
-\begin{nilemma}
+\begin{lemma}
   Translating from de Bruijn indices to names and back via $\func{fromRaw} \circ \func{toRaw}$ results in the same process modulo internal variable name hints.
-\end{nilemma}
+\end{lemma}
 
-\begin{nilemma}
+\begin{lemma}
   Translating from names to de Bruijn indices and back via $\func{toRaw} \circ \func{fromRaw}$ results in the same process modulo $\alpha$-conversion.
-\end{nilemma}
+\end{lemma}
 
@@
 
-\section{Definitions}
-\label{app:definitions}
-
-\begin{nidefinition}
-  A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
-
-  \begin{mathpar}
-    \datatype
-    {i : \Var_n \\ P : \Process_n}
-    {\Unused_i \; P}
-    \; \textsc{Unused}
-
-    \inferrule
-    { }
-    {\Unused_i \; \PO}
-
-    \inferrule
-    {\Unused_{\suc i} \; P}
-    {\Unused_i \; \new{}P}
-
-    \inferrule
-    {\Unused_i \; P \\ \Unused_i \; Q}
-    {\Unused_i \; \comp{P}{Q}}
-
-    \inferrule
-    {i \type{\not\equiv} x \\ \Unused_{\suc i} \; P}
-    {\Unused_i \; \recv{x}{} P}
-
-    \inferrule
-    {i \type{\not\equiv} x \\ i \type{\not\equiv} y \\ \Unused_i \; P}
-    {\Unused_i \; \send{x}{y} P}
-  \end{mathpar}
-
-  The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
-\end{nidefinition}
-
 \section{Substitution}
@@
 
-\begin{nitheorem}[Generalised substitution]
+\begin{theorem}[Generalised substitution]
   \label{thm:subst-generalization1}
@@
   Refer to \autoref{fig:subst} for a diagramatic representation.
-\end{nitheorem}
+\end{theorem}
 

### 2020-12-14 paper/paper.tex
Remove the index from process termination
Adapt to LLNCS style

@@
-\documentclass[]{lipics-v2019}
+\documentclass[]{llncs}
 
@@
 % Do not use italics in definitions and theorems
-\theoremstyle{definition}
-\newtheorem{nidefinition}{Definition}
-\newtheorem{nitheorem}{Theorem}
-\newtheorem{nilemma}{Lemma}
-% \newtheorem{example}{Example}
-\renewcommand{\nidefinitionautorefname}{Definition}%
-\renewcommand{\nitheoremautorefname}{Theorem}%
-\renewcommand{\nilemmaautorefname}{Lemma}%
-\renewcommand{\sectionautorefname}{\S}%
-\renewcommand{\subsectionautorefname}{\S}%
+% \theoremstyle{definition}
+% \newtheorem{definition}{Definition}
+% \newtheorem{theorem}{Theorem}
+% \newtheorem{lemma}{Lemma}
+% % \newtheorem{example}{Example}
+% \renewcommand{\definitionautorefname}{Definition}%
+% \renewcommand{\theoremautorefname}{Theorem}%
+% \renewcommand{\lemmaautorefname}{Lemma}%
+% \renewcommand{\sectionautorefname}{\S}%
+% \renewcommand{\subsectionautorefname}{\S}%
 % \renewcommand{\exampleautorefname}{Example}
@@
 \newcommand{\lo}{\func{\ell_o}}
-\newcommand{\lz}{\func{\ell_{\varnothing}}}
+\newcommand{\lz}{\func{\ell_{\varepsilon}}}
 \newcommand{\lio}{\func{\ell_{\#}}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\title{
+  $\pi$ with leftovers: \\ a mechanisation in Agda
+  \thanks{The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
+This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
+}
 \titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
-\author{Uma Zalakain}{University of Glasgow, Scotland}{u.zalakain.1@research.gla.ac.uk}{https://orcid.org/0000-0002-3268-9338}{}
-\author{Ornela Dardha}{University of Glasgow, Scotland}{ornela.dardha@glasgow.ac.uk}{https://orcid.org/0000-0001-9927-7875}{}
-\keywords{pi-calculus, linear types, leftover typing, concurrency, mechanisation, Agda}
-\Copyright{U. Zalakain, and O. Dardha}
-\ccsdesc[300]{Theory of computation~Process calculi}
-% \supplement{\url{https://github.com/umazalakain/typing-linear-pi}}
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10003752.10003753.10003761.10003764</concept_id>
-<concept_desc>Theory of computation~Process calculi</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
+\author{
+  Uma Zalakain\inst{1}\orcidID{0000-0002-3268-9338}
+  \and
+  Ornela Dardha\inst{2}\orcidID{0000-0001-9927-7875}
+}
+\institute{
+  University of Glasgow, Scotland \\ \email{u.zalakain.1@research.gla.ac.uk}
+  \and
+  University of Glasgow, Scotland \\ \email{ornela.dardha@glasgow.ac.uk}
+}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\acknowledgements{We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.}
-
-\funding{The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
-This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
-
 \maketitle
@@
 \begin{abstract}
+  \todo{the only references here are to session types}
+  \todo{fix acknowledgments}
+  \todo{fix lemma, definition, theorem, example envs}
   The \picalc{} is a computational model for communication and concurrency.
@@
   Our formalisation is fully mechanised in Agda.
+
+  \keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
 \end{abstract}
@@
 \begin{enumerate}
-  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
+  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
     Every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
@@
  \begin{itemize}
-   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972, Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
    For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
@@
     \begin{aligned}
-      \Process_n ::=& \; \PO_n               &&\text{(inaction)}   \\ 
+      \Process_n ::=& \; \PO                &&\text{(inaction)}   \\ 
       |& \; \new{} \Process_{\suc n}         &&\text{(restriction)}          \\ 
@@
 
-\begin{nidefinition}
+\begin{definition}
   \label{def:unused}
@@
   The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
-\end{nidefinition}
+\end{definition}
 
@@
 
-\begin{nilemma}
+\begin{lemma}
   \label{lm:substitution-unused}
   For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
-\end{nilemma}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{nidefinition}[Usage algebra]
+\begin{definition}[Usage algebra]
   A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
@@
   \end{figure}
-\end{nidefinition}
+\end{definition}
 
@@
 We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
-\begin{nidefinition}[Indexed set of usage algebras]
+\begin{definition}[Indexed set of usage algebras]
   An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type (\autoref{fig:indexed-multiplicities}).
@@
 
-\begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
+\begin{definition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
   A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two rows of \autoref{fig:prectx}).
@@
 A \emph{typing context} (last two rows of \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
-\end{nidefinition}
-
-\begin{nidefinition}[\textsc{Ctx}: usage contexts]
+\end{definition}
+
+\begin{definition}[\textsc{Ctx}: usage contexts]
   A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
@@
 
-\begin{nidefinition}[\textsc{VarRef}: typing variable references]
+\begin{definition}[\textsc{VarRef}: typing variable references]
   \label{def:varref}
@@
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts. 
-\end{nidefinition}
+\end{definition}
 
@@
 
-\begin{nidefinition}[\textsc{Types}: typing processes]
+\begin{definition}[\textsc{Types}: typing processes]
   \label{def:types}
@@
   By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide a top-down proof of context split.
-\end{nidefinition}
+\end{definition}
 
@@
 
-\begin{nitheorem}[Framing]
+\begin{theorem}[Framing]
   \label{thm:framing}
@@
   Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{nitheorem}
+\end{theorem}
 
@@
 Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
-\begin{nitheorem}[Weakening]
+\begin{theorem}[Weakening]
   \label{thm:weakening}
@@
   $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
-\end{nitheorem}
+\end{theorem}
 
@@
 Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
-\begin{nitheorem}[Strengthening]
+\begin{theorem}[Strengthening]
   \label{thm:strengthening} 
@@
   Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
-\end{nitheorem}
+\end{theorem}
 
@@
 We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
-\begin{nitheorem}[Exchange]
+\begin{theorem}[Exchange]
   \label{thm:exchange}
@@
   Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
-\end{nitheorem}
+\end{theorem}
 
@@
 
-\begin{nilemma}
+\begin{lemma}
   \label{lm:lower-lift}
   The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
-\end{nilemma}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{nilemma}
+\begin{lemma}
   \label{lm:exchange-exchange}
   The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
-\end{nilemma}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{nilemma}
+\begin{lemma}
   \label{lm:types-unused}
   For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{nilemma}
+\end{lemma}
 \begin{proof}
@@
 
-\begin{nitheorem}[Subject congruence]
+\begin{theorem}[Subject congruence]
   \label{thm:subject-congruence}
   If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-\end{nitheorem}
+\end{theorem}
 
@@
       Case $\constr{scope-comm}$:
-      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged.
-\qedhere
+      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qed
   \end{itemize}
@@
 
-\begin{nitheorem}[Generalised substitution]
+\begin{theorem}[Generalised substitution]
   \label{thm:substitution-generalization}
@@
   Then substituting $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-\end{nitheorem}
+\end{theorem}
 
@@
    \item
-     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qedhere
+     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qed
  \end{itemize}
@@
 
-\begin{nitheorem}[Substitution]
+\begin{theorem}[Substitution]
   \label{thm:substitution}
@@
   Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
-\end{nitheorem}
+\end{theorem}
 \begin{proof}[Proof]
@@
 
-\begin{nilemma}
+\begin{lemma}
   \label{lm:comm-capable}
   Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
+\end{lemma}
 
@@
 
-\begin{nitheorem}[Subject reduction]
+\begin{theorem}[Subject reduction]
   \label{thm:subject-reduction}
@@
   \end{itemize}
-\end{nitheorem}
+\end{theorem}
 
@@
     \item
-    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qedhere
+    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qed
   \end{itemize}
@@
 We generalise the usage algebra, leading to linear, graded and shared type systems.
-Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016, Atkey2018}), in our work we too are able to talk about fully consumed resources --- e.g., we can transmit $\lz$ multiplicities of a fully exhausted channel.
+Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016,Atkey2018}), in our work we too are able to talk about fully consumed resources --- e.g., we can transmit $\lz$ multiplicities of a fully exhausted channel.
 
@@
 
-Based on contextual type theory \cite{Pientkaa, Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
+Based on contextual type theory \cite{Pientkaa,Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
 The result is a meta-theory in which HOAS encodings with both linear and dependent types can be described.
@@
 
-Further work on mechanising the \picalc{} \cite{Henry-Gerard1999, Honsell2001a, Bengtson2013, Despeyroux2000, Affeldt2008}, focuses on non-linear variations, whereas we present a range of linear, graded and shared types.
+Further work on mechanising the \picalc{} \cite{Henry-Gerard1999,Honsell2001a,Bengtson2013,Despeyroux2000,Affeldt2008}, focuses on non-linear variations, whereas we present a range of linear, graded and shared types.
 
@@
 
-\bibliographystyle{plainurl}
+\section*{Acknowledgments}
+
+We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.
+
+\bibliographystyle{splncs04}
 \bibliography{paper}

### 2020-12-14 paper/paper.tex
Fix cursive in definitions, lemmas and theorems

@@
 % Continue examples later on
-\usepackage{thmtools}
-\renewcommand\thmcontinues[1]{Continued}
+% \usepackage{thmtools}
+% \renewcommand\thmcontinues[1]{Continued}
 
 % Do not use italics in definitions and theorems
-% \theoremstyle{definition}
-% \newtheorem{definition}{Definition}
-% \newtheorem{theorem}{Theorem}
-% \newtheorem{lemma}{Lemma}
-% % \newtheorem{example}{Example}
-% \renewcommand{\definitionautorefname}{Definition}%
-% \renewcommand{\theoremautorefname}{Theorem}%
-% \renewcommand{\lemmaautorefname}{Lemma}%
-% \renewcommand{\sectionautorefname}{\S}%
-% \renewcommand{\subsectionautorefname}{\S}%
-% \renewcommand{\exampleautorefname}{Example}
+\spnewtheorem{nidefinition}[definition]{Definition}{\bfseries}{}
+\spnewtheorem{nilemma}[lemma]{Lemma}{\bfseries}{}
+\spnewtheorem{nitheorem}[theorem]{Theorem}{\bfseries}{}
+
+\renewcommand{\sectionautorefname}{\S}%
+\renewcommand{\subsectionautorefname}{\S}%
 
@@
   \todo{fix acknowledgments}
-  \todo{fix lemma, definition, theorem, example envs}
   The \picalc{} is a computational model for communication and concurrency.
@@
 
-\begin{definition}
+\begin{nidefinition}
   \label{def:unused}
@@
   The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
-\end{definition}
+\end{nidefinition}
 
@@
 
-\begin{lemma}
+\begin{nilemma}
   \label{lm:substitution-unused}
   For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
-\end{lemma}
+\end{nilemma}
 \begin{proof}
@@
 
-\begin{definition}[Usage algebra]
+\begin{nidefinition}[Usage algebra]
   A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
@@
   \end{figure}
-\end{definition}
+\end{nidefinition}
 
@@
 We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
-\begin{definition}[Indexed set of usage algebras]
+\begin{nidefinition}[Indexed set of usage algebras]
   An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type (\autoref{fig:indexed-multiplicities}).
@@
 
-\begin{definition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
+\todo{definition captions reset textsc}
+\begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
   A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two rows of \autoref{fig:prectx}).
@@
 A \emph{typing context} (last two rows of \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
-\end{definition}
-
-\begin{definition}[\textsc{Ctx}: usage contexts]
+\end{nidefinition}
+
+\begin{nidefinition}[\textsc{Ctx}: usage contexts]
   A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
@@
 
-\begin{definition}[\textsc{VarRef}: typing variable references]
+\begin{nidefinition}[\textsc{VarRef}: typing variable references]
   \label{def:varref}
@@
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts. 
-\end{definition}
+\end{nidefinition}
 
@@
 
-\begin{definition}[\textsc{Types}: typing processes]
+\begin{nidefinition}[\textsc{Types}: typing processes]
   \label{def:types}
@@
   By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide a top-down proof of context split.
-\end{definition}
+\end{nidefinition}
 
@@
 
-\begin{theorem}[Framing]
+\begin{nitheorem}[Framing]
   \label{thm:framing}
@@
   Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
-\begin{theorem}[Weakening]
+\begin{nitheorem}[Weakening]
   \label{thm:weakening}
@@
   $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
-\begin{theorem}[Strengthening]
+\begin{nitheorem}[Strengthening]
   \label{thm:strengthening} 
@@
   Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
-\begin{theorem}[Exchange]
+\begin{nitheorem}[Exchange]
   \label{thm:exchange}
@@
   Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 
-\begin{lemma}
+\begin{nilemma}
   \label{lm:lower-lift}
   The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
-\end{lemma}
+\end{nilemma}
 \begin{proof}
@@
 
-\begin{lemma}
+\begin{nilemma}
   \label{lm:exchange-exchange}
   The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
-\end{lemma}
+\end{nilemma}
 \begin{proof}
@@
 
-\begin{lemma}
+\begin{nilemma}
   \label{lm:types-unused}
   For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{lemma}
+\end{nilemma}
 \begin{proof}
@@
 
-\begin{theorem}[Subject congruence]
+\begin{nitheorem}[Subject congruence]
   \label{thm:subject-congruence}
   If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 
-\begin{theorem}[Generalised substitution]
+\begin{nitheorem}[Generalised substitution]
   \label{thm:substitution-generalization}
@@
   Then substituting $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-\end{theorem}
+\end{nitheorem}
 
@@
 
-\begin{theorem}[Substitution]
+\begin{nitheorem}[Substitution]
   \label{thm:substitution}
@@
   Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
-\end{theorem}
+\end{nitheorem}
 \begin{proof}[Proof]
@@
 
-\begin{lemma}
+\begin{nilemma}
   \label{lm:comm-capable}
   Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{lemma}
+\end{nilemma}
 
@@
 
-\begin{theorem}[Subject reduction]
+\begin{nitheorem}[Subject reduction]
   \label{thm:subject-reduction}
@@
   \end{itemize}
-\end{theorem}
+\end{nitheorem}
 

### 2020-12-15 paper/paper.tex
A much more to-the-point introduction

@@
   $\pi$ with leftovers: \\ a mechanisation in Agda
-  \thanks{The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
+  \thanks{
+    The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
+    We further want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
     This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
@@
 \begin{abstract}
-  \todo{the only references here are to session types}
-  \todo{fix acknowledgments}
-  The \picalc{} is a computational model for communication and concurrency.
-  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
-  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, such as \emph{session types} \cite{H93,THK94,HVK98}.
-
-  We present an untyped syntax and semantics for the \picalc{}, on top of which we provide a resource-aware type system parametrised by a set of multiplicity algebras.
-  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types, under the same unified framework.
-  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs common when dealing with linear type systems.
-  We provide a framing theorem, generalise the weakening and strengthening theorems, and use them to prove subject congruence.
-  We show that the type system is stable under substitution and prove subject reduction.
-%
-  Our formalisation is fully mechanised in Agda.
-
+%  \todo{the only references here are to session types}
+%  The \picalc{} is a computational model for communication and concurrency.
+%  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
+%  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, such as \emph{session types} \cite{H93,THK94,HVK98}.
+%
+%  We present an untyped syntax and semantics for the \picalc{}, on top of which we provide a resource-aware type system parametrised by a set of multiplicity algebras.
+%  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types, under the same unified framework.
+%  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs common when dealing with linear type systems.
+%  We provide a framing theorem, generalise the weakening and strengthening theorems, and use them to prove subject congruence.
+%  We show that the type system is stable under substitution and prove subject reduction.
+%%
+%  Our formalisation is fully mechanised in Agda.
+%
   \keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
@@
 
-The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
-Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
-The \picalc{} is at the basis of the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and more recently Go \footnote{https://golang.org}.
-At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, session types, and types for liveness properties, such as deadlock freedom, livelock freedom or termination.
-These make of the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
-
-In the early '90s, \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018} were introduced, following J.-Y. Girard's development of linear logic \cite{Girard87}.
-Linear type systems guarantee that resources are used \emph{exactly once}.
-Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Later on, linearity also inspired unique types, as in Clean \cite{BarendsenS96}, and ownership types, as in Rust \cite{MatsakisK14}.
-
-\todo{introduce shared and graded types sooner}
-
-Linearity influenced the $\pi$-calculus as well.
-Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
-%
-With the rise of session types \cite{H93,THK94,HVK98} linearity acquired a different flavour: linearity ensures that each channel endpoint is owned by exactly one communicating participant, even though the channel itself may be used multiple times, as specified by its session type.
-Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
-
-In recent years, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} \cite{KPT96} into the spotlight again.
-The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, thus allowing to use the target linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be defined and implemented.
-\todo{go back to linear \picalc{}}
-
-In this paper we make the following contributions:
-\begin{enumerate}
-  \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
-    Every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-    \todo{Do Ciccone \& Padovani also use type level de Bruijn indices?}
-
-  \item We \textbf{parametrise our type system over a set of resource-aware usage algebras}:
-    A \emph{usage algebra} is defined as a ternary relation that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative} and \emph{commutative}.
-    The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-    This allows for linear, graded and shared types to be mixed in the same type system.
-
-  \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system}:
-    \todo{along the lines of Allais \cite{Allais2018a}}.
-\end{enumerate}
-
-\subsection{Summary}
-\todo{shorten}
-\begin{enumerate}
-\item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
+Recent years have seen an increase in the efforts to mechanise resource-aware process algebras \todo{cite}.
+A resource-aware type system needs to keep track of what resources are used in which parts of the program.
+To do so, the standard approach is to use context splits\todo{cite}: a typing rule for a term with multiple subterms will require an extra arguments that specifies what resources to allocate to each of the subterms.
+The typing derivations for the subterms must then use the entirety of resources allocated to them.
+A key observation here is that each subterm already \emph{knows} about the resources it needs, especially if these are linear.
+\emph{Context splits do little more than repeat information that is already contained in the subterms.}
+Moreover, it is impossible to provide typing derivations for the subterms until after the context splits have been provided.
+On top of that, because in their usual presentation context splits are binary, a typing rule for a term with $n$ subterms requires $n - 1$ context splits, which adds to the clutter of the type system.
+
+An alternative approach is to using context splits is using leftover typing \todo{cite}.
+Leftover typing changes the form of the typing judgments themselves by including a second output \emph{leftover} context that contains the resources that were unused by the term.
+As a result, typing rules can now \emph{thread the resources through} subterms without the need of context splits.
+Each subterm uses the resources it needs, and leaves the rest for its siblings.
+The first subterm in this chain of resources immediately knows about the resources it has available.
+
+Leftover typing has been used to formulate intuitionistic linear logic \todo{cite} and to mechanise the linear \lambdacalc.
+We make two main contributions in this paper:
 \begin{itemize}
-   \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-   For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
-   These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
-
-   \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
-   Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
-   The reduction relation tracks at the type level the channel on which communication occurs.
+  \item We use \textbf{leftover typing to type a resource-aware \picalc{} and prove subject reduction}.
+  \item We \textbf{generalise the resource counting to a set of usage algebras that can be mixed within the same type system}.
+\end{itemize}
+
+% The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
+% Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
+% The \picalc{} is at the basis of the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and more recently Go \footnote{https://golang.org}.
+% At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, session types, and types for liveness properties, such as deadlock freedom, livelock freedom or termination.
+% These make of the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
+% 
+% In the early '90s, \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018} were introduced, following J.-Y. Girard's development of linear logic \cite{Girard87}.
+% Linear type systems guarantee that resources are used \emph{exactly once}.
+% Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
+% Later on, linearity also inspired unique types, as in Clean \cite{BarendsenS96}, and ownership types, as in Rust \cite{MatsakisK14}.
+% 
+% \todo{introduce shared and graded types sooner}
+% 
+% Linearity influenced the $\pi$-calculus as well.
+% Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
+% %
+% With the rise of session types \cite{H93,THK94,HVK98} linearity acquired a different flavour: linearity ensures that each channel endpoint is owned by exactly one communicating participant, even though the channel itself may be used multiple times, as specified by its session type.
+% Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
+% 
+% In recent years, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} \cite{KPT96} into the spotlight again.
+% The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, thus allowing to use the target linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be defined and implemented.
+% \todo{go back to linear \picalc{}}
+% 
+% In this paper we make the following contributions:
+% \begin{enumerate}
+%   \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
+%     Every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+%     \todo{Do Ciccone \& Padovani also use type level de Bruijn indices?}
+% 
+%   \item We \textbf{parametrise our type system over a set of resource-aware usage algebras}:
+%     A \emph{usage algebra} is defined as a ternary relation that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative} and \emph{commutative}.
+%     The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
+%     This allows for linear, graded and shared types to be mixed in the same type system.
+% 
+%   \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system}:
+%     \todo{along the lines of Allais \cite{Allais2018a}}.
+% \end{enumerate}
+% 
+% 
+% 
+% \subsection{Summary}
+% \todo{shorten}
+% \begin{enumerate}
+% \item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
+%  \begin{itemize}
+%    \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+%    For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
+%    These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
+%    
+%    \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
+%    Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
+%    The reduction relation tracks at the type level the channel on which communication occurs.
 This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
-   The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
- \end{itemize}
-
-  \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
-  we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
-  \begin{itemize}
-    \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
-
-    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, graded and shared types under the same framework.
-
-    \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    This approach adds a leftover usage context to the typing judgements.
-    Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
-    Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
-  \end{itemize}
-
-  \item \textbf{Fully mechanised formalisation and type safety results}:
-    The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-
-    We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
-    We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
-    We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
-\end{enumerate}
+%    The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+%  \end{itemize}
+%   
+%   \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
+%   we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
+%   \begin{itemize}
+%     \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
+%     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
+% 
+%     Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
+%     This allows for type systems combining linear, graded and shared types under the same framework.
+%     
+%     \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+%     This approach adds a leftover usage context to the typing judgements.
+%     Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
+%     Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
+%   \end{itemize}
+% 
+%   \item \textbf{Fully mechanised formalisation and type safety results}:
+%     The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
+% 
+%     We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
+%     We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
+%     We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
+% \end{enumerate}
 
@@
 
-\section*{Acknowledgments}
-
-We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their thoughts, patience, work, education and camaraderie.
-
 \bibliographystyle{splncs04}

### 2020-12-15 paper/paper.tex
Add walkthrough to the introduction

@@
   \item We use \textbf{leftover typing to type a resource-aware \picalc{} and prove subject reduction}.
+    Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+    Leftover typing \textbf{makes context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
+
   \item We \textbf{generalise the resource counting to a set of usage algebras that can be mixed within the same type system}.
+    The user can provide \emph{resource-aware} algebras that are then applied to the type system (\autoref{multiplicities}).
+    Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
+    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
+    This allows for type systems combining linear, graded (exact number of times $n$) and shared (free to reuse and dicard) types under the same framework.
 \end{itemize}
+
+We start by using type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+We then provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
+This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
+The reduction relation tracks at the type level the channel on which communication occurs.
+This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
+The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+
+We then introduce an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
+Finally, we present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
+we have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
+The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
 

### 2021-01-05 paper/paper.tex
Another go through the introduction

@@
 A resource-aware type system needs to keep track of what resources are used in which parts of the program.
-To do so, the standard approach is to use context splits\todo{cite}: a typing rule for a term with multiple subterms will require an extra arguments that specifies what resources to allocate to each of the subterms.
-The typing derivations for the subterms must then use the entirety of resources allocated to them.
-A key observation here is that each subterm already \emph{knows} about the resources it needs, especially if these are linear.
-\emph{Context splits do little more than repeat information that is already contained in the subterms.}
-Moreover, it is impossible to provide typing derivations for the subterms until after the context splits have been provided.
-On top of that, because in their usual presentation context splits are binary, a typing rule for a term with $n$ subterms requires $n - 1$ context splits, which adds to the clutter of the type system.
-
-An alternative approach is to using context splits is using leftover typing \todo{cite}.
-Leftover typing changes the form of the typing judgments themselves by including a second output \emph{leftover} context that contains the resources that were unused by the term.
-As a result, typing rules can now \emph{thread the resources through} subterms without the need of context splits.
+To do so, the standard approach is to use context splits\todo{cite}: typing rules for terms with multiple subterms use a side condition that specifies what resources to allocate to each of the subterms.
+The typing derivations for the subterms must then use the entirety of the resources allocated to them.
+A key observation here is that each subterm already \emph{knows} about the resources it needs.
+\emph{Context splits contain usage information already defined in the subterms.}
+Moreover, the subterms cannot be typed until the context splits have been defined.
+On top of that, context splits are usually binary, and thus typing rules with $n$ subterms require $n - 1$ context splits, which considerably clutters the type system.
+
+An alternative approach to using context splits is using leftover typing \todo{cite}.
+Leftover typing changes the shape of the typing judgments by including a second \emph{leftover} output context that contains the resources that were not used by the term.
+As a result, typing rules \emph{thread the resources through} subterms without the need of context splits.
 Each subterm uses the resources it needs, and leaves the rest for its siblings.
@@
     Leftover typing \textbf{makes context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
-
-  \item We \textbf{generalise the resource counting to a set of usage algebras that can be mixed within the same type system}.
+    \todo{revise paragraph}
+
+  \item We \textbf{generalise resource counting to a set of usage algebras that can be mixed within the same type system}.
     The user can provide \emph{resource-aware} algebras that are then applied to the type system (\autoref{multiplicities}).
     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
-    Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-    This allows for type systems combining linear, graded (exact number of times $n$) and shared (free to reuse and dicard) types under the same framework.
+    Multiple algebras can be mixed in the type system --- usage contexts keep information about what algebra to for each type (\autoref{contexts}).
+    In particular, this allows for type systems combining linear, graded (exact number of times $n$) and shared (free to reuse and dicard) types under the same framework.
 \end{itemize}
@@
 
-We then introduce an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
+We then define an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
 Finally, we present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
-we have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
+We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
 We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).

### 2021-01-05 paper/paper.tex
Present an example of context splits vs leftover typing in intro
Minor changes to the intro

@@
 Moreover, the subterms cannot be typed until the context splits have been defined.
-On top of that, context splits are usually binary, and thus typing rules with $n$ subterms require $n - 1$ context splits, which considerably clutters the type system.
+On top of that, using binary context splits means that typing rules with $n$ subterms require $n - 1$ context splits, which considerably clutters the type system.
 
 An alternative approach to using context splits is using leftover typing \todo{cite}.
-Leftover typing changes the shape of the typing judgments by including a second \emph{leftover} output context that contains the resources that were not used by the term.
-As a result, typing rules \emph{thread the resources through} subterms without the need of context splits.
-Each subterm uses the resources it needs, and leaves the rest for its siblings.
-The first subterm in this chain of resources immediately knows about the resources it has available.
+Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
+As a result, typing rules \emph{thread the resources through} subterms without needing context splits: each subterm uses the resources it needs, and leaves the rest for its siblings.
+The first subterm in this chain of resources immediately knows what resources it has available.
+Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right uses leftover typing.
+\begin{mathpar}
+  \inferrule
+  {\opctx{\Gamma}{\Delta}{\Xi} \\ \Delta \; \type{\vdash} \; P \\ \Xi \; \type{\vdash} \; Q}
+  {\Gamma \; \type{\vdash} \; \comp{P}{Q}}
+
+  \inferrule
+  {\Gamma \; \type{\vdash} \; P \; \type{\triangleright} \; \Delta \\ \Delta \; \type{\vdash} \; P \; \type{\triangleright} \; \Xi}
+  {\Gamma \; \type{\vdash} \; \comp{P}{Q} \; \type{\triangleright} \; \Xi}
+\end{mathpar}
 
@@
     \todo{revise paragraph}
+    \todo{add context around the picalc{}}
 
@@
 Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look like:
+\todo{we have this in th intro already}
 \begin{mathpar}

### 2021-01-05 paper/paper.tex
Intro in relatively good shape

@@
 
-Recent years have seen an increase in the efforts to mechanise resource-aware process algebras \todo{cite}.
-A resource-aware type system needs to keep track of what resources are used in which parts of the program.
-To do so, the standard approach is to use context splits\todo{cite}: typing rules for terms with multiple subterms use a side condition that specifies what resources to allocate to each of the subterms.
+Recent years have seen an increase in the efforts to mechanise linear process algebras \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}
+A linear type system guarantees that resources are used \emph{exactly once}.
+Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
+Such a linear type system needs to keep track of what resources are used in which parts of the program.
+To do so, the standard approach is to use context splits: typing rules for terms with multiple subterms use a side condition that specifies what resources to allocate to each of the subterms.
 The typing derivations for the subterms must then use the entirety of the resources allocated to them.
@@
 
-An alternative approach to using context splits is using leftover typing \todo{cite}.
+An alternative approach to using context splits is using leftover typing \cite{Mackie,Allais2018a}.
 Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
@@
 
-Leftover typing has been used to formulate intuitionistic linear logic \todo{cite} and to mechanise the linear \lambdacalc.
+Leftover typing has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
+In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
+The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
+Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
+Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
 We make two main contributions in this paper:
@@
   \item We use \textbf{leftover typing to type a resource-aware \picalc{} and prove subject reduction}.
-    Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    Leftover typing \textbf{makes context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
-    \todo{revise paragraph}
-    \todo{add context around the picalc{}}
+    Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+    In addition to making context splits unnecessary, leftover typing allows for an otherwise unstateable \emph{framing} theorem (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
+    In addition t that, leftover typing is naturally associative, making type safety properties considerably easier to reason about.
 
   \item We \textbf{generalise resource counting to a set of usage algebras that can be mixed within the same type system}.
-    The user can provide \emph{resource-aware} algebras that are then applied to the type system (\autoref{multiplicities}).
+    We do not instantiate our type system to only work with linear resources, instead we present an algebra-agnostic type system, and admit a mix of user-defined \emph{resource aware} algebras (\autoref{multiplicities}).
     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
     Multiple algebras can be mixed in the type system --- usage contexts keep information about what algebra to for each type (\autoref{contexts}).
-    In particular, this allows for type systems combining linear, graded (exact number of times $n$) and shared (free to reuse and dicard) types under the same framework.
+    In particular, this allows for type systems combining linear (use exactly once), graded (exact number of times $n$) and shared (free to reuse and discard) types under the same framework.
 \end{itemize}
@@
 
-% The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
-% Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
-% The \picalc{} is at the basis of the design and implementation of programming languages for concurrency, such as Pict \cite{PierceT00} and more recently Go \footnote{https://golang.org}.
-% At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, session types, and types for liveness properties, such as deadlock freedom, livelock freedom or termination.
-% These make of the \picalc{} a fully-fledged tool for modelling and verifying concurrent and distributed systems.
-% 
-% In the early '90s, \emph{linear types} for (functional) programming languages \cite{Wadler90,Bernardy2018} were introduced, following J.-Y. Girard's development of linear logic \cite{Girard87}.
-% Linear type systems guarantee that resources are used \emph{exactly once}.
-% Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-% Later on, linearity also inspired unique types, as in Clean \cite{BarendsenS96}, and ownership types, as in Rust \cite{MatsakisK14}.
-% 
 % \todo{introduce shared and graded types sooner}
-% 
-% Linearity influenced the $\pi$-calculus as well.
-% Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
-% %
-% With the rise of session types \cite{H93,THK94,HVK98} linearity acquired a different flavour: linearity ensures that each channel endpoint is owned by exactly one communicating participant, even though the channel itself may be used multiple times, as specified by its session type.
-% Linearity in session types is the key ingredient that guarantees communication safety and session fidelity.
-% 
-% In recent years, an encoding of session types into linear types \cite{DardhaGS12,Dardha14,DardhaGS17} has pushed the linear \picalc{} \cite{KPT96} into the spotlight again.
-% The encoding has been used as a technique to implement session types in mainstream programming languages such as OCaml \cite{Padovani17} and Scala \cite{ScalasY16,ScalasDHY17}, thus allowing to use the target linear \picalc{} as an underlying theoretical and practical framework on top of which session types can be defined and implemented.
-% \todo{go back to linear \picalc{}}
-% 
-% In this paper we make the following contributions:
-% \begin{enumerate}
 %   \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
@@
 % \end{enumerate}
-% 
-% 
-% 
-% \subsection{Summary}
-% \todo{shorten}
-% \begin{enumerate}
 % \item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
@@
 It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$ (see \autoref{fig:example-process}).
-\begin{flalign*}
-  & \func{carry} \; : \; \Var_n && \to \Var_n \to \Var_n \to \Process_n & \\
-  & \func{carry} \; in0 \; in1 \; out \; &&= \\
-  & && \recv{in0}{} \\
-  & && \recv{(\suc \; in1)}{} \\
-  & && \send{(\suc \suc \; out)}{\suc \constr{0}} \\
-  & && \send{(\suc \suc \; out)}{\constr{0}} \PO
-\end{flalign*}
+$$
+\func{carry} \; : \; \Var_n \to \Var_n \to \Var_n \to \Process_n
+$$
+\begin{alignat}{5}
+  \func{carry} \; in0 \; in1 \; out \; &&= \\
+  \recv{in0 &&}{} \\
+  \recv{(\suc \; in1) &&}{} \\
+  \send{(\suc \suc \; out) &&}{\suc \constr{0}} \\
+  \send{(\suc \suc \; out) &&}{\constr{0}} \PO
+\end{alignat}
 
@@
 Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
-As Gay does \cite{Gay2001}, context splits are given extrinsically, and exhausted resources are removed from typing contexts altogether.
+As Gay does in \cite{Gay2001}, context splits are given extrinsically, and exhausted resources are removed from typing contexts altogether.
 The runtime semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.

### 2021-01-05 paper/paper.tex
Get rid of most of the figures

@@
 \documentclass[]{llncs}
+% TODO: get rid of page numbers
+\pagestyle{plain}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h]
-  \begin{mathpar}
-    \datatype
-    { }
-    {\type{\N} : \Set}
-    \; \textsc{Nat}
-
-    \inferrule
-    { }
-    {\constr{0} : \type{\N}}
-
-    \inferrule
-    {n : \type{\N}}
-    {\suc n : \type{\N}}
-  \end{mathpar}
-  \caption{Notation used in this paper.}
-  \label{fig:notation}
-\end{figure}
-
-\autoref{fig:notation} illustrates the notation used in this paper.
+
+We briefly illustrate the notation used in this paper.
 Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.
@@
 Implicit arguments are mentioned in type definitions but omitted by constructors.
+\begin{mathpar}
+  \datatype
+  { }
+  {\type{\N} : \Set}
+  \; \textsc{Nat}
+
+  \inferrule
+  { }
+  {\constr{0} : \type{\N}}
+
+  \inferrule
+  {n : \type{\N}}
+  {\suc n : \type{\N}}
+\end{mathpar}
 
@@
 
-\begin{figure}[h]
+In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
+Names are cumbersome to mechanise: they are not inherently well scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
+To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
+That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
+
+A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
+References outside of that range are meaningless.
+It is useful to rule out these ill-scoped terms syntactically.
+We do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
 \begin{mathpar}
@@
 \end{equation*}
-  \caption{Well-scoped grammar with type-level de Bruijn indices.}
-  \label{fig:syntax}
-\end{figure}
-
-In order to mechanise the \picalc{} syntax in Agda, we need to deal with bound names in continuation processes.
-Names are cumbersome to mechanise: they are not inherently well scoped, one has to deal with alpha-conversion, and inserting new variables into a context entails proving that their names differ from all other names in context.
-To overcome these challenges, we use de Bruijn indices \cite{deBruijn1972}, where a natural number $n$ (aka \emph{index}) is used to refer to the variable introduced $n$ binders ago.
-That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
-
-A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
-References outside of that range are meaningless.
-It is useful to rule out these ill-scoped terms syntactically.
-In \autoref{fig:syntax} --- which presents the syntax of the \picalc{} using de Bruijn indices --- we do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
-We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
-Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
 
@@
 Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to.
-
-\begin{figure}[t]
+\begin{center}
   \small
-  \centering
   \begin{tikzpicture}[draw,circle]
@@
   \end{tikzpicture}
-  \caption{The courier system: two \sender{} $S_1$ and $S_2$, one \receiver{} $R$ and one \courier{} $C$.}
-  \label{fig:example-process}
-\end{figure}
+\end{center}
+
 For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
@@
 
-We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ in \autoref{fig:struct-cong-base}.
-
-\begin{figure}[h]
+We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ as follows:
 \begin{mathpar}
@@
 \end{mathpar}
-\caption{Base cases of structural congruence.}
-\label{fig:struct-cong-base}
-\end{figure}
 
@@
   A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
-  \todo{should we have this as a figure instead?}
-
   \begin{mathpar}
@@
 In order to remedy this, we model the recursive call structure syntactically and index \textsc{Equals} by a type \textsc{Rec}.
-
-\begin{figure}[h]
 \begin{mathpar}
@@
 \end{mathpar}
-\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.}
-\label{fig:struct-cong1}
-\end{figure}
 
@@
 
-The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs (\autoref{fig:reduction}).
+The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
 We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
+\begin{mathpar}
+  \datatype
+  {n : \N}
+  {\Channel_n : \Set}
+  \; \textsc{Channel}
+
+  \inferrule
+  { }
+  {\constr{internal} : \Channel_n}
+
+  \inferrule
+  {i : \Var_n}
+  {\constr{external} \; i : \Channel_n}
+
+  \datatype
+  {c : \Channel_n \\ P \; Q : \Process_n}
+  {P \reduce{c} Q : \Set}
+  \; \textsc{Reduces}
+
+  \inferrule
+  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
+  {\constr{comm} : \comp{\recv{i}{}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
+
+  \inferrule
+  {red : P \reduce{c} P'}
+  {\constr{par} \; red : \comp{P}{Q} \reduce{c} \comp{P'}{Q}}
+
+  \inferrule
+  {red : P \reduce{c} Q}
+  {\constr{res} \; red : \new P \reduce{\func{dec}\; c} \new Q}
+
+  \inferrule
+  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
+  {\constr{struct} \; eq \; red : P \reduce{c} Q}
+\end{mathpar}
 
@@
 --- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
-
-\begin{figure}[h]
-\begin{mathpar}
-  \datatype
-  {n : \N}
-  {\Channel_n : \Set}
-  \; \textsc{Channel}
-
-  \inferrule
-  { }
-  {\constr{internal} : \Channel_n}
-
-  \inferrule
-  {i : \Var_n}
-  {\constr{external} \; i : \Channel_n}
-
-  \datatype
-  {c : \Channel_n \\ P \; Q : \Process_n}
-  {P \reduce{c} Q : \Set}
-  \; \textsc{Reduces}
-
-  \inferrule
-  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n}
-  {\constr{comm} : \comp{\recv{i}{}P}{\send{i}{j}{Q}} \reduce{\constr{external} \; i} \comp{\func{lower}_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}}) \; uP'}{Q}}
-
-  \inferrule
-  {red : P \reduce{c} P'}
-  {\constr{par} \; red : \comp{P}{Q} \reduce{c} \comp{P'}{Q}}
-
-  \inferrule
-  {red : P \reduce{c} Q}
-  {\constr{res} \; red : \new P \reduce{\func{dec}\; c} \new Q}
-
-  \inferrule
-  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
-  {\constr{struct} \; eq \; red : P \reduce{c} Q}
-\end{mathpar}
-\caption{Operational semantics indexed by the channel over which reduction occurs.}
-\label{fig:reduction}
-\end{figure}
 
@@
   It has an element $\one$ that is used to count inputs and outputs.
-  \autoref{fig:multiplicities} defines such an algebra as a record $\Algebra_C$ on a carrier $C$.
-
-  \begin{figure}[h]
+  Below we define such an algebra as a record $\Algebra_C$ on a carrier $C$.
+  We use $\forall$ for universal quantification.
+  The dependent product $\type{\exists}$ uses the value of its first argument in the type of its second.
+  The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.
   \begin{equation*}
@@
   \end{equation*}
-  \caption{
-    Usage algebra $\Algebra_C$ on a carrier $C$.
-    We use $\forall$ for universal quantification.
-    The dependent product $\type{\exists}$ uses the value of its first argument in the type of its second.
-    The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.
-  }
-  \label{fig:multiplicities}
-  \end{figure}
 \end{nidefinition}
 
-\autoref{fig:algebra-instances} sketches the implementation of linear, graded and shared types as instances of our usage algebra.
+We sketch the implementation of linear, graded and shared types as instances of our usage algebra below.
 Their use in typing derivations is illustrated in \autoref{example-derivation} (Continued).
-
-\begin{figure}[h]
+\begin{center}
 \begin{tabular}{l | l | l}
@@
 \end{tabular}
-\caption{Example instances of $\Usage$ and a ternary relation that satisfies our $\Algebra$ interface.}
-\label{fig:algebra-instances}
-\end{figure}
+\end{center}
 
@@
 \begin{nidefinition}[Indexed set of usage algebras]
-  An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type (\autoref{fig:indexed-multiplicities}).
-  \begin{figure}[h]
+  An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type.
   \begin{equation*}
@@
   \end{equation*}
-    \caption{Indexed set of usage algebras.}
-    \label{fig:indexed-multiplicities}
-  \end{figure}
 \end{nidefinition}
 
-We keep typing contexts ($\PreCtx$, in \autoref{fig:prectx}) and usage contexts ($\Ctx$ in \autoref{fig:ctx}) separate.
+\todo{definition captions reset textsc}
+We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
 The former are preserved throughout typing derivations; the latter are transformed as a result of input, output, and context splits.
-
-\todo{definition captions reset textsc}
 \begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
-  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two rows of \autoref{fig:prectx}).
-
-  \begin{figure}[h]
+  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two constructors below).
   \begin{mathpar}
@@
   \end{mathpar}
-  \caption{Types and length-indexed typing contexts.}
-  \label{fig:prectx}
-  \end{figure}
+
   The unit type $\unit$ serves as a proof of inhabitance for types.
@@
 
-A \emph{typing context} (last two rows of \autoref{fig:prectx}) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$.
+  A \emph{typing context} (last two rules above) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$r
 \end{nidefinition}
@@
 \begin{nidefinition}[\textsc{Ctx}: usage contexts]
-  A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$ (\autoref{fig:ctx}).
-
+  A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$.
   A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to an existing context $\Gamma$.
-
-  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
-  Henceforth, we use $\lz$ to denote the multiplicity pair $0 \comma 0$, $\li$ for the pair $1 \comma 0$, $\lo$ for $0 \comma 1$, and $\lio$ for $1 \comma 1$.
-  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
-
-  \begin{figure}[h]
   \begin{mathpar}
@@
   \end{mathpar}
-  \caption{Length-indexed context of carrier indices with a context of usage annotations on top.}
-  \label{fig:ctx}
-  \end{figure}
+
+  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
+  Henceforth, we use $\lz$ to denote the multiplicity pair $0 \comma 0$, $\li$ for the pair $1 \comma 0$, $\lo$ for $0 \comma 1$, and $\lio$ for $1 \comma 1$.
+  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
 \end{nidefinition}
@@
 
-  The \textsc{VarRef} typing relation for variable references is presented in \autoref{fig:variable-references}.
-
-  \begin{figure}[h]
+  The \textsc{VarRef} typing relation for variable references is presented below.
   \begin{mathpar}
@@
   \end{mathpar}
-  \caption{Typing rules for variable references.}
-  \label{fig:variable-references}
-  \end{figure}
-
-  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{(x_l \comma x_r)}{(y_l \comma y_r)}{(z_l \comma z_r)}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
 
@@
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.
+
+  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{(x_l \comma x_r)}{(y_l \comma y_r)}{(z_l \comma z_r)}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
 \end{nidefinition}
@@
 
-  The \textsc{Types} typing relation for \picalc{} processes is presented in \autoref{fig:types}.
-  For convenience, we reuse the constructor names introduced for the syntax in \autoref{fig:syntax}.
-
-  \begin{figure}[h]
+  The \textsc{Types} typing relation for \picalc{} processes is presented below.
+  For convenience, we reuse the constructor names introduced for the syntax in \autoref{syntax}.
   \begin{mathpar}
@@
   \end{mathpar}
-  \caption{Leftover typing for a resource-aware type system.}
-  \label{fig:types}
-  \end{figure}
 

### 2021-01-06 paper/paper.tex
Shorten example in syntax

@@
 \end{itemize}
+\todo{mention we don't support recursion, why}
 
@@
 The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-
-% \todo{introduce shared and graded types sooner}
-%   \item We use \textbf{type-level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax for \picalc{} processes that is well-scoped by construction}:
-%     Every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-%     \todo{Do Ciccone \& Padovani also use type level de Bruijn indices?}
-% 
-%   \item We \textbf{parametrise our type system over a set of resource-aware usage algebras}:
-%     A \emph{usage algebra} is defined as a ternary relation that is \emph{partial}, \emph{deterministic}, \emph{cancellative}, \emph{associative} and \emph{commutative}.
-%     The user is able to choose any mix of algebras for the type system --- as long as they comply with the specification.
-%     This allows for linear, graded and shared types to be mixed in the same type system.
-% 
-%   \item We \textbf{apply leftover typing to the \picalc{} and prove type preservation for the resulting type system}:
-%     \todo{along the lines of Allais \cite{Allais2018a}}.
-% \end{enumerate}
-% \item \textbf{Formalisation of the syntax and the semantics of the \picalc{}}:
-%  \begin{itemize}
-%    \item \textbf{Syntax}: we use type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
-%    For convenience and readability, the user can convert from and to a syntax using channel names (\autoref{from_to_deBruijn}).
-%    These conversion functions are shown to be inverses of one another up to $\alpha$-conversion.
-%    
-%    \item \textbf{Semantics}: we provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
-%    Thanks to our well-scoped grammar  (\autoref{syntax}), the operational semantics is defined on \emph{the totality of the syntax} as a reduction relation on processes (\autoref{operational-semantics}).
-%    The reduction relation tracks at the type level the channel on which communication occurs.
-This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
-%    The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
-%  \end{itemize}
-%   
-%   \item \textbf{Leftover typing for the \picalc{} with support for linear, graded and shared types}:
-%   we present our type system for a resource-aware \picalc{} in \autoref{type-system}.
-%   \begin{itemize}
-%     \item \textbf{Algebras on multiplicities}: The user can provide \emph{resource-aware} algebras, which are then applied to the type system (\autoref{multiplicities}).
-%     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra --- linear types, graded types and shared types all satisfy these algebraic laws.
-% 
-%     Multiple algebras can be simultaneously used in a single type system --- usage contexts keep information about what algebra to use on which element (\autoref{contexts}).
-%     This allows for type systems combining linear, graded and shared types under the same framework.
-%     
-%     \item \textbf{Leftover typing}: Our type system uses \emph{leftover typing} to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-%     This approach adds a leftover usage context to the typing judgements.
-%     Typing derivations take the resources of their input usage context, consume some of them, and leave the remaining as leftovers in the output usage context.
-%     Leftover typing \textbf{makes top-down context splits unnecessary}, allows for \emph{framing} (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the \picalc{}.
-%   \end{itemize}
-% 
-%   \item \textbf{Fully mechanised formalisation and type safety results}:
-%     The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
-% 
-%     We present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
-%     We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
-%     We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
-% \end{enumerate}
 
@@
 A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
-References outside of that range are meaningless.
-It is useful to rule out these ill-scoped terms syntactically.
+References outside of that range are meaningless, and it is useful to rule them out syntactically.
 We do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
-We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains variables that can refer to $n$ distinct elements.
+We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains free variables that can refer to $n$ distinct elements.
 Every time we go under a binder, we increase the index of the continuation process, allowing the variable references within to refer to one more thing.
@@
 
-\todo{adapt to de bruijn}
-Process $\PO$ denotes the terminated process, where no further communications can occur.
-Process $\new{}\; P$ creates a new channel bound with scope $P$.
-Process $\comp{P}{Q}$ is the parallel composition of processes $P$ and $Q$.
-Processes $\recv{x}{} P$ and $\send{x}{y} P$ denote respectively, the input and output processes of a variable $y$ over a channel $x$, with continuation $P$.
-Scope restriction $(\new{}x) \; P$ and input $\recv{x}{y} \; P$ are \emph{binders}, they are the only constructs that introduce bound names --- $x$ and $y$ in $P$, respectively.
+Process $\PO$ denotes the terminated process, where no further communications can occur;
+process $\new{} P$ creates a new channel and binds it at index $0$ in the continuation process $P$;
+$\comp{P}{Q}$ composes processes $P$ and $Q$ in parallel;
+process $\recv{x}{}P$ receives data along channel $x$ and makes that data available at index $0$ in the continuation process $P$;
+process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
 
@@
 
-We present an example that highlights the ways in which linear, graded and shared types can be used to prevent errors in a concurrent and distributed system defined by \picalc{} processes in parallel.
-
-A courier system consists of three \emph{roles}:
+As an example use of this syntax, we present a courier system that consists of three \emph{roles}:
 a \sender{}, who wants to send a package;
@@
 and a \courier{}, who carries the package from the sender to the receiver.
-
-Our courier system (given in \autoref{fig:example-process}) is defined by four \picalc{} processes in parallel instantiating the above three roles:
-we have two \sender{} processes, $S_1$ and $S_2$, sending data over channels $x$ and $y$, respectively;
-one \receiver{} process, $R$, which receives over channel $z$ the data sent from each of the senders -- hence receives twice;
-and a \courier{} process $C$, which synchronises communication among $S_1$, $S_2$ and $R$.
-The \courier{} process $C$ first receives data from $S_1$ and $S_2$ on its input channels $x$ and $y$, respectively, and then sends the received data to $R$ along its output channel $z$.
-
-We will now define the courier system as \picalc{} processes.
-In our work, the syntax of \picalc{} processes adopts \emph{de Bruijn indices} \cite{deBruijn1972} for channel names and variables (details in \autoref{syntax}).
-In short, for the purposes of this example:
-process $\new{} P$ creates a new channel and makes it available at index $0$ in the continuation process $P$;
-$\comp{P}{Q}$ composes processes $P$ and $Q$ in parallel; process $\recv{x}{}P$ receives data along channel $x$ and makes that data available at index $0$ in the continuation process $P$;
-process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
-Both variable references $\Var_n$ and processes $\Process_n$ are indexed by the number $n$ of free variables they can refer to.
 \begin{center}
@@
 
-For the sake of clarity, we will define the four subprocesses in our courier system separately and parametrise them by the channels on which they operate.
-Each role \sender{}, \receiver{} and \courier{} is defined as a function respectively, $\func{send}, \func{recv}$ and $\func{carry}$.
-
-The \sender{} \emph{role} is defined below.
-The \sender{} creates a new channel to be sent as data, and sends it over channel $c$, and then terminates, denoted by process $\PO$.
+\todo{the colors are confusing: function color for send recv carry, individual colors for sender receiver and carrier}
+Our courier system is defined by four \picalc{} processes composed in parallel instantiating the above three roles:
+we have two \sender{} processes, $\func{send} \, x$ and $\func{send} \, y$, sending data over channels $x$ and $y$, respectively;
+one \receiver{} process, $\func{recv} \, z$, which receives over channel $z$ the data sent from each of the senders -- hence receives twice;
+and a \courier{} process $\func{carry} \, x \, y \, z$, which synchronises communication among the senders and the receiver.
+The \courier{} process first receives data from the two senders along its input channels $x$ and $y$, and then sends the two received bits of data to the receiver along its output channel $z$.
+
+The \sender{} and \receiver{} \emph{roles} are defined below, parametrised by the channels on which they operate.
+The \sender{} creates a new channel to be sent as data, and sends it over channel $c$, and then terminates.
+Each \sender{} process $\func{send} \, x$ and $\func{send} \, y$ is an instantiation of $\func{send} \, c$.
+The \receiver{} receives data \emph{twice} on a channel $c$ and then terminates.
+The \receiver{} process $\func{recv} \, z$ is an instantiation of $\func{recv} \, c$.
+\todo{get rid of vertical space}
+\begin{minipage}[t]{0.5\textwidth}
 \begin{flalign*}
@@
 \end{flalign*}
-Each \sender{} \emph{process} $S_1$ and $S_2$ is an instantiation of $\func{send} \; c$.
-In particular,
-$S_1$ is defined as $\func{send} \; x$ and $S_2$ as $\func{send} \; y$, where data is sent over channels $x$ and $y$, respectively (see \autoref{fig:example-process}).
-
-The \receiver{} role is defined below, where $\func{recv} \; c \;$ receives data \emph{twice} on a channel $c$ and then terminates with $\PO$.
+\end{minipage}
+\begin{minipage}[t]{0.5\textwidth}
 \begin{flalign*}
@@
 \end{flalign*}
-The \receiver{} process $R$ is defined as  $\func{recv} \; z$, with channel $z$ instantiating $c$, which is the channel used to receive data from \courier{} (see \autoref{fig:example-process}).
-
-The \courier{} role is defined below.
-The \courier{} process $C$ is defined as $\func{carry} \; x \; y \; z$.
-It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$ (see \autoref{fig:example-process}).
-$$
-\func{carry} \; : \; \Var_n \to \Var_n \to \Var_n \to \Process_n
-$$
-\begin{alignat}{5}
-  \func{carry} \; in0 \; in1 \; out \; &&= \\
-  \recv{in0 &&}{} \\
-  \recv{(\suc \; in1) &&}{} \\
-  \send{(\suc \suc \; out) &&}{\suc \constr{0}} \\
-  \send{(\suc \suc \; out) &&}{\constr{0}} \PO
-\end{alignat}
-
-Finally, we compose these processes together and create three communication channels: the first channel is shared between the \sender{} $S_1$ and the \courier{} $C$, the second between the \sender{} $S_2$ and the \courier{} $C$, and the third between the \receiver{} $R$ and the \courier{} $C$.
-The result is the courier systems defined below.
+\end{minipage}
+
+The \courier{} role is defined below as $\func{carry} \; x \; y \; z$.
+It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$.
+Finally, we create three communication channels and compose all four processes together: the first channel is shared between the one \sender{} and the \courier{}, the second between the other \sender{} and the \courier{}, and the third between the \receiver{}  and the \courier{}.
+The result is the courier $\func{system}$ defined below.
+\begin{minipage}[t]{0.5\textwidth}
+\begin{flalign*}
+  \func{carry} \; : \; \Var_n \to \Var_n \to \Var_n \to \Process_n%
+\end{flalign*}%
+\begin{flalign*}
+  & \func{carry} \; in0 \; in1 \; out \; &&= \\
+  & && \recv{in0 &&}{} \\
+  & && \recv{(\suc \; in1) &&}{} \\
+  & && \send{(\suc \suc \; out) &&}{\suc \constr{0}} \\
+  & && \send{(\suc \suc \; out) &&}{\constr{0}} \PO
+\end{flalign*}
+\end{minipage}
+\begin{minipage}[t]{0.5\textwidth}
 \begin{flalign*}
@@
 \end{flalign*}
-
-In order to type the courier system we will use all three typing disciplines: linear, graded and shared typing.
-We will continue this example and provide the typing derivations in \autoref{leftover-typing}.
-We use \textbf{linear types} for the \emph{usage annotations} of data and \textbf{shared types} for the \emph{data} itself that is being sent from the \sender{} processes to the \receiver{}.
-We use \textbf{graded types} for the usage annotations of the communication channels $x$, $y$, and $z$: the \sender{} processes $\func{send} \; x$ and $\func{send} \; y$ each consume $1$ output multiplicity and the \receiver{} process $\func{recv} \; z$ consumes $2$ input multiplicities, as it receives twice on channel $z$.
-Dually, the \courier{} process $\func{carry} \; x \; y \; z$ consumes one input multiplicity on $x$, one input multiplicity on $y$, and two output multiplicities on $z$.
+\end{minipage}
+\todo{fix}
+
+We continue this example and by providing typing derivations for these processes in \autoref{leftover-typing}.
+There we use a mix of linear, graded and shared typing to type check $\func{system}$.
 \end{example}

### 2021-01-06 paper/paper.tex
Shorten structural congruence

@@
 \label{semantics}
-Thanks to our well-scoped grammar in \autoref{fig:syntax}, the semantics of our language can now be defined on the totality of the syntax.
+Thanks to our well-scoped grammar in \autoref{syntax}, the semantics of our language can now be defined on the totality of the syntax.
 We define structural congruence in \autoref{structural-congruence} and provide a reduction relation in \autoref{operational-semantics}.
@@
 
+\begin{nidefinition}
+  \label{def:unused}
+  We consider a variable $i$ to be unused in $P$ ($\Unused_i \; P$) if none of the inputs nor the outputs refer to it.
+  $Unused_i \; P$ is defined as a recursive predicate on $P$, incrementing $i$ every time we go under a binder, and using $i \type{\not\equiv} x$ ( which unfolds to the negation of propositional equality on \textsc{Var}, i.e.
+$i \type{\equiv} x \to \type{\bot}$) to compare variables.
+\end{nidefinition}
+
+\begin{nidefinition}
 We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ as follows:
@@
 The last three ($\constr{scope}-$*) state garbage collection, scope extrusion and commutativity of restrictions, respectively.
-In $\constr{scope-ext}$ the type $\Unused_i \; Q$ is an inductive proof asserting that the variable index $i$ does not appear neither in the inputs nor in the outputs of $Q$ (see \autoref{def:unused}).
+In $\constr{scope-ext}$ the side condition $\Unused_i \; Q$ makes sure that $i$ is unused in $Q$ (see \autoref{def:unused}).
 The function $\func{lower}_i \; Q \; uQ$ traverses $Q$ decrementing every index greater than $i$.
@@
 In all the above, $i$ is incremented every time we go under a binder.
+\end{nidefinition}
 
 \begin{nidefinition}
-  \label{def:unused}
-  A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
-  \begin{mathpar}
-    \datatype
-    {i : \Var_n \\ P : \Process_n}
-    {\Unused_i \; P}
-    \; \textsc{Unused}
-
-    \inferrule
-    { }
-    {\Unused_i \; \PO}
-
-    \inferrule
-    {\Unused_{\suc i} \; P}
-    {\Unused_i \; \new{}P}
-
-    \inferrule
-    {\Unused_i \; P \\ \Unused_i \; Q}
-    {\Unused_i \; \comp{P}{Q}}
-
-    \inferrule
-    {i \type{\not\equiv} x \\ \Unused_{\suc i} \; P}
-    {\Unused_i \; \recv{x}{} P}
-
-    \inferrule
-    {i \type{\not\equiv} x \\ i \type{\not\equiv} y \\ \Unused_i \; P}
-    {\Unused_i \; \send{x}{y} P}
-  \end{mathpar}
-
-  The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
-\end{nidefinition}
-
-We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$ as shown in \autoref{fig:struct-cong1}.
+We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$.
 The $\constr{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively scope restriction, parallel composition, input and output.
@@
 \end{mathpar}
+\end{nidefinition}
 
@@
 \end{equation*}
-\todo{introduce $\hookrightarrow$ notation?}
+\todo{as inference rules, better}
 

### 2021-01-06 paper/paper.tex
Get rid of REC

@@
 \newcommand{\Algebra}{\type{ALGEBRA}}
-\newcommand{\eq}[1]{\, \type{\simeq}_{#1} \,}
+\newcommand{\eq}{\, \type{\simeq} \,}
 \newcommand{\eqeq}{\, \type{\cong} \,}
@@
 The remaining three rules close structural congruence under reflexivity, symmetry and transitivity, respectively. 
-The transitivity rule makes it tricky to prove predicates by induction on \textsc{Equals}: Agda cannot immediately see that such doubly recursive calls terminate.
-In order to remedy this, we model the recursive call structure syntactically and index \textsc{Equals} by a type \textsc{Rec}.
 \begin{mathpar}
   \datatype
-  { }
-  {\Rec : \Set}
-  \; \textsc{Rec}
-
-  \inferrule
-  { }
-  {\constr{zero} : \Rec}
-
-  \inferrule
-  {r : \Rec}
-  {\constr{one} \; r : \Rec}
-
-  \inferrule
-  {r \; s : \Rec}
-  {\constr{two} \; r \; s : \Rec}
-
-  \datatype
   {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq{r} Q : \Set}
+  {P \eq Q : \Set}
   \; \textsc{Equals}
@@
   {eq : P \eqeq Q}
-  {\constr{struct} \; eq : P \eq{\constr{zero}} Q}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-scope} \; eq : \new P \eq{\constr{one} \; r} \new P'}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-comp} \; eq : \comp{P}{Q} \eq{\constr{one} \; r} \comp{P'}{Q}}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-recv} \; eq : \recv{x}{}P \eq{\constr{one} \; r} \recv{x}{}P'}
-
-  \inferrule
-  {eq : P \eq{r} P'}
-  {\constr{cong-send} \; eq : \send{x}{y}P \eq{\constr{one} \; r} \send{x}{y}P'}
-
-  \inferrule
-  { }
-  {\constr{refl} : P \eq{\constr{zero}} P}
-
-  \inferrule
-  {eq : P \eq{r} Q}
-  {\constr{sym} \; eq : Q \eq{\constr{one} \; r} P}
-
-  \inferrule
-  {eq_1 : P \eq{r} Q \\ \; eq_2 : Q \eq{s} R}
-  {\constr{trans} \; eq_1 \; eq_2 : P \eq{\constr{two} \; r \; s} R}
+  {\constr{struct} \; eq : P \eq Q}
+
+  \inferrule
+  {eq : P \eq P'}
+  {\constr{cong-scope} \; eq : \new P \eq \new P'}
+
+  \inferrule
+  {eq : P \eq P'}
+  {\constr{cong-comp} \; eq : \comp{P}{Q} \eq \comp{P'}{Q}}
+
+  \inferrule
+  {eq : P \eq P'}
+  {\constr{cong-recv} \; eq : \recv{x}{}P \eq \recv{x}{}P'}
+
+  \inferrule
+  {eq : P \eq P'}
+  {\constr{cong-send} \; eq : \send{x}{y}P \eq \send{x}{y}P'}
+
+  \inferrule
+  { }
+  {\constr{refl} : P \eq P}
+
+  \inferrule
+  {eq : P \eq Q}
+  {\constr{sym} \; eq : Q \eq P}
+
+  \inferrule
+  {eq_1 : P \eq Q \\ \; eq_2 : Q \eq R}
+  {\constr{trans} \; eq_1 \; eq_2 : P \eq R}
 \end{mathpar}
@@
 --- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
+\todo{why do we not apply structural congruence on both sides?}
+
+\begin{nilemma}
+  \label{lm:substitution-unused}
+  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 
@@
 \end{equation*}
-\todo{as inference rules, better}
-
-\begin{nilemma}
-  \label{lm:substitution-unused}
-  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
+\todo{introduce function definition syntax, fix alignment}
 

### 2021-01-11 paper/paper.tex
Shorten the intro to the type system

@@
 
-We use \emph{leftover typing} \cite{Allais2018a} for our type system, an approach that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
+We present a resource-aware type system for the \picalc{} based on \emph{leftover typing} \cite{Allais2018a}, a technique that, in addition to the usual typing context $\PreCtx_n$ and (input) usage context $\Ctx_{idxs}$, adds an extra \emph{(output) usage context} $\Ctx_{idxs}$ to the typing rules.
 This output context contains the \emph{leftovers} (the unused multiplicities) of the process being typed.
-These leftovers can then be used as an input to another typing derivation.
-
-Typing judgments for linear calculi typically carry a single context of usage annotations and thus the typing rules require frequent top-down proofs of context splitting --- e.g., the typing rule for parallel composition would look like:
-\todo{we have this in th intro already}
-\begin{mathpar}
-  \inferrule
-  {\opctx{\Gamma}{\Delta}{\Xi} \\ \Delta \; \type{\vdash} \; P \\ \Xi \; \type{\vdash} \; Q}
-  {\Gamma \; \type{\vdash} \; \comp{P}{Q}}
-\end{mathpar}
-Input requires a similar context split; output requires two such context splits.
-This is rather cumbersome: for input and output, the proof of context split carries \emph{exactly} the same information as the typing rule for variables; for parallel composition it means that the typing derivations of both subprocesses will have no context information whatsoever until the context split is determined.
-
-Leftover typing solves this problem by inverting the flow of information so that \textbf{the typing derivations of the subprocesses determine the context split}.
-As a result, usage information is threaded bottom-up through the typing derivation, and top-down context split proofs are no longer needed.
-In addition, it allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.
-
-Our type system with leftovers is composed of two typing relations: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
-Both relations are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).
+These leftovers can then be used as input to another typing derivation.
+
+Leftover typing inverts the information flow concerning usage annotations so that it is the typing derivations of subprocesses which determine how resources are allocated.
+As a result, context split proofs are no longer necessary.
+Leftover typing also allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.
+
+Our type system is composed of two typing judgments: one for variable references (\autoref{def:varref}) and one for processes (\autoref{def:types}).
+Both judgments are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).
 

### 2021-01-11 paper/paper.tex
Shorten variable reference example

@@
 \begin{example}[Variable reference]
-  \autoref{lst:channels} types a variable reference under a linear algebra.
-  \begin{lstlisting}[label=lst:channels,xleftmargin=15pt,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
-$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
-$\_ \; = \; \suc \constr{0}$
-
-$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
-$\_ \; = \; \constr{0}$
-  \end{lstlisting}
-  In Agda, the underscore \_ introduces an anonymous declaration immediately followed by its definition.
-  We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
-  The \textsc{VarRef} constructors we must use are completely determined by $\suc \; \constr{0}$ in the type.
-  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output, the result is the obligation on line 4.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to fulfill this proof obligation automatically.
+  We type a variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.
+  \begin{flalign*}
+  & \func{egVar} \; : \; \contains{(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)} {\constr{[]} \comma \lio \comma \lio)} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {(\constr{[]} \comma \lo \comma \lio)}\\
+  & \func{egVar} \; = \; \suc \constr{0}
+  \end{flalign*}
+  We must show that the variable with index $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)$ and a usage context $(\constr{[]} \comma \lio \comma \lio)$.
+  The \textsc{VarRef} constructors that we must use are completely determined by the variable index $\suc \; \constr{0}$ in the type.
+  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to automatically fulfill this proof obligation.
 \end{example}

### 2021-01-11 paper/paper.tex
Move acknowledgments to the end

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{
-  $\pi$ with leftovers: \\ a mechanisation in Agda
-  \thanks{
-    The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions.
-    We further want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
-    This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
-}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
 \titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
@@
 }
+% Support
+% This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+\section*{Acknowledgments}
+
+We further want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
+
 \bibliographystyle{splncs04}

### 2021-01-11 paper/paper.tex
Add abstract

@@
 \begin{abstract}
-%  \todo{the only references here are to session types}
-%  The \picalc{} is a computational model for communication and concurrency.
-%  The \emph{linear} \picalc{} is a typed version of the \picalc{} where channels must be used exactly once.
-%  It is an underlying theoretical and practical framework on top of which more advanced types and theories can be built, such as \emph{session types} \cite{H93,THK94,HVK98}.
-%
-%  We present an untyped syntax and semantics for the \picalc{}, on top of which we provide a resource-aware type system parametrised by a set of multiplicity algebras.
-%  This allows the user to mix \emph{linear}, \emph{graded} and \emph{shared} types, under the same unified framework.
-%  We use leftover typing \cite{Allais2018a} to encode our typing rules, thus avoiding the ubiquitous context splitting proofs common when dealing with linear type systems.
-%  We provide a framing theorem, generalise the weakening and strengthening theorems, and use them to prove subject congruence.
-%  We show that the type system is stable under substitution and prove subject reduction.
-%%
-%  Our formalisation is fully mechanised in Agda.
-%
-  \keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
+  Linear type systems need to keep track of how programs use their resources.
+  The standard approach is to use context splits, side conditions that specify how resources are split across subterms.
+  With this approach, context splits echo information that is already present within subterms, which already know about the resources they need.
+  An alternative approach is to use leftover typing \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have an additional output usage context: the leftovers.
+  With this approach, the leftovers of one typing derivation can be fed as the inputs of the next, threading through linear resources while avoiding context splits.
+  We use this technique to define a type system for a resource-aware \picalc{} \cite{MilnerPW92,Milner99}.
+  This type system is parametrised over a set of usage algebras that are general enough to encompass shared types (free to reuse and discard), graded types (use an $n$ number of times) and linear types (use exactly once).
+  We provide a framing theorem, generalise the weakening and strengthening theorems, and show subject reduction.
+  Our formalisation is mechanised in about 1800 lines of Agda.
+
+%  \keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
 \end{abstract}
@@
 A key observation here is that each subterm already \emph{knows} about the resources it needs.
-\emph{Context splits contain usage information already defined in the subterms.}
+\emph{Context splits contain usage information already present in the subterms.}
 Moreover, the subterms cannot be typed until the context splits have been defined.
@@
     Multiple algebras can be mixed in the type system --- usage contexts keep information about what algebra to for each type (\autoref{contexts}).
-    In particular, this allows for type systems combining linear (use exactly once), graded (exact number of times $n$) and shared (free to reuse and discard) types under the same framework.
+    In particular, this allows for type systems combining linear (use exactly once), graded (exact number of $n$ times) and shared (free to reuse and discard) types under the same framework.
 \end{itemize}

### 2021-01-18 paper/appendix.tex
Fix cursive in definitions, lemmas and theorems
Meta theory section

--- 2020-12-03 paper/appendix.tex
@@
 
-\section{Definitions}
-\label{app:definitions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Type Safety} \label{app:type-safety}
 
-\begin{nidefinition}
-  A variable $i$ is unused in $P$ ($\Unused_i \; P$) if it appears in none of its inputs and none of its outputs.
+\subsection{Auxiliary Results}
+We will start this section with auxiliary results which are needed to prove two main theorems: subject congruence (\autoref{thm:subject-congruence}) and subject reduction (\autoref{thm:subject-reduction}) for our \picalc{} with leftovers.
 
-  \begin{mathpar}
-    \datatype
-    {i : \Var_n \\ P : \Process_n}
-    {\Unused_i \; P}
-    \; \textsc{Unused}
+\paragraph*{Framing}
+This property asserts that the well-typedness of a process is independent of its leftover resources.
 
-    \inferrule
-    { }
-    {\Unused_i \; \PO}
+\begin{nitheorem}[Framing]
+  \label{thm:framing}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
+  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+\end{nitheorem}
 
-    \inferrule
-    {\Unused_{\suc i} \; P}
-    {\Unused_i \; \new{}P}
+\paragraph*{Weakening}
+This property states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
+\begin{nitheorem}[Weakening]
+  \label{thm:weakening}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
+  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
+\end{nitheorem}
 
-    \inferrule
-    {\Unused_i \; P \\ \Unused_i \; Q}
-    {\Unused_i \; \comp{P}{Q}}
+\paragraph*{Strengthening}
+This property states that removing an unused variable preserves the well-typedness of a process.
+Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
+\begin{nitheorem}[Strengthening]
+  \label{thm:strengthening} 
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $i$ be a variable not in $P$, such that $uP \; : \; \Unused_i \; P$.
+  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
+\end{nitheorem}
 
-    \inferrule
-    {i \type{\not\equiv} x \\ \Unused_{\suc i} \; P}
-    {\Unused_i \; \recv{x}{} P}
+\paragraph*{Exchange}
+This property states that the exchange of two variables preserves the well-typedness of a process.
+We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
+\begin{nitheorem}[Exchange]
+  \label{thm:exchange}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
+\end{nitheorem}
 
-    \inferrule
-    {i \type{\not\equiv} x \\ i \type{\not\equiv} y \\ \Unused_i \; P}
-    {\Unused_i \; \send{x}{y} P}
-  \end{mathpar}
+\begin{proof}[Proof]
+  All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
+  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
+\end{proof}
 
-  The type $i \type{\not\equiv} x$ unfolds to the negation of propositional equality on \textsc{Var}, i.e. $i \type{\equiv} x \to \type{\bot}$.
-\end{nidefinition}
+\subsection{Towards Subject Reduction}
+\paragraph*{Subject Congruence}
+This property states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
+To prove this result, we must first introduce lemmas that establish that certain syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
 
-\section{Substitution}
-\label{app:substitution-generalization}
+\begin{nilemma}
+  \label{lm:lower-lift}
+  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:exchange-exchange}
+  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+\begin{nilemma}
+  \label{lm:types-unused}
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
+\end{nilemma}
+\begin{proof}
+  By induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
+
+We are now in a position to prove subject congruence.
+
+\begin{nitheorem}[Subject congruence]
+  \label{thm:subject-congruence}
+  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+\end{nitheorem}
+
+\begin{proof}[Proof]
+  The proof is by induction on \textsc{Equals} $\eq{}$.
+  Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
+  Full proof in \cite{Zalakain2020Agda}.
+  We proceed by induction on \textsc{StructCong} $\eqeq$:
+  \begin{itemize}
+    \item
+      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
+    \item
+      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
+    \item
+      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
+    \item
+      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
+      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$.
+    \item
+      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
+      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
+    \item
+      Case $\constr{scope-comm}$:
+      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qed
+  \end{itemize}
+\end{proof}
+
+\paragraph*{Substitution}
+This result is key to proving subject reduction.
+In \autoref{thm:substitution-generalization} we prove a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
+Then, in \autoref{thm:substitution} we instantiate the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
 
@@
 \end{proof}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+      \node (gamma-i) at (0,4)   {$\Gamma_i$};
+      \node (gamma-m) at (0,2)   {$\Gamma$};
+      \node (gamma-j) at (0,0)   {$\Gamma_j$};
+      \node (xi-i)    at (3,3.5) {$\Xi_i$};
+      \node (theta)   at (3,2.3) {$\Theta$};
+      \node (delta-m) at (3,1.7) {};
+      \node (xi-j)    at (3,0.5) {$\Xi_j$};
+      \node (psi-i)   at (6,3)   {$\Psi_i$};
+      \node (psi-m)   at (6,2)   {$\Psi$};
+      \node (psi-j)   at (6,1)   {$\Psi_j$};
+
+      \draw[-]  (gamma-m) -- (delta-m.center);
+      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \lz$}(psi-m);
+
+      \draw[->,densely dotted] (gamma-m) -- (theta);
+      \draw[->,densely dotted] (theta) -- (psi-m);
+
+      \draw[->] (gamma-i) -- node[left] {$\ni_i m$} (gamma-m);
+      \draw[->] (gamma-j) -- node[left] {$\ni_j m$} (gamma-m);
+      \draw[->] (psi-i) -- node[right] {$\ni_i n$} (psi-m);
+      \draw[->] (psi-j) -- node[right] {$\ni_j n$} (psi-m);
+
+      \draw[->] (gamma-i) -- node[above] {$\vdash P$} (xi-i);
+      \draw[->] (xi-i) -- node[above] {$\vdash Q$} (psi-i);
+      \draw[->,densely dotted] (gamma-j) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-j);
+      \draw[->,densely dotted] (xi-j) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-j);
+      \draw[->,densely dotted] (xi-i) -- node[left] {$\ni_i l$} (theta);
+      \draw[->,densely dotted] (xi-j) -- node[left] {$\ni_j l$} (theta);
+    \end{tikzpicture}
+    \caption{
+      Diagrammatic representation of the $\comp{}{}$ case for substitution.
+      Continuous lines represent known facts, dotted lines proof obligations.
+    }
+    \label{fig:subst}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}
+      \node (gamma-i) at (0,2)   {};
+      \node (gamma-m) at (0,1)   {};
+      \node (gamma-j) at (0,0)   {};
+      \node (xi-i)    at (2,3)   {};
+      \node (xi-m)    at (2,2) {};
+      \node (xi-j)    at (2,1)   {};
+      \node (psi-m)   at (4,2) {};
+
+      \draw[->] (gamma-i) -- (xi-j);
+      \draw[->] (gamma-m) -- (xi-j);
+      \draw[->,densely dotted] (gamma-j) -- (xi-j);
+      \draw[->] (gamma-i) -- (gamma-m);
+      \draw[->] (gamma-j) -- (gamma-m);
+      \draw[->] (xi-i)    -- (xi-m);
+      \draw[->] (xi-j)    -- (xi-m);
+      \draw[->] (xi-i)    -- (psi-m);
+      \draw[->] (xi-m)    -- (psi-m);
+      \draw[->,densely dotted] (xi-j)    -- (psi-m);
+    \end{tikzpicture}
+    \caption{
+      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
+      This approach makes composing typing derivations before and after subsitution more difficult.
+    }
+    \label{fig:subst-alternative}
+\end{figure}
+
+\begin{nitheorem}[Substitution]
+  \label{thm:substitution}
+  Let process $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
+  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
+  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
+\end{nitheorem}
+\begin{proof}[Proof]
+  For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
+  Then, we use these to apply \autoref{thm:substitution-generalization}.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \begin{note}
+  We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+% \end{note}
+
+\paragraph*{Subject Reduction}
+Finally we are ready to present our main result, stating  that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
+The relation between the typing contexts used to type $P$ and $Q$ will be explained in \autoref{thm:subject-reduction}.
+In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
+We first introduce an auxiliary lemma:
+
+\begin{nilemma}
+  \label{lm:comm-capable}
+  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
+\end{nilemma}
+
+\begin{proof}
+  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
+\end{proof}
+
+\begin{nitheorem}[Subject reduction]
+  \label{thm:subject-reduction}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
+  \begin{itemize}
+    \item If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
+  \end{itemize}
+\end{nitheorem}
+
+\begin{proof}[Proof]
+  By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  
+  \begin{itemize}
+    \item
+    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
+
+    \item
+    Case $\constr{par}$: by induction on the process that is being reduced.
+
+    \item
+    Case $\constr{res}$: case split on channel $c$:
+    if $\constr{internal}$ proceed inductively;
+    if $\constr{external}\; \constr{0}$ (i.e. the channel introduced by scope restriction) use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and proceed inductively;
+    if $\constr{external}\; (\suc i)$ proceed inductively.
+
+    \item
+    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qed
+  \end{itemize}
+\end{proof}

### 2021-01-18 paper/paper.tex
Meta theory section

@@
 \usepackage{todonotes}
+\usepackage{appendix}
 
@@
 \renewcommand{\subsectionautorefname}{\S}%
+\renewcommand{\appendixautorefname}{Appendix}
 
@@
   Linear type systems need to keep track of how programs use their resources.
-  The standard approach is to use context splits, side conditions that specify how resources are split across subterms.
+  The standard approach is to use context splits, side conditions that specify how resources are disjointly split across subterms.
   With this approach, context splits echo information that is already present within subterms, which already know about the resources they need.
@@
 We then define an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
-Finally, we present the type safety results of our \picalc{} with leftovers in \autoref{type-safety}.
+Finally, we present the meta theoretical properties of our \picalc{} with leftovers in \autoref{meta-theory}.
 We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
@@
 Once the types are given, the typing derivation is completely syntax directed.
+\begin{multicols}{2}
 \begin{flalign*}
-  & \func{\vdash recv} \; \types{\gamma \comma \channel{t}{\lz}}{\Gamma \comma (\suc \suc l \comma r)}{\func{recv} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash recv} \; = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO
+  & \func{\vdash recv} \; && : \; \types{\gamma \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc \suc l \comma r)\\ & &&}{\func{recv} \; \constr{0}\\ & &&}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash recv} \; && = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO
 \end{flalign*}
+\begin{flalign*}
+  & \func{\vdash send} \; && : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}\\ & &&}{\Gamma \comma (l \comma \suc r)\\ & &&}{\func{send} \; \constr{0}\\ & &&}{\Gamma \comma (l \comma r)} & \\
+  & \func{\vdash send} \; && = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
+\end{flalign*}
+\end{multicols}
 
@@
 Agda can uniquely determine the arguments required by the $\new{}$ constructor, and thus we can omit them using underscores.
-\begin{flalign*}
-  & \func{\vdash send} \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}}{\Gamma \comma (l \comma \suc r)}{\func{send} \; \constr{0}}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash send} \; = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
-\end{flalign*}
 
@@
 The typing derivation here is uniquely determined by the type and syntax of the process.
+\begin{multicols}{2}
 \begin{flalign*}
@@
 \end{flalign*}
-
-We can now compose these processes in parallel and type the courier system.
-Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
 \begin{flalign*}
@@
 \end{flalign*}
+\end{multicols}
+
+We can now compose these processes in parallel and type the courier system.
+Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
 \end{example}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Type Safety}
-\label{type-safety}
-\subsection{Auxiliary Results}
-We will start this section with auxiliary results which are needed to prove two main theorems: subject congruence (\autoref{thm:subject-congruence}) and subject reduction (\autoref{thm:subject-reduction}) for our \picalc{} with leftovers.
-
-\paragraph*{Framing}
-This property asserts that the well-typedness of a process is independent of its leftover resources.
-
-\begin{nitheorem}[Framing]
-  \label{thm:framing}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
-  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
-  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{nitheorem}
-
-\paragraph*{Weakening}
-This property states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
-Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
-\begin{nitheorem}[Weakening]
-  \label{thm:weakening}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
-  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
-\end{nitheorem}
-
-\paragraph*{Strengthening}
-This property states that removing an unused variable preserves the well-typedness of a process.
-Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
-\begin{nitheorem}[Strengthening]
-  \label{thm:strengthening} 
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable not in $P$, such that $uP \; : \; \Unused_i \; P$.
-  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
-\end{nitheorem}
-
-\paragraph*{Exchange}
-This property states that the exchange of two variables preserves the well-typedness of a process.
-We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
-\begin{nitheorem}[Exchange]
-  \label{thm:exchange}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, $\types{\func{exchange}_i \; \gamma}{\func{exchange}_i \; \Gamma}{\allowbreak \func{exchange}_i \; P}{\func{exchange}_i \; \Xi}$.
-\end{nitheorem}
-
-\begin{proof}[Proof]
-  All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
-  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
-\end{proof}
-
-\subsection{Towards Subject Reduction}
-\paragraph*{Subject Congruence}
-This property states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
-To prove this result, we must first introduce lemmas that establish that certain syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
-
-\begin{nilemma}
-  \label{lm:lower-lift}
-  The function $\func{lower}_i \; P \; uP$ has an inverse $\func{lift}_i \; P$ that increments every $\textsc{Var}$ greater than or equal to $i$, such that $\func{lift}_i \; (\func{lower}_i \; P \; uP) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:exchange-exchange}
-  The function $\func{exchange}_i \; P$ is its own inverse: $\func{exchange}_i \; (\func{exchange}_i \; P) \equiv P$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-\begin{nilemma}
-  \label{lm:types-unused}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
-\end{nilemma}
-\begin{proof}
-  By induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
-We are now in a position to prove subject congruence.
-
-\begin{nitheorem}[Subject congruence]
-  \label{thm:subject-congruence}
-  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-\end{nitheorem}
-
-\begin{proof}[Proof]
-  The proof is by induction on \textsc{Equals} $\eq{}$.
-  Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
-  Full proof in \cite{Zalakain2020Agda}.
-  We proceed by induction on \textsc{StructCong} $\eqeq$:
-  \begin{itemize}
-    \item
-      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
-    \item
-      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
-    \item
-      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
-    \item
-      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
-      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$.
-    \item
-      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
-      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
-    \item
-      Case $\constr{scope-comm}$:
-      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qed
-  \end{itemize}
-\end{proof}
-
-\paragraph*{Substitution}
-This result is key to proving subject reduction.
-In \autoref{thm:substitution-generalization} we prove a generalised version of substitution, where the substitition $\subst{P}{j}{i}$ is on any variable $i$.
-Then, in \autoref{thm:substitution} we instantiate the generalised version to the concrete case where $i$ is the most recently introduced variable $\constr{0}$, as required by subject reduction.
-
-\begin{nitheorem}[Generalised substitution]
-  \label{thm:substitution-generalization}
-  Let process $P$ be well typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Then, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
-  \begin{multicols}{2}
-  \begin{itemize}
-    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
-    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
-    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
-    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
-  \end{itemize}
-  \end{multicols}
-  Let $\Gamma$ and $\Psi$ be such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
-  Let $\Delta$ at position $i$ have usage $\lz$, meaning all consumption from $m$ to $n$ must happen in $P$.
-  Then substituting $i$ to $j$ in $P$ will be well typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-\end{nitheorem}
-
-\begin{proof}[Proof]
-  By induction on $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  Full proof can be found in \autoref{app:substitution-generalization}.
-  \begin{itemize}
-   \item
-     Constructor $\PO$: observe that $\Gamma_i \equiv \Psi_i$ and thus $\Gamma_j \equiv \Psi_j$.
-   \item
-     Constructor $\new$: we proceed inductively.
-   \item
-     Constructors $\recv{}{}$, $\send{}{}$ and $\comp{}{}$: we must find $\Theta$ in \autoref{fig:subst} and split the diagram along its vertical axis to proceed by induction. \qed
- \end{itemize}
-\end{proof}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}
-      \node (gamma-i) at (0,4)   {$\Gamma_i$};
-      \node (gamma-m) at (0,2)   {$\Gamma$};
-      \node (gamma-j) at (0,0)   {$\Gamma_j$};
-      \node (xi-i)    at (3,3.5) {$\Xi_i$};
-      \node (theta)   at (3,2.3) {$\Theta$};
-      \node (delta-m) at (3,1.7) {};
-      \node (xi-j)    at (3,0.5) {$\Xi_j$};
-      \node (psi-i)   at (6,3)   {$\Psi_i$};
-      \node (psi-m)   at (6,2)   {$\Psi$};
-      \node (psi-j)   at (6,1)   {$\Psi_j$};
-
-      \draw[-]  (gamma-m) -- (delta-m.center);
-      \draw[->] (delta-m.center) -- node[align=center,below] {$\Delta$\\$\Delta_i = \lz$}(psi-m);
-
-      \draw[->,densely dotted] (gamma-m) -- (theta);
-      \draw[->,densely dotted] (theta) -- (psi-m);
-
-      \draw[->] (gamma-i) -- node[left] {$\ni_i m$} (gamma-m);
-      \draw[->] (gamma-j) -- node[left] {$\ni_j m$} (gamma-m);
-      \draw[->] (psi-i) -- node[right] {$\ni_i n$} (psi-m);
-      \draw[->] (psi-j) -- node[right] {$\ni_j n$} (psi-m);
-
-      \draw[->] (gamma-i) -- node[above] {$\vdash P$} (xi-i);
-      \draw[->] (xi-i) -- node[above] {$\vdash Q$} (psi-i);
-      \draw[->,densely dotted] (gamma-j) -- node[below] {$\vdash \subst{P}{j}{i}$} (xi-j);
-      \draw[->,densely dotted] (xi-j) -- node[below] {$\vdash \subst{Q}{j}{i}$} (psi-j);
-      \draw[->,densely dotted] (xi-i) -- node[left] {$\ni_i l$} (theta);
-      \draw[->,densely dotted] (xi-j) -- node[left] {$\ni_j l$} (theta);
-    \end{tikzpicture}
-    \caption{
-      Diagrammatic representation of the $\comp{}{}$ case for substitution.
-      Continuous lines represent known facts, dotted lines proof obligations.
-    }
-    \label{fig:subst}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}
-      \node (gamma-i) at (0,2)   {};
-      \node (gamma-m) at (0,1)   {};
-      \node (gamma-j) at (0,0)   {};
-      \node (xi-i)    at (2,3)   {};
-      \node (xi-m)    at (2,2) {};
-      \node (xi-j)    at (2,1)   {};
-      \node (psi-m)   at (4,2) {};
-
-      \draw[->] (gamma-i) -- (xi-j);
-      \draw[->] (gamma-m) -- (xi-j);
-      \draw[->,densely dotted] (gamma-j) -- (xi-j);
-      \draw[->] (gamma-i) -- (gamma-m);
-      \draw[->] (gamma-j) -- (gamma-m);
-      \draw[->] (xi-i)    -- (xi-m);
-      \draw[->] (xi-j)    -- (xi-m);
-      \draw[->] (xi-i)    -- (psi-m);
-      \draw[->] (xi-m)    -- (psi-m);
-      \draw[->,densely dotted] (xi-j)    -- (psi-m);
-    \end{tikzpicture}
-    \caption{
-      Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-      This approach makes composing typing derivations before and after subsitution more difficult.
-    }
-    \label{fig:subst-alternative}
-\end{figure}
-
-\begin{nitheorem}[Substitution]
-  \label{thm:substitution}
-  Let process $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
-  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
-  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
-\end{nitheorem}
-\begin{proof}[Proof]
-  For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
-  Then, we use these to apply \autoref{thm:substitution-generalization}.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{note}
-  We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-% \end{note}
-
-\paragraph*{Subject Reduction}
-Finally we are ready to present our main result, stating  that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
-The relation between the typing contexts used to type $P$ and $Q$ will be explained in \autoref{thm:subject-reduction}.
-In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
-We first introduce an auxiliary lemma:
-
-\begin{nilemma}
-  \label{lm:comm-capable}
-  Every input usage context $\Gamma$ of a well-typed process $\types{\gamma}{\Gamma}{P}{\Delta}$ that reduces by communicating on a channel external (that is, $P \reduce{\constr{external} \; i} Q$ for some $Q$) has a multiplicity of at least $\lio$ at index $i$.
-\end{nilemma}
-
-\begin{proof}
-  By induction on the reduction derivation $P \reduce{\constr{external \; i}}Q$.
-\end{proof}
-
-\begin{nitheorem}[Subject reduction]
-  \label{thm:subject-reduction}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
-  \begin{itemize}
-    \item If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-    \item If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
-  \end{itemize}
-\end{nitheorem}
-
-\begin{proof}[Proof]
-  By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  
-  \begin{itemize}
-    \item
-    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
-
-    \item
-    Case $\constr{par}$: by induction on the process that is being reduced.
-
-    \item
-    Case $\constr{res}$: case split on channel $c$:
-    if $\constr{internal}$ proceed inductively;
-    if $\constr{external}\; \constr{0}$ (i.e. the channel introduced by scope restriction) use \autoref{lm:comm-capable} to subtract $\lio$ from the channel's usage annotation and proceed inductively;
-    if $\constr{external}\; (\suc i)$ proceed inductively.
-
-    \item
-    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qed
-  \end{itemize}
-\end{proof}
+\section{Meta-Theory} \label{meta-theory}
+
+We have mechanised subject reduction for our \picalc{} with leftovers in 850 lines of Agda code.
+The meta-theory of resource-aware type systems often needs to reason on typing derivations modulo associativity in the allocation of resources.
+For type systems using context splitting side conditions, this means applying associativity lemmas to recompute context splits; for type systems using leftover typing, this comes naturally \todo{rephrase?}.
+An example use of this is showing that the application of $\constr{comp-asssoc}$ preserves typing: the proof proceeds by deconstructing the input derivation into $\comp{P}{(\comp{Q}{R})}$ and reassembling it as $\comp{(\comp{P}{Q})}{R}$ without the need of any extra reasoning.
+
+All the reasoning carried out in our type safety proofs is based on the algebraic properties introduced in \autoref{multiplicities} -- an exception to this is $\func{\cdot -compute^r}$, which is only ever used for the user's convenience.
+This algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose arrows with the same typing context and a matching output and input usage contexts.
+
+Leftover typing also allows us to prove a \emph{framing} theorem, which states that arbitrary usage annotations can be added to the input usage context, as long as they are added to the output usage context as well -- it is a way of understanding a typing derivation independently from its unused resources.
+An example use of this is showing that $\constr{comp-comm}$ preserves typing: we use framing to show that in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
+
+We refer to \autoref{app:type-safety} \todo{autoref} for a more detailed account of the mechanised type-safety proofs.
 

### 2021-01-18 paper/paper.tex
Do not explicitly list all type safety theorems

@@
     Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}).
-    In addition to making context splits unnecessary, leftover typing allows for an otherwise unstateable \emph{framing} theorem (\autoref{thm:framing}) to be stated, and makes it possible to generalise \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}).
-    In addition t that, leftover typing is naturally associative, making type safety properties considerably easier to reason about.
+    In addition to making context splits unnecessary, leftover typing allows for a \emph{framing} theorem to be stated and is naturally associative, making type safety properties considerably easier to reason about (\autoref{meta-theory}).
 
@@
 The reduction relation tracks at the type level the channel on which communication occurs.
-This information is later used to state the subject reduction theorem --- aka type preservation (\autoref{thm:subject-reduction}).
+This information is later used to state the subject reduction theorem --- aka type preservation.
 The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
@@
 Finally, we present the meta theoretical properties of our \picalc{} with leftovers in \autoref{meta-theory}.
-We have fully mechanised \emph{framing} (\autoref{thm:framing}), \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) theorems.
-We use these results to then prove and mechanise \emph{subject congruence} (\autoref{thm:subject-congruence}), and together with \emph{substitution} (\autoref{thm:substitution}), to prove and mechanise \emph{subject reduction} (\autoref{thm:subject-reduction}).
+We have fully mechanised a proof of subject reduction, the details of which are found in \autoref{app:type-safety}
 The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
@@
 The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
-We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
+We keep track of channel $c$ so we can state subject reduction (\autoref{app:type-safety}).
 \begin{mathpar}
@@
 As a result, context split proofs are no longer necessary.
-Leftover typing also allows \emph{framing} (\autoref{thm:framing}) to be stated, and \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) to acquire a more general form.
+Leftover typing also allows \emph{framing} to be stated, and \emph{weakening} and \emph{strengthening} to acquire a more general form.
 
@@
 An example use of this is showing that $\constr{comp-comm}$ preserves typing: we use framing to show that in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
+Moreover, leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form, in which variables can freely be added or removed from context regardless of their usage annotation -- as long as this usage annotation is added or removed from the leftover context as well.
 

### 2021-01-18 paper/paper.tex
Allow for structural congruence after reduction
No need to spell out what a congruent equivalence relation is

@@
 We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$.
-The $\constr{cong-}$* rules define structural congruence under a context $\mathcal{C}[\cdot]$ \cite{Sangio01}, respectively scope restriction, parallel composition, input and output.
-The remaining three rules close structural congruence under reflexivity, symmetry and transitivity, respectively. 
-\begin{mathpar}
-  \datatype
-  {P \, Q : \Process_n \\ r : \Rec}
-  {P \eq Q : \Set}
-  \; \textsc{Equals}
-
-  \inferrule
-  {eq : P \eqeq Q}
-  {\constr{struct} \; eq : P \eq Q}
-
-  \inferrule
-  {eq : P \eq P'}
-  {\constr{cong-scope} \; eq : \new P \eq \new P'}
-
-  \inferrule
-  {eq : P \eq P'}
-  {\constr{cong-comp} \; eq : \comp{P}{Q} \eq \comp{P'}{Q}}
-
-  \inferrule
-  {eq : P \eq P'}
-  {\constr{cong-recv} \; eq : \recv{x}{}P \eq \recv{x}{}P'}
-
-  \inferrule
-  {eq : P \eq P'}
-  {\constr{cong-send} \; eq : \send{x}{y}P \eq \send{x}{y}P'}
-
-  \inferrule
-  { }
-  {\constr{refl} : P \eq P}
-
-  \inferrule
-  {eq : P \eq Q}
-  {\constr{sym} \; eq : Q \eq P}
-
-  \inferrule
-  {eq_1 : P \eq Q \\ \; eq_2 : Q \eq R}
-  {\constr{trans} \; eq_1 \; eq_2 : P \eq R}
-\end{mathpar}
+The relation \textsc{Equals} is structurally congruent under a context $\mathcal{C}[\cdot]$ \cite{Sangio01} and is reflexive, symmetric and transitive. 
 \end{nidefinition}
@@
   \inferrule
-  {eq : P \eq{} P' \\ red : P' \reduce{c} Q}
+  {eq_1 : P \eq{} P' \\ red : P' \reduce{c} Q' \\ eq_2 : Q' \eq{} Q}
   {\constr{struct} \; eq \; red : P \reduce{c} Q}

### 2021-01-19 paper/paper.tex
Add a bit on recursion

@@
 % Do not use italics in definitions and theorems
-\spnewtheorem{nidefinition}[definition]{Definition}{\bfseries}{}
+\spnewtheorem{nidefinition}[definition]{Definition}{}{}
 \spnewtheorem{nilemma}[lemma]{Lemma}{\bfseries}{}
@@
 
+\subsection{On recursion}
+\todo{put somewhere}
+
+Recursion can be added to a resource-algebra-agnostic type system with the following rule:
+\begin{mathpar}
+  \inferrule
+  {\types{\gamma}{\Gamma}{P}{\Gamma}}
+  {\types{\gamma}{\Gamma}{!P}{\Gamma}}
+\end{mathpar}
+That is, if a process consumes no resources (e.g. they are shared), then the process can be executed infinitely many times.
+To show subject congruence, one has to prove that $P \eqeq \comp{P}{!P}$ preserves types in both directions.
+In our setting, that means showing that $\types{\gamma}{\Gamma}{\comp{P}{!P}}{\Delta}$ then $\types{\gamma}{\Gamma}{!P}{\Delta}$.
+Eliminating the hypothesis gives $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{!P}{\Delta}$, eliminating the latter unifies $\Xi$ and $\Delta$ resulting in $\types{\gamma}{\Xi}{P}{\Xi}$.
+To satisfy the goal $\types{\gamma}{\Gamma}{!P}{\Xi}$, one has to show that $\Gamma$ and $\Xi$ are equivalent too.
+Note that the process $P$ does not syntactically restrict the types of the channels created within $P$.
+That means that within the typing derivations $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$ one can create channels that transmit different amounts of data in $\Gamma$ and in $\Xi$, which are then not necessarily the same anymore.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 

### 2021-01-19 paper/paper.tex
Mention replication section

@@
     In particular, this allows for type systems combining linear (use exactly once), graded (exact number of $n$ times) and shared (free to reuse and discard) types under the same framework.
+    We leave adding replication to our type system for future work (\autoref{replication}).
 \end{itemize}
-\todo{mention we don't support recursion, why}
 
@@
 As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
-We intend to add support for products, sum types and recursion to both our syntax and our type system.
+We intend to add support for products, sum types and replication to both our syntax and our type system.
 Orthogonally, making our typing rules bidirectional would allow us to provide a decision procedure for type checking processes in a given set of algebras.
@@
 
-\subsection{On recursion}
-\todo{put somewhere}
+\subsection{Replication}
+\label{replication}
 
-Recursion can be added to a resource-algebra-agnostic type system with the following rule:
+Replication can be added to a resource-algebra-agnostic type system with the following rule:
 \begin{mathpar}

### 2021-01-20 paper/paper.tex
update abstract

@@
   The standard approach is to use context splits, side conditions that specify how resources are disjointly split across subterms.
-  With this approach, context splits echo information that is already present within subterms, which already know about the resources they need.
+  With this approach, context splits echo information which is already present within subterms, thus leading to redundancy.
 An alternative approach is to use leftover typing \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have an additional output usage context: the leftovers.
@@
   We provide a framing theorem, generalise the weakening and strengthening theorems, and show subject reduction.
-  Our formalisation is mechanised in about 1800 lines of Agda.
+  Our formalisation is fully mechanised in about 1800 lines of Agda.
 

### 2021-01-20 paper/paper.tex
update intro

@@
 
-Recent years have seen an increase in the efforts to mechanise linear process algebras \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}
+%Recent years have seen an increase in the efforts to mechanise linear process algebras \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}
 A linear type system guarantees that resources are used \emph{exactly once}.
@@
 
-An alternative approach to using context splits is using leftover typing \cite{Mackie,Allais2018a}.
+An alternative approach to using context splits is using leftover typing, which has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
+
+In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
+The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
+Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
+%Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
+
 Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
@@
 
-Leftover typing has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
-In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
-The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
-Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
-Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
-We make two main contributions in this paper:
-\begin{itemize}
-  \item We use \textbf{leftover typing to type a resource-aware \picalc{} and prove subject reduction}.
-    Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}).
+We make the following main contributions in this paper:
+\begin{enumerate}
+  \item \textbf{Leftover typing for a resource-aware \picalc{}}. Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}) and satisfies subject reduction.
     In addition to making context splits unnecessary, leftover typing allows for a \emph{framing} theorem to be stated and is naturally associative, making type safety properties considerably easier to reason about (\autoref{meta-theory}).
 
-  \item We \textbf{generalise resource counting to a set of usage algebras that can be mixed within the same type system}.
+  \item \textbf{Shared, graded and linear unified \picalc{}}.
+  We {generalise resource counting to a set of usage algebras that can be mixed within the same type system}.
     We do not instantiate our type system to only work with linear resources, instead we present an algebra-agnostic type system, and admit a mix of user-defined \emph{resource aware} algebras (\autoref{multiplicities}).
@@
     In particular, this allows for type systems combining linear (use exactly once), graded (exact number of $n$ times) and shared (free to reuse and discard) types under the same framework.
-    We leave adding replication to our type system for future work (\autoref{replication}).
-\end{itemize}
+    Thanks to leftovers, our type system satisfies \emph{weakening} and \emph{strengthening} for the whole framework. In particular, these are stated for the first time for the linear \picalc{}.
+
+    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in \autoref{app:type-safety}
+The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
+
+\end{enumerate}
 
@@
 The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
-
 We then define an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
 Finally, we present the meta theoretical properties of our \picalc{} with leftovers in \autoref{meta-theory}.
-We have fully mechanised a proof of subject reduction, the details of which are found in \autoref{app:type-safety}
-The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}
 

### 2021-01-20 paper/paper.tex
Bring theorems back into the meta theory section

@@
 
-Leftover typing also allows us to prove a \emph{framing} theorem, which states that arbitrary usage annotations can be added to the input usage context, as long as they are added to the output usage context as well -- it is a way of understanding a typing derivation independently from its unused resources.
+Leftover typing also allows us to prove a \emph{framing} theorem, which states that arbitrary usage annotations can be added to the input usage context, as long as they are added to the output usage context as well -- one can thus understand a typing derivation independently from its unused resources.
 An example use of this is showing that $\constr{comp-comm}$ preserves typing: we use framing to show that in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
-Moreover, leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form, in which variables can freely be added or removed from context regardless of their usage annotation -- as long as this usage annotation is added or removed from the leftover context as well.
+\begin{nitheorem}[Framing]
+  \label{thm:framing}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
+  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
+  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
+  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+\end{nitheorem}
+
+Leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form, in which variables can freely be added or removed from context regardless of their usage annotation -- as long as this usage annotation is added or removed from the leftover context as well.
+Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
+Strengthening states that removing an unused variable preserves the well-typedness of a process.
+
+\begin{nitheorem}[Weakening]
+  \label{thm:weakening}
+  Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
+  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
+\end{nitheorem}
+
+\begin{nitheorem}[Strengthening]
+  \label{thm:strengthening} 
+  Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
+  Let $i$ be a variable not in $P$, such that $uP \; : \; \Unused_i \; P$.
+  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
+\end{nitheorem}
+
+\subsubsection{Type Safety}
+Subject congruence states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
+Subject reduction states that a well-typed process that communicates on a channel $c$ (\autoref{operational-semantics}) preserves its well-typedness and consumes $\lio$ from $c$ if $c$ is a channel in the environment.
+
+\begin{nitheorem}[Subject congruence]
+  \label{thm:subject-congruence}
+  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+\end{nitheorem}
+
+\begin{nitheorem}[Subject reduction]
+  \label{thm:subject-reduction}
+  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
+  \begin{itemize}
+    \item If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+    \item If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
+  \end{itemize}
+\end{nitheorem}
 

### 2021-01-20 paper/paper.tex
Rearrange dec definition
Rework intro

@@
   With this approach, context splits echo information which is already present within subterms, thus leading to redundancy.
-An alternative approach is to use leftover typing \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have an additional output usage context: the leftovers.
-  With this approach, the leftovers of one typing derivation can be fed as the inputs of the next, threading through linear resources while avoiding context splits.
-  We use this technique to define a type system for a resource-aware \picalc{} \cite{MilnerPW92,Milner99}.
+An alternative approach is to use leftover typing \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have also an output usage context: the leftovers.
+  With this approach, the leftovers of one typing derivation are fed as the inputs of the next, threading through linear resources while avoiding context splits.
+  We use this technique to define a type system for a \emph{resource-aware \picalc{}} \cite{MilnerPW92,Milner99}.
   This type system is parametrised over a set of usage algebras that are general enough to encompass shared types (free to reuse and discard), graded types (use an $n$ number of times) and linear types (use exactly once).
-  We provide a framing theorem, generalise the weakening and strengthening theorems, and show subject reduction.
+  In the \picalc{} linear types are of fundamental importance: they ensure the privacy and safety of the communication and avoid race conditions.
+  We provide a framing theorem for our type system, generalise the weakening and strengthening theorems, and show subject reduction.
   Our formalisation is fully mechanised in about 1800 lines of Agda.
@@
 
-%Recent years have seen an increase in the efforts to mechanise linear process algebras \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}
-A linear type system guarantees that resources are used \emph{exactly once}.
-Enforcing this ``no duplicating'' and ``no discarding'' of resources via linearity allows for resource-aware programming and more efficient implementations \cite{Wadler90}.
-Such a linear type system needs to keep track of what resources are used in which parts of the program.
-To do so, the standard approach is to use context splits: typing rules for terms with multiple subterms use a side condition that specifies what resources to allocate to each of the subterms.
-The typing derivations for the subterms must then use the entirety of the resources allocated to them.
+The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
+Notably, it features channel mobility: channels themselves are first class values and can be sent and received.
+Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} with linear channel types, where channels have to be used exactly once.
+In this setting, linearity guarantees the lack of race conditions and the privacy and safety of the communication.
+
+A linear type system must keep track of what resources are used in which parts of the program, and guarantee that they are \emph{neither duplicated nor discarded}.
+To do so, the standard approach is to use context splits: typing rules for terms with multiple subterms add an extra a side condition specifying what resources to allocate to each of the subterms.
+The typing derivations for the subterms must then use the entirety of their allocated resources.
 A key observation here is that each subterm already \emph{knows} about the resources it needs.
-\emph{Context splits contain usage information already present in the subterms.}
+\emph{Context splits contain usage information that is already present in the subterms.}
 Moreover, the subterms cannot be typed until the context splits have been defined.
 On top of that, using binary context splits means that typing rules with $n$ subterms require $n - 1$ context splits, which considerably clutters the type system.
-
-An alternative approach to using context splits is using leftover typing, which has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
-
-In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
-The \picalc{} \cite{MilnerPW92,Milner99} is a computational model for communication and concurrency that boils concurrent processing down to the sending and receiving of data over communication channels.
-Notably, it features channel mobility: channels themselves are first class values, and can therefore be sent and received.
-%Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} where the linear type system restricts the usage of channels to exactly once.
-
+There is a plethora of mechanisations of linear process calculi, all of them follow this approach \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
+
+An alternative approach is leftover typing, a technique that has already been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
 Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
@@
 
-We make the following main contributions in this paper:
+In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
+We make the following main contributions:
 \begin{enumerate}
@@
 --- notably, not under input nor output, as doing so would not preserve the sequencing of actions \cite{Sangio01}.
-\todo{why do we not apply structural congruence on both sides?}
-
-\begin{nilemma}
-  \label{lm:substitution-unused}
-  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
-
 Rule $\constr{res}$ uses $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
@@
 \todo{introduce function definition syntax, fix alignment}
+
+\begin{nilemma}
+  \label{lm:substitution-unused}
+  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
+\end{nilemma}
+\begin{proof}
+  By structural induction on \textsc{Process} and \textsc{Var}.
+\end{proof}
 

### 2021-01-20 paper/paper.tex
Shorten contrib 2

@@
     In particular, this allows for type systems combining linear (use exactly once), graded (exact number of $n$ times) and shared (free to reuse and discard) types under the same framework.
-    Thanks to leftovers, our type system satisfies \emph{weakening} and \emph{strengthening} for the whole framework. In particular, these are stated for the first time for the linear \picalc{}.
+    Thanks to leftover typing, we can now define \emph{weakening} and \emph{strengthening} for the whole framework, not just the shared fragment.
 

### 2021-01-20 paper/paper.tex
Remove typing rules for PreCtx, Idxs, and Ctx

@@
 
-  \begin{mathpar}
-    \datatype
-    {n : \N}
-    {\PreCtx_n : \Set}
-    \; \textsc{PreCtx}
-
-    \inferrule
-        { }
-        {[] : \PreCtx_{\constr{0}}}
-
-        \inferrule
-            {\gamma : \PreCtx_n \\ t : \Type}
-            {\gamma \comma t : \PreCtx_{\suc n}}
-  \end{mathpar}
-
   The unit type $\unit$ serves as a proof of inhabitance for types.
@@
 
-  A \emph{typing context} (last two rules above) is a length-indexed list of types, and is either empty $[]$ or the result of appending a type $t$ to an existing context $\gamma$r
+  A \emph{typing context} $\PreCtx_n$ is a length-indexed list of types that is either empty $\constr{[]}$ or the result of appending a type $\Type$ to an existing context.
 \end{nidefinition}
@@
 \begin{nidefinition}[\textsc{Ctx}: usage contexts]
-  A \emph{usage context} is a context $\Ctx_{idxs}$ of pairs of usage annotations that is indexed by a length-indexed context of indices $\Idxs_n$.
-  A usage context is either empty $[]$ or the result or appending a usage annotation $x$ to an existing context $\Gamma$.
-  \begin{mathpar}
-    \datatype
-    {n : \N}
-    {\Idxs_n : \Set}
-    \; \textsc{Idxs}
-
-    \inferrule
-        { }
-        {[] : \Idxs_{\constr{0}}}
-
-    \inferrule
-        {idxs : \Idxs_n \\ idx : \Idx}
-        {idxs \comma idx : \Idxs_{\suc n}}
-
-    \datatype
-    {idxs : \Idxs_n}
-    {\Ctx_{idxs} : \Set}
-    \; \textsc{Ctx}
-
-    \inferrule
-        { }
-        {[] : \Ctx_{[]}}
-
-    \inferrule
-        {\Gamma : \Ctx_{idxs} \\ x : \Usage_{idx} ^{\func{2}}}
-        {\Gamma \comma x : \Ctx_{idxs \comma idx}}
-  \end{mathpar}
-
+  A context of indices $\Idxs_n$ is a length-indexed list that is either empty $\constr{[]}$ or the result of appending an index $\Idx$ to an existing context.
+  A \emph{usage context} is a context $\Ctx_{idxs}$ indexed by a context of indices $\Idxs_n$ that is either empty $\constr{[]}$ or the result or appending a usage annotation pair $\Usage_{idx}^{\func{2}}$ with index $idx$ to an existing context.
   We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.

### 2021-01-20 paper/paper.tex
Nail down the explanations on PreCtx, Idxs, Ctx

@@
 \begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
-  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$) (first two constructors below).
+  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$).
   \begin{mathpar}
@@
 
-  A \emph{typing context} $\PreCtx_n$ is a length-indexed list of types that is either empty $\constr{[]}$ or the result of appending a type $\Type$ to an existing context.
+  A \emph{typing context} $\PreCtx_n$ is a length-indexed list of types that is either empty ($\constr{[]}$) or the result of appending a type $t : \Type$ to an existing context ($\gamma \constr{,} t$).
 \end{nidefinition}
 
-\begin{nidefinition}[\textsc{Ctx}: usage contexts]
-  A context of indices $\Idxs_n$ is a length-indexed list that is either empty $\constr{[]}$ or the result of appending an index $\Idx$ to an existing context.
-  A \emph{usage context} is a context $\Ctx_{idxs}$ indexed by a context of indices $\Idxs_n$ that is either empty $\constr{[]}$ or the result or appending a usage annotation pair $\Usage_{idx}^{\func{2}}$ with index $idx$ to an existing context.
+\begin{nidefinition}[\textsc{Idxs} and \textsc{Ctx}: contexts of indices and usage contexts]
+  A context of indices $\Idxs_n$ is a length-indexed list that is either empty ($\constr{[]}$) or the result of appending an index $i : \Idx$ to an existing context ($idxs \constr{,} i$).
+  A \emph{usage context} is a context $\Ctx_{idxs}$ indexed by a context of indices $idxs : \Idxs_n$ that is either empty ($\constr{[]}$) or the result or appending a usage annotation pair $u : \Usage_{idx}^{\func{2}}$ with index $idx : \Idx$ to an existing context ($\Gamma \constr{,} u$).
   We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.

### 2021-01-20 paper/paper.tex
Shorten bit on replication

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Related Work}
+\section{Conclusions, Related and Future Work}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section{Conclusion and Future Work}
+\subsection{Conclusions and Future Work}
 
@@
 
-\subsection{Replication}
+\subsubsection{Replication}
 \label{replication}
@@
 \end{mathpar}
-That is, if a process consumes no resources (e.g. they are shared), then the process can be executed infinitely many times.
-To show subject congruence, one has to prove that $P \eqeq \comp{P}{!P}$ preserves types in both directions.
-In our setting, that means showing that $\types{\gamma}{\Gamma}{\comp{P}{!P}}{\Delta}$ then $\types{\gamma}{\Gamma}{!P}{\Delta}$.
-Eliminating the hypothesis gives $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{!P}{\Delta}$, eliminating the latter unifies $\Xi$ and $\Delta$ resulting in $\types{\gamma}{\Xi}{P}{\Xi}$.
-To satisfy the goal $\types{\gamma}{\Gamma}{!P}{\Xi}$, one has to show that $\Gamma$ and $\Xi$ are equivalent too.
-Note that the process $P$ does not syntactically restrict the types of the channels created within $P$.
-That means that within the typing derivations $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$ one can create channels that transmit different amounts of data in $\Gamma$ and in $\Xi$, which are then not necessarily the same anymore.
+That is, if a process consumes no resources (e.g. because they are shared), then the process can be executed infinitely many times.
+To show subject congruence, one has to prove that $P \eqeq \comp{P}{!P}$ preserves typing in both directions.
+In our setting, that means showing that if $\types{\gamma}{\Gamma}{\comp{P}{!P}}{\Delta}$ then $\types{\gamma}{\Gamma}{!P}{\Delta}$.
+Eliminating the hypothesis gives $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$, where one has to show that $\Gamma$ and $\Xi$ are equivalent.
+However the process $P$ does not syntactically restrict the types of the channels created within $P$: within the typing derivations $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$ one can create channels that transmit different amounts of data in $\Gamma$ and in $\Xi$, which are then not necessarily the same anymore.
 

### 2021-01-20 paper/paper.tex
Join paragraphs in related work
Wrap definitions in definitions

@@
 
+\begin{nidefinition}[\textsc{Var} and \textsc{Process}]
+
 A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
@@
 process $\send{x}{y}{P}$ sends variable $y$ over channel $x$ and continues as process $P$.
+\end{nidefinition}
 
@@
 
-\begin{nidefinition}
+\begin{nidefinition}[\textsc{Unused}]
   \label{def:unused}
@@
 
-\begin{nidefinition}
-We define the base cases of a structural congruence relation \textsc{StructCong} $\eqeq$ as follows:
+\begin{nidefinition}[\textsc{StructCong}]
+We define the base cases of a structural congruence relation $\eqeq$ as follows:
 \begin{mathpar}
@@
 
-\begin{nidefinition}
-We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in \textsc{Equals} $\eq{}$.
-The relation \textsc{Equals} is structurally congruent under a context $\mathcal{C}[\cdot]$ \cite{Sangio01} and is reflexive, symmetric and transitive. 
+\begin{nidefinition}[\textsc{Equals}]
+We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in $\eq{}$.
+This relation is structurally congruent under a context $\mathcal{C}[\cdot]$ \cite{Sangio01} and is reflexive, symmetric and transitive. 
 \end{nidefinition}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Reduction Relation}
+\subsection{Operational Semantics}
 \label{operational-semantics}
 
-The operational semantics of the \picalc{} is defined as a reduction relation \textsc{Reduces} $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
+\begin{nidefinition}[\textsc{Reduces}]
+The operational semantics of the \picalc{} is defined as a reduction relation $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
 We keep track of channel $c$ so we can state subject reduction (\autoref{app:type-safety}).
@@
 \todo{introduce function definition syntax, fix alignment}
+\end{nidefinition}
 
@@
 
-\begin{nidefinition}[Usage algebra]
+\begin{nidefinition}[\textsc{Algebra}]
   A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
@@
 We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
-\begin{nidefinition}[Indexed set of usage algebras]
+\begin{nidefinition}[\textsc{Algebras}]
   An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type.
@@
 This allows one to talk about ill-typed terms, and defers the proof of subject reduction to a later stage.
-
+%
 To the best of our knowledge, leftover typing makes its appearance in 1994, when Ian Mackie first uses it to formulate intuitionistic linear logic \cite{Mackie}.
@@
 Drawing from quantitative type theory (by McBride and Atkey \cite{McBride2016,Atkey2018}), in our work we too are able to talk about fully consumed resources --- e.g., we can transmit $\lz$ multiplicities of a fully exhausted channel.
-
+%
 Recent years have seen an increase in the efforts to mechanise resource-aware process algebras, but one of the earliest works is the mechanisation of the linear \picalc{} in Isabelle/HOL by Gay \cite{Gay2001}.
@@
 We remove the demand for context splits, preserve the ability to talk about consumed resources, and adopt a more general usage algebra.
-
+%
 Orchard et al. introduce Granule \cite{Orchard}, a fully-fledged functional language with graded modal types, linear types, indexed types and polymorphism.
@@
 The authors provide bidirectional typing rules, and show the type safety of their semantics.
-
+%
 The work by Goto et al. \cite{Goto2016a} is, to the best of our knowledge, the first formalisation of session types which comes along with a mechanised proof of type safety in Coq.
@@
 Significantly, by merging the syntax and static semantics of the object language one can fully use the expressive power of the host language.
-
+%
 Thiemann formalises in Agda the MicroSession (minimal GV \cite{Gay2010}) calculus with support for recursion and subtyping \cite{Thiemann2019}.
@@
 The runtime semantics are given as an intrinsically typed CEK machine with a global context of session-typed channels.
-
+%
 In their recent paper, Ciccone and Padovani mechanise a dependently-typed linear \picalc{} in Agda \cite{Ciccone}.
@@
 In contrast to our work, their algebra is on the multiplicities $0$, $1$, $\omega$, and top-down context splitting proofs must be provided.
-
+%
 In another recent work, Rouvoet et al. provide an intrinsic type system for a \lambdacalc{} with session types \cite{Rouvoet2020}.
@@
 As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
-
+%
 Orchard and Yoshida \cite{OrchardY16} embed a small effectul imperative language into the session-typed \picalc{}, showing that session types are expressive enough to encode effect systems.
-
+%
 Based on contextual type theory \cite{Pientkaa,Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
@@
 The developer obtains for free an equational theory of substitution and decidable typechecking without having to encode context splits within the object language.
-
+%
 Further work on mechanising the \picalc{} \cite{Henry-Gerard1999,Honsell2001a,Bengtson2013,Despeyroux2000,Affeldt2008}, focuses on non-linear variations, whereas we present a range of linear, graded and shared types.
 
-\subsection{Conclusions and Future Work}
+\subsubsection*{Conclusions}
 
@@
 
+\subsubsection*{Future Work}
+\label{replication}
+
 As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
@@
 Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
-
-\subsubsection{Replication}
-\label{replication}
 

### 2021-01-20 paper/paper.tex
Use equals instead of righthookarrow

@@
     & \func{dec} : \Channel_(\suc n) && \to \Channel_n \\
-    & \func{dec} \; \constr{internal}                 && \hookrightarrow \constr{internal} \\
-    & \func{dec} \; (\constr{external} \; \constr{0}) && \hookrightarrow \constr{internal} \\
-    & \func{dec} \; (\constr{external} \; (\suc n))   && \hookrightarrow \constr{external} \; n \\
+    & \func{dec} \; \constr{internal}                 && = \constr{internal} \\
+    & \func{dec} \; (\constr{external} \; \constr{0}) && = \constr{internal} \\
+    & \func{dec} \; (\constr{external} \; (\suc n))   && = \constr{external} \; n \\
   \end{aligned}
 \end{equation*}
-\todo{introduce function definition syntax, fix alignment}
 \end{nidefinition}

### 2021-01-20 paper/paper.tex
Get rid of minor lemma

@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (we apply \autoref{lm:substitution-unused} to obtain a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
+After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (and a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$ is constructed).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 \end{nidefinition}
-
-\begin{nilemma}
-  \label{lm:substitution-unused}
-  For every variables $i$ and $j$, if $i \type{\not\equiv} j$ then $\Unused_i \allowbreak (\subst{P}{j}{i})$.
-\end{nilemma}
-\begin{proof}
-  By structural induction on \textsc{Process} and \textsc{Var}.
-\end{proof}
 

### 2021-01-20 paper/paper.tex
Get rid of unecessary subsubsections

@@
 
-\subsubsection{Type Safety}
 Subject congruence states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
@@
 
-\subsubsection*{Conclusions}
+\paragraph*{Conclusions}
 
@@
 
-\subsubsection*{Future Work}
-\label{replication}
+\paragraph*{Future Work}
 

### 2021-01-20 paper/paper.tex
Minor brushups

@@
 The operational semantics of the \picalc{} is defined as a reduction relation $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
-We keep track of channel $c$ so we can state subject reduction (\autoref{app:type-safety}).
+We keep track of channel $c$ so we can state subject reduction (\autoref{thm:subject-reduction}).
+
 \begin{mathpar}
@@
   \begin{aligned}
-    & \func{dec} : \Channel_(\suc n) && \to \Channel_n \\
     & \func{dec} \; \constr{internal}                 && = \constr{internal} \\
@@
 
-\begin{example}[Typing derivation, continues=example-process]
+\begin{example}[Typing derivation (Continued)]
 \label{example-derivation}

### 2021-01-22 paper/paper.tex
passing quickly

@@
 \begin{abstract}
-  Linear type systems need to keep track of how programs use their resources.
-  The standard approach is to use context splits, side conditions that specify how resources are disjointly split across subterms.
-  With this approach, context splits echo information which is already present within subterms, thus leading to redundancy.
-An alternative approach is to use leftover typing \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have also an output usage context: the leftovers.
-  With this approach, the leftovers of one typing derivation are fed as the inputs of the next, threading through linear resources while avoiding context splits.
-  We use this technique to define a type system for a \emph{resource-aware \picalc{}} \cite{MilnerPW92,Milner99}.
+  Linear type systems must keep track of how programs use their resources.
+  The standard approach is to use \emph{context splits} specifying how resources are disjointly split across subterms.
+  In this approach, context splits echo information which is already present within subterms, thus leading to redundancy.
+An alternative approach is to use \emph{leftover typing} \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have also an output usage context: the leftovers.
+  In this approach, the leftovers of one typing derivation are fed as the input of the next, threading through linear resources while avoiding context splits.
+  We use leftover typing to define a type system for a \emph{resource-aware \picalc{}} \cite{MilnerPW92,Milner99}.
   This type system is parametrised over a set of usage algebras that are general enough to encompass shared types (free to reuse and discard), graded types (use an $n$ number of times) and linear types (use exactly once).
-  In the \picalc{} linear types are of fundamental importance: they ensure the privacy and safety of the communication and avoid race conditions.
-  We provide a framing theorem for our type system, generalise the weakening and strengthening theorems, and show subject reduction.
-  Our formalisation is fully mechanised in about 1800 lines of Agda.
+   Linear types are of fundamental importance in the \picalc{}: they ensure privacy and safety of communication and avoid race conditions.
+  We provide a framing theorem for our type system, generalise the weakening and strengthening theorems to include linear types, and prove subject reduction.
+  Our formalisation is fully mechanised in about 1800 lines of Agda \cite{Zalakain2020Agda}.
 
@@
 Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} with linear channel types, where channels have to be used exactly once.
-In this setting, linearity guarantees the lack of race conditions and the privacy and safety of the communication.
+Linearity guarantees privacy and safety of communication and avoids race conditions.
 
@@
 On top of that, using binary context splits means that typing rules with $n$ subterms require $n - 1$ context splits, which considerably clutters the type system.
-There is a plethora of mechanisations of linear process calculi, all of them follow this approach \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
 
-An alternative approach is leftover typing, a technique that has already been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
+An alternative approach is leftover typing, a technique which has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
 Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
@@
 \end{mathpar}
+\todo{Ornela: Say something about the rules and why you presented them here. Did you want to convey that the second is so much easier and elegant etc etc?}
 
-In this paper, we use leftover typing to define a resource-aware type system for the \picalc{}.
+In this paper, we use leftover typing for the first time to define a resource-aware type system for the \picalc{} and fully mechanise it in Agda. 
+With regards to mechanisation, all previous work on mechanisation of linear process calculi follows the approach of context splits \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
+
 We make the following main contributions:
@@
 
-\subsection{Notation}
+\paragraph{Notation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Castro et al. \cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq, where de Bruijn indices are not as popular as in Agda or Idris.
-As a use-case, they use their tool to help automate proofs of subject reduction for a type system with session types.
+Castro et al. \cite{Castro2020} provide tooling for locally-nameless representations of process calculi in Coq, where de Bruijn indices are less popular than in Agda or Idris.
+They use their tool to help automate proofs of subject reduction for a type system with session types.
 %
@@
 %
-Further work on mechanising the \picalc{} \cite{Henry-Gerard1999,Honsell2001a,Bengtson2013,Despeyroux2000,Affeldt2008}, focuses on non-linear variations, whereas we present a range of linear, graded and shared types.
+Further work on mechanisation of the \picalc{} \cite{Henry-Gerard1999,Honsell2001a,Bengtson2013,Despeyroux2000,Affeldt2008}, focuses on non-linear variations, differently from our range of linear, graded and shared types.
 
-\paragraph*{Conclusions}
+\paragraph*{Conclusions and Future Work}
 
@@
 
-\paragraph*{Future Work}
-
-As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
-We intend to add support for products, sum types and replication to both our syntax and our type system.
-Orthogonally, making our typing rules bidirectional would allow us to provide a decision procedure for type checking processes in a given set of algebras.
-Furthermore, it might also be worth identifying correspondences between our usage algebra and particular state machines.
-Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
-
-Replication can be added to a resource-algebra-agnostic type system with the following rule:
+%As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
+%We intend to add support for products, sum types and replication to both our syntax and our type system.
+As future work we intend to expand our framework to include infinite behaviour by adding replication, which can be added to a resource-algebra-agnostic type system with the following rule:
 \begin{mathpar}
@@
 However the process $P$ does not syntactically restrict the types of the channels created within $P$: within the typing derivations $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$ one can create channels that transmit different amounts of data in $\Gamma$ and in $\Xi$, which are then not necessarily the same anymore.
+Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.
+Furthermore, it might also be worth identifying correspondences between our usage algebra and particular state machines.
+Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section*{Acknowledgments}
+\paragraph*{Acknowledgments}
 
-We further want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
+We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
 

### 2021-01-22 paper/paper.tex
more space

@@
 }
+\authorrunning{U. Zalakain \and O. Dardha}
+
 \institute{
-  University of Glasgow, Scotland \\ \email{u.zalakain.1@research.gla.ac.uk}
+  University of Glasgow, Scotland \ \ \email{u.zalakain.1@research.gla.ac.uk}
   \and
-  University of Glasgow, Scotland \\ \email{ornela.dardha@glasgow.ac.uk}
+  University of Glasgow, Scotland \ \ \email{ornela.dardha@glasgow.ac.uk}
 }

### 2021-01-25 paper/paper.tex
Update paper.tex
one always escapes
Rearrenge first example listings

@@
-\documentclass[]{llncs}
+\documentclass[runningheads]{llncs}
 % TODO: get rid of page numbers
-\pagestyle{plain}
 
@@
 \usepackage{appendix}
+\usepackage[strict]{changepage}
 
@@
 
-\renewcommand{\sectionautorefname}{\S}%
-\renewcommand{\subsectionautorefname}{\S}%
-\renewcommand{\appendixautorefname}{Appendix}
+%% remove for the camera ready
+\AtBeginDocument{%
+\renewcommand*{\figureautorefname}{Fig.\@\!}
+\renewcommand*{\sectionautorefname}{\S\!\!\,}
+\renewcommand*{\subsectionautorefname}{\S\!\!\,}
+\newcommand*{\nitheoremautorefname}{Thm.}
+\newcommand*{\nidefinitionautorefname}{Def.}
+\newcommand*{\nilemmaautorefname}{Lem.}
+\newcommand*{\exampleautorefname}{Eg.}
+}
 
@@
 The \receiver{} process $\func{recv} \, z$ is an instantiation of $\func{recv} \, c$.
-\todo{get rid of vertical space}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flalign*}
-  & \func{send} \; : \; \Var_n \to \Process_n & \\
-  & \func{send} \; c = \new{} \send{(\suc c)}{\constr{0}} \PO
-\end{flalign*}
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flalign*}
-  & \func{recv} \; : \; \Var_n \to \Process_n & \\
-  & \func{recv} \; c \; = \recv{c}{} \recv{(\suc c)} \PO
-\end{flalign*}
-\end{minipage}
+\begin{align*}
+  & \func{send} \; c = \new{} \send{(\suc c)}{\constr{0}} \PO &&
+    \func{recv} \; c \; = \recv{c}{} \recv{(\suc c)} \PO
+\end{align*}
 
@@
 The result is the courier $\func{system}$ defined below.
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flalign*}
-  \func{carry} \; : \; \Var_n \to \Var_n \to \Var_n \to \Process_n%
-\end{flalign*}%
-\begin{flalign*}
-  & \func{carry} \; in0 \; in1 \; out \; &&= \\
-  & && \recv{in0 &&}{} \\
-  & && \recv{(\suc \; in1) &&}{} \\
-  & && \send{(\suc \suc \; out) &&}{\suc \constr{0}} \\
-  & && \send{(\suc \suc \; out) &&}{\constr{0}} \PO
-\end{flalign*}
-\end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{flalign*}
-  & \func{system} && : \; \Process_{\constr{0}} & \\
-  & \func{system} && = \new{} ( \func{send} \; \constr{0} \\
-  & && \comp{}{ \new{} ( \func{send} \; \constr{0} } \\
-  & && \comp{}{ \new{} ( \func{recv} \; \constr{0} } \\
-  & && \comp{}{ \func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } \\
-\end{flalign*}
-\end{minipage}
-\todo{fix}
+\begin{align*}
+  & \func{carry} \; in0 \; in1 \; out \; =  \recv{in0 }{} 
+  \recv{(\suc \; in1) }{} 
+  \send{(\suc \suc \; out) }{\suc \constr{0}} 
+  \send{(\suc \suc \; out) }{\constr{0}} \PO \\
+  & \func{system} =  \new{} ( \func{send} \; \constr{0} 
+  \comp{}{\new{} ( \func{send} \; \constr{0} } 
+  \comp{}{\new{} ( \func{recv} \; \constr{0} } 
+  \comp{}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) } 
+\end{align*}
 
@@
 
+%%
+\clearpage
 \bibliographystyle{splncs04}
@@
 %%
+\clearpage
 \appendix
-\input{appendix.tex}
+\changetext{}{10em}{-5em}{-5em}{}\input{appendix.tex}
 \end{document}

### 2021-01-25 paper/paper.tex
Get rid of extra colors in example: it confuses notation colors

@@
 \newcommand{\eqeq}{\, \type{\cong} \,}
-
-% Example
-\newcommand{\sender}{\textcolor{brown}{sender}}
-\newcommand{\receiver}{\textcolor{purple}{receiver}}
-\newcommand{\courier}{\textcolor{cyan}{courier}}
 
@@
 As an example use of this syntax, we present a courier system that consists of three \emph{roles}:
-a \sender{}, who wants to send a package;
-a \receiver{}, who receives the package sent by the sender;
-and a \courier{}, who carries the package from the sender to the receiver.
+a sender, who wants to send a package;
+a receiver, who receives the package sent by the sender;
+and a courier, who carries the package from the sender to the receiver.
 \begin{center}
@@
 
-\todo{the colors are confusing: function color for send recv carry, individual colors for sender receiver and carrier}
 Our courier system is defined by four \picalc{} processes composed in parallel instantiating the above three roles:
-we have two \sender{} processes, $\func{send} \, x$ and $\func{send} \, y$, sending data over channels $x$ and $y$, respectively;
-one \receiver{} process, $\func{recv} \, z$, which receives over channel $z$ the data sent from each of the senders -- hence receives twice;
-and a \courier{} process $\func{carry} \, x \, y \, z$, which synchronises communication among the senders and the receiver.
-The \courier{} process first receives data from the two senders along its input channels $x$ and $y$, and then sends the two received bits of data to the receiver along its output channel $z$.
+we have two sender processes, $\func{send} \, x$ and $\func{send} \, y$, sending data over channels $x$ and $y$, respectively;
+one receiver process, $\func{recv} \, z$, which receives over channel $z$ the data sent from each of the senders -- hence receives twice;
+and a courier process $\func{carry} \, x \, y \, z$, which synchronises communication among the senders and the receiver.
+The courier process first receives data from the two senders along its input channels $x$ and $y$, and then sends the two received bits of data to the receiver along its output channel $z$.
 
-The \sender{} and \receiver{} \emph{roles} are defined below, parametrised by the channels on which they operate.
-The \sender{} creates a new channel to be sent as data, and sends it over channel $c$, and then terminates.
-Each \sender{} process $\func{send} \, x$ and $\func{send} \, y$ is an instantiation of $\func{send} \, c$.
-The \receiver{} receives data \emph{twice} on a channel $c$ and then terminates.
-The \receiver{} process $\func{recv} \, z$ is an instantiation of $\func{recv} \, c$.
+The sender and receiver \emph{roles} are defined below, parametrised by the channels on which they operate.
+The sender creates a new channel to be sent as data, and sends it over channel $c$, and then terminates.
+Each sender process $\func{send} \, x$ and $\func{send} \, y$ is an instantiation of $\func{send} \, c$.
+The receiver receives data \emph{twice} on a channel $c$ and then terminates.
+The receiver process $\func{recv} \, z$ is an instantiation of $\func{recv} \, c$.
 \begin{align*}
@@
 
-The \courier{} role is defined below as $\func{carry} \; x \; y \; z$.
+The courier role is defined below as $\func{carry} \; x \; y \; z$.
 It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$.
-Finally, we create three communication channels and compose all four processes together: the first channel is shared between the one \sender{} and the \courier{}, the second between the other \sender{} and the \courier{}, and the third between the \receiver{}  and the \courier{}.
+Finally, we create three communication channels and compose all four processes together: the first channel is shared between the one sender and the courier, the second between the other sender and the courier, and the third between the receiver  and the courier.
 The result is the courier $\func{system}$ defined below.
@@
 
-The \receiver{} defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
+The receiver defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
 After receiving twice, the process ends, and we should not be left with any unused multiplicities: $u$ needs to be $\lz$.
@@
 
-The \sender{} defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
-We instantiate $t$ (the type of data that the \sender{} sends) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
+The sender defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
+We instantiate $t$ (the type of data that the sender sends) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
 As per the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
@@
 
-Dually, the \courier{} defined by the $\func{carry}$ process expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
+Dually, the courier defined by the $\func{carry}$ process expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
 The typing derivation here is uniquely determined by the type and syntax of the process.

### 2021-01-25 paper/paper.tex
Get rid of structural congruence and reduction subsections

@@
 We then provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
-This operational semantics is defined as a reduction relation on processes (\autoref{operational-semantics}).
+This operational semantics is defined as a reduction relation on processes.
 The reduction relation tracks at the type level the channel on which communication occurs.
 This information is later used to state the subject reduction theorem --- aka type preservation.
-The reduction relation is defined modulo \emph{structural congruence} (\autoref{structural-congruence}) --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
+The reduction relation is defined modulo \emph{structural congruence} --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
 We then define an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
@@
 Thanks to our well-scoped grammar in \autoref{syntax}, the semantics of our language can now be defined on the totality of the syntax.
-We define structural congruence in \autoref{structural-congruence} and provide a reduction relation in \autoref{operational-semantics}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Structural Congruence}
-\label{structural-congruence}
 
@@
 \begin{nidefinition}[\textsc{StructCong}]
+  \label{def:structcong}
 We define the base cases of a structural congruence relation $\eqeq$ as follows:
@@
 
-\begin{nidefinition}[\textsc{Equals}]
+\begin{nidefinition}[\textsc{Equals}] \label{equals}
 We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in $\eq{}$.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Operational Semantics}
-\label{operational-semantics}
-
-\begin{nidefinition}[\textsc{Reduces}]
+\begin{nidefinition}[\textsc{Reduces}]\label{def:reduces}
 The operational semantics of the \picalc{} is defined as a reduction relation $\reduce{c}$ indexed by the channel $c$ on which communication occurs.
@@
   The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.
+
   \begin{equation*}
@@
   An \emph{indexed set of usage algebras} is a type $\Idx$ of indices that is nonempty ($\type{\exists IDX}$) together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras of the corresponding type.
+
   \begin{equation*}
@@
 
-Subject congruence states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
-Subject reduction states that a well-typed process that communicates on a channel $c$ (\autoref{operational-semantics}) preserves its well-typedness and consumes $\lio$ from $c$ if $c$ is a channel in the environment.
+Subject congruence states that applying structural congruence (\autoref{def:equals}) to a well-typed process preserves its well-typedness.
+Subject reduction states that a well-typed process that communicates on a channel $c$ (\autoref{def:reduces}) preserves its well-typedness and consumes $\lio$ from $c$ if $c$ is a channel in the environment.
 

### 2021-01-25 paper/paper.tex
Shorten bit on replication

@@
 We define the base cases of a structural congruence relation $\eqeq$ as follows:
+
 \begin{mathpar}
@@
 
-%As future work, we intend to prove further properties of our type system, such as that reduction preserves the balancing of channels.
-%We intend to add support for products, sum types and replication to both our syntax and our type system.
-As future work we intend to expand our framework to include infinite behaviour by adding replication, which can be added to a resource-algebra-agnostic type system with the following rule:
+As future work we intend to expand our framework to include infinite behaviour by adding replication, which can be added to a resource-algebra-agnostic type system like ours with the following rule:
 \begin{mathpar}
@@
   {\types{\gamma}{\Gamma}{P}{\Gamma}}
-  {\types{\gamma}{\Gamma}{!P}{\Gamma}}
+  {\types{\gamma}{\Gamma}{\constr{!}P}{\Gamma}}
 \end{mathpar}
-That is, if a process consumes no resources (e.g. because they are shared), then the process can be executed infinitely many times.
-To show subject congruence, one has to prove that $P \eqeq \comp{P}{!P}$ preserves typing in both directions.
-In our setting, that means showing that if $\types{\gamma}{\Gamma}{\comp{P}{!P}}{\Delta}$ then $\types{\gamma}{\Gamma}{!P}{\Delta}$.
-Eliminating the hypothesis gives $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$, where one has to show that $\Gamma$ and $\Xi$ are equivalent.
-However the process $P$ does not syntactically restrict the types of the channels created within $P$: within the typing derivations $\types{\gamma}{\Gamma}{P}{\Xi}$ and $\types{\gamma}{\Xi}{P}{\Xi}$ one can create channels that transmit different amounts of data in $\Gamma$ and in $\Xi$, which are then not necessarily the same anymore.
+That is, a process that consumes no resources (e.g. because they are shared) can be executed infinitely many times.
+To show subject congruence, one has to prove that $P \eqeq \comp{P}{\constr{!}P}$ preserves typing in both directions.
+To do so the resource consumption of a process has to be determined by e.g. adding type annotations to the syntax.
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.

### 2021-01-25 paper/paper.tex
Shorten typing examples

@@
 
-\todo{definition captions reset textsc}
 We keep typing contexts ($\PreCtx$) and usage contexts ($\Ctx$) separate.
@@
     \inferrule
-    { }
-    {\unit : \Type}
-
-    \inferrule
     {n : \N}
@@
 
-  The unit type $\unit$ serves as a proof of inhabitance for types.
-  The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as booleans, $1$ as natural numbers) so that we avoid having to deal with universe polymorphism.
+  The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as the unit type, $1$ as booleans) so that we avoid dealing with universe polymorphism.
   The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.
-  This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the multiplicities of input $i$ and output $o$ capabilities, respectively \cite{K07}.
+  This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the input and output multiplicities, respectively \cite{K07}.
 
@@
   For convenience, we reuse the constructor names introduced for the syntax in \autoref{syntax}.
+
   \begin{mathpar}
@@
 Once the types are given, the typing derivation is completely syntax directed.
-\begin{multicols}{2}
-\begin{flalign*}
-  & \func{\vdash recv} \; && : \; \types{\gamma \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc \suc l \comma r)\\ & &&}{\func{recv} \; \constr{0}\\ & &&}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash recv} \; && = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO
-\end{flalign*}
-\begin{flalign*}
-  & \func{\vdash send} \; && : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}\\ & &&}{\Gamma \comma (l \comma \suc r)\\ & &&}{\func{send} \; \constr{0}\\ & &&}{\Gamma \comma (l \comma r)} & \\
-  & \func{\vdash send} \; && = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
-\end{flalign*}
-\end{multicols}
+\begin{align*}
+   & \func{recvWT} \;  : \; \types{\gamma \comma \channel{t}{\lz}  }{\Gamma \comma (\suc \suc l \comma r)  }{\func{recv} \; \constr{0}  }{\Gamma \comma (l \comma r)} \\
+   & \func{recvWT} \;  = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
+   & \func{sendWT} \;  : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}  }{\Gamma \comma (l \comma \suc r)  }{\func{send} \; \constr{0}  }{\Gamma \comma (l \comma r)}  \\
+   & \func{sendWT} \;  = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
+\end{align*}
 
@@
 The typing derivation here is uniquely determined by the type and syntax of the process.
-\begin{multicols}{2}
-\begin{flalign*}
-& \func{\vdash carry} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
-& \func{\vdash carry} && = \\
-& && \recv{(\suc \suc \constr{0})}{} \\
-& && \recv{(\suc \suc \constr{0})}{} \\
-& && \send{(\suc \suc \constr{0})}{\suc \constr{0}} \\
-& && \send{(\suc \suc \constr{0})}{\constr{0}} \PO
-\end{flalign*}
-\begin{flalign*}
-& \func{\vdash system} && : \; \types{\constr{[]}}{\constr{[]}}{\func{system}}{\constr{[]}} & \\
-& \func{\vdash system} && = \new{} \; \_ \; \_ \; \_ \; ( \func{\vdash send} \\
-& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash send} \\
-& && \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{\vdash recv} \\
-& && \comp{}{ \func{\vdash carry} })))
-\end{flalign*}
-\end{multicols}
-
 We can now compose these processes in parallel and type the courier system.
 Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
+\begin{align*}
+& \func{carryWT} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
+& \func{carryWT} && = \recv{(\suc \suc \constr{0})}{} \recv{(\suc \suc \constr{0})}{} \send{(\suc \suc \constr{0})}{\suc \constr{0}} \send{(\suc \suc \constr{0})}{\constr{0}} \PO \\
+& \func{systemWT} && : \; \types{\constr{[]}}{\constr{[]}}{\func{system}}{\constr{[]}} \\
+& \func{systemWT} && = \new{} \; \_ \; \_ \; \_ \; ( \func{sendWT} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{sendWT} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{recvWT} \comp{}{ \func{carryWT} })))
+\end{align*}
 \end{example}

### 2021-01-25 paper/paper.tex
Fix references to appendices

@@
 
-    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in \autoref{app:type-safety}
+    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in Appendix \ref{app:type-safety}
 The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
@@
 
-We refer to \autoref{app:type-safety} \todo{autoref} for a more detailed account of the mechanised type-safety proofs.
+We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised type-safety proofs.
 

### 2021-01-25 paper/paper.tex
Compress things up

@@
 \begin{nidefinition}[\textsc{Var} and \textsc{Process}]
-
 A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
@@
 
-\begin{nidefinition}[\textsc{Unused}]
-  \label{def:unused}
+\begin{nidefinition}[\textsc{Unused}] \label{def:unused}
   We consider a variable $i$ to be unused in $P$ ($\Unused_i \; P$) if none of the inputs nor the outputs refer to it.
@@
 
-\begin{nidefinition}[\textsc{StructCong}]
-  \label{def:structcong}
+\begin{nidefinition}[\textsc{StructCong}] \label{def:structcong}
 We define the base cases of a structural congruence relation $\eqeq$ as follows:
@@
 
-\begin{nidefinition}[\textsc{VarRef}: typing variable references]
-  \label{def:varref}
-
+\begin{nidefinition}[\textsc{VarRef}: typing variable references] \label{def:varref}
   The \textsc{VarRef} typing relation for variable references is presented below.
@@
 
-\begin{nidefinition}[\textsc{Types}: typing processes]
-  \label{def:types}
-
+\begin{nidefinition}[\textsc{Types}: typing processes] \label{def:types}
   The \textsc{Types} typing relation for \picalc{} processes is presented below.
@@
 \label{example-derivation}
-
 We provide the typing derivation for the courier system defined in \autoref{example-process}.
@@
 
-The receiver defined by the $\func{recv}$ process receives data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
-After receiving twice, the process ends, and we should not be left with any unused multiplicities: $u$ needs to be $\lz$.
+The receiver defined by the $\func{recv}$ process receives data along the channel with index $\constr{0}$, which needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+After receiving twice, the process ends: we must not be left with any unused multiplicities, thus $u = \lz$.
 We will use graded types to keep track of the exact number of times communication happens.

### 2021-01-25 paper/paper.tex
Everything fits!

@@
 
-\begin{nidefinition}[\textsc{Equals}] \label{equals}
+\begin{nidefinition}[\textsc{Equals}] \label{def:equals}
 We lift the relation \textsc{StructCong} $\eqeq{}$ and close it under equivalence and congruence in $\eq{}$.
@@
 Whatever the input multiplicity of the channel, we will consume $2$ of it and leave the remaining as leftovers.
-Once the types are given, the typing derivation is completely syntax directed.
+The sender defined by the $\func{send}$ process sends data along the channel with index $\constr{0}$, which needs to be of type $\channel{t}{u}$ for some $t$ and $u$.
+We instantiate $t$ (the type of data that the sender sends) to the trivial channel $\channel{\unit}{\func{\omega}}$.
+As per the type of the process $\func{recv}$, $u = \lz$.
+We will transmit once, thus use $\suc{\constr{0}}$ output multiplicity, and leave the rest as leftovers.
+Agda can uniquely determine the arguments required by the $\new{}$ constructor.
 \begin{align*}
@@
 
-The sender defined by the $\func{send}$ process sends data along the channel with index $0$, thus that variable needs to be of channel type $\channel{t}{u}$ for some $t$ and $u$.
-We instantiate $t$ (the type of data that the sender sends) to $\channel{\unit}{\func{\omega}}$, as that is a value we can easily construct.
-As per the type of the process $\func{recv}$, the transmitted multiplicities $u$ are $\lz$.
-We will transmit once, thus use a single output multiplicity, and leave the rest as leftovers.
-Agda can uniquely determine the arguments required by the $\new{}$ constructor, and thus we can omit them using underscores.
-
 Dually, the courier defined by the $\func{carry}$ process expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
-The typing derivation here is uniquely determined by the type and syntax of the process.
 We can now compose these processes in parallel and type the courier system.
-Specifying the types of the channels is not required: Agda can infer them from the types of the processes.
 \begin{align*}
@@
 The meta-theory of resource-aware type systems often needs to reason on typing derivations modulo associativity in the allocation of resources.
-For type systems using context splitting side conditions, this means applying associativity lemmas to recompute context splits; for type systems using leftover typing, this comes naturally \todo{rephrase?}.
-An example use of this is showing that the application of $\constr{comp-asssoc}$ preserves typing: the proof proceeds by deconstructing the input derivation into $\comp{P}{(\comp{Q}{R})}$ and reassembling it as $\comp{(\comp{P}{Q})}{R}$ without the need of any extra reasoning.
+For type systems using context splitting side conditions, this means applying associativity lemmas to recompute context splits; for type systems using leftover typing it does not.
+As an example, the proof that $\constr{comp-asssoc}$ preserves typing proceeds by deconstructing the input derivation into $\comp{P}{(\comp{Q}{R})}$ and reassembling it as $\comp{(\comp{P}{Q})}{R}$ without the need of any extra reasoning.
 
-All the reasoning carried out in our type safety proofs is based on the algebraic properties introduced in \autoref{multiplicities} -- an exception to this is $\func{\cdot -compute^r}$, which is only ever used for the user's convenience.
+All the reasoning carried out in our type safety proofs is based on the algebraic properties introduced in \autoref{multiplicities} -- the exception to this is $\func{\cdot -compute^r}$, only ever used for the user's convenience.
 This algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose arrows with the same typing context and a matching output and input usage contexts.
 
-Leftover typing also allows us to prove a \emph{framing} theorem, which states that arbitrary usage annotations can be added to the input usage context, as long as they are added to the output usage context as well -- one can thus understand a typing derivation independently from its unused resources.
-An example use of this is showing that $\constr{comp-comm}$ preserves typing: we use framing to show that in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
+Leftover typing also allows us to state a \emph{framing} theorem showing that adding arbitrary usage annotations to the input and output usage contexts preserves typing -- one can understand a typing derivation independently from its unused resources.
+With framing one can show that $\constr{comp-comm}$ preserves typing: in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
 \begin{nitheorem}[Framing]
   \label{thm:framing}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
+  Let $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
   Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
-  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
+  Then for any $\Gamma_r$ and $\Xi_r$ where $\opctx{\Gamma_r}{\Delta}{\Xi_r}$ it holds that $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
 \end{nitheorem}
 
-Leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form, in which variables can freely be added or removed from context regardless of their usage annotation -- as long as this usage annotation is added or removed from the leftover context as well.
-Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
-Strengthening states that removing an unused variable preserves the well-typedness of a process.
+Leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form where linear variables can freely be added or removed from context too -- as long as they are added and remove to or from both input and output contexts.
 
@@
   \label{thm:weakening}
-  Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
+  Let $\func{ins}_i$ insert an element into a context at position $i$ --- we use it both to insert types into typing contexts and usage annotations into usage contexts.
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
@@
   \label{thm:strengthening} 
-  Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
+  Let $\func{del}_i$ delete the element at position $i$ from a context --- we use it both to delete types from typing contexts and usage annotations from usage contexts.
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
@@
 
-Subject congruence states that applying structural congruence (\autoref{def:equals}) to a well-typed process preserves its well-typedness.
-Subject reduction states that a well-typed process that communicates on a channel $c$ (\autoref{def:reduces}) preserves its well-typedness and consumes $\lio$ from $c$ if $c$ is a channel in the environment.
+Subject congruence states that structural congruence (\autoref{def:equals}) preserves the well-typedness of a process.
+Subject reduction states that reducing on a channel $c$ (\autoref{def:reduces}) preserves the well-typedness of a process (after consuming $\lio$ from $c$ if $c$ is an $\constr{external}$ channel).
+We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised type-safety proofs.
 
@@
   \label{thm:subject-reduction}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
-  \begin{itemize}
-    \item If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
-    \item If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
-  \end{itemize}
+  Let $\types{\gamma}{\Gamma}{P}{\Xi}$ and $P \reduce{c} Q$.
+  If $c$ is $\constr{internal}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
 \end{nitheorem}
-
-We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised type-safety proofs.
 

### 2021-01-25 paper/paper.tex
Put the unit type in, take base type out

@@
 The result is the courier $\func{system}$ defined below.
+We continue this example and provide typing derivations for these processes in \autoref{leftover-typing}, where we use a mix of linear, graded and shared typing to type $\func{system}$.
 \begin{align*}
@@
 \end{align*}
-
-We continue this example and by providing typing derivations for these processes in \autoref{leftover-typing}.
-There we use a mix of linear, graded and shared typing to type check $\func{system}$.
 \end{example}
@@
 \label{semantics}
-Thanks to our well-scoped grammar in \autoref{syntax}, the semantics of our language can now be defined on the totality of the syntax.
+Thanks to our well-scoped grammar in \autoref{syntax}, we now define the semantics of our language on the totality of the syntax.
 
@@
 \begin{nidefinition}[\textsc{Type} and \textsc{PreCtx}: types and typing contexts]
-  A \emph{type} is either a unit type ($\unit$), a base type ($\base{n}$) or a channel type ($\channel{t}{x}$).
+  A \emph{type} is either a unit type ($\unit$),
+  % a base type ($\base{n}$)
+  or a channel type ($\channel{t}{x}$).
   \begin{mathpar}
@@
     \inferrule
-    {n : \N}
-    {\base{n} : \Type}
+    { }
+    {\unit : \Type}
 
@@
 
-  The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as the unit type, $1$ as booleans) so that we avoid dealing with universe polymorphism.
+  The unit type $\unit$ serves as a proof of inhabitance for types.
+  % The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as the unit type, $1$ as booleans) so that we avoid dealing with universe polymorphism.
   The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.

### 2021-01-26 paper/paper.tex
Fix capitalisation in references

@@
 The first subterm in this chain of resources immediately knows what resources it has available.
-Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right uses leftover typing.
+Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right does not and uses leftover typing instead:
 \begin{mathpar}
@@
 \end{mathpar}
-\todo{Ornela: Say something about the rules and why you presented them here. Did you want to convey that the second is so much easier and elegant etc etc?}
 
 In this paper, we use leftover typing for the first time to define a resource-aware type system for the \picalc{} and fully mechanise it in Agda. 
-With regards to mechanisation, all previous work on mechanisation of linear process calculi follows the approach of context splits \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
+All previous work on the mechanisation of linear process calculi uses context splits \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
 

### 2021-01-26 paper/appendix.tex
Minor fixes from the reviews

@@
 
-In order to demonstrate the correspondence between an \picalc{} that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
+In order to demonstrate the correspondence between a \picalc{} that uses names and one that uses de Bruijn indices, we provide conversion functions in both directions and prove that they are inverses of each other up to $\alpha$-conversion.
 
@@
 As some widely-used languages do, this translation function produces unique variable names.
-These unique variable names use the naming scheme $<name-hint>^{<n>}$, where $^{<n>}$ denotes that the name $<name-hint>$ has already been bound $n$ times before.
+These unique variable names use the naming scheme $<namehint>^{<n>}$, where $^{<n>}$ denotes that the name $<namehint>$ has already been bound $n$ times before.
 \begin{equation*}
@@
 \begin{proof}[Proof]
-  All the above results are proved by induction on \textsc{Process} and \textsc{Var} (\autoref{fig:syntax}).
+  All the above results are proved by induction on \textsc{Process}, \textsc{Var}, and \textsc{Raw} (\autoref{fig:syntax}).
   Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
@@
   \label{lm:types-unused}
-  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equivalent to $\Xi$ at $i$.
+  For all well-typed processes $\types{\gamma}{\Gamma}{P}{\Xi}$, if the variable $i$ is unused within $P$, then $\Gamma$ at $i$ is equal to $\Xi$ at $i$.
 \end{nilemma}
@@
     \item
-      Case $\constr{scope-ext}$ for $\new \comp{P}{Q}$: we to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
+      Case $\constr{scope-ext}$ for $\new (\comp{P}{Q})$: we need to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
       In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).

### 2021-01-26 paper/paper.tex
Minor fixes from the reviews

@@
 \newcommand{\opctx}[3]{#1 \, \func{\coloneqq} \, #2 \, \func{\otimes} \, #3}
-\newcommand{\zero}{\func{0\cdot}}
-\newcommand{\one}{\func{1\cdot}}
+\newcommand{\zero}{\func{\cdot-0}}
+\newcommand{\one}{\func{\cdot-1}}
 \newcommand{\li}{\func{\ell_i}}
@@
 process $\new{} P$ creates a new channel and binds it at index $0$ in the continuation process $P$;
-$\comp{P}{Q}$ composes processes $P$ and $Q$ in parallel;
+process $\comp{P}{Q}$ composes $P$ and $Q$ in parallel;
 process $\recv{x}{}P$ receives data along channel $x$ and makes that data available at index $0$ in the continuation process $P$;
@@
 \begin{align*}
-  & \func{send} \; c = \new{} \send{(\suc c)}{\constr{0}} \PO &&
-    \func{recv} \; c \; = \recv{c}{} \recv{(\suc c)} \PO
+  & \func{send} \; c = \new{} (\send{\suc c}{\constr{0}} \PO) &&
+    \func{recv} \; c \; = \recv{c}{} \recv{(\suc c)}{} \PO
 \end{align*}
@@
 The courier role is defined below as $\func{carry} \; x \; y \; z$.
-It sequentially receives on the two input channels $x$ and $y$ -- instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$ -- instantiating $out$.
+It sequentially receives on the two input channels $x$ and $y$, instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$, instantiating $out$.
 Finally, we create three communication channels and compose all four processes together: the first channel is shared between the one sender and the courier, the second between the other sender and the courier, and the third between the receiver  and the courier.
@@
   \begin{aligned}
-    &\zero                   &:{} &                 &        & C \\
-    &\one                    &:{} &                 &        & C \\
     &\op{\_}{\_}{\_}         &:{} &                 &        & C \to C \to C \to \Set \\
-    &\func{\cdot-compute^r}  &:{} &\forall x y       & \to \; & \type{DEC} \; (\type{\exists} z  \; (\op{x}{y}{z})) \\
     &\func{\cdot-unique}     &:{} &\forall x x' y z  & \to \; & \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
     &\func{\cdot-unique^l}   &:{} &\forall x y y' z  & \to \; & \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
+    &\func{\cdot-assoc}      &:{} &\forall x y z u v & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z}) \\
+    &\func{\cdot-comm}       &:{} &\forall x y z     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
+    &\func{\cdot-compute^r}  &:{} &\forall x y       & \to \; & \type{DEC} \; (\type{\exists} z  \; (\op{x}{y}{z})) \\
+    &\zero                   &:{} &                 &        & C \\
+    &\func{\cdot-id^l}       &:{} &\forall x         & \to \; & \op{x}{\zero}{x} \\
     &\func{\cdot-min^l}      &:{} &\forall y z        & \to \; & \op{\zero}{y}{z} \to y \equiv \zero \\
-    &\func{\cdot-id^l}       &:{} &\forall x         & \to \; & \op{x}{\zero}{x} \\
-    &\func{\cdot-comm}       &:{} &\forall x y z     & \to \; & \op{x}{y}{z} \to \op{x}{z}{y} \\
-    &\func{\cdot-assoc}      &:{} &\forall x y z u v & \to \; & \op{x}{y}{z} \to \op{y}{u}{v} \to \\
-    &                        &    &                  &        & \type{\exists} w  \; (\op{x}{u}{w} \times \op{w}{v}{z})
+    &\one                    &:{} &                 &        & C \\
   \end{aligned}
@@
   % The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as the unit type, $1$ as booleans) so that we avoid dealing with universe polymorphism.
-  The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel.
+  The type $\channel{t}{x}$ of a channel determines what type ${t}$ of data and what usage annotations ${x}$ are sent over that channel --- we use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
   This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the input and output multiplicities, respectively \cite{K07}.
@@
   A \emph{usage context} is a context $\Ctx_{idxs}$ indexed by a context of indices $idxs : \Idxs_n$ that is either empty ($\constr{[]}$) or the result or appending a usage annotation pair $u : \Usage_{idx}^{\func{2}}$ with index $idx : \Idx$ to an existing context ($\Gamma \constr{,} u$).
-  We use the notation $\type{C}^{\func{2}}$ to stand for a $\type{C} \constr{\times} \type{C}$ pair of input and output multiplicities, respectively.
-  Henceforth, we use $\lz$ to denote the multiplicity pair $0 \comma 0$, $\li$ for the pair $1 \comma 0$, $\lo$ for $0 \comma 1$, and $\lio$ for $1 \comma 1$.
+  Henceforth, we use $\lz$ to denote the multiplicity pair $\zero \comma \zero$, $\li$ for the pair $\one \comma \zero$, $\lo$ for $\zero \comma \one$, and $\lio$ for $\one \comma \one$.
   This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
@@
     \inferrule
-    {loc_i : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
-    {\suc \; loc_i : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
+    {v : \contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}}
+    {\suc \; v : \contains{\gamma \comma t'}{\Gamma \comma x'}{\suc i}{t}{x}{\Delta \comma x'}}
   \end{mathpar}
@@
         {\stacked{
-            chan_i : \contains{\gamma \hspace{1.0em}}{\Gamma \hspace{1.0em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
-            cont \hspace{0.5em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.2em}}{\Theta \comma \lz}}}
-        {\recv{chan_i}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
+            chan : \contains{\gamma \hspace{1.0em}}{\Gamma \hspace{1.0em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
+            cont \hspace{0.3em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{3.8em}}{\Theta \comma \lz}}}
+        {\recv{chan}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}
 
@@
         {\stacked{
-            chan_i : \contains{\gamma}{\Gamma \hspace{0.1em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
-            loc_j \hspace{0.7em} : \contains{\gamma}{\Delta}{j}{t \hspace{2.8em}}{x \hspace{0.2em}}{\Xi} \\\\
-            cont \hspace{0.5em} : \types{\gamma}{\Xi}{\hspace{0.4em}P\hspace{4.0em}}{\Theta}}}
-        {\send{chan_i}{loc_j} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
+            chan : \contains{\gamma}{\Gamma \hspace{0.1em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
+            loc \hspace{0.9em} : \contains{\gamma}{\Delta}{j}{t \hspace{2.8em}}{x \hspace{0.2em}}{\Xi} \\\\
+            cont \hspace{0.3em} : \types{\gamma}{\Xi}{\hspace{0.4em}P\hspace{3.8em}}{\Theta}}}
+        {\send{chan}{loc} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}
 
@@
    & \func{sendWT} \;  : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}  }{\Gamma \comma (l \comma \suc r)  }{\func{send} \; \constr{0}  }{\Gamma \comma (l \comma r)}  \\
-   & \func{sendWT} \;  = \; \new{} \; \_ \; \_ \; \zero \; \send{(\suc \constr{0})}{\constr{0}} \PO
+   & \func{sendWT} \;  = \; \new{} \; \_ \; \_ \; \zero \; (\send{\suc \constr{0}}{\constr{0}} \PO)
 \end{align*}

### 2021-01-26 paper/appendix.tex
Fix appendices

@@
 
-\begin{figure}[h]
+The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar below:
 \begin{mathpar}
@@
 \end{equation*}
-  \caption{Grammar with names.}
-  \label{fig:syntax-names}
-\end{figure}
-
-The syntax of the \picalc{} \cite{Sangio01} using channel names is given by the $\Raw$ grammar in \autoref{fig:syntax-names}.
 
@@
 \begin{example}[$\func{fromRaw}$ and $\func{toRaw}$]
-  \begin{figure}[t]
+  We illustrate the conversion functions from names to de Bruijn indices $(\func{fromRaw}$) and back ($\func{toRaw}$) with three processes $P,Q,R$  below.
   \begin{alignat*}{7}
@@
   \end{alignat*}
-    \caption{From names to de Bruijn indices and back}
-    \label{fig:conversion}
-    \end{figure}
-  We illustrate the conversion functions from names to de Bruijn indices $(\func{fromRaw}$) and back ($\func{toRaw}$) with three processes $P,Q,R$  in \autoref{fig:conversion}.
 
@@
 \begin{proof}[Proof]
-  All the above results are proved by induction on \textsc{Process}, \textsc{Var}, and \textsc{Raw} (\autoref{fig:syntax}).
+  All the above results are proved by induction on \textsc{Process}, \textsc{Var}, and \textsc{Raw} (\autoref{def:syntax}).
   Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
@@
 
-\subsection{Auxiliary Results}
-We will start this section with auxiliary results which are needed to prove two main theorems: subject congruence (\autoref{thm:subject-congruence}) and subject reduction (\autoref{thm:subject-reduction}) for our \picalc{} with leftovers.
-
-\paragraph*{Framing}
-This property asserts that the well-typedness of a process is independent of its leftover resources.
-
-\begin{nitheorem}[Framing]
-  \label{thm:framing}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
-  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
-  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
-  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
-\end{nitheorem}
-
-\paragraph*{Weakening}
-This property states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
-Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
-\begin{nitheorem}[Weakening]
-  \label{thm:weakening}
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Then, lifting every variable greater than or equal to $i$ in $P$ is well typed in
-  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
-\end{nitheorem}
-
-\paragraph*{Strengthening}
-This property states that removing an unused variable preserves the well-typedness of a process.
-Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
-\begin{nitheorem}[Strengthening]
-  \label{thm:strengthening} 
-  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
-  Let $i$ be a variable not in $P$, such that $uP \; : \; \Unused_i \; P$.
-  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P \; uP}{\func{del}_i \; \Xi}$.
-\end{nitheorem}
-
 \paragraph*{Exchange}
 This property states that the exchange of two variables preserves the well-typedness of a process.
-We extend $\func{exchange}_i$  introduced in \autoref{structural-congruence} to exchange types in typing contexts and usage annotations in usage contexts.
+We extend $\func{exchange}_i$  introduced in \autoref{def:equals} to exchange types in typing contexts and usage annotations in usage contexts.
 \begin{nitheorem}[Exchange]
@@
 
-\subsection{Towards Subject Reduction}
 \paragraph*{Subject Congruence}
-This property states that applying structural congruence (\autoref{structural-congruence}) to a well-typed process preserves its well-typedness.
+This property states that applying structural congruence (\autoref{def:equals}) to a well-typed process preserves its well-typedness.
 To prove this result, we must first introduce lemmas that establish that certain syntactic manipulations can be inverted (\autoref{lm:lower-lift}, \autoref{lm:exchange-exchange}) and how unused variables relate to the preservation of leftovers (\autoref{lm:types-unused}).
@@
 \begin{nitheorem}[Subject congruence]
-  \label{thm:subject-congruence}
+  \label{thm:subject-congruence1}
   If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
@@
 \begin{nitheorem}[Generalised substitution]
-  \label{thm:subst-generalization1}
+  \label{thm:substitution-generalization}
   Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
       We can now make two inductive calls (on the derivation of $P$ and $Q$) and compose their results.
-  \end{itemize}  
-\end{proof}
 
-\begin{figure}[h]
+      \begin{minipage}[b]{0.5\linewidth}
         \centering
@@
         \end{tikzpicture}
-    \caption{
+
         Diagrammatic representation of the $\comp{}{}$ case for substitution.
         Continuous lines represent known facts, dotted lines proof obligations.
-    }
-    \label{fig:subst}
-\end{figure}
-
-\begin{figure}[h]
+      \end{minipage}
+      \begin{minipage}[b]{0.5\linewidth}
         \centering
@@
         \end{tikzpicture}
-    \caption{
+
         Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
         This approach makes composing typing derivations before and after subsitution more difficult.
-    }
-    \label{fig:subst-alternative}
-\end{figure}
+      \end{minipage}
+
+  \end{itemize}  
+\end{proof}
 
@@
 Finally we are ready to present our main result, stating  that if $P$ is well typed and it reduces to $Q$, then $Q$ is well typed.
-The relation between the typing contexts used to type $P$ and $Q$ will be explained in \autoref{thm:subject-reduction}.
-In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{operational-semantics}).
+The relation between the typing contexts used to type $P$ and $Q$ will be explained in \autoref{thm:subject-reduction1}.
+In the \picalc{} we distinguish between a reduction $P \reduce{\constr{internal}} Q$ on a channel internal to $P$, and a reduction $P \reduce{\constr{external} \; i} Q$ on a channel $i$ external to $P$ (refer to \autoref{semantics}).
 We first introduce an auxiliary lemma:
@@
 \begin{nitheorem}[Subject reduction]
-  \label{thm:subject-reduction}
+  \label{thm:subject-reduction1}
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$ and reduce such that $P \reduce{c} Q$.
@@
     \item
-    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence}) and proceed inductively. \qed
+    Case $\constr{struct}$: we apply subject congruence (\autoref{thm:subject-congruence1}) and proceed inductively. \qed
   \end{itemize}

### 2021-01-26 paper/paper.tex
Fix appendices

@@
 
-\begin{nidefinition}[\textsc{Var} and \textsc{Process}]
+\begin{nidefinition}[\textsc{Var} and \textsc{Process}]\label{def:syntax}
 A variable reference occurring under $n$ binders can refer to $n$ distinct variables.

### 2021-01-26 paper/paper.tex
Fix ald clarify typing rules

@@
 %
-  The input process in rule $\recv{}{}$ requires a channel $chan_i$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  The input process in rule $\recv{}{}$ requires a channel $chan$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
   Note that the index $i$ is used in the syntax of the typed process.
@@
 %
-  Similarly to input, the output process in rule $\send{}{}$ requires a channel $chan_i$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
-  We use the leftover context $\Delta$ to type the transmitted data, which needs an element $loc_j$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan_i$.
+  Similarly to input, the output process in rule $\send{}{}$ requires a channel $chan$ at index $i$ with usage $\lo$ available, such that data with type $t$ and usage $x$ can be sent over it.
+  We use the leftover context $\Delta$ to type the transmitted data, which needs an element $loc$ at index $j$ with type $t$ and usage $x$, as per the type of the channel $chan$.
   The leftovers $\Xi$ are used to type the continuation process.
@@
   Parallel composition in rule $\comp{}{}$ uses the leftovers of the left-hand process to type the right-hand process.
-  By keeping track in the typing derivation of $P$ of the resources $P$ uses, we can use them to type $Q$ and save the user from having to provide a top-down proof of context split.
+  Indeed, \autoref{thm:subject-congruence} shows that the alternative rule where the resources are threaded through $Q$ first is admissible too.
 \end{nidefinition}

### 2021-01-26 paper/paper.tex
Clarify context splitting argument to var

@@
 \begin{nidefinition}[\textsc{Algebra}]
-  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning), \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition), and in which the leftovers can be \emph{computed} (as to automatically update the usage context every time input and output occurs).
+  A \emph{usage algebra} is a ternary relation $\op{x}{y}{z}$ that is \emph{partial} (as not any two multiplicities can be combined), \emph{deterministic} and \emph{cancellative} (to aid equational reasoning) and \emph{associative} and \emph{commutative} (following directly from subject congruence for parallel composition).
+  In addition, we ask that the leftovers can be \emph{computed} so that we can automatically update the usage context every time input and output occurs --- this is purely for usability.
   It has a \emph{neutral element} $\zero$ that is absorbed on either side, and that is also \emph{minimal} (so that new resources cannot arbitrarily spring into life).
@@
   Below we define such an algebra as a record $\Algebra_C$ on a carrier $C$.
-  We use $\forall$ for universal quantification.
+  (We use $\forall$ for universal quantification.
   The dependent product $\type{\exists}$ uses the value of its first argument in the type of its second.
-  The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.
+  The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.)
 
@@
 
-  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$ (the leftovers).
+  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{(x_l \comma x_r)}{(y_l \comma y_r)}{(z_l \comma z_r)}$ on pairs of multiplicities.
+  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}^{\func{2}}$ into $y$ and $z$ (the leftovers).
   Note that the remaining context $\Gamma$ is preserved unused as a leftover.
   This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
-  We use the \emph{computability} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
+  In our Agda implementation, $\opsquared{x}{y}{z}$ is actually a trivially satisfiable implicit argument if $\opsquared{x}{y}{z}$ is inhabited and an unsatisfiable argument otherwise.
   The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.
 
-  We lift the operation $\op{x}{y}{z}$ and its algebraic properties to an operation $\opsquared{(x_l \comma x_r)}{(y_l \comma y_r)}{(z_l \comma z_r)}$ on pairs of multiplicities and an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts with the same underlying context of indices.
 \end{nidefinition}
@@
 
-All the reasoning carried out in our type safety proofs is based on the algebraic properties introduced in \autoref{multiplicities} -- the exception to this is $\func{\cdot -compute^r}$, only ever used for the user's convenience.
-This algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose arrows with the same typing context and a matching output and input usage contexts.
+All the reasoning carried out in our type safety proofs is based on the algebraic properties introduced in \autoref{multiplicities} -- the exception to this is $\func{\cdot -compute^r}$, only there for the user's convenience.
+  We lift the operation $\opsquared{x}{y}{z}$ and its algebraic properties to an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have the same underlying context of indices.
+The algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose arrows with the same typing context and a matching output and input usage contexts.
 
@@
   \label{thm:weakening}
-  Let $\func{ins}_i$ insert an element into a context at position $i$ --- we use it both to insert types into typing contexts and usage annotations into usage contexts.
+  Let $\func{ins}_i$ insert an element into a context at position $i$.
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
@@
   \label{thm:strengthening} 
-  Let $\func{del}_i$ delete the element at position $i$ from a context --- we use it both to delete types from typing contexts and usage annotations from usage contexts.
+  Let $\func{del}_i$ delete the element at position $i$ from a context.
   Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.

### 2021-01-26 paper/appendix.tex
Apply remaining fixes

@@
 
-\begin{proof}[Proof]
-  All the above results are proved by induction on \textsc{Process}, \textsc{Var}, and \textsc{Raw} (\autoref{def:syntax}).
+\begin{proof}
+  All the above results are proved by induction on \textsc{Process}, \textsc{Var} (\autoref{def:syntax}) and \textsc{Raw}.
   Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
@@
 
-\begin{proof}[Proof]
+\begin{proof}
   All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
@@
 
-\begin{proof}[Proof]
+\begin{proof}
   The proof is by induction on \textsc{Equals} $\eq{}$.
@@
   Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
-  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
+  The substituted variable at position $i$ can be split into $m$ in $\Gamma_i$, and into $n$ in $\Psi_i$.
   Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
-  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
+  In other words, let there be some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
   \begin{multicols}{2}
@@
   Then substituting $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
-  Refer to \autoref{fig:subst} for a diagramatic representation.
 \end{nitheorem}
 
-\begin{proof}[Proof]
+\begin{proof}
   By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
@@
 
-      \begin{minipage}[b]{0.5\linewidth}
       \centering
@@
       Continuous lines represent known facts, dotted lines proof obligations.
-      \end{minipage}
-      \begin{minipage}[b]{0.5\linewidth}
-        \centering
-        \begin{tikzpicture}
-          \node (gamma-i) at (0,2)   {};
-          \node (gamma-m) at (0,1)   {};
-          \node (gamma-j) at (0,0)   {};
-          \node (xi-i)    at (2,3)   {};
-          \node (xi-m)    at (2,2) {};
-          \node (xi-j)    at (2,1)   {};
-          \node (psi-m)   at (4,2) {};
-
-          \draw[->] (gamma-i) -- (xi-j);
-          \draw[->] (gamma-m) -- (xi-j);
-          \draw[->,densely dotted] (gamma-j) -- (xi-j);
-          \draw[->] (gamma-i) -- (gamma-m);
-          \draw[->] (gamma-j) -- (gamma-m);
-          \draw[->] (xi-i)    -- (xi-m);
-          \draw[->] (xi-j)    -- (xi-m);
-          \draw[->] (xi-i)    -- (psi-m);
-          \draw[->] (xi-m)    -- (psi-m);
-          \draw[->,densely dotted] (xi-j)    -- (psi-m);
-        \end{tikzpicture}
-
-        Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
-        This approach makes composing typing derivations before and after subsitution more difficult.
-      \end{minipage}
 
@@
 \end{nitheorem}
-\begin{proof}[Proof]
+\begin{proof}
   For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
@@
 \end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{note}
-  We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
-% \end{note}
 
@@
 
-\begin{proof}[Proof]
+\begin{proof}
   By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  

### 2021-01-26 paper/paper.tex
Apply remaining fixes
Last pass

@@
 \newcommand{\lo}{\func{\ell_o}}
-\newcommand{\lz}{\func{\ell_{\varepsilon}}}
+\newcommand{\lz}{\func{\ell_{\emptyset}}}
 \newcommand{\lio}{\func{\ell_{\#}}}
@@
   We provide a framing theorem for our type system, generalise the weakening and strengthening theorems to include linear types, and prove subject reduction.
-  Our formalisation is fully mechanised in about 1800 lines of Agda \cite{Zalakain2020Agda}.
+  Our formalisation is fully mechanised in about 1850 lines of Agda \cite{Zalakain2020Agda}.
 
@@
 A linear type system must keep track of what resources are used in which parts of the program, and guarantee that they are \emph{neither duplicated nor discarded}.
-To do so, the standard approach is to use context splits: typing rules for terms with multiple subterms add an extra a side condition specifying what resources to allocate to each of the subterms.
+To do so, the standard approach is to use context splits: typing rules for terms with multiple subterms add an extra side condition specifying what resources to allocate to each of the subterms.
 The typing derivations for the subterms must then use the entirety of their allocated resources.
@@
 
-An alternative approach is leftover typing, a technique which has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
-Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were not used by the term.
-As a result, typing rules \emph{thread the resources through} subterms without needing context splits: each subterm uses the resources it needs, and leaves the rest for its siblings.
+An alternative approach is leftover typing, a technique that has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
+Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were left unused by the term.
+As a result, typing rules \emph{thread} the resources \emph{through} subterms without needing context splits: each subterm uses the resources it needs, and leaves the rest for its siblings.
 The first subterm in this chain of resources immediately knows what resources it has available.
-Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right does not and uses leftover typing instead:
+Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right does not, and uses leftover typing instead:
 \begin{mathpar}
@@
     In addition to making context splits unnecessary, leftover typing allows for a \emph{framing} theorem to be stated and is naturally associative, making type safety properties considerably easier to reason about (\autoref{meta-theory}).
+    Thanks to leftover typing, we can now also define \emph{weakening} and \emph{strengthening} for the whole framework, not just the shared fragment.
 
@@
     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
-    Multiple algebras can be mixed in the type system --- usage contexts keep information about what algebra to for each type (\autoref{contexts}).
+    Multiple algebras can be mixed in the type system --- usage contexts keep information about what algebra to use for each type (\autoref{contexts}).
     In particular, this allows for type systems combining linear (use exactly once), graded (exact number of $n$ times) and shared (free to reuse and discard) types under the same framework.
-    Thanks to leftover typing, we can now define \emph{weakening} and \emph{strengthening} for the whole framework, not just the shared fragment.
 
-    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in Appendix \ref{app:type-safety}
-The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our GitHub repository \cite{Zalakain2020Agda}.
+    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in Appendix \ref{app:type-safety}.
+The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our \cite{Zalakain2020Agda}.
 
@@
 We then define an interface for resource-aware algebras (\autoref{multiplicities}) and use it to parametrise a type system based on leftover typing (\autoref{leftover-typing}).
-Finally, we present the meta theoretical properties of our \picalc{} with leftovers in \autoref{meta-theory}.
+Finally, we present the meta theoretical properties of our type system in \autoref{meta-theory}.
 
@@
 Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.
-Constructors ($\constr{0}$ and $\suc$) are given as inference rules.
+Constructors ($\constr{0}$ and $\suc$) are used as inference rule names.
 We maintain a close correspondence between the definitions presented in this paper and our mechanised definitions in Agda: inference rules become type constructors, premises become argument types and conclusions return types.
@@
 A variable reference occurring under $n$ binders can refer to $n$ distinct variables.
-References outside of that range are meaningless, and it is useful to rule them out syntactically.
-We do so by introducing the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
+We introduce the indexed family of types \cite{Dybjer1994} $\Var_n$: for all naturals $n$, the type $\Var_n$ has $n$ distinct elements.
 We index processes according to their \emph{depth}: for all naturals $n$, a process of type $\Process_n$ contains free variables that can refer to $n$ distinct elements.
@@
 
-As an example use of this syntax, we present a courier system that consists of three \emph{roles}:
+We present a courier system that consists of three \emph{roles}:
 a sender, who wants to send a package;
@@
 The sender creates a new channel to be sent as data, and sends it over channel $c$, and then terminates.
-Each sender process $\func{send} \, x$ and $\func{send} \, y$ is an instantiation of $\func{send} \, c$.
+Processes $\func{send} \, x$ and $\func{send} \, y$ are an instantiation of $\func{send} \, c$.
 The receiver receives data \emph{twice} on a channel $c$ and then terminates.
@@
 The courier role is defined below as $\func{carry} \; x \; y \; z$.
-It sequentially receives on the two input channels $x$ and $y$, instantiating $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$, instantiating $out$.
+It sequentially receives on the two input channels $x$ and $y$, instantiated as $in0$ and $in1$, and then outputs the two pieces of received data on the output channel $z$, instantiated as $out$.
 Finally, we create three communication channels and compose all four processes together: the first channel is shared between the one sender and the courier, the second between the other sender and the courier, and the third between the receiver  and the courier.
@@
   We consider a variable $i$ to be unused in $P$ ($\Unused_i \; P$) if none of the inputs nor the outputs refer to it.
-  $Unused_i \; P$ is defined as a recursive predicate on $P$, incrementing $i$ every time we go under a binder, and using $i \type{\not\equiv} x$ ( which unfolds to the negation of propositional equality on \textsc{Var}, i.e.
+  $\Unused_i \; P$ is defined as a recursive predicate on $P$, incrementing $i$ every time we go under a binder, and using $i \type{\not\equiv} x$ ( which unfolds to the negation of propositional equality on \textsc{Var}, i.e.
 $i \type{\equiv} x \to \type{\bot}$) to compare variables.
@@
   { }
-  {\constr{comp-end} : \comp{P}{\PO_n} \eqeq P}
+  {\constr{comp-id} : \comp{P}{\PO_n} \eqeq P}
 
@@
 As a result of that communication, the continuation of the input process $P$ has all the references to its most immediate variable substituted with references to $\suc j$, the variable sent by the output process $\send{i}{j}{Q}$.
-After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (and a proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$ is constructed).
+After this substitution, $\subst{P}{\suc j}{\constr{0}}$ is \emph{lowered} --- all variable references are decreased by one (and we derive the proof $\Unused_{\constr{0}} \; (\subst{P}{\suc j}{\constr{0}})$).
 Reduction is closed under parallel composition (rule $\constr{par}$), restriction (rule $\constr{res}$) and structural congruence (rule $\constr{struct}$) 
@@
 
-  The unit type $\unit$ serves as a proof of inhabitance for types.
+  The unit type $\unit$ serves as a base case for types.
   % The base type $\base{n}$ uses natural numbers as placeholders for types (the host language can then interpret the former into the latter --- e.g., $0$ as the unit type, $1$ as booleans) so that we avoid dealing with universe polymorphism.
@@
   This channel notation aligns with $[t]\ \mathbf{chan}_{(i^{y},o^{z})}$, where $y,z$ are the input and output multiplicities, respectively \cite{K07}.
-
+  Henceforth, we use $\lz$ to denote the multiplicity pair $\zero \comma \zero$, $\li$ for the pair $\one \comma \zero$, $\lo$ for $\zero \comma \one$, and $\lio$ for $\one \comma \one$.
+  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
   A \emph{typing context} $\PreCtx_n$ is a length-indexed list of types that is either empty ($\constr{[]}$) or the result of appending a type $t : \Type$ to an existing context ($\gamma \constr{,} t$).
@@
   A \emph{usage context} is a context $\Ctx_{idxs}$ indexed by a context of indices $idxs : \Idxs_n$ that is either empty ($\constr{[]}$) or the result or appending a usage annotation pair $u : \Usage_{idx}^{\func{2}}$ with index $idx : \Idx$ to an existing context ($\Gamma \constr{,} u$).
-  Henceforth, we use $\lz$ to denote the multiplicity pair $\zero \comma \zero$, $\li$ for the pair $\one \comma \zero$, $\lo$ for $\zero \comma \one$, and $\lio$ for $\one \comma \one$.
-  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
 \end{nidefinition}
@@
 
-Leftover typing inverts the information flow concerning usage annotations so that it is the typing derivations of subprocesses which determine how resources are allocated.
+Leftover typing inverts the information flow of usage annotations so that it is the typing derivations of subprocesses which determine how resources are allocated.
 As a result, context split proofs are no longer necessary.
-Leftover typing also allows \emph{framing} to be stated, and \emph{weakening} and \emph{strengthening} to acquire a more general form.
+Leftover typing also allows \emph{framing} to be stated, and \emph{weakening} and \emph{strengthening} to cover linear types too.
 
@@
 \begin{example}[Variable reference]
-  We type a variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.
+  In $\func{egVar}$ we type a variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.
+  We must show that the variable with index $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)$ and a usage context $(\constr{[]} \comma \lio \comma \lio)$.
+  The \textsc{VarRef} constructors that we must use are completely determined by the variable index $\suc \; \constr{0}$ in the type.
+  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to automatically fulfill this proof obligation.
   \begin{flalign*}
@@
   \end{flalign*}
-  We must show that the variable with index $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)$ and a usage context $(\constr{[]} \comma \lio \comma \lio)$.
-  The \textsc{VarRef} constructors that we must use are completely determined by the variable index $\suc \; \constr{0}$ in the type.
-  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to automatically fulfill this proof obligation.
 \end{example}
@@
   The input process in rule $\recv{}{}$ requires a channel $chan$ at index $i$ with usage $\li$ available, such that data with type $t$ and usage $x$ can be sent over it.
-  Note that the index $i$ is used in the syntax of the typed process.
+  Note that the index $i$ is determined by the syntax of the typed process.
   We use the leftovers $\Xi$ to type the continuation process, which is also provided with the received element --- of type $t$ and multiplicity $x$ --- at index $\constr{0}$.
@@
   The leftovers $\Xi$ are used to type the continuation process.
-  Note that both indices $i$ and $j$ are used in the syntax of the typed process.
+  Note that both indices $i$ and $j$ are determined by the syntax of the typed process.
 %
   Parallel composition in rule $\comp{}{}$ uses the leftovers of the left-hand process to type the right-hand process.
-  Indeed, \autoref{thm:subject-congruence} shows that the alternative rule where the resources are threaded through $Q$ first is admissible too.
+  Indeed, \autoref{thm:subject-congruence} shows that an alternative rule where the resources are first threaded through $Q$ is admissible too.
 \end{nidefinition}
@@
   We lift the operation $\opsquared{x}{y}{z}$ and its algebraic properties to an operation $\opctx{\Gamma}{\Delta}{\Xi}$ on usage contexts that have the same underlying context of indices.
-The algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose arrows with the same typing context and a matching output and input usage contexts.
+The algebraic properties of the algebras allow us to see a typing derivation $\types{\gamma}{\Gamma}{P}{\Delta}$ as a unique \emph{arrow} from $\Gamma$ to $\Delta$, and to freely compose and reason with arrows with the same typing context and a matching output and input usage contexts.
 
-Leftover typing also allows us to state a \emph{framing} theorem showing that adding arbitrary usage annotations to the input and output usage contexts preserves typing -- one can understand a typing derivation independently from its unused resources.
+Leftover typing also allows us to state a \emph{framing} theorem showing that adding or subtracting arbitrary usage annotations to the input and output usage contexts preserves typing -- one can understand a typing derivation independently from its unused resources.
 With framing one can show that $\constr{comp-comm}$ preserves typing: in $\comp{P}{Q}$ the typing of $P$ and $Q$ is independent of one another.
@@
 
-Leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form where linear variables can freely be added or removed from context too -- as long as they are added and remove to or from both input and output contexts.
+Leftover typing allows \emph{weakening} and \emph{strengthening} to acquire a more general form where linear variables can freely be added or removed from context too -- as long as they are added and removed to and from both the input and output contexts.
 
@@
 Subject congruence states that structural congruence (\autoref{def:equals}) preserves the well-typedness of a process.
-Subject reduction states that reducing on a channel $c$ (\autoref{def:reduces}) preserves the well-typedness of a process (after consuming $\lio$ from $c$ if $c$ is an $\constr{external}$ channel).
+Subject reduction states that reducing on a channel $c$ (\autoref{def:reduces}) preserves the well-typedness of a process --- after consuming $\lio$ from $c$ if $c$ is an $\constr{external}$ channel.
 We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised type-safety proofs.
@@
   If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
+  (We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.)
 \end{nitheorem}
@@
 To show subject congruence, one has to prove that $P \eqeq \comp{P}{\constr{!}P}$ preserves typing in both directions.
-To do so the resource consumption of a process has to be determined by e.g. adding type annotations to the syntax.
+To do so the resource consumption of a process has to be uniquely determined by e.g. adding type annotations to the syntax.
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.

### 2021-01-27 paper/paper.tex
pass on abstract

@@
 \begin{abstract}
-  Linear type systems must keep track of how programs use their resources.
-  The standard approach is to use \emph{context splits} specifying how resources are disjointly split across subterms.
-  In this approach, context splits echo information which is already present within subterms, thus leading to redundancy.
+  Linear type systems need to keep track of how programs use their resources.
+  The standard approach is to use \emph{context splits} specifying how resources are (disjointly) split across subterms.
+  In this approach, context splits redundantly echo information which is already present within subterms.
 An alternative approach is to use \emph{leftover typing} \cite{Mackie,Allais2018a}, where in addition to the usual (input) usage context, typing judgments have also an output usage context: the leftovers.
-  In this approach, the leftovers of one typing derivation are fed as the input of the next, threading through linear resources while avoiding context splits.
-  We use leftover typing to define a type system for a \emph{resource-aware \picalc{}} \cite{MilnerPW92,Milner99}.
-  This type system is parametrised over a set of usage algebras that are general enough to encompass shared types (free to reuse and discard), graded types (use an $n$ number of times) and linear types (use exactly once).
-   Linear types are of fundamental importance in the \picalc{}: they ensure privacy and safety of communication and avoid race conditions.
+  In this approach, the leftovers of one typing derivation are fed as input to the next, threading through linear resources while avoiding context splits.
+  We use leftover typing to define a type system for a {resource-aware \picalc{}} \cite{MilnerPW92,Milner99}, a process algebra used to  model concurrent systems.
+  Our type system is parametrised over a set of \emph{usage algebras} that are general enough to encompass \emph{shared types} (free to reuse and discard), \emph{graded types} (use exactly $n$ number of times) and \emph{linear types} (use exactly once).
+   Linear types are important in the \picalc{}: they ensure privacy and safety of communication and avoid race conditions, while graded and shared types allow for more flexible programming.
   We provide a framing theorem for our type system, generalise the weakening and strengthening theorems to include linear types, and prove subject reduction.

### 2021-01-27 paper/paper.tex
pass on intro

@@
 Notably, it features channel mobility: channels themselves are first class values and can be sent and received.
-Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} with linear channel types, where channels have to be used exactly once.
-Linearity guarantees privacy and safety of communication and avoids race conditions.
+Kobayashi et al. \cite{KPT96} introduced a typed version of the \picalc{} with linear channel types, where channels must be used \emph{exactly} once.
+Linearity in the \picalc{} guarantees privacy and safety of communication and avoids race conditions.
 
+More broadly, linearity allows for resource-aware programming and more \emph{efficient} implementations \cite{Wadler90}, and it  inspired unique types (as in Clean \cite{BarendsenS96}), and ownership types (as in Rust \cite{MatsakisK14}).
 A linear type system must keep track of what resources are used in which parts of the program, and guarantee that they are \emph{neither duplicated nor discarded}.
@@
 
-An alternative approach is leftover typing, a technique that has been used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc \cite{Allais2018a}.
+An alternative approach is \emph{leftover typing}, a technique used to formulate intuitionistic linear logic \cite{Mackie} and to mechanise the linear \lambdacalc~\cite{Allais2018a}.
 Leftover typing changes the shape of the typing judgments and includes a second \emph{leftover} output context that contains the resources that were left unused by the term.
@@
 The first subterm in this chain of resources immediately knows what resources it has available.
-Below we present two alternative typing rules for process composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right does not, and uses leftover typing instead:
+
+In this paper, we use leftover typing to define for the first time a resource-aware type system for the \picalc{}, and we fully mechanise our work in Agda \cite{Zalakain2020Agda}. 
+All previous work on mechanisation of linear process calculi uses context splits instead \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
+We will further highlight the benefits of leftover typing as opposed to context splits in contributions and the rest of the paper.
+
+Below we present two alternative typing rules for parallel composition in the linear \picalc{}: the one on the left uses context splits, while the one on the right does not, and uses leftover typing instead:
 \begin{mathpar}
@@
 
-In this paper, we use leftover typing for the first time to define a resource-aware type system for the \picalc{} and fully mechanise it in Agda. 
-All previous work on the mechanisation of linear process calculi uses context splits \cite{Gay2001,Goto2016a,Gay2010,Thiemann2019,Ciccone}.
-
-We make the following main contributions:
+\paragraph*{Contributions and Structure of the Paper}
 \begin{enumerate}
-  \item \textbf{Leftover typing for a resource-aware \picalc{}}. Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}) and satisfies subject reduction.
-    In addition to making context splits unnecessary, leftover typing allows for a \emph{framing} theorem to be stated and is naturally associative, making type safety properties considerably easier to reason about (\autoref{meta-theory}).
-    Thanks to leftover typing, we can now also define \emph{weakening} and \emph{strengthening} for the whole framework, not just the shared fragment.
-
+  \item \textbf{Leftover typing for resource-aware \picalc{}}.
+Our type system uses leftover typing to model the resource-aware \picalc{} (\autoref{leftover-typing}) and satisfies subject reduction (\autoref{thm:subject-reduction}).
+    In addition to making context splits unnecessary, leftover typing allows for a \emph{framing} theorem (\autoref{thm:framing}) to be stated and is naturally associative, making type safety properties considerably easier to reason about (\autoref{meta-theory}).
+    Thanks to leftover typing, we can now state \emph{weakening} (\autoref{thm:weakening}) and \emph{strengthening} (\autoref{thm:strengthening}) for the whole framework, not just the shared fragment.
+This give a uniform and complete presentation of all the meta-theory for the resource-aware \picalc{}.
   \item \textbf{Shared, graded and linear unified \picalc{}}.
@@
 
-    \item \textbf{Full mechanisation in Agda}. We have fully mechanised a proof of subject reduction, the details of which are found in Appendix \ref{app:type-safety}.
-The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available in our \cite{Zalakain2020Agda}.
-
+    \item \textbf{Full mechanisation in Agda}.
+The formalisation of the \picalc{} with leftover typing, from the syntax to the semantics and the type system, has been fully mechanised in Agda in about 1850 lines of code, and is publicly available at \cite{Zalakain2020Agda}.
+We have fully mechanised all meta-theory and the details of a proof of subject reduction can be found in Appendix \ref{app:type-safety}.
 \end{enumerate}
 
-We start by using type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to introduce a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
+We use type level {de Bruijn indices} \cite{deBruijn1972,Dybjer1994} to define a syntax of \picalc{} processes that is \emph{well scoped by construction}: every free variable is accounted for in the type of the process that uses it (\autoref{syntax}).
 We then provide an operational semantics for the \picalc{}, prior to any typing (\autoref{semantics}).
@@
 The reduction relation tracks at the type level the channel on which communication occurs.
-This information is later used to state the subject reduction theorem --- aka type preservation.
+This information is later used to state the subject reduction theorem.
 The reduction relation is defined modulo \emph{structural congruence} --- a relation defined on processes that acts as a quotient type to remove unnecessary syntactic minutiae introduced by the syntax of the \picalc{}.
@@
 
-We briefly illustrate the notation used in this paper.
+%We briefly illustrate the notation used in this paper.
 Data type definitions ($\type{\N}$) use double inference lines and index-free synonyms (\textsc{Nat}) as rule names for ease of reference.

### 2021-01-27 paper/paper.tex
a last pass to the rest of the paper

@@
 The result is the courier $\func{system}$ defined below.
-We continue this example and provide typing derivations for these processes in \autoref{leftover-typing}, where we use a mix of linear, graded and shared typing to type $\func{system}$.
 \begin{align*}
@@
 \end{align*}
+
+We continue this running example in  \autoref{leftover-typing}, where we provide typing derivations for the above processes and use a mix of linear, graded and shared typing to type the courier $\func{system}$.
 \end{example}
@@
 
-We distinguish between channels that are created inside of the process ($\constr{internal}$), and channels that are created outside ($\constr{external}\ i$), where $i$ is the index of the channel variable.
+We distinguish between channels that are created inside the process ($\constr{internal}$), and channels that are created outside ($\constr{external}\ i$), where $i$ is the index of the channel variable.
 In rule $\constr{comm}$, parallel processes reduce when they communicate over a common channel with index ${i}$.
@@
 Rule $\constr{res}$ uses $\func{dec}$ to decrement the index of channel $c$ as we wrap processes $P$ and $Q$ inside a binder.
-It is defined as follows:
+It is defined as expected below:
 \begin{equation*}
@@
 We sketch the implementation of linear, graded and shared types as instances of our usage algebra below.
-Their use in typing derivations is illustrated in \autoref{example-derivation} (Continued).
+Their use in typing derivations is illustrated in \autoref{example-derivation}.
 \begin{center}
@@
 \begin{nidefinition}[\textsc{Types}: typing processes] \label{def:types}
-  The \textsc{Types} typing relation for \picalc{} processes is presented below.
+  The \textsc{Types} typing relation for the resource-aware \picalc{} processes is presented below.
   For convenience, we reuse the constructor names introduced for the syntax in \autoref{syntax}.
@@
 Dually, the courier defined by the $\func{carry}$ process expects input multiplicities for the channels shared with $\func{send}$ and output multiplicities for the channel shared with $\func{recv}$.
-We can now compose these processes in parallel and type the courier system.
+We can now compose these processes in parallel and type the courier $\func{system}$.
 \begin{align*}
@@
 Subject congruence states that structural congruence (\autoref{def:equals}) preserves the well-typedness of a process.
-Subject reduction states that reducing on a channel $c$ (\autoref{def:reduces}) preserves the well-typedness of a process --- after consuming $\lio$ from $c$ if $c$ is an $\constr{external}$ channel.
-We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised type-safety proofs.
-
-\begin{nitheorem}[Subject congruence]
+\begin{nitheorem}[Subject Congruence]
   \label{thm:subject-congruence}
-  If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
+  Let $P$ and $Q$ be processes. If $P \eq{} Q$ and $\types{\gamma}{\Gamma}{P}{\Xi}$, then $\types{\gamma}{\Gamma}{Q}{\Xi}$.
 \end{nitheorem}
 
-\begin{nitheorem}[Subject reduction]
+Finally, subject reduction states that reducing on a channel $c$ (\autoref{def:reduces}) preserves the well-typedness of a process --- after consuming $\lio$ from $c$ if $c$ is an $\constr{external}$ channel.
+Below we use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
+\begin{nitheorem}[Subject Reduction]
   \label{thm:subject-reduction}
@@
   If $c$ is $\constr{external} \; i$ and $\containsusage{\Gamma}{i}{\lio}{\Delta}$, then $\types{\gamma}{\Delta}{Q}{\Xi}$.
-  (We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.)
 \end{nitheorem}
+We refer to Appendix \ref{app:type-safety} for a more detailed account of the mechanised proofs.
 
@@
 
-In this paper we provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
-We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}, which is defined here for the first time for the \picalc{}.
-As a result, theorems like framing, weakening and strengthening can now be stated for the linear \picalc{}.
+We provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
+We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.
+As a result, theorems like framing, weakening and strengthening can now be stated also for the linear \picalc{}.
 Our work is fully mechanised in around 1850 lines of code in Agda \cite{Zalakain2020Agda}.
 
-As future work we intend to expand our framework to include infinite behaviour by adding replication, which can be added to a resource-algebra-agnostic type system like ours with the following rule:
-\begin{mathpar}
-  \inferrule
-  {\types{\gamma}{\Gamma}{P}{\Gamma}}
-  {\types{\gamma}{\Gamma}{\constr{!}P}{\Gamma}}
-\end{mathpar}
-That is, a process that consumes no resources (e.g. because they are shared) can be executed infinitely many times.
-To show subject congruence, one has to prove that $P \eqeq \comp{P}{\constr{!}P}$ preserves typing in both directions.
-To do so the resource consumption of a process has to be uniquely determined by e.g. adding type annotations to the syntax.
+As future work we intend to expand our framework to include infinite behaviour by adding process replication, which is challenging due to a combination of proving subject congruence and the need for type annotations in the syntax of processes.
+%which can be added to a resource-algebra-agnostic type system like ours with the expected following rule:
+%\begin{mathpar}
+%  \inferrule
+%  {\types{\gamma}{\Gamma}{P}{\Gamma}}
+%  {\types{\gamma}{\Gamma}{\constr{!}P}{\Gamma}}
+%\end{mathpar}
+%That is, a process that consumes no resources (e.g. because they are shared) can be executed infinitely many times.
+%To show subject congruence, one has to prove that $P \eqeq \comp{P}{\constr{!}P}$ preserves typing in both directions.
+%To do so the resource consumption of a process has to be uniquely determined by e.g. adding type annotations to the syntax.
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.

### 2021-01-27 paper/paper.tex
adding conversion functions paragraph

@@
 That is, binders no longer introduce names; terms at different \emph{depths} use different indices to refer to the same binding.
+
+While de Bruijn indices are useful for mechanisation, they are not as readable as names.
+To overcome this difficulty and demonstrate the correspondence between a \picalc{} that uses names and one that uses de Bruijn indices,
+we provide \emph{conversion functions} in both directions and prove that they are inverses of each other up to $\alpha$-conversion. Further details can be found in Appendix~\ref{from_to_deBruijn}.
 
@@
 
-We provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show that reduction preserves the well-typedness of processes.
-We avoid the need for extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.
+We provide a well-scoped syntax and a semantics for the \picalc{}, extrinsically define a type system on top of the syntax capable of handling linear, graded and shared types under the same unified framework and show subject reduction.
+We avoid extrinsic context splits by defining a type system based on leftover typing \cite{Allais2018a}.
 As a result, theorems like framing, weakening and strengthening can now be stated also for the linear \picalc{}.
@@
 
-We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their work and thoughts.
+We want to thank Erika Kreuter, Wen Kokke, James Wood, Guillaume Allais, Bob Atkey, and Conor McBride for their valuable suggestions.
 

### 2021-01-27 paper/paper.tex
Fix replication comment

@@
 
-As future work we intend to expand our framework to include infinite behaviour by adding process replication, which is challenging due to a combination of proving subject congruence and the need for type annotations in the syntax of processes.
-%which can be added to a resource-algebra-agnostic type system like ours with the expected following rule:
-%\begin{mathpar}
-%  \inferrule
-%  {\types{\gamma}{\Gamma}{P}{\Gamma}}
-%  {\types{\gamma}{\Gamma}{\constr{!}P}{\Gamma}}
-%\end{mathpar}
-%That is, a process that consumes no resources (e.g. because they are shared) can be executed infinitely many times.
-%To show subject congruence, one has to prove that $P \eqeq \comp{P}{\constr{!}P}$ preserves typing in both directions.
-%To do so the resource consumption of a process has to be uniquely determined by e.g. adding type annotations to the syntax.
+As future work we intend to expand our framework to include infinite behaviour by adding process replication, which is challenging, as to prove subject congruence one needs to uniquely determine the resources consumed by a process --- e.g., by adding type annotations to the syntax.
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.

### 2021-01-27 paper/paper.tex
Use wt instead of WT

@@
 \begin{align*}
-   & \func{recvWT} \;  : \; \types{\gamma \comma \channel{t}{\lz}  }{\Gamma \comma (\suc \suc l \comma r)  }{\func{recv} \; \constr{0}  }{\Gamma \comma (l \comma r)} \\
-   & \func{recvWT} \;  = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
-   & \func{sendWT} \;  : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}  }{\Gamma \comma (l \comma \suc r)  }{\func{send} \; \constr{0}  }{\Gamma \comma (l \comma r)}  \\
-   & \func{sendWT} \;  = \; \new{} \; \_ \; \_ \; \zero \; (\send{\suc \constr{0}}{\constr{0}} \PO)
+   & \func{recvwt} \;  : \; \types{\gamma \comma \channel{t}{\lz}  }{\Gamma \comma (\suc \suc l \comma r)  }{\func{recv} \; \constr{0}  }{\Gamma \comma (l \comma r)} \\
+   & \func{recvwt} \;  = \; \recv{\constr{0}}{} \recv{(\suc \constr{0})}{} \PO \\
+   & \func{sendwt} \;  : \; \types{\gamma \comma \channel{\channel{\unit}{\func{\omega}}}{\lz}  }{\Gamma \comma (l \comma \suc r)  }{\func{send} \; \constr{0}  }{\Gamma \comma (l \comma r)}  \\
+   & \func{sendwt} \;  = \; \new{} \; \_ \; \_ \; \zero \; (\send{\suc \constr{0}}{\constr{0}} \PO)
 \end{align*}
@@
 \begin{align*}
-& \func{carryWT} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
-& \func{carryWT} && = \recv{(\suc \suc \constr{0})}{} \recv{(\suc \suc \constr{0})}{} \send{(\suc \suc \constr{0})}{\suc \constr{0}} \send{(\suc \suc \constr{0})}{\constr{0}} \PO \\
-& \func{systemWT} && : \; \types{\constr{[]}}{\constr{[]}}{\func{system}}{\constr{[]}} \\
-& \func{systemWT} && = \new{} \; \_ \; \_ \; \_ \; ( \func{sendWT} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{sendWT} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{recvWT} \comp{}{ \func{carryWT} })))
+& \func{carrywt} && : \types{\gamma \comma \channel{t}{\lz} \comma \channel{t}{\lz} \comma \channel{t}{\lz}\\ & &&}{\Gamma \comma (\suc lx \comma rx) \comma (\suc ly \comma ry) \comma (lz \comma \suc \suc rz)\\ & &&}{\func{carry} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}\\ & &&}{\Gamma \comma (lx \comma rx) \comma (ly \comma ry) \comma (lz \comma rz)} & \\
+& \func{carrywt} && = \recv{(\suc \suc \constr{0})}{} \recv{(\suc \suc \constr{0})}{} \send{(\suc \suc \constr{0})}{\suc \constr{0}} \send{(\suc \suc \constr{0})}{\constr{0}} \PO \\
+& \func{systemwt} && : \; \types{\constr{[]}}{\constr{[]}}{\func{system}}{\constr{[]}} \\
+& \func{systemwt} && = \new{} \; \_ \; \_ \; \_ \; ( \func{sendwt} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{sendwt} \comp{}{ \new{} \; \_ \; \_ \; \_ } \; ( \func{recvwt} \comp{}{ \func{carrywt} })))
 \end{align*}

### 2021-01-27 paper/paper.tex
Shorten a little

@@
 Both judgments are indexed by a typing context $\gamma$, an input usage context $\Gamma$, and an output usage context $\Delta$ (the leftovers).
-
 The \textbf{typing judgement for variables} $\contains{\gamma}{\Gamma}{i}{t}{y}{\Delta}$ asserts that ``index $i$ in typing context $\gamma$ is of type $t$, and subtracting $y$ at position $i$ from input usage context $\Gamma$ results in leftovers $\Delta$''.
@@
   We must show that the variable with index $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)$ and a usage context $(\constr{[]} \comma \lio \comma \lio)$.
-  The \textsc{VarRef} constructors that we must use are completely determined by the variable index $\suc \; \constr{0}$ in the type.
+  The \textsc{VarRef} constructors are completely determined by the variable index $\suc \; \constr{0}$ in the type.
   The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to automatically fulfill this proof obligation.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- $\func{\cdot-compute^r}$ automatically fulfills this proof obligation.
   \begin{flalign*}
@@
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.
-Furthermore, it might also be worth identifying correspondences between our usage algebra and particular state machines.
 Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.

### 2021-01-27 paper/paper.tex
Fix spacing

@@
 We define the base cases of a structural congruence relation $\eqeq$ as follows:
-
 \begin{mathpar}
@@
 \end{mathpar}
-
 The first three rules ($\constr{comp}-$*) state associativity, symmetry, and $\PO$ as being the neutral element of parallel composition, respectively.
@@
   The type $\type{DEC} \; P$ is a witness of either $P$ or $P \to \type{\bot}$, where $\type{\bot}$ is the empty type with no constructors.)
-
   \begin{equation*}
@@
   For convenience, we reuse the constructor names introduced for the syntax in \autoref{syntax}.
-
   \begin{mathpar}

### 2021-04-05 paper/paper.tex
Fix typo
Address minor comments
Clarify example 2

@@
 \newcommand*{\nilemmaautorefname}{Lem.}
-\newcommand*{\exampleautorefname}{Eg.}
+\newcommand*{\exampleautorefname}{Example}
 }
@@
 \begin{example}[Variable reference]
-  In $\func{egVar}$ we type a variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.
-  We must show that the variable with index $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)$ and a usage context $(\constr{[]} \comma \lio \comma \lio)$.
+  $\func{egVar}$ defines a variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.
+  We must show that this variable is well typed in an environment with a typing context $\gamma = \constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\Gamma = \constr{[]} \comma \lio \comma \lio$.
   The \textsc{VarRef} constructors are completely determined by the variable index $\suc \; \constr{0}$ in the type.
   The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output.
-  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- $\func{\cdot-compute^r}$ automatically fulfills this proof obligation.
+  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- using $\func{\cdot-compute^r}$ to automatically fulfill the proof obligation.
   \begin{flalign*}
-  & \func{egVar} \; : \; \contains{(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)} {\constr{[]} \comma \lio \comma \lio)} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {(\constr{[]} \comma \lo \comma \lio)}\\
+  & \func{egVar} \; : \; \contains{(\constr{[]} \comma \channel{\unit}{\li} \comma \unit)} {(\constr{[]} \comma \lio \comma \lio)} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {(\constr{[]} \comma \lo \comma \lio)}\\
   & \func{egVar} \; = \; \suc \constr{0}

### 2021-04-13 paper/paper.tex
Add references to work on usage algebras

@@
   We use leftover typing to define a type system for a {resource-aware \picalc{}} \cite{MilnerPW92,Milner99}, a process algebra used to  model concurrent systems.
-  Our type system is parametrised over a set of \emph{usage algebras} that are general enough to encompass \emph{shared types} (free to reuse and discard), \emph{graded types} (use exactly $n$ number of times) and \emph{linear types} (use exactly once).
+  Our type system is parametrised over a set of \emph{usage algebras} \cite{JungSSSTBD15,TuronTABD13} that are general enough to encompass \emph{shared types} (free to reuse and discard), \emph{graded types} (use exactly $n$ number of times) and \emph{linear types} (use exactly once).
    Linear types are important in the \picalc{}: they ensure privacy and safety of communication and avoid race conditions, while graded and shared types allow for more flexible programming.
@@
   We {generalise resource counting to a set of usage algebras that can be mixed within the same type system}.
-    We do not instantiate our type system to only work with linear resources, instead we present an algebra-agnostic type system, and admit a mix of user-defined \emph{resource aware} algebras (\autoref{multiplicities}).
+    We do not instantiate our type system to only work with linear resources, instead we present an algebra-agnostic type system, and admit a mix of user-defined \emph{resource aware} algebras \cite{JungSSSTBD15,TuronTABD13} (\autoref{multiplicities}).
     Any \emph{partial commutative monoid} that is \emph{decidable}, \emph{deterministic}, \emph{cancellative} and has a \emph{minimal element} is a valid such algebra.
@@
 In the linear \picalc{} each channel has an input and an output \emph{capability}, and each capability has a given \emph{multiplicity} of 0 (exhausted) or 1 (available).
-We generalise over this notion by defining an algebra for multiplicities that is satisfied by linear, graded and shared types alike.
+We generalise over this notion by defining an algebra for multiplicities \cite{JungSSSTBD15,TuronTABD13} that is satisfied by linear, graded and shared types alike.
 We then use pairs of multiplicities as usage annotations for a channel's input and output capabilities.
@@
 %
-Based on contextual type theory \cite{Pientkaa,Pientka}, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
+Based on contextual type theory, LINCX \cite{Georges2017} extends the linear logical framework LLF \cite{Cervesato1996} by internalising the notion of bindings and contexts.
 The result is a meta-theory in which HOAS encodings with both linear and dependent types can be described.

### 2021-04-21 paper/paper.tex
Add keywords and thanks
Update paper.tex

@@
 \AtBeginDocument{%
-\renewcommand*{\figureautorefname}{Fig.\@\!}
+\renewcommand*{\figureautorefname}{Figure\@\!}
 \renewcommand*{\sectionautorefname}{\S\!\!\,}
 \renewcommand*{\subsectionautorefname}{\S\!\!\,}
-\newcommand*{\nitheoremautorefname}{Thm.}
-\newcommand*{\nidefinitionautorefname}{Def.}
-\newcommand*{\nilemmaautorefname}{Lem.}
+\newcommand*{\nitheoremautorefname}{Theorem}
+\newcommand*{\nidefinitionautorefname}{Definition}
+\newcommand*{\nilemmaautorefname}{Lemma}
 \newcommand*{\exampleautorefname}{Example}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{$\pi$ with leftovers: \\ a mechanisation in Agda}
+\title{$\pi$ with leftovers: \\ a mechanisation in Agda
+\thanks{This work is supported by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}}
 \titlerunning{$\pi$ with leftovers: a mechanisation in Agda}
@@
 }
-% Support
-% This work is supported by the UK EPSRC grant EP/K034413/1 ``From Data Types to Session Types: A Basis for Concurrency and Distribution'' (ABCD), and by the EU HORIZON 2020 MSCA RISE project 778233 ``Behavioural Application Program Interfaces'' (BehAPI).}
 
@@
 
-%  \keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
+\keywords{Pi-calculus \and Linear types \and Leftover typing \and Concurrency \and Mechanisation \and Agda}
 \end{abstract}
@@
 %
-Orchard and Yoshida \cite{OrchardY16} embed a small effectul imperative language into the session-typed \picalc{}, showing that session types are expressive enough to encode effect systems.
+Orchard and Yoshida \cite{OrchardY16} embed a small effecftul imperative language into the session-typed \picalc{}, showing that session types are expressive enough to encode effect systems.
 %
@@
 Orthogonally, we aim to investigate making our typing rules bidirectional which would allow us to provide a decision procedure for type checking processes in a given set of algebras.
-Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types and other advanced type theories.
+Finally, we will use our \picalc{} with leftovers as an underlying framework on top of which we can implement session types, via their encodings into linear types \cite{Dardha14,DardhaGS17,ScalasDHY17} and other advanced type theories.
 
